### 2022-11-08 paper.tex
Add style file and example.

\documentclass[letterpaper,12pt,leqno]{article}
\usepackage{paper,math}
\pdfoutput=1
\newcommand{\bib}{paper.bib}
\newcommand{\pdf}{figures.pdf}
\available{https://www.pascalmichaillat.org/13.html}
\hypersetup{pdftitle={u*=√uv}}

\begin{document}

\title{$\bm{u^*=\sqrt{uv}}$}
\author{Pascal Michaillat, Emmanuel Saez
\thanks{Pascal Michaillat: Brown University. Emmanuel Saez: University of California--Berkeley. We thank David Baqaee, Daniel Bogart, Varanya Chaubey, Olivia Lattus, Romain Ranciere, Guillaume Rocheteau, and Pierre-Olivier Weill for helpful comments and discussions.}}
\date{June 2022}

\begin{titlepage}\maketitle

Most governments are mandated to maintain their economies at full employment. We propose that the best marker of full employment is the efficient unemployment rate, $u^*$. We define $u^*$ as the unemployment rate that minimizes the nonproductive use of labor---both jobseeking and recruiting. The nonproductive use of labor is well measured by the number of jobseekers and vacancies, $u + v$. Through the Beveridge curve, the number of vacancies is inversely related to the number of jobseekers. With such symmetry, the labor market is efficient when there are as many jobseekers as vacancies ($u = v$), too tight when there are more vacancies than jobseekers ($v > u$), and too slack when there are more jobseekers than vacancies ($u > v$). Accordingly, the efficient unemployment rate is the geometric average of the unemployment and vacancy rates: $u^* = \sqrt{uv}$. We compute $u^*$ for the United States between 1930 and 2022. We find for instance that the US labor market has been over full employment ($u < u^*$) since May 2021.

\end{titlepage}\section{Introduction}

\paragraph{Full-employment mandate} Policymakers in many countries are mandated to maintain the economy at full employment. In the United States this legislative mandate comes from the 1978 Full Employment and Balanced Growth Act---itself an amendment of the 1946 Employment Act.\footnote{See \url{https://perma.cc/N4N5-BGJD} and \url{https://perma.cc/QK6U-C73K} for the history of these laws.} Despite the long existence of the full-employment mandate, there are no employment or unemployment rates that governments can use to capture full employment. In this paper, we produce an unemployment rate $u^*$ that characterizes a state of full employment.

\paragraph{Full employment does not mean no unemployment} Reaching full employment should be interpreted not as reaching zero unemployment. This is because it is not physically possible to reduce unemployment to zero. For instance, during World War 2, \citet[p. 125]{B44} noted that ``Even under full employment, there will be some unemployment \dots as there is some unemployment even in Britain at war to-day.''

\paragraph{So what is full employment} Rather, reaching full employment should be interpreted as reaching a socially desirable amount of unemployment. We define the socially efficient amount of unemployment as the amount that minimizes the nonproductive use of labor---both recruiting and jobseeking. It is because people constantly move in and out of jobs that unemployment cannot be reduced to zero, as noted by \citet[p. 125]{B44}: ``However great the unsatisfied demand for labor, there is an irreducible minimum of unemployment, a margin in the labor force required to make change and movement possible.'' Hence, while labor market flows are inevitable, the goal is that workers spend as much time as possible producing useful things and waste as little time as possible searching for employees or jobs.

\paragraph{Measuring jobseeking activity} The number of jobseekers is well measured by the number of unemployed workers. In theory, unemployed workers might also produce useful things at home beside looking for jobs. But in practice home production is minimal, so we count unemployed workers full-time jobseekers \citep[pp. 9--11]{MS16}.

\paragraph{Measuring recruiting activity} The number of recruiters is well measured by the number of vacancies. Indeed, it takes about one full-time worker to service a vacancy, so there are as many recruiters as vacancies \citep[p. 11]{MS16}.

\paragraph{Social objective} The social objective is to minimize the time spent on jobseeking and recruiting at the expense of producing. The amount of jobseeking is well measured by the number of unemployed workers, and the amount of recruiting by the number of job vacancies. Hence, the objective is to minimize the sum of the unemployment and vacancy rates, $u + v$.

\paragraph{Beveridge curve} Because of the Beveridge curve, it is not possible to reduce the numbers of jobseekers and vacancies at the same time. When the economy is depressed, there are many jobseekers and few vacancies; conversely, when the economy is booming, there are few jobseekers and many vacancies. The Beveridge curve is approximately a rectangular hyperbola:  $vu = A$, where $A$ is a constant \citep[figure 6]{MS16}. Hence, unemployment rate $u$ and vacancy rate $v$ are inversely related.

\paragraph{Efficiency criterion} Because of the symmetrical roles played by unemployed workers and vacant jobs, we find that the nonproductive use of labor is minimized when there are as many unemployed workers as vacant jobs ($u = v$). The labor market is inefficiently tight when there are more vacancies than jobseekers ($v > u$), and inefficiently slack when there are more jobseekers than vacancies ($u > v$).

\paragraph{Efficient labor-market tightness} The efficiency criterion can be reformulated in terms of labor-market tightness. The labor-market tightness is the ratio of vacancies to unemployment, $\t = v/u$. It measures the number of vacancies per unemployed worker. Our analysis implies that the efficient labor-market tightness is $\t^* = 1$. The labor market is inefficiently tight when $\t > 1$ and inefficiently slack when $\t < 1$.

\paragraph{Efficient unemployment rate} Because the unemployment and vacancy rates play a symmetric role on the labor market, the efficient unemployment rate is the geometric average of the unemployment and vacancy rates: $u^* = \sqrt{uv}$. This formula arises by combining the Beveridge curve, $uv = A$, and the result that the efficient unemployment and vacancy rates are equal.

\paragraph{Measuring the efficient unemployment rate the United States} Because it only requires unemployment and vacancy rates, the formula $u^* = \sqrt{uv}$ is easy to apply. In the United States, we can combine the historical unemployment and vacancy series produced by \citet{B10d} and \citet{PZ21} with the data produced by the Bureau of Labor Statistics to assemble monthly unemployment and vacancy series from 1930 and 2022. We use these series to compute the efficient unemployment rate $u^*$ between 1930 and 2022. We find that $u^*$ averages $4.2\%$ over the period and is very stable.

\paragraph{Application to the post-pandemic labor market} We also apply our formula in the post-pandemic period in the United States. There has been more vacant jobs than unemployed workers since May 2021, so the labor market has been too tight since then. In March 2022, there is twice as many vacant jobs as unemployed workers. Looking at historical data going back to 1930, we see that the last time that the labor-market tightness was so high, and the last time that the unemployment rate was so much below its efficient level, was at the end of World War 2. The post-pandemic labor market is in a very unusual situation: by any measure, 2022 is experiencing the tightest labor market since World War 2.

\paragraph{Application to modern monetary policy} During a press conference in May 2022, Jerome Powell, the chair of the Federal Reserve, discussed labor-market tightness at length.\footnote{See pp. 12--13 at \url{https://perma.cc/4EGU-XZYG}.} In response to a question from journalist Howard Schneider asking which tightness the Fed might target, he responded: ``So in terms of the vacancy-to-unemployment ratio, we don't have a goal in mind. There's no specific number that we're saying, `We've got to get to that.'  \dots I think when we got to one-to-one in the, you know, in the late teens, we thought that was a pretty good number. But, again, we're not shooting for any particular number.'' We find that a vacancy-to-unemployment ratio of $1$ is indeed what the Fed should target.

\section{Derivation of the efficient-unemployment formula}\label{s:formula}

This section derives the formula for the efficient unemployment rate: $u^* = \sqrt{uv}$.

\subsection{Definition of efficiency}

From a social perspective, it is optimal to maximize the number of people who spend time producing useful things rather than shuffling employees or jobs. This perspective is consistent with the view expressed by \citet[p. 20]{B44} that ``The material end of all human activity is consumption. Employment is wanted as a means to more consumption \dots as a means to a higher standard of life.'' We therefore define the socially efficient amount of unemployment as the amount that minimizes the nonproductive use of labor---both jobseeking and recruiting. Because people constantly move in and out of jobs, it is impossible to have zero unemployment, and it is impossible to have all workers devoted to production. While labor market flows are inevitable, the goal is that workers waste as little time as possible searching for jobs or employees.

\subsection{Measuring jobseeking activity}

The number of jobseekers is well measured by the number of unemployed workers. In theory, beside looking for jobs, unemployed workers might also produce useful things at home. Such home production would be included into aggregate production and contribute to social welfare. But in practice home production is minimal, as was already noted by \citet[p. 11]{R49}: ``The most important aspect of unemployment is its wastefulness. It is the existence of unused productive resources side by side with unsatisfied human needs that is the intolerable condition.'' \citet[pp. 9--11]{MS16} measure the fraction of nonwork time devoted to home production in the United States. The measure is based on the results by \citet{BM15}. Using administrative data from the US military, \citeauthor{BM15} study how servicemembers choose between reenlisting and leaving the military. The choices allow them to estimate the difference between market production and the sum of home production and public benefits during unemployment. Subtracting the value of public benefits from these estimates, \citet[p. 11]{MS16} find that the value of home production relative to market production could be as low as $0.03$. Given such low value, we assume that unemployed workers do not engage in home production at all, and we count them as full-time jobseekers.

\subsection{Measuring recruiting activity}

The number of recruiters is well measured by the number of vacancies. In theory it might take more or less than one full-time worker to service a vacancy. But in practice it takes about one full-time worker to service a vacancy, so the numbers of recruiters and vacancies are about the same. In the United States, the amount of labor required to service a vacancy can be measured from the National Employer Survey, which was conducted by the Census Bureau in 1997 \citep{V10}. \citet[p. 11]{MS16} estimate that servicing a vacancy requires $0.92$ worker at any point in time. So it takes about $1$ worker to service a vacancy, which is what we assume here by equating the number of recruiters to the number of vacancies.

\subsection{Shape of the Beveridge curve}

\paragraph{First observations of the Beveridge curve} Looking at labor-market statistics for the United Kingdom, \citet{B44} first noted that the number of vacancies and the number of jobseekers move in opposite directions. When the economy is in a slump, there are lots of jobseekers and few vacancies. Conversely, when the economy is in a boom, there are few jobseekers and many vacancies. This negative relationship between unemployment and vacancies---known as the Beveridge curve---was first plotted by \citet[figures 1 and 2]{DD58} using UK data for 1946--1956. Moreover, \citet[p. 22]{DD58} noted that the Beveridge curve looked like a rectangular hyperbola in an unemployment-vacancy plane.

\begin{figure}[t]
\subcaptionbox{Regular scale\label{f:regular1951}}{\includegraphics[scale=\sfig,page=1]{\pdf}}\hfill
\subcaptionbox{Log scale\label{f:log1951}}{\includegraphics[scale=\sfig,page=2]{\pdf}}
\caption{Unemployment and vacancy rates in the United States, 1951--2019}
\note{The unemployment rate is constructed by the \citet{UNRATE}. For 1951--2000, the vacancy rate is constructed by \citet{B10d}; for 2001--2019, the vacancy rate is the number of job openings divided by the civilian labor force, both measured by the \citet{CLF16OV,JTSJOL}. Unemployment and vacancy rates are quarterly averages of monthly series. The gray areas are NBER-dated recessions.}
\label{f:1951}\end{figure}

\paragraph{Unemployment and vacancy rates move in opposite directions} The Beveridge curve holds remarkably well in the United States as well \citep{BD89,EMR15}. Figure~\ref{f:regular1951} depicts the unemployment rate $u$ (all jobseekers divided by the labor force) and the vacancy rate $v$ (all vacancies divided by the same labor force) in the United States for 1951--2019. The unemployment rate is constructed by the \citet{UNRATE}. For 1951--2000, the vacancy rate is constructed by \citet{B10d}; for 2001--2019, the vacancy rate is the number of job openings divided by the civilian labor force, both measured by the \citet{CLF16OV,JTSJOL}. The figure shows clearly that unemployment and vacancy rates move in opposite directions.

\paragraph{Unemployment and vacancy rates are inversely related} In fact, unemployment and vacancy appear to be the inverse of each other, so that doubling the unemployment rate cuts the vacancy rate in half, and inversely, doubling the vacancy rate cuts the unemployment rate in half. Figure \ref{f:log1951} displays again unemployment and vacancy, but now in log scale. The fluctuations of the unemployment and vacancy rates are almost the mirror image of each other. This empirical regularity is particularly striking because the two series are constructed independently of each other.

\begin{figure}[p]
\subcaptionbox{1951Q1--1961Q1: elasticity $= -0.85$\label{f:branch1}}{\includegraphics[scale=\sfig,page=3]{\pdf}}\hfill
\subcaptionbox{1961Q2--1971Q4: elasticity $= -1.02$\label{f:branch2}}{\includegraphics[scale=\sfig,page=4]{\pdf}}\vfig
\subcaptionbox{1972Q1--1989Q1: elasticity $= -0.84$\label{f:branch3}}{\includegraphics[scale=\sfig,page=5]{\pdf}}\hfill
\subcaptionbox{1989Q2--1999Q2: elasticity $= -0.94$\label{f:branch4}}{\includegraphics[scale=\sfig,page=6]{\pdf}}\vfig
\subcaptionbox{1999Q3--2009Q3: elasticity $= -1.00$\label{f:branch5}}{\includegraphics[scale=\sfig,page=7]{\pdf}}\hfill
\subcaptionbox{2009Q4--2019Q4: elasticity $= -0.84$\label{f:branch6}}{\includegraphics[scale=\sfig,page=8]{\pdf}}
\caption{Estimated Beveridge curve in the United States, 1951--2019}
\note{This figure is constructed from figures 5 and 6 in \citet{MS16}. Unemployment and vacancy rates are the same as in figure~\ref{f:1951}. The number of structural breaks, dates of structural breaks, and elasticities of the Beveridge curve between breaks are estimated with the algorithm of \citet{BP98,BP03}.}
\label{f:hyperbola}\end{figure}

\paragraph{The Beveridge curve is a rectangular hyperbola} Mathematically, the property that unemployment and vacancy rates are inversely related implies that in an unemployment-vacancy plane, the Beveridge curve is a rectangular hyperbola:  
\begin{equation*}
vu = A,
\end{equation*}
where $A>0$ is a constant.

\paragraph{Estimates of the Beveridge elasticity} We can formally establish that the Beveridge curve is a rectangular hyperbola by estimating the elasticity of the vacancy rate with respect to the unemployment rate, $\odlx{v}{u}$. An elasticity of $-1$ corresponds to an hyperbola. \citet[figures 5 and 6]{MS16} use the algorithm of \citet{BP98,BP03} to estimate the structural breaks of the US Beveridge curve, and the elasticity of the Beveridge curve between these breaks. They find that the Beveridge elasticity remains between $-0.84$ and $-1.02$, so never far from $-1$ (figure~\ref{f:hyperbola}). This finding confirms that the Beveridge curve is close to an hyperbola.

\paragraph{Informal foundation for the hyperbolic Beveridge curve} Why is the Beveridge curve an hyperbola? Empirically, the number of new matches depends symmetrically on the number of jobseekers and vacancies. Doubling jobseekers increases new matches much the same as doubling vacancies. This makes sense as both jobseekers and vacancies are striving to find each other. This is obvious for jobseekers who spend time looking for jobs, applying, doing interviews. But this is also the case for vacancies as recruiters have to advertise vacancies, and screen and interview applicants. Therefore, as new matches have to meet the number of job separations, cutting in half the number of jobseekers requires doubling the number of vacancies, explaining why unemployment and vacancy rates are the inverse of each other.

\paragraph{Formal foundation for the hyperbolic Beveridge curve} In matching models the Beveridge curve says labor market flows are balanced: the number of job separations equals the number of jobs found. The employment rate $1-u$ is approximately constant around 1 since the unemployment rate $u$ is an order of magnitude less than 1. The job-separation rate $\l$ is also constant, so the number of job separations $\l (1-u)$ is approximately constant around $\l$. So at any point of the business cycle, the number of jobs found is approximately constant at $\l$.  With the standard symmetric Cobb-Douglas matching function, $m = \o \cdot\sqrt{uv}$, the number of jobs found at any point in time is proportional to $\sqrt{uv}$. \footnote{The US matching function appears to have a Cobb-Douglas form with exponents of $0.5$ on unemployment and vacancies. See \citet[p. 9]{MS16} for a survey of the US estimates based on aggregate data, and \citet{PP01} for a broader survey.} Hence $\sqrt{uv}$ is constant along the Beveridge curve, which implies $uv = \text{constant}$ along the Beveridge curve.

\subsection{Efficiency criterion}

The social objective is to minimize the time spent on jobseeking and recruiting at the expense of producing. Since the amount of jobseeking can be measured by the number of unemployed workers, and the amount of recruiting by the number of vacancies, the social objective is to minimize the sum of the unemployment and vacancy rates, $u + v$. This minimization is subject to the Beveridge curve constraint, $u v = \text{constant}$. Because of the Beveridge curve, it is not possible to reduce unemployment and vacancies at the same time. Therefore, the social planner must trade off unemployment and vacancies.

\paragraph{Symmetry argument} Because of the symmetrical roles played by unemployed workers and vacant jobs, the nonproductive use of labor is minimized when there are as many unemployed workers as vacant jobs ($u = v$). Indeed, we would like to minimize wasted resources from unemployment and from vacancies, subject to the Beveridge curve.  Since this problem is perfectly symmetric in $u$ and $v$, the solution to minimize waste is to have $u = v$.

\paragraph{Departure from efficiency} When the number of jobseekers is not equal to the number of vacancies, the labor market is operating inefficiently. The labor market is inefficiently tight when there are more vacancies than jobseekers ($v > u$). In that case, increasing $u$ and reducing $v$ would reduce waste. The labor market is inefficiently slack when there are more jobseekers than vacancies ($u > v$). Then, reducing $u$ and increasing $v$ would reduce waste.

\paragraph{First-order condition} We can also establish these results mathematically. The planner minimizes nonproduction $u + v$ subject to Beveridge curve $uv = A$, or $v = A/u$. This is equivalent to minimizing $u+A/u$, which is convex in $u$. A first-order condition is necessary and sufficient to find the minimum. We take the derivative of $u+A/u$ with respect to $u$ and set it to $0$. We obtain $1-A/u^2 = 0$. This condition implies that the minimum occurs when $u = \sqrt{A}$. By the Beveridge curve we have $v = A/u$, so at the minimum $v = A/ \sqrt{A} = \sqrt{A}$. Accordingly, the efficient unemployment and vacancy rates satisfy $u^* = v^* = \sqrt{A}$.

\paragraph{Formulation in terms of labor-market tightness} The efficiency criterion can be reformulated in terms of labor-market tightness. The labor-market tightness is the ratio of vacancies to unemployment, $\t = v/u$. It measures the number of vacancies per unemployed workers. The analysis implies that the efficient labor-market tightness is $\t^* = 1$; that the labor market is inefficiently tight when $\t > 1$; and that the labor market is inefficiently slack when $\t < 1$.

\begin{figure}[t]
\subcaptionbox{Unemployment and vacancy rates\label{f:uv1951}}{\includegraphics[scale=\sfig,page=9]{\pdf}}\hfill
\subcaptionbox{Labor-market tightness\label{f:tightness1951}}{\includegraphics[scale=\sfig,page=10]{\pdf}}
\caption{State of the labor market in the United States, 1951--2019}
\note{The unemployment and vacancy rates come from figure~\ref{f:1951}. The labor-market tightness is the vacancy rate divided by the unemployment rate. The gray areas are NBER-dated recessions. The labor market is inefficiently slack whenever the unemployment rate is above the vacancy rate (purple shade); it is inefficiently tight whenever the unemployment rate is below the vacancy rate (orange shade). Between 1951 and 2019, the US labor market is always inefficiently slack except during three episodes: the Korean War (1951Q1--1953Q3), the Vietnam War (1965Q4--1970Q1), and the end of the Trump presidency (2018Q1--2019Q4).}
\label{f:state1951}\end{figure}

\subsection{Efficient unemployment rate}

We have seen that the efficient unemployment rate is given by $u^* = \sqrt{A}$ where the parameter $A$ determines the location of Beveridge curve, $uv = A$. Accordingly, the efficient unemployment rate is the geometric average of the unemployment and vacancy rates: $u^* = \sqrt{uv}$. Intuitively, the efficient unemployment rate is the average of the unemployment and vacancy rates because unemployment and vacancy rates play a symmetric role on the labor market.

\subsection{Similarity with previous markers of full employment}

Our criterion provide a theoretical underpinning for the old idea that the state of full employment could be assessed by comparing the number of vacant jobs to that of unemployed workers. For example,  \citet[p. 18]{B44} states that ``Full employment \dots means having always more vacant jobs than unemployed men,'' and before his report, the prevailing view was that full employment was ``a state of affairs in which the number of unfilled vacancies is not appreciably below the number of unemployed persons.'' \citet[p. 39 and chart 5]{R57} applies Beveridge' s verbal criterion to the United States. He computes the ratio between number of vacancies and number of unemployed and examines when the ratio crosses $1$. In the same spirit, the \citet{BLS22} scrutinizes the jobseeker-per-job-opening statistic and flags when it crosses $1$.

\begin{figure}[t]
\subcaptionbox{Efficient unemployment rate\label{f:efficient1951}}{\includegraphics[scale=\sfig,page=11]{\pdf}}\hfill
\subcaptionbox{Unemployment gap\label{f:gap1951}}{\includegraphics[scale=\sfig,page=12]{\pdf}}
\caption{Efficient unemployment in the United States, 1951--2019}
\note{The unemployment rate $u$ and vacancy rate $v$ come from figure~\ref{f:1951}. The efficient unemployment rate is $u^* = \sqrt{uv}$. The unemployment gap is the difference between the actual unemployment rate and the efficient unemployment rate, $u-u^*$. The gray areas are NBER-dated recessions.}
\label{f:squareroot1951}\end{figure}

\begin{figure}[t]
\includegraphics[scale=\sfig,page=13]{\pdf}
\caption{Comparison with efficient unemployment from \citet{MS16}}
\note{The Michaillat \& Saez (2021) line represents the efficient unemployment rate constructed by \citet[figure 7B]{MS16}. The $\sqrt{uv}$ line reproduces the efficient unemployment rate from figure~\ref{f:efficient1951}. The gray areas are NBER-dated recessions.}
\label{f:comparison}\end{figure}

\section{Efficient unemployment in the United States}\label{s:postwar}

We compute the efficient unemployment rate in the United States in three periods: the standard postwar period, 1951--2019; the coronavirus pandemic, 2020--2022; and the Great Depression and World War 2, 1930--1950.

\subsection{Postwar period}

\paragraph{State of the labor market} We assess the state of the US labor market between 1951 and 2019. We use the unemployment and vacancy rates from figure~\ref{f:1951}. The labor market is inefficiently slack whenever the unemployment rate is above the vacancy rate; it is inefficiently tight whenever the unemployment rate is below the vacancy rate. Between 1951 and 2019, the US labor market is always inefficiently slack except in three episodes (figure~\ref{f:uv1951}): the Korean War (1951Q1--1953Q3), the Vietnam War (1965Q4--1970Q1), and the end of the Trump presidency (2018Q1--2019Q4).

\paragraph{Visualization with labor-market tightness} The state of the US labor market can also be visualized by plotting the labor-market tightness (figure~\ref{f:tightness1951}). The labor market is inefficiently slack whenever tightness is below $1$. It is inefficiently tight whenever tightness is above $1$. And it is efficient when the tightness equals $1$---when there is just one vacancy per jobseeker. The tightness averages $0.62$ between 1951 and 2019, and it is sharply procyclical. The tightness peaked at $1.49$ in 1953Q1, during the Korean War, and it bottomed at $0.16$ in 2009Q3, during the Great Recession.

\paragraph{Efficient unemployment rate} We then compute the efficient unemployment rate between 1951 and 2019 using the formula $u^* = \sqrt{uv}$. The efficient unemployment rate is fairly stable: it remains between $3.0\%$ and $5.3\%$, with an average value of $4.2\%$ (figure~\ref{f:efficient1951}).

\paragraph{Unemployment gap} We also compute the unemployment gap, which is the difference $u-u^*$ between the actual unemployment rate and the efficient unemployment rate. The unemployment gap indicates the distance from full employment at any given time. The unemployment gap is sharply countercyclical and generally positive. The unemployment gap peaked at $+6.0$ percentage points in 2009Q4, during the Great Recession, and bottomed at $-0.6$ percentage points in 1969Q1, during the Vietnam War. The unemployment gap averages $1.6$ percentage points over the 1951--2019 period. Hence, the economy is generally not at full employment, and it is especially far from full employment in recessions.

\paragraph{Comparison with the formula from \citet{MS16}} In a previous paper, \citet[proposition 3]{MS16} compute a more sophisticated formula for the efficient unemployment rate:
\begin{equation}
u^* = \bp{\frac{\k \e}{1-\z}\cdot\frac{v}{u^{-\e}}}^{1/(1+\e)},
\label{e:ms16}\end{equation}
where  $\e$ is the Beveridge elasticity, $\z$ is the social value of nonwork, and $\k$ is the recruiting cost. The formula requires to keep track of three statistics in addition to the unemployment and vacancy rates ($\e$, $\z$, $\k$), so it is harder for policymakers to compute than the simple $u^* = \sqrt{uv}$. But $u^* = \sqrt{uv}$ is a special case of \eqref{e:ms16} when $\e=1$, $\k=1$, and $\z=0$. The two formulas are so closely related because the model of the labor market presented here is a special case of the model in \citet{MS16}. By specializing the model and therefore the values of the sufficient statistics, we obtain an extremely simple and user-friendly formula. In the United States, however, the two formulas yield almost identical unemployment rates (figure~\ref{f:comparison}). Over 1951--2019, the two efficient unemployment rates depart by $0.2$ percentage points on average, and they never depart by more than $0.5$ percentage points.

\begin{figure}[t]
\subcaptionbox{Unemployment and vacancy rates\label{f:uv2020}}{\includegraphics[scale=\sfig,page=14]{\pdf}}\hfill
\subcaptionbox{Labor-market tightness\label{f:tightness2020}}{\includegraphics[scale=\sfig,page=15]{\pdf}}
\caption{State of the labor market in the United States, 2020M1--2022M3}
\note{The unemployment rate is constructed by the \citet{UNRATE}. The vacancy rate is the number of job openings divided by the civilian labor force, both measured by the \citet{CLF16OV,JTSJOL}. The labor-market tightness is the vacancy rate divided by the unemployment rate. The gray area is the NBER-dated pandemic recession. The labor market is inefficiently slack whenever the unemployment rate is above the vacancy rate (purple shade); it is inefficiently tight whenever the unemployment rate is below the vacancy rate (orange shade). The US labor market is inefficiently slack from 2020M3 to 2021M4; it is inefficiently tight in 2020M1--2020M2 and from 2021M5 to 2022M3.}
\label{f:state2020}\end{figure}

\begin{figure}[t]
\subcaptionbox{Efficient unemployment rate\label{f:efficient2020}}{\includegraphics[scale=\sfig,page=16]{\pdf}}\hfill
\subcaptionbox{Unemployment gap\label{f:gap2020}}{\includegraphics[scale=\sfig,page=17]{\pdf}}
\caption{Efficient unemployment in the United States, 2020M1--2022M3}
\note{The unemployment rate $u$ and vacancy rate $v$ come from figure~\ref{f:uv2020}.  The efficient unemployment rate is $u^* = \sqrt{uv}$. The unemployment gap is the difference between the actual unemployment rate and the efficient unemployment rate, $u-u^*$. The gray area is the NBER-dated pandemic recession.}
\label{f:squareroot2020}\end{figure}

\begin{figure}[t]
\includegraphics[scale=\sfig,page=18]{\pdf}
\caption{Effect of the pandemic on the US Beveridge curve}
\note{Each dot represents the unemployment and vacancy rates in a month between 2001M1 and 2022M3. The unemployment rate is constructed by the \citet{UNRATE}. The vacancy rate is the number of job openings divided by the civilian labor force, both measured by the \citet{CLF16OV,JTSJOL}. The labor market is inefficiently slack when the unemployment rate is above the vacancy rate (purple shade); it is inefficiently tight when the unemployment rate is below the vacancy rate (orange shade); it is efficient when the unemployment rate equals the vacancy rate (pink line).}
\label{f:pandemic}\end{figure}

\subsection{Coronavirus pandemic}\label{s:pandemic}

We now apply our efficiency criterion and efficient-unemployment formula to the coronavirus pandemic in the United States. We focus on the period between January 2020 and March 2022.

\paragraph{Unemployment and vacancy rates} The unemployment rate is constructed by the \citet{UNRATE}. The vacancy rate is the number of job openings divided by the civilian labor force, both measured by the \citet{CLF16OV,JTSJOL}. Both series are displayed on figure~\ref{f:uv2020}.

\paragraph{State of the labor market} The labor market is inefficiently slack whenever the unemployment rate is above the vacancy rate; it is inefficiently tight whenever the unemployment rate is below the vacancy rate. We find that the US labor market is inefficiently slack from March 2020 to April 2021; it is inefficiently tight from May 2021 to March 2022 (figure~\ref{f:uv2020}).

\paragraph{Visualization with labor-market tightness} The state of the US labor market can also be visualized by plotting labor-market tightness (figure~\ref{f:tightness2020}). The labor market is inefficiently slack whenever tightness is below $1$. And it is inefficiently tight whenever tightness is above $1$. The tightness cratered to $0.20$ in April 2020, at the beginning of the pandemic. It then steadily recovered to reach $2.0$ in March 2022---the highest tightness recorded in the postwar period.

\paragraph{Efficient unemployment rate} We then compute the efficient unemployment rate between 2020 and 2022 using the formula $u^* = \sqrt{uv}$. The efficient unemployment rate was at $4.0\%$ at the onset of the pandemic, but it sharply increased above $6.0\%$ in the first months of the pandemic (figure~\ref{f:efficient2020}). It hovered around $6.0\%$ during the rest of the 2020--2022 period.

\paragraph{Unemployment gap} We also compute the unemployment gap, which is the difference $u-u^*$ between the actual unemployment rate and the efficient unemployment rate (figure~\ref{f:gap2020}). The unemployment gap was initially positive and large: the labor market was much too slack in the first pandemic year. The unemployment gap reached $+8.1$ percentage points in April 2020---its highest level in the postwar period. But the economy recovered quickly, and the unemployment gap turned negative in May 2021. The unemployment gap reached $-1.5$ percentage points in March 2022. This is the lowest unemployment gap in the postwar period. So in 2022, the labor market is well beyond full employment.

\paragraph{Why did the efficient unemployment rate increase during the pandemic?} The efficient unemployment rate increased by more than 2 percentage points at the onset of the pandemic. Such a sharp increase is unprecedented. It can be explained by the gigantic outward shift of the Beveridge curve that took place in the spring 2020. Graphically, the efficient unemployment rate appears at the intersection of the Beveridge curve and the pink identity line (figure~\ref{f:pandemic}). In February 2020, at the onset of the pandemic, the labor market was close to efficient, and the unemployment rate was around $4\%$. In May 2021, about a year later, the labor market had returned to efficiency, but the unemployment rate was now $6\%$. The increase of the efficient unemployment rate from $4\%$ to $6\%$ was caused by the outward shit of the Beveridge curve that occurred between March and April 2020. Mathematically, the efficient unemployment rate is determined by the location of the Beveridge curve ($u^* = \sqrt{A}$), so only a sharp outward shift of the Beveridge curve (increase in $A$) can increase the efficient unemployment rate.\footnote{The previous shifts of the Beveridge curve during the postwar period (figure~\ref{f:hyperbola}) were not large enough to produce noteworthy changes in the efficient unemployment rate (figure~\ref{f:squareroot1951}).}

\begin{figure}[t]
\subcaptionbox{Regular scale\label{f:regular1930}}{\includegraphics[scale=\sfig,page=19]{\pdf}}\hfill
\subcaptionbox{Log scale\label{f:log1930}}{\includegraphics[scale=\sfig,page=20]{\pdf}}
\caption{Unemployment and vacancy rates in the United States, 1930--1950}
\note{The unemployment and vacancy rates are constructed by \citet{PZ21}. Unemployment and vacancy rates are quarterly averages of monthly series. The gray areas are NBER-dated recessions.}
\label{f:1930}\end{figure}

\begin{figure}[t]
\subcaptionbox{Unemployment and vacancy rates\label{f:uv1930}}{\includegraphics[scale=\sfig,page=21]{\pdf}}\hfill
\subcaptionbox{Labor-market tightness\label{f:tightness1930}}{\includegraphics[scale=\sfig,page=22]{\pdf}}
\caption{State of the labor market in the United States, 1930--1950}
\note{The unemployment and vacancy rates come from figure~\ref{f:1930}.  The labor-market tightness is the vacancy rate divided by the unemployment rate. The gray areas are NBER-dated recessions. The labor market is inefficiently slack whenever the unemployment rate is above the vacancy rate (purple shade); it is inefficiently tight whenever the unemployment rate is below the vacancy rate (orange shade). Between 1930 and 1950, the US labor market is always inefficiently slack except during World War 2 (1942Q3--1946Q3).}
\label{f:state1930}\end{figure}

\begin{figure}[t]
\subcaptionbox{Efficient unemployment rate\label{f:efficient1930}}{\includegraphics[scale=\sfig,page=23]{\pdf}}\hfill
\subcaptionbox{Unemployment gap\label{f:gap1930}}{\includegraphics[scale=\sfig,page=24]{\pdf}}
\caption{Efficient unemployment in the United States, 1930--1950}
\note{The unemployment rate $u$ and vacancy rate $v$ come from figure~\ref{f:1930}. The efficient unemployment rate is $u^* = \sqrt{uv}$. The unemployment gap is the difference between the actual unemployment rate and the efficient unemployment rate, $u-u^*$. The gray areas are NBER-dated recessions.}
\label{f:squareroot1930}\end{figure}

\subsection{Great Depression and World War 2}\label{s:depression}

Finally, we apply our efficiency criterion and efficient-unemployment formula to the historical period 1930--1950, which covers both the Great Depression and World War 2.

\paragraph{Unemployment and vacancy rates} The unemployment and vacancy rates are constructed by \citet{PZ21} from NBER macrohistory files. Between 1930 and 1950 it remains true that unemployment and vacancy rates move in opposite directions (figure~\ref{f:regular1930}). Using a log scale, it also appears that unemployment and vacancy rates are inversely related (figure~\ref{f:log1930}). These fluctuations indicate that just as in the postwar era, the hyperbolic Beveridge curve holds in 1930--1950.\footnote{See also figure 3 in \citet{PZ21}.}

\paragraph{State of the labor market} The labor market is inefficiently slack whenever the unemployment rate is above the vacancy rate; it is inefficiently tight whenever the unemployment rate is below the vacancy rate. The US labor market is always inefficiently slack between 1930 and 1950 except at the end World War 2, during 1942Q3--1946Q3 (figure~\ref{f:uv1930}).

\paragraph{Visualization with labor-market tightness} The state of the labor market can also be visualized by plotting labor-market tightness (figure~\ref{f:tightness1930}). The labor market is inefficiently slack whenever tightness is below $1$; it is inefficiently tight whenever tightness is above $1$. Tightness is extremely volatile during the 1930--1950 period: it plunged to $0.03$ in 1932Q3, during the Great Depression, and climbed all the way to $6.8$ in 1944Q4, at the end of World War 2. These are the most extreme tightness fluctuations on record.

\paragraph{Efficient unemployment rate} We then compute the efficient unemployment rate between 1930 and 1950 using the formula $u^* = \sqrt{uv}$. The efficient unemployment rate is again fairly stable: it remains between $2.5\%$ and $4.6\%$, with an average value of $3.5\%$ (figure~\ref{f:efficient1930}). This value of the efficient unemployment rate is actually pretty close from the full-employment rate of unemployment advocated by \citet[p. 21]{B44} just after World War 2, which was $3\%$.

\paragraph{Unemployment gap} Finally, we compute the unemployment gap, which is the difference $u-u^*$ between the actual unemployment rate and the efficient unemployment rate (figure~\ref{f:gap1930}). The unemployment gap was of course positive and very large during the Great Depression: the labor market was much too slack then. The unemployment gap reached $+20.9$ percentage points in 1932Q3---its highest level on record. The economy recovered only slowly from the Great Depression . The unemployment gap finally turned negative during World War 2, in 1942Q3. The unemployment gap reached $-1.6$ percentage points in 1945Q1---its lowest level on record. The unemployment gap turned positive again during the 1948--1949 recession.

\newlength{\imageheight}
\settoheight{\imageheight}{\includegraphics[page=24]{\pdf}}
\begin{figure}[p]
\subcaptionbox{Unemployment and vacancy rates, $u$ and $v$\label{f:uv}}{\includegraphics[trim=0 0 0 0.5\imageheight, clip, scale=\lfig,page=25]{\pdf}}\vfig
\subcaptionbox{Labor-market tightness, $v/u$\label{f:tightness}}{\includegraphics[trim=0 0 0 0.5\imageheight, clip, scale=\lfig,page=26]{\pdf}}\vfig
\subcaptionbox{Efficient unemployment rate, $u^* = \sqrt{uv}$ \label{f:squareroot}}{\includegraphics[trim=0 0 0 0.5\imageheight, clip, scale=\lfig,page=27]{\pdf}}
\caption{State of the US labor market, 1930Q1--2022Q1}
\note{The unemployment and vacancy rates are obtained by splicing the unemployment and vacancy rates from figures~\ref{f:uv1951}, \ref{f:uv2020}, and \ref{f:uv1930}. Labor-market tightness and efficient unemployment rate are constructed from the unemployment and vacancy rates. The gray areas are NBER-dated recessions.}
\label{f:summary}\end{figure}

\subsection{Summary}

To summarize, we combine the US data for 1930Q1--2022Q1.

\paragraph{State of the labor market} The unemployment rate is generally above the vacancy rate, and this gap is exacerbated in recessions (figure~\ref{f:uv}). This means that the labor market does not operate efficiently, unlike what economists often assume. It also means that monetary and fiscal policies are needed to reduce unemployment in bad times, and adhere more closely to the full-employment mandate. Before 2018, the labor market is only inefficiently tight in wartime---World War 2, Korean War, Vietnam War. Since then the labor market has been inefficiently tight at the end of the Trump presidency and during the recovery of the coronavirus pandemic. The state of the labor market in 2018--2019 and 2021--2022 is a rarity: it is the only peacetime labor market that is exceedingly tight. \citet[p. 322]{K36} doubted that the labor market could reach full employment in peacetime. It seems that he was essentially right: before 2018 the US labor market never reached full employment in peacetime.

\paragraph{Labor-market tightness} The state of the labor market can also be visualized by plotting labor-market tightness (figure~\ref{f:tightness}). Over 1930Q1--2022Q1, labor-market tightness averages $0.68$. In 2022, tightness reached a value of $2.0$, which it had last reached in 1945. In the recovery of the pandemic, the US labor market has reached historical levels of tightness.

\paragraph{Full employment} Over 1930Q1--2022Q1, the efficient unemployment rate averages $4.2\%$ (figure~\ref{f:squareroot}). The efficient unemployment rate is stable over time. It hovered around $4\%$ between 1930 and 1970. It raised to about $5\%$ in the 1970s and stayed there in the 1980s. It then remained around $4\%$ again between 1990 and 2020. Finally, it rose to $6\%$ during the pandemic. This means that in the United States, the full-employment unemployment rate is stable over time, around $4\%$.

\section{Frequently asked questions}\label{s:faq}

To conclude, we address questions that are frequently asked about the analysis. We cover questions about the analysis, about the policy implications of the results, about the theoretical underpinnings of the results, and about possible extensions.

\subsection{Shouldn't unemployment dynamics play a role?}

\paragraph{Informal argument} Unemployment perpetually evolves through a dynamic process driven by differences in inflow into and outflow from unemployment. Yet, our analysis of labor-market efficiency is static. But because gross flows are so large relative to net flows---$4\%$ of employment per month compared to a fraction of a percent of employment per month---this dynamic unemployment process converges extremely quickly to the Beveridge curve. This explains why the static analysis is an excellent approximation to the dynamic process.

 \paragraph{Formal argument in a matching model \citep[p. 7]{MS16}} In a matching model the unemployment rate evolves according to the following differential equation:
\begin{equation}
\dot{u}(t) = \l [1-u(t)] - f u(t),
\label{e:uDot}\end{equation}
where $\l$ is the job-separation rate and $f$ is the job-finding rate. The term $\l [1-u(t)]$ gives the number of workers who lose or quit their jobs and enter unemployment during a unit time. The term $f u(t)$ gives the number of workers who find a job and leave unemployment during a unit time. The difference between inflow into and outflow from unemployment determines the change in the unemployment rate, $\dot{u}(t)$. The solution to the differential equation~\eqref{e:uDot} is
\begin{equation}
u(t) - u^b = [u(0) - u^b] e^{-(\l+f)t},
\label{e:uSolution}\end{equation}
where
\begin{equation*}
u^b = \frac{\l}{\l+f}.
\end{equation*}
The unemployment rate $u^b$ is the critical point of differential equation~\eqref{e:uDot}, so it corresponds to the unemployment rate given by the Beveridge curve when the job-finding rate is $f$. Equation \eqref{e:uSolution} shows that the distance between the unemployment rate $u(t)$ and the Beveridgean unemployment rate $u^b$ decays at an exponential rate $\l+f$. In the United States, the rate of decay is really fast: between 1951 and 2019, the job-finding rate averages $f = 55.8\%$ per month, the job-separation rate averages $\l = 3.2\%$ per month, so the rate of decay averages $\l+f = 59.0 \%$ per month. Accordingly, the half-life of the deviation from the Beveridgean unemployment rate, $u(t)-u^b$, is $\ln(2)/0.590 = 1.17$ month. Since about $50\%$ of the deviation evaporates within one month, about $90\%$ of it evaporates within one quarter, which implies that the unemployment rate is always close to the Beveridgean unemployment rate.

\subsection{Shouldn't workers out of the labor force be part of the calculation?}

If the labor-force participation rate systematically responded to the unemployment rate, then it would make sense to endogenize labor-force participation and include it in the welfare analysis. And in theory, two kinds of effects are possible. It would be possible that additional workers enter the labor force in bad times to supplement their household's income. It would also be possible that jobseekers become discouraged in bad times and leave the labor force. In practice, however, the labor-force participation rate appears acyclical, so variations in unemployment do not lead to systematic changes in participation \citep[p. 294]{S09}. Shimer's result is based on US data for 1960--2006. It is in line with earlier evidence from the United States. Using data for 1946--1954: \citet[p. 32]{R57} found that ``There is no substantial evidence upholding the view that labor-force participation declines when there are moderate decreases in the demand for labor.''

\subsection{Shouldn't the hardship from unemployment be included in the calculation?}

Beyond production lost, unemployment can also create psychological hardship for unemployed workers. This psychological toll has been understood for a long time, even though it is often neglected in economics. \citet[p. 11]{R49} for instance noted that ``The most striking aspect of unemployment is the suffering of the unemployed and their families---the loss of health and morale that follows loss of income and occupation.'' In contrast, unemployed workers might enjoy more leisure than employed workers. \citet{MS16} account for such considerations and find that psychological costs from unemployment offset the value of home production and leisure by unemployed workers. In net, it is as if unemployed workers do not engage in home production---which is what we assume here.

\subsection{Are recruiting and jobseeking useless activities?}

Of course recruiting and jobseeking are necessary for workers and firms to match with each other. But they do not generate any social welfare by themselves. In academia, faculty members have to spend time recruiting new colleagues. But during that time they do not teach or do research. Similarly, when graduate students are on the job market looking for a faculty position, they cannot teach or do research. In this paper we are looking for the unemployment rate that minimizes disruptions caused by recruiting and jobseeking.

\subsection{Shouldn't we eliminate unemployment to almost zero?}

Unemployment is clearly a waste of economic resources as people who would like to work and produce are forced to look for jobs and are not able to be productive. Yet, reducing the unemployment rate to zero is not desirable because, by the logic of the Beveridge curve, it would require a huge number of vacancies. Servicing vacancies takes work--- posting advertisements, screening and interviewing candidates---work that has to be diverted away from other productive tasks. And machines cannot do recruiting for firms. Therefore, vacancies also consume economic resources. It turns out that job recruiting is as labor intensive as job seeking is: one vacancy consumes as much resources as what is wasted when one person is looking for a job instead of working. As a result, it is efficient to reduce the sum of the unemployment and vacancy rates, and not simply the unemployment rate.

\subsection{Don't people move from job to job without going through unemployment?}

It is true that many workers change jobs without going through unemployment. This does reduce the waste on the jobseeking side of the labor market. Conceptually, such moves can be considered as unemployment of duration zero. While it would be ideal if all unemployment spells were of zero duration, empirically they are not: this is why the Beveridge curve exists. Our analysis follows from the Beveridge curve, and it applies even with job-to-job moves.

\subsection{What do the results imply for monetary policy?}

\paragraph{In practice} In the United States, the government is mandated to stabilize the economy at full employment. We have argued that our efficient unemployment rate $u^* = \sqrt{uv}$ is the appropriate measure of full employment. Hence, the Fed could use it as their full-employment target---especially since $\sqrt{uv}$ can be measured in real time.

\paragraph{In theory} Moreover, targeting $u^*$ is optimal in certain monetary models with unemployment. This happens for instance in the model developed by \citet{MS19}. In that model, it is optimal to eliminate the unemployment gap by setting the nominal interest rate appropriately. As the unemployment gap is countercyclical, and a reduction in interest rate lowers unemployment, it is optimal to lower the interest rate in slumps, when the unemployment gap is high, and to raise it in booms, when the unemployment gap turns negative.

\paragraph{Sufficient-statistic formula} In fact, \citet[section 5]{MS19} show that starting from a nominal interest rate $i>0$ and an inefficient unemployment rate $u \neq u^*$, the Fed should set the federal funds rate to $i^*$ such that
\begin{equation}
 i - i^* =  \frac{u-u^*}{du/di}.
\label{e:monetary}\end{equation}
The statistic $i-i^*$ indicates the decrease in interest rate required to reach the optimal policy. The statistic $u-u^*$ is the prevailing unemployment gap. And the statistic $du/di>0$ is the monetary multiplier: the percentage-point decrease in unemployment achieved by lowering the nominal interest rate by 1 percentage point. This sufficient-statistic formula ensures that the unemployment rate is efficient, and therefore also that the economy is at full employment.

\paragraph{Monetary multiplier} A large literature estimates the effect of the federal funds rate on unemployment \citep{CEE99,C12,R15}. A midrange estimate of the US monetary multiplier is $du/di = 0.5$ \citep[p. 402]{MS19}. Combined with a monetary multiplier of $0.5$, formula~\eqref{e:monetary} says that the Fed should cut its interest rate by $2$ percentage points for each positive percentage point of unemployment gap, and raise its interest rate by $2$ percentage points for each negative percentage point of unemployment gap. For instance in March 2022, the economy is exceedingly tight: the unemployment gap is $-1.6$ percentage points. The Fed should raise its interest rate by $1.6\times 2 = 3.2$ percentage points in total to cool the economy back to full employment.

\paragraph{At the zero lower bound} If the zero lower bound on nominal interest rates becomes binding, conventional monetary policy will not be able to sustain full employment. But other policies, such as government spending, can bring the economy closer to full employment. In a fiscal model with unemployment, \citet{MS15} find that optimal government spending deviates from the \citet{S54} rule to reduce---but not eliminate---the unemployment gap. Since the unemployment gap is countercyclical, and an increase in government spending reduces unemployment \citep{Ra13}, optimal government spending is countercyclical.

\subsection{Shouldn't redistributive considerations be included in the analysis?}

Our simple efficiency criterion minimizes the nonproductive use of labor. However, it is reasonable to think that waste on the jobseekers' side from looking for a job is more harmful than waste on the employer side for spending resources to fill vacancies. In other words, for distributive reasons, we might prefer having more waste if this waste is pushed toward employers. It is indeed the case that labor likes tight labor markets. Under some conditions, it is theoretically preferable to have direct redistribution (for example through more generous unemployment insurance funded by taxes on the wealthy) rather than through inefficiently tight labor markets. Practically, if taxes and transfers are set and cannot be made more generous, then an excessively tight labor market is an indirect way to redistribute toward labor.

\subsection{Is the hyperbolic Beveridge curve consistent with a matching function?}

\citet[equation (A12)]{MS16} relate the matching elasticity $\h$ to the Beveridge elasticity $\e$ in a matching model:
\begin{equation*}
\h = \frac{1}{1+\e}\bp{\e-\frac{u}{1-u}}.
\end{equation*}
An hyperbolic Beveridge curve implies a Beveridge elasticity $\e=1$. With an unemployment rate around $u=5\%$, we obtain a matching elasticity of $\h = 0.47$. Such matching elasticity falls squarely in the range of estimates obtained with aggregate US data, which span $0.30$--$0.76$ \citet[p. 9]{MS16}. So the hyperbolic Beveridge curve is consistent with US estimates of the matching function.

\subsection{What causes the observed fluctuations along the Beveridge curve?}

Over the business cycle, unemployment and vacancy rates move along the Beveridge curve (such as in figure~\ref{f:pandemic}). Different models offer different explanations for these fluctuations? In the DMP model, shocks to workers' bargaining power lead to fluctuations along the Beveridge curve \citep[table~6]{S05}. In the variants of the DMP model by \citet{H05}, \citet{HM08}, and \citet{M09}, real wages are rigid, and shocks to labor productivity generate realistic fluctuations along the Beveridge curve. In the mismatch model by \citet{S07} and stock-flow matching model by \citet{ES10}, shocks to labor productivity also generate sizable fluctuations along the Beveridge curve. Finally, in the business-cycle model by \citet{MS19}, prices are rigid and shocks to aggregate demand generate fluctuations along the Beveridge curve. Such shocks may be caused by changes in monetary policy, or by changes in people's willingness to save (modeled as changes in people's marginal utility of wealth).

\subsection{Should the government really take the Beveridge curve as given?}

\paragraph{Empirical argument} Empirically the Beveridge curve is very stable over the business cycle. It does not seem to respond to stabilization policies (either monetary or fiscal policy) in any systematic way. Given this stability, it makes sense that the Fed takes the Beveridge curve as given.

\paragraph{Theoretical argument} Most Beveridgean models of monetary and fiscal policy are built around a matching function \citep{W05,BG08,RW11,M12,MS15,MS19}. In these models the Beveridge curve is determined by the matching function and job-separation rate. Neither respond to monetary or fiscal policy, so the Beveridge curve is unaffected by policy. In the context of business-cycle stabilization, it therefore seems appropriate for the government to take the Beveridge curve as given.\footnote{Stabilization policies do not seem to affect the Beveridge curve, but some structural policies might. For instance, Germany implemented labor-market reforms in 2003--2005 that were designed to improve how the labor market operates (Hartz reforms). Such policies might shift the Beveridge curve inward.}

\subsection{What explains the inefficiency of the labor market?}

In most Beveridgean models, unlike in Walrasian models, the market economy does not overlap with the planning economy. This is because most price and wage mechanisms do not ensure efficiency. For instance, in matching models of the labor market, the wage is determined in a situation of bilateral monopoly, so a range of wages is theoretically possible. A wage mechanism picks one wage within the range. There is only an infinitesimal chance that this wage is the one ensuring efficiency \citep[p.~185]{P00}. Accordingly, theory does not guarantee that unemployment is efficient---or that full employment occurs.

\subsection{What does labor-market tightness imply for inflation?}

In the recovery of the coronavirus pandemic in the United States, the labor market has reached a tightness never seen since World War 2 (figure~\ref{f:tightness}). At the same time, wage inflation has also been very high. The Employment Cost Index for wages and salaries of workers in the private sector grew at an annual rate of $6.3\%$ in 2021Q3 and $5.0\%$ in 2022Q1, the highest rates recorded since the Employment Cost Index was created in 2001 \citep{ECIWAG}. It is therefore quite possible that wage inflation accelerates when the labor market is exceedingly tight. \citet{DS22} provide some support for this hypothesis. They find that both unemployment and vacancies are good predictors of wage inflation. Since labor-market tightness is the vacancy-to-unemployment ratio, it could be a good predictor of wage inflation as well.

\bibliography{\bib}

\end{document}

### 2023-01-16 appendix.tex
Add template for paper and appendix

\documentclass[letterpaper,12pt,leqno]{article}
\usepackage{paper,appendix}
\bibliographystyle{bibliography}
\hypersetup{pdftitle={Paper Title: Online Appendices}}
% Enter journal of publication and URL to paper
\published{Journal Name}{https://github.com/pmichaillat/latex-paper}
% Enter BibTeX file with references here:
\newcommand{\bib}{bibliography.bib}
% Enter PDF file with figures here:
\newcommand{\pdf}{figures.pdf}
% Enter TeX file with the main text so references can be used in appendix file (this is especially useful if the appendix has to be split away from the main text)
\externaldocument{paper}

\begin{document}

\title{Paper Title: Online Appendices}
\author{Author 1, Author 2}
\date{Month Year}

\begin{titlepage}
\maketitle
\tableofcontents
\end{titlepage}

\section{A section}\label{a:appendix1}

We extend the p-hacking model of section~\ref{s:section} by introducing a cost of research, incurred at each new p-hacking step. Obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.

\subsection{Links}

The references from the paper file are available in the appendix as long as the appendix is compiled after the paper and intermediary LaTeX-related files are not deleted.

\begin{itemize}    
    \item Equation \ref{e:type1} was very helpful.
    \item Figure \ref{f:graph1} provided a lot of information---especially the plot on figure \ref{f:panel2}. 
\end{itemize}

\subsection{Assumptions}

At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga. A corrolary in the appendix is as follows:

\begin{corollary} The number of p-hacking steps might be
\begin{equation*}
\mathbb{E}(\Omega) = \mathbb{P}(\omega\cdot \mu - \xi) - \sum_{i=0}^{m}\sum_{j=-\infty}^{n} \sigma(i,j) + 123^{56}.
\end{equation*}\end{corollary}

\subsection{Optimal stopping time and robust critical value}

\paragraph{Significant result} Since it is optimal to engage in research, the researcher starts a first p-hacking step. With probability $\xi$, the step can be completed, and the researcher obtains a test result \citep{MS22a}.

\paragraph{Another paragraph with some math} When a researcher decides to start research, three scenarios are again possible. With probability $1-\gamma$, the researcher cannot complete the first research step and cannot submit any result; she then collects $0$. With probability $\gamma$, she can complete the first research step. Then with probability $S(z^*)$, her result is significant and she collects $v^s$. With probability $1-S(z^*)$, her result is insignificant and the continuation value at this point is $\mathcal{V}^i$. Aggregating these scenarios, we obtain the initial continuation value:
\begin{equation}
\mathbb{V}^r = (1-\gamma) \times 0 +\gamma S(z^*) v^s+\gamma [1-S(z^*)] \mathcal{V}^i-c.
\label{e:appendix1}\end{equation}

\paragraph{Paragraph with links to appendix equations} What does the researcher decide if the result is insignificant? It depends on the research cost $\mathcal{C}$. Equation \eqref{e:appendix1} shows that if the cost is high enough, the researcher stops right away. This happens when the possibility of obtaining a significant result in the future does not compensate the research cost.

\subsection{A larger figure, without panel, in the appendix}

At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga. This is showed in figure \ref{f:appendix1}.

\begin{figure}[t]
\includegraphics[scale=\mfig,page=1]{\pdf}
\caption{A caption for the larger graph}
\note{A note for the larger graph. Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus.}
\label{f:appendix1}\end{figure}

\section{Another section}\label{a:appendix2}

At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti.

\subsection{A even larger figure, without panel, in the appendix}

At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.

\begin{figure}[t]
\includegraphics[scale=\lfig,page=3]{\pdf}
\caption{A caption for the even larger graph}
\note{A note for the larger graph. Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus.}
\label{f:appendix2}\end{figure}

When a researcher decides to continue p-hacking, three scenarios are possible, as showed in figure \ref{f:appendix2}. The vector probability $1-\bm{\gamma}$ then gives $v^i$. The vector $\bm{\gamma}$ contains the p-hacking steps. Then with probability $S(z^*)$, her result is significant and she collects $v^s$. With probability $1-S(z^*)$, her result is insignificant once again and the continuation value at this point is $\mathcal{V}^i$. In any case, she must incur a cost $c$ to conduct the research step.

We rewrite the initial continuation value as
\begin{equation}
\mathbb{V}^r = \zeta \mathcal{V}^i + \zeta S(z^*) (v^s - \mathcal{V}^i) - c.
\label{e:appendix2}\end{equation}

Equation \eqref{e:appendix2} shows when there is no p-hacking. The robust critical value is then just the classical critical value.

\subsection{Condition for p-hacking}\label{a:subappendix}

When a researcher has obtained one insignificant result, it is optimal to continue p-hacking if $\mathcal{V}^i > \mathcal{V}^r$. This is related to the results by \citet{MS21b}.\footnote{The reference goes to its own reference list at the end of the appendix---unlike when the appendix was at the end of the main text.}

\subsection{Continuation value of research}

We first compute the continuation value of research for a researcher who has already recorded an insignificant result. We denote this value $\mathcal{B}^\theta$. Because the researcher's situation is invariant in time, the continuation value is the same at each p-hacking step.

\subsection{Condition for research}

We also compute the cost below which it is optimal to engage in research. Given that we have normalized the outside option of the researcher to $0$, it is optimal to engage in research if the expected value from it is positive. This results are summarized in a repository with the following URL: \url{https://github.com/pmichaillat/latex-paper}.

\bibliography{\bib}

\end{document}

### 2023-01-16 paper.tex
Upload TeX file as an example
Add template for paper and appendix

@@
 \documentclass[letterpaper,12pt,leqno]{article}
-\usepackage{paper,math}
-\pdfoutput=1
-\newcommand{\bib}{paper.bib}
+\usepackage{paper}
+\bibliographystyle{bibliography}
+\hypersetup{pdftitle={Paper Title}}
+% Enter permanent URL to paper
+\available{https://github.com/pmichaillat/latex-paper}
+% Enter BibTeX file with references here:
+\newcommand{\bib}{bibliography.bib}
+% Enter PDF file with figures here:
 \newcommand{\pdf}{figures.pdf}
-\available{https://www.pascalmichaillat.org/13.html}
-\hypersetup{pdftitle={u*=√uv}}
 
@@
 
-\title{$\bm{u^*=\sqrt{uv}}$}
-\author{Pascal Michaillat, Emmanuel Saez
-\thanks{Pascal Michaillat: Brown University.
-Emmanuel Saez: University of California--Berkeley.
-We thank David Baqaee, Daniel Bogart, Varanya Chaubey, Olivia Lattus, Romain Ranciere, Guillaume Rocheteau, and Pierre-Olivier Weill for helpful comments and discussions.}}
-\date{June 2022}
+\title{Paper Title}
+\author{Author 1, Author 2
+\thanks{Author 1: University 1.
+Author 2: University 2.
+We thank many colleagues for helpful comments and discussions.
+This work was supported by the a generous grant [grant number]; and another generous foundation.}}
+\date{Month Year}
 
@@
 
-Most governments are mandated to maintain their economies at full employment.
-We propose that the best marker of full employment is the efficient unemployment rate, $u^*$.
-We define $u^*$ as the unemployment rate that minimizes the nonproductive use of labor---both jobseeking and recruiting.
-The nonproductive use of labor is well measured by the number of jobseekers and vacancies, $u + v$.
-Through the Beveridge curve, the number of vacancies is inversely related to the number of jobseekers.
-With such symmetry, the labor market is efficient when there are as many jobseekers as vacancies ($u = v$), too tight when there are more vacancies than jobseekers ($v > u$), and too slack when there are more jobseekers than vacancies ($u > v$).
-Accordingly, the efficient unemployment rate is the geometric average of the unemployment and vacancy rates: $u^* = \sqrt{uv}$.
-We compute $u^*$ for the United States between 1930 and 2022.
-We find for instance that the US labor market has been over full employment ($u < u^*$) since May 2021.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
 
-\end{titlepage}\section{Introduction}
+\end{titlepage}\section{Introduction}\label{s:introduction}
 
-\paragraph{Full-employment mandate} Policymakers in many countries are mandated to maintain the economy at full employment.
-In the United States this legislative mandate comes from the 1978 Full Employment and Balanced Growth Act---itself an amendment of the 1946 Employment Act.\footnote{See \url{https://perma.cc/N4N5-BGJD} and \url{https://perma.cc/QK6U-C73K} for the history of these laws.} Despite the long existence of the full-employment mandate, there are no employment or unemployment rates that governments can use to capture full employment.
-In this paper, we produce an unemployment rate $u^*$ that characterizes a state of full employment.
+\paragraph{Research question} Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.\footnote{Excepteur sint, sunt in culpa qui officia deserunt mollit anim id est laborum.}
 
-\paragraph{Full employment does not mean no unemployment} Reaching full employment should be interpreted not as reaching zero unemployment.
-This is because it is not physically possible to reduce unemployment to zero.
-For instance, during World War 2, \citet[p.
-125]{B44} noted that ``Even under full employment, there will be some unemployment \dots as there is some unemployment even in Britain at war to-day.''
+\paragraph{Positioning in the literature} Lorem ipsum dolor sit amet \citep{MS15}, consectetur adipiscing elit \citep[p.
+1305]{MS19}, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua .
+\citet{M12,M14} enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+\citeauthor{EMM21} aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.\footnote{In reprehenderit in eu fugiat nulla pariatur \citep{LMS18a}.}
 
-\paragraph{So what is full employment} Rather, reaching full employment should be interpreted as reaching a socially desirable amount of unemployment.
-We define the socially efficient amount of unemployment as the amount that minimizes the nonproductive use of labor---both recruiting and jobseeking.
-It is because people constantly move in and out of jobs that unemployment cannot be reduced to zero, as noted by \citet[p.
-125]{B44}: ``However great the unsatisfied demand for labor, there is an irreducible minimum of unemployment, a margin in the labor force required to make change and movement possible.'' Hence, while labor market flows are inevitable, the goal is that workers spend as much time as possible producing useful things and waste as little time as possible searching for employees or jobs.
+\paragraph{Answer to the question} Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.\footnote{Sunt in culpa qui officia deserunt mollit anim id est laborum.
+Excepteur sint occaecat cupidatat non proident.
+Sunt in culpa qui officia deserunt mollit anim id est laborum.
+Excepteur sint occaecat cupidatat non proident.} Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
 
-\paragraph{Measuring jobseeking activity} The number of jobseekers is well measured by the number of unemployed workers.
-In theory, unemployed workers might also produce useful things at home beside looking for jobs.
-But in practice home production is minimal, so we count unemployed workers full-time jobseekers \citep[pp.
-9--11]{MS16}.
+\paragraph{Another paragraph} Sed ut perspiciatis, unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam eaque ipsa, quae ab illo inventore veritatis et quasi architecto beatae vitae dicta sunt, explicabo.
+Nemo enim ipsam voluptatem, quia voluptas sit, aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos, qui ratione voluptatem sequi nesciunt, neque porro quisquam est, qui dolorem ipsum, quia dolor sit amet consectetur adipiscing velit, sed quia non numquam eius modi tempora incidunt, ut labore et dolore magnam aliquam quaerat voluptatem.
 
-\paragraph{Measuring recruiting activity} The number of recruiters is well measured by the number of vacancies.
-Indeed, it takes about one full-time worker to service a vacancy, so there are as many recruiters as vacancies \citep[p.
-11]{MS16}.
+\section{Section}\label{s:section}
 
-\paragraph{Social objective} The social objective is to minimize the time spent on jobseeking and recruiting at the expense of producing.
-The amount of jobseeking is well measured by the number of unemployed workers, and the amount of recruiting by the number of job vacancies.
-Hence, the objective is to minimize the sum of the unemployment and vacancy rates, $u + v$.
+Ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat.
 
-\paragraph{Beveridge curve} Because of the Beveridge curve, it is not possible to reduce the numbers of jobseekers and vacancies at the same time.
-When the economy is depressed, there are many jobseekers and few vacancies; conversely, when the economy is booming, there are few jobseekers and many vacancies.
-The Beveridge curve is approximately a rectangular hyperbola:  $vu = A$, where $A$ is a constant \citep[figure 6]{MS16}.
-Hence, unemployment rate $u$ and vacancy rate $v$ are inversely related.
+\subsection{A subsection with lists}\label{s:lists}
 
-\paragraph{Efficiency criterion} Because of the symmetrical roles played by unemployed workers and vacant jobs, we find that the nonproductive use of labor is minimized when there are as many unemployed workers as vacant jobs ($u = v$).
-The labor market is inefficiently tight when there are more vacancies than jobseekers ($v > u$), and inefficiently slack when there are more jobseekers than vacancies ($u > v$).
+At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint.
 
-\paragraph{Efficient labor-market tightness} The efficiency criterion can be reformulated in terms of labor-market tightness.
-The labor-market tightness is the ratio of vacancies to unemployment, $\t = v/u$.
-It measures the number of vacancies per unemployed worker.
-Our analysis implies that the efficient labor-market tightness is $\t^* = 1$.
-The labor market is inefficiently tight when $\t > 1$ and inefficiently slack when $\t < 1$.
+Oobcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga. Here is a list:
+\begin{itemize}
+ 	\item Et harum quidem rerum facilis est et expedita distinctio.
+ 	\item Nam libero tempore, cum soluta nobis est eligendi optio.
+ 	\item emporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet.
+ 		\begin{itemize}
+ 			\item cumque nihil impedit
+ 			\item quo minus id, quod maxime placeat
+ 			\item facere possimus, omnis voluptas assumenda est
+ 		\end{itemize}
+\end{itemize}
 
-\paragraph{Efficient unemployment rate} Because the unemployment and vacancy rates play a symmetric role on the labor market, the efficient unemployment rate is the geometric average of the unemployment and vacancy rates: $u^* = \sqrt{uv}$.
-This formula arises by combining the Beveridge curve, $uv = A$, and the result that the efficient unemployment and vacancy rates are equal.
+Omnis dolor repellendus. And here is a numbered list:
+\begin{enumerate}
+ 	\item Et harum quidem rerum facilis est et expedita distinctio.
+ 	\item Nam libero tempore, cum soluta nobis est eligendi optio.
+ 	\item emporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet.
+	\begin{enumerate}
+ 			\item cumque nihil impedit
+ 			\item quo minus id, quod maxime placeat
+ 			\item facere possimus, omnis voluptas assumenda est
+ 	\end{enumerate}
+\end{enumerate}
+Itaque earum rerum hic tenetur a sapiente delectus, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat.
 
-\paragraph{Measuring the efficient unemployment rate the United States} Because it only requires unemployment and vacancy rates, the formula $u^* = \sqrt{uv}$ is easy to apply.
-In the United States, we can combine the historical unemployment and vacancy series produced by \citet{B10d} and \citet{PZ21} with the data produced by the Bureau of Labor Statistics to assemble monthly unemployment and vacancy series from 1930 and 2022.
-We use these series to compute the efficient unemployment rate $u^*$ between 1930 and 2022.
-We find that $u^*$ averages $4.2\%$ over the period and is very stable.
+\subsection{Another subsection with text}
 
-\paragraph{Application to the post-pandemic labor market} We also apply our formula in the post-pandemic period in the United States.
-There has been more vacant jobs than unemployed workers since May 2021, so the labor market has been too tight since then.
-In March 2022, there is twice as many vacant jobs as unemployed workers.
-Looking at historical data going back to 1930, we see that the last time that the labor-market tightness was so high, and the last time that the unemployment rate was so much below its efficient level, was at the end of World War 2.
-The post-pandemic labor market is in a very unusual situation: by any measure, 2022 is experiencing the tightest labor market since World War 2.
+Sed ut perspiciatis, as showed in section \ref{s:lists}, unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam eaque ipsa, quae ab illo inventore veritatis et quasi architecto beatae vitae dicta sunt, explicabo.
+Nemo enim ipsam voluptatem, quia voluptas sit, aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos, qui ratione voluptatem sequi nesciunt, neque porro quisquam est, qui dolorem ipsum, quia dolor sit amet consectetur adipiscing velit, sed quia non numquam eius modi tempora incidunt, ut labore et dolore magnam aliquam quaerat voluptatem.
+We will explore more in section \ref{s:graphs}.
 
-\paragraph{Application to modern monetary policy} During a press conference in May 2022, Jerome Powell, the chair of the Federal Reserve, discussed labor-market tightness at length.\footnote{See pp.
-12--13 at \url{https://perma.cc/4EGU-XZYG}.} In response to a question from journalist Howard Schneider asking which tightness the Fed might target, he responded: ``So in terms of the vacancy-to-unemployment ratio, we don't have a goal in mind.
-There's no specific number that we're saying, `We've got to get to that.'  \dots I think when we got to one-to-one in the, you know, in the late teens, we thought that was a pretty good number.
-But, again, we're not shooting for any particular number.'' We find that a vacancy-to-unemployment ratio of $1$ is indeed what the Fed should target.
+\subsection{Some math and theorems}
 
-\section{Derivation of the efficient-unemployment formula}\label{s:formula}
+\begin{proposition}\label{p:type1}  When the critical value is set to $z \in \mathbb{R}$, the probability of finding a type 1 error in a reported study is
+\begin{equation}
+S^*(z) = \frac{\bm{S(z)}}{ 1 -\zeta [1- \bm{S(z)}]}.
+\label{e:type1}\end{equation}
+The probability of type 1 error is larger when researchers p-hack ($S^*(z) > \bm{S(z)}$).
+In fact, the probability of type 1 error grows linearly with the expected number of p-hacking steps, which is givne by $\mathbb{E}(N(z))$:
+\begin{equation}
+S^*(z) = S(z) \times \mathbb{E}(N(z)).
+\label{e:type1steps}\end{equation}
+\end{proposition}
 
-This section derives the formula for the efficient unemployment rate: $u^* = \sqrt{uv}$.
+\begin{proof} There is a simple proof to this result. It is very simple. It is so simple. It gives equation \eqref{e:type1steps}.\end{proof}
 
-\subsection{Definition of efficiency}
+ Quis autem vel eum iure reprehenderit, qui in ea voluptate velit esse, quam nihil molestiae consequatur $\bm{\alpha(X)}$ when researchers p-hack.
+Since the probability of type 1 error with p-hacking is given by \eqref{e:type1}, the robust critical value $z^*$ is implicitly defined by
+\begin{equation*}
+\frac{\mathcal{S}(z^*)}{1 -\theta + \theta \mathcal{S}(z^*)} = \sum_{\alpha=-\infty}^{+\infty}F^{\alpha} - \Lambda(\alpha).
+\end{equation*}
+From this implicit definition we obtain the following results (details of the proof are relegated to appendix~\ref{a:appendix1}):
 
-From a social perspective, it is optimal to maximize the number of people who spend time producing useful things rather than shuffling employees or jobs.
-This perspective is consistent with the view expressed by \citet[p.
-20]{B44} that ``The material end of all human activity is consumption.
-Employment is wanted as a means to more consumption \dots as a means to a higher standard of life.'' We therefore define the socially efficient amount of unemployment as the amount that minimizes the nonproductive use of labor---both jobseeking and recruiting.
-Because people constantly move in and out of jobs, it is impossible to have zero unemployment, and it is impossible to have all workers devoted to production.
-While labor market flows are inevitable, the goal is that workers waste as little time as possible searching for jobs or employees.
+\begin{lemma}\label{p:cv} For any hypothesis test with significance level $\alpha$, the robust critical value is
+\begin{equation}
+z^* = \int_{0}^{\infty} \alpha(i) \cdot \frac{1-\beta}{1-\alpha(i)\beta}\,di.
+\label{e:cv}\end{equation}
+The robust critical value is always larger than the classical critical value $\mathcal{Z}(\alpha)$. \end{lemma}
 
-\subsection{Measuring jobseeking activity}
+\paragraph{P-hacking under the robust critical value} Ut enim ad minima veniam, quis nostrum exercitationem ullam corporis suscipit laboriosam, nisi ut aliquid ex ea commodi consequatur?
+Combining equation \eqref{e:cv} with other results, we obtain the following corollary (the proof is in appendix~\ref{a:subappendix}):
 
-The number of jobseekers is well measured by the number of unemployed workers.
-In theory, beside looking for jobs, unemployed workers might also produce useful things at home.
-Such home production would be included into aggregate production and contribute to social welfare.
-But in practice home production is minimal, as was already noted by \citet[p.
-11]{R49}: ``The most important aspect of unemployment is its wastefulness.
-It is the existence of unused productive resources side by side with unsatisfied human needs that is the intolerable condition.'' \citet[pp.
-9--11]{MS16} measure the fraction of nonwork time devoted to home production in the United States.
-The measure is based on the results by \citet{BM15}.
-Using administrative data from the US military, \citeauthor{BM15} study how servicemembers choose between reenlisting and leaving the military.
-The choices allow them to estimate the difference between market production and the sum of home production and public benefits during unemployment.
-Subtracting the value of public benefits from these estimates, \citet[p.
-11]{MS16} find that the value of home production relative to market production could be as low as $0.03$.
-Given such low value, we assume that unemployed workers do not engage in home production at all, and we count them as full-time jobseekers.
+\begin{corollary} The number of p-hacking steps under the robust critical value satisfies
+\begin{equation*}
+\mathbb{E}(N(z^*)) \approx \frac{1-\mathbb{P}(\alpha\pi)}{1-\pi}- \frac{f(y)}{z(p)^*} +P(\Gamma).
+\end{equation*}\end{corollary}
 
-\subsection{Measuring recruiting activity}
+\section{A section with graphs}\label{s:graphs}
 
-The number of recruiters is well measured by the number of vacancies.
-In theory it might take more or less than one full-time worker to service a vacancy.
-But in practice it takes about one full-time worker to service a vacancy, so the numbers of recruiters and vacancies are about the same.
-In the United States, the amount of labor required to service a vacancy can be measured from the National Employer Survey, which was conducted by the Census Bureau in 1997 \citep{V10}.
-\citet[p.
-11]{MS16} estimate that servicing a vacancy requires $0.92$ worker at any point in time.
-So it takes about $1$ worker to service a vacancy, which is what we assume here by equating the number of recruiters to the number of vacancies.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
 
-\subsection{Shape of the Beveridge curve}
+\subsection{A subsection with graphs}
 
-\paragraph{First observations of the Beveridge curve} Looking at labor-market statistics for the United Kingdom, \citet{B44} first noted that the number of vacancies and the number of jobseekers move in opposite directions.
-When the economy is in a slump, there are lots of jobseekers and few vacancies.
-Conversely, when the economy is in a boom, there are few jobseekers and many vacancies.
-This negative relationship between unemployment and vacancies---known as the Beveridge curve---was first plotted by \citet[figures 1 and 2]{DD58} using UK data for 1946--1956.
-Moreover, \citet[p.
-22]{DD58} noted that the Beveridge curve looked like a rectangular hyperbola in an unemployment-vacancy plane.
+At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
+Et harum quidem rerum facilis est et expedita distinctio.
+Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus, omnis voluptas assumenda est, omnis dolor repellendus.
+Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet, ut et voluptates repudiandae sint et molestiae non recusandae.
+Itaque earum rerum hic tenetur a sapiente delectus, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat.
 
 \begin{figure}[t]
-\subcaptionbox{Regular scale\label{f:regular1951}}{\includegraphics[scale=\sfig,page=1]{\pdf}}\hfill
-\subcaptionbox{Log scale\label{f:log1951}}{\includegraphics[scale=\sfig,page=2]{\pdf}}
-\caption{Unemployment and vacancy rates in the United States, 1951--2019}
-\note{The unemployment rate is constructed by the \citet{UNRATE}.
-For 1951--2000, the vacancy rate is constructed by \citet{B10d}; for 2001--2019, the vacancy rate is the number of job openings divided by the civilian labor force, both measured by the \citet{CLF16OV,JTSJOL}.
-Unemployment and vacancy rates are quarterly averages of monthly series.
-The gray areas are NBER-dated recessions.}
-\label{f:1951}\end{figure}
+\subcaptionbox{First panel\label{f:panel1}}{\includegraphics[scale=\sfig,page=1]{\pdf}}\hfill
+\subcaptionbox{Second panel\label{f:panel2}}{\includegraphics[scale=\sfig,page=2]{\pdf}}
+\caption{A caption for the graph}
+\note{A note for the graph. Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus.}
+\label{f:graph1}\end{figure}
 
-\paragraph{Unemployment and vacancy rates move in opposite directions} The Beveridge curve holds remarkably well in the United States as well \citep{BD89,EMR15}.
-Figure~\ref{f:regular1951} depicts the unemployment rate $u$ (all jobseekers divided by the labor force) and the vacancy rate $v$ (all vacancies divided by the same labor force) in the United States for 1951--2019.
-The unemployment rate is constructed by the \citet{UNRATE}.
-For 1951--2000, the vacancy rate is constructed by \citet{B10d}; for 2001--2019, the vacancy rate is the number of job openings divided by the civilian labor force, both measured by the \citet{CLF16OV,JTSJOL}.
-The figure shows clearly that unemployment and vacancy rates move in opposite directions.
+\paragraph{Description of a result} Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus, omnis voluptas assumenda est, omnis dolor repellendus.
+Figure~\ref{f:graph1} depicts the result.
 
-\paragraph{Unemployment and vacancy rates are inversely related} In fact, unemployment and vacancy appear to be the inverse of each other, so that doubling the unemployment rate cuts the vacancy rate in half, and inversely, doubling the vacancy rate cuts the unemployment rate in half.
-Figure \ref{f:log1951} displays again unemployment and vacancy, but now in log scale.
-The fluctuations of the unemployment and vacancy rates are almost the mirror image of each other.
-This empirical regularity is particularly striking because the two series are constructed independently of each other.
+\paragraph{Another result in a panel} In fact, nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus, omnis voluptas assumenda est, omnis dolor repellendus.
+Figure \ref{f:panel1} displays again this fact, but now in log scale.
+This empirical regularity is particularly striking in figure \ref{f:panel2} because the two series are constructed independently of each other.
+
+\subsection{A full-page graphs}
+
+A full-page graph is on figure \ref{f:graph2}.
+Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus, omnis voluptas assumenda est, omnis dolor repellendus.
 
 \begin{figure}[p]
-\subcaptionbox{1951Q1--1961Q1: elasticity $= -0.85$\label{f:branch1}}{\includegraphics[scale=\sfig,page=3]{\pdf}}\hfill
-\subcaptionbox{1961Q2--1971Q4: elasticity $= -1.02$\label{f:branch2}}{\includegraphics[scale=\sfig,page=4]{\pdf}}\vfig
-\subcaptionbox{1972Q1--1989Q1: elasticity $= -0.84$\label{f:branch3}}{\includegraphics[scale=\sfig,page=5]{\pdf}}\hfill
-\subcaptionbox{1989Q2--1999Q2: elasticity $= -0.94$\label{f:branch4}}{\includegraphics[scale=\sfig,page=6]{\pdf}}\vfig
-\subcaptionbox{1999Q3--2009Q3: elasticity $= -1.00$\label{f:branch5}}{\includegraphics[scale=\sfig,page=7]{\pdf}}\hfill
-\subcaptionbox{2009Q4--2019Q4: elasticity $= -0.84$\label{f:branch6}}{\includegraphics[scale=\sfig,page=8]{\pdf}}
-\caption{Estimated Beveridge curve in the United States, 1951--2019}
-\note{This figure is constructed from figures 5 and 6 in \citet{MS16}.
-Unemployment and vacancy rates are the same as in figure~\ref{f:1951}.
-The number of structural breaks, dates of structural breaks, and elasticities of the Beveridge curve between breaks are estimated with the algorithm of \citet{BP98,BP03}.}
-\label{f:hyperbola}\end{figure}
+\subcaptionbox{Caption for panel}{\includegraphics[scale=\sfig,page=1]{\pdf}}\hfill
+\subcaptionbox{Caption for panel}{\includegraphics[scale=\sfig,page=2]{\pdf}}\vfig
+\subcaptionbox{Caption for panel}{\includegraphics[scale=\sfig,page=3]{\pdf}}\hfill
+\subcaptionbox{Caption for panel}{\includegraphics[scale=\sfig,page=4]{\pdf}}\vfig
+\subcaptionbox{Caption for panel}{\includegraphics[scale=\sfig,page=5]{\pdf}}\hfill
+\subcaptionbox{Caption for panel}{\includegraphics[scale=\sfig,page=1]{\pdf}}
+\caption{A caption for the graph}
+\note{A note for the graph. Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat. The figures are borrowed from \citet[section 3]{MS22b}.}
+\label{f:graph2}\end{figure}
 
-\paragraph{The Beveridge curve is a rectangular hyperbola} Mathematically, the property that unemployment and vacancy rates are inversely related implies that in an unemployment-vacancy plane, the Beveridge curve is a rectangular hyperbola:
-\begin{equation*}
-vu = A,
-\end{equation*}
-where $A>0$ is a constant.
+\paragraph{Informal description of the results} At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
+Et harum quidem rerum facilis est et expedita distinctio.
+Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus, omnis voluptas assumenda est, omnis dolor repellendus.
+Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet, ut et voluptates repudiandae sint et molestiae non recusandae.
+Itaque earum rerum hic tenetur a sapiente delectus, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat.
 
-\paragraph{Estimates of the Beveridge elasticity} We can formally establish that the Beveridge curve is a rectangular hyperbola by estimating the elasticity of the vacancy rate with respect to the unemployment rate, $\odlx{v}{u}$.
-An elasticity of $-1$ corresponds to an hyperbola.
-\citet[figures 5 and 6]{MS16} use the algorithm of \citet{BP98,BP03} to estimate the structural breaks of the US Beveridge curve, and the elasticity of the Beveridge curve between these breaks.
-They find that the Beveridge elasticity remains between $-0.84$ and $-1.02$, so never far from $-1$ (figure~\ref{f:hyperbola}).
-This finding confirms that the Beveridge curve is close to an hyperbola.
+\paragraph{Formal description of the results} At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
+Et harum quidem rerum facilis est et expedita distinctio.
+Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus, omnis voluptas assumenda est, omnis dolor repellendus.
+Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet, ut et voluptates repudiandae sint et molestiae non recusandae.
+Itaque earum rerum hic tenetur a sapiente delectus, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat.
 
-\paragraph{Informal foundation for the hyperbolic Beveridge curve} Why is the Beveridge curve an hyperbola?
-Empirically, the number of new matches depends symmetrically on the number of jobseekers and vacancies.
-Doubling jobseekers increases new matches much the same as doubling vacancies.
-This makes sense as both jobseekers and vacancies are striving to find each other.
-This is obvious for jobseekers who spend time looking for jobs, applying, doing interviews.
-But this is also the case for vacancies as recruiters have to advertise vacancies, and screen and interview applicants.
-Therefore, as new matches have to meet the number of job separations, cutting in half the number of jobseekers requires doubling the number of vacancies, explaining why unemployment and vacancy rates are the inverse of each other.
+\section{A section with table}
 
-\paragraph{Formal foundation for the hyperbolic Beveridge curve} In matching models the Beveridge curve says labor market flows are balanced: the number of job separations equals the number of jobs found.
-The employment rate $1-u$ is approximately constant around 1 since the unemployment rate $u$ is an order of magnitude less than 1.
-The job-separation rate $\l$ is also constant, so the number of job separations $\l (1-u)$ is approximately constant around $\l$.
-So at any point of the business cycle, the number of jobs found is approximately constant at $\l$.
-With the standard symmetric Cobb-Douglas matching function, $m = \o \cdot\sqrt{uv}$, the number of jobs found at any point in time is proportional to $\sqrt{uv}$.
-\footnote{The US matching function appears to have a Cobb-Douglas form with exponents of $0.5$ on unemployment and vacancies.
-See \citet[p.
-9]{MS16} for a survey of the US estimates based on aggregate data, and \citet{PP01} for a broader survey.} Hence $\sqrt{uv}$ is constant along the Beveridge curve, which implies $uv = \text{constant}$ along the Beveridge curve.
+Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet, ut et voluptates repudiandae sint et molestiae non recusandae.
+Itaque earum rerum hic tenetur a sapiente delectus, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat.
 
-\subsection{Efficiency criterion}
+\subsection{The table}
 
-The social objective is to minimize the time spent on jobseeking and recruiting at the expense of producing.
-Since the amount of jobseeking can be measured by the number of unemployed workers, and the amount of recruiting by the number of vacancies, the social objective is to minimize the sum of the unemployment and vacancy rates, $u + v$.
-This minimization is subject to the Beveridge curve constraint, $u v = \text{constant}$.
-Because of the Beveridge curve, it is not possible to reduce unemployment and vacancies at the same time.
-Therefore, the social planner must trade off unemployment and vacancies.
+Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet, ut et voluptates repudiandae sint et molestiae non recusandae.
+Itaque earum rerum hic tenetur a sapiente delectus, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat.
+Table \ref{t:table} shows many things.
 
-\paragraph{Symmetry argument} Because of the symmetrical roles played by unemployed workers and vacant jobs, the nonproductive use of labor is minimized when there are as many unemployed workers as vacant jobs ($u = v$).
-Indeed, we would like to minimize wasted resources from unemployment and from vacancies, subject to the Beveridge curve.
-Since this problem is perfectly symmetric in $u$ and $v$, the solution to minimize waste is to have $u = v$.
+At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
 
-\paragraph{Departure from efficiency} When the number of jobseekers is not equal to the number of vacancies, the labor market is operating inefficiently.
-The labor market is inefficiently tight when there are more vacancies than jobseekers ($v > u$).
-In that case, increasing $u$ and reducing $v$ would reduce waste.
-The labor market is inefficiently slack when there are more jobseekers than vacancies ($u > v$).
-Then, reducing $u$ and increasing $v$ would reduce waste.
-
-\paragraph{First-order condition} We can also establish these results mathematically.
-The planner minimizes nonproduction $u + v$ subject to Beveridge curve $uv = A$, or $v = A/u$.
-This is equivalent to minimizing $u+A/u$, which is convex in $u$.
-A first-order condition is necessary and sufficient to find the minimum.
-We take the derivative of $u+A/u$ with respect to $u$ and set it to $0$.
-We obtain $1-A/u^2 = 0$.
-This condition implies that the minimum occurs when $u = \sqrt{A}$.
-By the Beveridge curve we have $v = A/u$, so at the minimum $v = A/ \sqrt{A} = \sqrt{A}$.
-Accordingly, the efficient unemployment and vacancy rates satisfy $u^* = v^* = \sqrt{A}$.
-
-\paragraph{Formulation in terms of labor-market tightness} The efficiency criterion can be reformulated in terms of labor-market tightness.
-The labor-market tightness is the ratio of vacancies to unemployment, $\t = v/u$.
-It measures the number of vacancies per unemployed workers.
-The analysis implies that the efficient labor-market tightness is $\t^* = 1$; that the labor market is inefficiently tight when $\t > 1$; and that the labor market is inefficiently slack when $\t < 1$.
-
-\begin{figure}[t]
-\subcaptionbox{Unemployment and vacancy rates\label{f:uv1951}}{\includegraphics[scale=\sfig,page=9]{\pdf}}\hfill
-\subcaptionbox{Labor-market tightness\label{f:tightness1951}}{\includegraphics[scale=\sfig,page=10]{\pdf}}
-\caption{State of the labor market in the United States, 1951--2019}
-\note{The unemployment and vacancy rates come from figure~\ref{f:1951}.
-The labor-market tightness is the vacancy rate divided by the unemployment rate.
-The gray areas are NBER-dated recessions.
-The labor market is inefficiently slack whenever the unemployment rate is above the vacancy rate (purple shade); it is inefficiently tight whenever the unemployment rate is below the vacancy rate (orange shade).
-Between 1951 and 2019, the US labor market is always inefficiently slack except during three episodes: the Korean War (1951Q1--1953Q3), the Vietnam War (1965Q4--1970Q1), and the end of the Trump presidency (2018Q1--2019Q4).}
-\label{f:state1951}\end{figure}
-
-\subsection{Efficient unemployment rate}
-
-We have seen that the efficient unemployment rate is given by $u^* = \sqrt{A}$ where the parameter $A$ determines the location of Beveridge curve, $uv = A$.
-Accordingly, the efficient unemployment rate is the geometric average of the unemployment and vacancy rates: $u^* = \sqrt{uv}$.
-Intuitively, the efficient unemployment rate is the average of the unemployment and vacancy rates because unemployment and vacancy rates play a symmetric role on the labor market.
-
-\subsection{Similarity with previous markers of full employment}
-
-Our criterion provide a theoretical underpinning for the old idea that the state of full employment could be assessed by comparing the number of vacant jobs to that of unemployed workers.
-For example,  \citet[p.
-18]{B44} states that ``Full employment \dots means having always more vacant jobs than unemployed men,'' and before his report, the prevailing view was that full employment was ``a state of affairs in which the number of unfilled vacancies is not appreciably below the number of unemployed persons.'' \citet[p.
-39 and chart 5]{R57} applies Beveridge' s verbal criterion to the United States.
-He computes the ratio between number of vacancies and number of unemployed and examines when the ratio crosses $1$.
-In the same spirit, the \citet{BLS22} scrutinizes the jobseeker-per-job-opening statistic and flags when it crosses $1$.
-
-\begin{figure}[t]
-\subcaptionbox{Efficient unemployment rate\label{f:efficient1951}}{\includegraphics[scale=\sfig,page=11]{\pdf}}\hfill
-\subcaptionbox{Unemployment gap\label{f:gap1951}}{\includegraphics[scale=\sfig,page=12]{\pdf}}
-\caption{Efficient unemployment in the United States, 1951--2019}
-\note{The unemployment rate $u$ and vacancy rate $v$ come from figure~\ref{f:1951}.
-The efficient unemployment rate is $u^* = \sqrt{uv}$.
-The unemployment gap is the difference between the actual unemployment rate and the efficient unemployment rate, $u-u^*$.
-The gray areas are NBER-dated recessions.}
-\label{f:squareroot1951}\end{figure}
-
-\begin{figure}[t]
-\includegraphics[scale=\sfig,page=13]{\pdf}
-\caption{Comparison with efficient unemployment from \citet{MS16}}
-\note{The Michaillat \& Saez (2021) line represents the efficient unemployment rate constructed by \citet[figure 7B]{MS16}.
-The $\sqrt{uv}$ line reproduces the efficient unemployment rate from figure~\ref{f:efficient1951}.
-The gray areas are NBER-dated recessions.}
-\label{f:comparison}\end{figure}
-
-\section{Efficient unemployment in the United States}\label{s:postwar}
-
-We compute the efficient unemployment rate in the United States in three periods: the standard postwar period, 1951--2019; the coronavirus pandemic, 2020--2022; and the Great Depression and World War 2, 1930--1950.
-
-\subsection{Postwar period}
-
-\paragraph{State of the labor market} We assess the state of the US labor market between 1951 and 2019.
-We use the unemployment and vacancy rates from figure~\ref{f:1951}.
-The labor market is inefficiently slack whenever the unemployment rate is above the vacancy rate; it is inefficiently tight whenever the unemployment rate is below the vacancy rate.
-Between 1951 and 2019, the US labor market is always inefficiently slack except in three episodes (figure~\ref{f:uv1951}): the Korean War (1951Q1--1953Q3), the Vietnam War (1965Q4--1970Q1), and the end of the Trump presidency (2018Q1--2019Q4).
-
-\paragraph{Visualization with labor-market tightness} The state of the US labor market can also be visualized by plotting the labor-market tightness (figure~\ref{f:tightness1951}).
-The labor market is inefficiently slack whenever tightness is below $1$.
-It is inefficiently tight whenever tightness is above $1$.
-And it is efficient when the tightness equals $1$---when there is just one vacancy per jobseeker.
-The tightness averages $0.62$ between 1951 and 2019, and it is sharply procyclical.
-The tightness peaked at $1.49$ in 1953Q1, during the Korean War, and it bottomed at $0.16$ in 2009Q3, during the Great Recession.
-
-\paragraph{Efficient unemployment rate} We then compute the efficient unemployment rate between 1951 and 2019 using the formula $u^* = \sqrt{uv}$.
-The efficient unemployment rate is fairly stable: it remains between $3.0\%$ and $5.3\%$, with an average value of $4.2\%$ (figure~\ref{f:efficient1951}).
-
-\paragraph{Unemployment gap} We also compute the unemployment gap, which is the difference $u-u^*$ between the actual unemployment rate and the efficient unemployment rate.
-The unemployment gap indicates the distance from full employment at any given time.
-The unemployment gap is sharply countercyclical and generally positive.
-The unemployment gap peaked at $+6.0$ percentage points in 2009Q4, during the Great Recession, and bottomed at $-0.6$ percentage points in 1969Q1, during the Vietnam War.
-The unemployment gap averages $1.6$ percentage points over the 1951--2019 period.
-Hence, the economy is generally not at full employment, and it is especially far from full employment in recessions.
-
-\paragraph{Comparison with the formula from \citet{MS16}} In a previous paper, \citet[proposition 3]{MS16} compute a more sophisticated formula for the efficient unemployment rate:
-\begin{equation}
-u^* = \bp{\frac{\k \e}{1-\z}\cdot\frac{v}{u^{-\e}}}^{1/(1+\e)},
-\label{e:ms16}\end{equation}
-where  $\e$ is the Beveridge elasticity, $\z$ is the social value of nonwork, and $\k$ is the recruiting cost.
-The formula requires to keep track of three statistics in addition to the unemployment and vacancy rates ($\e$, $\z$, $\k$), so it is harder for policymakers to compute than the simple $u^* = \sqrt{uv}$.
-But $u^* = \sqrt{uv}$ is a special case of \eqref{e:ms16} when $\e=1$, $\k=1$, and $\z=0$.
-The two formulas are so closely related because the model of the labor market presented here is a special case of the model in \citet{MS16}.
-By specializing the model and therefore the values of the sufficient statistics, we obtain an extremely simple and user-friendly formula.
-In the United States, however, the two formulas yield almost identical unemployment rates (figure~\ref{f:comparison}).
-Over 1951--2019, the two efficient unemployment rates depart by $0.2$ percentage points on average, and they never depart by more than $0.5$ percentage points.
-
-\begin{figure}[t]
-\subcaptionbox{Unemployment and vacancy rates\label{f:uv2020}}{\includegraphics[scale=\sfig,page=14]{\pdf}}\hfill
-\subcaptionbox{Labor-market tightness\label{f:tightness2020}}{\includegraphics[scale=\sfig,page=15]{\pdf}}
-\caption{State of the labor market in the United States, 2020M1--2022M3}
-\note{The unemployment rate is constructed by the \citet{UNRATE}.
-The vacancy rate is the number of job openings divided by the civilian labor force, both measured by the \citet{CLF16OV,JTSJOL}.
-The labor-market tightness is the vacancy rate divided by the unemployment rate.
-The gray area is the NBER-dated pandemic recession.
-The labor market is inefficiently slack whenever the unemployment rate is above the vacancy rate (purple shade); it is inefficiently tight whenever the unemployment rate is below the vacancy rate (orange shade).
-The US labor market is inefficiently slack from 2020M3 to 2021M4; it is inefficiently tight in 2020M1--2020M2 and from 2021M5 to 2022M3.}
-\label{f:state2020}\end{figure}
-
-\begin{figure}[t]
-\subcaptionbox{Efficient unemployment rate\label{f:efficient2020}}{\includegraphics[scale=\sfig,page=16]{\pdf}}\hfill
-\subcaptionbox{Unemployment gap\label{f:gap2020}}{\includegraphics[scale=\sfig,page=17]{\pdf}}
-\caption{Efficient unemployment in the United States, 2020M1--2022M3}
-\note{The unemployment rate $u$ and vacancy rate $v$ come from figure~\ref{f:uv2020}.
-The efficient unemployment rate is $u^* = \sqrt{uv}$.
-The unemployment gap is the difference between the actual unemployment rate and the efficient unemployment rate, $u-u^*$.
-The gray area is the NBER-dated pandemic recession.}
-\label{f:squareroot2020}\end{figure}
-
-\begin{figure}[t]
-\includegraphics[scale=\sfig,page=18]{\pdf}
-\caption{Effect of the pandemic on the US Beveridge curve}
-\note{Each dot represents the unemployment and vacancy rates in a month between 2001M1 and 2022M3.
-The unemployment rate is constructed by the \citet{UNRATE}.
-The vacancy rate is the number of job openings divided by the civilian labor force, both measured by the \citet{CLF16OV,JTSJOL}.
-The labor market is inefficiently slack when the unemployment rate is above the vacancy rate (purple shade); it is inefficiently tight when the unemployment rate is below the vacancy rate (orange shade); it is efficient when the unemployment rate equals the vacancy rate (pink line).}
-\label{f:pandemic}\end{figure}
-
-\subsection{Coronavirus pandemic}\label{s:pandemic}
-
-We now apply our efficiency criterion and efficient-unemployment formula to the coronavirus pandemic in the United States. We focus on the period between January 2020 and March 2022.
-
-\paragraph{Unemployment and vacancy rates} The unemployment rate is constructed by the \citet{UNRATE}.
-The vacancy rate is the number of job openings divided by the civilian labor force, both measured by the \citet{CLF16OV,JTSJOL}.
-Both series are displayed on figure~\ref{f:uv2020}.
-
-\paragraph{State of the labor market} The labor market is inefficiently slack whenever the unemployment rate is above the vacancy rate; it is inefficiently tight whenever the unemployment rate is below the vacancy rate.
-We find that the US labor market is inefficiently slack from March 2020 to April 2021; it is inefficiently tight from May 2021 to March 2022 (figure~\ref{f:uv2020}).
-
-\paragraph{Visualization with labor-market tightness} The state of the US labor market can also be visualized by plotting labor-market tightness (figure~\ref{f:tightness2020}).
-The labor market is inefficiently slack whenever tightness is below $1$.
-And it is inefficiently tight whenever tightness is above $1$.
-The tightness cratered to $0.20$ in April 2020, at the beginning of the pandemic.
-It then steadily recovered to reach $2.0$ in March 2022---the highest tightness recorded in the postwar period.
-
-\paragraph{Efficient unemployment rate} We then compute the efficient unemployment rate between 2020 and 2022 using the formula $u^* = \sqrt{uv}$.
-The efficient unemployment rate was at $4.0\%$ at the onset of the pandemic, but it sharply increased above $6.0\%$ in the first months of the pandemic (figure~\ref{f:efficient2020}).
-It hovered around $6.0\%$ during the rest of the 2020--2022 period.
-
-\paragraph{Unemployment gap} We also compute the unemployment gap, which is the difference $u-u^*$ between the actual unemployment rate and the efficient unemployment rate (figure~\ref{f:gap2020}).
-The unemployment gap was initially positive and large: the labor market was much too slack in the first pandemic year.
-The unemployment gap reached $+8.1$ percentage points in April 2020---its highest level in the postwar period.
-But the economy recovered quickly, and the unemployment gap turned negative in May 2021.
-The unemployment gap reached $-1.5$ percentage points in March 2022.
-This is the lowest unemployment gap in the postwar period.
-So in 2022, the labor market is well beyond full employment.
-
-\paragraph{Why did the efficient unemployment rate increase during the pandemic?} The efficient unemployment rate increased by more than 2 percentage points at the onset of the pandemic.
-Such a sharp increase is unprecedented.
-It can be explained by the gigantic outward shift of the Beveridge curve that took place in the spring 2020.
-Graphically, the efficient unemployment rate appears at the intersection of the Beveridge curve and the pink identity line (figure~\ref{f:pandemic}).
-In February 2020, at the onset of the pandemic, the labor market was close to efficient, and the unemployment rate was around $4\%$.
-In May 2021, about a year later, the labor market had returned to efficiency, but the unemployment rate was now $6\%$.
-The increase of the efficient unemployment rate from $4\%$ to $6\%$ was caused by the outward shit of the Beveridge curve that occurred between March and April 2020.
-Mathematically, the efficient unemployment rate is determined by the location of the Beveridge curve ($u^* = \sqrt{A}$), so only a sharp outward shift of the Beveridge curve (increase in $A$) can increase the efficient unemployment rate.\footnote{The previous shifts of the Beveridge curve during the postwar period (figure~\ref{f:hyperbola}) were not large enough to produce noteworthy changes in the efficient unemployment rate (figure~\ref{f:squareroot1951}).}
-
-\begin{figure}[t]
-\subcaptionbox{Regular scale\label{f:regular1930}}{\includegraphics[scale=\sfig,page=19]{\pdf}}\hfill
-\subcaptionbox{Log scale\label{f:log1930}}{\includegraphics[scale=\sfig,page=20]{\pdf}}
-\caption{Unemployment and vacancy rates in the United States, 1930--1950}
-\note{The unemployment and vacancy rates are constructed by \citet{PZ21}. Unemployment and vacancy rates are quarterly averages of monthly series. The gray areas are NBER-dated recessions.}
-\label{f:1930}\end{figure}
-
-\begin{figure}[t]
-\subcaptionbox{Unemployment and vacancy rates\label{f:uv1930}}{\includegraphics[scale=\sfig,page=21]{\pdf}}\hfill
-\subcaptionbox{Labor-market tightness\label{f:tightness1930}}{\includegraphics[scale=\sfig,page=22]{\pdf}}
-\caption{State of the labor market in the United States, 1930--1950}
-\note{The unemployment and vacancy rates come from figure~\ref{f:1930}.
-The labor-market tightness is the vacancy rate divided by the unemployment rate.
-The gray areas are NBER-dated recessions.
-The labor market is inefficiently slack whenever the unemployment rate is above the vacancy rate (purple shade); it is inefficiently tight whenever the unemployment rate is below the vacancy rate (orange shade).
-Between 1930 and 1950, the US labor market is always inefficiently slack except during World War 2 (1942Q3--1946Q3).}
-\label{f:state1930}\end{figure}
-
-\begin{figure}[t]
-\subcaptionbox{Efficient unemployment rate\label{f:efficient1930}}{\includegraphics[scale=\sfig,page=23]{\pdf}}\hfill
-\subcaptionbox{Unemployment gap\label{f:gap1930}}{\includegraphics[scale=\sfig,page=24]{\pdf}}
-\caption{Efficient unemployment in the United States, 1930--1950}
-\note{The unemployment rate $u$ and vacancy rate $v$ come from figure~\ref{f:1930}.
-The efficient unemployment rate is $u^* = \sqrt{uv}$.
-The unemployment gap is the difference between the actual unemployment rate and the efficient unemployment rate, $u-u^*$.
-The gray areas are NBER-dated recessions.}
-\label{f:squareroot1930}\end{figure}
-
-\subsection{Great Depression and World War 2}\label{s:depression}
-
-Finally, we apply our efficiency criterion and efficient-unemployment formula to the historical period 1930--1950, which covers both the Great Depression and World War 2.
-
-\paragraph{Unemployment and vacancy rates} The unemployment and vacancy rates are constructed by \citet{PZ21} from NBER macrohistory files.
-Between 1930 and 1950 it remains true that unemployment and vacancy rates move in opposite directions (figure~\ref{f:regular1930}).
-Using a log scale, it also appears that unemployment and vacancy rates are inversely related (figure~\ref{f:log1930}).
-These fluctuations indicate that just as in the postwar era, the hyperbolic Beveridge curve holds in 1930--1950.\footnote{See also figure 3 in \citet{PZ21}.}
-
-\paragraph{State of the labor market} The labor market is inefficiently slack whenever the unemployment rate is above the vacancy rate; it is inefficiently tight whenever the unemployment rate is below the vacancy rate.
-The US labor market is always inefficiently slack between 1930 and 1950 except at the end World War 2, during 1942Q3--1946Q3 (figure~\ref{f:uv1930}).
-
-\paragraph{Visualization with labor-market tightness} The state of the labor market can also be visualized by plotting labor-market tightness (figure~\ref{f:tightness1930}).
-The labor market is inefficiently slack whenever tightness is below $1$; it is inefficiently tight whenever tightness is above $1$.
-Tightness is extremely volatile during the 1930--1950 period: it plunged to $0.03$ in 1932Q3, during the Great Depression, and climbed all the way to $6.8$ in 1944Q4, at the end of World War 2.
-These are the most extreme tightness fluctuations on record.
-
-\paragraph{Efficient unemployment rate} We then compute the efficient unemployment rate between 1930 and 1950 using the formula $u^* = \sqrt{uv}$.
-The efficient unemployment rate is again fairly stable: it remains between $2.5\%$ and $4.6\%$, with an average value of $3.5\%$ (figure~\ref{f:efficient1930}).
-This value of the efficient unemployment rate is actually pretty close from the full-employment rate of unemployment advocated by \citet[p.
-21]{B44} just after World War 2, which was $3\%$.
-
-\paragraph{Unemployment gap} Finally, we compute the unemployment gap, which is the difference $u-u^*$ between the actual unemployment rate and the efficient unemployment rate (figure~\ref{f:gap1930}).
-The unemployment gap was of course positive and very large during the Great Depression: the labor market was much too slack then.
-The unemployment gap reached $+20.9$ percentage points in 1932Q3---its highest level on record.
-The economy recovered only slowly from the Great Depression .
-The unemployment gap finally turned negative during World War 2, in 1942Q3.
-The unemployment gap reached $-1.6$ percentage points in 1945Q1---its lowest level on record.
-The unemployment gap turned positive again during the 1948--1949 recession.
-
-\newlength{\imageheight}
-\settoheight{\imageheight}{\includegraphics[page=24]{\pdf}}
-\begin{figure}[p]
-\subcaptionbox{Unemployment and vacancy rates, $u$ and $v$\label{f:uv}}{\includegraphics[trim=0 0 0 0.5\imageheight, clip, scale=\lfig,page=25]{\pdf}}\vfig
-\subcaptionbox{Labor-market tightness, $v/u$\label{f:tightness}}{\includegraphics[trim=0 0 0 0.5\imageheight, clip, scale=\lfig,page=26]{\pdf}}\vfig
-\subcaptionbox{Efficient unemployment rate, $u^* = \sqrt{uv}$ \label{f:squareroot}}{\includegraphics[trim=0 0 0 0.5\imageheight, clip, scale=\lfig,page=27]{\pdf}}
-\caption{State of the US labor market, 1930Q1--2022Q1}
-\note{The unemployment and vacancy rates are obtained by splicing the unemployment and vacancy rates from figures~\ref{f:uv1951}, \ref{f:uv2020}, and \ref{f:uv1930}.
-Labor-market tightness and efficient unemployment rate are constructed from the unemployment and vacancy rates.
-The gray areas are NBER-dated recessions.}
-\label{f:summary}\end{figure}
+\begin{table}[t]
+\caption{A caption for the table}
+\begin{tabular*}{\textwidth}[]{p{3.3cm}@{\extracolsep\fill}cccc}
+\toprule
+& Timeline & Natural rate & Monetary  & Government \\
+&  & of interest &  policy & spending\\
+\midrule
+\multicolumn{5}{l}{A. ZLB episode}\\
+ZLB: & $t\in (0,T) $ & $r^n<0$ & $i=0$  & -- \\
+Normal times: & $t>T$ &  $r^n>0$ & $i= r^n + \phi \pi$  & --   \\
+\midrule
+\multicolumn{5}{l}{B. ZLB episode with forward guidance}\\
+ZLB: & $t\in (0,T)$ & $r^n<0$ & $i=0$  & -- \\
+Forward guidance: & $t\in (T,T+\Delta)$ & $r^n>0$ & $i=0$  & --\\
+Normal times: & $t>T+\Delta$ & $r^n>0$ & $i= r^n + \phi \pi$  & -- \\
+\midrule
+\multicolumn{5}{l}{C. ZLB episode with government spending}\\
+ZLB: & $t\in (0,T) $ & $r^n<0$ & $i=0$  & $g>0$  \\
+Normal times: & $t>T$ & $r^n>0$ & $i= r^n + \phi \pi$  & $g=0$   \\
+\bottomrule
+\end{tabular*}
+\note{This table describes three scenarios: Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet, ut et voluptates repudiandae sint et molestiae non recusandae.
+Ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat.
+The table is borrowed from \citet[table 1]{MS21a}.}
+\label{t:table}\end{table}
 
@@
 To summarize, we combine the US data for 1930Q1--2022Q1.
+Qui dolorem eum fugiat, quo voluptas nulla pariatur?
+At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
+Et harum quidem rerum facilis est et expedita distinctio.
+Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus, omnis voluptas assumenda est, omnis dolor repellendus.
+Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet, ut et voluptates repudiandae sint et molestiae non recusandae.
+Itaque earum rerum hic tenetur a sapiente delectus, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat (figure~\ref{f:graph2}).
+Over 1930Q1--2022Q1, labor-market tightness averages $0.68$.
+In 2022, tightness reached a value of 2.0, which it had last reached in 1945.The efficient unemployment rate is stable over time.
+It hovered around 4\%--5\% between 1930 and 1970.
 
-\paragraph{State of the labor market} The unemployment rate is generally above the vacancy rate, and this gap is exacerbated in recessions (figure~\ref{f:uv}).
-This means that the labor market does not operate efficiently, unlike what economists often assume.
-It also means that monetary and fiscal policies are needed to reduce unemployment in bad times, and adhere more closely to the full-employment mandate.
-Before 2018, the labor market is only inefficiently tight in wartime---World War 2, Korean War, Vietnam War.
-Since then the labor market has been inefficiently tight at the end of the Trump presidency and during the recovery of the coronavirus pandemic.
-The state of the labor market in 2018--2019 and 2021--2022 is a rarity: it is the only peacetime labor market that is exceedingly tight.
-\citet[p.
-322]{K36} doubted that the labor market could reach full employment in peacetime.
-It seems that he was essentially right: before 2018 the US labor market never reached full employment in peacetime.
-
-\paragraph{Labor-market tightness} The state of the labor market can also be visualized by plotting labor-market tightness (figure~\ref{f:tightness}).
-Over 1930Q1--2022Q1, labor-market tightness averages $0.68$.
-In 2022, tightness reached a value of $2.0$, which it had last reached in 1945.
-In the recovery of the pandemic, the US labor market has reached historical levels of tightness.
-
-\paragraph{Full employment} Over 1930Q1--2022Q1, the efficient unemployment rate averages $4.2\%$ (figure~\ref{f:squareroot}).
-The efficient unemployment rate is stable over time.
-It hovered around $4\%$ between 1930 and 1970.
-It raised to about $5\%$ in the 1970s and stayed there in the 1980s.
-It then remained around $4\%$ again between 1990 and 2020.
-Finally, it rose to $6\%$ during the pandemic.
-This means that in the United States, the full-employment unemployment rate is stable over time, around $4\%$.
-
-\section{Frequently asked questions}\label{s:faq}
+\section{Conclusion}\label{s:ccl}
 
@@
 
-\subsection{Shouldn't unemployment dynamics play a role?}
+At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
 
-\paragraph{Informal argument} Unemployment perpetually evolves through a dynamic process driven by differences in inflow into and outflow from unemployment.
-Yet, our analysis of labor-market efficiency is static.
-But because gross flows are so large relative to net flows---$4\%$ of employment per month compared to a fraction of a percent of employment per month---this dynamic unemployment process converges extremely quickly to the Beveridge curve.
-This explains why the static analysis is an excellent approximation to the dynamic process.
-
- \paragraph{Formal argument in a matching model \citep[p. 7]{MS16}} In a matching model the unemployment rate evolves according to the following differential equation:
-\begin{equation}
-\dot{u}(t) = \l [1-u(t)] - f u(t),
-\label{e:uDot}\end{equation}
-where $\l$ is the job-separation rate and $f$ is the job-finding rate.
-The term $\l [1-u(t)]$ gives the number of workers who lose or quit their jobs and enter unemployment during a unit time.
-The term $f u(t)$ gives the number of workers who find a job and leave unemployment during a unit time.
-The difference between inflow into and outflow from unemployment determines the change in the unemployment rate, $\dot{u}(t)$.
-The solution to the differential equation~\eqref{e:uDot} is
-\begin{equation}
-u(t) - u^b = [u(0) - u^b] e^{-(\l+f)t},
-\label{e:uSolution}\end{equation}
-where
-\begin{equation*}
-u^b = \frac{\l}{\l+f}.
-\end{equation*}
-The unemployment rate $u^b$ is the critical point of differential equation~\eqref{e:uDot}, so it corresponds to the unemployment rate given by the Beveridge curve when the job-finding rate is $f$.
-Equation \eqref{e:uSolution} shows that the distance between the unemployment rate $u(t)$ and the Beveridgean unemployment rate $u^b$ decays at an exponential rate $\l+f$.
-In the United States, the rate of decay is really fast: between 1951 and 2019, the job-finding rate averages $f = 55.8\%$ per month, the job-separation rate averages $\l = 3.2\%$ per month, so the rate of decay averages $\l+f = 59.0 \%$ per month.
-Accordingly, the half-life of the deviation from the Beveridgean unemployment rate, $u(t)-u^b$, is $\ln(2)/0.590 = 1.17$ month.
-Since about $50\%$ of the deviation evaporates within one month, about $90\%$ of it evaporates within one quarter, which implies that the unemployment rate is always close to the Beveridgean unemployment rate.
-
-\subsection{Shouldn't workers out of the labor force be part of the calculation?}
-
-If the labor-force participation rate systematically responded to the unemployment rate, then it would make sense to endogenize labor-force participation and include it in the welfare analysis.
-And in theory, two kinds of effects are possible.
-It would be possible that additional workers enter the labor force in bad times to supplement their household's income.
-It would also be possible that jobseekers become discouraged in bad times and leave the labor force.
-In practice, however, the labor-force participation rate appears acyclical, so variations in unemployment do not lead to systematic changes in participation \citep[p.
-294]{S09}.
-Shimer's result is based on US data for 1960--2006.
-It is in line with earlier evidence from the United States.
-Using data for 1946--1954: \citet[p.
-32]{R57} found that ``There is no substantial evidence upholding the view that labor-force participation declines when there are moderate decreases in the demand for labor.''
-
-\subsection{Shouldn't the hardship from unemployment be included in the calculation?}
-
-Beyond production lost, unemployment can also create psychological hardship for unemployed workers.
-This psychological toll has been understood for a long time, even though it is often neglected in economics.
-\citet[p.
-11]{R49} for instance noted that ``The most striking aspect of unemployment is the suffering of the unemployed and their families---the loss of health and morale that follows loss of income and occupation.'' In contrast, unemployed workers might enjoy more leisure than employed workers.
-\citet{MS16} account for such considerations and find that psychological costs from unemployment offset the value of home production and leisure by unemployed workers.
-In net, it is as if unemployed workers do not engage in home production---which is what we assume here.
-
-\subsection{Are recruiting and jobseeking useless activities?}
-
-Of course recruiting and jobseeking are necessary for workers and firms to match with each other.
-But they do not generate any social welfare by themselves.
-In academia, faculty members have to spend time recruiting new colleagues.
-But during that time they do not teach or do research.
-Similarly, when graduate students are on the job market looking for a faculty position, they cannot teach or do research.
-In this paper we are looking for the unemployment rate that minimizes disruptions caused by recruiting and jobseeking.
-
-\subsection{Shouldn't we eliminate unemployment to almost zero?}
-
-Unemployment is clearly a waste of economic resources as people who would like to work and produce are forced to look for jobs and are not able to be productive.
-Yet, reducing the unemployment rate to zero is not desirable because, by the logic of the Beveridge curve, it would require a huge number of vacancies.
-Servicing vacancies takes work--- posting advertisements, screening and interviewing candidates---work that has to be diverted away from other productive tasks.
-And machines cannot do recruiting for firms.
-Therefore, vacancies also consume economic resources.
-It turns out that job recruiting is as labor intensive as job seeking is: one vacancy consumes as much resources as what is wasted when one person is looking for a job instead of working.
-As a result, it is efficient to reduce the sum of the unemployment and vacancy rates, and not simply the unemployment rate.
-
-\subsection{Don't people move from job to job without going through unemployment?}
-
-It is true that many workers change jobs without going through unemployment.
-This does reduce the waste on the jobseeking side of the labor market.
-Conceptually, such moves can be considered as unemployment of duration zero.
-While it would be ideal if all unemployment spells were of zero duration, empirically they are not: this is why the Beveridge curve exists.
-Our analysis follows from the Beveridge curve, and it applies even with job-to-job moves.
-
-\subsection{What do the results imply for monetary policy?}
-
-\paragraph{In practice} In the United States, the government is mandated to stabilize the economy at full employment.
-We have argued that our efficient unemployment rate $u^* = \sqrt{uv}$ is the appropriate measure of full employment.
-Hence, the Fed could use it as their full-employment target---especially since $\sqrt{uv}$ can be measured in real time.
-
-\paragraph{In theory} Moreover, targeting $u^*$ is optimal in certain monetary models with unemployment.
-This happens for instance in the model developed by \citet{MS19}.
-In that model, it is optimal to eliminate the unemployment gap by setting the nominal interest rate appropriately.
-As the unemployment gap is countercyclical, and a reduction in interest rate lowers unemployment, it is optimal to lower the interest rate in slumps, when the unemployment gap is high, and to raise it in booms, when the unemployment gap turns negative.
-
-\paragraph{Sufficient-statistic formula} In fact, \citet[section 5]{MS19} show that starting from a nominal interest rate $i>0$ and an inefficient unemployment rate $u \neq u^*$, the Fed should set the federal funds rate to $i^*$ such that
-\begin{equation}
- i - i^* =  \frac{u-u^*}{du/di}.
-\label{e:monetary}\end{equation}
-The statistic $i-i^*$ indicates the decrease in interest rate required to reach the optimal policy.
-The statistic $u-u^*$ is the prevailing unemployment gap.
-And the statistic $du/di>0$ is the monetary multiplier: the percentage-point decrease in unemployment achieved by lowering the nominal interest rate by 1 percentage point.
-This sufficient-statistic formula ensures that the unemployment rate is efficient, and therefore also that the economy is at full employment.
-
-\paragraph{Monetary multiplier} A large literature estimates the effect of the federal funds rate on unemployment \citep{CEE99,C12,R15}.
-A midrange estimate of the US monetary multiplier is $du/di = 0.5$ \citep[p.
-402]{MS19}.
-Combined with a monetary multiplier of $0.5$, formula~\eqref{e:monetary} says that the Fed should cut its interest rate by $2$ percentage points for each positive percentage point of unemployment gap, and raise its interest rate by $2$ percentage points for each negative percentage point of unemployment gap.
-For instance in March 2022, the economy is exceedingly tight: the unemployment gap is $-1.6$ percentage points.
-The Fed should raise its interest rate by $1.6\times 2 = 3.2$ percentage points in total to cool the economy back to full employment.
-
-\paragraph{At the zero lower bound} If the zero lower bound on nominal interest rates becomes binding, conventional monetary policy will not be able to sustain full employment.
-But other policies, such as government spending, can bring the economy closer to full employment.
-In a fiscal model with unemployment, \citet{MS15} find that optimal government spending deviates from the \citet{S54} rule to reduce---but not eliminate---the unemployment gap.
-Since the unemployment gap is countercyclical, and an increase in government spending reduces unemployment \citep{Ra13}, optimal government spending is countercyclical.
-
-\subsection{Shouldn't redistributive considerations be included in the analysis?}
-
-Our simple efficiency criterion minimizes the nonproductive use of labor.
-However, it is reasonable to think that waste on the jobseekers' side from looking for a job is more harmful than waste on the employer side for spending resources to fill vacancies.
-In other words, for distributive reasons, we might prefer having more waste if this waste is pushed toward employers.
-It is indeed the case that labor likes tight labor markets.
-Under some conditions, it is theoretically preferable to have direct redistribution (for example through more generous unemployment insurance funded by taxes on the wealthy) rather than through inefficiently tight labor markets.
-Practically, if taxes and transfers are set and cannot be made more generous, then an excessively tight labor market is an indirect way to redistribute toward labor.
-
-\subsection{Is the hyperbolic Beveridge curve consistent with a matching function?}
-
-\citet[equation (A12)]{MS16} relate the matching elasticity $\h$ to the Beveridge elasticity $\e$ in a matching model:
-\begin{equation*}
-\h = \frac{1}{1+\e}\bp{\e-\frac{u}{1-u}}.
-\end{equation*}
-An hyperbolic Beveridge curve implies a Beveridge elasticity $\e=1$.
-With an unemployment rate around $u=5\%$, we obtain a matching elasticity of $\h = 0.47$.
-Such matching elasticity falls squarely in the range of estimates obtained with aggregate US data, which span $0.30$--$0.76$ \citet[p.
-9]{MS16}.
-So the hyperbolic Beveridge curve is consistent with US estimates of the matching function.
-
-\subsection{What causes the observed fluctuations along the Beveridge curve?}
-
-Over the business cycle, unemployment and vacancy rates move along the Beveridge curve (such as in figure~\ref{f:pandemic}).
-Different models offer different explanations for these fluctuations?
-In the DMP model, shocks to workers' bargaining power lead to fluctuations along the Beveridge curve \citep[table~6]{S05}.
-In the variants of the DMP model by \citet{H05}, \citet{HM08}, and \citet{M09}, real wages are rigid, and shocks to labor productivity generate realistic fluctuations along the Beveridge curve.
-In the mismatch model by \citet{S07} and stock-flow matching model by \citet{ES10}, shocks to labor productivity also generate sizable fluctuations along the Beveridge curve.
-Finally, in the business-cycle model by \citet{MS19}, prices are rigid and shocks to aggregate demand generate fluctuations along the Beveridge curve.
-Such shocks may be caused by changes in monetary policy, or by changes in people's willingness to save (modeled as changes in people's marginal utility of wealth).
-
-\subsection{Should the government really take the Beveridge curve as given?}
-
-\paragraph{Empirical argument} Empirically the Beveridge curve is very stable over the business cycle.
-It does not seem to respond to stabilization policies (either monetary or fiscal policy) in any systematic way.
-Given this stability, it makes sense that the Fed takes the Beveridge curve as given.
-
-\paragraph{Theoretical argument} Most Beveridgean models of monetary and fiscal policy are built around a matching function \citep{W05,BG08,RW11,M12,MS15,MS19}.
-In these models the Beveridge curve is determined by the matching function and job-separation rate.
-Neither respond to monetary or fiscal policy, so the Beveridge curve is unaffected by policy.
-In the context of business-cycle stabilization, it therefore seems appropriate for the government to take the Beveridge curve as given.\footnote{Stabilization policies do not seem to affect the Beveridge curve, but some structural policies might.
-For instance, Germany implemented labor-market reforms in 2003--2005 that were designed to improve how the labor market operates (Hartz reforms).
-Such policies might shift the Beveridge curve inward.}
-
-\subsection{What explains the inefficiency of the labor market?}
-
-In most Beveridgean models, unlike in Walrasian models, the market economy does not overlap with the planning economy.
-This is because most price and wage mechanisms do not ensure efficiency.
-For instance, in matching models of the labor market, the wage is determined in a situation of bilateral monopoly, so a range of wages is theoretically possible.
-A wage mechanism picks one wage within the range.
-There is only an infinitesimal chance that this wage is the one ensuring efficiency \citep[p.~185]{P00}.
-Accordingly, theory does not guarantee that unemployment is efficient---or that full employment occurs.
-
-\subsection{What does labor-market tightness imply for inflation?}
-
-In the recovery of the coronavirus pandemic in the United States, the labor market has reached a tightness never seen since World War 2 (figure~\ref{f:tightness}).
-At the same time, wage inflation has also been very high.
-The Employment Cost Index for wages and salaries of workers in the private sector grew at an annual rate of $6.3\%$ in 2021Q3 and $5.0\%$ in 2022Q1, the highest rates recorded since the Employment Cost Index was created in 2001 \citep{ECIWAG}.
-It is therefore quite possible that wage inflation accelerates when the labor market is exceedingly tight.
-\citet{DS22} provide some support for this hypothesis.
-They find that both unemployment and vacancies are good predictors of wage inflation.
-Since labor-market tightness is the vacancy-to-unemployment ratio, it could be a good predictor of wage inflation as well.
+At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint.
 
@@
 
+\newpage
+\appendix
+
+\section{A section}\label{a:appendix1}
+
+We extend the p-hacking model of section~\ref{s:section} by introducing a cost of research, incurred at each new p-hacking step.
+Obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
+
+\subsection{Assumptions}
+
+At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
+A corrolary in the appendix is as follows:
+
+\begin{corollary} The number of p-hacking steps might be
+\begin{equation*}
+\mathbb{E}(\Omega) = \mathbb{P}(\omega\cdot \mu - \xi) - \sum_{i=0}^{m}\sum_{j=-\infty}^{n} \sigma(i,j) + 123^{56}.
+\end{equation*}\end{corollary}
+
+\subsection{Optimal stopping time and robust critical value}
+
+\paragraph{Significant result} Since it is optimal to engage in research, the researcher starts a first p-hacking step.
+With probability $\xi$, the step can be completed, and the researcher obtains a test result \citep{MS22a}.
+
+\paragraph{Another paragraph with some math} When a researcher decides to start research, three scenarios are again possible.
+With probability $1-\gamma$, the researcher cannot complete the first research step and cannot submit any result; she then collects $0$.
+With probability $\gamma$, she can complete the first research step.
+Then with probability $S(z^*)$, her result is significant and she collects $v^s$.
+With probability $1-S(z^*)$, her result is insignificant and the continuation value at this point is $\mathcal{V}^i$.
+Aggregating these scenarios, we obtain the initial continuation value:
+\begin{equation}
+\mathbb{V}^r = (1-\gamma) \times 0 +\gamma S(z^*) v^s+\gamma [1-S(z^*)] \mathcal{V}^i-c.
+\label{e:appendix1}\end{equation}
+
+\paragraph{Paragraph with links to appendix equations} What does the researcher decide if the result is insignificant?
+It depends on the research cost $\mathcal{C}$.
+Equation \eqref{e:appendix1} shows that if the cost is high enough, the researcher stops right away.
+This happens when the possibility of obtaining a significant result in the future does not compensate the research cost.
+
+\subsection{A larger figure, without panel, in the appendix}
+
+At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
+This is showed in figure \ref{f:appendix1}.
+
+\begin{figure}[t]
+\includegraphics[scale=\mfig,page=1]{\pdf}
+\caption{A caption for the larger graph}
+\note{A note for the larger graph. Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus.}
+\label{f:appendix1}\end{figure}
+
+\section{Another section}\label{a:appendix2}
+
+At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti.
+
+\subsection{A even larger figure, without panel, in the appendix}
+
+At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
+
+\begin{figure}[t]
+\includegraphics[scale=\lfig,page=3]{\pdf}
+\caption{A caption for the even larger graph}
+\note{A note for the larger graph. Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus.}
+\label{f:appendix2}\end{figure}
+
+When a researcher decides to continue p-hacking, three scenarios are possible, as showed in figure \ref{f:appendix2}.
+The vector probability $1-\bm{\gamma}$ then gives $v^i$.
+The vector $\bm{\gamma}$ contains the p-hacking steps.
+Then with probability $S(z^*)$, her result is significant and she collects $v^s$.
+With probability $1-S(z^*)$, her result is insignificant once again and the continuation value at this point is $\mathcal{V}^i$.
+In any case, she must incur a cost $c$ to conduct the research step.
+
+We rewrite the initial continuation value as
+\begin{equation}
+\mathbb{V}^r = \zeta \mathcal{V}^i + \zeta S(z^*) (v^s - \mathcal{V}^i) - c.
+\label{e:appendix2}\end{equation}
+
+Equation \eqref{e:appendix2} shows when there is no p-hacking. The robust critical value is then just the classical critical value.
+
+\subsection{Condition for p-hacking}\label{a:subappendix}
+
+When a researcher has obtained one insignificant result, it is optimal to continue p-hacking if $\mathcal{V}^i > \mathcal{V}^r$.
+This is related to the results by \citet{MS21b}.\footnote{The reference goes to the reference list at the end of the main text.}
+
+\paragraph{Continuation value of research} We first compute the continuation value of research for a researcher who has already recorded an insignificant result.
+We denote this value $\mathcal{B}^\theta$.
+Because the researcher's situation is invariant in time, the continuation value is the same at each p-hacking step.
+
+\paragraph{Condition for research} We also compute the cost below which it is optimal to engage in research.
+Given that we have normalized the outside option of the researcher to $0$, it is optimal to engage in research if the expected value from it is positive.
+This results are summarized in a repository with the following URL: \url{https://github.com/pmichaillat/latex-paper}.
+
 \end{document}

### 2023-01-29 paper.tex
Update paper.tex
Update paper template

@@
 \bibliographystyle{bibliography}
-\hypersetup{pdftitle={Paper Title}}
+\hypersetup{pdftitle={Paper Example}}
 % Enter permanent URL to paper
@@
 \paragraph{Positioning in the literature} Lorem ipsum dolor sit amet \citep{MS15}, consectetur adipiscing elit \citep[p.
-1305]{MS19}, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua .
+1305]{MS19}, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
 \citet{M12,M14} enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
@@
 
-\begin{proposition}\label{p:type1}  When the critical value is set to $z \in \mathbb{R}$, the probability of finding a type 1 error in a reported study is
+\begin{proposition}\label{p:type1}  Lorem ipsum dolor sit amet $z \in \mathbb{R}$, consectetur adipiscing elit
 \begin{equation}
@@
 \label{e:type1}\end{equation}
-The probability of type 1 error is larger when researchers p-hack ($S^*(z) > \bm{S(z)}$).
-In fact, the probability of type 1 error grows linearly with the expected number of p-hacking steps, which is givne by $\mathbb{E}(N(z))$:
+Ut enim ad minim veniam, sunt in culpa qui officia deserunt mollit anim id est laborum:
 \begin{equation}
@@
 
-\begin{proof} There is a simple proof to this result. It is very simple. It is so simple. It gives equation \eqref{e:type1steps}.\end{proof}
+\begin{proof} There is a simple proof to this result. It gives equation \eqref{e:type1steps}.\end{proof}
 
- Quis autem vel eum iure reprehenderit, qui in ea voluptate velit esse, quam nihil molestiae consequatur $\bm{\alpha(X)}$ when researchers p-hack.
-Since the probability of type 1 error with p-hacking is given by \eqref{e:type1}, the robust critical value $z^*$ is implicitly defined by
+ Quis autem vel eum iure reprehenderit, qui in ea voluptate velit esse, quam nihil molestiae consequatur $\bm{\alpha(X)}$.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum \eqref{e:type1}, excepteur sint occaecat cupidatat non proident $z^*$ is implicitly defined by
 \begin{equation*}
@@
 
-\begin{lemma}\label{p:cv} For any hypothesis test with significance level $\alpha$, the robust critical value is
+\begin{lemma}\label{p:cv} Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat $\alpha$:
 \begin{equation}
@@
 \label{e:cv}\end{equation}
-The robust critical value is always larger than the classical critical value $\mathcal{Z}(\alpha)$. \end{lemma}
+Consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua $\mathcal{Z}(\alpha)$. \end{lemma}
 
-\paragraph{P-hacking under the robust critical value} Ut enim ad minima veniam, quis nostrum exercitationem ullam corporis suscipit laboriosam, nisi ut aliquid ex ea commodi consequatur?
+Ut enim ad minima veniam, quis nostrum exercitationem ullam corporis suscipit laboriosam, nisi ut aliquid ex ea commodi consequatur?
 Combining equation \eqref{e:cv} with other results, we obtain the following corollary (the proof is in appendix~\ref{a:subappendix}):
 
-\begin{corollary} The number of p-hacking steps under the robust critical value satisfies
+\begin{corollary} Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua:
 \begin{equation*}
@@
 Itaque earum rerum hic tenetur a sapiente delectus, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat (figure~\ref{f:graph2}).
-Over 1930Q1--2022Q1, labor-market tightness averages $0.68$.
-In 2022, tightness reached a value of 2.0, which it had last reached in 1945.The efficient unemployment rate is stable over time.
-It hovered around 4\%--5\% between 1930 and 1970.
 
@@
 
-We extend the p-hacking model of section~\ref{s:section} by introducing a cost of research, incurred at each new p-hacking step.
+We extend the model of section~\ref{s:section} by introducing a quadratic cost.
 Obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
@@
 At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
+
+\subsection{New results}
+
+ Et harum quidem rerum facilis est et expedita distinctio.
+Nam libero tempore, cum soluta nobis est eligendi optio cumque nihil impedit quo minus id quod maxime placeat facere possimus, omnis voluptas assumenda est, omnis dolor repellendus \citep{MS22a}.
 A corrolary in the appendix is as follows:
 
-\begin{corollary} The number of p-hacking steps might be
+\begin{corollary} Similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga:
 \begin{equation*}
@@
 
-\subsection{Optimal stopping time and robust critical value}
-
-\paragraph{Significant result} Since it is optimal to engage in research, the researcher starts a first p-hacking step.
-With probability $\xi$, the step can be completed, and the researcher obtains a test result \citep{MS22a}.
-
-\paragraph{Another paragraph with some math} When a researcher decides to start research, three scenarios are again possible.
-With probability $1-\gamma$, the researcher cannot complete the first research step and cannot submit any result; she then collects $0$.
-With probability $\gamma$, she can complete the first research step.
-Then with probability $S(z^*)$, her result is significant and she collects $v^s$.
-With probability $1-S(z^*)$, her result is insignificant and the continuation value at this point is $\mathcal{V}^i$.
-Aggregating these scenarios, we obtain the initial continuation value:
+\paragraph{Another paragraph with some math} Temporibus autem quibusdam $\xi$ et aut officiis debitis aut rerum necessitatibus saepe eveniet ut et voluptates repudiandae sint et molestiae non recusandae $1-\gamma$.
+Itaque earum rerum hic $S(z^*)$ tenetur a sapiente delectus $\mathcal{B}^\theta$, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat $\mathcal{V}^i$.
+Aggregating these scenarios, we obtain the continuation value:
 \begin{equation}
@@
 
-\paragraph{Paragraph with links to appendix equations} What does the researcher decide if the result is insignificant?
-It depends on the research cost $\mathcal{C}$.
-Equation \eqref{e:appendix1} shows that if the cost is high enough, the researcher stops right away.
-This happens when the possibility of obtaining a significant result in the future does not compensate the research cost.
+\paragraph{Paragraph with links to appendix equations} Ut enim ad minima veniam, quis nostrum exercitationem ullam corporis suscipit laboriosam, nisi ut aliquid ex ea commodi consequatur $\mathcal{C}$?
+Quis autem vel eum iure reprehenderit qui in ea voluptate velit esse quam nihil molestiae consequatur, vel illum qui dolorem eum fugiat quo voluptas nulla pariatur?
+Equation \eqref{e:appendix1} shows that autem vel eum iure reprehenderit qui in ea voluptate velit esse quam nihil molestiae consequatur.
 
@@
 
-At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
+At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga, as showed in figure \ref{f:appendix2}.
 
@@
 
-When a researcher decides to continue p-hacking, three scenarios are possible, as showed in figure \ref{f:appendix2}.
-The vector probability $1-\bm{\gamma}$ then gives $v^i$.
-The vector $\bm{\gamma}$ contains the p-hacking steps.
-Then with probability $S(z^*)$, her result is significant and she collects $v^s$.
-With probability $1-S(z^*)$, her result is insignificant once again and the continuation value at this point is $\mathcal{V}^i$.
-In any case, she must incur a cost $c$ to conduct the research step.
+\subsection{Some final discussion with footnote and references}\label{a:subappendix}
 
-We rewrite the initial continuation value as
-\begin{equation}
-\mathbb{V}^r = \zeta \mathcal{V}^i + \zeta S(z^*) (v^s - \mathcal{V}^i) - c.
-\label{e:appendix2}\end{equation}
-
-Equation \eqref{e:appendix2} shows when there is no p-hacking. The robust critical value is then just the classical critical value.
-
-\subsection{Condition for p-hacking}\label{a:subappendix}
-
-When a researcher has obtained one insignificant result, it is optimal to continue p-hacking if $\mathcal{V}^i > \mathcal{V}^r$.
-This is related to the results by \citet{MS21b}.\footnote{The reference goes to the reference list at the end of the main text.}
-
-\paragraph{Continuation value of research} We first compute the continuation value of research for a researcher who has already recorded an insignificant result.
-We denote this value $\mathcal{B}^\theta$.
-Because the researcher's situation is invariant in time, the continuation value is the same at each p-hacking step.
-
-\paragraph{Condition for research} We also compute the cost below which it is optimal to engage in research.
-Given that we have normalized the outside option of the researcher to $0$, it is optimal to engage in research if the expected value from it is positive.
+Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt if $\mathcal{V}^i > \mathcal{V}^r$.
+This is related to the results by \citet{MS21b}.\footnote{The reference goes to the reference list at the end of the main text.} Sed ut perspiciatis unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam, eaque ipsa quae ab illo inventore veritatis et quasi architecto beatae vitae dicta sunt explicabo.
 This results are summarized in a repository with the following URL: \url{https://github.com/pmichaillat/latex-paper}.

### 2023-01-29 appendix.tex
Update paper and appendix templates

@@
 
-When a researcher decides to continue p-hacking, three scenarios are possible, as showed in figure \ref{f:appendix2}.
-The vector probability $1-\bm{\gamma}$ then gives $v^i$.
-The vector $\bm{\gamma}$ contains the p-hacking steps.
-Then with probability $S(z^*)$, her result is significant and she collects $v^s$.
-With probability $1-S(z^*)$, her result is insignificant once again and the continuation value at this point is $\mathcal{V}^i$.
-In any case, she must incur a cost $c$ to conduct the research step.
+\subsection{Some final discussion with footnote and references}\label{a:subappendix}
 
-We rewrite the initial continuation value as
-\begin{equation}
-\mathbb{V}^r = \zeta \mathcal{V}^i + \zeta S(z^*) (v^s - \mathcal{V}^i) - c.
-\label{e:appendix2}\end{equation}
-
-Equation \eqref{e:appendix2} shows when there is no p-hacking. The robust critical value is then just the classical critical value.
-
-\subsection{Condition for p-hacking}\label{a:subappendix}
-
-When a researcher has obtained one insignificant result, it is optimal to continue p-hacking if $\mathcal{V}^i > \mathcal{V}^r$.
-This is related to the results by \citet{MS21b}.\footnote{The reference goes to its own reference list at the end of the appendix---unlike when the appendix was at the end of the main text.}
-
-\subsection{Continuation value of research}
-
-We first compute the continuation value of research for a researcher who has already recorded an insignificant result.
-We denote this value $\mathcal{B}^\theta$.
-Because the researcher's situation is invariant in time, the continuation value is the same at each p-hacking step.
-
-\subsection{Condition for research}
-
-We also compute the cost below which it is optimal to engage in research.
-Given that we have normalized the outside option of the researcher to $0$, it is optimal to engage in research if the expected value from it is positive.
+Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt if $\mathcal{V}^i > \mathcal{V}^r$.
+This is related to the results by \citet{MS21b}.\footnote{The reference goes to its own reference list at the end of the appendix---unlike when the appendix was at the end of the main text.} Sed ut perspiciatis unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam, eaque ipsa quae ab illo inventore veritatis et quasi architecto beatae vitae dicta sunt explicabo.
 This results are summarized in a repository with the following URL: \url{https://github.com/pmichaillat/latex-paper}.

### 2023-01-29 paper.tex
Update paper and appendix templates

@@
 
-Oobcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga. Here is a list:
+\paragraph{Itemized lists} Oobcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga. Here is a list:
 \begin{itemize}
@@
 
-Omnis dolor repellendus. And here is a numbered list:
+\paragraph{Numbered lists} Omnis dolor repellendus. And here is a numbered list:
 \begin{enumerate}
@@
 \begin{equation}
-S^*(z) = \frac{\bm{S(z)}}{ 1 -\zeta [1- \bm{S(z)}]}.
+\bm{S}^*(z) = \frac{S(z)}{23 -\zeta\gamma [45- S(z)] + \ln(y)}.
 \label{e:type1}\end{equation}
@@
 \begin{equation}
-S^*(z) = S(z) \times \mathbb{E}(N(z)).
+\mathcal{S}^*(z) = S(z) \times \mathbb{E}(N(z)) + \alpha \beta - \sum_{i=0}^{+\infty}\mathcal{F}^{i}-\exp(\lambda \mu).
 \label{e:type1steps}\end{equation}
@@
 \begin{equation*}
-\frac{\mathcal{S}(z^*)}{1 -\theta + \theta \mathcal{S}(z^*)} = \sum_{\alpha=-\infty}^{+\infty}F^{\alpha} - \Lambda(\alpha).
+\frac{\mathcal{S}(z^*)}{1 -\theta + \theta \mathcal{S}(z^*)} = \sum_{i=0}^{+\infty}F^{i} - \Lambda(i).
 \end{equation*}
@@
 At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
-Et harum quidem rerum facilis est et expedita distinctio.
+
+\paragraph{A paragraph} Et harum quidem rerum facilis est et expedita distinctio.
 Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus, omnis voluptas assumenda est, omnis dolor repellendus.
 Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet, ut et voluptates repudiandae sint et molestiae non recusandae.
-Itaque earum rerum hic tenetur a sapiente delectus, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat.
+
+\paragraph{Another paragraph} Itaque earum rerum hic tenetur a sapiente delectus, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat.
+Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus, omnis voluptas assumenda est, omnis dolor repellendus.
+Figure~\ref{f:graph1} depicts the result.
 
@@
 \label{f:graph1}\end{figure}
-
-\paragraph{Description of a result} Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus, omnis voluptas assumenda est, omnis dolor repellendus.
-Figure~\ref{f:graph1} depicts the result.
 

### 2023-03-10 paper.tex
Update readme

@@
 \bibliographystyle{bibliography}
+
+% Enter paper title:
 \hypersetup{pdftitle={Paper Example}}
+
 % Enter permanent URL to paper
 \available{https://github.com/pmichaillat/latex-paper}
-% Enter BibTeX file with references here:
+
+% Enter BibTeX file with references:
 \newcommand{\bib}{bibliography.bib}
+
 % Enter PDF file with figures here:
@@
 
+% Fill out paper:
 \begin{document}
-
 \title{Paper Title}
@@
 
-\section{Section}\label{s:section}
+\section{Section title}\label{s:section}
 
@@
 
-\subsection{A subsection with lists}\label{s:lists}
+\subsection{Subsection title}
+
+Nemo enim ipsam voluptatem, quia voluptas sit, aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos, qui ratione voluptatem sequi nesciunt, neque porro quisquam est, qui dolorem ipsum, quia dolor sit amet consectetur adipiscing velit, sed quia non numquam eius modi tempora incidunt, ut labore et dolore magnam aliquam quaerat voluptatem.
+Sed ut perspiciatis, unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam eaque ipsa, quae ab illo inventore veritatis et quasi architecto beatae vitae dicta sunt, explicabo.
+
+\subsection{Subsection with lists}\label{s:lists}
 
@@
 Sed ut perspiciatis, as showed in section \ref{s:lists}, unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam eaque ipsa, quae ab illo inventore veritatis et quasi architecto beatae vitae dicta sunt, explicabo.
-Nemo enim ipsam voluptatem, quia voluptas sit, aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos, qui ratione voluptatem sequi nesciunt, neque porro quisquam est, qui dolorem ipsum, quia dolor sit amet consectetur adipiscing velit, sed quia non numquam eius modi tempora incidunt, ut labore et dolore magnam aliquam quaerat voluptatem.
-We will explore more in section \ref{s:graphs}.
+Nemo enim ipsam voluptatem, quia voluptas sit, aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos, qui ratione voluptatem sequi nesciunt, neque porro quisquam est, qui dolorem ipsum, quia dolor sit amet consectetur adipiscing velit, sed quia non numquam eius modi tempora incidunt, ut labore et dolore magnam aliquam quaerat voluptatem in section \ref{s:graphs}.
 
@@
 Quis autem vel eum iure reprehenderit, qui in ea voluptate velit esse, quam nihil molestiae consequatur $\bm{\alpha(X)}$.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum \eqref{e:type1}, excepteur sint occaecat cupidatat non proident $z^*$ is implicitly defined by
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum \eqref{e:type1}, excepteur sint occaecat cupidatat non proident $z^*$:
 \begin{equation*}
@@
 \end{equation*}
-From this implicit definition we obtain the following results (details of the proof are relegated to appendix~\ref{a:appendix1}):
+From this definition we obtain the following results (details of the proof are relegated to appendix~\ref{a:appendix1}):
 
@@
 
-\section{A section with graphs}\label{s:graphs}
+\section{Section with graphs}\label{s:graphs}
 
@@
 
-\subsection{A subsection with graphs}
+\subsection{Subsection with graphs}
 
@@
 
-\paragraph{A paragraph} Et harum quidem rerum facilis est et expedita distinctio.
+\paragraph{Paragraph title} Et harum quidem rerum facilis est et expedita distinctio.
 Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus, omnis voluptas assumenda est, omnis dolor repellendus.
@@
 \begin{figure}[t]
-\subcaptionbox{First panel\label{f:panel1}}{\includegraphics[scale=\sfig,page=1]{\pdf}}\hfill
-\subcaptionbox{Second panel\label{f:panel2}}{\includegraphics[scale=\sfig,page=2]{\pdf}}
-\caption{A caption for the graph}
-\note{A note for the graph. Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus.}
+\subcaptionbox{Caption for first panel\label{f:panel1}}{\includegraphics[scale=\sfig,page=1]{\pdf}}\hfill
+\subcaptionbox{Caption for second panel\label{f:panel2}}{\includegraphics[scale=\sfig,page=2]{\pdf}}
+\caption{Caption for the graph}
+\note{Note for the graph. Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus.}
 \label{f:graph1}\end{figure}
 
-\paragraph{Another result in a panel} In fact, nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus, omnis voluptas assumenda est, omnis dolor repellendus.
-Figure \ref{f:panel1} displays again this fact, but now in log scale.
-This empirical regularity is particularly striking in figure \ref{f:panel2} because the two series are constructed independently of each other.
+\paragraph{Another result in a panel} Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus, omnis voluptas assumenda est, omnis dolor repellendus.
+Figure \ref{f:panel1} displays cum soluta nobis est eligendi optio, cumque nihil impedit.
+Quod maxime placeat, facere possimus, omnis voluptas assumenda est in figure \ref{f:panel2}.
 
@@
 \subcaptionbox{Caption for panel}{\includegraphics[scale=\sfig,page=1]{\pdf}}
-\caption{A caption for the graph}
-\note{A note for the graph. Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat. The figures are borrowed from \citet[section 3]{MS22b}.}
+\caption{Caption for the graph}
+\note{Note for the graph. Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat from \citet[section 3]{MS22b}.}
 \label{f:graph2}\end{figure}
 
-\paragraph{Informal description of the results} At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
+\paragraph{Another paragraph} At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
 Et harum quidem rerum facilis est et expedita distinctio.
@@
 
-\paragraph{Formal description of the results} At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
+\paragraph{Yet another paragraph} At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
 Et harum quidem rerum facilis est et expedita distinctio.
@@
 
-\subsection{The table}
+\subsection{Table}
 
@@
 \begin{table}[t]
-\caption{A caption for the table}
+\caption{Caption for the table}
 \begin{tabular*}{\textwidth}[]{p{3.3cm}@{\extracolsep\fill}cccc}
@@
 \midrule
-\multicolumn{5}{l}{A. ZLB episode}\\
+\multicolumn{5}{l}{A. Qui dolorem eum fugiat}\\
 ZLB: & $t\in (0,T) $ & $r^n<0$ & $i=0$  & -- \\
@@
 \midrule
-\multicolumn{5}{l}{B. ZLB episode with forward guidance}\\
+\multicolumn{5}{l}{B. Nam libero tempore, cum soluta nobis}\\
 ZLB: & $t\in (0,T)$ & $r^n<0$ & $i=0$  & -- \\
@@
 \midrule
-\multicolumn{5}{l}{C. ZLB episode with government spending}\\
+\multicolumn{5}{l}{C. At vero eos et accusamus}\\
 ZLB: & $t\in (0,T) $ & $r^n<0$ & $i=0$  & $g>0$  \\
@@
 \end{tabular*}
-\note{This table describes three scenarios: Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet, ut et voluptates repudiandae sint et molestiae non recusandae.
-Ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat.
-The table is borrowed from \citet[table 1]{MS21a}.}
+\note{Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet, ut et voluptates repudiandae sint et molestiae non recusandae.
+Ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat from \citet[table 1]{MS21a}.}
 \label{t:table}\end{table}
@@
 
-To summarize, we combine the US data for 1930Q1--2022Q1.
 Qui dolorem eum fugiat, quo voluptas nulla pariatur?
@@
 
-To conclude, we address questions that are frequently asked about the analysis.
-We cover questions about the analysis, about the policy implications of the results, about the theoretical underpinnings of the results, and about possible extensions.
-
-At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
+To conclude, at vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
 
@@
 
+% Fill out appendix:
 \newpage
 \appendix
-
-\section{A section}\label{a:appendix1}
+\section{Appendix title}\label{a:appendix1}
 
@@
 
-\subsection{Assumptions}
+\subsection{Subsection}
 
@@
 
-\subsection{New results}
+\subsection{Another subsection}
 
@@
 
-\paragraph{Another paragraph with some math} Temporibus autem quibusdam $\xi$ et aut officiis debitis aut rerum necessitatibus saepe eveniet ut et voluptates repudiandae sint et molestiae non recusandae $1-\gamma$.
+\paragraph{A paragraph with some math} Temporibus autem quibusdam $\xi$ et aut officiis debitis aut rerum necessitatibus saepe eveniet ut et voluptates repudiandae sint et molestiae non recusandae $1-\gamma$.
 Itaque earum rerum hic $S(z^*)$ tenetur a sapiente delectus $\mathcal{B}^\theta$, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat $\mathcal{V}^i$.
@@
 
-\subsection{A larger figure, without panel, in the appendix}
+\subsection{Larger figure, without panel, in the appendix}
 
-At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
-This is showed in figure \ref{f:appendix1}.
+At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga in figure \ref{f:appendix1}.
 
@@
 \includegraphics[scale=\mfig,page=1]{\pdf}
-\caption{A caption for the larger graph}
-\note{A note for the larger graph. Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus.}
+\caption{Caption for the larger graph}
+\note{Note for the larger graph. Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus.}
 \label{f:appendix1}\end{figure}
@@
 
-\subsection{A even larger figure, without panel, in the appendix}
+\subsection{Even larger figure, without panel, in the appendix}
 
@@
 \includegraphics[scale=\lfig,page=3]{\pdf}
-\caption{A caption for the even larger graph}
-\note{A note for the larger graph. Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus.}
+\caption{Caption for the even larger graph}
+\note{Note for the larger graph. Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus.}
 \label{f:appendix2}\end{figure}
@@
 
-Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt if $\mathcal{V}^i > \mathcal{V}^r$.
-This is related to the results by \citet{MS21b}.\footnote{The reference goes to the reference list at the end of the main text.} Sed ut perspiciatis unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam, eaque ipsa quae ab illo inventore veritatis et quasi architecto beatae vitae dicta sunt explicabo.
+Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt $\mathcal{V}^i > \mathcal{V}^r$ \citep{MS21b}.\footnote{The reference goes to the reference list at the end of the main text.} Sed ut perspiciatis unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam, eaque ipsa quae ab illo inventore veritatis et quasi architecto beatae vitae dicta sunt explicabo.
 This results are summarized in a repository with the following URL: \url{https://github.com/pmichaillat/latex-paper}.

### 2023-08-31 appendix.tex
Update math symbol font; improve formatting of title pages

@@
 \bibliographystyle{bibliography}
+% Enter paper title:
 \hypersetup{pdftitle={Paper Title: Online Appendices}}
-% Enter journal of publication and URL to paper
-\published{Journal Name}{https://github.com/pmichaillat/latex-paper}
-% Enter BibTeX file with references here:
+% Enter journal of publication and permanent URL to paper
+\available[Journal Name]{https://github.com/pmichaillat/latex-paper}
+% Enter BibTeX file with references:
 \newcommand{\bib}{bibliography.bib}
@@
 \newcommand{\pdf}{figures.pdf}
-% Enter TeX file with the main text so references can be used in appendix file (this is especially useful if the appendix has to be split away from the main text)
+% Enter TeX file with the main text so references can be used in appendix:
 \externaldocument{paper}
 
+% Fill out content of online appendix:
 \begin{document}
-
 \title{Paper Title: Online Appendices}
@@
 \date{Month Year}
-
 \begin{titlepage}
@@
 
-\section{A section}\label{a:appendix1}
+\section{Appendix title}\label{a:appendix1}
 
@@
 
-\subsection{Links}
+\subsection{Subsection title}
 
@@
 
-\subsection{Assumptions}
+\subsection{Another subsection}
 
@@
 
-\subsection{New results}
+\subsection{Yet another subsection}
 
@@
 \begin{equation*}
-\mathbb{E}(\Omega) = \mathbb{P}(\omega\cdot \mu - \xi) - \sum_{i=0}^{m}\sum_{j=-\infty}^{n} \sigma(i,j) + 123^{56}.
+\mathbb{E}(\Omega) = \mathbb{P}(\omega\cdot \mu - \xi) - \sum_{i=0}^{m}\sum_{j=i}^{n} \sigma(i,j) + 123^{56}.
 \end{equation*}\end{corollary}
 
-\paragraph{Another paragraph with some math} Temporibus autem quibusdam $\xi$ et aut officiis debitis aut rerum necessitatibus saepe eveniet ut et voluptates repudiandae sint et molestiae non recusandae $1-\gamma$.
-Itaque earum rerum hic $S(z^*)$ tenetur a sapiente delectus $\mathcal{B}^\theta$, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat $\mathcal{V}^i$.
+\paragraph{Paragraph with some math} Temporibus autem quibusdam $\xi$ et aut officiis debitis aut rerum necessitatibus saepe eveniet ut et voluptates repudiandae sint et molestiae non recusandae $1-\gamma$.
+Itaque earum rerum hic $S(z^*)$ tenetur a sapiente delectus $\mathcal{B}^\theta$, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat $\prod_i\mathcal{V}^i \mapsto \mathbb{Q} + \mathbb{\Lambda(\beta + \gamma)}$.
 Aggregating these scenarios, we obtain the continuation value:
 \begin{equation}
-\mathbb{V}^r = (1-\gamma) \times 0 +\gamma S(z^*) v^s+\gamma [1-S(z^*)] \mathcal{V}^i-c.
+\mathbb{V}^r = (1-\gamma) \times 0 +\gamma S(z^*) v^s+\gamma [1-S(z^*)] \times \frac{\sum_i\mathcal{V}^i}{2-c}.
 \label{e:appendix1}\end{equation}
@@
 
-\section{Another section}\label{a:appendix2}
+\section{Another appendix}\label{a:appendix2}
 
@@
 
-\subsection{Some final discussion with footnote and references}\label{a:subappendix}
+\subsection{Subsection with footnote and references}\label{a:subappendix}
 
-Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt if $\mathcal{V}^i > \mathcal{V}^r$.
-This is related to the results by \citet{MS21b}.\footnote{The reference goes to its own reference list at the end of the appendix---unlike when the appendix was at the end of the main text.} Sed ut perspiciatis unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam, eaque ipsa quae ab illo inventore veritatis et quasi architecto beatae vitae dicta sunt explicabo.
+Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt if $\mathcal{V}^i > \mathcal{V}^r$ \citep{MS21b}.\footnote{The reference goes to its own reference list at the end of the appendix---unlike when the appendix was at the end of the main text.} Sed ut perspiciatis unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam, eaque ipsa quae ab illo inventore veritatis et quasi architecto beatae vitae dicta sunt explicabo.
 This results are summarized in a repository with the following URL: \url{https://github.com/pmichaillat/latex-paper}.
@@
 \bibliography{\bib}
-
 \end{document}

### 2023-08-31 paper.tex
Update math symbol font; improve formatting of title pages

@@
 \bibliographystyle{bibliography}
-
 % Enter paper title:
 \hypersetup{pdftitle={Paper Example}}
-
 % Enter permanent URL to paper
 \available{https://github.com/pmichaillat/latex-paper}
-
 % Enter BibTeX file with references:
 \newcommand{\bib}{bibliography.bib}
-
 % Enter PDF file with figures here:
@@
 
-% Fill out paper:
+% Fill out content of paper:
 \begin{document}
@@
 Author 2: University 2.
-We thank many colleagues for helpful comments and discussions.
-This work was supported by the a generous grant [grant number]; and another generous foundation.}}
+We thank colleagues for helpful comments and discussions.
+This work was supported by a grant [grant number]; another grant [grant number]; and a foundation.}}
 \date{Month Year}                       
-
 \begin{titlepage}\maketitle
@@
 Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
 
@@
 Nemo enim ipsam voluptatem, quia voluptas sit, aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos, qui ratione voluptatem sequi nesciunt, neque porro quisquam est, qui dolorem ipsum, quia dolor sit amet consectetur adipiscing velit, sed quia non numquam eius modi tempora incidunt, ut labore et dolore magnam aliquam quaerat voluptatem.
+
 Sed ut perspiciatis, unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam eaque ipsa, quae ab illo inventore veritatis et quasi architecto beatae vitae dicta sunt, explicabo.
@@
 
-\begin{proposition}\label{p:type1}  Lorem ipsum dolor sit amet $z \in \mathbb{R}$, consectetur adipiscing elit
+\begin{proposition}\label{p:type1}  Lorem ipsum dolor sit amet $z \subset \mathbb{R}$, consectetur adipiscing elit
 \begin{equation}
-\bm{S}^*(z) = \frac{S(z)}{23 -\zeta\gamma [45- S(z)] + \ln(y)}.
+\bm{S}_{k_x}(z) \approx \frac{S(z)^x}{k / 23 -\zeta\gamma [45- S(z)] + \ln(y) - j^2+x(l)}.
 \label{e:type1}\end{equation}
@@
 \begin{equation}
-\mathcal{S}^*(z) = S(z) \times \mathbb{E}(N(z)) + \alpha \beta - \sum_{i=0}^{+\infty}\mathcal{F}^{i}-\exp(\lambda \mu).
+\sum_{j=0}^{+\infty}\mathcal{S}^j(z^{\ast}) = S(z) \times \mathbb{E}(N(z)) + \alpha \beta - \left[\sum_{i=0}^{+\infty}\mathcal{F}^{i}-\exp(\lambda \mu)\right].
 \label{e:type1steps}\end{equation}
@@
 
-Quis autem vel eum iure reprehenderit, qui in ea voluptate velit esse, quam nihil molestiae consequatur $\bm{\alpha(X)}$.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum \eqref{e:type1}, excepteur sint occaecat cupidatat non proident $z^*$:
+Quis autem vel eum iure reprehenderit, qui in ea voluptate velit esse, quam nihil molestiae consequatur $\prod_i\alpha_i(\bm{X}) \geq 2 \implies \mathcal{X} \to 5$ qui in ea voluptate velit esse $\sum_{x=1}^4 \xi^x(j) \quad \forall j\in \mathbb{R}$.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum \eqref{e:type1}, excepteur sint occaecat cupidatat non proident $z^\ast/4$:
 \begin{equation*}
-\frac{\mathcal{S}(z^*)}{1 -\theta + \theta \mathcal{S}(z^*)} = \sum_{i=0}^{+\infty}F^{i} - \Lambda(i).
+\frac{\mathcal{S}(z^*)}{1 -\theta + \theta \mathcal{S}(z^*)} = \prod_{i=0}^{+\infty}F^{i} - \Lambda(i) = \left\lbrace\prod_{i=0}^{+\infty}G^{i} - \Omega(i)\right\rbrace.
 \end{equation*}
@@
 
-\subsection{Table}
+\subsection{A subsection title}
 
@@
 \note{Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet, ut et voluptates repudiandae sint et molestiae non recusandae.
-Ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat from \citet[table 1]{MS21a}.}
+Ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat \citep[table 1]{MS21a}.}
 \label{t:table}\end{table}
 
-\subsection{Summary}
+\subsection{Another subsection title}
 
@@
 
-\subsection{Subsection}
+\subsection{Subsection title}
 
@@
 
-\subsection{Some final discussion with footnote and references}\label{a:subappendix}
+\subsection{Final subsection with footnote and references}\label{a:subappendix}
 

### 2023-09-01 paper.tex
Update PDF title
Explain better figure references

@@
 % Enter paper title:
-\hypersetup{pdftitle={Paper Example}}
+\hypersetup{pdftitle={Paper Title}}
 % Enter permanent URL to paper
@@
 
-\paragraph{Paragraph title} Et harum quidem rerum facilis est et expedita distinctio.
-Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus, omnis voluptas assumenda est, omnis dolor repellendus.
-Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet, ut et voluptates repudiandae sint et molestiae non recusandae.
-
-\paragraph{Another paragraph} Itaque earum rerum hic tenetur a sapiente delectus, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat.
-Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus, omnis voluptas assumenda est, omnis dolor repellendus.
-Figure~\ref{f:graph1} depicts the result.
+\paragraph{Paragraph with references to figures and panels} It is possible to refer to a specific panel in a figure, for instance figure \ref{f:panel1} or figure \ref{f:panel2}.
+Its also possible to refer to the entire figure, for instance figure \ref{f:graph1} or figure \ref{f:graph2}.
+It is also possible to refer to a panel within a figure by itself, for instance panel \subref{f:panel1} or panel \subref{f:panel2} in figure \ref{f:graph1}.
 
@@
 
-\paragraph{Another result in a panel} Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus, omnis voluptas assumenda est, omnis dolor repellendus.
-Figure \ref{f:panel1} displays cum soluta nobis est eligendi optio, cumque nihil impedit.
-Quod maxime placeat, facere possimus, omnis voluptas assumenda est in figure \ref{f:panel2}.
+\paragraph{Paragraph title} Et harum quidem rerum facilis est et expedita distinctio.
+Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus, omnis voluptas assumenda est, omnis dolor repellendus.
+Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet, ut et voluptates repudiandae sint et molestiae non recusandae.
 
-\subsection{A full-page graphs}
+\subsection{A full-page graph}
 

### 2023-09-03 paper.tex
Make paper LaTeX file easier to use

@@
 
+This is the abstract.
 Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
@@
 
-\paragraph{Positioning in the literature} Lorem ipsum dolor sit amet \citep{MS15}, consectetur adipiscing elit \citep[p.
-1305]{MS19}, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-\citet{M12,M14} enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-\citeauthor{EMM21} aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.\footnote{In reprehenderit in eu fugiat nulla pariatur \citep{LMS18a}.}
+\paragraph{Positioning in the literature} Sed ut perspiciatis, unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam eaque ipsa, quae ab illo inventore veritatis et quasi architecto beatae vitae dicta sunt, explicabo.
+Nemo enim ipsam voluptatem, quia voluptas sit, aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos, qui ratione voluptatem sequi nesciunt, neque porro quisquam est, qui dolorem ipsum, quia dolor sit amet consectetur adipiscing velit, sed quia non numquam eius modi tempora incidunt, ut labore et dolore magnam aliquam quaerat voluptatem.
 
@@
 
-\paragraph{Another paragraph} Sed ut perspiciatis, unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam eaque ipsa, quae ab illo inventore veritatis et quasi architecto beatae vitae dicta sunt, explicabo.
-Nemo enim ipsam voluptatem, quia voluptas sit, aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos, qui ratione voluptatem sequi nesciunt, neque porro quisquam est, qui dolorem ipsum, quia dolor sit amet consectetur adipiscing velit, sed quia non numquam eius modi tempora incidunt, ut labore et dolore magnam aliquam quaerat voluptatem.
+\paragraph{Paragraph with citations} Citations can be appended at the end of a sentence, in parenthesis \citep{LMS18a}.
+Citations can also include page numbers or other bibliographical information \citep[p.
+1305]{MS19}.
+Citations can be in text: for instance, \citet{M12} found this and \citet{M14} found that.
+It's also possible to collate several citations to the same author: for instance, \citet{M12,M14} found things.
+It is also possible to cite the authors of a study by name, or the year of a study: for instance, \citeauthor{EMM21} wrote a paper in \citeyear{EMM21}.
 
@@
 
-Ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat.
+This is a section. This is the section introduction. Below are the subsections.
 
@@
 
-At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint.
+Lists are not very common in scientific papers, but they are sometimes useful.
 
-\paragraph{Itemized lists} Oobcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga. Here is a list:
+\paragraph{Here is an itemized lists (with two levels)} Oobcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
 \begin{itemize}
@@
 
-\paragraph{Numbered lists} Omnis dolor repellendus. And here is a numbered list:
+\paragraph{Here is a numbered lists (with two levels)} Oobcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
 \begin{enumerate}
@@
 
-\subsection{Another subsection with text}
+\subsection{Some math}
 
-Sed ut perspiciatis, as showed in section \ref{s:lists}, unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam eaque ipsa, quae ab illo inventore veritatis et quasi architecto beatae vitae dicta sunt, explicabo.
-Nemo enim ipsam voluptatem, quia voluptas sit, aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos, qui ratione voluptatem sequi nesciunt, neque porro quisquam est, qui dolorem ipsum, quia dolor sit amet consectetur adipiscing velit, sed quia non numquam eius modi tempora incidunt, ut labore et dolore magnam aliquam quaerat voluptatem in section \ref{s:graphs}.
+Here are some Greek letters in math: $\alpha + \Gamma - \zeta \times \Lambda$.
+Here are some blackboard bold letters: $\mathbb{R} + \mathbb{Q} - \mathbb{Z}$.
+Here are some calligraphic letters: $\mathcal{A} - \mathcal{B} + \mathcal{Q}$.
+Here are some symbols: $Z \in Q \to R$ and $\sum_i [x(i)]^2 = \prod_j [y(j)]^2 = \int [g(h)]^2dh $.
+Here are some operators: $\max_x \pi(x) = \ln(u) \mapsto \cos(\Theta)$.
 
-\subsection{Some math and theorems}
+Here is some more random math, including display equations: Quis autem vel eum iure reprehenderit, qui in ea voluptate velit esse, quam nihil molestiae consequatur $\prod_i\alpha_i(\bm{X}) \geq 2 \implies \mathcal{X} \to 5$ qui in ea voluptate velit esse $\sum_{x=1}^4 \xi^x(j) \quad \forall j\in \mathbb{R}$.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum \eqref{e:type1}, excepteur sint occaecat cupidatat non proident $z^\ast/4$:
+\begin{align*}
+\frac{\mathcal{S}(z^*)}{1 -\theta + \theta \mathcal{S}(z^*)} &= \sum_{i=0}^{+\infty}F^{i} - \Lambda(i) = \left\lbrace\sum_{i=0}^{+\infty}G^{i} - \Omega(i)\right\rbrace\\
+\frac{\mathbb{S}(z^*)}{1 -\mu + \xi \mathbb{S}(z^*)} &= \prod_{i=0}^{+\infty}F^{i} - \Phi(i) = \left\lbrace\int_{i=0}^{+\infty}\frac{G^{i}}{\epsilon(i)} di\right\rbrace.
+\end{align*}
+
+\subsection{Some theorems}
+
+Here is a proposition with some more math:
 
@@
 \begin{equation}
-\bm{S}_{k_x}(z) \approx \frac{S(z)^x}{k / 23 -\zeta\gamma [45- S(z)] + \ln(y) - j^2+x(l)}.
+\sum_k\bm{S}_{k_x}(z) \approx \frac{S(z)^x}{k / 23 -\zeta\gamma [45- S(z)] + \ln(y) - j^2+x(l)}.
 \label{e:type1}\end{equation}
@@
 \begin{equation}
-\sum_{j=0}^{+\infty}\mathcal{S}^j(z^{\ast}) = S(z) \times \mathbb{E}(N(z)) + \alpha \beta - \left[\sum_{i=0}^{+\infty}\mathcal{F}^{i}-\exp(\lambda \mu)\right].
+\sum_{j=0}^{+\infty}\mathcal{S}^j(z^{\ast}) = S(z) \times \mathbb{E}(N(z)) + \mathbb{\mu} \bm{\beta} - \prod_{i=0}^{+\infty}\mathcal{F}^{i}-\exp(\lambda \mu).
 \label{e:type1steps}\end{equation}
@@
 
-\begin{proof} There is a simple proof to this result. It gives equation \eqref{e:type1steps}.\end{proof}
+\begin{proof} Here is the proof to the proposition.\end{proof}
 
-Quis autem vel eum iure reprehenderit, qui in ea voluptate velit esse, quam nihil molestiae consequatur $\prod_i\alpha_i(\bm{X}) \geq 2 \implies \mathcal{X} \to 5$ qui in ea voluptate velit esse $\sum_{x=1}^4 \xi^x(j) \quad \forall j\in \mathbb{R}$.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum \eqref{e:type1}, excepteur sint occaecat cupidatat non proident $z^\ast/4$:
-\begin{equation*}
-\frac{\mathcal{S}(z^*)}{1 -\theta + \theta \mathcal{S}(z^*)} = \prod_{i=0}^{+\infty}F^{i} - \Lambda(i) = \left\lbrace\prod_{i=0}^{+\infty}G^{i} - \Omega(i)\right\rbrace.
-\end{equation*}
-From this definition we obtain the following results (details of the proof are relegated to appendix~\ref{a:appendix1}):
+Here is a lemma:
 
@@
 \label{e:cv}\end{equation}
-Consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua $\mathcal{Z}(\alpha)$. \end{lemma}
+Consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua $\mathcal{Z}(\alpha)$. 
+\begin{equation}
+\frac{\sum_i z^i}{\prod_i q^i} = \frac{\int_{0}^{\infty} \alpha(i) \cdot [1-\beta]\,di}{1-\exp(\alpha)\sin(\beta)}.
+\label{e:cv}\end{equation}
+\end{lemma}
 
-Ut enim ad minima veniam, quis nostrum exercitationem ullam corporis suscipit laboriosam, nisi ut aliquid ex ea commodi consequatur?
-Combining equation \eqref{e:cv} with other results, we obtain the following corollary (the proof is in appendix~\ref{a:subappendix}):
+\begin{proof} Here is the proof to the lemma.\end{proof}
+
+Here is a corollary:
 
@@
 \begin{equation*}
-\mathbb{E}(N(z^*)) \approx \frac{1-\mathbb{P}(\alpha\pi)}{1-\pi}- \frac{f(y)}{z(p)^*} +P(\Gamma).
+\mathbb{E}(N(z^*)) \approx \frac{1-\mathbb{P}(\alpha\pi)}{1-\pi}- \frac{f(y)}{z(p)^*} + P(\Gamma).
 \end{equation*}\end{corollary}
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-
-\subsection{Subsection with graphs}
-
+Here is a section with a variety of graphs.
 At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
 
-\paragraph{Paragraph with references to figures and panels} It is possible to refer to a specific panel in a figure, for instance figure \ref{f:panel1} or figure \ref{f:panel2}.
-Its also possible to refer to the entire figure, for instance figure \ref{f:graph1} or figure \ref{f:graph2}.
-It is also possible to refer to a panel within a figure by itself, for instance panel \subref{f:panel1} or panel \subref{f:panel2} in figure \ref{f:graph1}.
+\subsection{Subsection with graphs at the top of the page}
+
+A simple two-panel graph is on figure \ref{f:graph1}. It will be placed at the top of the page, just about here.
 
@@
 
-\paragraph{Paragraph title} Et harum quidem rerum facilis est et expedita distinctio.
+Et harum quidem rerum facilis est et expedita distinctio.
 Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus, omnis voluptas assumenda est, omnis dolor repellendus.
@@
 
-\subsection{A full-page graph}
+\subsection{Subsection with references to figures and panels}
 
-A full-page graph is on figure \ref{f:graph2}.
-Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus, omnis voluptas assumenda est, omnis dolor repellendus.
+As usual, it is easy to refer to a figure: see figure \ref{f:graph1}.
+It is possible to refer to a specific panel in a figure, for instance figure \ref{f:panel1} or figure \ref{f:panel2}.
+Its also possible to refer to the entire figure, for instance figure \ref{f:graph1} or figure \ref{f:graph2}.
+It is also possible to refer to a panel within a figure by itself, for instance panel \subref{f:panel1} or panel \subref{f:panel2} in figure \ref{f:graph1}.
+
+\subsection{A subsection with a full-page graph}
+
+A full-page graph is on figure \ref{f:graph2}. It will be placed in a full page about here.
 
@@
 \caption{Caption for the graph}
-\note{Note for the graph. Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat from \citet[section 3]{MS22b}.}
+\note{Note for the graph. Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat.}
 \label{f:graph2}\end{figure}
 
-\paragraph{Another paragraph} At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
-Et harum quidem rerum facilis est et expedita distinctio.
-Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus, omnis voluptas assumenda est, omnis dolor repellendus.
-Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet, ut et voluptates repudiandae sint et molestiae non recusandae.
-Itaque earum rerum hic tenetur a sapiente delectus, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat.
-
-\paragraph{Yet another paragraph} At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
+At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
 Et harum quidem rerum facilis est et expedita distinctio.
@@
 
+Here is a section with a variety of tables.
 Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet, ut et voluptates repudiandae sint et molestiae non recusandae.
@@
 
-\subsection{A subsection title}
+\subsection{A subsection with a simple table}
+
+Table \ref{t:table1} is a simple table with one panel.
+Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet, ut et voluptates repudiandae sint et molestiae non recusandae.
+Itaque earum rerum hic tenetur a sapiente delectus, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat.
+
+\begin{table}[t]
+\caption{Table caption}
+\begin{tabular*}{\textwidth}[]{p{3.3cm}@{\extracolsep\fill}cccc}
+\toprule
+& Column 1 &  Column 2 &  Column 3  &  Column 4 \\
+\midrule
+Line 1: & A & B & C  & d \\
+Line 2: & E &  F & G  & H   \\
+Line 3: & K & V & P  & K  \\
+Line 4: & J & M & N  & K  \\
+\bottomrule
+\end{tabular*}
+\note{Note for table.}
+\label{t:table1}\end{table}
 
@@
 Itaque earum rerum hic tenetur a sapiente delectus, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat.
-Table \ref{t:table} shows many things.
 
-At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
+\subsection{A subsection with a multi-panel table}
+
+Table \ref{t:table2} is a more sophisticated table with several panels.
+Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet, ut et voluptates repudiandae sint et molestiae non recusandae.
+Itaque earum rerum hic tenetur a sapiente delectus, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat.
 
 \begin{table}[t]
-\caption{Caption for the table}
-\begin{tabular*}{\textwidth}[]{p{3.3cm}@{\extracolsep\fill}cccc}
+\caption{Table caption}
+\begin{tabular*}{\textwidth}[]{p{2.5cm}@{\extracolsep\fill}ccccccc}
 \toprule
-& Timeline & Natural rate & Monetary  & Government \\
-&  & of interest &  policy & spending\\
+    & Column 1 &  Column 2 &  Column 3  &  Column 4 &  Column 5 &  Column 6 &  Column 7 \\
 \midrule
-\multicolumn{5}{l}{A. Qui dolorem eum fugiat}\\
-ZLB: & $t\in (0,T) $ & $r^n<0$ & $i=0$  & -- \\
-Normal times: & $t>T$ &  $r^n>0$ & $i= r^n + \phi \pi$  & --   \\
+\multicolumn{8}{l}{A. First panel}\\
+Line 1: & A & C & V  & -- & -- & K & A\\
+Line 2: & X &  H & O  & -- & -- & K & A  \\
 \midrule
-\multicolumn{5}{l}{B. Nam libero tempore, cum soluta nobis}\\
-ZLB: & $t\in (0,T)$ & $r^n<0$ & $i=0$  & -- \\
-Forward guidance: & $t\in (T,T+\Delta)$ & $r^n>0$ & $i=0$  & --\\
-Normal times: & $t>T+\Delta$ & $r^n>0$ & $i= r^n + \phi \pi$  & -- \\
+\multicolumn{8}{l}{B. Second panel}\\
+Line 3: & U & B & J  & K & A & K & A\\
+Line 4: & N & Y & T  & L & T & K & A\\
+Line 5: & G & S & Q  & P & Q & K & A\\
 \midrule
-\multicolumn{5}{l}{C. At vero eos et accusamus}\\
-ZLB: & $t\in (0,T) $ & $r^n<0$ & $i=0$  & $g>0$  \\
-Normal times: & $t>T$ & $r^n>0$ & $i= r^n + \phi \pi$  & $g=0$   \\
+\multicolumn{8}{l}{C. Third panel}\\
+Line 3: & U & B & J  & K & K & K & A\\
+Line 4: & N & Y & T  & L & T & K & A\\
+Line 5: & G & S & Q  & P & Q & K & A\\
 \bottomrule
 \end{tabular*}
-\note{Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet, ut et voluptates repudiandae sint et molestiae non recusandae.
-Ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat \citep[table 1]{MS21a}.}
-\label{t:table}\end{table}
+\note{Note for table.
+Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet, ut et voluptates repudiandae sint et molestiae non recusandae.
+Ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat.}
+\label{t:table2}\end{table}
 
-\subsection{Another subsection title}
+Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet, ut et voluptates repudiandae sint et molestiae non recusandae.
+Itaque earum rerum hic tenetur a sapiente delectus, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat.
 
-Qui dolorem eum fugiat, quo voluptas nulla pariatur?
-At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
-Et harum quidem rerum facilis est et expedita distinctio.
-Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus, omnis voluptas assumenda est, omnis dolor repellendus.
-Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet, ut et voluptates repudiandae sint et molestiae non recusandae.
-Itaque earum rerum hic tenetur a sapiente delectus, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat (figure~\ref{f:graph2}).
+\section{A section with cross-references}
+
+As usual, it is possible to reference an equation, such as equation \eqref{e:cv}.
+It is also possible to reference a section, such as section \ref{s:graphs}, or a subsection, such as section \ref{s:lists}.
+It is also possible to reference an appendix, such as appendix \ref{a:appendix1}, or a subsection in an appendix, such as appendix~\ref{a:subappendix}.
+Of course it is possible to reference figures, such as figure \ref{f:graph2}, or tables, such as table~\ref{t:table1}.
 
@@
 
-To conclude, at vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
-
+Here is a conclusion.
+At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
 At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint.
@@
 
-We extend the model of section~\ref{s:section} by introducing a quadratic cost.
-Obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
+This is a first appendix. Obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
 
@@
 
+This is a subsection in the appendix.
 At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
 
-\subsection{Another subsection}
+\subsection{A subsection with results}
 
- Et harum quidem rerum facilis est et expedita distinctio.
-Nam libero tempore, cum soluta nobis est eligendi optio cumque nihil impedit quo minus id quod maxime placeat facere possimus, omnis voluptas assumenda est, omnis dolor repellendus \citep{MS22a}.
-A corrolary in the appendix is as follows:
+Here is a result in the appendix:
 
@@
 
-\paragraph{A paragraph with some math} Temporibus autem quibusdam $\xi$ et aut officiis debitis aut rerum necessitatibus saepe eveniet ut et voluptates repudiandae sint et molestiae non recusandae $1-\gamma$.
-Itaque earum rerum hic $S(z^*)$ tenetur a sapiente delectus $\mathcal{B}^\theta$, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat $\mathcal{V}^i$.
-Aggregating these scenarios, we obtain the continuation value:
+Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
+tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
+quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
+consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
+cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
+proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+
+\subsection{A subsection with math}
+
+Here are math and an equation in the appendix---see equation \eqref{e:appendix1}.
+Temporibus autem quibusdam $\xi$ et aut officiis debitis aut rerum necessitatibus saepe eveniet ut et voluptates repudiandae sint et molestiae non recusandae $1-\gamma$.
+Itaque earum rerum hic $S(z^*)$ tenetur a sapiente delectus $\mathcal{B}^\theta$, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat $\mathcal{V}^i$:
 \begin{equation}
@@
 
-\paragraph{Paragraph with links to appendix equations} Ut enim ad minima veniam, quis nostrum exercitationem ullam corporis suscipit laboriosam, nisi ut aliquid ex ea commodi consequatur $\mathcal{C}$?
+Ut enim ad minima veniam, quis nostrum exercitationem ullam corporis suscipit laboriosam, nisi ut aliquid ex ea commodi consequatur $\mathcal{C}$?
 Quis autem vel eum iure reprehenderit qui in ea voluptate velit esse quam nihil molestiae consequatur, vel illum qui dolorem eum fugiat quo voluptas nulla pariatur?
-Equation \eqref{e:appendix1} shows that autem vel eum iure reprehenderit qui in ea voluptate velit esse quam nihil molestiae consequatur.
+Autem vel eum iure reprehenderit qui in ea voluptate velit esse quam nihil molestiae consequatur.
 
@@
 
-At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga in figure \ref{f:appendix1}.
+Here is a large, simple figure in the appendix (see figure \ref{f:appendix1}).
+At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
 
@@
 
-\section{Another section}\label{a:appendix2}
+\section{Another appendix}\label{a:appendix2}
 
-At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti.
+Here is a second appendix. At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti.
 
@@
 
-At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga, as showed in figure \ref{f:appendix2}.
+Figure \ref{f:appendix2} shows an even larger figure in the appendix.
+At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
+Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus.
 
@@
 
-\subsection{Final subsection with footnote and references}\label{a:subappendix}
+\subsection{Final subsection with footnote, references, and URL}\label{a:subappendix}
 
-Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt $\mathcal{V}^i > \mathcal{V}^r$ \citep{MS21b}.\footnote{The reference goes to the reference list at the end of the main text.} Sed ut perspiciatis unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam, eaque ipsa quae ab illo inventore veritatis et quasi architecto beatae vitae dicta sunt explicabo.
-This results are summarized in a repository with the following URL: \url{https://github.com/pmichaillat/latex-paper}.
+Here is a sentence with two citations \citep{MS19,MS21b}.
+The references go to the reference list at the end of the main text.
+Here is a sentence with a footnote.\footnote{Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt.}
+
+It is possible to insert an URL: \url{https://github.com/pmichaillat/latex-paper}. Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
+tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
+quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
+consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
+cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
+proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
 

### 2023-09-03 appendix.tex
Use Source Code Pro as monospace font; clean templates

@@
 
-The references from the paper file are available in the appendix as long as the appendix is compiled after the paper and intermediary LaTeX-related files are not deleted.
+This is a subsection in the appendix.
+Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
+consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
+cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
+proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
 
+\subsection{A subsection with references from the main text}
+
+The references from the main text (in the \texttt{paper.tex} file) are available in the online appendix as long as the appendix is compiled after the main text and intermediary LaTeX-related files (especially the \texttt{paper.aux} file) are not deleted.
+Here are some references to the main text:
 \begin{itemize}    
@@
     \item Figure \ref{f:graph1} provided a lot of information---especially the plot on figure \ref{f:panel2}. 
+    \item Section \ref{s:graphs} presented new results, and so did section \ref{s:lists}.
+    \item Table~\ref{t:table1} contained interesting information.
 \end{itemize}
 
-\subsection{Another subsection}
+\subsection{A subsection with results}
 
-At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
-
-\subsection{Yet another subsection}
-
-Et harum quidem rerum facilis est et expedita distinctio.
-Nam libero tempore, cum soluta nobis est eligendi optio cumque nihil impedit quo minus id quod maxime placeat facere possimus, omnis voluptas assumenda est, omnis dolor repellendus \citep{MS22a}.
-A corrolary in the appendix is as follows:
+Here is a result in the appendix:
 
@@
 \begin{equation*}
-\mathbb{E}(\Omega) = \mathbb{P}(\omega\cdot \mu - \xi) - \sum_{i=0}^{m}\sum_{j=i}^{n} \sigma(i,j) + 123^{56}.
+\mathbb{E}(\Omega) = \mathbb{P}(\omega\cdot \mu - \xi) - \sum_{i=0}^{m}\sum_{j=-\infty}^{n} \sigma(i,j) + 123^{56}.
 \end{equation*}\end{corollary}
 
-\paragraph{Paragraph with some math} Temporibus autem quibusdam $\xi$ et aut officiis debitis aut rerum necessitatibus saepe eveniet ut et voluptates repudiandae sint et molestiae non recusandae $1-\gamma$.
-Itaque earum rerum hic $S(z^*)$ tenetur a sapiente delectus $\mathcal{B}^\theta$, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat $\prod_i\mathcal{V}^i \mapsto \mathbb{Q} + \mathbb{\Lambda(\beta + \gamma)}$.
-Aggregating these scenarios, we obtain the continuation value:
+Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
+tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
+quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
+consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
+cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
+proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+
+\subsection{A subsection with math}
+
+Here are math and an equation in the appendix---see equation \eqref{e:appendix1}.
+Temporibus autem quibusdam $\xi$ et aut officiis debitis aut rerum necessitatibus saepe eveniet ut et voluptates repudiandae sint et molestiae non recusandae $1-\gamma$.
+Itaque earum rerum hic $S(z^*)$ tenetur a sapiente delectus $\mathcal{B}^\theta$, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat $\mathbb{V}^i$:
 \begin{equation}
-\mathbb{V}^r = (1-\gamma) \times 0 +\gamma S(z^*) v^s+\gamma [1-S(z^*)] \times \frac{\sum_i\mathcal{V}^i}{2-c}.
+\mathbb{V}^r = (1-\gamma) \times 0 +\gamma S(z^*) v^s+\gamma [1-S(z^*)] \mathcal{V}^i-c.
 \label{e:appendix1}\end{equation}
 
-\paragraph{Paragraph with links to appendix equations} Ut enim ad minima veniam, quis nostrum exercitationem ullam corporis suscipit laboriosam, nisi ut aliquid ex ea commodi consequatur $\mathcal{C}$?
+Ut enim ad minima veniam, quis nostrum exercitationem ullam corporis suscipit laboriosam, nisi ut aliquid ex ea commodi consequatur?
 Quis autem vel eum iure reprehenderit qui in ea voluptate velit esse quam nihil molestiae consequatur, vel illum qui dolorem eum fugiat quo voluptas nulla pariatur?
-Equation \eqref{e:appendix1} shows that autem vel eum iure reprehenderit qui in ea voluptate velit esse quam nihil molestiae consequatur.
+Autem vel eum iure reprehenderit qui in ea voluptate velit esse quam nihil molestiae consequatur.
 
-\subsection{A larger figure, without panel, in the appendix}
+\section{Another appendix}\label{a:appendix2}
 
+Here is a second appendix. At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti.
+
+\subsection{Larger figure, without panel, in the appendix}
+
+Here is a large, simple figure in the appendix (see figure \ref{f:appendix1}).
 At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
-This is showed in figure \ref{f:appendix1}.
 
@@
 \includegraphics[scale=\mfig,page=1]{\pdf}
-\caption{A caption for the larger graph}
-\note{A note for the larger graph. Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus.}
+\caption{Caption for the larger graph}
+\note{Note for the larger graph. Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus.}
 \label{f:appendix1}\end{figure}
 
-\section{Another appendix}\label{a:appendix2}
-
-At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti.
+Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
+tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
+quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
+consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
+cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
+proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
 
@@
 
-At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga, as showed in figure \ref{f:appendix2}.
+Figure \ref{f:appendix2} shows an even larger figure in the appendix.
+At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
+Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus.
 
-\begin{figure}[t]
+\begin{figure}[p]
 \includegraphics[scale=\lfig,page=3]{\pdf}
-\caption{A caption for the even larger graph}
-\note{A note for the larger graph. Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus.}
+\caption{Caption for the even larger graph}
+\note{Note for the larger graph. Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus.}
 \label{f:appendix2}\end{figure}
 
-\subsection{Subsection with footnote and references}\label{a:subappendix}
+Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
+tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
+quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
+consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
+cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
+proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
 
-Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt if $\mathcal{V}^i > \mathcal{V}^r$ \citep{MS21b}.\footnote{The reference goes to its own reference list at the end of the appendix---unlike when the appendix was at the end of the main text.} Sed ut perspiciatis unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam, eaque ipsa quae ab illo inventore veritatis et quasi architecto beatae vitae dicta sunt explicabo.
-This results are summarized in a repository with the following URL: \url{https://github.com/pmichaillat/latex-paper}.
+\subsection{Subsection with references}\label{a:subappendix}
+
+Here is a sentence with some citations: \citet{MS19,MS21b} found something but that is an uncommon result \citep{M12,M14}.
+The references go to their own reference list at the end of the appendix---unlike when the appendix was at the end of the main text.
+
+Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
+tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
+quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
+consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
+cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
+proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+
+\subsection{Subsection with a foonote}
+
+Here is a sentence with a footnote.\footnote{Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt.}
+
+Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
+quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
+consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
+cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
+proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+
+\subsection{Subsection with an URL}
+
+It is possible to insert an URL: \url{https://github.com/pmichaillat/latex-paper}. Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
+tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
+quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
+consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
+cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
+proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
 

### 2023-09-03 paper.tex
Use Source Code Pro as monospace font; clean templates

@@
 Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet, ut et voluptates repudiandae sint et molestiae non recusandae.
-Itaque earum rerum hic tenetur a sapiente delectus, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat.
 
@@
 This is a subsection in the appendix.
-At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
+Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
+consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
+cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
+proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
 
@@
 Temporibus autem quibusdam $\xi$ et aut officiis debitis aut rerum necessitatibus saepe eveniet ut et voluptates repudiandae sint et molestiae non recusandae $1-\gamma$.
-Itaque earum rerum hic $S(z^*)$ tenetur a sapiente delectus $\mathcal{B}^\theta$, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat $\mathcal{V}^i$:
+Itaque earum rerum hic $S(z^*)$ tenetur a sapiente delectus $\mathcal{B}^\theta$, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat $\mathbb{V}^i$:
 \begin{equation}
@@
 
-Ut enim ad minima veniam, quis nostrum exercitationem ullam corporis suscipit laboriosam, nisi ut aliquid ex ea commodi consequatur $\mathcal{C}$?
+Ut enim ad minima veniam, quis nostrum exercitationem ullam corporis suscipit laboriosam, nisi ut aliquid ex ea commodi consequatur?
 Quis autem vel eum iure reprehenderit qui in ea voluptate velit esse quam nihil molestiae consequatur, vel illum qui dolorem eum fugiat quo voluptas nulla pariatur?
 Autem vel eum iure reprehenderit qui in ea voluptate velit esse quam nihil molestiae consequatur.
+
+\section{Another appendix}\label{a:appendix2}
+
+Here is a second appendix. At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti.
 
@@
 
-\section{Another appendix}\label{a:appendix2}
-
-Here is a second appendix. At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti.
+Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
+tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
+quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
+consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
+cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
+proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
 
@@
 
-\begin{figure}[t]
+\begin{figure}[p]
 \includegraphics[scale=\lfig,page=3]{\pdf}
@@
 
-\subsection{Final subsection with footnote, references, and URL}\label{a:subappendix}
+Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
+tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
+quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
+consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
+cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
+proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
 
-Here is a sentence with two citations \citep{MS19,MS21b}.
-The references go to the reference list at the end of the main text.
-Here is a sentence with a footnote.\footnote{Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt.}
+\subsection{Subsection with references}\label{a:subappendix}
+
+Here is a sentence with some citations: \citet{MS19,MS21b} found something but that is an uncommon result \citep{M12,M14}.
+The references go to the reference list at the end of the main text---so appendix and main text share a reference list.
+
+Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
+tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
+quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
+consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
+cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
+proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+
+\subsection{Subsection with a foonote}
+
+Here is a sentence with a footnote.\footnote{Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt.} The numbering of the footnotes, just like the numbering of the pages, continues from the main text to the appendix.
+
+Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
+quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
+consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
+cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
+proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+
+\subsection{Subsection with an URL}
 

### 2023-09-06 paper.tex
Improve list spacing

@@
 
-\paragraph{Here is an itemized lists (with two levels)} Oobcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
+\paragraph{Here is an itemized lists (with two levels)} Oobcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga:
 \begin{itemize}
@@
 \item Nam libero tempore, cum soluta nobis est eligendi optio.
-\item emporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet.
+\item Emporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet. Nam libero tempore, cum soluta nobis est eligendi optio:
 \begin{itemize}
-	\item cumque nihil impedit
-	\item quo minus id, quod maxime placeat
-	\item facere possimus, omnis voluptas assumenda est
+	\item Cumque nihil impedit
+	\item Quo minus id, quod maxime placeat
+	\item Facere possimus, omnis voluptas assumenda est
 \end{itemize}
@@
 \item Nam libero tempore, cum soluta nobis est eligendi optio.
-\item emporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet.
+\item Emporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet. Nam libero tempore, cum soluta nobis est eligendi optio:
 \begin{enumerate}
-	\item cumque nihil impedit
-	\item quo minus id, quod maxime placeat
-	\item facere possimus, omnis voluptas assumenda est
+	\item Cumque nihil impedit
+	\item Quo minus id, quod maxime placeat
+	\item Facere possimus, omnis voluptas assumenda est
 \end{enumerate}
@@
 
-Here are some Greek letters in math: $\alpha + \Gamma - \zeta \times \Lambda$.
-Here are some blackboard bold letters: $\mathbb{R} + \mathbb{Q} - \mathbb{Z}$.
-Here are some calligraphic letters: $\mathcal{A} - \mathcal{B} + \mathcal{Q}$.
-Here are some symbols: $Z \in Q \to R$ and $\sum_i [x(i)]^2 = \prod_j [y(j)]^2 = \int [g(h)]^2dh $.
-Here are some operators: $\max_x \pi(x) = \ln(u) \mapsto \cos(\Theta)$.
+\begin{itemize}
+\item Here are some Greek letters in math: $\alpha + \Gamma - \zeta \times \Lambda$.
+\item Here are some blackboard bold letters: $\mathbb{R} + \mathbb{Q} - \mathbb{Z}$.
+\item Here are some calligraphic letters: $\mathcal{A} - \mathcal{B} + \mathcal{Q}$. 
+\item Here are some symbols: $Z \in Q \to R$ and $\sum_i [x(i)]^2 = \prod_j [y(j)]^2 = \int [g(h)]^2dh $.
+\item Here are some operators: $\max_x \pi(x) = \ln(u) \mapsto \cos(\Theta)$.
+\end{itemize}
 

### 2023-09-07 paper.tex
Update format of numbered lists

@@
 Sunt in culpa qui officia deserunt mollit anim id est laborum.
+Excepteur sint occaecat cupidatat non proident.
+Sunt in culpa qui officia deserunt mollit anim id est laborum.
 Excepteur sint occaecat cupidatat non proident.} Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
@@
 
-\subsection{Subsection with lists}\label{s:lists}
+\subsection{Subsection with itemized lists}\label{s:lists}
 
-Lists are not very common in scientific papers, but they are sometimes useful.
-
-\paragraph{Here is an itemized lists (with two levels)} Oobcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga:
+Lists are not very common in scientific papers, but they are sometimes useful. Here is an itemized list with two levels:
 \begin{itemize}
@@
 \end{itemize}
+\item Et harum quidem rerum facilis est et expedita distinctio.
+\item Nam libero tempore, cum soluta nobis est eligendi optio.
 \end{itemize}
 
-\paragraph{Here is a numbered lists (with two levels)} Oobcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
+\subsection{Subsection with numbered lists}\label{s:lists}
+
+And here is a numbered lists with two levels:
 \begin{enumerate}
@@
 \end{enumerate}
+\item Et harum quidem rerum facilis est et expedita distinctio.
+\item Nam libero tempore, cum soluta nobis est eligendi optio.
 \end{enumerate}
@@
 \item Here are some calligraphic letters: $\mathcal{A} - \mathcal{B} + \mathcal{Q}$. 
-\item Here are some symbols: $Z \in Q \to R$ and $\sum_i [x(i)]^2 = \prod_j [y(j)]^2 = \int [g(h)]^2dh $.
-\item Here are some operators: $\max_x \pi(x) = \ln(u) \mapsto \cos(\Theta)$.
+\item Here are some symbols: $Z \in Q \to R$ and $\sum_i [x(i)]^2 \subset \prod_j [y(j)]^2 \geq \int [g(h)]^2dh $.
+\item Here are some operators: $\max_x \pi(x) = \ln(u)/\exp(v) \mapsto \cos(\Theta) + \sin(\Omega)$.
 \end{itemize}
 
-Here is some more random math, including display equations: Quis autem vel eum iure reprehenderit, qui in ea voluptate velit esse, quam nihil molestiae consequatur $\prod_i\alpha_i(\bm{X}) \geq 2 \implies \mathcal{X} \to 5$ qui in ea voluptate velit esse $\sum_{x=1}^4 \xi^x(j) \quad \forall j\in \mathbb{R}$.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum \eqref{e:type1}, excepteur sint occaecat cupidatat non proident $z^\ast/4$:
+\subsection{Some equations}
+
+Here are equations in text: $\prod_i\alpha_i(\bm{X}) \geq 2 \implies \mathcal{X} \to 5$ but  $\sum_{x=1}^4 \xi^x(j) \quad \forall j\in \mathbb{R}$. Here are equations without labels:
 \begin{align*}
@@
 \end{align*}
+And here is a single equation with a label:
+\begin{equation}
+\sum_{j=0}^{+\infty}\mathcal{S}^j(z^{\ast}) = S(z) \times \mathbb{E}(N(z)) + \mathbb{\mu} \bm{\beta} - \prod_{i=0}^{+\infty}\mathcal{F}^{i}-\exp(\lambda \mu).
+\label{e:type1steps}\end{equation}
+Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
+tempor incididunt ut labore et dolore magna aliqua.
 
@@
 
-\begin{proposition}\label{p:type1}  Lorem ipsum dolor sit amet $z \subset \mathbb{R}$, consectetur adipiscing elit
+\begin{proposition}\label{p:type1}  Lorem ipsum dolor sit amet, consectetur adipiscing elit:
 \begin{equation}
@@
 \label{e:type1}\end{equation}
-Ut enim ad minim veniam, sunt in culpa qui officia deserunt mollit anim id est laborum:
-\begin{equation}
-\sum_{j=0}^{+\infty}\mathcal{S}^j(z^{\ast}) = S(z) \times \mathbb{E}(N(z)) + \mathbb{\mu} \bm{\beta} - \prod_{i=0}^{+\infty}\mathcal{F}^{i}-\exp(\lambda \mu).
-\label{e:type1steps}\end{equation}
+Ut enim ad minim veniam, sunt in culpa qui officia deserunt mollit anim id est laborum. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
 \end{proposition}
@@
 
-\begin{lemma}\label{p:cv} Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat $\alpha$:
+\begin{lemma}\label{p:cv} Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat:
 \begin{equation}
@@
 \label{e:cv}\end{equation}
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. 
 \end{lemma}
 
-\begin{proof} Here is the proof to the lemma.\end{proof}
-
-Here is a corollary:
+And here is a corollary following the lemma:
 
@@
 \mathbb{E}(N(z^*)) \approx \frac{1-\mathbb{P}(\alpha\pi)}{1-\pi}- \frac{f(y)}{z(p)^*} + P(\Gamma).
-\end{equation*}\end{corollary}
+\end{equation*}
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.\end{corollary}
 

### 2023-09-17 appendix.tex
Update URL font to monospaced
Add hyperlinks to footnotes

@@
 
+\subsection{Subsection with a foonote}
+
+Here is a sentence with a footnote.\footnote{Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt.}
+
+Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
+quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
+consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
+cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
+proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+
 \section{Another appendix}\label{a:appendix2}
 
-Here is a second appendix. At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti.
+Here is a second appendix.
+At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti.
+Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non
+proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+
+\subsection{Subsection with references}\label{a:subappendix}
+
+Here is a sentence with some citations: \citet{MS19,MS21b} found something but that is an uncommon result \citep{M12,M14}.
+The references go to their own reference list at the end of the appendix.
+By contrast, when the appendix is at the end of the main text, the appendix shares the list of references with the main text.
+
+Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
+tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
+quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
+consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
+cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
+proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+
+\subsection{Subsection with a URL}
+
+Here is a URL: \url{https://github.com/pmichaillat/latex-paper}. Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
+tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
+quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
+consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
+cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
+proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
 
@@
 
-Figure \ref{f:appendix2} shows an even larger figure in the appendix.
-At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
-Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus.
-
 \begin{figure}[p]
@@
 \label{f:appendix2}\end{figure}
+
+Figure \ref{f:appendix2} shows an even larger figure in the appendix.
+At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
+Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus.
 
@@
 
-\subsection{Subsection with references}\label{a:subappendix}
-
-Here is a sentence with some citations: \citet{MS19,MS21b} found something but that is an uncommon result \citep{M12,M14}.
-The references go to their own reference list at the end of the appendix---unlike when the appendix was at the end of the main text.
-
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
-tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
-quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
-consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
-cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
-proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
-
-\subsection{Subsection with a foonote}
-
-Here is a sentence with a footnote.\footnote{Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt.}
-
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
-quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
-consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
-cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
-proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
-
-\subsection{Subsection with an URL}
-
-It is possible to insert an URL: \url{https://github.com/pmichaillat/latex-paper}. Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
-tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
-quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
-consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
-cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
-proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
-
 \bibliography{\bib}

### 2023-09-17 paper.tex
Correct typos
Add hyperlinks to footnotes

@@
 Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
 Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
 Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
 
@@
 
-\subsection{Subsection with itemized lists}\label{s:lists}
+\subsection{Subsection with itemized lists}
 
@@
 Consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua $\mathcal{Z}(\alpha)$. 
-\begin{equation}
+\begin{equation*}
 \frac{\sum_i z^i}{\prod_i q^i} = \frac{\int_{0}^{\infty} \alpha(i) \cdot [1-\beta]\,di}{1-\exp(\alpha)\sin(\beta)}.
-\label{e:cv}\end{equation}
+\end{equation*}
 Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. 
@@
 
-A simple two-panel graph is on figure \ref{f:graph1}. It will be placed at the top of the page, just about here.
+A simple two-panel graph is on figure \ref{f:graph1}.
+It will be placed at the top of the page, just about here.
+Et harum quidem rerum facilis est et expedita distinctio.
+Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus, omnis voluptas assumenda est, omnis dolor repellendus.
+Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet, ut et voluptates repudiandae sint et molestiae non recusandae.
 
@@
 \label{f:graph1}\end{figure}
-
-Et harum quidem rerum facilis est et expedita distinctio.
-Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus, omnis voluptas assumenda est, omnis dolor repellendus.
-Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet, ut et voluptates repudiandae sint et molestiae non recusandae.
 
@@
 % Fill out appendix:
-\newpage
 \appendix

### 2023-10-14 paper.tex
Remove unnecessary figure commands

@@
 \begin{figure}[t]
-\subcaptionbox{Caption for first panel\label{f:panel1}}{\includegraphics[scale=\sfig,page=1]{\pdf}}\hfill
-\subcaptionbox{Caption for second panel\label{f:panel2}}{\includegraphics[scale=\sfig,page=2]{\pdf}}
+\subcaptionbox{Caption for first panel\label{f:panel1}}{\includegraphics[scale=0.2,page=1]{\pdf}}\hfill
+\subcaptionbox{Caption for second panel\label{f:panel2}}{\includegraphics[scale=0.2,page=2]{\pdf}}
 \caption{Caption for the graph}
@@
 \begin{figure}[p]
-\subcaptionbox{Caption for panel}{\includegraphics[scale=\sfig,page=1]{\pdf}}\hfill
-\subcaptionbox{Caption for panel}{\includegraphics[scale=\sfig,page=2]{\pdf}}\vfig
-\subcaptionbox{Caption for panel}{\includegraphics[scale=\sfig,page=3]{\pdf}}\hfill
-\subcaptionbox{Caption for panel}{\includegraphics[scale=\sfig,page=4]{\pdf}}\vfig
-\subcaptionbox{Caption for panel}{\includegraphics[scale=\sfig,page=5]{\pdf}}\hfill
-\subcaptionbox{Caption for panel}{\includegraphics[scale=\sfig,page=1]{\pdf}}
+\subcaptionbox{Caption for panel}{\includegraphics[scale=0.2,page=1]{\pdf}}\hfill
+\subcaptionbox{Caption for panel}{\includegraphics[scale=0.2,page=2]{\pdf}}\fspace
+\subcaptionbox{Caption for panel}{\includegraphics[scale=0.2,page=3]{\pdf}}\hfill
+\subcaptionbox{Caption for panel}{\includegraphics[scale=0.2,page=4]{\pdf}}\fspace
+\subcaptionbox{Caption for panel}{\includegraphics[scale=0.2,page=5]{\pdf}}\hfill
+\subcaptionbox{Caption for panel}{\includegraphics[scale=0.2,page=1]{\pdf}}
 \caption{Caption for the graph}
@@
 \begin{figure}[t]
-\includegraphics[scale=\mfig,page=1]{\pdf}
+\includegraphics[scale=0.3,page=1]{\pdf}
 \caption{Caption for the larger graph}
@@
 \begin{figure}[p]
-\includegraphics[scale=\lfig,page=3]{\pdf}
+\includegraphics[scale=0.4,page=3]{\pdf}
 \caption{Caption for the even larger graph}

### 2023-12-05 paper_contents/paper.tex
data collection tool: Included data colletion tool

--- 2023-10-14 paper.tex
@@
 
-% Fill out content of paper:
 \begin{document}
-\title{Paper Title}
-\author{Author 1, Author 2
-\thanks{Author 1: University 1.
-Author 2: University 2.
-We thank colleagues for helpful comments and discussions.
-This work was supported by a grant [grant number]; another grant [grant number]; and a foundation.}}
-\date{Month Year}                       
+\title{A high quality X-ray diffraction dataset}
+\author{Daniel Hollarek
+\thanks{Adie Alwen}
+\date{Nov 2023}                       
 \begin{titlepage}\maketitle
@@
 
-\subsection{A subsection with math}
-
-Here are math and an equation in the appendix---see equation \eqref{e:appendix1}.
-Temporibus autem quibusdam $\xi$ et aut officiis debitis aut rerum necessitatibus saepe eveniet ut et voluptates repudiandae sint et molestiae non recusandae $1-\gamma$.
-Itaque earum rerum hic $S(z^*)$ tenetur a sapiente delectus $\mathcal{B}^\theta$, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat $\mathbb{V}^i$:
-\begin{equation}
-\mathbb{V}^r = (1-\gamma) \times 0 +\gamma S(z^*) v^s+\gamma [1-S(z^*)] \mathcal{V}^i-c.
-\label{e:appendix1}\end{equation}
-
-Ut enim ad minima veniam, quis nostrum exercitationem ullam corporis suscipit laboriosam, nisi ut aliquid ex ea commodi consequatur?
-Quis autem vel eum iure reprehenderit qui in ea voluptate velit esse quam nihil molestiae consequatur, vel illum qui dolorem eum fugiat quo voluptas nulla pariatur?
-Autem vel eum iure reprehenderit qui in ea voluptate velit esse quam nihil molestiae consequatur.
-
-\section{Another appendix}\label{a:appendix2}
-
-Here is a second appendix. At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti.
-
-\subsection{Larger figure, without panel, in the appendix}
-
-Here is a large, simple figure in the appendix (see figure \ref{f:appendix1}).
-At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
-
-\begin{figure}[t]
-\includegraphics[scale=0.3,page=1]{\pdf}
-\caption{Caption for the larger graph}
-\note{Note for the larger graph. Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus.}
-\label{f:appendix1}\end{figure}
-
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
-tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
-quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
-consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
-cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
-proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
-
-\subsection{Even larger figure, without panel, in the appendix}
-
-Figure \ref{f:appendix2} shows an even larger figure in the appendix.
-At vero eos et accusamus et iusto odio dignissimos ducimus, qui blanditiis praesentium voluptatum deleniti atque corrupti, quos dolores et quas molestias excepturi sint, obcaecati cupiditate non provident, similique sunt in culpa, qui officia deserunt mollitia animi, id est laborum et dolorum fuga.
-Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus.
-
-\begin{figure}[p]
-\includegraphics[scale=0.4,page=3]{\pdf}
-\caption{Caption for the even larger graph}
-\note{Note for the larger graph. Nam libero tempore, cum soluta nobis est eligendi optio, cumque nihil impedit, quo minus id, quod maxime placeat, facere possimus.}
-\label{f:appendix2}\end{figure}
-
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
-tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
-quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
-consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
-cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
-proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
-
-\subsection{Subsection with references}\label{a:subappendix}
-
-Here is a sentence with some citations: \citet{MS19,MS21b} found something but that is an uncommon result \citep{M12,M14}.
-The references go to the reference list at the end of the main text---so appendix and main text share a reference list.
-
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
-tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
-quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
-consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
-cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
-proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
-
-\subsection{Subsection with a foonote}
-
-Here is a sentence with a footnote.\footnote{Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt.} The numbering of the footnotes, just like the numbering of the pages, continues from the main text to the appendix.
-
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
-quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
-consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
-cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
-proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
-
-\subsection{Subsection with an URL}
-
-It is possible to insert an URL: \url{https://github.com/pmichaillat/latex-paper}. Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
-tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
-quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
-consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
-cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
-proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
-
 \end{document}

### 2023-12-05 data_collection_tool/datacollectapp.py

@@ -1,0 +1,147 @@
+import os
+from datetime import datetime
+from typing import Optional
+from kivy.uix.slider import Slider
+from kivy.uix.widget import Widget
+from kivy.uix.boxlayout import BoxLayout
+from kivy.uix.popup import Popup
+from kivy.app import App
+from kivy.core.window import Window
+
+
+from data_collector.filesystem import zip_file_list, produce_csv_file
+from elements import PathCheckbox, InputDialog
+from elements import get_select_layout, get_scroll_view
+from elements import get_file_count_widget, get_ok_button, get_feedback_widget, get_scrollable_checkboxes_layout
+
+# -------------------------------------------
+
+class DataCollectApp(App):
+    def __init__(self, target_folder : Optional[str] = None,
+                 override_input_folder : Optional[str]  = None):
+        super(DataCollectApp, self).__init__()
+
+        # Properties
+        self.title : str = 'AiMAT XRD data collect'
+        self.target_folder : str = target_folder
+        self.input_folder_override : Optional[str]  = override_input_folder
+        self.default_font_size = Window.width*0.018
+
+        # PathCheckbox
+        self.root_checkbox : Optional[PathCheckbox] = None
+
+        # GUI Elements
+        self.slider : Optional[Widget] = None
+        self.feedback_widget : Optional[Widget] = None
+        self.scroll_view : Optional[Widget] = None
+        self.filecount_label : Optional[Widget] = None
+        self.input_popup : Optional[Popup] = InputDialog(callback=self.set_select_layout_content)
+
+
+    def build(self):
+        # Main layout
+        self.scroll_view = get_scroll_view()
+
+        self.filecount_label = get_file_count_widget(num_elements=0)
+        select_layout = get_select_layout(file_count_label=self.filecount_label,
+                                          scroll_view=self.scroll_view)
+        self.slider = Slider(orientation='vertical',min=0, max=1, value=1, size_hint=(0.1, 1))
+
+        main_layout = BoxLayout(orientation='horizontal', size_hint=(1, 0.9))
+        main_layout.add_widget(widget=select_layout)
+        main_layout.add_widget(widget=self.slider)
+
+        # Finish layout
+        ok_button = get_ok_button()
+        self.feedback_widget = get_feedback_widget(font_size=self.default_font_size)
+
+        finish_layout = BoxLayout(orientation='horizontal',size_hint =(1,0.1))
+        finish_layout.add_widget(ok_button)
[…]

### 2023-12-05 data_collection_tool/elements/__init__.py

@@ -1,0 +1,3 @@
+from .layouts import get_scrollable_checkboxes_layout,get_select_layout, get_scroll_view
+from .widgets import get_ok_button,get_feedback_widget,get_file_count_widget, InputDialog
+from .path_checkbox import PathCheckbox

### 2023-12-05 data_collection_tool/elements/layouts.py

@@ -1,0 +1,42 @@
+from kivy.uix.boxlayout import BoxLayout
+from kivy.uix.scrollview import ScrollView
+# from kivy.core.window import Window
+from kivy.uix.label import Label
+# from kivy.uix.slider import Slider
+from data_collector.elements.path_checkbox import PathCheckbox
+
+# -------------------------------------------
+
+def get_select_layout(file_count_label: Label, scroll_view: ScrollView):
+    select_layout = BoxLayout(orientation='vertical')
+    select_layout.add_widget(widget=file_count_label)
+    select_layout.add_widget(widget=scroll_view)
+
+    return select_layout
+
+
+def get_scroll_view():
+    scroll_view = ScrollView(size_hint=(1, 1), bar_width=0)
+    return scroll_view
+
+
+def get_scrollable_checkboxes_layout(root_checkbox : PathCheckbox):
+    scroll_layout = BoxLayout(orientation='vertical', size_hint_y=None)
+    scroll_layout.bind(minimum_height=scroll_layout.setter('height'))
+    recursively_add_boxes(gui_parent=scroll_layout, root_box=root_checkbox, indent=0)
+
+    return scroll_layout
+
+
+def recursively_add_boxes(gui_parent, root_box : PathCheckbox, indent : int):
+    if not root_box.get_is_file() and len(root_box.get_file_descendants()) == 0:
+        return
+
+    root_box.initialize_gui(gui_parent, level=indent)
+
+    for child_box in root_box.active_children:
+        recursively_add_boxes(gui_parent=root_box.child_container,
+                              root_box=child_box,
+                              indent=indent + 1)
+
+

### 2023-12-05 data_collection_tool/elements/path_checkbox.py

@@ -1,0 +1,152 @@
+from __future__ import annotations
+from kivy.uix.boxlayout import  BoxLayout
+
+import copy
+import os.path
+from PIL import Image
+from data_collector.elements.widgets import LabeledCheckBox, ImageToggleButton
+from data_collector.resources import get_foldericon_path, get_fileicon_path
+from data_collector.filesystem import Folder
+
+from typing import Optional
+# -------------------------------------------
+
+class PathCheckbox:
+    dim = 20
+    file_img = Image.open(fp=get_fileicon_path()).resize((dim, dim))
+    folder_img = Image.open(fp=get_foldericon_path()).resize((dim, dim))
+
+    def __init__(self,path : str):
+        self.path : str = path
+        self.active_children: list[PathCheckbox] = self.get_children_from_path()
+        self.potential_des : list[PathCheckbox] = []
+        self.relevant_des : list[PathCheckbox] =[]
+        self.widget : Optional[LabeledCheckBox] = None
+
+        self.child_container = None
+        self.total_container = None
+
+    def initialize_gui(self, master_frame, level : int):
+        self.total_container = BoxLayout(orientation='vertical', size_hint=(1, None))
+        self.total_container.bind(minimum_height=self.total_container.setter('height'))
+
+        # Item container
+        item_container = self.make_toggleitem_container(level=level)
+        self.total_container.add_widget(item_container)
+
+
+        # Child container
+        if not self.get_is_file():
+            self.child_container = self.get_child_container()
+            self.total_container.add_widget(self.child_container)
+
+        # Master
+        master_frame.add_widget(self.total_container)
+
+
+    def on_toggle(self, instance, value):
+        if value == 'down':
+            self.total_container.remove_widget(self.child_container)
+        else:
+            self.total_container.add_widget(self.child_container)
+
+
+    def recursively_initialize_descendants(self):
+        descendants = copy.copy(self.active_children)
+        for child in self.active_children:
+            child.recursively_initialize_descendants()
+            descendants += child.potential_des
+
+
+        self.potential_des = descendants
+
+        relevant_des = []
+        for outer_des in self.potential_des:
+            is_relevant = any([des.get_is_file() for des in outer_des.potential_des]) or outer_des.get_is_file()
+            relevant_des += [outer_des] if is_relevant else []
+
+        self.relevant_des = relevant_des
+
+    # -------------------------------------------
+    # State
+
+    def toggle(self, *args)-> None:
+        _ = args
[…]

### 2023-12-05 data_collection_tool/elements/widgets.py

@@ -1,0 +1,180 @@
+from __future__ import annotations
+import os
+
+from PIL import Image
+from kivy.clock import Clock
+from kivy.uix.boxlayout import BoxLayout
+from kivy.uix.checkbox import CheckBox
+from kivy.uix.image import Image
+from kivy.uix.relativelayout import RelativeLayout
+from kivy.uix.togglebutton import ToggleButton
+from kivy.uix.popup import Popup
+from kivy.uix.label import Label
+from kivy.uix.textinput import TextInput
+from kivy.uix.button import Button
+from kivy.core.window import Window
+from data_collector.filesystem import Folder
+from kivy.uix.scrollview import ScrollView
+
+from data_collector.resources import get_foldericon_path,get_fileicon_path
+# -------------------------------------------
+from data_collector.resources import get_collapsed_icon_path, get_expanded_icon_path
+
+class LabeledCheckBox(BoxLayout):
+    indent = 50
+
+    def __init__(self, text, toggle_callback, is_file : bool, indent=0, **kwargs):
+        super().__init__(**kwargs)
+        # Properties
+        self.size_hint_y = None
+        self.height = 30  # Fixed height for each LabeledCheckBox
+
+        # Create the CheckBox
+        self.check_box = CheckBox(size_hint=(None, None), size=(30, 30))
+        self.check_box.bind(active=toggle_callback)
+
+        # Create the Icon
+        icon_path = get_fileicon_path() if is_file else get_foldericon_path()
+        self.icon = Image(source=icon_path, size_hint=(None, None), size=(30, 30))
+        self.add_widget(self.icon)
+
+        # Create the Label
+        self.label = BlackLabel(text=text,
+                                size_hint_x=None,
+                                size_hint_y=None,
+                                height=30,
+                                halign='left',
+                                valign='middle')
+
+        self.add_widget(self.check_box)
+        self.add_widget(self.label)
+        self.padding = [indent * LabeledCheckBox.indent, 0, 0, 0]  # Left padding for indentation
+
+
+class InputDialog(Popup):
+    def __init__(self, callback  : callable, **kwargs):
+        super(InputDialog, self).__init__(**kwargs)
+        self.size_hint = (0.8, 0.6)
+        self.title = 'XRD data collector'
+        self.callback = callback
+
+        self.scroll_view = ScrollView(size_hint=(1, 0.8))
+        self.hint = AutoSizeLabel(text='This application is designed to scan a given input folder for all files that fit any of the following common XRD formats: '
[…]

### 2023-12-05 data_collection_tool/filesystem/__init__.py

@@ -1,0 +1,2 @@
+from .folder import Folder
+from .out_files import zip_file_list, produce_csv_file

### 2023-12-05 data_collection_tool/filesystem/folder.py

@@ -1,0 +1,45 @@
+from __future__ import annotations
+import os
+# from typing import Optional
+
+# -------------------------------------------
+
+
+class Folder:
+    xrd_formats : list = ['raw','dif','gdf','dat','ras','cpi','txt','plv','xrdml']
+
+    def __init__(self, path : str):
+        self.path : str = path
+        self.name : str = os.path.basename(path)
+        self.sub_paths : list[str] = [os.path.join(self.path,name) for name in os.listdir(self.path)]
+
+    def get_descendant_xrdfilepaths(self) -> list[str]:
+        xrd_file_paths = []
+
+        xrd_file_paths += self._get_children_xrdfilepaths()
+        for folder_path in self._get_children_folderpaths():
+            xrd_folder = Folder(path=folder_path)
+            xrd_file_paths += xrd_folder.get_descendant_xrdfilepaths()
+
+        return xrd_file_paths
+
+    def get_children_xrdpaths(self):
+        return self._get_children_xrdfilepaths() + self._get_children_folderpaths()
+
+    def _get_children_xrdfilepaths(self) -> list[str]:
+        return [path for path in self._get_children_filepaths() if self._get_path_is_xrd_file(path=path)]
+
+
+    def _get_path_is_xrd_file(self, path : str):
+        is_file = os.path.isfile(path)
+        is_xrd_format = any([path.endswith(f'.{the_format}') for the_format in self.xrd_formats])
+        return is_file and is_xrd_format
+
+
+    # -------------------------------------------
+
+    def _get_children_filepaths(self) -> list[str]:
+        return [path for path in self.sub_paths if os.path.isfile(path)]
+
+    def _get_children_folderpaths(self) -> list[str]:
+        return [path for path in self.sub_paths if os.path.isdir(path)]

### 2023-12-05 data_collection_tool/filesystem/out_files.py

@@ -1,0 +1,77 @@
+# import logging
+from collections import defaultdict
+
+import os
+import zipfile, csv
+
+def log(to_log : str):
+    print(to_log)
+
+
+def get_zip_path_mapping(path_list: list[str]) -> dict:
+    name_counter = defaultdict(int)
+    zip_path_dict = {}
+
+    for path in path_list:
+        try:
+            file_name = os.path.basename(path)
+            foldername = os.path.basename(os.path.dirname(path))
+            base_path = os.path.join(foldername,file_name)
+
+            if not base_path in name_counter:
+                name_counter[base_path] = 1
+            else:
+                name_counter[base_path] += 1
+
+            name = f"{foldername}_{name_counter[base_path]}"
+            zip_path = os.path.join(name, file_name)
+
+        except Exception as e:
+            zip_path = path
+            print(f"Error processing path '{path}': {e}")
+
+        zip_path_dict[path] = zip_path
+
+    return zip_path_dict
+
+
+
+def zip_file_list(path_list: list[str], zipfile_path : str):
+    zip_path_dict = get_zip_path_mapping(path_list=path_list)
+
+    if os.path.isfile(zipfile_path):
+        log(to_log=f'Target path {zipfile_path} already exists. Aborting ...')
+        return
+
+    with zipfile.ZipFile(zipfile_path, 'w') as zipf:
+        for file_path in path_list:
+            zip_path = zip_path_dict[file_path]
+            zipf.write(file_path, zip_path)
+
+    print(f"ZIP archive created at {zipfile_path}")
+
+
+
+def produce_csv_file(path_list : list[str], target_path : str):
+    zip_path_dict = get_zip_path_mapping(path_list=path_list)
+
+    if os.path.isfile(target_path):
+        print(f'Target CSV path {target_path} already exists. Aborting ...')
+        return
+
+    with open(target_path, mode='w', newline='') as file:
+        writer = csv.writer(file)
+        writer.writerow(['Filepath relative to zip root'])
+        for file_path in path_list:
+            zip_file_path = zip_path_dict[file_path]
+            writer.writerow([zip_file_path])
+
+    print(f"CSV file created at {target_path}")
+
+
+if __name__ == "__main__":
+    source_path = ["/home/work/Desktop/test2/asdf2.png"]
+    the_target_path = "/home/work/Desktop/target"
+
+    zip_file_list(source_path, the_target_path)
+

### 2023-12-05 data_collection_tool/resources/__init__.py

@@ -1,0 +1,1 @@
+from .resource_manager import get_foldericon_path, get_fileicon_path, get_collapsed_icon_path, get_expanded_icon_path

### 2023-12-05 data_collection_tool/resources/resource_manager.py

@@ -1,0 +1,15 @@
+import os
+file_path = __file__
+dir_path = os.path.dirname(file_path)
+
+def get_foldericon_path() -> str:
+    return os.path.join(dir_path,'folder.png')
+
+def get_fileicon_path() -> str:
+    return os.path.join(dir_path,'file.png')
+
+def get_collapsed_icon_path() -> str:
+    return os.path.join(dir_path,'collapsed.png')
+
+def get_expanded_icon_path() -> str:
+    return os.path.join(dir_path,'expanded.png')

### 2023-12-05 data_collection_tool/run.py

@@ -1,0 +1,28 @@
+from kivy.config import Config
+Config.set('graphics', 'fullscreen', 0)  # Can use 'auto' or '1'
+Config.set('graphics', 'resizable', '0')
+
+import os
+from kivy.core.window import Window
+from data_collector.datacollectapp import DataCollectApp
+
+# -------------------------------------------
+
+def main():
+    # win_test_dir = 'C:\\Users\\daniel\\OneDrive\\Desktop\\test_folder'
+    # win_target_test = 'C:\\Users\\daniel\\OneDrive\\Desktop'
+
+    # _, __ = win_test_dir, win_target_test
+
+    linux_input_test = '/home/work/Desktop'
+    # linux_target_test = '/home/work/Desktop'
+
+    Window.clearcolor = (1, 1, 1, 1)  # RGBA for white
+    app = DataCollectApp(target_folder=os.path.dirname(os.path.abspath(__file__)),
+                         override_input_folder= linux_input_test)
+
+
+    app.run()
+
+if __name__ == "__main__":
+    main()

### 2023-12-05 data_collector/filesystem/out_files.py
data colletor: Renamed data tool

@@ -70,7 +70,7 @@ def produce_csv_file(path_list : list[str], target_path : str):
 
 
 if __name__ == "__main__":
-    source_path = ["/home/work/Desktop/test2/asdf2.png"]
+    source_path = ["/home/work/Desktop/test1/asdf2.png"]
     the_target_path = "/home/work/Desktop/target"
 
     zip_file_list(source_path, the_target_path)

### 2023-12-05 run.py
Added setup.py

@@ -4,7 +4,7 @@ Config.set('graphics', 'resizable', '0')
 
 import os
 from kivy.core.window import Window
-from data_collector.datacollectapp import DataCollectApp
+from datacollectapp import DataCollectApp
 
 # -------------------------------------------
 
@@ -24,5 +24,4 @@ def main():
 
     app.run()
 
-if __name__ == "__main__":
-    main()
+main()

### 2023-12-05 setup.py

@@ -1,0 +1,21 @@
+from cx_Freeze import setup, Executable
+import sys
+
+# GUI applications require a different base on Windows
+base = None
+if sys.platform == "win32":
+    base = "Win32GUI"
+
+build_exe_options = {
+    "packages": ["kivy", "os","PIL"],
+    "excludes": [],
+    "include_files": []
+}
+
+setup(
+    name="SimpleKivyApp",
+    version="0.1",
+    description="My Kivy GUI application!",
+    options={"build_exe": build_exe_options},
+    executables=[Executable("run.py", base=base)]
+)

### 2023-12-05 data_collector/datacollectapp.py
data_collector: Fixed imports

@@ -10,9 +10,9 @@ from kivy.core.window import Window
 
 
 from data_collector.filesystem import zip_file_list, produce_csv_file
-from elements import PathCheckbox, InputDialog
-from elements import get_select_layout, get_scroll_view
-from elements import get_file_count_widget, get_ok_button, get_feedback_widget, get_scrollable_checkboxes_layout
+from data_collector.elements import PathCheckbox, InputDialog
+from data_collector.elements import get_select_layout, get_scroll_view
+from data_collector.elements import get_file_count_widget, get_ok_button, get_feedback_widget, get_scrollable_checkboxes_layout
 
 # -------------------------------------------
 

### 2023-12-05 run.py

@@ -4,7 +4,7 @@ Config.set('graphics', 'resizable', '0')
 
 import os
 from kivy.core.window import Window
-from datacollectapp import DataCollectApp
+from data_collector.datacollectapp import DataCollectApp
 
 # -------------------------------------------
 

### 2024-01-22 data_collector/resources/resource_manager.py
data_collector: Implemented custom version of checkbox for DPI aware images

@@ -48,6 +48,8 @@ def get_unchecked_box_path() -> str:
 def get_template_csv() -> str:
     return os.path.join(doc_folder_path, 'template.csv')
 
+def get_empty_path() -> str:
+    return os.path.join(images_folder_path, 'empty.png')
 
 def pil_to_kivy_image(pil_image, **kwargs):
     data = BytesIO()

### 2024-01-22 data_collector/elements/node_widget.py
data_collector: Removed original kivy images added onto custom checkmark

@@ -105,7 +105,8 @@ class ImageCheckBox(CheckBox):
         self.checked_image = CoreImage(get_checked_box_path()).texture
         self.unchecked_image = CoreImage(get_unchecked_box_path()).texture
 
-        self._checkbox_image = get_empty_path()
+        self.background_checkbox_down = get_empty_path()
+        self.background_checkbox_normal = get_empty_path()
 
         with self.canvas:
             self.rect = Rectangle(texture=self.unchecked_image, pos=self.pos, size=self.size)
@@ -118,8 +119,8 @@ class ImageCheckBox(CheckBox):
             return True
         return super().on_touch_down(touch)
 
-    def _on_state(self, instance, value):
-        pass
+    # def _on_state(self, instance, value):
+    #     self._checkbox_image = get_empty_path()
 
     def update_rect(self, *args):
         self.rect.pos = self.pos

### 2024-01-22 data_collector/layouts/dialogue.py
data_collector: Adjsuted spacing in input popup to be DPI aware

@@ -30,6 +30,9 @@ class InputDialog(Popup):
 
         l = 12*get_app_width()/1600.
         self.content = BoxLayout(orientation='vertical', padding=[l, l, l, l], spacing=l)
+
+        print(f'App width is: {get_app_width()}')
+
         self.background_color = (100 / 255, 255 / 255, 255 / 255, 1)  # (R, G, B, A)
 
         # logo_image = Image(source=get_blended_logo_path(), size_hint=(1, 0.3))

### 2024-01-22 data_collector/layouts/finish.py

@@ -18,7 +18,8 @@ class FinishLayout(BoxLayout):
         l = 15 * get_app_width() / 1600.
         super(FinishLayout, self).__init__(orientation='vertical', size_hint=(1, 0.085),padding=(0, l, 0, l),**kwargs)
 
-        upper_finish = BoxLayout(orientation='horizontal', size_hint=(1, 0.3),spacing=10)
+        s = 10 * get_app_width() / 1600.
+        upper_finish = BoxLayout(orientation='horizontal', size_hint=(1, 0.3),spacing=s)
         note = BlackLabel(text='Target folder:', size_hint=(0.2, 1), font_size=Window.width * 0.02, bold=True)
         self.target_path_input = TextInput(text=f'{os.getcwd()}',
                                            size_hint=(0.7, 1),

### 2024-01-22 data_collector/configs/__init__.py
data_collector: Updated app dimension functions for scaling factors

@@ -1,2 +1,3 @@
 from .conf import set_configs, get_line_height
-from .conf import get_app_height, get_app_width
+from .conf import get_scaled_down_app_height, get_scaled_down_app_width
+from .conf import get_true_height, get_true_width

### 2024-01-22 data_collector/configs/conf.py

@@ -33,7 +33,7 @@ def set_configs():
     Config.set('graphics', 'fullscreen', 0)
     # Config.set('graphics', 'resizable', '1')
 
-    app_width, app_height = get_app_width(), get_app_height()
+    app_width, app_height = get_scaled_down_app_width(), get_scaled_down_app_height()
 
     print(f'[INFO]: Unscaled width, height: {app_width},{app_height}')
 
@@ -41,21 +41,31 @@ def set_configs():
     Config.set('graphics', 'height', str(app_height))
 
 
-def get_app_width() -> int:
+def get_scaled_down_app_width() -> int:
+    width = int(get_true_width() / get_scaling_factor())
+    return width
+
+
+def get_scaled_down_app_height() -> int:
+    height = int(get_true_height() / get_scaling_factor())
+    return height
+
+
+def get_true_width() -> int:
     primary_monitor = get_primary_monitor()
-    scaling_factor = get_scaling_factor()
     relative_width = 0.75
 
-    app_width = int(primary_monitor.width * relative_width / scaling_factor)
+    app_width = int(primary_monitor.width * relative_width)
     return app_width
 
 
-def get_app_height() -> int:
+def get_true_height() -> int:
     primary_monitor = get_primary_monitor()
-    scaling_factor = get_scaling_factor()
     relative_height = 0.8
-    app_height = int(primary_monitor.height * relative_height / scaling_factor)
+
+    app_height = int(primary_monitor.height * relative_height)
     return app_height
+
 
 def get_line_height():
     std_width = 1920

### 2024-01-22 data_collector/elements/types.py

@@ -9,7 +9,7 @@ from kivy.uix.textinput import TextInput
 from kivy.core.window import Window
 
 from kivy.uix.widget import Widget
-from data_collector.configs import get_app_width
+from data_collector.configs import get_scaled_down_app_width
 from data_collector.resources import get_logo_path, get_kivy_image
 from typing import List
 
@@ -89,7 +89,7 @@ class HeaderWidget(BoxLayout):
         file_count_label.bind(size=file_count_label.setter('text_size'))
 
         # logo_image = Image(source=get_logo_path(), size_hint=(0.3, 1))
-        img_width = get_app_width()*0.3
+        img_width = get_scaled_down_app_width() * 0.3
         logo_image = get_kivy_image(width=img_width, imgPath=get_logo_path(), size_hint=(0.3,1))
         print(f'Width of the header widget:{img_width}')
 

### 2024-01-22 data_collector/layouts/dialogue.py

@@ -10,7 +10,7 @@ from kivy.uix.button import Button
 from kivy.uix.label import Label
 from kivy.uix.popup import Popup
 
-from data_collector.configs import get_app_width
+from data_collector.configs import get_scaled_down_app_width, get_true_width
 from data_collector.elements.types import FocusTextInput
 from data_collector.filesystem import FsNode, get_initial_path
 from data_collector.resources.resource_manager import get_blended_logo_path, get_kivy_image
@@ -28,10 +28,10 @@ class InputDialog(Popup):
         self.auto_dismiss = False
         self.callback = callback
 
-        l = 12*get_app_width()/1600.
+        l = 12 * get_true_width () / 1600.
         self.content = BoxLayout(orientation='vertical', padding=[l, l, l, l], spacing=l)
 
-        print(f'App width is: {get_app_width()}')
+        print(f'App width is: {get_scaled_down_app_width()}')
 
         self.background_color = (100 / 255, 255 / 255, 255 / 255, 1)  # (R, G, B, A)
 

### 2024-01-22 data_collector/layouts/finish.py

@@ -10,15 +10,16 @@ from kivy.uix.widget import Widget
 from data_collector.elements.types import BlackLabel
 from kivy.uix.popup import Popup
 from kivy.uix.label import Label
-from data_collector.configs import get_app_width
+from data_collector.configs import get_scaled_down_app_width, get_true_width
 # -------------------------------------------
 
 class FinishLayout(BoxLayout):
     def __init__(self, callback: callable, exit_funct : callable, **kwargs):
-        l = 15 * get_app_width() / 1600.
+
+        l = 15 * get_true_width() / 1600.
         super(FinishLayout, self).__init__(orientation='vertical', size_hint=(1, 0.085),padding=(0, l, 0, l),**kwargs)
 
-        s = 10 * get_app_width() / 1600.
+        s = 10 * get_true_width() / 1600.
         upper_finish = BoxLayout(orientation='horizontal', size_hint=(1, 0.3),spacing=s)
         note = BlackLabel(text='Target folder:', size_hint=(0.2, 1), font_size=Window.width * 0.02, bold=True)
         self.target_path_input = TextInput(text=f'{os.getcwd()}',

### 2024-01-22 data_collector/app.py
data_collector: Implemented out file production in different thread to prevent GUI freeze

@@ -1,9 +1,12 @@
 import os
+import threading
 from datetime import datetime
 from typing import Optional
 from kivy.uix.boxlayout import BoxLayout
 from kivy.app import App
 from typing import List
+
+from kivy.clock import Clock
 
 from data_collector.filesystem import zip_file_list, produce_csv_file
 from data_collector.layouts import SelectionLayout, FinishLayout, InputDialog
@@ -69,8 +72,8 @@ class DataCollectApp(App):
         zipfile_path = os.path.join(target_folder_path, f'xrd_data_collected_on_{datetime_stamp}.zip')
         csv_file_path = os.path.join(target_folder_path,f'xrd_labels_generated_on_{datetime_stamp}.csv')
 
-        print("Checked Paths:", checked_paths)
-        try:
+
+        def produce_files():
             zip_file_list(path_list=checked_paths,
                           zipfile_path=zipfile_path,
                           root_path=self.get_rootfolder_path())
@@ -78,14 +81,20 @@ class DataCollectApp(App):
             produce_csv_file(abs_path_list=checked_paths,
                              target_path=csv_file_path,
                              root_path=self.get_rootfolder_path())
+            self.finish_layout.o
 
+            Clock.schedule_once(self.show_feedback)
+
+        print("Checked Paths:", checked_paths)
+        try:
+            produce_files_thread = threading.Thread(target=produce_files)
+            produce_files_thread.start()
 
             self.finish_layout.feedback_widget.text = (f'Done! Wrote {len(checked_paths)} xrd file(s) to .zip file and produced corresponding label .csv file at:\n'
                                          f'{zipfile_path} \n'
                                          f'{csv_file_path}')
         except:
             self.finish_layout.feedback_widget.text = f'An error occured during the creating of the zip archive.\nIs "{target_folder_path}" a valid folder path?'
-        self.show_feedback()
 
 
     def get_checked_filepaths(self) -> List[str]:
@@ -100,7 +109,8 @@ class DataCollectApp(App):
         return self.finish_layout.target_path_input.text if self.targt_folder_override is None else self.targt_folder_override
 
 
-    def show_feedback(self):
+    def show_feedback(self, *args, **kwargs):
+        _, __ = args, kwargs
         self.finish_layout.show()
         # self.finish_layout.feedback_widget.opacity = 1
         # print(self.finish_layout.feedback_widget.text)

### 2024-01-22 data_collector/app.py
data_collector: Included is working message on bundle files button press

@@ -81,7 +81,7 @@ class DataCollectApp(App):
             produce_csv_file(abs_path_list=checked_paths,
                              target_path=csv_file_path,
                              root_path=self.get_rootfolder_path())
-            self.finish_layout.o
+            self.finish_layout.ok_button.text = 'working ...'
 
             Clock.schedule_once(self.show_feedback)
 
@@ -112,6 +112,7 @@ class DataCollectApp(App):
     def show_feedback(self, *args, **kwargs):
         _, __ = args, kwargs
         self.finish_layout.show()
+        self.finish_layout.ok_button.text = 'bundle files'
         # self.finish_layout.feedback_widget.opacity = 1
         # print(self.finish_layout.feedback_widget.text)
 

### 2024-01-22 data_collector/layouts/finish.py

@@ -27,12 +27,12 @@ class FinishLayout(BoxLayout):
                                            font_size=Window.width * 0.02,
                                            multiline=False)
         buffer = BlackLabel(text='', size_hint=(0.075, 1))
-        ok_button = self.get_ok_button()
-        ok_button.bind(on_press=callback)
+        self.ok_button = self.get_ok_button()
+        self.ok_button.bind(on_press=callback)
 
         upper_finish.add_widget(note)
         upper_finish.add_widget(self.target_path_input)
-        upper_finish.add_widget(ok_button)
+        upper_finish.add_widget(self.ok_button)
         upper_finish.add_widget(buffer)
 
         self.exit_funct = exit_funct

### 2024-01-24 data_collector/app.py
data_collector: Fixed bug to display working... on button

@@ -74,6 +74,8 @@ class DataCollectApp(App):
 
 
         def produce_files():
+            self.finish_layout.ok_button.text = 'working ...'
+
             zip_file_list(path_list=checked_paths,
                           zipfile_path=zipfile_path,
                           root_path=self.get_rootfolder_path())
@@ -81,7 +83,6 @@ class DataCollectApp(App):
             produce_csv_file(abs_path_list=checked_paths,
                              target_path=csv_file_path,
                              root_path=self.get_rootfolder_path())
-            self.finish_layout.ok_button.text = 'working ...'
 
             Clock.schedule_once(self.show_feedback)
 

### 2024-01-24 data_collector/layouts/dialogue.py
data_collector: Decreased size of logo in popup

@@ -36,7 +36,7 @@ class InputDialog(Popup):
         self.background_color = (100 / 255, 255 / 255, 255 / 255, 1)  # (R, G, B, A)
 
         # logo_image = Image(source=get_blended_logo_path(), size_hint=(1, 0.3))
-        logo_image = get_kivy_image(width=Window.width*0.4, imgPath=get_blended_logo_path(), size_hint =(1,0.25))
+        logo_image = get_kivy_image(width=Window.width*0.4, imgPath=get_blended_logo_path(), size_hint =(1,0.2))
         print(f'Width of the blended logo: {Window.width*0.4}')
         first_hint = self.make_hint()
         self.format_input =self.make_format_input()

### 2024-01-24 data_collector/layouts/dialogue.py
data_collector: Made font size relative to true app width

@@ -18,7 +18,7 @@ from data_collector.resources.resource_manager import get_blended_logo_path, get
 
 class InputDialog(Popup):
 
-    font_size = Window.width * 0.0145
+    font_size = get_true_width() * 0.0145
 
     def __init__(self, callback: callable, **kwargs):
         super(InputDialog, self).__init__(**kwargs)
@@ -36,7 +36,7 @@ class InputDialog(Popup):
         self.background_color = (100 / 255, 255 / 255, 255 / 255, 1)  # (R, G, B, A)
 
         # logo_image = Image(source=get_blended_logo_path(), size_hint=(1, 0.3))
-        logo_image = get_kivy_image(width=Window.width*0.4, imgPath=get_blended_logo_path(), size_hint =(1,0.2))
+        logo_image = get_kivy_image(width=Window.width*0.4, imgPath=get_blended_logo_path(), size_hint =(1,0.25))
         print(f'Width of the blended logo: {Window.width*0.4}')
         first_hint = self.make_hint()
         self.format_input =self.make_format_input()

### 2024-01-24 data_collector/layouts/dialogue.py
data_collector: Tied vertical padding to height rather than width

@@ -10,7 +10,7 @@ from kivy.uix.button import Button
 from kivy.uix.label import Label
 from kivy.uix.popup import Popup
 
-from data_collector.configs import get_scaled_down_app_width, get_true_width
+from data_collector.configs import get_scaled_down_app_width, get_true_width, get_true_height
 from data_collector.elements.types import FocusTextInput
 from data_collector.filesystem import FsNode, get_initial_path
 from data_collector.resources.resource_manager import get_blended_logo_path, get_kivy_image
@@ -28,8 +28,11 @@ class InputDialog(Popup):
         self.auto_dismiss = False
         self.callback = callback
 
-        l = 12 * get_true_width () / 1600.
-        self.content = BoxLayout(orientation='vertical', padding=[l, l, l, l], spacing=l)
+        horizontal_spacing = 12 * get_true_width() / 1600.
+        vertical_spacing = 12 * get_true_height() / 900.  # Assuming 900 is a relevant measure for height
+        self.content = BoxLayout(orientation='vertical',
+                                 padding=[horizontal_spacing, vertical_spacing, horizontal_spacing, vertical_spacing],
+                                 spacing=vertical_spacing)
 
         print(f'App width is: {get_scaled_down_app_width()}')
 

### 2024-01-24 data_collector/configs/conf.py
data_collector: Tied line height to monitor height rather than width

@@ -7,16 +7,6 @@ import logging
 logging.basicConfig(level=logging.INFO)
 # -------------------------------------------
 
-def get_primary_monitor():
-    monitors = get_monitors()
-    for monitor in monitors:
-        if monitor.is_primary:
-            return monitor
-    return None
-
-
-def get_primary_monitor_width():
-    return get_primary_monitor().width
 
 def get_scaling_factor():
     ui_scale = 1
@@ -68,12 +58,24 @@ def get_true_height() -> int:
 
 
 def get_line_height():
-    std_width = 1920
+    std_width = 1080
     std_line_height = 40
-
-    line_height =int(std_line_height * get_primary_monitor_width()/std_width)
-
+    line_height =int(std_line_height * get_primary_monitor_height()/std_width)
     print(line_height)
 
     return line_height
 
+
+def get_primary_monitor():
+    monitors = get_monitors()
+    for monitor in monitors:
+        if monitor.is_primary:
+            return monitor
+    return None
+
+
+def get_primary_monitor_width():
+    return get_primary_monitor().width
+
+def get_primary_monitor_height():
+    return get_primary_monitor().height

### 2024-01-24 data_collector/app.py
data_collector: Disabled overscrolling

@@ -114,6 +114,4 @@ class DataCollectApp(App):
         _, __ = args, kwargs
         self.finish_layout.show()
         self.finish_layout.ok_button.text = 'bundle files'
-        # self.finish_layout.feedback_widget.opacity = 1
-        # print(self.finish_layout.feedback_widget.text)
 

### 2024-01-24 data_collector/layouts/selection.py

@@ -1,5 +1,6 @@
 from __future__ import annotations
 
+import copy
 from typing import Optional
 from kivy.uix.label import Label
 from kivy.uix.scrollview import ScrollView
@@ -103,14 +104,23 @@ class SelectionLayout(BoxLayout):
 
     def on_scroll_view_scroll(self, instance, value):
         _ = instance
+
         self.slider.unbind(value=self.adjust_scroll_view)
-        self.slider.value = value
+
+        if self.scroll_view.scroll_y < 0:
+            self.scroll_view.scroll_y = 0
+
+        if self.scroll_view.scroll_y > 1:
+            self.scroll_view.scroll_y = 1
+
+        self.slider.value = copy.copy(self.scroll_view.scroll_y)
         self.slider.bind(value=self.adjust_scroll_view)
 
 
     def adjust_scroll_view(self, instance, value):
         _ = instance
         self.scroll_view.scroll_y = value
+
 
     def check_for_dismiss(self, *args, **kwargs):
         _, __ = args, kwargs

### 2024-01-24 data_collector/app.py
data_collector: Updated error handling of out file production

@@ -59,21 +59,25 @@ class DataCollectApp(App):
         _ = args
         print('Triggered callback of ok button')
         checked_paths = self.get_checked_filepaths()
-
-        current_date = datetime.now()
-        datetime_stamp = current_date.strftime('%d_%m_%Y_%H_%M_%S')
-
         target_folder_path = self.get_targetfolder_path()
 
         if target_folder_path is None:
             print(f'No target folder provided. Check setup')
             return
 
+        print("Checked Paths:", checked_paths)
+        produce_files_thread = threading.Thread(target=self.produce_files, args=(target_folder_path,checked_paths,))
+        produce_files_thread.start()
+
+
+
+    def produce_files(self, target_folder_path: str, checked_paths : List[str]):
+        current_date = datetime.now()
+        datetime_stamp = current_date.strftime('%d_%m_%Y_%H_%M_%S')
         zipfile_path = os.path.join(target_folder_path, f'xrd_data_collected_on_{datetime_stamp}.zip')
-        csv_file_path = os.path.join(target_folder_path,f'xrd_labels_generated_on_{datetime_stamp}.csv')
+        csv_file_path = os.path.join(target_folder_path, f'xrd_labels_generated_on_{datetime_stamp}.csv')
 
-
-        def produce_files():
+        try:
             self.finish_layout.ok_button.text = 'working ...'
 
             zip_file_list(path_list=checked_paths,
@@ -84,19 +88,18 @@ class DataCollectApp(App):
                              target_path=csv_file_path,
                              root_path=self.get_rootfolder_path())
 
-            Clock.schedule_once(self.show_feedback)
+            self.finish_layout.feedback_widget.text = (
+                f'Done! Wrote {len(checked_paths)} xrd file(s) to .zip file and produced corresponding label .csv file at:\n'
+                f'{zipfile_path} \n'
+                f'{csv_file_path}')
 
-        print("Checked Paths:", checked_paths)
-        try:
-            produce_files_thread = threading.Thread(target=produce_files)
-            produce_files_thread.start()
 
-            self.finish_layout.feedback_widget.text = (f'Done! Wrote {len(checked_paths)} xrd file(s) to .zip file and produced corresponding label .csv file at:\n'
-                                         f'{zipfile_path} \n'
-                                         f'{csv_file_path}')
-        except:
-            self.finish_layout.feedback_widget.text = f'An error occured during the creating of the zip archive.\nIs "{target_folder_path}" a valid folder path?'
+        except Exception as e:
[…]

### 2024-01-24 data_collector/app.py
data_collector: Tied vertical padding of finish layout to height instead of width

@@ -88,6 +88,7 @@ class DataCollectApp(App):
                              target_path=csv_file_path,
                              root_path=self.get_rootfolder_path())
 
+            self.finish_layout.feedback_popup.title = 'Success!'
             self.finish_layout.feedback_widget.text = (
                 f'Done! Wrote {len(checked_paths)} xrd file(s) to .zip file and produced corresponding label .csv file at:\n'
                 f'{zipfile_path} \n'
@@ -95,6 +96,7 @@ class DataCollectApp(App):
 
 
         except Exception as e:
+            self.finish_layout.feedback_popup.title = 'Failed to bundle files!'
             self.finish_layout.feedback_widget.text = (f'An error occured during the creating of the zip archive: {e}\n'
                                                        f'Is "{target_folder_path}" a valid folder path?'
                                                        f'Does the executable have permission to view all selected files?')

### 2024-01-24 data_collector/layouts/finish.py

@@ -10,13 +10,13 @@ from kivy.uix.widget import Widget
 from data_collector.elements.types import BlackLabel
 from kivy.uix.popup import Popup
 from kivy.uix.label import Label
-from data_collector.configs import get_scaled_down_app_width, get_true_width
+from data_collector.configs import get_true_height, get_true_width
 # -------------------------------------------
 
 class FinishLayout(BoxLayout):
     def __init__(self, callback: callable, exit_funct : callable, **kwargs):
 
-        l = 15 * get_true_width() / 1600.
+        l = 15 * get_true_height() / 900.
         super(FinishLayout, self).__init__(orientation='vertical', size_hint=(1, 0.085),padding=(0, l, 0, l),**kwargs)
 
         s = 10 * get_true_width() / 1600.
@@ -24,7 +24,7 @@ class FinishLayout(BoxLayout):
         note = BlackLabel(text='Target folder:', size_hint=(0.2, 1), font_size=Window.width * 0.02, bold=True)
         self.target_path_input = TextInput(text=f'{os.getcwd()}',
                                            size_hint=(0.7, 1),
-                                           font_size=Window.width * 0.02,
+                                           font_size=get_true_width() * 0.02,
                                            multiline=False)
         buffer = BlackLabel(text='', size_hint=(0.075, 1))
         self.ok_button = self.get_ok_button()
@@ -36,7 +36,7 @@ class FinishLayout(BoxLayout):
         upper_finish.add_widget(buffer)
 
         self.exit_funct = exit_funct
-        self.default_font_size = Window.width * 0.018
+        self.default_font_size = get_true_width() * 0.018
         self.feedback_widget = self.get_feedback_widget(font_size=self.default_font_size)
         self.feedback_popup = self.get_feedback_popup()
 

### 2024-01-24 data_collector/app.py
data_collector: Adjusted feedback popup size

@@ -89,7 +89,7 @@ class DataCollectApp(App):
                              root_path=self.get_rootfolder_path())
 
             self.finish_layout.feedback_popup.title = 'Success!'
-            self.finish_layout.feedback_widget.text = (
+            self.finish_layout.feedback_label.text = (
                 f'Done! Wrote {len(checked_paths)} xrd file(s) to .zip file and produced corresponding label .csv file at:\n'
                 f'{zipfile_path} \n'
                 f'{csv_file_path}')
@@ -97,9 +97,9 @@ class DataCollectApp(App):
 
         except Exception as e:
             self.finish_layout.feedback_popup.title = 'Failed to bundle files!'
-            self.finish_layout.feedback_widget.text = (f'An error occured during the creating of the zip archive: {e}\n'
-                                                       f'Is "{target_folder_path}" a valid folder path?'
-                                                       f'Does the executable have permission to view all selected files?')
+            self.finish_layout.feedback_label.text = (f'An error occured during the creating of the zip archive:\n {e}'
+                                                       f'\nIs "{target_folder_path}" a valid folder path?'
+                                                       f'\nDoes the executable have permission to view all selected files?')
 
         Clock.schedule_once(self.show_feedback)
 

### 2024-01-24 data_collector/layouts/finish.py

@@ -37,21 +37,20 @@ class FinishLayout(BoxLayout):
 
         self.exit_funct = exit_funct
         self.default_font_size = get_true_width() * 0.018
-        self.feedback_widget = self.get_feedback_widget(font_size=self.default_font_size)
+        self.feedback_label = self.get_feedback_label(font_size=self.default_font_size)
         self.feedback_popup = self.get_feedback_popup()
 
         self.add_widget(upper_finish)
 
 
     def show(self):
-        self.feedback_widget.opacity = 1
         self.feedback_popup.open()
 
     def get_feedback_popup(self):
         container = BoxLayout(orientation='vertical')
-        container.add_widget(self.feedback_widget)
+        container.add_widget(self.feedback_label)
 
-        button_layout = BoxLayout(size_hint_y=None, height=50)
+        button_layout = BoxLayout(size_hint_y=0.2)
         dismiss_button = Button(text='Dismiss', on_press=lambda instance: self.feedback_popup.dismiss())
         exit_button = Button(text='Exit', on_press=self.exit_funct)
         button_layout.add_widget(dismiss_button)
@@ -60,13 +59,15 @@ class FinishLayout(BoxLayout):
         container.add_widget(button_layout)
 
         # Create and return the popup
-        return Popup(title="Success!", content=container, size_hint=(0.95, 0.25),title_align='center', title_size=self.default_font_size)
+        return Popup(title="Success!", content=container, size_hint=(0.95, 0.4),title_align='center', title_size=self.default_font_size)
 
     @staticmethod
-    def get_feedback_widget(font_size: float) -> Widget:
-        return Label(size_hint=(1, 1),
-                          opacity=0,
-                          font_size=font_size)
+    def get_feedback_label(font_size: float) -> Widget:
+        text_label =  Label(size_hint=(1, 0.8),
+                            halign='left',
+                            font_size=font_size)
+        text_label.bind(size=text_label.setter('text_size'))
+        return text_label
 
     @staticmethod
     def get_ok_button() -> Widget:

### 2024-01-24 data_collector/elements/node_element.py
data_collector: Added node initialization readiness check

@@ -97,7 +97,7 @@ class NodeElement(FsNode, LoadableElem):
 
 
     def unload(self):
-        if self.is_loaded:
+        if self.is_loaded and self.is_initialized:
             self.root_container.remove_widget(self.get_node_widget())
             self.root_container.add_widget(self.place_holder, index=1)
             self.is_loaded = False
@@ -106,7 +106,7 @@ class NodeElement(FsNode, LoadableElem):
 
 
     def load(self):
-        if not self.is_loaded:
+        if not self.is_loaded and self.is_initialized:
             self.root_container.remove_widget(self.place_holder)
             self.root_container.add_widget(self.get_node_widget(), index=1)
             self.is_loaded = True

### 2024-01-24 data_collector/layouts/finish.py

@@ -64,7 +64,6 @@ class FinishLayout(BoxLayout):
     @staticmethod
     def get_feedback_label(font_size: float) -> Widget:
         text_label =  Label(size_hint=(1, 0.8),
-                            halign='left',
                             font_size=font_size)
         text_label.bind(size=text_label.setter('text_size'))
         return text_label

### 2024-01-24 data_collector/app.py
data_collector: Centered feedback text

@@ -97,9 +97,10 @@ class DataCollectApp(App):
 
         except Exception as e:
             self.finish_layout.feedback_popup.title = 'Failed to bundle files!'
-            self.finish_layout.feedback_label.text = (f'An error occured during the creating of the zip archive:\n {e}'
-                                                       f'\nIs "{target_folder_path}" a valid folder path?'
-                                                       f'\nDoes the executable have permission to view all selected files?')
+            self.finish_layout.feedback_label.text = (f'An error occured during the creation of the zip archive:'
+                                                       f'\n{e}'
+                                                       f'\n\nIs "{target_folder_path}" a valid folder path?'
+                                                       f'Does the executable have permission to view all selected files?')
 
         Clock.schedule_once(self.show_feedback)
 

### 2024-01-24 data_collector/layouts/finish.py

@@ -64,6 +64,8 @@ class FinishLayout(BoxLayout):
     @staticmethod
     def get_feedback_label(font_size: float) -> Widget:
         text_label =  Label(size_hint=(1, 0.8),
+                            valign='middle',
+                            halign='left',
                             font_size=font_size)
         text_label.bind(size=text_label.setter('text_size'))
         return text_label

### 2024-01-24 data_collector/layouts/selection.py
data_collector: Introduced spacing to checkboxes layout

@@ -13,7 +13,7 @@ from data_collector.filesystem.fsnode import FsNode
 from data_collector.elements.node_element import NodeElement
 from data_collector.elements.types import ThickVerticalSlider, HeaderWidget
 
-from data_collector.configs import get_line_height
+from data_collector.configs import get_line_height, get_true_height
 import intervals as I
 
 # -------------------------------------------
@@ -142,7 +142,7 @@ class SelectionLayout(BoxLayout):
 
     @staticmethod
     def get_checkboxes_layout(file_count_label: Label, scroll_view: ScrollView):
-        select_layout = BoxLayout(orientation='vertical')
+        select_layout = BoxLayout(orientation='vertical', spacing = 15 * get_true_height() / 900.)
         select_layout.add_widget(widget=file_count_label)
         select_layout.add_widget(widget=scroll_view)
 

### 2024-01-24 data_collector/layouts/selection.py
data_collector: Made scrolling dpi aware

@@ -103,7 +103,7 @@ class SelectionLayout(BoxLayout):
 
 
     def on_scroll_view_scroll(self, instance, value):
-        _ = instance
+        _, __ = instance, value
 
         self.slider.unbind(value=self.adjust_scroll_view)
 
@@ -151,7 +151,7 @@ class SelectionLayout(BoxLayout):
     @staticmethod
     def get_scroll_view():
         scroll_view = ScrollView(size_hint=(1, 1), bar_width=0)
-        scrol_unit = 200
+        scrol_unit = get_true_height()/10.
         scroll_view.scroll_distance = scrol_unit
         scroll_view.scroll_wheel_distance = scrol_unit
 

### 2024-01-24 data_collector/layouts/selection.py
data_collector: Added placeholder widget to scrollview to prevent overscroll

@@ -11,7 +11,7 @@ from kivy.clock import Clock
 
 from data_collector.filesystem.fsnode import FsNode
 from data_collector.elements.node_element import NodeElement
-from data_collector.elements.types import ThickVerticalSlider, HeaderWidget
+from data_collector.elements.types import ThickVerticalSlider, HeaderWidget, Placeholder
 
 from data_collector.configs import get_line_height, get_true_height
 import intervals as I
@@ -160,6 +160,7 @@ class SelectionLayout(BoxLayout):
     def get_scroll_view_content(self, root_checkbox: NodeElement):
         scroll_layout = BoxLayout(orientation='vertical', size_hint_y=None)
         scroll_layout.bind(minimum_height=scroll_layout.setter('height'))
+        scroll_layout.add_widget(Placeholder(height=get_line_height()/4.))
         self.recursively_add_boxes(gui_parent=scroll_layout, root_box=root_checkbox, indent=0)
 
         return scroll_layout

### 2024-01-24 data_collector/layouts/dialogue.py
data_collector: Introduced size hints to labels in input dialog

@@ -107,7 +107,7 @@ class InputDialog(Popup):
             'You can edit this list of formats to your preferences:')
 
         hint = Label(text=first_hint_text,
-                          size_hint=(1, None),
+                          size_hint=(1, 0.2),
                           font_size=InputDialog.font_size)
         hint.bind(width=lambda *x: setattr(hint, 'text_size', (hint.width, None)))
         hint.bind(texture_size=self._update_text_height)
@@ -124,7 +124,7 @@ class InputDialog(Popup):
         )
 
         second_hint = Label(text=second_hint_text,
-                                         size_hint=(1, None),
+                                         size_hint=(1, 0.35),
                                          font_size=InputDialog.font_size)
         second_hint.bind(width=lambda *x: setattr(second_hint, 'text_size', (second_hint.width, None)))
         second_hint.bind(texture_size=self._update_text_height)

### 2024-01-24 data_collector/layouts/dialogue.py
data_collector: Adjusted input dialog confirm font size

@@ -178,7 +178,7 @@ class InputDialog(Popup):
 
 
     def make_confirm_button(self) -> Widget:
-        confirm = Button(text='start search', size_hint=(1, 0.08), )
+        confirm = Button(text='start search', size_hint=(1, 0.08), font_size=InputDialog.font_size/1.2)
         confirm.background_color = (0, 1, 0, 0.6)  # (R, G, B, A)
         confirm.bind(on_press=self.on_confirm)
         return confirm

### 2024-01-24 data_collector/layouts/dialogue.py
data_collector: Adjusted font size in input dialog

@@ -160,7 +160,7 @@ class InputDialog(Popup):
 
     def make_input_folder_hint(self):
         hint = Label(text='Data folder path:',
-                                         size_hint=(1, None),
+                                         size_hint=(1, 0.1),
                                          font_size=InputDialog.font_size)
         hint.bind(width=lambda *x: setattr(hint, 'text_size', (hint.width, None)))
         hint.bind(texture_size=self._update_text_height)

### 2024-01-24 data_collector/layouts/dialogue.py
data_collector: Adjusted confirm button color

@@ -179,6 +179,6 @@ class InputDialog(Popup):
 
     def make_confirm_button(self) -> Widget:
         confirm = Button(text='start search', size_hint=(1, 0.08), font_size=InputDialog.font_size/1.2)
-        confirm.background_color = (0, 1, 0, 0.6)  # (R, G, B, A)
+        confirm.background_color = (0, 1, 0, 1)  # (R, G, B, A)
         confirm.bind(on_press=self.on_confirm)
         return confirm

### 2024-01-24 data_collector/layouts/dialogue.py
data_collector: Adjusted input dialog input text fields heights

@@ -140,7 +140,7 @@ class InputDialog(Popup):
             default_xrd_text += f'.{xrd_format},'
 
         format_input = FocusTextInput(text=f'{default_xrd_text}',
-                                           size_hint=(1, 0.08),
+                                           size_hint=(1, 0.085),
                                            font_size=InputDialog.font_size,
                                            multiline=False)
         Clock.schedule_once(lambda dt: setattr(format_input, 'focus', True), 0.1)
@@ -169,7 +169,7 @@ class InputDialog(Popup):
 
     def make_path_input(self) -> Widget:
         path_input = FocusTextInput(text=f'{get_initial_path()}',
-                                         size_hint=(1, 0.08),
+                                         size_hint=(1, 0.085),
                                          font_size=InputDialog.font_size,
                                          multiline=False)
 

### 2024-01-24 data_collector/elements/node_widget.py
data_collector: Removed unnecessary code

@@ -14,6 +14,11 @@ from data_collector.elements.types import BlackLabel
 from data_collector.resources import get_unchecked_box_path, get_checked_box_path, get_fileicon_path, \
     get_foldericon_path, get_collapsed_icon_path, get_expanded_icon_path, get_kivy_image, get_empty_path
 from kivy.uix.widget import Widget
+from kivy.uix.widget import Widget
+from kivy.graphics import Rectangle, Color
+from kivy.core.image import Image as CoreImage
+from kivy.properties import BooleanProperty, ObjectProperty
+
 
 # -------------------------------------------
 
@@ -66,11 +71,14 @@ class IconToggleButton(RelativeLayout):
     def __init__(self, **kwargs):
         super(IconToggleButton, self).__init__(**kwargs)
 
-        height = self.height
-        # self.collapsed = Image(source=get_collapsed_icon_path(), size_hint=(None, None),size=(height,height))
-        # self.expanded = Image(source=get_expanded_icon_path(), size_hint=(None, None),size=(height,height))
-        self.collapsed = get_kivy_image(width=height, imgPath=get_collapsed_icon_path(), size_hint=(None, None), size=(height,height))
-        self.expanded = get_kivy_image(width=height, imgPath=get_expanded_icon_path(), size_hint=(None, None), size=(height,height))
+        self.collapsed = get_kivy_image(width=self.height, imgPath=get_collapsed_icon_path(),
+                                        size_hint=(None, None),
+                                        size=(height,height))
+
+        self.expanded = get_kivy_image(width=self.height,
+                                       imgPath=get_expanded_icon_path(),
+                                       size_hint=(None, None),
+                                       size=(height,height))
 
         self.btn = ToggleButton(size_hint=(1, 1), background_color=(0, 0, 0, 0))
         self.add_widget(self.btn)
@@ -90,10 +98,6 @@ class IconToggleButton(RelativeLayout):
             self.icon = self.expanded
         self.add_widget(self.icon)
 
-from kivy.uix.widget import Widget
-from kivy.graphics import Rectangle, Color
-from kivy.core.image import Image as CoreImage
-from kivy.properties import BooleanProperty, ObjectProperty
 
 class ImageCheckBox(CheckBox):
     active = BooleanProperty(False)
@@ -118,9 +122,6 @@ class ImageCheckBox(CheckBox):
             self.active = not self.active
             return True
         return super().on_touch_down(touch)
-
[…]

### 2024-01-24 data_collector/filesystem/out_files.py

@@ -62,12 +62,3 @@ def get_rel_path_dict(abs_path_list: List[str], root_path : str) -> dict:
 
     return rel_path_dict
 
-
-
-
-# if __name__ == "__main__":
-#     source_path = ["/home/work/Desktop/test1/asdf2.png"]
-#     the_target_path = "/home/work/Desktop/target"
-#
-#     zip_file_list(source_path, the_target_path)
-

### 2024-01-24 data_collector/layouts/dialogue.py

@@ -17,7 +17,6 @@ from data_collector.resources.resource_manager import get_blended_logo_path, get
 # -------------------------------------------
 
 class InputDialog(Popup):
-
     font_size = get_true_width() * 0.0145
 
     def __init__(self, callback: callable, **kwargs):
@@ -27,20 +26,15 @@ class InputDialog(Popup):
         self.title_align = 'center'
         self.auto_dismiss = False
         self.callback = callback
+        self.background_color = (100 / 255, 1, 1, 1)  # (R, G, B, A)
 
-        horizontal_spacing = 12 * get_true_width() / 1600.
-        vertical_spacing = 12 * get_true_height() / 900.  # Assuming 900 is a relevant measure for height
+        h_space = 12 * get_true_width() / 1600.
+        v_space = 12 * get_true_height() / 900.
         self.content = BoxLayout(orientation='vertical',
-                                 padding=[horizontal_spacing, vertical_spacing, horizontal_spacing, vertical_spacing],
-                                 spacing=vertical_spacing)
+                                 padding=[h_space, v_space, h_space, v_space],
+                                 spacing=v_space)
 
-        print(f'App width is: {get_scaled_down_app_width()}')
-
-        self.background_color = (100 / 255, 255 / 255, 255 / 255, 1)  # (R, G, B, A)
-
-        # logo_image = Image(source=get_blended_logo_path(), size_hint=(1, 0.3))
         logo_image = get_kivy_image(width=Window.width*0.4, imgPath=get_blended_logo_path(), size_hint =(1,0.25))
-        print(f'Width of the blended logo: {Window.width*0.4}')
         first_hint = self.make_hint()
         self.format_input =self.make_format_input()
         second_hint = self.make_second_hint()
@@ -70,10 +64,10 @@ class InputDialog(Popup):
     def _update_text_height(instance, texture_size):
         instance.height = texture_size[1]  # Set the height to the text height
 
-    def on_confirm(self, instance):
-        FsNode.set_default_formats(self.format_input.text)
 
+    def on_confirm(self, instance):
         _ = instance
+        FsNode.set_default_formats(self.format_input.text)
         user_input = self.path_input.text
 
         print(f"User input: {user_input}")
@@ -83,7 +77,6 @@ class InputDialog(Popup):
             return
         else:
             self.print_wait()
-
 
         Clock.schedule_once(lambda dt: self.callback(user_input),0.1)
 

### 2024-01-24 data_collector/layouts/selection.py
data_collector: Removed overscroll prevention

@@ -106,13 +106,6 @@ class SelectionLayout(BoxLayout):
         _, __ = instance, value
 
         self.slider.unbind(value=self.adjust_scroll_view)
-
-        if self.scroll_view.scroll_y < 0:
-            self.scroll_view.scroll_y = 0
-
-        if self.scroll_view.scroll_y > 1:
-            self.scroll_view.scroll_y = 1
-
         self.slider.value = copy.copy(self.scroll_view.scroll_y)
         self.slider.bind(value=self.adjust_scroll_view)
 

### 2024-01-25 data_collector/layouts/selection.py
data_collector: Renamed and deleted unnecessary code in selection

@@ -25,7 +25,7 @@ class SelectionLayout(BoxLayout):
         # attributes
         self.root_checkbox : Optional[NodeElement] = None
         self.last_load_range = I.closed(float('-inf'),float('inf'))
-        self.callback = None
+        self.dismiss_popup = None
 
         # GUI elements
         self.slider = ThickVerticalSlider(orientation='vertical', min=0, max=1, value=1, size_hint=(0.1, 1))
@@ -57,15 +57,14 @@ class SelectionLayout(BoxLayout):
 
         scroll_view_content = self.get_scroll_view_content(root_checkbox=self.root_checkbox)
         self.scroll_view.add_widget(widget=scroll_view_content)
-        self.scroll_view.bind(scroll_y=self.update_node_population)
 
         update_rate = 0.2
         Clock.schedule_interval(self.update_node_population, update_rate)
         Clock.schedule_interval(self.update_do_scroll, update_rate)
-        Clock.schedule_interval(self.check_for_dismiss, update_rate)
+        Clock.schedule_interval(self.check_selection_readiness, update_rate)
 
     def register_content_done_callback(self, callback : callable):
-        self.callback = callback
+        self.dismiss_popup = callback
 
     # -------------------------------------------
     # logic
@@ -77,15 +76,14 @@ class SelectionLayout(BoxLayout):
 
         vp_ypos = (1-self.scroll_view.scroll_y)*(total_height-vp_height)
         buffer_range = vp_height/2.
-
         new_load_range = I.closed(lower=vp_ypos-buffer_range,upper=vp_ypos+vp_height+buffer_range)
         unload_range_list = self.last_load_range - new_load_range
 
+        nodes_to_load = self.root_checkbox.get_nodes_in_interval(node=self.root_checkbox, interval=new_load_range)
         nodes_to_unload = []
         for unload_range in unload_range_list:
             nodes_to_unload += self.root_checkbox.get_nodes_in_interval(node=self.root_checkbox, interval=unload_range)
 
-        nodes_to_load = self.root_checkbox.get_nodes_in_interval(node=self.root_checkbox, interval=new_load_range)
         for node in nodes_to_load:
             node.load()
 
@@ -96,10 +94,7 @@ class SelectionLayout(BoxLayout):
 
     def update_do_scroll(self, *args, **kwargs):
         _, __ = args, kwargs
-        total_height = self.get_total_height()
-        vp_height = self.get_vp_height()
-
[…]

### 2024-01-25 data_collector/elements/node_widget.py
data_collector: Streamlined scrolling mechanism in selection

@@ -2,23 +2,20 @@ from __future__ import annotations
 
 from kivy.core.window import Window
 from kivy.uix.boxlayout import BoxLayout
-
 from kivy.uix.checkbox import CheckBox
 from kivy.uix.image import Image
 from kivy.uix.relativelayout import RelativeLayout
 from kivy.uix.togglebutton import ToggleButton
+from kivy.uix.widget import Widget
+from kivy.graphics import Rectangle
+from kivy.core.image import Image as CoreImage
+from kivy.properties import BooleanProperty, ObjectProperty
 
 from data_collector.configs import get_line_height
 from data_collector.elements.types import BlackLabel
 
 from data_collector.resources import get_unchecked_box_path, get_checked_box_path, get_fileicon_path, \
     get_foldericon_path, get_collapsed_icon_path, get_expanded_icon_path, get_kivy_image, get_empty_path
-from kivy.uix.widget import Widget
-from kivy.uix.widget import Widget
-from kivy.graphics import Rectangle, Color
-from kivy.core.image import Image as CoreImage
-from kivy.properties import BooleanProperty, ObjectProperty
-
 
 # -------------------------------------------
 
@@ -71,11 +68,12 @@ class IconToggleButton(RelativeLayout):
     def __init__(self, **kwargs):
         super(IconToggleButton, self).__init__(**kwargs)
 
-        self.collapsed = get_kivy_image(width=self.height, imgPath=get_collapsed_icon_path(),
+        height = self.height
+        self.collapsed = get_kivy_image(width=height, imgPath=get_collapsed_icon_path(),
                                         size_hint=(None, None),
                                         size=(height,height))
 
-        self.expanded = get_kivy_image(width=self.height,
+        self.expanded = get_kivy_image(width=height,
                                        imgPath=get_expanded_icon_path(),
                                        size_hint=(None, None),
                                        size=(height,height))
@@ -124,6 +122,7 @@ class ImageCheckBox(CheckBox):
         return super().on_touch_down(touch)
 
     def update_rect(self, *args):
+        _ = args
         self.rect.pos = self.pos
         self.rect.size = self.size
         self.rect.texture = self.checked_image if self.active else self.unchecked_image

### 2024-01-25 data_collector/layouts/selection.py

@@ -1,6 +1,5 @@
 from __future__ import annotations
 
-import copy
 from typing import Optional
 from kivy.uix.label import Label
 from kivy.uix.scrollview import ScrollView
@@ -29,10 +28,10 @@ class SelectionLayout(BoxLayout):
 
         # GUI elements
         self.slider = ThickVerticalSlider(orientation='vertical', min=0, max=1, value=1, size_hint=(0.1, 1))
-        self.slider.bind(value=self.adjust_scroll_view)
+        self.slider.bind(value=self.do_scroll)
 
         self.scroll_view = self.get_scroll_view()
-        self.scroll_view.bind(scroll_y=self.on_scroll_view_scroll)
+        self.scroll_view.bind(scroll_y=self.do_scroll)
 
         self.header_layout = self.get_header_widget(num_elements=0)
         self.checkboxes_layout = self.get_checkboxes_layout(file_count_label=self.header_layout,
@@ -54,13 +53,10 @@ class SelectionLayout(BoxLayout):
 
         new_label = self.get_header_widget(num_elements=len(self.root_checkbox.get_xrd_file_des())).children[2]
         self.header_layout.children[2].text =  new_label.text
-
-        scroll_view_content = self.get_scroll_view_content(root_checkbox=self.root_checkbox)
-        self.scroll_view.add_widget(widget=scroll_view_content)
+        self.scroll_view.add_widget(widget=self.get_scroll_view_content(root_checkbox=self.root_checkbox))
 
         update_rate = 0.2
         Clock.schedule_interval(self.update_node_population, update_rate)
-        Clock.schedule_interval(self.update_do_scroll, update_rate)
         Clock.schedule_interval(self.check_selection_readiness, update_rate)
 
     def register_content_done_callback(self, callback : callable):
@@ -92,22 +88,13 @@ class SelectionLayout(BoxLayout):
 
         self.last_load_range = new_load_range
 
-    def update_do_scroll(self, *args, **kwargs):
-        _, __ = args, kwargs
+    def do_scroll(self, initiator, value):
         self.scroll_view.do_scroll_y = self.get_total_height() > self.get_vp_height()
 
-
-    def on_scroll_view_scroll(self, instance, value):
-        _, __ = instance, value
-
-        self.slider.unbind(value=self.adjust_scroll_view)
-        self.slider.value = copy.copy(self.scroll_view.scroll_y)
-        self.slider.bind(value=self.adjust_scroll_view)
-
-
-    def adjust_scroll_view(self, instance, value):
[…]

### 2024-01-25 data_collector/elements/types.py
data_collector: Updated setting num elements in header widget

@@ -77,26 +77,35 @@ class Placeholder(Widget):
 class HeaderWidget(BoxLayout):
     def __init__(self, num_elements: int, format_list : List[str]):
         super().__init__(orientation='horizontal', size_hint=(1, 0.125))
-        with_leading_dot_list = [f'.{the_format}' for the_format in format_list]
+        self.format_list = format_list
+        self.num_elements = num_elements
 
-        file_count_label = BlackLabel(
-            text=f"Found {num_elements} files that match specified XRD formats:\n {with_leading_dot_list}",
-            size_hint=(0.6, 1),  # Adjusted size hint
+        self.file_count_label = BlackLabel(
+            text=self._get_text(),
+            size_hint=(0.6, 1),
             font_size=Window.width * 0.0145,
             halign="center",
             valign="middle"
         )
-        file_count_label.bind(size=file_count_label.setter('text_size'))
-
-        # logo_image = Image(source=get_logo_path(), size_hint=(0.3, 1))
-        img_width = get_scaled_down_app_width() * 0.3
-        logo_image = get_kivy_image(width=img_width, imgPath=get_logo_path(), size_hint=(0.3,1))
-        print(f'Width of the header widget:{img_width}')
+        self.file_count_label.bind(size=self.file_count_label.setter('text_size'))
+        logo_image = get_kivy_image(width=get_scaled_down_app_width() * 0.3,
+                                    imgPath=get_logo_path(),
+                                    size_hint=(0.3,1))
 
         left_placeholder = Widget(size_hint=(0.4, 1))
         right_placeholder = Widget(size_hint=(0.2, 1))
 
-        self.add_widget(left_placeholder)  # Placeholder to balance the self
-        self.add_widget(file_count_label)
-        self.add_widget(right_placeholder)  # Placeholder to balance the self
+        self.add_widget(left_placeholder)
+        self.add_widget(self.file_count_label)
+        self.add_widget(right_placeholder)
         self.add_widget(logo_image)
+
+    def update_number_of_elements(self, num_elements : int):
+        self.num_elements = num_elements
+        self.file_count_label.text = self._get_text()
+
+
+
+    def _get_text(self) -> str:
+        format_with_leading_dots = [f'.{the_format}' for the_format in self.format_list]
+        return f"Found {self.num_elements} files that match specified XRD formats:\n {format_with_leading_dots}"

### 2024-01-25 data_collector/layouts/selection.py

@@ -51,8 +51,7 @@ class SelectionLayout(BoxLayout):
         print(f'Time taken to initialize filesystem: {self.root_checkbox.time_filesystem_searching}')
         print(f'Time taken to search relevancy: {self.root_checkbox.time_relevancy_sorting}')
 
-        new_label = self.get_header_widget(num_elements=len(self.root_checkbox.get_xrd_file_des())).children[2]
-        self.header_layout.children[2].text =  new_label.text
+        self.header_layout.update_number_of_elements(num_elements=len(self.root_checkbox.get_xrd_file_des()))
         self.scroll_view.add_widget(widget=self.get_scroll_view_content(root_checkbox=self.root_checkbox))
 
         update_rate = 0.2
@@ -87,6 +86,7 @@ class SelectionLayout(BoxLayout):
             node.unload()
 
         self.last_load_range = new_load_range
+
 
     def do_scroll(self, initiator, value):
         self.scroll_view.do_scroll_y = self.get_total_height() > self.get_vp_height()

### 2024-01-25 data_collector/layouts/dialogue.py
data_collector: Remved unused variable warnnig

@@ -10,7 +10,7 @@ from kivy.uix.button import Button
 from kivy.uix.label import Label
 from kivy.uix.popup import Popup
 
-from data_collector.configs import get_scaled_down_app_width, get_true_width, get_true_height
+from data_collector.configs import get_true_width, get_true_height
 from data_collector.elements.types import FocusTextInput
 from data_collector.filesystem import FsNode, get_initial_path
 from data_collector.resources.resource_manager import get_blended_logo_path, get_kivy_image

### 2024-01-25 data_collector/run/dev_run.py

@@ -13,10 +13,12 @@ def main():
     win_test_dir = 'C:\\Users\\daniel\\OneDrive\\Desktop\\test_folder'
     win_target_test = 'C:\\Users\\daniel\\OneDrive\\Desktop'
 
-    _, __ = win_test_dir, win_target_test
-
     linux_input_test = '/home/daniel/OneDrive'
     linux_target_test = f'/home/daniel/aimat/tool_target_folder'
+
+
+    _, __ = win_test_dir, win_target_test
+    _, __ = linux_input_test, linux_target_test
 
     app = DataCollectApp(override_input_folder=None,
                          override_target_folder=linux_target_test)

### 2024-01-25 package.sh
general: Renamed data_collector to submission helper

@@ -23,6 +23,6 @@ DIST_DIR="$ROOT_DIR/dist"
 BUILD_DIR="$ROOT_DIR/build"
 script_dir="$(cd "$(dirname "$0")" && pwd)"
 
-pyinstaller "$script_dir/data_collector/prod_run.spec" --distpath "$DIST_DIR" --workpath "$BUILD_DIR"
+pyinstaller "$script_dir/submission_helper/prod_run.spec" --distpath "$DIST_DIR" --workpath "$BUILD_DIR"
 echo "-> Built executable under $ROOT_DIR/dist"
 echo "done"

### 2024-01-25 submission_helper/app.py

@@ -8,8 +8,8 @@ from typing import List
 
 from kivy.clock import Clock
 
-from data_collector.filesystem import zip_file_list, produce_csv_file
-from data_collector.layouts import SelectionLayout, FinishLayout, InputDialog
+from submission_helper.filesystem import zip_file_list, produce_csv_file
+from submission_helper.layouts import SelectionLayout, FinishLayout, InputDialog
 
 # -------------------------------------------
 

### 2024-01-25 submission_helper/elements/loadable_element.py

@@ -1,6 +1,6 @@
 from __future__ import annotations
 
-from data_collector.elements.node_widget import LabeledCheckBox
+from submission_helper.elements.node_widget import LabeledCheckBox
 from abc import abstractmethod
 from intervals import Interval
 

### 2024-01-25 submission_helper/elements/node_element.py

@@ -5,12 +5,12 @@ from PIL import Image
 from typing import Optional, List
 from kivy.uix.boxlayout import  BoxLayout
 
-from data_collector.elements.types import Placeholder
-from data_collector.elements.node_widget import LabeledCheckBox, NodeWidget
-from data_collector.elements.loadable_element import LoadableElem
-from data_collector.resources import get_foldericon_path, get_fileicon_path
-from data_collector.filesystem import FsNode
-from data_collector.configs import get_line_height
+from submission_helper.elements.types import Placeholder
+from submission_helper.elements.node_widget import LabeledCheckBox, NodeWidget
+from submission_helper.elements.loadable_element import LoadableElem
+from submission_helper.resources import get_foldericon_path, get_fileicon_path
+from submission_helper.filesystem import FsNode
+from submission_helper.configs import get_line_height
 # -------------------------------------------
 
 class NodeElement(FsNode, LoadableElem):

### 2024-01-25 submission_helper/elements/node_widget.py

@@ -11,10 +11,10 @@ from kivy.graphics import Rectangle
 from kivy.core.image import Image as CoreImage
 from kivy.properties import BooleanProperty, ObjectProperty
 
-from data_collector.configs import get_line_height
-from data_collector.elements.types import BlackLabel
+from submission_helper.configs import get_line_height
+from submission_helper.elements.types import BlackLabel
 
-from data_collector.resources import get_unchecked_box_path, get_checked_box_path, get_fileicon_path, \
+from submission_helper.resources import get_unchecked_box_path, get_checked_box_path, get_fileicon_path, \
     get_foldericon_path, get_collapsed_icon_path, get_expanded_icon_path, get_kivy_image, get_empty_path
 
 # -------------------------------------------

### 2024-01-25 submission_helper/elements/types.py

@@ -9,8 +9,8 @@ from kivy.uix.textinput import TextInput
 from kivy.core.window import Window
 
 from kivy.uix.widget import Widget
-from data_collector.configs import get_scaled_down_app_width
-from data_collector.resources import get_logo_path, get_kivy_image
+from submission_helper.configs import get_scaled_down_app_width
+from submission_helper.resources import get_logo_path, get_kivy_image
 from typing import List
 
 class FocusTextInput(TextInput):

### 2024-01-25 submission_helper/filesystem/fsnode.py

@@ -7,7 +7,7 @@ from abc import abstractmethod
 from typing import Optional
 import time
 
-from data_collector.filesystem.lib import make_fsys_dict
+from submission_helper.filesystem.lib import make_fsys_dict
 from kivy.clock import Clock
 
 # -------------------------------------------

### 2024-01-25 submission_helper/filesystem/out_files.py

@@ -1,4 +1,4 @@
-from data_collector.resources import get_template_csv
+from submission_helper.resources import get_template_csv
 
 import os
 import zipfile, csv

### 2024-01-25 submission_helper/layouts/dialogue.py

@@ -10,10 +10,10 @@ from kivy.uix.button import Button
 from kivy.uix.label import Label
 from kivy.uix.popup import Popup
 
-from data_collector.configs import get_true_width, get_true_height
-from data_collector.elements.types import FocusTextInput
-from data_collector.filesystem import FsNode, get_initial_path
-from data_collector.resources.resource_manager import get_blended_logo_path, get_kivy_image
+from submission_helper.configs import get_true_width, get_true_height
+from submission_helper.elements.types import FocusTextInput
+from submission_helper.filesystem import FsNode, get_initial_path
+from submission_helper.resources.resource_manager import get_blended_logo_path, get_kivy_image
 # -------------------------------------------
 
 class InputDialog(Popup):

### 2024-01-25 submission_helper/layouts/finish.py

@@ -7,10 +7,10 @@ from kivy.core.window import Window
 from kivy.uix.textinput import TextInput
 from kivy.uix.widget import Widget
 
-from data_collector.elements.types import BlackLabel
+from submission_helper.elements.types import BlackLabel
 from kivy.uix.popup import Popup
 from kivy.uix.label import Label
-from data_collector.configs import get_true_height, get_true_width
+from submission_helper.configs import get_true_height, get_true_width
 # -------------------------------------------
 
 class FinishLayout(BoxLayout):

### 2024-01-25 submission_helper/layouts/selection.py

@@ -8,11 +8,11 @@ from kivy.uix.boxlayout import BoxLayout
 from kivy.clock import Clock
 
 
-from data_collector.filesystem.fsnode import FsNode
-from data_collector.elements.node_element import NodeElement
-from data_collector.elements.types import ThickVerticalSlider, HeaderWidget, Placeholder
+from submission_helper.filesystem.fsnode import FsNode
+from submission_helper.elements.node_element import NodeElement
+from submission_helper.elements.types import ThickVerticalSlider, HeaderWidget, Placeholder
 
-from data_collector.configs import get_line_height, get_true_height
+from submission_helper.configs import get_line_height, get_true_height
 import intervals as I
 
 # -------------------------------------------

### 2024-01-25 submission_helper/run/dev_run.py

@@ -1,11 +1,11 @@
-from data_collector.configs import set_configs
+from submission_helper.configs import set_configs
 set_configs()
 
 from kivy.core.window import Window
 Window.clearcolor = (1, 1, 1, 1)
 
 from kivy.modules import inspector
-from data_collector.app import DataCollectApp
+from submission_helper.app import DataCollectApp
 # -------------------------------------------
 
 

### 2024-01-25 submission_helper/run/prod_run.py

@@ -1,9 +1,9 @@
-from data_collector.configs import set_configs
+from submission_helper.configs import set_configs
 set_configs()
 
 from kivy.core.window import Window
 Window.clearcolor = (1, 1, 1, 1)
-from data_collector.app import DataCollectApp
+from submission_helper.app import DataCollectApp
 # -------------------------------------------
 
 

### 2024-01-25 submission_helper/layouts/finish.py
submission_helper: Added sectioning to resource_manager

@@ -26,6 +26,7 @@ class FinishLayout(BoxLayout):
                                            size_hint=(0.7, 1),
                                            font_size=get_true_width() * 0.02,
                                            multiline=False)
+
         buffer = BlackLabel(text='', size_hint=(0.075, 1))
         self.ok_button = self.get_ok_button()
         self.ok_button.bind(on_press=callback)
@@ -53,6 +54,7 @@ class FinishLayout(BoxLayout):
         button_layout = BoxLayout(size_hint_y=0.2)
         dismiss_button = Button(text='Dismiss', on_press=lambda instance: self.feedback_popup.dismiss())
         exit_button = Button(text='Exit', on_press=self.exit_funct)
+
         button_layout.add_widget(dismiss_button)
         button_layout.add_widget(exit_button)
 

### 2024-01-25 submission_helper/resources/resource_manager.py

@@ -7,7 +7,6 @@ from io import BytesIO
 from PIL import Image
 
 
-# Determine the base path for resources based on whether the app is bundled or run as a script
 if getattr(sys, 'frozen', False):
     # The app is bundled with PyInstaller
     base_path = sys._MEIPASS  # PyInstaller's base directory
@@ -17,6 +16,9 @@ else:
 
 images_folder_path = os.path.join(base_path, 'images')
 doc_folder_path = os.path.join(base_path, 'documents')
+
+# -------------------------------------------
+# images
 
 def get_blended_logo_path() -> str:
     return os.path.join(images_folder_path, 'blended_logo.png')
@@ -45,11 +47,17 @@ def get_checked_box_path() -> str:
 def get_unchecked_box_path() -> str:
     return os.path.join(images_folder_path, 'unchecked_box.png')
 
+# -------------------------------------------
+# documents
+
 def get_template_csv() -> str:
     return os.path.join(doc_folder_path, 'template.csv')
 
 def get_empty_path() -> str:
     return os.path.join(images_folder_path, 'empty.png')
+
+# -------------------------------------------
+# resizing with PIL instead of kivy
 
 def pil_to_kivy_image(pil_image, **kwargs):
     data = BytesIO()

### 2024-01-25 submission_helper/package.sh
submission_helper: Moved rest of code inside directory

@@ -23,6 +23,6 @@ DIST_DIR="$ROOT_DIR/dist"
 BUILD_DIR="$ROOT_DIR/build"
 script_dir="$(cd "$(dirname "$0")" && pwd)"
 
-pyinstaller "$script_dir/submission_helper/prod_run.spec" --distpath "$DIST_DIR" --workpath "$BUILD_DIR"
+pyinstaller "prod_run.spec" --distpath "$DIST_DIR" --workpath "$BUILD_DIR"
 echo "-> Built executable under $ROOT_DIR/dist"
 echo "done"

### 2024-04-09 paper/00_abstract.tex
Imported structure from Henrik thesis paper; Included author details

We, the AiMat group at the Karlsruhe Institute of Technology (KIT), are collecting a dataset of both labeled and unlabeled experimental powder X-ray diffraction (pXRD) data, which will be made publicly available once sufficient data has been collected. We look forward to publishing a paper surveying the database and discussing its uses, and would be happy to include anyone contributing data as co-authors. You can view the current draft here: State-of-the-database

The main motivation for collecting and publishing this dataset is to assist researchers working on machine learning based approaches for pXRD analysis. Any data, even unlabeled data, can provide clues how simulations can be made to better reflect real-world experiments and labeled data can be used to test the developed algorithms. We hope to provide both types of data in a wide variety and large quantity.

While we do not collect single crystal data, we are interested in any kind of polycrystalline samples, including thin films. If your samples exhibit preferred crystallite orientation, please make mention of this in the description field of the upload form. So far, we have received upwards of 4155 unique powder X-ray diffraction patterns from institutions including:

%Machine learning techniques have successfully been used to extract structural information such as the crystal space group from powder X-ray diffractograms.
%However, training directly on simulated diffractograms from databases such as
%the ICSD is challenging due to its limited size, class-inhomogeneity, and bias
%toward certain structure types. We propose an alternative approach of generating
%synthetic crystals with random coordinates by using the symmetry operations of
%each space group. Based on this approach, we 
%demonstrate online training of deep ResNet-like models on up to a few
%million unique on-the-fly generated synthetic diffractograms per hour. For our chosen task of
%space group classification, we achieved a test accuracy of 79.9\% on unseen ICSD
%structure types from most space groups. This surpasses the 56.1\% accuracy of the current state-of-the-art approach of training on ICSD crystals directly. Our results demonstrate that synthetically generated crystals can be
%used to extract structural information from ICSD powder diffractograms, which makes it possible to apply very large state-of-the-art machine learning models in the area of powder X-ray diffraction.
%We further show first steps toward applying our methodology to experimental data, where automated XRD data analysis is crucial, especially in high-throughput settings.
%While we focused on the prediction of the space group, our approach has the potential to be extended to related tasks in the future.

### 2024-04-09 paper/01_motivation.tex
Imported structure from Henrik thesis paper; Included author details

Machine learning techniques have emerged as a powerful tool in the toolkit of
materials scientists. While they are often used to make predictions on
the properties of materials or find new materials with certain properties, an increasingly interesting domain is the automated analysis of raw experimental measurements
guided by machine learning\supercite{radovicMachineLearningEnergy2018}.

With the advent of high-throughput experiments, the amount of gathered data is
vast and the analysis often becomes a bottleneck in the processing pipeline
\supercite{rahmanianEnablingModularAutonomous2022}. Powder X-ray diffraction
(XRD) is an important measurement technique used to obtain structural
information from polycrystalline samples\supercite{harrisContemporaryAdvancesUse2001}. 
The diffractograms are an
information-dense fingerprint of the structure of the material. However,
analyzing these diffractograms is not an easy task\supercite{holderTutorialPowderXray2019}. 
Full structure solutions and
Rietveld refinement take time and require expert knowledge, both about the
analysis technique and the materials class at hand. This is not feasible in high-throughput experiments on a larger
scale. Therefore, the question arises whether it is possible to automatically 
analyze powder diffractograms with machine learning models trained on large
amounts of data, making it possible to run inference almost instantaneously.

During the last few years, there have been several studies tackling this
objective by applying machine learning models to various tasks concerning the
analysis of powder diffractograms such as phase
classification\supercite{leeDeeplearningTechniquePhase2020,maffettoneCrystallographyCompanionAgent2021,schuetzkeEnhancingDeeplearningTraining2021,szymanskiProbabilisticDeepLearning2021,wangRapidIdentificationXray2020},
phase fraction determination\supercite{leeDatadrivenXRDAnalysis2021}, space
group
classification\supercite{parkClassificationCrystalStructure2017,oviedoFastInterpretableClassification2019,zalogaCrystalSymmetryClassification2020,vecseiNeuralNetworkBased2019,suzukiSymmetryPredictionKnowledge2020,chakrabortyDeepCrystalStructure2022},
machine-learning-guided Rietveld
refinement\supercite{ozakiAutomatedCrystalStructure2020,fengMethodArtificialIntelligence2019},
extraction of lattice
parameters\supercite{dongDeepConvolutionalNeural2021,chitturiAutomatedPredictionLattice2021,chakrabortyDeepCrystalStructure2022,
habershonPowderDiffractionIndexing2004} and crystallite
sizes\supercite{dongDeepConvolutionalNeural2021,chakrabortyDeepCrystalStructure2022},
and also novelty detection based on unsupervised techniques\supercite{bankoDeepLearningVisualization2021}.
Since an abundant source of experimental diffractograms is hard to come by, most
applications train their models on simulated diffractograms from the Inorganic
Crystal Structure Database (ICSD)\supercite{bergerhoff1987}, which contains a total of 272\,260 structures (October 2022).

\citeauthor{leeDeeplearningTechniquePhase2020} used a deep convolutional neural
network (CNN) trained on a large dataset of multiphase compositions from the
quaternary \ce{Sr-Li-Al-O} pool to classify present phases in the
diffractogram\supercite{leeDeeplearningTechniquePhase2020}. In a follow-up
study, they further showed good results for phase fraction inference in the
quaternary \ce{Li-La-Zr-O} pool\supercite{leeDatadrivenXRDAnalysis2021}.
\citeauthor{schuetzkeEnhancingDeeplearningTraining2021} performed phase
classification on iron ores and cement compounds and used data augmentation with
respect to lattice parameters, crystallite sizes, and preferred orientation
\supercite{schuetzkeEnhancingDeeplearningTraining2021}. They showed that
especially the lattice parameter variations enhance the classification accuracy
significantly.

Instead of the analysis of phase composition,
\citeauthor{dongDeepConvolutionalNeural2021} performed regression of scale
factors, lattice parameters, and crystallite sizes in a five-phase catalytic
materials system\supercite{dongDeepConvolutionalNeural2021}. In contrast to
supervised tasks, \citeauthor{bankoDeepLearningVisualization2021} used a
variational autoencoder to visualize variations in space group, preferred
orientation, crystallite size, and peak
shifts\supercite{bankoDeepLearningVisualization2021}.
\citeauthor{parkClassificationCrystalStructure2017} used a deep CNN to classify
space groups of single-phase diffractograms, reaching a test accuracy of 81.14\%
on simulated diffractograms.\supercite{parkClassificationCrystalStructure2017} 
However, as we will show later in this paper, this accuracy is highly overestimated and drops to 56.1\% when test splits are designed in a way to reduce data leakage in non-IID datasets such as the ICSD.
\citeauthor{vecseiNeuralNetworkBased2019}\supercite{vecseiNeuralNetworkBased2019} developed
a similar approach and applied their classifier to experimental diffractograms
from the RRUFF mineral database\supercite{lafuentePowerDatabasesRRUFF2015},
reaching an experimental test accuracy of 54\%.

While the ICSD contains a large number of structures spanning many different
classes of materials, it still falls short in size, distribution, and generality
compared to the datasets used to train very large state-of-the-art models of
other fields such as computer vision.
%for image classification (ImageNet
%challenge\supercite{russakovskyImageNetLargeScale2015}). 
Furthermore, the ICSD
database is highly imbalanced with respect to space groups, as can be seen
in the histogram in Figure~\ref{fig:histogram_dist_spgs}. This makes the classification of
space groups more difficult, as shown and discussed by
\citeauthor{zalogaCrystalSymmetryClassification2020}
\supercite{zalogaCrystalSymmetryClassification2020}. The ICSD also contains a
limited number of different structure types that may not adequately represent
the crystal structures analyzed in future experiments.

To overcome these shortcomings, we propose to train machine learning models on
diffractograms simulated from synthetic crystal structures randomly generated
based on the symmetry operations of the space groups. This makes it possible to
train on structures with new structure types not present in the ICSD. We used
the crystals from the ICSD only to determine vague statistics guiding the random
generation and for calculating the test accuracy. Our approach goes one step
further than classical data augmentation by fully detaching itself from the
individual entries in the ICSD database.
\replaced{The generated synthetic crystals form a training dataset that includes stable ICSD crystal structures, unstable crystal structures,
but also stable structures that are not yet present in the ICSD. By training a model on the full dataset, we can also expect improvements on the unknown stable crystal structures.}{It is our hypothesis that when trying
to solve more general tasks concerning the extraction of structural details from
powder diffractograms, the crystals from which the training dataset is simulated
do not need to be stable or chemically viable.} \replaced{Furthermore, we}{We} propose viewing the problem
\deleted{simply} as a mathematical task of getting back some of the real-space information
leading to given powder X-ray diffractograms. \added{Therefore, even the unstable structures included in our generated dataset will help to learn to classify the stable structures.}

Here, we applied this approach to the classification of the crystal symmetry,
namely the space group. The space group is usually one of the first structural
pieces of information needed after synthesizing a new material. This task is
well-suited to showcase the strengths of using a synthetic dataset and to
benchmark it. We further show the results of using our methodology to infer space group labels of an experimental dataset.

We embedded our synthetic generation algorithm in a framework with
distributed computing capabilities to generate and simulate diffractograms on
multiple nodes in parallel using the \emph{Python} library \emph{Ray}
\supercite{moritzRayDistributedFramework2018a}. In contrast to the traditional
approach of generating a simulated dataset before training, we used this
distributed computing architecture to build an infinite stream of \replaced{synthetically}{randomly}
generated and simulated diffractograms to perform batch-wise online learning.
This increases the generalization performance, eliminates the problem of
overfitting, and allows very large models to be trained.

### 2024-04-09 paper/02_existing_datasets.tex
Imported structure from Henrik thesis paper; Included author details

\subsection{Generating synthetic crystals} \label{sec:generate_crystals}

    To generate synthetic crystals, we randomly place atoms on the Wyckoff
    positions of a given space group \added{following the Wyckoff occupation probabilities extracted from the ICSD} and
    then apply the respective symmetry operations. The algorithm is
    explained in the following (see also Figure~\ref{fig:generation_and_distributed}a for a flow diagram of the algorithm).
    We only explain the most important steps, details can be found in Section~S1 of the SI.

    \begin{enumerate}
        \itemsep0em
        \item \replaced{Sampling of a space group from the space group distribution of the ICSD.}{Random selection of a space group. We follow the space group distribution of the ICSD to allow comparison with previous work.}
        \item Sample unique elements of the crystal and their number of repetitions in the asymmetric unit.
        \item Place atoms onto the Wyckoff positions and draw uniform coordinates for each.
        \item Draw lattice parameters from a \replaced{kernel density estimate}{KDE} based on the ICSD.
        \item Apply space group symmetry operations.
    \end{enumerate}

    Parts of this algorithm were inspired by the generation algorithm of the
    \emph{Python} library \emph{PyXtal}
    \supercite{fredericksPyXtalPythonLibrary2021}. 
    We only keep generated crystals for training if the
    conventional unit cell volume is below \SI{7000}{\angstrom\cubed} and if
    there are less than 100 atoms in the asymmetric unit. 
    We did not employ any form of
    distance checks on the coordinates, as we found this to have no meaningful
    impact on space group classification accuracy. We only prevented the
    algorithm from placing more than one atom onto a Wyckoff position that does not
    have a degree of freedom. We also did not use partial occupancies. We chose
    this algorithm for its simplicity and its capability to reproduce many
    important characteristics of ICSD crystals adequately (see Section~
    \ref{sec:synth_distribution}).

    \begin{figure}[h]
    \centering
    \includegraphics{./figures/distribution_spgs_arrows.pdf}
    \caption{Distribution (logarithmic scale in blue, linear scale in red)
    of space groups in the ICSD. Space groups are sorted by count\added{ (see Figure S13 in the SI for the distribution without sorting by count)}. The population of the space groups varies by
    multiple orders of magnitude, showing that the ICSD is a highly imbalanced
    dataset regarding space groups. The space groups excluded due to insufficient
    statistics are visualized with black stripes. \replaced{The histogram displays the distribution of the full ICSD, while the exclusion of space groups that do not contain enough samples is based on the statistics dataset (which does not include the test dataset, see Section~\ref{sec:dataset_split}) that we used to guide the random crystal generation. Therefore, the excluded space groups are not exactly the first 85 counted from the left.}{Since the exclusion is based on 
    the statistics dataset and this histogram shows the full ICSD,
    the excluded space groups are not exactly the first 85 counted from the left.}}
    \label{fig:histogram_dist_spgs}
    \end{figure}

    For some space groups, there are not enough crystals in the ICSD to form a
    representative \replaced{kernel density estimate}{KDE} for the volume or to calculate suitable occupation
    probabilities for individual Wyckoff positions. Therefore, we chose to only
    perform the classification on space groups with 50 or more crystals
    available in the statistics dataset we used to extract the probabilities
    (see Section~\ref{sec:dataset_split}), resulting in the exclusion of 85
    space groups (see Figure~\ref{fig:histogram_dist_spgs}). A classifier trained directly on ICSD data of all space groups will likely not be able to properly identify these space
    groups containing very few samples.

    If a similar performance for all space groups is desired, a uniform
    distribution of space groups in the training dataset is needed. This is
    trivially possible with our synthetic approach, in contrast to training
    directly on the ICSD, where weighting, over-, or undersampling methods are
    needed \supercite{sunClassificationImbalancedData2009}. To allow a direct
    and fair comparison between our approach and the original approach of
    training directly on ICSD entries, we still followed the same distribution
    of space groups of the ICSD in our synthetic training dataset. This
    eliminates the problem that the effective number of total space groups is
    smaller when training on a highly imbalanced dataset, making it easier to
    reach high accuracies.

    \added{Our choice of not sampling the space groups uniformly and using
    general statistics extracted from the ICSD to guide the crystal generation
    algorithm further builds upon the hypothesis that future crystals will
    roughly follow the more general statistics already present in the ICSD. With
    the chosen crystal generation algorithm we tried to find a middle ground
    between being much more general than using the ICSD crystals directly and
    not being too general such that it is very hard to extract structural
    information at all.}

    \begin{figure}[!htb] 
    \centering
    \includegraphics{./figures/combined_generation_and_distributed_system.pdf}
    \caption{a) Flowchart of how the generation algorithm produces synthetic
    crystals. Atoms are independently placed on the Wyckoff positions and random
    coordinates are drawn. b) Overview of the distributed computing system
    implemented using the \emph{Python} library
    \emph{Ray}\supercite{moritzRayDistributedFramework2018a}. Two compute nodes
    (that generate and simulate diffractograms) are connected to the \emph{Ray}
    head node using a \emph{Ray queue} object.}
    \label{fig:generation_and_distributed}
    \end{figure}

    \subsection{Simulating diffractograms} \label{sec:simulation}

    To simulate powder X-ray diffractograms based on the generated crystals, we used
    the implementation found in the \emph{Python} library \emph{Pymatgen}
    \supercite{ongPythonMaterialsGenomics2013}. We optimized the simulation code
    using the LLVM just-in-time compiler \emph{Numba}
    \supercite{lamNumbaLLVMbasedPython2015}. This increases the performance of
    the main loop over the reciprocal lattice vectors of the crystal
    significantly and makes the continuous simulation while training (discussed
    in the next section) possible.

    We used the wavelength \SI{1.5406}{\angstrom} (\replaced{Cu K$\alpha_1$}{Cu K$_\alpha$} line) to
    simulate all diffractograms. The obtained peaks were further broadened with
    a Gaussian peak profile to form the full diffractogram. To obtain the peak
    widths, we used the Scherrer
    equation\supercite{gilmoreInternationalTablesCrystallography2019}
    \begin{equation}
        \beta=\frac{K \lambda}{L \cos \theta} \text{,} \label{eq:scherrer}
    \end{equation}
    where $ \beta $ is the line broadening at half maximum intensity (on the
    $2\theta$-scale), $K$ is a shape factor, $\lambda$ is the wavelength, and
    $L$ is the (average) thickness of crystallites. We drew crystallite sizes
    from the range $ \left[ 20, 100 \right] $ \si{\nano \metre} and used 
    $K=0.9$.

    Diffractograms were generated in the range $ 2 \theta \in \left[ 5, 90
    \right] \si{\degree} $ with step size \SI{0.01}{\degree}. After generating
    each diffractogram, it was rescaled to fit in the intensity range $ \left[
    0,1 \right]$. \added{In Figure S9 of the SI we show an exemplary diffractogram simulated 
    from the ICSD, Figure S10 shows an exemplary diffractogram simulated from a synthetic 
    crystal.}

    \subsection{Continuous generation of training data}\label{sec:online_learning}

    Typically, machine learning models are trained with a fixed dataset predefined at the beginning of training. Sometimes, data augmentation
    is applied to further increase the effective size of this dataset. In
    contrast to that, we generated our dataset on-the-fly, parallel to model training. The
    main advantage of using this approach compared to a fixed-size dataset is
    the eliminated possibility to overfit to individual diffractograms since
    every diffractogram is only used once. Furthermore, not having to
    pre-simulate a dataset before training makes this approach more flexible
    when changing simulation parameters.

    We used a distributed architecture on multiple nodes using the \emph{Python}
    framework \emph{Ray} \supercite{moritzRayDistributedFramework2018a}, which enabled the training on 1-2 GPUs and simultaneous generation of training data on more than 200 CPU cores (see Figure~\ref{fig:generation_and_distributed}b and SI Section~S2.2). 
    Depending on
    the model size and corresponding training speed, this setup allows
    training with up to millions of unique diffractograms per hour.

    \subsection{Dataset split} \label{sec:dataset_split}

    The ICSD database contains many structures that are very similar with
    slightly different lattice parameters and coordinates. For example, there
    are 25 entries for \ce{NaCl} (October 2022). Furthermore,
    there are 3898 entries that have the same structure type as \ce{NaCl} and
    thus also similar powder diffractograms. If some of them appear in the
    training dataset and some in the test dataset, the classification will be
    simplified to recognizing the structure type or structure. In that case, the
    test set accuracy will not represent the true generalization performance of the neural
    network. To quantify the true generalization performance, we split the
    dataset in such a way that the same structure type appears either only in
    the training or in the test dataset. We used the structure type definitions
    provided by the ICSD. The obtained accuracy on the test dataset reflects the
    accuracy of our network when being used on a novel sample with a structure
    type not yet present in the ICSD database.

    We want to emphasize that the used test split is very important for the task of
    space group classification and not a trivial choice. The ICSD contains many
    subtypes of structure types (for example, subtypes of perovskites), which we
    regarded as separate structure types in our split. Considering the subtypes
    as the same structure type may also be a viable option when performing the
    split. A combination of a split based on structure type and sum formula or
    similar approaches are also possible. \deleted{However, we think that our choice of
    splitting the dataset is able to effectively measure the generalization
    performance while still being relatively simple.}

    \added{Depending on the experimental setting, it further might make more sense in some cases to not do a structure type-based split. If the likelihood of finding structures similar to already-discovered structure types in the planned experiment is high, training should definitely include those structure types to evaluate the performance of the model. However, in a pure discovery setting, new structure types can appear. To evaluate the expected model performance in this scenario and thus quantify the true generalization error to unseen data, we chose the most strict structure type-based split.}

    We divided the ICSD (database version 2021, June 15) in a 70:30 split. For
    our synthetic crystal approach, the 70\% part (\replaced{which we
    call}{called} statistics or training dataset) was only used to create the
    \replaced{kernel density estimates}{KDEs} and to calculate the Wyckoff
    occupation probabilities needed for the generation algorithm. Since we can
    judge the performance of the synthetic generation algorithm by comparing the
    training accuracy (on synthetic crystals) with the accuracy tested on
    diffractograms simulated directly from the statistics dataset, an additional
    validation dataset was not needed. For comparison with the original approach
    of directly training on ICSD
    crystals\supercite{parkClassificationCrystalStructure2017}, we simulated
    crystals directly from the statistics dataset and trained on them.

    Analogous to the synthetic generation, we only used crystals with a
    conventional unit cell volume below \SI{7000}{\angstrom\cubed} and with less
    than 100 atoms in the asymmetric unit for the statistics and test dataset.
    This covers $ \approx 94\% $ of the ICSD crystals.

    \subsection{Models and \added{computational} experiments} \label{sec:models_and_experiments}

    \subsubsection*{Models}
    We will briefly introduce the models we used for the classification
    of space groups. A more detailed description can be found in the SI
    Section~S2.1.

    As a baseline, we first used the CNN models introduced by
    \citeauthor{parkClassificationCrystalStructure2017}\supercite{parkClassificationCrystalStructure2017}.
    They used three models, one for the classification of crystal systems
    (``parkCNN small''), one for extinction groups (``parkCNN medium''), and one
    for space groups (``parkCNN big''). All models have three convolution layers
    with two hidden fully connected layers and one output layer. The
    three models differ in the number of neurons in the hidden fully connected
    layers, increasing the number of model parameters with the number of target
    labels. Here, we only used the models ``parkCNN medium'' and ``parkCNN big''
    and applied both to the classification of space groups. When using ICSD
    crystals to train the ``parkCNN'' models, dropout was used, while the
    training of the ``parkCNN'' models on synthetic crystals did not use
    dropout.

    Since the approach of using an infinite stream of generated training data
    eliminates the problem of overfitting, we further used deeper models with a
    higher number of model parameters. For this, we used the deep convolutional
    neural networks ResNet-10, ResNet-50, and ResNet-101, which were introduced
    by
    \citeauthor{heDeepResidualLearning2016}\supercite{heDeepResidualLearning2016}
    in 2015.

    Details of the machine learning setup can be found in the SI Section~2.2. Overall, our setup allowed us the training of models over up to 2000 epochs with more than $100\,000$ unique, newly generated crystals and corresponding diffractograms in each epoch \added{(see the upper x-axis of Figure \ref{fig:training_curve})}.

    \subsubsection*{\replaced{Computational experiments}{Experiments}}

    We performed two sets of experiments to evaluate our new dataset split as well as our synthetic crystal generation approach and compare it to state-of-the-art models in literature: Firstly, we trained and tested models on ICSD crystals only, and secondly, we trained on synthetic crystals and tested on ICSD crystals.

    In particular, we first performed an experiment with the ``parkCNN medium''
    model trained directly on the diffractograms simulated from the ICSD
    statistics dataset with a fully random train-test split (similar to
    \supercite{parkClassificationCrystalStructure2017}), instead of splitting by
    the structure type of the crystals. This experiment makes a comparison of
    the two different methods of train-test split possible. We then trained the
    ``parkCNN big'' model using the structure type-based split, again directly
    on ICSD diffractograms. We further repeated the same experiment using the
    smaller model ``parkCNN medium'' to resolve potential overfitting to the
    ICSD diffractograms.

    For the experiments performed on our continuously generated
    dataset based on synthetic crystals, we used the structure type-
    based split. As discussed in Section~\ref{sec:dataset_split}, the training / 
    statistics dataset was only used to extract more general statistics, such as the 
    element distribution.
    First, we trained the ``parkCNN big'' model. For each
    batch, we generated $ 435 $ random crystals and simulated two diffractograms
    with different crystallite sizes for each of them. This resulted in the
    batch size of $ 870 $. 
    Since our synthetic crystal generation algorithm yields an infinite stream
    of unique diffractograms to train on, using much larger models than for the
    fixed ICSD dataset is possible without overfitting. We performed experiments
    for the ResNet-10, ResNet-50, and ResNet-101 models. Instead of generating
    two diffractograms with different uniformly sampled crystallite sizes for each generated crystal (as
    we did for the ``parkCNN big'' model), we now created only one diffractogram
    for each of the $ 145 $ crystals generated for one batch. This is due to 
    the slower training of the ResNet models, which means that reusing the same 
    diffractogram with different crystallite sizes is not necessary to generate
    training data fast enough.

    To obtain the
    highest-possible ICSD test accuracy, we further applied the square root
    function as a preprocessing step to the input diffractograms of the
    network when using the ResNet models. This was suggested by
    \citeauthor{zalogaCrystalSymmetryClassification2020}
    \supercite{zalogaCrystalSymmetryClassification2020} and in their case
    improved classification accuracy by approximately 2 percentage points. Some
    initial tests suggested that this approach also yields a higher accuracy in our
    case, so we used this preprocessing step to train the ResNet
    models.

    While we focused mainly on the methodology of using synthetic crystals to
    extract structural information from powder diffractograms, we also show some
    initial steps toward applying our methods to experimental data. We used the
    publicly available RRUFF mineral database
    \supercite{lafuentePowerDatabasesRRUFF2015} which provides experimental
    measurements, including powder diffractograms \added{(see Figure S11 in the
    SI for an exemplary diffractogram from the RRUFF)}. In order to imitate
    experimental diffractograms, we added Gaussian additive and multiplicative
    noise \added{(similar to
    \supercite{vecseiNeuralNetworkBased2019,szymanskiProbabilisticDeepLearning2021})}
    and a background function based on samples from a Gaussian process to our
    simulated diffractograms. Furthermore, we added a small amount of an
    impurity phase to each diffractogram. Details about the experimental data
    generation protocol can be found in the SI Section~S4\added{, Figure S12
    shows an exemplary synthetic diffractogram with noise, background and an
    impurity phase.} Using the ResNet-50 model, we performed two experiments for
    experimental data, one with the mentioned impurity phase, and one without.

### 2024-04-09 paper/03_our_dataset.tex
Imported structure from Henrik thesis paper; Included author details

    \subsection{Synthetic distribution} \label{sec:synth_distribution}

    \begin{figure}[!htb]
    \centering
    \includegraphics{./figures/example_structures.pdf}
    \caption{a) Some \replaced{randomly chosen and thus representative}{randomly picked} examples of ICSD crystals. b) Some \replaced{randomly chosen and thus representative}{randomly
    picked} examples of synthetically generated crystals. While coordination and
    distances are not chemically correct for the synthetic crystals, crystal
    symmetries are reproduced correctly.}
    \label{fig:example_structures}
    \end{figure}

    \begin{figure}[!htb]
    \centering
    \includegraphics{./figures/histograms/histograms.png} 
    \caption{Histograms comparing the distributions of descriptors of the
    synthetically generated crystals with the ICSD distribution in the test
    dataset. a) density factor $ \frac{V_\text{unit cell}}{\sum_i 4/3 \pi
    \left(\frac{r_{i\text{;cov}} + r_{i\text{;VdW}}}{2}\right)^3} =
    \frac{V_\text{unit cell}}{V_\text{atomic}}$, b) crystallite sizes, c) unit cell
    volume (conventional cell settings), d) number of atoms in the asymmetric
    unit. The probability density of the ICSD is visualized by a stacked bar
    histogram, where the green portion of the bar was correctly classified and
    the red portion was incorrectly classified. The probability density of the
    synthetic crystals is visualized by the dark blue line. The portion between the
    dark blue line and the light blue line was correctly classified, the portion below the
    light blue line was incorrectly classified. The reported classification
    performance is based on the ResNet-101 model trained on diffractograms from
    synthetic crystals.}
    \label{fig:histograms}
    \end{figure}

    We first present an analysis of the generated synthetic crystals.
    Figure~\ref{fig:example_structures} shows some \replaced{randomly chosen and thus representative examples of}{randomly selected} ICSD and
    synthetic crystal structures side-by-side. Visually, the crystals appear very
    similar. However, no physical or chemical considerations regarding stability, clashing atoms, and element combinations are taken into account in the generation of synthetic crystals. As discussed earlier, our goal is to demonstrate that this is
    not problematic when using these crystals for the extraction of structural
    information from powder diffractograms. On the contrary, we expect the synthetic crystals to be a better basis for generalization to fundamentally new crystal structures than existing finite databases.

    To compare the distribution of ICSD crystals with the synthetic
    distribution, we evaluated structural descriptors, i.e. density factors, crystallite sizes, unit cell volumes, and numbers of atoms in the asymmetric unit, and compare their histograms
    in Figure~\ref{fig:histograms}.
    One can see that the overall distributions of
    the synthetic and ICSD crystals are very similar for all four descriptors.
    This shows that our chosen generation algorithm reproduces crystals that are
    similar to ICSD crystals in terms of these more general descriptors.

    \subsection{Classification results} \label{sec:classification_results}

    The main results of our experiments (see Section~\ref{sec:models_and_experiments}) to
    classify the space group of powder diffractograms can be
    found in Table~\ref{tab:results}. \added{In SI Table S2, we further provide the training time and total number of unique diffractograms for each computational experiment.}
    The goal of our experiments is to systematically analyse and quantify the changes in classification accuracy introduced by our two main contributions: A more challenging dataset split, and training on continuously generated synthetic data.

    We started by repeating previously reported
    experiments\supercite{parkClassificationCrystalStructure2017} trained
    directly using ICSD crystals with a random train-test split instead of the
    split based on structure types. This model achieved a very high test
    accuracy of 83.2\%. We note that the previous publication that we compare
    our results to\supercite{parkClassificationCrystalStructure2017} removed
    data from the training dataset, ``[...] heavily duplicated data
    [...]''\supercite{parkClassificationCrystalStructure2017}, but did not
    specify the exact criterions used. In contrast, we did not exclude any
    duplicates in this experiment based on a random train-test split.
    Furthermore, as discussed in Section~\ref{sec:generate_crystals}, we
    excluded crystals with a very high unit cell volume and a very high number
    of atoms in the asymmetric unit. This is likely the reason for the slightly
    higher classification accuracy that we observed, compared to the originally
    reported 81.1\%.

    When splitting randomly, the model merely needs to recognize structures or
    structure types and assign the correct space group. This task is much easier
    than actually extracting the space group using more general patterns. When
    going from random splits to structure type-based splits (see
    Section~\ref{sec:dataset_split}), it becomes obvious that both the ``parkCNN
    big'' as well as the ``parkCNN medium'' models overfit the training data and
    do not generalize well to unseen structure types in the test set \added{(see
    Table \ref{tab:results})}. \added{The ``parkCNN medium'' model, which
    achieved 83.2\% on a random split, now only yields 55.9\% with the structure
    type-based split.}

    \replaced{Training the models by
    \citeauthor{parkClassificationCrystalStructure2017}}{Training the same
    model}, in particular the ``parkCNN big'' model, on synthetic crystals leads
    to a 1.6 percentage points higher test accuracy than the ``parkCNN big''
    model trained on ICSD diffractograms. At the same time, the training
    accuracy drops from \added{the} 87.2\% \added{when we trained the model
    directly on the ICSD} to 74.2\% \added{on the synthetic distribution}
    indicating that the model is now limited more by missing capacity rather
    than by overfitting, which is why we explored larger models, which will be
    discussed later. The gap between training and test accuracy is $31.1$
    percentage points when training on ICSD data, while for training using
    synthetic crystals, the gap is only $16.5$ percentage points. We note that
    this gap between training using synthetic crystals and testing using ICSD
    crystals cannot stem from overfitting, since no diffractograms are repeated
    for the synthetic training. The difference rather stems from the differences
    between the synthetic distribution and the ICSD distribution of crystals.

    \begin{table}[!htb]
        \begin{center}
        \caption{Results of training on diffractograms simulated from ICSD crystals (random splits as well as structure type-based splits) compared to when training on diffractograms from synthetic
        crystals. Test accuracy
        in all cases refers to the accuracy when testing on the ICSD test
        dataset. The training accuracies are averaged over the last 10 epochs of
        the respective run.
        Experiments trained directly on ICSD data
        overfitted to the training data. Training longer would have further
        increased the training accuracy, while not increasing the test
        accuracy.}
        \vspace*{2mm}

        \begin{tabular}{ccccccc} 
            \toprule
            Split & \begin{tabular}{@{}c@{}}Training \\ dataset\end{tabular} & \begin{tabular}{@{}c@{}}Testing \\ dataset\end{tabular} & Model & \begin{tabular}{@{}c@{}}Number of \\ parameters\end{tabular}& \begin{tabular}{@{}c@{}}Training acc. \\ / \%\end{tabular} & \begin{tabular}{@{}c@{}}Test acc. \\ / \%\end{tabular} \\
            \midrule
            Random & ICSD & ICSD & parkCNN medium & 4\,246\,797 & 88.4 & 83.2 \\
            \midrule
            Structure& \multirow{2}{*}{ICSD} & \multirow{2}{*}{ICSD}& parkCNN big & 4\,959\,585 & 87.2 & 56.1 \\
            type&&& parkCNN medium & 4\,246\,797 & 90.9 & 55.9 \\
            \midrule
            & \multirow{4}{*}{synthetic} & \multirow{4}{*}{ICSD} &parkCNN big & 4\,959\,585 & 74.2 & 57.7 \\
            Structure&&& ResNet-10 & 9\,395\,025 & 87.2 & 73.4 \\
            type$^{1}$ &&& ResNet-50 & 41\,362\,385 & 91.8 & 79.3 \\
            &&& ResNet-101 & 60\,354\,513 & 92.2 & \textbf{79.9} \\
            \bottomrule
        \end{tabular}\label{tab:results}
        \end{center} 
        \footnotesize{$^{1}$Here, the split type refers to the statistics and the test dataset, rather than the training and the test dataset.}
    \end{table}

    While the ``parkCNN big'' model trained on synthetic crystals outperforms
    the approach of training directly on ICSD crystals by only 1.6 percentage
    points, the advantage of training on an infinite stream of synthetic data increases when using models with more parameters and thus higher capacity. In
    contrast to training directly on a finite set of ICSD crystals, it is possible to train
    very large models using the infinite synthetic data stream without the
    potential of overfitting. As found in the last lines of Table~\ref{tab:results}, ResNet-10, ResNet-50, and ResNet-101 based models achieve ICSD test accuracies of 73.4\%,
    79.3\%, and 79.9\%. This is a significant increase from the 57.7\% achieved
    by the ``parkCNN big'' model. Figure~S4 in the SI further shows the top-$k$ accuracy over $k$ for the ResNet-101 model. With increasing $k$ the accuracy exceeds 95\% at $k=5$.
    This means that our model can not only determine the correct space group with a high probability but can also generate an almost complete list of possible space group candidates.

    Figure~\ref{fig:training_curve} shows the ICSD test accuracy, the training
    accuracy (on synthetic data), and the ICSD top-5 test accuracy for all three ResNet variants
    as a function of epochs trained. For all three metrics, the difference
    between ResNet-50 and ResNet-101 is comparably small, while the step from
    ResNet-10 to ResNet-50 is substantial (5.9 percentage points in ICSD test
    accuracy, see Table~\ref{tab:results}). This shows that going beyond the
    model size of the ResNet-101 will likely not yield a big improvement in
    accuracy. In contrast to the 79.9\% accuracy reached in the top-1 ICSD test accuracy, the top-5 ICSD test accuracy of the ResNet-101 model reaches
    96\%. However, for all three ResNet variants, a gap between training using synthetic
    crystals and testing using the ICSD remains (12.3 percentage points for
    ResNet-101). As also shown in Figure~S5 in the SI, the accuracy convergence can be approximately described by a power law, indicating that exponentially more training time will substantially reduce classification errors and thus potentially lead to top-1 accuracies of 90\% and above, at the cost of a 100-fold increase in training times\replaced{. Considering the current training times provided in Table S2 of the SI, this is currently infeasible or only possible with tremendous hardware resources.}{, which is currently infeasible.}

    \begin{figure}[!htbp] 
    \centering
    \includegraphics{figures/training_curves/training_curve_main.png}
    \caption{Test accuracy (ICSD), training accuracy (synthetic crystals), and
    test top-5 accuracy (ICSD) as a function of epochs\added{ (bottom axis). Since each additional epoch contains newly generated unique diffractograms, we further show the accuracies as a function of the total number of unique synthetic diffractograms (top axis)}. We show all three
    metrics for the models ResNet-101, ResNet-50, and ResNet-10. To better show
    the scaling behavior, both axes use logarithmic scaling. \added{Figure S6
    shows the same plot but without logarithmic scaling.} To better see the
    exponential behaviour, see Figure~S5 in the SI. }
    \label{fig:training_curve} 
    \end{figure}

    The histograms in Figure~\ref{fig:histograms} show, next to the overall
    distribution, also the fraction of diffractograms classified wrongly for
    testing on the ICSD (red bar) and on the synthetic data (below the light blue line)
    for the ResNet-101 model. First, one can see that throughout almost all
    regions of the distributions, the accuracy on the synthetic data is slightly
    higher than that on the ICSD. This is related to the aforementioned gap of 12.3
    percentage points between train and test accuracy and can be attributed to
    differences between the synthetic and ICSD distribution of crystals. This
    will be discussed in detail in the next section. It is surprising to see that the
    dependence on crystallite sizes is rather weak, as smaller crystallite sizes
    result in broader peaks (see Scherrer equation, Eq.~\ref{eq:scherrer}), potentially making the
    classification harder due to more peak overlaps.

    In summary, the maximum ICSD test accuracy of 79.9\% that we achieved using
    the ResNet-101 model almost reaches the previously
    reported\supercite{parkClassificationCrystalStructure2017} 81.14\% for the
    space group classification. However, our accuracy is based on a train-test
    split based on structure types, in contrast to a random split. This creates
    a much harder but also realistic task to solve since the model needs to
    generalize to other structure types without merely recognizing
    diffractograms or structure types that it has already seen during training.
    This becomes especially apparent from our experiment directly trained on
    diffractograms from ICSD crystals with the split based on structure types,
    which reached only 56.1\% instead of the previously
    reported\supercite{parkClassificationCrystalStructure2017} 81.14\%.

    \subsubsection*{Experimental results}
    To go beyond simulated diffractograms, we trained ResNet-50 models on
    calculated diffractograms with background, noise, and impurities and applied
    the trained models to the RRUFF mineral database. Our results (see Figure~S3
    in the SI) show that it is essential to include impurity phases in the
    training data. By doing so, we obtain a top-1 accuracy of 25.2\% and a
    top-10 accuracy of over 60\%. This is of high practical relevance since
    having a short list of potential space groups is often sufficient as a first
    step to further refinement and analysis.

    \citeauthor{vecseiNeuralNetworkBased2019} performed similar experiments of
    space group classification on the same database. Using an ensemble of 10
    fully connected neural networks, they reached a classification accuracy of
    54\%\supercite{vecseiNeuralNetworkBased2019}. While our obtained accuracy is
    significantly lower, our approach is much more general: In contrast to our
    approach, the training dataset was based on simulated diffractograms of
    structures of the ICSD\supercite{vecseiNeuralNetworkBased2019}, which
    contains almost all RRUFF structures, leading to high similarities of
    training and \replaced{test}{testing} data. Therefore, the model needed to
    simply recognize the minerals, instead of directly inferring the space group
    using the symmetry elements - as our method needs to do.

    We want to emphasize that our efforts to apply the methodology to
    experimental data are only preliminary. We expect improved results with an
    improved data generation protocol since the procedure contains many
    parameters to be tuned. Ideally, one would use a generative machine learning
    approach to add the experimental effects (noise, background, impurities) to
    the pure diffractograms. We also want to point out that the noise level and
    quality of data in the RRUFF dataset are limited. Application of the
    presented methodology \replaced{to}{on} other experimental datasets is
    desirable. \deleted{We will address these points in future work, where we
    focus on improved ways of modeling experimental imperfections.}
    \added{As discussed above, for the
    classification of pure diffractograms we observed the ResNet-50 to have the
    best cost-benefit ratio, since the ResNet-101 yielded only slight
    improvements. For the more complicated problem of classifying diffractograms
    with experimental imperfections, bigger models and longer training times
    might be necessary.}

    \added{Next to improving the modeling of experimental imperfections and therefore the overall accuracy on experimental data, the practical application of deep neural networks for analyzing powder diffractograms yields further challenges that we want to discuss. Since experimental setups differ, e.g., concerning the used wavelength, a different $2\theta$ step size, or a different $2\theta$ range, a new neural network would need to be trained for each situation. Since our largest model requires a significant computational investment, this might not be feasible in all situations. Arguably, though, for large high-throughput experiments, the 11-day training of a ResNet-50 should not be unreasonable, especially if it can speed up the data analysis significantly and allow in-loop adaptive experimentation. For smaller setups, where this is not feasible, other solutions must be found. First, one can use a form of transfer learning from a pre-trained model to fine-tune to the desired experimental setup. This, however, would only work for a change in wavelength, since a change in step size or $2\theta$ range would change the input dimensions of the network. However, to handle a change in the $2\theta$ range, it might be possible to include a form of zero-masking in the synthetic training data, such that different input ranges (with zeros where no measurement was made) can be used, which would lead to a more flexible model, not requiring new training data when applied to a new $2\theta$ range. For a change in the step size, a cubic spline interpolation might be helpful. We plan to address these challenges in future work.}

    \deleted{As discussed above, for the classification of pure diffractograms we
    observed the ResNet-50 to have the best cost-benefit ratio, since the
    ResNet-101 yielded only slight improvements. For the more complicated
    problem of classifying diffractograms with experimental imperfections,
    bigger models and longer training times might be necessary.}

    \added{Furthermore, analysis of the loss value or gradient norm associated with particular samples, i.e. crystal structures, during training on synthesis crystals or during transfer learning from synthetic to experimental data can help to better understand the relevance and informativeness of given samples for the model. This can help in generating more relevant synthetic data based on experimental crystal structures that are underrepresented in the synthetic data distribution.}

    \subsection{Differences between synthetic crystals and ICSD crystals}
    We showed that training directly on crystals from the ICSD yields a gap
    between the training and test accuracy due to overfitting. The training on
    the synthetic dataset also shows a gap between the training and test
    accuracy (see Table~\ref{tab:results}), but it is smaller than when training
    directly on ICSD crystals. Furthermore, this gap is not due to overfitting,
    since overfitting to singular diffractograms is not possible when the model
    is trained using an infinite stream of generated synthetic crystals. The gap
    rather stems from systematic differences between the synthetic and ICSD
    distribution of crystals.

    To analyze those differences, we created three modifications of the ICSD
    test dataset \added{(see SI Section~S3 for details)}. In the first modification, the fractional coordinates of the
    atoms in the asymmetric unit of the crystals of the ICSD test dataset were
    randomly uniformly resampled (as in the synthetic crystal generation
    algorithm). In the second modification, the lattice parameters were
    randomized following the \replaced{kernel density estimate}{KDE} used in the synthetic generation algorithm. The
    third modification combines both previous modifications, i.e. both the
    coordinates and the lattice parameters were resampled. These three modified
    test datasets bring the ICSD test dataset closer to the distribution used
    for training and let us quantify which factors contribute to the gap between
    training \added{on synthetic crystals} and testing \added{on the ICSD}.

    We evaluated the test accuracies on \replaced{the randomized}{these}
    datasets for the experiment using the ResNet-101 model trained using
    synthetic crystals. We found that randomizing the coordinates yields an
    increase in test accuracy of \replaced{$4.89$}{$5.96$} percentage points.
    Randomizing the lattice parameters results in an increase of
    \replaced{$0.79$}{$1.31$} percentage points. Randomizing both the
    coordinates and the lattice parameters leads to an increase of
    \replaced{$5.70$}{$6.32$} percentage points, explaining
    \replaced{almost}{approximately} half of the gap of 12.3 percentage points
    between synthetic training and ICSD test accuracy.

    \added{So far, we have randomized the lattice parameters and coordinates of the test dataset, such that they follow a distribution that is based on the statistics extracted from the statistics dataset. However, this does not take into account the different Wyckoff position occupation probabilities between the test and statistics datasets. For this, we repeated
    a similar analysis, for which we applied the
    randomizations to the statistics dataset rather than the test dataset.
    Without any modifications, testing on the statistics dataset instead of the test dataset yielded $3.89$ percentage points higher accuracy. This can be explained by slight differences in the overall statistics between the test and statistics datasets.
    Randomizing the coordinates yields a further increase of $4.72$
    percentage points, randomizing the lattice $1.16$ percentage points, and
    randomizing both the coordinates and the lattice parameters $6.68$ percentage points.
    In total, testing on the statistics dataset with randomized coordinates and
    lattice parameters yields a $10.57$ percentage points higher accuracy than on the
    unmodified test dataset. This almost completely explains the gap
    of $12.3$ percentage points between the training accuracy on synthetic crystals and the test accuracy on the ICSD. The
    remaining part is likely due to our algorithm that places atoms on Wyckoff
    positions 
    not reproducing the ICSD distribution exactly. 
    However, the
    remaining difference is remarkably small.}

    \deleted{We also tested the ResNet-101 model on diffractograms simulated
    based on the statistics dataset from which the statistics for our generation
    algorithm have been extracted. This returns a $4.33$ percentage points
    higher accuracy than on the test dataset. This is due to the different
    structure types present in the test and statistics dataset yielding
    different occupation probabilities (potentially also 0) for Wyckoff
    positions and overall slightly different statistics for the synthetic
    generation algorithm.}
    \deleted{To summarize, the contributions to the gap stem from differences in the
    distribution of lattice parameters and coordinates with $6.32$ percentage
    points and from the differences between the statistics dataset and test
    dataset with $4.33$ percentage points. Added together (assuming the effects of geometry and Wyckoff position occupation probabilities are unrelated and thus additive), the contributions
    almost completely explain the gap of $12.3$ percentage points between
    training and test accuracy. The remaining part is likely due to our
    algorithm that places atoms on Wyckoff positions not reproducing the ICSD
    distribution exactly. However, the remaining difference is remarkably small.}

    \begin{figure}[!htb]
    \centering
    \includegraphics{./figures/histograms/histograms_rel.png}
    \caption{Classification \replaced{error}{performance} for each bin of a) the unit cell volume
    (conventional cell settings) and b) the number of atoms in the asymmetric
    unit. \deleted{The classification performance when testing on the ICSD is shown using
    a stacked bar plot, where the red bar indicates the classification error and
    the green bar indicates the correctly classified portion of each bin. When
    testing on the synthetic distribution, the classification error is given by
    the light blue line, the rest (between dark blue and light blue line) was correctly
    classified.} The reported classification performance is based on the
    ResNet-101 model trained on diffractograms from synthetic crystals. This
    visualization clearly shows the error rate within each bin, in contrast to
    Figure~\ref{fig:histograms}, which additionally includes the relative
    proportion of the crystals of the respective bin to the total amount of
    crystals.}
    \label{fig:histograms_rel}
    \end{figure}

    In Figure~\ref{fig:histograms_rel} we show the test classification error in
    each bin for the unit cell volume and the number of atoms in the asymmetric
    unit using the ResNet-101 model trained on diffractograms of synthetic
    crystals. The classification error is shown both for testing on
    diffractograms from synthetic crystals and on ICSD diffractograms. One can
    see that for small volumes and a small number of atoms in the asymmetric
    unit, the difference between classifying ICSD diffractograms and
    diffractograms from synthetic crystals is relatively small. As the volume
    and number of atoms in the asymmetric unit increase, the gap between the two
    errors increases, too. We already identified the uniformly sampled atom
    coordinates in the synthetic distribution as the main contributor to the gap
    in accuracy between the synthetic crystals and ICSD crystals. Therefore, it
    seems that the uniform sampling of atom coordinates works well for small 
    number of atoms in the asymmetric unit and small volumes, while the error 
    due to this sampling strategy increases slightly for higher volumes and higher
    number of atoms in the asymmetric unit.

    When looking at the distribution of crystals in the ICSD, the number of
    atoms in the asymmetric unit tends to be larger for lower-symmetry space
    groups (for example, in the triclinic crystal system) than for
    higher-symmetry space groups such as those from the cubic crystal system.
    Therefore, the increasing test error on diffractograms from ICSD crystals
    with a higher number of atoms in the asymmetric unit is especially relevant
    for these lower-symmetry space groups.
    It might be possible that a different scheme of generating atom positions in
    the unit cell (compared to the independent uniform sampling that we used)
    works better for a high number of atoms in the asymmetric unit.

    Overall, it is important to note that the distribution of ICSD crystals is (apart from a few Wyckoff position occupation probabilities which are exactly zero in the statistics dataset\footnote{Setting them to small non-zero values typically leads to the generation of rather large unit cells, as the general Wyckoff positions have high multiplicities.}) almost completely encompassed by the much larger distribution of snynthetic crystals that we used for training.
    However, due to finite training times and model capacity, a performance gap remains. This gap can be improved by using (substantially) more computing power or by narrowing the very general synthetic distribution, e.g., by using a different algorithm to generate atom positions.
    This indicates an inherent challenge in XRD classification but more generally in materials property prediction:
    Machine learning models are ultimately trained to be employed in real-world tasks, which are typically related to novel, i.e. yet unseen materials and structures.
    At the same time, the machine learning models are tested based on an IID assumption, i.e. the assumption that the distribution of training and testing data is the same.
    While not being a contradiction in the limit of infinite training data and model capacity, this becomes an (unsolvable) challenge in reality, when facing finite datasets and models.
    In our case, our model trained on a large distribution of synthetic crystal structures will likely generalize better to completely new crystal structures different from any crystal structure contained in the ICSD database.
    At the same time, it suffers from smaller ICSD test set errors, even though the ICSD distribution is contained in the synthetic data generation distribution.

### 2024-04-09 paper/04_contributions.tex
Imported structure from Henrik thesis paper; Included author details

We developed an algorithm based on the symmetry operations of the space groups
to generate synthetic crystals that follow the distribution found in the ICSD
database in terms of general descriptors like volume, density, or types of
elements. The generated crystals have randomly sampled coordinates and span a
wide range of structure types, many of which do not appear in the ICSD. We
showed that, compared to using ICSD crystals directly, simulating the training
data based on the synthetic crystals can improve the performance of tasks that
extract structural information from powder diffractograms, in this case, the
space group. \replaced{The more general dataset that also contains unstable structures helps to classify unseen stable crystal structures.}{Our results verify our hypothesis that crystals forming the
training dataset do not need to be stable or chemically viable.}

We trained on an infinite on-the-fly generated stream of synthetic crystals and simulated
batches of diffractograms using a distributed framework based on the
\emph{Python} library \emph{Ray}\supercite{moritzRayDistributedFramework2018a}.
This allows the training of very large networks without overfitting. The
best-performing model (ResNet-101) reached a space group classification accuracy
of 79.9\% vs. 56.1\% when training on ICSD structures directly. By performing
the train-test split using the structure type, we forced our models to not just
recognize structure types or individual structures, but to actually learn rules
to distinguish different space groups by their symmetry elements. This shows the
true generalization capabilities to new structure types and novel classes of
materials. We also demonstrated first steps toward applying the presented methodology to an experimental dataset. We expect further improvements in this area using improved models of experimental imperfections, as well as larger ML models and longer training times.

Even though models trained on the synthetic distribution transfer well when
tested on ICSD crystals, we found a gap of 12.3 percentage points (ResNet-101)
between the training accuracy on synthetic crystals and test accuracy on the
ICSD. We showed that the main contribution to this gap stems from the
independently uniformly sampled atom coordinates. An improved approach may be
needed to artificially generate more ordered structures, which contain more
ordered diffraction planes than a cloud of uniformly sampled points. This might
be especially important for crystals with a high number of atoms in the
asymmetric unit.

Lastly, the developed algorithm to synthetically generate crystals can be used
for other XRD-related tasks in the future, such as the extraction of crystallite sizes, lattice
parameters, information about the occupation of Wyckoff positions, etc.
Furthermore, instead of generating \replaced{synthetic}{completely random} crystals of all space
groups, one can also generate \deleted{random} crystals of given structure types to solve
more specialized tasks. This would allow the use of very large models for tasks
that are typically strongly limited by the dataset size when using only the entries of
the ICSD. Also, tasks concerning multi-phase diffractograms or augmentations such as
strain in given crystal structures can benefit from our batch-wise online
learning approach.

### 2024-04-09 paper/_main.tex
Imported structure from Henrik thesis paper; Included author details

\documentclass[a4paper]{article}

\pdfoutput=1
\usepackage[utf8]{inputenc}

%\usepackage[
%backend=biber,
%%style=numeric,
%style=nature,
%sorting=none,
%doi=true,
%maxbibnames=99,
%date=iso,
%seconds=true,
%urldate=iso
%%citestyle=numeric-comp
%]{biblatex}

\usepackage[super,sort&compress,comma]{natbib} 
\newcommand{\supercite}[1]{\cite{#1}}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{csquotes}

%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{colorlinks=true, allcolors=blue}
\usepackage{authblk}
\usepackage{multirow}
\usepackage{nomencl}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{textcomp} 
\usepackage{amsfonts}
\usepackage{outlines} %%same
\usepackage{subfig}
\usepackage[version=4]{mhchem}
\usepackage{siunitx}
\usepackage{bm}
\usepackage[normalem]{ulem} %%temporary, for organization purposes

\date{}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\DeclareSIUnit\angstrom{\text {Å}}

\usepackage[final]{changes}
%\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}
\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}

\interfootnotelinepenalty=10000
\title{Towards a high quality public powder X-ray diffraction dataset}

\author[1,2]{Daniel Hollarek}
\author[1,2]{Henrik Schopmans}
\author[1,2]{Pasc1al Friederich}
\author[2]{Ben Breitung}
\author[2]{Simon Schweidler}
\author[3]{Andrea Hodge}
\author[3]{Adie Alwen}
\author[4]{Mriganka Singh}
\author[4]{Carolin Sutter-Fella}
\author[4]{Tim Kodalle}
\author[4]{Maged Abdelsamie}

% I made the affiliations smaller because there's a lot of them
\renewcommand\Affilfont{\small}

\affil[1]{Institute of Theoretical Informatics, Karlsruhe Institute of Technology, Engler-Bunte-Ring 8, 76131 Karlsruhe, Germany}
\affil[2]{Institute of Nanotechnology, Karlsruhe Institute of Technology, Hermann-von-Helmholtz-Platz 1, 76344 Eggenstein-Leopoldshafen, Germany}
\affil[3]{Department of Aerospace and Mechanical Engineering, University of Southern California, 3650 McClintock Ave, Los Angeles, CA 90089, USA}
\affil[4]{Department of Chemical Engineering and Materials Science, University of Southern California, 925 Bloom Walk, HED 216, Los Angeles, CA 90089, USA}
\affil[5]{Lawrence Berkeley National Laboratory, Molecular Foundry Division, 1 Cyclotron Rd., Berkeley, 94720 CA, USA}

\begin{document}

\maketitle
\begin{abstract}
    \input{00_abstract.tex}
\end{abstract}

\section{Motivation}
\input{paper/01_motivation}

\section{Existing datasets}
\input{paper/02_existing_datasets}

\section{Our dataset}
\input{paper/03_our_dataset}

\section{Contributions}
\input{paper/04_survey}

\subsection*{Acknowledgements}
\input{07_acknowledgements.tex}

\printnomenclature

\clearpage

%\listofchanges
%\clearpage

%\nocite{*}
%\TheBibliography
%\printbibliography[heading=bibintoc]

\bibliography{bib}
\bibliographystyle{rsc} %the RSC's .bst file

\end{document}

### 2024-04-09 paper/00_abstract.tex
Updated abstract

@@
-We, the AiMat group at the Karlsruhe Institute of Technology (KIT), are collecting a dataset of both labeled and unlabeled experimental powder X-ray diffraction (pXRD) data, which will be made publicly available once sufficient data has been collected.
-We look forward to publishing a paper surveying the database and discussing its uses, and would be happy to include anyone contributing data as co-authors.
-You can view the current draft here: State-of-the-database
+% Autonomous laboratory
+% https://www.nature.com/articles/s41586-023-06734-w
 
-The main motivation for collecting and publishing this dataset is to assist researchers working on machine learning based approaches for pXRD analysis.
-Any data, even unlabeled data, can provide clues how simulations can be made to better reflect real-world experiments and labeled data can be used to test the developed algorithms.
-We hope to provide both types of data in a wide variety and large quantity.
+% Accelerating material discovery
+% https://www.nature.com/articles/s41524-022-00765-z
 
-While we do not collect single crystal data, we are interested in any kind of polycrystalline samples, including thin films.
-If your samples exhibit preferred crystallite orientation, please make mention of this in the description field of the upload form.
-So far, we have received upwards of 4155 unique powder X-ray diffraction patterns from institutions including:
+%High-throughput experiments facilitate materials innovation: A review
+%https://link.springer.com/article/10.1007/s11431-018-9369-9#:~:text=High%2Dthroughput%20materials%20experiments%20including,at%20a%20fraction%20of%20cost.
+
+%The advent of high-throughput experiments, holds the prospect of accelerating the speed materials discovery multiple times over.
+The synthesis and probing of novel materials is becoming increasingly efficient and automated, which has correspondingly increased the throughput of samples in these steps of the experimentation pipeline.
+
+% You need a good initial guess, it can only optimize to the nearest minimum from the initial guess
+
+% Comments -> Why having to do rietveld refinement sucks
+% FullProf tutorial on fitting XRD peaks -> Tedious, requires lots of manual input
+% https://pranabdas.github.io/research/fullprof/
+
+% Topas tutorial -> 44 page tutorial, many boring and potentially frustrating steps, lots of points offer potential for task failure
+% https://topas.webspace.durham.ac.uk/topas_user_menu/
+
+%  Methods and Tutorials – Powder Diffraction -> Overview of rietveld refinement data
+%https://neutrons.ornl.gov/sites/default/files/Methods%20and%20Tutorials%20Powder%20Diffraction.pdf
+
+% Rietveld Refinement from Powder Diffraction Data -> 50 pg tutorial that no one wants to read, overwhelming amout of quantities/figures that needs to be manually looked at in the process
+% https://www.fkf.mpg.de/4112052/cpd26.pdf
+
+In powder x-ray diffraction, the analysis of the diffractograms still presents a significant bottleneck in high-throughput experimentation that remains to be solved, in order to match the increased throughput of the prior stages of experimentation.
+While Rietveld refinement can identify the phases and extract the structural information found in the diffractograms, each sample requires manual inputs and careful configuration of the analysis software from trained staff, which is infeasible for highly automated high-throughput experiments.
+
+In the current landscape of powder x-ray diffraction analysis, machine learning has emerged as a very promising direction of research resolving this bottleneck by enabling fully automated powder diffraction analysis.
+A noteable difficulty in applying machine learning to this domain is the lack of labeled datasets i.e.
+raw experimental diffraction patterns with accompanying structure solutions, which has relegated machine learning researches to training on simulated data.
+Since these simulations largely fail to accurately reflect experiment, the performance of models trained on this simulated data, also mostly fails to transfer to experiment and provide value in practice.
+We hope to remedy this by providing a freely available and easily accessible dataset of powder diffraction data, which provides machine learning researches with a large quantity of real experimental diffractograms collected from a broad range of samples.
+We hope that this can guide machine learning research towards more realistic simulations and potentially enable training and testing on real experimental data.
+
+% mention below:
+% uniform, easily readable
 

### 2024-04-09 paper/_main.tex
Updated abstract

@@
 \usepackage[utf8]{inputenc}
-
-%\usepackage[
-%backend=biber,
-%%style=numeric,
-%style=nature,
-%sorting=none,
-%doi=true,
-%maxbibnames=99,
-%date=iso,
-%seconds=true,
-%urldate=iso
-%%citestyle=numeric-comp
-%]{biblatex}
-
-\usepackage[super,sort&compress,comma]{natbib} 
-\newcommand{\supercite}[1]{\cite{#1}}
+\usepackage[super,sort\&compress,comma]{natbib}
+\renewcommand{\supercite}[1]{\cite{#1}}
 
@@
 \DeclareSIUnit\angstrom{\text {Å}}
-
 \usepackage[final]{changes}
-%\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}
-\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}
 
 \interfootnotelinepenalty=10000
-\title{Towards a high quality public powder X-ray diffraction dataset}
+\title{Towards a free high quality powder X-ray diffraction dataset}
 
@@
 \author[3]{Adie Alwen}
+\author[4,5]{Carolin Sutter-Fella}
 \author[4]{Mriganka Singh}
-\author[4]{Carolin Sutter-Fella}
 \author[4]{Tim Kodalle}
@@
 
-\affil[1]{Institute of Theoretical Informatics, Karlsruhe Institute of Technology, Engler-Bunte-Ring 8, 76131 Karlsruhe, Germany}
-\affil[2]{Institute of Nanotechnology, Karlsruhe Institute of Technology, Hermann-von-Helmholtz-Platz 1, 76344 Eggenstein-Leopoldshafen, Germany}
-\affil[3]{Department of Aerospace and Mechanical Engineering, University of Southern California, 3650 McClintock Ave, Los Angeles, CA 90089, USA}
-\affil[4]{Department of Chemical Engineering and Materials Science, University of Southern California, 925 Bloom Walk, HED 216, Los Angeles, CA 90089, USA}
-\affil[5]{Lawrence Berkeley National Laboratory, Molecular Foundry Division, 1 Cyclotron Rd., Berkeley, 94720 CA, USA}
+\affil[1]{Institute of Theoretical Informatics, Karlsruhe Institute of Technology,\protect \\
+Engler-Bunte-Ring 8, 76131 Karlsruhe, Germany}
+\affil[2]{Institute of Nanotechnology, Karlsruhe Institute of Technology,\protect \\ Hermann-von-Helmholtz-Platz 1, 76344 Eggenstein-Leopoldshafen, Germany}
+\affil[3]{Department of Aerospace and Mechanical Engineering, University of Southern California, \protect \\3650 McClintock Ave, Los Angeles, CA 90089, USA}
+\affil[4]{Department of Chemical Engineering and Materials Science, University of Southern California,\protect \\ 925 Bloom Walk, HED 216, Los Angeles, CA 90089, USA}
+\affil[5]{Lawrence Berkeley National Laboratory, Molecular Foundry Division,\protect \\ 1 Cyclotron Rd., Berkeley, 94720 CA, USA}
 
@@
 \begin{abstract}
-    \input{00_abstract.tex}
+\input{00_abstract}
 \end{abstract}
 
-\section{Motivation}
-\input{paper/01_motivation}
+\section{Motivation}\label{sec:motivation}
+\input{01_motivation}
 
-\section{Existing datasets}
-\input{paper/02_existing_datasets}
+\section{Existing datasets}\label{sec:existing_datasets}
+\input{02_existing_datasets}
 
-\section{Our dataset}
-\input{paper/03_our_dataset}
+\section{Our dataset}\label{sec:our_dataset}
+\input{03_our_dataset}
 
-\section{Contributions}
-\input{paper/04_survey}
-
-\subsection*{Acknowledgements}
-\input{07_acknowledgements.tex}
+\section{Contributions}\label{sec:contributions}
+\input{04_contributions}
 

### 2024-04-10 paper/00_abstract.tex
Added rietveld refinement section of motivation

--- 2024-04-09 paper/00_abstract.tex
@@
-We, the AiMat group at the Karlsruhe Institute of Technology (KIT), are collecting a dataset of both labeled and unlabeled experimental powder X-ray diffraction (pXRD) data, which will be made publicly available once sufficient data has been collected.
-We look forward to publishing a paper surveying the database and discussing its uses, and would be happy to include anyone contributing data as co-authors.
-You can view the current draft here: State-of-the-database
+In powder x-ray diffraction, the analysis of the diffractograms still presents a significant bottleneck
+in high-throughput experimentation that remains to be solved, in order to match the increased throughput
+of the prior stages of experimentation.
+While Rietveld refinement can identify the phases and extract the structural information found
+in the diffractograms, each sample requires manual inputs and careful configuration of the analysis software
+from trained staff, which is infeasible for highly automated high-throughput experiments.
 
-The main motivation for collecting and publishing this dataset is to assist researchers working on machine learning based approaches for pXRD analysis.
-Any data, even unlabeled data, can provide clues how simulations can be made to better reflect real-world experiments and labeled data can be used to test the developed algorithms.
-We hope to provide both types of data in a wide variety and large quantity.
+In the current landscape of powder x-ray diffraction analysis, machine learning has emerged as a very promising direction of research resolving this bottleneck by enabling fully automated powder diffraction analysis.
+A noteable difficulty in applying machine learning to this domain is the lack of labeled datasets i.e.\
+raw experimental diffraction patterns with accompanying structure solutions, which has relegated machine learning
+researches to training on simulated data.
+Since these simulations largely fail to accurately reflect experiment, the performance of models trained
+on this simulated data, also by in large fail to transfer to experiment and provide value in practice.
+We aim to remedy this by providing a freely available and easily accessible dataset of powder diffraction data
+, providing machine learning researches with a large quantity of real experimental diffractograms collected
+from a broad range of samples.
+We hope that this can guide machine learning research towards more realistic
+simulations and potentially enable training and testing models on real experimental data.
 
-While we do not collect single crystal data, we are interested in any kind of polycrystalline samples, including thin films.
-If your samples exhibit preferred crystallite orientation, please make mention of this in the description field of the upload form.
-So far, we have received upwards of 4155 unique powder X-ray diffraction patterns from institutions including:
+% mention below:
+% uniform, easily readable
 

### 2024-04-10 paper/01_motivation.tex
Updated abstract
Added rietveld refinement section of motivation

@@
+%but it requires a good initial guess of the structure possibly combined with manual adjustments. Both \textbf{}
+
+% Autonomous laboratory
+% https://www.nature.com/articles/s41586-023-06734-w
+
+% Accelerating material discovery
+% https://www.nature.com/articles/s41524-022-00765-z
+
+%High-throughput experiments facilitate materials innovation: A review
+%https://link.springer.com/article/10.1007/s11431-018-9369-9#:~:text=High%2Dthroughput%20materials%20experiments%20including,at%20a%20fraction%20of%20cost.
+
+The advent of high-throughput experiments, holds the prospect of accelerating the speed materials discovery multiple times over.
+The synthesis and probing of novel materials is becoming increasingly efficient and automated,
+which has correspondingly increased the throughput of samples in these steps
+of the experimentation pipeline.
+
+% Rietveld Refinement (3.9 Profile fitting and rietveld analysis)
+%https://www.sciencedirect.com/topics/biochemistry-genetics-and-molecular-biology/rietveld-refinement
+
+% Rietveld Refinement: Half a century anniversary
+% https://pubs.acs.org/doi/10.1021/acs.cgd.1c00854
+
+% On Rietveld refinement
+The objective of powder x-ray diffraction experiments is to extract as much information about the crystal structure
+of the sample as possible, or if the sample is multi-phase, also identify the phases present in the sample and quantify
+their fractions.
+This is done through a technique called Rietveld refinement: A working model of the crystal structure, starting
+from an initial guess provided by the operator of the analysis software, is assumed and the corresponding diffractogram
+is simulated.
+The parameters of the working model of the sample are then adjusted to minimize the difference between the simulated
+diffraction pattern and the one observed in experiment.
+This workflow often works well, since usually the operator of the experiment will be able
+to provide an at least somewhat accurate initial guess of the crystal structure.
+
+However, employing manually operated Rietveld Refinement is not without issues and it is certainly not
+scalable to the degree that would be required to keep up with advances
+in throughput and efficiency in other steps of the experimentation pipeline, if just due to the manual effort involved.
+
+% Reference library required and analysis software required, reference library may be incomplete
+% software training required, doesnt always converge or not to true minimum,
+% sometimes sample is truely unknown,  manual labor required in any case
+
+% -> This is too detailed for this paper, but can use it for thesis
+%For one, under normal circumstances, the initial guess of the structure requires manually or computationally consulting
+%a reference library of pairs of patterns and structures.
+%The reference library is an undesirable overhead that needs to be distributed along with the analysis software and
+%maintained to remain up to date and interoperable with the anlysis software.
+%The industry standard in terms of such reference libraries is the Powder Diffraction File and costs around 1000\$ per
+%year to license from its maintainer, the International Centre for Diffraction Data.
+%Another issue is that since the number of chemically viable crystal structures is infinite any reference library, no
+%matter how
+
+% Crystallographic open database
+% https://www.crystallography.net/cod/
+
 Machine learning techniques have emerged as a powerful tool in the toolkit of
@@
 overfitting, and allows very large models to be trained.
+
+% PDF pricing
+%https://www.icdd.com/assets/files/2023-Purchase-Options.pdf
+
+% Computationally intensive -> Probably outdated. If it used to take minutes in 2010
+% it probably only takes seconds by now.
+% https://www.rigaku.com/de/node/707
+
+% A simple solution to the Rietveld refinement recipe problem
+% -> Rietveld refinement requires "expert input"
+% https://www.ncbi.nlm.nih.gov/pmc/articles/PMC10840306/
+
+% Comments -> Why having to do rietveld refinement sucks
+% FullProf tutorial on fitting XRD peaks -> Tedious, requires lots of manual input
+% https://pranabdas.github.io/research/fullprof/
+
+% Topas tutorial -> 44 page tutorial, many boring and potentially frustrating steps, lots of points offer potential for task failure
+% https://topas.webspace.durham.ac.uk/topas_user_menu/
+
+%  Methods and Tutorials – Powder Diffraction -> Overview of rietveld refinement data
+%https://neutrons.ornl.gov/sites/default/files/Methods%20and%20Tutorials%20Powder%20Diffraction.pdf
+
+% Rietveld Refinement from Powder Diffraction Data -> 50 pg tutorial that no one wants to read, overwhelming amout of quantities/figures that needs to be manually looked at in the process
+% https://www.fkf.mpg.de/4112052/cpd26.pdf

### 2024-04-10 paper/01_motivation.tex
Highlighted reference library problem

@@
 diffraction pattern and the one observed in experiment.
-This workflow often works well, since usually the operator of the experiment will be able
+Critically, this process does not generally guarantee to guide the working model towards a global minimum of the difference
+in diffractograms and hence arrive at an accurate model of the sample.
+In practice the refinement process may only guide the working model towards the nearest local minimum
+starting from the initial guess.
+Hence, the result of the refinement procedure is in general only as good as the initial guess it was provided with.
+Still this workflow often works well, since usually the operator of the experiment will be able
 to provide an at least somewhat accurate initial guess of the crystal structure.
 
-However, employing manually operated Rietveld Refinement is not without issues and it is certainly not
-scalable to the degree that would be required to keep up with advances
-in throughput and efficiency in other steps of the experimentation pipeline, if just due to the manual effort involved.
+However, employing manually operated Rietveld Refinement is not without issues.
+It is certainly not scalable to the degree that would be required to keep up with advances
+in throughput and efficiency in other steps of the experimentation pipeline, even just due to the manual effort involved.
+But beyond the lacking scalability of this workflow we also wish to highlight the following issue:
+Providing the analysis software with an initial guess of the structure requires consulting a reference library
+of crystal structures with corresponding simulated diffractograms and
+identifying a crystal structure that is likely to be similar to that of the investigated sample.
+Since the number of chemically viable crystal structures is infinite, any such library will always be incomplete.
+The best that can be done to remedy this is to continually update such a reference library both on the side
+of the distributor and the user, which in itself represents an undesirable overhead.
+Especially for novel material discovery it may be impossible to find a sufficiently accurate match in the reference
+library to apply Rietveld Refinement successfully.
+
+Machine learning is a very promising candidate to eliminate both the need for manual inputs and the need for reference
+libraries in powder x-ray diffraction analysis.
 
@@
 
-While the ICSD contains a large number of structures spanning many different
-classes of materials, it still falls short in size, distribution, and generality
-compared to the datasets used to train very large state-of-the-art models of
-other fields such as computer vision.
-%for image classification (ImageNet
-%challenge\supercite{russakovskyImageNetLargeScale2015}). 
-Furthermore, the ICSD
-database is highly imbalanced with respect to space groups, as can be seen
-in the histogram in Figure~\ref{fig:histogram_dist_spgs}. This makes the classification of
-space groups more difficult, as shown and discussed by
-\citeauthor{zalogaCrystalSymmetryClassification2020}
-\supercite{zalogaCrystalSymmetryClassification2020}. The ICSD also contains a
-limited number of different structure types that may not adequately represent
-the crystal structures analyzed in future experiments.
-
-To overcome these shortcomings, we propose to train machine learning models on
-diffractograms simulated from synthetic crystal structures randomly generated
-based on the symmetry operations of the space groups. This makes it possible to
-train on structures with new structure types not present in the ICSD. We used
-the crystals from the ICSD only to determine vague statistics guiding the random
-generation and for calculating the test accuracy. Our approach goes one step
-further than classical data augmentation by fully detaching itself from the
-individual entries in the ICSD database.
-\replaced{The generated synthetic crystals form a training dataset that includes stable ICSD crystal structures, unstable crystal structures,
-but also stable structures that are not yet present in the ICSD.
-By training a model on the full dataset, we can also expect improvements on the unknown stable crystal structures.}{It is our hypothesis that when trying
-to solve more general tasks concerning the extraction of structural details from
-powder diffractograms, the crystals from which the training dataset is simulated
-do not need to be stable or chemically viable.} \replaced{Furthermore, we}{We} propose viewing the problem
-\deleted{simply} as a mathematical task of getting back some of the real-space information
-leading to given powder X-ray diffractograms. \added{Therefore, even the unstable structures included in our generated dataset will help to learn to classify the stable structures.}
-
-Here, we applied this approach to the classification of the crystal symmetry,
-namely the space group. The space group is usually one of the first structural
-pieces of information needed after synthesizing a new material. This task is
-well-suited to showcase the strengths of using a synthetic dataset and to
-benchmark it. We further show the results of using our methodology to infer space group labels of an experimental dataset.
-
-We embedded our synthetic generation algorithm in a framework with
-distributed computing capabilities to generate and simulate diffractograms on
-multiple nodes in parallel using the \emph{Python} library \emph{Ray}
-\supercite{moritzRayDistributedFramework2018a}. In contrast to the traditional
-approach of generating a simulated dataset before training, we used this
-distributed computing architecture to build an infinite stream of \replaced{synthetically}{randomly}
-generated and simulated diffractograms to perform batch-wise online learning.
-This increases the generalization performance, eliminates the problem of
-overfitting, and allows very large models to be trained.
-
 % PDF pricing

### 2024-04-10 paper/01_motivation.tex
Section on training the model

@@
-%but it requires a good initial guess of the structure possibly combined with manual adjustments. Both \textbf{}
-
 % Autonomous laboratory
@@
 The synthesis and probing of novel materials is becoming increasingly efficient and automated,
-which has correspondingly increased the throughput of samples in these steps
-of the experimentation pipeline.
+which has correspondingly increased the throughput of samples in these steps of the experimentation pipeline.
 
@@
 The parameters of the working model of the sample are then adjusted to minimize the difference between the simulated
-diffraction pattern and the one observed in experiment.
+diffraction pattern and the pattern observed in experiment.
 Critically, this process does not generally guarantee to guide the working model towards a global minimum of the difference
 in diffractograms and hence arrive at an accurate model of the sample.
-In practice the refinement process may only guide the working model towards the nearest local minimum
+In general one can only assume the refinement process to guide the working model towards the nearest local minimum
 starting from the initial guess.
@@
 Still this workflow often works well, since usually the operator of the experiment will be able
-to provide an at least somewhat accurate initial guess of the crystal structure.
+to provide an at least somewhat accurate initial guess of the crystal structure.\\
 
@@
 But beyond the lacking scalability of this workflow we also wish to highlight the following issue:
-Providing the analysis software with an initial guess of the structure requires consulting a reference library
-of crystal structures with corresponding simulated diffractograms and
-identifying a crystal structure that is likely to be similar to that of the investigated sample.
+Providing the analysis software with an initial guess of the structure requires manually or algorithmically
+consulting a reference library of crystal structures with corresponding simulated or sometimes real diffractograms.
+From that reference library a crystal structure must be identified that is likely to be similar
+to that of the investigated sample which can then serve as the initial guess provided to the analysis software.
 Since the number of chemically viable crystal structures is infinite, any such library will always be incomplete.
-The best that can be done to remedy this is to continually update such a reference library both on the side
-of the distributor and the user, which in itself represents an undesirable overhead.
-Especially for novel material discovery it may be impossible to find a sufficiently accurate match in the reference
+The best that can be done to remedy this is to continually update the library both on the side
+of the distributor and the user, which is an undesirable overhead.
+Especially for novel material discovery, it may be impossible to find a sufficiently accurate match in the reference
 library to apply Rietveld Refinement successfully.
@@
 Machine learning is a very promising candidate to eliminate both the need for manual inputs and the need for reference
-libraries in powder x-ray diffraction analysis.
+libraries in powder x-ray diffraction analysis, resulting in a streamlined and automated analysis process that can
+keep pace with an automated high-throughput experimentation environment. \\
+
+Models can be trained to predict crystal structure properties given only the diffractogram itself, given that
+the model can be trained on dataset consisting of pairs of diffractograms and corresponding crystal structures.
+So far however, due to an absence of sufficiently large labeled datasets with diffractograms from experiment,
+machine learning in this domain has largely relied on simulated diffractograms.
+Note, that reference libraries which are used for Rietveld Refinement by in large use simulated diffractograms
+rather than diffractograms from experiment and hence cannot either provide a dataset with diffractograms coming
+from experiment.
+
+Within the domain of predicting the properties of simulated diffractograms
 

### 2024-04-10 paper/00_abstract.tex
Added performance translation section

@@
 A noteable difficulty in applying machine learning to this domain is the lack of labeled datasets i.e.\
-raw experimental diffraction patterns with accompanying structure solutions, which has relegated machine learning
+diffraction patterns originating from experiment with accompanying structure solutions, which has relegated machine learning
 researches to training on simulated data.
 Since these simulations largely fail to accurately reflect experiment, the performance of models trained
-on this simulated data, also by in large fail to transfer to experiment and provide value in practice.
-We aim to remedy this by providing a freely available and easily accessible dataset of powder diffraction data
-, providing machine learning researches with a large quantity of real experimental diffractograms collected
-from a broad range of samples.
+on this simulated data, also by in large fails to transfer to experiment and provide value in practice.
+We aim to remedy this by providing a freely available and easily accessible dataset of partially labeled
+powder diffraction data, providing machine learning researches with a large quantity of real experimental diffractograms
+collected from a broad range of samples.
 We hope that this can guide machine learning research towards more realistic
 simulations and potentially enable training and testing models on real experimental data.
-
-% mention below:
-% uniform, easily readable
 

### 2024-04-10 paper/01_motivation.tex
Added performance translation section

@@
 rather than diffractograms from experiment and hence cannot either provide a dataset with diffractograms coming
-from experiment.
+from experiment. \\
 
-Within the domain of predicting the properties of simulated diffractograms
+Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting
+lattice parameters, spacegroup and crystallite size from simulated diffractograms.
+There is every reason to believe, that this performance can translate to predicting crystal structure properties
+of diffractograms originating from experiment, given that the models are trained on a dataset that accurately
+reflects diffractograms as they appear in experiment.\\
 

### 2024-04-10 paper/01_motivation.tex
Added dataset usefulness

--- 2024-04-10 paper/01_motivation.tex
@@
 rather than diffractograms from experiment and hence cannot either provide a dataset with diffractograms coming
-from experiment.
+from experiment. \\
 
-Within the domain of predicting the properties of simulated diffractograms
+Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting
+lattice parameters, spacegroup and crystallite size from simulated diffractograms but the performance drops off immensely
+when these models are applied to data originating from experiment.
+There is every reason to believe, that the strong performance observed in the domain of purely simulated diffractograms
+can translate to predicting crystal structure properties of diffractograms originating from experiment, given that
+models are trained on a dataset that accurately reflects diffractograms as they appear in experiment. \\
+
+However due to the scarcity of labeled diffractograms originating from experiment, collecting a labeled dataset
+of real experimental powder x-ray diffraction data of sufficient size to train a model
+that can be useful in practice is next to impossible and is not the aim of this dataset.
+Rather we hope that this dataset can enable machine learning researchers to evaluate how closely their simulations
+represent actual experimental data, identify what features are missing and modify their simulations algorithms
+accordingly to arrive at a simulation that nearly perfectly mirrors how data appears in reality.
+Secondly we would like to also present this dataset as a challenge: The fraction of labeled data can be used to
+easily test the performance of the model on a broad range of data as it appears in reality, which can be used
+to gauge how well models would perform in practice.
 

### 2024-04-10 paper/01_motivation.tex
Added dataset as challenge

@@
 represent actual experimental data, identify what features are missing and modify their simulations algorithms
-accordingly to arrive at a simulation that nearly perfectly mirrors how data appears in reality.
+accordingly to arrive at a simulation that nearly perfectly mirrors how data appears in reality. \\
+
 Secondly we would like to also present this dataset as a challenge: The fraction of labeled data can be used to
-easily test the performance of the model on a broad range of data as it appears in reality, which can be used
-to gauge how well models would perform in practice.
+easily numerically evalute the performance of the model on a broad range of data as it appears in experiment, which can be used
+to gauge how well models would perform in practice. \\
+But even the unlabeled portion of the dataset can be used to test the performance of models: The least a model
+has to do in order to prove useful in practice is to make a guess about the crystal structure that is good enough
+for Rietveld refinement to converge toward the true structure. \\
+Hence, the performance of the model can be tested by evaluating if the inital guess of the model is sufficient for
+Rietveld refinment to refine the initial guess to a structure that produces a diffractogram matching the one observed in experiment.
+Finally, by inviting the application of models to data as it appears in experiment we would like to bring attention
+to some of the practical difficulties that need to be addressed in applying a model to a wide variety of x-ray diffractograms
+collected from different sources: For example the angle and the step size of the diffractograms will in general be variable
+which is something that any model that aims to perform well on this dataset but be equipped to handle.
 

### 2024-04-12 paper/00_abstract.tex
New title idea

@@
 diffraction patterns originating from experiment with accompanying structure solutions, which has relegated machine learning
-researches to training on simulated data.
+researchers to training on simulated data.
 Since these simulations largely fail to accurately reflect experiment, the performance of models trained
@@
 We aim to remedy this by providing a freely available and easily accessible dataset of partially labeled
-powder diffraction data, providing machine learning researches with a large quantity of real experimental diffractograms
+powder diffraction data, providing machine learning researchers with a large quantity of real experimental diffractograms
 collected from a broad range of samples.

### 2024-04-12 paper/01_introduction.tex
New title idea

--- 2024-04-10 paper/01_motivation.tex
@@
 represent actual experimental data, identify what features are missing and modify their simulations algorithms
-accordingly to arrive at a simulation that nearly perfectly mirrors how data appears in reality. \\
+accordingly to arrive at a simulation that nearly perfectly mirrors how data appears in reality.\\
+One very simple approach which could give a first hint of how closely simulations reflect experiment would be to attempt
+to train a classifier on simulated and real diffractograms.
+If the classifier performs well that would be an indicator that the simulation doesn't reflect experiment sufficiently well.
+Note that the converse doesn't necessarily hold as well.
+For a set of given sample properties there is only one corresponding simulated diffractogram, but there may an infinite amount of
+variations in how the diffractogram of a sample could look like in experiment for example due varying measurement apperatus, noise
+, sample impurities and lattice defects.
+In order for a simulated dataset to closely resemble experiment it is critical that the simulatio also represents these
+statistical variations.
+Determining the nature of these variations is a much a harder task than making individual diffractograms look like they
+originated from experiment and cannot be accomplished by simply training a classifier.
+Nevertheless, we hope that this dataset can at least provide clues of how to take this aspect into account during
+simulation as well.\\
 
@@
 to some of the practical difficulties that need to be addressed in applying a model to a wide variety of x-ray diffractograms
-collected from different sources: For example the angle and the step size of the diffractograms will in general be variable
-which is something that any model that aims to perform well on this dataset but be equipped to handle.
+collected from different sources: For example the angle and the step size of the diffractograms will differ between sources
+, which is something that any model that aims to perform well on this dataset must be equipped to handle.
 
@@
 
-Machine learning techniques have emerged as a powerful tool in the toolkit of
-materials scientists. While they are often used to make predictions on
-the properties of materials or find new materials with certain properties, an increasingly interesting domain is the automated analysis of raw experimental measurements
-guided by machine learning\supercite{radovicMachineLearningEnergy2018}.
-
-With the advent of high-throughput experiments, the amount of gathered data is
-vast and the analysis often becomes a bottleneck in the processing pipeline
-\supercite{rahmanianEnablingModularAutonomous2022}. Powder X-ray diffraction
-(XRD) is an important measurement technique used to obtain structural
-information from polycrystalline samples\supercite{harrisContemporaryAdvancesUse2001}. 
-The diffractograms are an
-information-dense fingerprint of the structure of the material. However,
-analyzing these diffractograms is not an easy task\supercite{holderTutorialPowderXray2019}. 
-Full structure solutions and
-Rietveld refinement take time and require expert knowledge, both about the
-analysis technique and the materials class at hand. This is not feasible in high-throughput experiments on a larger
-scale. Therefore, the question arises whether it is possible to automatically 
-analyze powder diffractograms with machine learning models trained on large
-amounts of data, making it possible to run inference almost instantaneously.
-
-During the last few years, there have been several studies tackling this
-objective by applying machine learning models to various tasks concerning the
-analysis of powder diffractograms such as phase
-classification\supercite{leeDeeplearningTechniquePhase2020,maffettoneCrystallographyCompanionAgent2021,schuetzkeEnhancingDeeplearningTraining2021,szymanskiProbabilisticDeepLearning2021,wangRapidIdentificationXray2020},
-phase fraction determination\supercite{leeDatadrivenXRDAnalysis2021}, space
-group
-classification\supercite{parkClassificationCrystalStructure2017,oviedoFastInterpretableClassification2019,zalogaCrystalSymmetryClassification2020,vecseiNeuralNetworkBased2019,suzukiSymmetryPredictionKnowledge2020,chakrabortyDeepCrystalStructure2022},
-machine-learning-guided Rietveld
-refinement\supercite{ozakiAutomatedCrystalStructure2020,fengMethodArtificialIntelligence2019},
-extraction of lattice
-parameters\supercite{dongDeepConvolutionalNeural2021,chitturiAutomatedPredictionLattice2021,chakrabortyDeepCrystalStructure2022,
-habershonPowderDiffractionIndexing2004} and crystallite
-sizes\supercite{dongDeepConvolutionalNeural2021,chakrabortyDeepCrystalStructure2022},
-and also novelty detection based on unsupervised techniques\supercite{bankoDeepLearningVisualization2021}.
-Since an abundant source of experimental diffractograms is hard to come by, most
-applications train their models on simulated diffractograms from the Inorganic
-Crystal Structure Database (ICSD)\supercite{bergerhoff1987}, which contains a total of 272\,260 structures (October 2022).
-
-\citeauthor{leeDeeplearningTechniquePhase2020} used a deep convolutional neural
-network (CNN) trained on a large dataset of multiphase compositions from the
-quaternary \ce{Sr-Li-Al-O} pool to classify present phases in the
-diffractogram\supercite{leeDeeplearningTechniquePhase2020}. In a follow-up
-study, they further showed good results for phase fraction inference in the
-quaternary \ce{Li-La-Zr-O} pool\supercite{leeDatadrivenXRDAnalysis2021}.
-\citeauthor{schuetzkeEnhancingDeeplearningTraining2021} performed phase
-classification on iron ores and cement compounds and used data augmentation with
-respect to lattice parameters, crystallite sizes, and preferred orientation
-\supercite{schuetzkeEnhancingDeeplearningTraining2021}. They showed that
-especially the lattice parameter variations enhance the classification accuracy
-significantly.
-
-Instead of the analysis of phase composition,
-\citeauthor{dongDeepConvolutionalNeural2021} performed regression of scale
-factors, lattice parameters, and crystallite sizes in a five-phase catalytic
-materials system\supercite{dongDeepConvolutionalNeural2021}. In contrast to
-supervised tasks, \citeauthor{bankoDeepLearningVisualization2021} used a
-variational autoencoder to visualize variations in space group, preferred
-orientation, crystallite size, and peak
-shifts\supercite{bankoDeepLearningVisualization2021}.
-\citeauthor{parkClassificationCrystalStructure2017} used a deep CNN to classify
-space groups of single-phase diffractograms, reaching a test accuracy of 81.14\%
-on simulated diffractograms.\supercite{parkClassificationCrystalStructure2017}
-However, as we will show later in this paper, this accuracy is highly overestimated and drops to 56.1\% when test splits are designed in a way to reduce data leakage in non-IID datasets such as the ICSD.
-\citeauthor{vecseiNeuralNetworkBased2019}\supercite{vecseiNeuralNetworkBased2019} developed
-a similar approach and applied their classifier to experimental diffractograms
-from the RRUFF mineral database\supercite{lafuentePowerDatabasesRRUFF2015},
-reaching an experimental test accuracy of 54\%.
+%
+%
+%Machine learning techniques have emerged as a powerful tool in the toolkit of
+%materials scientists. While they are often used to make predictions on
+%the properties of materials or find new materials with certain properties, an increasingly interesting domain is the automated analysis of raw experimental measurements
+%guided by machine learning\supercite{radovicMachineLearningEnergy2018}.
+%
+%With the advent of high-throughput experiments, the amount of gathered data is
+%vast and the analysis often becomes a bottleneck in the processing pipeline
+%\supercite{rahmanianEnablingModularAutonomous2022}. Powder X-ray diffraction
+%(XRD) is an important measurement technique used to obtain structural
+%information from polycrystalline samples\supercite{harrisContemporaryAdvancesUse2001}.
+%The diffractograms are an
+%information-dense fingerprint of the structure of the material. However,
+%analyzing these diffractograms is not an easy task\supercite{holderTutorialPowderXray2019}.
+%Full structure solutions and
+%Rietveld refinement take time and require expert knowledge, both about the
+%analysis technique and the materials class at hand. This is not feasible in high-throughput experiments on a larger
+%scale. Therefore, the question arises whether it is possible to automatically
+%analyze powder diffractograms with machine learning models trained on large
+%amounts of data, making it possible to run inference almost instantaneously.
+%
+%During the last few years, there have been several studies tackling this
+%objective by applying machine learning models to various tasks concerning the
+%analysis of powder diffractograms such as phase
+%classification\supercite{leeDeeplearningTechniquePhase2020,maffettoneCrystallographyCompanionAgent2021,schuetzkeEnhancingDeeplearningTraining2021,szymanskiProbabilisticDeepLearning2021,wangRapidIdentificationXray2020},
+%phase fraction determination\supercite{leeDatadrivenXRDAnalysis2021}, space
+%group
+%classification\supercite{parkClassificationCrystalStructure2017,oviedoFastInterpretableClassification2019,zalogaCrystalSymmetryClassification2020,vecseiNeuralNetworkBased2019,suzukiSymmetryPredictionKnowledge2020,chakrabortyDeepCrystalStructure2022},
+%machine-learning-guided Rietveld
+%refinement\supercite{ozakiAutomatedCrystalStructure2020,fengMethodArtificialIntelligence2019},
+%extraction of lattice
+%parameters\supercite{dongDeepConvolutionalNeural2021,chitturiAutomatedPredictionLattice2021,chakrabortyDeepCrystalStructure2022,
+%habershonPowderDiffractionIndexing2004} and crystallite
+%sizes\supercite{dongDeepConvolutionalNeural2021,chakrabortyDeepCrystalStructure2022},
+%and also novelty detection based on unsupervised techniques\supercite{bankoDeepLearningVisualization2021}.
+%Since an abundant source of experimental diffractograms is hard to come by, most
+%applications train their models on simulated diffractograms from the Inorganic
+%Crystal Structure Database (ICSD)\supercite{bergerhoff1987}, which contains a total of 272\,260 structures (October 2022).
+%
+%\citeauthor{leeDeeplearningTechniquePhase2020} used a deep convolutional neural
+%network (CNN) trained on a large dataset of multiphase compositions from the
+%quaternary \ce{Sr-Li-Al-O} pool to classify present phases in the
+%diffractogram\supercite{leeDeeplearningTechniquePhase2020}. In a follow-up
+%study, they further showed good results for phase fraction inference in the
+%quaternary \ce{Li-La-Zr-O} pool\supercite{leeDatadrivenXRDAnalysis2021}.
+%\citeauthor{schuetzkeEnhancingDeeplearningTraining2021} performed phase
+%classification on iron ores and cement compounds and used data augmentation with
+%respect to lattice parameters, crystallite sizes, and preferred orientation
+%\supercite{schuetzkeEnhancingDeeplearningTraining2021}. They showed that
+%especially the lattice parameter variations enhance the classification accuracy
+%significantly.
+%
+%Instead of the analysis of phase composition,
+%\citeauthor{dongDeepConvolutionalNeural2021} performed regression of scale
+%factors, lattice parameters, and crystallite sizes in a five-phase catalytic
+%materials system\supercite{dongDeepConvolutionalNeural2021}. In contrast to
+%supervised tasks, \citeauthor{bankoDeepLearningVisualization2021} used a
+%variational autoencoder to visualize variations in space group, preferred
+%orientation, crystallite size, and peak
+%shifts\supercite{bankoDeepLearningVisualization2021}.
+%\citeauthor{parkClassificationCrystalStructure2017} used a deep CNN to classify
+%space groups of single-phase diffractograms, reaching a test accuracy of 81.14\%
+%on simulated diffractograms.\supercite{parkClassificationCrystalStructure2017}
+%However, as we will show later in this paper, this accuracy is highly overestimated and drops to 56.1\% when test splits are designed in a way to reduce data leakage in non-IID datasets such as the ICSD.
+%\citeauthor{vecseiNeuralNetworkBased2019}\supercite{vecseiNeuralNetworkBased2019} developed
+%a similar approach and applied their classifier to experimental diffractograms
+%from the RRUFF mineral database\supercite{lafuentePowerDatabasesRRUFF2015},
+%reaching an experimental test accuracy of 54\%.
 

### 2024-04-28 paper/00_abstract.tex
Moved main tex file and references to repo root

@@
-In powder x-ray diffraction, the analysis of the diffractograms still presents a significant bottleneck
-in high-throughput experimentation that remains to be solved, in order to match the increased throughput
+Powder X-ray diffraction (pXRD) experiments are a cornerstone for materials structure characterization.
+Despite its widespread application, the analysis of the diffractograms still presents a significant bottleneck
+in high-throughput experimentation that remains to be solved in order to match the increased throughput
 of the prior stages of experimentation.
-While Rietveld refinement can identify the phases and extract the structural information found
-in the diffractograms, each sample requires manual inputs and careful configuration of the analysis software
-from trained staff, which is infeasible for highly automated high-throughput experiments.
-
-In the current landscape of powder x-ray diffraction analysis, machine learning has emerged as a very promising direction of research resolving this bottleneck by enabling fully automated powder diffraction analysis.
-A noteable difficulty in applying machine learning to this domain is the lack of labeled datasets i.e.\
+%While Rietveld refinement can identify the phases and extract the structural information found
+%in the diffractograms, each sample requires manual inputs and careful configuration of the analysis software
+%from trained staff, which is infeasible for highly automated high-throughput experiments.
+In the current landscape of powder X-ray diffraction analysis, machine learning has emerged as a promising research direction to resolve this bottleneck by enabling automated powder diffraction analysis.
+A notable difficulty in applying machine learning to this domain is the lack of labeled datasets, i.e.\
 diffraction patterns originating from experiment with accompanying structure solutions, which has relegated machine learning
-researchers to training on simulated data.
-Since these simulations largely fail to accurately reflect experiment, the performance of models trained
-on this simulated data, also by in large fails to transfer to experiment and provide value in practice.
+researchers to train on simulated data.
+Since these simulations largely fail to accurately reflect the experiment, the performance of models trained
+on this simulated data by and large fails to transfer to experiment and provide value in practice.
 We aim to remedy this by providing a freely available and easily accessible dataset of partially labeled
-powder diffraction data, providing machine learning researchers with a large quantity of real experimental diffractograms
-collected from a broad range of samples.
-We hope that this can guide machine learning research towards more realistic
-simulations and potentially enable training and testing models on real experimental data.
-
-%Machine learning techniques have successfully been used to extract structural information such as the crystal space group from powder X-ray diffractograms.
-%However, training directly on simulated diffractograms from databases such as
-%the ICSD is challenging due to its limited size, class-inhomogeneity, and bias
-%toward certain structure types. We propose an alternative approach of generating
-%synthetic crystals with random coordinates by using the symmetry operations of
-%each space group. Based on this approach, we 
-%demonstrate online training of deep ResNet-like models on up to a few
-%million unique on-the-fly generated synthetic diffractograms per hour. For our chosen task of
-%space group classification, we achieved a test accuracy of 79.9\% on unseen ICSD
-%structure types from most space groups.
-This surpasses the 56.1\% accuracy of the current state-of-the-art approach of training on ICSD crystals directly.
-Our results demonstrate that synthetically generated crystals can be
-%used to extract structural information from ICSD powder diffractograms, which makes it possible to apply very large state-of-the-art machine learning models in the area of powder X-ray diffraction.
-%We further show first steps toward applying our methodology to experimental data, where automated XRD data analysis is crucial, especially in high-throughput settings.
-%While we focused on the prediction of the space group, our approach has the potential to be extended to related tasks in the future.
+powder diffraction data, providing machine learning researchers with a large quantity of real experimental diffractograms collected from a broad range of samples.
+We hope that this can guide machine learning research toward more realistic handling of experimental imperfections to enable the improvement of current automated analysis workflows.

### 2024-04-28 paper/02_existing_datasets.tex
Highlighted reference library problem
Added performance translation section
Moved main tex file and references to repo root

@@
-\subsection{Generating synthetic crystals} \label{sec:generate_crystals}
+% Source of png : https://www.sciencedirect.com/topics/biochemistry-genetics-and-molecular-biology/rietveld-refinement
 
-    To generate synthetic crystals, we randomly place atoms on the Wyckoff
-    positions of a given space group \added{following the Wyckoff occupation probabilities extracted from the ICSD} and
-    then apply the respective symmetry operations. The algorithm is
-    explained in the following (see also Figure~\ref{fig:generation_and_distributed}a for a flow diagram of the algorithm).
-    We only explain the most important steps, details can be found in Section~S1 of the SI.
-
-    \begin{enumerate}
-        \itemsep0em
-        \item \replaced{Sampling of a space group from the space group distribution of the ICSD.}{Random selection of a space group.
-We follow the space group distribution of the ICSD to allow comparison with previous work.}
-        \item Sample unique elements of the crystal and their number of repetitions in the asymmetric unit.
-        \item Place atoms onto the Wyckoff positions and draw uniform coordinates for each.
-        \item Draw lattice parameters from a \replaced{kernel density estimate}{KDE} based on the ICSD.
-        \item Apply space group symmetry operations.
-    \end{enumerate}
-
-    Parts of this algorithm were inspired by the generation algorithm of the
-    \emph{Python} library \emph{PyXtal}
-    \supercite{fredericksPyXtalPythonLibrary2021}. 
-    We only keep generated crystals for training if the
-    conventional unit cell volume is below \SI{7000}{\angstrom\cubed} and if
-    there are less than 100 atoms in the asymmetric unit. 
-    We did not employ any form of
-    distance checks on the coordinates, as we found this to have no meaningful
-    impact on space group classification accuracy. We only prevented the
-    algorithm from placing more than one atom onto a Wyckoff position that does not
-    have a degree of freedom. We also did not use partial occupancies. We chose
-    this algorithm for its simplicity and its capability to reproduce many
-    important characteristics of ICSD crystals adequately (see Section~
-    \ref{sec:synth_distribution}).
-
-    \begin{figure}[h]
-    \centering
-    \includegraphics{./figures/distribution_spgs_arrows.pdf}
-    \caption{Distribution (logarithmic scale in blue, linear scale in red)
-    of space groups in the ICSD. Space groups are sorted by count\added{ (see Figure S13 in the SI for the distribution without sorting by count)}. The population of the space groups varies by
-    multiple orders of magnitude, showing that the ICSD is a highly imbalanced
-    dataset regarding space groups. The space groups excluded due to insufficient
-    statistics are visualized with black stripes.
-\replaced{The histogram displays the distribution of the full ICSD, while the exclusion of space groups that do not contain enough samples is based on the statistics dataset (which does not include the test dataset, see Section~\ref{sec:dataset_split}) that we used to guide the random crystal generation.
-Therefore, the excluded space groups are not exactly the first 85 counted from the left.}{Since the exclusion is based on
-    the statistics dataset and this histogram shows the full ICSD,
-    the excluded space groups are not exactly the first 85 counted from the left.}}
-    \label{fig:histogram_dist_spgs}
-    \end{figure}
-
-    For some space groups, there are not enough crystals in the ICSD to form a
-    representative \replaced{kernel density estimate}{KDE} for the volume or to calculate suitable occupation
-    probabilities for individual Wyckoff positions. Therefore, we chose to only
-    perform the classification on space groups with 50 or more crystals
-    available in the statistics dataset we used to extract the probabilities
-    (see Section~\ref{sec:dataset_split}), resulting in the exclusion of 85
-    space groups (see Figure~\ref{fig:histogram_dist_spgs}). A classifier trained directly on ICSD data of all space groups will likely not be able to properly identify these space
-    groups containing very few samples.
-
-    If a similar performance for all space groups is desired, a uniform
-    distribution of space groups in the training dataset is needed. This is
-    trivially possible with our synthetic approach, in contrast to training
-    directly on the ICSD, where weighting, over-, or undersampling methods are
-    needed \supercite{sunClassificationImbalancedData2009}. To allow a direct
-    and fair comparison between our approach and the original approach of
-    training directly on ICSD entries, we still followed the same distribution
-    of space groups of the ICSD in our synthetic training dataset. This
-    eliminates the problem that the effective number of total space groups is
-    smaller when training on a highly imbalanced dataset, making it easier to
-    reach high accuracies.
-
-    \added{Our choice of not sampling the space groups uniformly and using
-    general statistics extracted from the ICSD to guide the crystal generation
-    algorithm further builds upon the hypothesis that future crystals will
-    roughly follow the more general statistics already present in the ICSD. With
-    the chosen crystal generation algorithm we tried to find a middle ground
-    between being much more general than using the ICSD crystals directly and
-    not being too general such that it is very hard to extract structural
-    information at all.}
-
-    \begin{figure}[!htb] 
-    \centering
-    \includegraphics{./figures/combined_generation_and_distributed_system.pdf}
-    \caption{a) Flowchart of how the generation algorithm produces synthetic
-    crystals. Atoms are independently placed on the Wyckoff positions and random
-    coordinates are drawn. b) Overview of the distributed computing system
-    implemented using the \emph{Python} library
-    \emph{Ray}\supercite{moritzRayDistributedFramework2018a}. Two compute nodes
-    (that generate and simulate diffractograms) are connected to the \emph{Ray}
-    head node using a \emph{Ray queue} object.}
-    \label{fig:generation_and_distributed}
-    \end{figure}
-
-    \subsection{Simulating diffractograms} \label{sec:simulation}
-
-    To simulate powder X-ray diffractograms based on the generated crystals, we used
-    the implementation found in the \emph{Python} library \emph{Pymatgen}
-    \supercite{ongPythonMaterialsGenomics2013}. We optimized the simulation code
-    using the LLVM just-in-time compiler \emph{Numba}
-    \supercite{lamNumbaLLVMbasedPython2015}. This increases the performance of
-    the main loop over the reciprocal lattice vectors of the crystal
-    significantly and makes the continuous simulation while training (discussed
-    in the next section) possible.
-
-    We used the wavelength \SI{1.5406}{\angstrom} (\replaced{Cu K$\alpha_1$}{Cu K$_\alpha$} line) to
-    simulate all diffractograms. The obtained peaks were further broadened with
-    a Gaussian peak profile to form the full diffractogram. To obtain the peak
-    widths, we used the Scherrer
-    equation\supercite{gilmoreInternationalTablesCrystallography2019}
-    \begin{equation}
-        \beta=\frac{K \lambda}{L \cos \theta} \text{,} \label{eq:scherrer}
-    \end{equation}
-    where $ \beta $ is the line broadening at half maximum intensity (on the
-    $2\theta$-scale), $K$ is a shape factor, $\lambda$ is the wavelength, and
-    $L$ is the (average) thickness of crystallites. We drew crystallite sizes
-    from the range $ \left[ 20, 100 \right] $ \si{\nano \metre} and used 
-    $K=0.9$.
-
-    Diffractograms were generated in the range $ 2 \theta \in \left[ 5, 90
-    \right] \si{\degree} $ with step size \SI{0.01}{\degree}. After generating
-    each diffractogram, it was rescaled to fit in the intensity range $ \left[
-    0,1 \right]$. \added{In Figure S9 of the SI we show an exemplary diffractogram simulated 
-    from the ICSD, Figure S10 shows an exemplary diffractogram simulated from a synthetic 
-    crystal.}
-
-    \subsection{Continuous generation of training data}\label{sec:online_learning}
-
-    Typically, machine learning models are trained with a fixed dataset predefined at the beginning of training. Sometimes, data augmentation
-    is applied to further increase the effective size of this dataset. In
-    contrast to that, we generated our dataset on-the-fly, parallel to model training. The
-    main advantage of using this approach compared to a fixed-size dataset is
-    the eliminated possibility to overfit to individual diffractograms since
-    every diffractogram is only used once. Furthermore, not having to
-    pre-simulate a dataset before training makes this approach more flexible
-    when changing simulation parameters.
-
-    We used a distributed architecture on multiple nodes using the \emph{Python}
-    framework \emph{Ray} \supercite{moritzRayDistributedFramework2018a}, which enabled the training on 1-2 GPUs and simultaneous generation of training data on more than 200 CPU cores (see Figure~\ref{fig:generation_and_distributed}b and SI Section~S2.2).
-    Depending on
-    the model size and corresponding training speed, this setup allows
-    training with up to millions of unique diffractograms per hour.
-
-    \subsection{Dataset split} \label{sec:dataset_split}
-
-    The ICSD database contains many structures that are very similar with
-    slightly different lattice parameters and coordinates. For example, there
-    are 25 entries for \ce{NaCl} (October 2022). Furthermore,
-    there are 3898 entries that have the same structure type as \ce{NaCl} and
-    thus also similar powder diffractograms. If some of them appear in the
-    training dataset and some in the test dataset, the classification will be
-    simplified to recognizing the structure type or structure. In that case, the
-    test set accuracy will not represent the true generalization performance of the neural
-    network. To quantify the true generalization performance, we split the
-    dataset in such a way that the same structure type appears either only in
-    the training or in the test dataset. We used the structure type definitions
-    provided by the ICSD. The obtained accuracy on the test dataset reflects the
-    accuracy of our network when being used on a novel sample with a structure
-    type not yet present in the ICSD database.
-
-    We want to emphasize that the used test split is very important for the task of
-    space group classification and not a trivial choice. The ICSD contains many
-    subtypes of structure types (for example, subtypes of perovskites), which we
-    regarded as separate structure types in our split. Considering the subtypes
-    as the same structure type may also be a viable option when performing the
-    split. A combination of a split based on structure type and sum formula or
-    similar approaches are also possible. \deleted{However, we think that our choice of
-    splitting the dataset is able to effectively measure the generalization
-    performance while still being relatively simple.}
-
-    \added{Depending on the experimental setting, it further might make more sense in some cases to not do a structure type-based split.
-If the likelihood of finding structures similar to already-discovered structure types in the planned experiment is high, training should definitely include those structure types to evaluate the performance of the model.
-However, in a pure discovery setting, new structure types can appear.
-To evaluate the expected model performance in this scenario and thus quantify the true generalization error to unseen data, we chose the most strict structure type-based split.}
-
-    We divided the ICSD (database version 2021, June 15) in a 70:30 split. For
-    our synthetic crystal approach, the 70\% part (\replaced{which we
-    call}{called} statistics or training dataset) was only used to create the
-    \replaced{kernel density estimates}{KDEs} and to calculate the Wyckoff
-    occupation probabilities needed for the generation algorithm. Since we can
-    judge the performance of the synthetic generation algorithm by comparing the
-    training accuracy (on synthetic crystals) with the accuracy tested on
-    diffractograms simulated directly from the statistics dataset, an additional
-    validation dataset was not needed. For comparison with the original approach
-    of directly training on ICSD
-    crystals\supercite{parkClassificationCrystalStructure2017}, we simulated
-    crystals directly from the statistics dataset and trained on them.
-
-    Analogous to the synthetic generation, we only used crystals with a
-    conventional unit cell volume below \SI{7000}{\angstrom\cubed} and with less
-    than 100 atoms in the asymmetric unit for the statistics and test dataset.
-    This covers $ \approx 94\% $ of the ICSD crystals.
-
-    \subsection{Models and \added{computational} experiments} \label{sec:models_and_experiments}
-
-    \subsubsection*{Models}
-    We will briefly introduce the models we used for the classification
-    of space groups. A more detailed description can be found in the SI
-    Section~S2.1.
-
-    As a baseline, we first used the CNN models introduced by
-    \citeauthor{parkClassificationCrystalStructure2017}\supercite{parkClassificationCrystalStructure2017}.
-    They used three models, one for the classification of crystal systems
-    (``parkCNN small''), one for extinction groups (``parkCNN medium''), and one
-    for space groups (``parkCNN big''). All models have three convolution layers
-    with two hidden fully connected layers and one output layer. The
-    three models differ in the number of neurons in the hidden fully connected
-    layers, increasing the number of model parameters with the number of target
-    labels. Here, we only used the models ``parkCNN medium'' and ``parkCNN big''
-    and applied both to the classification of space groups. When using ICSD
-    crystals to train the ``parkCNN'' models, dropout was used, while the
-    training of the ``parkCNN'' models on synthetic crystals did not use
-    dropout.
-
-    Since the approach of using an infinite stream of generated training data
-    eliminates the problem of overfitting, we further used deeper models with a
-    higher number of model parameters. For this, we used the deep convolutional
-    neural networks ResNet-10, ResNet-50, and ResNet-101, which were introduced
-    by
-    \citeauthor{heDeepResidualLearning2016}\supercite{heDeepResidualLearning2016}
-    in 2015.
-
-    Details of the machine learning setup can be found in the SI Section~2.2.
-Overall, our setup allowed us the training of models over up to 2000 epochs with more than $100\,000$ unique, newly generated crystals and corresponding diffractograms in each epoch \added{(see the upper x-axis of Figure \ref{fig:training_curve})}.
-
-    \subsubsection*{\replaced{Computational experiments}{Experiments}}
-
-    We performed two sets of experiments to evaluate our new dataset split as well as our synthetic crystal generation approach and compare it to state-of-the-art models in literature: Firstly, we trained and tested models on ICSD crystals only, and secondly, we trained on synthetic crystals and tested on ICSD crystals.
-
-    In particular, we first performed an experiment with the ``parkCNN medium''
-    model trained directly on the diffractograms simulated from the ICSD
-    statistics dataset with a fully random train-test split (similar to
-    \supercite{parkClassificationCrystalStructure2017}), instead of splitting by
-    the structure type of the crystals. This experiment makes a comparison of
-    the two different methods of train-test split possible. We then trained the
-    ``parkCNN big'' model using the structure type-based split, again directly
-    on ICSD diffractograms. We further repeated the same experiment using the
-    smaller model ``parkCNN medium'' to resolve potential overfitting to the
-    ICSD diffractograms.
-
-    For the experiments performed on our continuously generated
-    dataset based on synthetic crystals, we used the structure type-
-    based split. As discussed in Section~\ref{sec:dataset_split}, the training / 
-    statistics dataset was only used to extract more general statistics, such as the 
-    element distribution.
-    First, we trained the ``parkCNN big'' model. For each
-    batch, we generated $ 435 $ random crystals and simulated two diffractograms
-    with different crystallite sizes for each of them. This resulted in the
-    batch size of $ 870 $. 
-    Since our synthetic crystal generation algorithm yields an infinite stream
-    of unique diffractograms to train on, using much larger models than for the
-    fixed ICSD dataset is possible without overfitting. We performed experiments
-    for the ResNet-10, ResNet-50, and ResNet-101 models. Instead of generating
-    two diffractograms with different uniformly sampled crystallite sizes for each generated crystal (as
-    we did for the ``parkCNN big'' model), we now created only one diffractogram
-    for each of the $ 145 $ crystals generated for one batch. This is due to 
-    the slower training of the ResNet models, which means that reusing the same 
-    diffractogram with different crystallite sizes is not necessary to generate
-    training data fast enough.
-
-    To obtain the
-    highest-possible ICSD test accuracy, we further applied the square root
-    function as a preprocessing step to the input diffractograms of the
-    network when using the ResNet models. This was suggested by
-    \citeauthor{zalogaCrystalSymmetryClassification2020}
-    \supercite{zalogaCrystalSymmetryClassification2020} and in their case
-    improved classification accuracy by approximately 2 percentage points. Some
-    initial tests suggested that this approach also yields a higher accuracy in our
-    case, so we used this preprocessing step to train the ResNet
-    models.
-
-    While we focused mainly on the methodology of using synthetic crystals to
-    extract structural information from powder diffractograms, we also show some
-    initial steps toward applying our methods to experimental data. We used the
-    publicly available RRUFF mineral database
-    \supercite{lafuentePowerDatabasesRRUFF2015} which provides experimental
-    measurements, including powder diffractograms \added{(see Figure S11 in the
-    SI for an exemplary diffractogram from the RRUFF)}. In order to imitate
-    experimental diffractograms, we added Gaussian additive and multiplicative
-    noise \added{(similar to
-    \supercite{vecseiNeuralNetworkBased2019,szymanskiProbabilisticDeepLearning2021})}
-    and a background function based on samples from a Gaussian process to our
-    simulated diffractograms. Furthermore, we added a small amount of an
-    impurity phase to each diffractogram. Details about the experimental data
-    generation protocol can be found in the SI Section~S4\added{, Figure S12
-    shows an exemplary synthetic diffractogram with noise, background and an
-    impurity phase.} Using the ResNet-50 model, we performed two experiments for
-    experimental data, one with the mentioned impurity phase, and one without.
+% Zenodo data e.g. is not uniform, easily redable but we aimt ot be
+% uniform, easily readable

### 2024-04-28 paper/03_our_dataset.tex
New title idea
Moved main tex file and references to repo root

@@
-    \subsection{Synthetic distribution} \label{sec:synth_distribution}
+% Sharing raw powder x-ray diffraction data: Challenges and benefits https://www.researchgate.net/publication/329256004_Sharing_powder_diffraction_raw_data_Challenges_and_benefits
 
-    \begin{figure}[!htb]
-    \centering
-    \includegraphics{./figures/example_structures.pdf}
-    \caption{a) Some \replaced{randomly chosen and thus representative}{randomly picked} examples of ICSD crystals. b) Some \replaced{randomly chosen and thus representative}{randomly
-    picked} examples of synthetically generated crystals. While coordination and
-    distances are not chemically correct for the synthetic crystals, crystal
-    symmetries are reproduced correctly.}
-    \label{fig:example_structures}
-    \end{figure}
-
-    \begin{figure}[!htb]
-    \centering
-    \includegraphics{./figures/histograms/histograms.png} 
-    \caption{Histograms comparing the distributions of descriptors of the
-    synthetically generated crystals with the ICSD distribution in the test
-    dataset. a) density factor $ \frac{V_\text{unit cell}}{\sum_i 4/3 \pi
-    \left(\frac{r_{i\text{;cov}} + r_{i\text{;VdW}}}{2}\right)^3} =
-    \frac{V_\text{unit cell}}{V_\text{atomic}}$, b) crystallite sizes, c) unit cell
-    volume (conventional cell settings), d) number of atoms in the asymmetric
-    unit. The probability density of the ICSD is visualized by a stacked bar
-    histogram, where the green portion of the bar was correctly classified and
-    the red portion was incorrectly classified. The probability density of the
-    synthetic crystals is visualized by the dark blue line. The portion between the
-    dark blue line and the light blue line was correctly classified, the portion below the
-    light blue line was incorrectly classified. The reported classification
-    performance is based on the ResNet-101 model trained on diffractograms from
-    synthetic crystals.}
-    \label{fig:histograms}
-    \end{figure}
-
-    We first present an analysis of the generated synthetic crystals.
-    Figure~\ref{fig:example_structures} shows some \replaced{randomly chosen and thus representative examples of}{randomly selected} ICSD and
-    synthetic crystal structures side-by-side. Visually, the crystals appear very
-    similar.
-However, no physical or chemical considerations regarding stability, clashing atoms, and element combinations are taken into account in the generation of synthetic crystals.
-As discussed earlier, our goal is to demonstrate that this is
-    not problematic when using these crystals for the extraction of structural
-    information from powder diffractograms.
-On the contrary, we expect the synthetic crystals to be a better basis for generalization to fundamentally new crystal structures than existing finite databases.
-
-    To compare the distribution of ICSD crystals with the synthetic
-    distribution, we evaluated structural descriptors, i.e. density factors, crystallite sizes, unit cell volumes, and numbers of atoms in the asymmetric unit, and compare their histograms
-    in Figure~\ref{fig:histograms}.
-    One can see that the overall distributions of
-    the synthetic and ICSD crystals are very similar for all four descriptors.
-    This shows that our chosen generation algorithm reproduces crystals that are
-    similar to ICSD crystals in terms of these more general descriptors.
-
-    \subsection{Classification results} \label{sec:classification_results}
-
-    The main results of our experiments (see Section~\ref{sec:models_and_experiments}) to
-    classify the space group of powder diffractograms can be
-    found in Table~\ref{tab:results}. \added{In SI Table S2, we further provide the training time and total number of unique diffractograms for each computational experiment.}
-    The goal of our experiments is to systematically analyse and quantify the changes in classification accuracy introduced by our two main contributions: A more challenging dataset split, and training on continuously generated synthetic data.
-
-    We started by repeating previously reported
-    experiments\supercite{parkClassificationCrystalStructure2017} trained
-    directly using ICSD crystals with a random train-test split instead of the
-    split based on structure types. This model achieved a very high test
-    accuracy of 83.2\%. We note that the previous publication that we compare
-    our results to\supercite{parkClassificationCrystalStructure2017} removed
-    data from the training dataset, ``[...] heavily duplicated data
-    [...]''\supercite{parkClassificationCrystalStructure2017}, but did not
-    specify the exact criterions used. In contrast, we did not exclude any
-    duplicates in this experiment based on a random train-test split.
-    Furthermore, as discussed in Section~\ref{sec:generate_crystals}, we
-    excluded crystals with a very high unit cell volume and a very high number
-    of atoms in the asymmetric unit. This is likely the reason for the slightly
-    higher classification accuracy that we observed, compared to the originally
-    reported 81.1\%.
-
-    When splitting randomly, the model merely needs to recognize structures or
-    structure types and assign the correct space group. This task is much easier
-    than actually extracting the space group using more general patterns. When
-    going from random splits to structure type-based splits (see
-    Section~\ref{sec:dataset_split}), it becomes obvious that both the ``parkCNN
-    big'' as well as the ``parkCNN medium'' models overfit the training data and
-    do not generalize well to unseen structure types in the test set \added{(see
-    Table \ref{tab:results})}. \added{The ``parkCNN medium'' model, which
-    achieved 83.2\% on a random split, now only yields 55.9\% with the structure
-    type-based split.}
-
-    \replaced{Training the models by
-    \citeauthor{parkClassificationCrystalStructure2017}}{Training the same
-    model}, in particular the ``parkCNN big'' model, on synthetic crystals leads
-    to a 1.6 percentage points higher test accuracy than the ``parkCNN big''
-    model trained on ICSD diffractograms. At the same time, the training
-    accuracy drops from \added{the} 87.2\% \added{when we trained the model
-    directly on the ICSD} to 74.2\% \added{on the synthetic distribution}
-    indicating that the model is now limited more by missing capacity rather
-    than by overfitting, which is why we explored larger models, which will be
-    discussed later. The gap between training and test accuracy is $31.1$
-    percentage points when training on ICSD data, while for training using
-    synthetic crystals, the gap is only $16.5$ percentage points. We note that
-    this gap between training using synthetic crystals and testing using ICSD
-    crystals cannot stem from overfitting, since no diffractograms are repeated
-    for the synthetic training. The difference rather stems from the differences
-    between the synthetic distribution and the ICSD distribution of crystals.
-
-    \begin{table}[!htb]
-        \begin{center}
-        \caption{Results of training on diffractograms simulated from ICSD crystals (random splits as well as structure type-based splits) compared to when training on diffractograms from synthetic
-        crystals. Test accuracy
-        in all cases refers to the accuracy when testing on the ICSD test
-        dataset. The training accuracies are averaged over the last 10 epochs of
-        the respective run.
-        Experiments trained directly on ICSD data
-        overfitted to the training data. Training longer would have further
-        increased the training accuracy, while not increasing the test
-        accuracy.}
-        \vspace*{2mm}
-
-        \begin{tabular}{ccccccc} 
-            \toprule
-            Split & \begin{tabular}{@{}c@{}}Training \\ dataset\end{tabular} & \begin{tabular}{@{}c@{}}Testing \\ dataset\end{tabular} & Model & \begin{tabular}{@{}c@{}}Number of \\ parameters\end{tabular}& \begin{tabular}{@{}c@{}}Training acc.
-\\ / \%\end{tabular} & \begin{tabular}{@{}c@{}}Test acc.
-\\ / \%\end{tabular} \\
-            \midrule
-            Random & ICSD & ICSD & parkCNN medium & 4\,246\,797 & 88.4 & 83.2 \\
-            \midrule
-            Structure& \multirow{2}{*}{ICSD} & \multirow{2}{*}{ICSD}& parkCNN big & 4\,959\,585 & 87.2 & 56.1 \\
-            type&&& parkCNN medium & 4\,246\,797 & 90.9 & 55.9 \\
-            \midrule
-            & \multirow{4}{*}{synthetic} & \multirow{4}{*}{ICSD} &parkCNN big & 4\,959\,585 & 74.2 & 57.7 \\
-            Structure&&& ResNet-10 & 9\,395\,025 & 87.2 & 73.4 \\
-            type$^{1}$ &&& ResNet-50 & 41\,362\,385 & 91.8 & 79.3 \\
-            &&& ResNet-101 & 60\,354\,513 & 92.2 & \textbf{79.9} \\
-            \bottomrule
-        \end{tabular}\label{tab:results}
-        \end{center} 
-        \footnotesize{$^{1}$Here, the split type refers to the statistics and the test dataset, rather than the training and the test dataset.}
-    \end{table}
-
-    While the ``parkCNN big'' model trained on synthetic crystals outperforms
-    the approach of training directly on ICSD crystals by only 1.6 percentage
-    points, the advantage of training on an infinite stream of synthetic data increases when using models with more parameters and thus higher capacity. In
-    contrast to training directly on a finite set of ICSD crystals, it is possible to train
-    very large models using the infinite synthetic data stream without the
-    potential of overfitting. As found in the last lines of Table~\ref{tab:results}, ResNet-10, ResNet-50, and ResNet-101 based models achieve ICSD test accuracies of 73.4\%,
-    79.3\%, and 79.9\%. This is a significant increase from the 57.7\% achieved
-    by the ``parkCNN big'' model. Figure~S4 in the SI further shows the top-$k$ accuracy over $k$ for the ResNet-101 model. With increasing $k$ the accuracy exceeds 95\% at $k=5$.
-    This means that our model can not only determine the correct space group with a high probability but can also generate an almost complete list of possible space group candidates.
-
-    Figure~\ref{fig:training_curve} shows the ICSD test accuracy, the training
-    accuracy (on synthetic data), and the ICSD top-5 test accuracy for all three ResNet variants
-    as a function of epochs trained. For all three metrics, the difference
-    between ResNet-50 and ResNet-101 is comparably small, while the step from
-    ResNet-10 to ResNet-50 is substantial (5.9 percentage points in ICSD test
-    accuracy, see Table~\ref{tab:results}). This shows that going beyond the
-    model size of the ResNet-101 will likely not yield a big improvement in
-    accuracy. In contrast to the 79.9\% accuracy reached in the top-1 ICSD test accuracy, the top-5 ICSD test accuracy of the ResNet-101 model reaches
-    96\%. However, for all three ResNet variants, a gap between training using synthetic
-    crystals and testing using the ICSD remains (12.3 percentage points for
-    ResNet-101).
-As also shown in Figure~S5 in the SI, the accuracy convergence can be approximately described by a power law, indicating that exponentially more training time will substantially reduce classification errors and thus potentially lead to top-1 accuracies of 90\% and above, at the cost of a 100-fold increase in training times\replaced{.
-Considering the current training times provided in Table S2 of the SI, this is currently infeasible or only possible with tremendous hardware resources.}{, which is currently infeasible.}
-
-    \begin{figure}[!htbp] 
-    \centering
-    \includegraphics{figures/training_curves/training_curve_main.png}
-    \caption{Test accuracy (ICSD), training accuracy (synthetic crystals), and
-    test top-5 accuracy (ICSD) as a function of epochs\added{ (bottom axis).
-Since each additional epoch contains newly generated unique diffractograms, we further show the accuracies as a function of the total number of unique synthetic diffractograms (top axis)}.
-We show all three
-    metrics for the models ResNet-101, ResNet-50, and ResNet-10. To better show
-    the scaling behavior, both axes use logarithmic scaling. \added{Figure S6
-    shows the same plot but without logarithmic scaling.} To better see the
-    exponential behaviour, see Figure~S5 in the SI. }
-    \label{fig:training_curve} 
-    \end{figure}
-
-    The histograms in Figure~\ref{fig:histograms} show, next to the overall
-    distribution, also the fraction of diffractograms classified wrongly for
-    testing on the ICSD (red bar) and on the synthetic data (below the light blue line)
-    for the ResNet-101 model. First, one can see that throughout almost all
-    regions of the distributions, the accuracy on the synthetic data is slightly
-    higher than that on the ICSD. This is related to the aforementioned gap of 12.3
-    percentage points between train and test accuracy and can be attributed to
-    differences between the synthetic and ICSD distribution of crystals. This
-    will be discussed in detail in the next section. It is surprising to see that the
-    dependence on crystallite sizes is rather weak, as smaller crystallite sizes
-    result in broader peaks (see Scherrer equation, Eq.~\ref{eq:scherrer}), potentially making the
-    classification harder due to more peak overlaps.
-
-    In summary, the maximum ICSD test accuracy of 79.9\% that we achieved using
-    the ResNet-101 model almost reaches the previously
-    reported\supercite{parkClassificationCrystalStructure2017} 81.14\% for the
-    space group classification. However, our accuracy is based on a train-test
-    split based on structure types, in contrast to a random split. This creates
-    a much harder but also realistic task to solve since the model needs to
-    generalize to other structure types without merely recognizing
-    diffractograms or structure types that it has already seen during training.
-    This becomes especially apparent from our experiment directly trained on
-    diffractograms from ICSD crystals with the split based on structure types,
-    which reached only 56.1\% instead of the previously
-    reported\supercite{parkClassificationCrystalStructure2017} 81.14\%.
-
-    \subsubsection*{Experimental results}
-    To go beyond simulated diffractograms, we trained ResNet-50 models on
-    calculated diffractograms with background, noise, and impurities and applied
-    the trained models to the RRUFF mineral database. Our results (see Figure~S3
-    in the SI) show that it is essential to include impurity phases in the
-    training data. By doing so, we obtain a top-1 accuracy of 25.2\% and a
-    top-10 accuracy of over 60\%. This is of high practical relevance since
-    having a short list of potential space groups is often sufficient as a first
-    step to further refinement and analysis.
-
-    \citeauthor{vecseiNeuralNetworkBased2019} performed similar experiments of
-    space group classification on the same database. Using an ensemble of 10
-    fully connected neural networks, they reached a classification accuracy of
-    54\%\supercite{vecseiNeuralNetworkBased2019}. While our obtained accuracy is
-    significantly lower, our approach is much more general: In contrast to our
-    approach, the training dataset was based on simulated diffractograms of
-    structures of the ICSD\supercite{vecseiNeuralNetworkBased2019}, which
-    contains almost all RRUFF structures, leading to high similarities of
-    training and \replaced{test}{testing} data. Therefore, the model needed to
-    simply recognize the minerals, instead of directly inferring the space group
-    using the symmetry elements - as our method needs to do.
-
-    We want to emphasize that our efforts to apply the methodology to
-    experimental data are only preliminary. We expect improved results with an
-    improved data generation protocol since the procedure contains many
-    parameters to be tuned. Ideally, one would use a generative machine learning
-    approach to add the experimental effects (noise, background, impurities) to
-    the pure diffractograms. We also want to point out that the noise level and
-    quality of data in the RRUFF dataset are limited. Application of the
-    presented methodology \replaced{to}{on} other experimental datasets is
-    desirable. \deleted{We will address these points in future work, where we
-    focus on improved ways of modeling experimental imperfections.}
-    \added{As discussed above, for the
-    classification of pure diffractograms we observed the ResNet-50 to have the
-    best cost-benefit ratio, since the ResNet-101 yielded only slight
-    improvements. For the more complicated problem of classifying diffractograms
-    with experimental imperfections, bigger models and longer training times
-    might be necessary.}
-
-    \added{Next to improving the modeling of experimental imperfections and therefore the overall accuracy on experimental data, the practical application of deep neural networks for analyzing powder diffractograms yields further challenges that we want to discuss.
-Since experimental setups differ, e.g., concerning the used wavelength, a different $2\theta$ step size, or a different $2\theta$ range, a new neural network would need to be trained for each situation.
-Since our largest model requires a significant computational investment, this might not be feasible in all situations.
-Arguably, though, for large high-throughput experiments, the 11-day training of a ResNet-50 should not be unreasonable, especially if it can speed up the data analysis significantly and allow in-loop adaptive experimentation.
-For smaller setups, where this is not feasible, other solutions must be found.
-First, one can use a form of transfer learning from a pre-trained model to fine-tune to the desired experimental setup.
-This, however, would only work for a change in wavelength, since a change in step size or $2\theta$ range would change the input dimensions of the network.
-However, to handle a change in the $2\theta$ range, it might be possible to include a form of zero-masking in the synthetic training data, such that different input ranges (with zeros where no measurement was made) can be used, which would lead to a more flexible model, not requiring new training data when applied to a new $2\theta$ range.
-For a change in the step size, a cubic spline interpolation might be helpful.
-We plan to address these challenges in future work.}
-
-    \deleted{As discussed above, for the classification of pure diffractograms we
-    observed the ResNet-50 to have the best cost-benefit ratio, since the
-    ResNet-101 yielded only slight improvements. For the more complicated
-    problem of classifying diffractograms with experimental imperfections,
-    bigger models and longer training times might be necessary.}
-
-    \added{Furthermore, analysis of the loss value or gradient norm associated with particular samples, i.e.
-crystal structures, during training on synthesis crystals or during transfer learning from synthetic to experimental data can help to better understand the relevance and informativeness of given samples for the model.
-This can help in generating more relevant synthetic data based on experimental crystal structures that are underrepresented in the synthetic data distribution.}
-
-    \subsection{Differences between synthetic crystals and ICSD crystals}
-    We showed that training directly on crystals from the ICSD yields a gap
-    between the training and test accuracy due to overfitting. The training on
-    the synthetic dataset also shows a gap between the training and test
-    accuracy (see Table~\ref{tab:results}), but it is smaller than when training
-    directly on ICSD crystals. Furthermore, this gap is not due to overfitting,
-    since overfitting to singular diffractograms is not possible when the model
-    is trained using an infinite stream of generated synthetic crystals. The gap
-    rather stems from systematic differences between the synthetic and ICSD
-    distribution of crystals.
-
-    To analyze those differences, we created three modifications of the ICSD
-    test dataset \added{(see SI Section~S3 for details)}. In the first modification, the fractional coordinates of the
-    atoms in the asymmetric unit of the crystals of the ICSD test dataset were
-    randomly uniformly resampled (as in the synthetic crystal generation
-    algorithm). In the second modification, the lattice parameters were
-    randomized following the \replaced{kernel density estimate}{KDE} used in the synthetic generation algorithm. The
-    third modification combines both previous modifications, i.e. both the
-    coordinates and the lattice parameters were resampled. These three modified
-    test datasets bring the ICSD test dataset closer to the distribution used
-    for training and let us quantify which factors contribute to the gap between
-    training \added{on synthetic crystals} and testing \added{on the ICSD}.
-
-    We evaluated the test accuracies on \replaced{the randomized}{these}
-    datasets for the experiment using the ResNet-101 model trained using
-    synthetic crystals. We found that randomizing the coordinates yields an
-    increase in test accuracy of \replaced{$4.89$}{$5.96$} percentage points.
-    Randomizing the lattice parameters results in an increase of
-    \replaced{$0.79$}{$1.31$} percentage points. Randomizing both the
-    coordinates and the lattice parameters leads to an increase of
-    \replaced{$5.70$}{$6.32$} percentage points, explaining
-    \replaced{almost}{approximately} half of the gap of 12.3 percentage points
-    between synthetic training and ICSD test accuracy.
-
-    \added{So far, we have randomized the lattice parameters and coordinates of the test dataset, such that they follow a distribution that is based on the statistics extracted from the statistics dataset.
-However, this does not take into account the different Wyckoff position occupation probabilities between the test and statistics datasets.
-For this, we repeated
-    a similar analysis, for which we applied the
-    randomizations to the statistics dataset rather than the test dataset.
-    Without any modifications, testing on the statistics dataset instead of the test dataset yielded $3.89$ percentage points higher accuracy.
-This can be explained by slight differences in the overall statistics between the test and statistics datasets.
-    Randomizing the coordinates yields a further increase of $4.72$
-    percentage points, randomizing the lattice $1.16$ percentage points, and
-    randomizing both the coordinates and the lattice parameters $6.68$ percentage points.
-    In total, testing on the statistics dataset with randomized coordinates and
-    lattice parameters yields a $10.57$ percentage points higher accuracy than on the
-    unmodified test dataset. This almost completely explains the gap
-    of $12.3$ percentage points between the training accuracy on synthetic crystals and the test accuracy on the ICSD. The
-    remaining part is likely due to our algorithm that places atoms on Wyckoff
-    positions 
-    not reproducing the ICSD distribution exactly. 
-    However, the
-    remaining difference is remarkably small.}
-
-    \deleted{We also tested the ResNet-101 model on diffractograms simulated
-    based on the statistics dataset from which the statistics for our generation
-    algorithm have been extracted. This returns a $4.33$ percentage points
-    higher accuracy than on the test dataset. This is due to the different
-    structure types present in the test and statistics dataset yielding
-    different occupation probabilities (potentially also 0) for Wyckoff
-    positions and overall slightly different statistics for the synthetic
-    generation algorithm.}
-    \deleted{To summarize, the contributions to the gap stem from differences in the
-    distribution of lattice parameters and coordinates with $6.32$ percentage
-    points and from the differences between the statistics dataset and test
-    dataset with $4.33$ percentage points. Added together (assuming the effects of geometry and Wyckoff position occupation probabilities are unrelated and thus additive), the contributions
-    almost completely explain the gap of $12.3$ percentage points between
-    training and test accuracy. The remaining part is likely due to our
-    algorithm that places atoms on Wyckoff positions not reproducing the ICSD
-    distribution exactly. However, the remaining difference is remarkably small.}
-
-    \begin{figure}[!htb]
-    \centering
-    \includegraphics{./figures/histograms/histograms_rel.png}
-    \caption{Classification \replaced{error}{performance} for each bin of a) the unit cell volume
-    (conventional cell settings) and b) the number of atoms in the asymmetric
-    unit. \deleted{The classification performance when testing on the ICSD is shown using
-    a stacked bar plot, where the red bar indicates the classification error and
-    the green bar indicates the correctly classified portion of each bin. When
-    testing on the synthetic distribution, the classification error is given by
-    the light blue line, the rest (between dark blue and light blue line) was correctly
-    classified.} The reported classification performance is based on the
-    ResNet-101 model trained on diffractograms from synthetic crystals. This
-    visualization clearly shows the error rate within each bin, in contrast to
-    Figure~\ref{fig:histograms}, which additionally includes the relative
-    proportion of the crystals of the respective bin to the total amount of
-    crystals.}
-    \label{fig:histograms_rel}
-    \end{figure}
-
-    In Figure~\ref{fig:histograms_rel} we show the test classification error in
-    each bin for the unit cell volume and the number of atoms in the asymmetric
-    unit using the ResNet-101 model trained on diffractograms of synthetic
-    crystals. The classification error is shown both for testing on
-    diffractograms from synthetic crystals and on ICSD diffractograms. One can
-    see that for small volumes and a small number of atoms in the asymmetric
-    unit, the difference between classifying ICSD diffractograms and
-    diffractograms from synthetic crystals is relatively small. As the volume
-    and number of atoms in the asymmetric unit increase, the gap between the two
-    errors increases, too. We already identified the uniformly sampled atom
-    coordinates in the synthetic distribution as the main contributor to the gap
-    in accuracy between the synthetic crystals and ICSD crystals. Therefore, it
-    seems that the uniform sampling of atom coordinates works well for small 
-    number of atoms in the asymmetric unit and small volumes, while the error 
-    due to this sampling strategy increases slightly for higher volumes and higher
-    number of atoms in the asymmetric unit.
-
-    When looking at the distribution of crystals in the ICSD, the number of
-    atoms in the asymmetric unit tends to be larger for lower-symmetry space
-    groups (for example, in the triclinic crystal system) than for
-    higher-symmetry space groups such as those from the cubic crystal system.
-    Therefore, the increasing test error on diffractograms from ICSD crystals
-    with a higher number of atoms in the asymmetric unit is especially relevant
-    for these lower-symmetry space groups.
-    It might be possible that a different scheme of generating atom positions in
-    the unit cell (compared to the independent uniform sampling that we used)
-    works better for a high number of atoms in the asymmetric unit.
-
-    Overall, it is important to note that the distribution of ICSD crystals is (apart from a few Wyckoff position occupation probabilities which are exactly zero in the statistics dataset\footnote{Setting them to small non-zero values typically leads to the generation of rather large unit cells, as the general Wyckoff positions have high multiplicities.}) almost completely encompassed by the much larger distribution of snynthetic crystals that we used for training.
-    However, due to finite training times and model capacity, a performance gap remains.
-This gap can be improved by using (substantially) more computing power or by narrowing the very general synthetic distribution, e.g., by using a different algorithm to generate atom positions.
-    This indicates an inherent challenge in XRD classification but more generally in materials property prediction:
-    Machine learning models are ultimately trained to be employed in real-world tasks, which are typically related to novel, i.e. yet unseen materials and structures.
-    At the same time, the machine learning models are tested based on an IID assumption, i.e. the assumption that the distribution of training and testing data is the same.
-    While not being a contradiction in the limit of infinite training data and model capacity, this becomes an (unsolvable) challenge in reality, when facing finite datasets and models.
-    In our case, our model trained on a large distribution of synthetic crystal structures will likely generalize better to completely new crystal structures different from any crystal structure contained in the ICSD database.
-    At the same time, it suffers from smaller ICSD test set errors, even though the ICSD distribution is contained in the synthetic data generation distribution.
+% The subsections of this section should be the 
+% individual contributions that we get from labs

### 2024-04-29 paper/03_our_dataset.tex
Included first notes on dataset section

@@
+% Notes:
+% The subsections of this section should be the 
+% individual contributions that we get from labs
+
 % Sharing raw powder x-ray diffraction data: Challenges and benefits https://www.researchgate.net/publication/329256004_Sharing_powder_diffraction_raw_data_Challenges_and_benefits
 
-% The subsections of this section should be the 
-% individual contributions that we get from labs
+In collaboration with several other institutions we have collected a dataset of diffractograms stemming from experiment, some of them labeled with corresponding structural information. \\
+Currently, institutions that contributed towards our dataset include:
+\item Institute of Nanotechnology at Karlsruhe institute of technology
+\item University of Southern California
+\item Lawrence Berkeley National Laboratory
+
+We are still in the process of growing the dataset and additional contributions are still very welcome. 
+To find out more about how to contribute to this dataset, visit our website specially designed for the purpose of collecting this dataset, https://xrd.aimat.science \\
+
+A major advantage of this dataset over other comparable datasets is that is is very easy to handle in python and in PyTorch. \\
+Our accompanying python library xrdpattern (https://github.com/aimat-lab/xrdpattern) provides a means to read in as much of the dataset as fits in memory simply by using PatternDB.load(data_dirpath).
+The patterns attribute of this class is a list of the indivdual pattern diffractograms, each of which supports a standardize, plot and to tensor method.
+\\
+The standardize method returns a "standardized" version of the pattern with a fixed angle range, a fixed number of entries and intensities normalized to the $[0,1]$ interval.

### 2024-04-29 paper/01_introduction.tex
Weakened the statement about translating neural networks performance on data stemming from experiment

@@
 Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting
-lattice parameters, spacegroup and crystallite size from simulated diffractograms but the performance drops off immensely
-when these models are applied to data originating from experiment.
-There is every reason to believe, that the strong performance observed in the domain of purely simulated diffractograms
-can translate to predicting crystal structure properties of diffractograms originating from experiment, given that
-models are trained on a dataset that accurately reflects diffractograms as they appear in experiment. \\
+lattice parameters, spacegroup and crystallite size from simulated diffractograms but the performance drops off immensely when these models are applied to data originating from experiment.
+
+The patterns observed in experiment suffer from a host of imperfections that may result in a partial loss of information or make extracting the structural information more difficult. \\
+However, there is still reason to believe that the strong performance observed in the domain of purely simulated diffractograms can to a large extent translate to predicting crystal structure properties from diffractograms originating from experiment, given that models are trained on a dataset that accurately reflects diffractograms as they appear in experiment.
 

### 2024-04-29 paper/02_existing_datasets.tex
Included notes on existing datasets

@@
 
-% Zenodo data e.g. is not uniform, easily redable but we aimt ot be
-% uniform, easily readable
+% Sources to mention
+% Zenodo -> A good amount of raw powder data there but spread into smaller datasets; Inhomogenous, not specially prepared to be worked with by others, therefore difficult to work with especially at scale
+
+% Crystal structures with corresponding simulated diffractograms;
+% ->  ICSD (https://icsd.products.fiz-karlsruhe.de/) 
+% - 230k structures;
+% - Proprietary
+
+% -> PDF (https://www.icdd.com/) -> 
+%- 1200k "material data sets" (structures?) 
+% + 19 k entries with "Experimental raw data digital patterns"; 
+%  Proprietary
+%The PDF itself lists the following data sources:
+\begin{itemize}
+    \item \textbf{ICDD Powders (00) – International Centre for Diffraction Data Newtown Square, PA, USA}
+    \item \textbf{ICSD (01) – FIZ Karlsruhe – Leibniz Institute for Information Infrastructure (FIZ), Karlsruhe, Germany – Inorganic Crystal Structure Database}
+    \item \textbf{NIST (03) – National Institute of Standards and Technology, Gaithersburg, Maryland, USA}
+    \item \textbf{MPDS (04) – Material Phases Data System, Vitznau, Switzerland – Linus Pauling File (LPF)}
+    \item \textbf{ICDD Single Crystal Data (05) – International Centre for Diffraction Data Newtown Square, PA, USA}
+\end{itemize}
+
+% ->  COD (https://www.crystallography.net/cod/) -
+% -  513k structures,
+% - Fully open

### 2024-04-29 paper/01_introduction.tex
Commited changes by Henrik

@@
+\begin{itemize}
+    \item Plan for Figures / Tables
+    \begin{itemize}
+        \item Figure 1: Multiple panels showing histograms of some key metrics: Space group, unit cell volume (normalizable?), ...
+        \item Table 1: List of existing powder XRD datasets
+        \item Table 2: List of existing ML approaches for pXRD
+        \item Table 3: New pXRD datasets offered by this paper
+    \end{itemize}
+
+    \item Additional thoughts
+    \begin{itemize}
+        \item General structure of the paper:
+        \begin{itemize}
+            \item Introduction
+            \item Introduce existing datasets and already show why they are not enough / too small
+            \item Summarize SOTA ML that has been applied to experimental pXRD, discuss the limits and gap between performance on simulated data and exp.
+data (find some good example papers where this is very apparent)
+            \item Summary of our dataset, including paragraphs of the different contributions by other groups; comparison with existing datasets, especially in terms of size
+        \end{itemize} 
+    \end{itemize}
+\end{itemize}
+
 % Autonomous laboratory
@@
 to some of the practical difficulties that need to be addressed in applying a model to a wide variety of x-ray diffractograms
-collected from different sources: For example the angle and the step size of the diffractograms will differ between sources
-, which is something that any model that aims to perform well on this dataset must be equipped to handle.
+collected from different sources: For example the angle and the step size of the diffractograms will differ between sources, 
+which is something that any model that aims to perform well on this dataset must be equipped to handle.
 

### 2024-04-29 paper/02_existing_datasets.tex
Commited changes by Henrik

@@
 %The PDF itself lists the following data sources:
+
 \begin{itemize}
-    \item \textbf{ICDD Powders (00) – International Centre for Diffraction Data Newtown Square, PA, USA}
-    \item \textbf{ICSD (01) – FIZ Karlsruhe – Leibniz Institute for Information Infrastructure (FIZ), Karlsruhe, Germany – Inorganic Crystal Structure Database}
-    \item \textbf{NIST (03) – National Institute of Standards and Technology, Gaithersburg, Maryland, USA}
-    \item \textbf{MPDS (04) – Material Phases Data System, Vitznau, Switzerland – Linus Pauling File (LPF)}
-    \item \textbf{ICDD Single Crystal Data (05) – International Centre for Diffraction Data Newtown Square, PA, USA}
+\item Powder Diffraction File (ICDD)
+\begin{itemize}
+    \item ICDD Powders (00) – International Centre for Diffraction Data Newtown Square, PA, USA
+    \item ICSD (01) – FIZ Karlsruhe – Leibniz Institute for Information Infrastructure (FIZ), Karlsruhe, Germany – Inorganic Crystal Structure Database
+    \item NIST (03) – National Institute of Standards and Technology, Gaithersburg, Maryland, USA
+    \item MPDS (04) – Material Phases Data System, Vitznau, Switzerland – Linus Pauling File (LPF)
+    \item ICDD Single Crystal Data (05) – International Centre for Diffraction Data Newtown Square, PA, USA
+\end{itemize}
 \end{itemize}

### 2024-04-29 chapters/04_our_dataset.tex
Added todos in introduction

--- 2024-04-29 paper/03_our_dataset.tex
@@
 Currently, institutions that contributed towards our dataset include:
+\begin{itemize}
     \item Institute of Nanotechnology at Karlsruhe institute of technology
@@
     \item Lawrence Berkeley National Laboratory
+\end{itemize}
 
@@
 
-A major advantage of this dataset over other comparable datasets is that is is very easy to handle in python and in PyTorch. \\
-Our accompanying python library xrdpattern (https://github.com/aimat-lab/xrdpattern) provides a means to read in as much of the dataset as fits in memory simply by using PatternDB.load(data_dirpath).
+A major advantage of this dataset over other comparable datasets is that is is very easy to handle in python and in PyTorch.
+Our accompanying python library xrdpattern (https://github.com/aimat-lab/xrdpattern) provides a means to read in as much of the dataset as fits in memory simply by using PatternDB.load(data\_dirpath).
 The patterns attribute of this class is a list of the indivdual pattern diffractograms, each of which supports a standardize, plot and to tensor method.

### 2024-04-29 chapters/01_introduction.tex
Added todos in introduction
Added idea for paths towards improving simulation

--- 2024-04-29 paper/01_introduction.tex
@@
 Rather we hope that this dataset can enable machine learning researchers to evaluate how closely their simulations
-represent actual experimental data, identify what features are missing and modify their simulations algorithms
-accordingly to arrive at a simulation that nearly perfectly mirrors how data appears in reality.\\
-One very simple approach which could give a first hint of how closely simulations reflect experiment would be to attempt
-to train a classifier on simulated and real diffractograms.
-If the classifier performs well that would be an indicator that the simulation doesn't reflect experiment sufficiently well.
-Note that the converse doesn't necessarily hold as well.
-For a set of given sample properties there is only one corresponding simulated diffractogram, but there may an infinite amount of
-variations in how the diffractogram of a sample could look like in experiment for example due varying measurement apperatus, noise
+represent actual experimental data, identify what features are missing and modify their simulations algorithms accordingly to arrive at a simulation that nearly perfectly mirrors how data appears in reality.\\
+
+There are roughly speaking two paths towards improving the simulation: The first is to more accurately model the sample and the processes taking place within it during the simulation of the powder diffraction pattern.
+This would for example necessitate modelling absorption of the incoming X-ray radiation and its re-emission in the form of characteristic radiation and modeling the effect of stress and motion within the crystal on the diffractogram.
+\\
+An alternative route to obtaining a more accurate simulation is to try to learn these "imperfections" compared to an idealized simulation using machine learning techniques by extracting them from this dataset.
+\\
+As an example of this could be done we can look at results from other domains, where a transfer of core data between different "styles" of representation was attempted:
+%\cite{Ganin2015} -> Attempted to
+
+%A first hint of how closely a given simulation reflects experiment could be obtained by
+%training a machine learning model, which attempts to classify between simulated and real diffractograms. \\
+%If the simulation is very accurate, we would expect the classifier to not have few to no features which it can use to distinguish between simulated and real diffractograms and hence perform poorly.
+\\
+%Conversely, if the classifier performs well and is not difficult to train, that would be an indicator that the simulation doesn't reflect experiment sufficiently well. \\
+
+For a set of given sample properties there is only one corresponding simulated diffractogram, but there may an infinite amount of variations in how the diffractogram of a sample could look like in experiment for example due varying measurement apperatus, noise
 , sample impurities and lattice defects.
@@
 
-Secondly we would like to also present this dataset as a challenge: The fraction of labeled data can be used to
-easily numerically evalute the performance of the model on a broad range of data as it appears in experiment, which can be used
-to gauge how well models would perform in practice. \\
-But even the unlabeled portion of the dataset can be used to test the performance of models: The least a model
-has to do in order to prove useful in practice is to make a guess about the crystal structure that is good enough
-for Rietveld refinement to converge toward the true structure. \\
-Hence, the performance of the model can be tested by evaluating if the inital guess of the model is sufficient for
-Rietveld refinment to refine the initial guess to a structure that produces a diffractogram matching the one observed in experiment.
+Secondly we would like to also present this dataset as a challenge: The fraction of labeled data can be used to easily numerically evalute the performance of the model on a broad range of data as it appears in experiment, which can be used to gauge how well models would perform in practice.
+\\
+But even the unlabeled portion of the dataset can be used to test the performance of models, if the model attempts to predict the entire crystal structure.
+From the predicted crystal structure, the corresponding diffractogram can be simulated, which can then be compared to the real diffractogram that was used as input for the prediction.
+While a nearly matching diffractogram doesn't necessarily mean a correctly predicted structure, a non-matching diffractogram certainly means an incorrectly predicted structure.
+\\
+
 Finally, by inviting the application of models to data as it appears in experiment we would like to bring attention
@@
 
-% Reference library required and analysis software required, reference library may be incomplete
-% software training required, doesnt always converge or not to true minimum,
-% sometimes sample is truely unknown,  manual labor required in any case
+% TODOS:
+% - This paper talks about the importance of sharing raw powder x-ray diffraction data. I'd definitely cite it ->  \cite{Aranda2018}
+% - The image diff.png shows the difference between a powder diffractogram calculated from single x-ray diffraction data
+% and a real powder diffractogram of the same material. The difference is enormous. I'd include it to highlight
+% that simulating powder diffractino patterns from single crystal data only goes so far
+%https://www.sciencedirect.com/topics/biochemistry-genetics-and-molecular-biology/rietveld-refinement
 
-% -> This is too detailed for this paper, but can use it for thesis
-%For one, under normal circumstances, the initial guess of the structure requires manually or computationally consulting
-%a reference library of pairs of patterns and structures.
-%The reference library is an undesirable overhead that needs to be distributed along with the analysis software and
-%maintained to remain up to date and interoperable with the anlysis software.
-%The industry standard in terms of such reference libraries is the Powder Diffraction File and costs around 1000\$ per
-%year to license from its maintainer, the International Centre for Diffraction Data.
-%Another issue is that since the number of chemically viable crystal structures is infinite any reference library, no
-%matter how
-
-% Crystallographic open database
-% https://www.crystallography.net/cod/
-
-%
-%
-%Machine learning techniques have emerged as a powerful tool in the toolkit of
-%materials scientists. While they are often used to make predictions on
-%the properties of materials or find new materials with certain properties, an increasingly interesting domain is the automated analysis of raw experimental measurements
-%guided by machine learning\supercite{radovicMachineLearningEnergy2018}.
-%
-%With the advent of high-throughput experiments, the amount of gathered data is
-%vast and the analysis often becomes a bottleneck in the processing pipeline
-%\supercite{rahmanianEnablingModularAutonomous2022}. Powder X-ray diffraction
-%(XRD) is an important measurement technique used to obtain structural
-%information from polycrystalline samples\supercite{harrisContemporaryAdvancesUse2001}.
-%The diffractograms are an
-%information-dense fingerprint of the structure of the material. However,
-%analyzing these diffractograms is not an easy task\supercite{holderTutorialPowderXray2019}.
-%Full structure solutions and
-%Rietveld refinement take time and require expert knowledge, both about the
-%analysis technique and the materials class at hand. This is not feasible in high-throughput experiments on a larger
-%scale. Therefore, the question arises whether it is possible to automatically
-%analyze powder diffractograms with machine learning models trained on large
-%amounts of data, making it possible to run inference almost instantaneously.
-%
-%During the last few years, there have been several studies tackling this
-%objective by applying machine learning models to various tasks concerning the
-%analysis of powder diffractograms such as phase
-%classification\supercite{leeDeeplearningTechniquePhase2020,maffettoneCrystallographyCompanionAgent2021,schuetzkeEnhancingDeeplearningTraining2021,szymanskiProbabilisticDeepLearning2021,wangRapidIdentificationXray2020},
-%phase fraction determination\supercite{leeDatadrivenXRDAnalysis2021}, space
-%group
-%classification\supercite{parkClassificationCrystalStructure2017,oviedoFastInterpretableClassification2019,zalogaCrystalSymmetryClassification2020,vecseiNeuralNetworkBased2019,suzukiSymmetryPredictionKnowledge2020,chakrabortyDeepCrystalStructure2022},
-%machine-learning-guided Rietveld
-%refinement\supercite{ozakiAutomatedCrystalStructure2020,fengMethodArtificialIntelligence2019},
-%extraction of lattice
-%parameters\supercite{dongDeepConvolutionalNeural2021,chitturiAutomatedPredictionLattice2021,chakrabortyDeepCrystalStructure2022,
-%habershonPowderDiffractionIndexing2004} and crystallite
-%sizes\supercite{dongDeepConvolutionalNeural2021,chakrabortyDeepCrystalStructure2022},
-%and also novelty detection based on unsupervised techniques\supercite{bankoDeepLearningVisualization2021}.
-%Since an abundant source of experimental diffractograms is hard to come by, most
-%applications train their models on simulated diffractograms from the Inorganic
-%Crystal Structure Database (ICSD)\supercite{bergerhoff1987}, which contains a total of 272\,260 structures (October 2022).
-%
-%\citeauthor{leeDeeplearningTechniquePhase2020} used a deep convolutional neural
-%network (CNN) trained on a large dataset of multiphase compositions from the
-%quaternary \ce{Sr-Li-Al-O} pool to classify present phases in the
-%diffractogram\supercite{leeDeeplearningTechniquePhase2020}. In a follow-up
-%study, they further showed good results for phase fraction inference in the
-%quaternary \ce{Li-La-Zr-O} pool\supercite{leeDatadrivenXRDAnalysis2021}.
-%\citeauthor{schuetzkeEnhancingDeeplearningTraining2021} performed phase
-%classification on iron ores and cement compounds and used data augmentation with
-%respect to lattice parameters, crystallite sizes, and preferred orientation
-%\supercite{schuetzkeEnhancingDeeplearningTraining2021}. They showed that
-%especially the lattice parameter variations enhance the classification accuracy
-%significantly.
-%
-%Instead of the analysis of phase composition,
-%\citeauthor{dongDeepConvolutionalNeural2021} performed regression of scale
-%factors, lattice parameters, and crystallite sizes in a five-phase catalytic
-%materials system\supercite{dongDeepConvolutionalNeural2021}. In contrast to
-%supervised tasks, \citeauthor{bankoDeepLearningVisualization2021} used a
-%variational autoencoder to visualize variations in space group, preferred
-%orientation, crystallite size, and peak
-%shifts\supercite{bankoDeepLearningVisualization2021}.
-%\citeauthor{parkClassificationCrystalStructure2017} used a deep CNN to classify
-%space groups of single-phase diffractograms, reaching a test accuracy of 81.14\%
-%on simulated diffractograms.\supercite{parkClassificationCrystalStructure2017}
-%However, as we will show later in this paper, this accuracy is highly overestimated and drops to 56.1\% when test splits are designed in a way to reduce data leakage in non-IID datasets such as the ICSD.
-%\citeauthor{vecseiNeuralNetworkBased2019}\supercite{vecseiNeuralNetworkBased2019} developed
-%a similar approach and applied their classifier to experimental diffractograms
-%from the RRUFF mineral database\supercite{lafuentePowerDatabasesRRUFF2015},
-%reaching an experimental test accuracy of 54\%.
-
-% PDF pricing
-%https://www.icdd.com/assets/files/2023-Purchase-Options.pdf
+%-----------------------------------------------------------%
+% Sources on rietveld refinement
 

### 2024-04-29 chapters/01_introduction.tex
give ideas how to transition to more accurate simulation

@@
 \\
-As an example of this could be done we can look at results from other domains, where a transfer of core data between different "styles" of representation was attempted:
-%\cite{Ganin2015} -> Attempted to
+As an example of this could be done, we can investigate results from other areas of research, where a transfer of data content between different "styles" of representation was attempted: \\
+Ganin et al.\cite{Ganin2015} proposed a mechanism for training a classifier which performs well independently of which style the input data is represented in, which only requires labeled data from one style and unlabeled data from other styles.
+This corresponds precisely to the situation at hand, where the number of labeled simulated diffractograms is virtually unlimited and a large amount of unlabeled real diffractograms are available, but labeled real diffractogram are in short suppply.
+This algorithm could be applied to train a "style-invariant" structure properties prediction model using a combination of our dataset and simulated labeled patterns.
+\\
+Another approach by Gatys et al.
+\cite{Gatys2016} achieves style transfer of images, by requiring that the content of the output image remain the same, while the style must match that of another image, chosen to represent the target style.
+This procedure could be applied to transfer the style of simulated diffractorams to that of real diffractograms, which could then be used to train the model data that is more faithful to powder diffractograms as they appear in experiment.
 

### 2024-05-11 chapters/00_abstract.tex
Added todos in introduction
Commited updates from Henrik/Pascal

--- 2024-04-28 paper/00_abstract.tex
@@
 Powder X-ray diffraction (pXRD) experiments are a cornerstone for materials structure characterization.
-Despite its widespread application, the analysis of the diffractograms still presents a significant bottleneck
-in high-throughput experimentation that remains to be solved in order to match the increased throughput
-of the prior stages of experimentation.
-%While Rietveld refinement can identify the phases and extract the structural information found
-%in the diffractograms, each sample requires manual inputs and careful configuration of the analysis software
-%from trained staff, which is infeasible for highly automated high-throughput experiments.
+Despite their widespread application, the analysis of pXRD diffractograms still presents a significant automation challenge and thus a bottleneck in high-throughput experimentation that remains to be solved in order to match increasing throughput of automated materials synthesis in (self-driving) labs.
 In the current landscape of powder X-ray diffraction analysis, machine learning has emerged as a promising research direction to resolve this bottleneck by enabling automated powder diffraction analysis.
-A notable difficulty in applying machine learning to this domain is the lack of labeled datasets, i.e.\
-diffraction patterns originating from experiment with accompanying structure solutions, which has relegated machine learning
-researchers to train on simulated data.
-Since these simulations largely fail to accurately reflect the experiment, the performance of models trained
-on this simulated data by and large fails to transfer to experiment and provide value in practice.
-We aim to remedy this by providing a freely available and easily accessible dataset of partially labeled
-powder diffraction data, providing machine learning researchers with a large quantity of real experimental diffractograms collected from a broad range of samples.
-We hope that this can guide machine learning research toward more realistic handling of experimental imperfections to enable the improvement of current automated analysis workflows.
+A notable difficulty in applying machine learning to this domain is the lack of sufficiently sized experimental datasets, which has relegated machine learning researchers to train primarily on simulated data.
+Since simulations largely fail to accurately reflect the experiment, the performance of models trained on only simulated data often lacks transferability to experimental data and thus fails to provide value in practice.
+With the Open Experimental Powder X-Ray Diffraction Database (opXRD), we aim to remedy this by providing an openly available and easily accessible dataset of partially labeled experimental powder diffractograms, providing machine learning researchers with a large quantity of real experimental diffractograms collected from a broad range of samples.
+We provide almost barrier-free software tools that allow experimental researchers to find their data and make it accessible - in virtually any widely used format.
+We collected XXX labeled and YYY unlabeled diffractograms from a wide spectrum of materials classes, which establishes the first version of the opXRD database.
+We hope that this ongoing effort can guide machine learning research toward fully automated analysis of pXRD data and thus enable future self-driving materials labs.

### 2024-05-11 chapters/01_introduction.tex
Commited updates from Henrik/Pascal

--- 2024-04-29 chapters/01_introduction.tex
@@
-\begin{itemize}
-    \item Plan for Figures / Tables
-    \begin{itemize}
-        \item Figure 1: Multiple panels showing histograms of some key metrics: Space group, unit cell volume (normalizable?), ...
-        \item Table 1: List of existing powder XRD datasets
-        \item Table 2: List of existing ML approaches for pXRD
-        \item Table 3: New pXRD datasets offered by this paper
-    \end{itemize}
-
-    \item Additional thoughts
-    \begin{itemize}
-        \item General structure of the paper:
-        \begin{itemize}
-            \item Introduction
-            \item Introduce existing datasets and already show why they are not enough / too small
-            \item Summarize SOTA ML that has been applied to experimental pXRD, discuss the limits and gap between performance on simulated data and exp.
-data (find some good example papers where this is very apparent)
-            \item Summary of our dataset, including paragraphs of the different contributions by other groups; comparison with existing datasets, especially in terms of size
-        \end{itemize} 
-    \end{itemize}
-\end{itemize}
-
 % Autonomous laboratory
@@
 
-The advent of high-throughput experiments, holds the prospect of accelerating the speed materials discovery multiple times over.
-The synthesis and probing of novel materials is becoming increasingly efficient and automated,
-which has correspondingly increased the throughput of samples in these steps of the experimentation pipeline.
+\begin{figure*}[!ht]
+    \centering
+    \includegraphics{figures/overview.pdf}
+    \caption{Overview of the data collection pipeline.
+Datasets are submitted using an online submission form or our submission helper software.
+After post-processing and data homogenization, we create a Zenodo entry for each user submission and subsequently include the submission in the opXRD database.}
+    \label{fig:overview}
+\end{figure*}
+
+The advent of high-throughput experiments holds the prospect of significantly accelerating the speed of materials discovery.
+The synthesis and characterization of novel materials is becoming increasingly efficient and automated,
+which increases the throughput of samples in experimentation pipelines.
 
@@
 % On Rietveld refinement
-The objective of powder x-ray diffraction experiments is to extract as much information about the crystal structure
-of the sample as possible, or if the sample is multi-phase, also identify the phases present in the sample and quantify
-their fractions.
-This is done through a technique called Rietveld refinement: A working model of the crystal structure, starting
-from an initial guess provided by the operator of the analysis software, is assumed and the corresponding diffractogram
-is simulated.
-The parameters of the working model of the sample are then adjusted to minimize the difference between the simulated
-diffraction pattern and the pattern observed in experiment.
-Critically, this process does not generally guarantee to guide the working model towards a global minimum of the difference
-in diffractograms and hence arrive at an accurate model of the sample.
-In general one can only assume the refinement process to guide the working model towards the nearest local minimum
-starting from the initial guess.
-Hence, the result of the refinement procedure is in general only as good as the initial guess it was provided with.
-Still this workflow often works well, since usually the operator of the experiment will be able
-to provide an at least somewhat accurate initial guess of the crystal structure.\\
+After fabricating a new material, a number of analysis techniques can be used to characterize the sample.
+The most important method to determine the crystal structure of a new material is powder X-ray diffraction. 
+%In the case of multi-phase samples, this includes identifying the phases present in the sample and quantifying their fractions.
+To analyze pXRD data, an initial guess is optimized using Rietveld refinement.
+In this process, an initial crystal structure, including unit cell parameters, provided by the operator of the analysis software, is fitted to the observed diffractogram in an iterative optimization process based on simulated forward prediction of expected pXRD diffractograms given the current crystal structure.
+%The unit cell parameters of the crystal structure are adjusted to minimize the difference between the simulated diffraction pattern and the pattern observed in the experiment. 
+As Rietveld refinement is a local optimization method, the result of the refinement procedure is generally only as good as the initial guess it was provided with.
+Therefore, it strongly relies on the experience of the operator to provide an accurate initial guess of the crystal structure.
 
-However, employing manually operated Rietveld Refinement is not without issues.
-It is certainly not scalable to the degree that would be required to keep up with advances
-in throughput and efficiency in other steps of the experimentation pipeline, even just due to the manual effort involved.
-But beyond the lacking scalability of this workflow we also wish to highlight the following issue:
-Providing the analysis software with an initial guess of the structure requires manually or algorithmically
-consulting a reference library of crystal structures with corresponding simulated or sometimes real diffractograms.
-From that reference library a crystal structure must be identified that is likely to be similar
-to that of the investigated sample which can then serve as the initial guess provided to the analysis software.
-Since the number of chemically viable crystal structures is infinite, any such library will always be incomplete.
-The best that can be done to remedy this is to continually update the library both on the side
-of the distributor and the user, which is an undesirable overhead.
-Especially for novel material discovery, it may be impossible to find a sufficiently accurate match in the reference
-library to apply Rietveld Refinement successfully.
+Employing manually operated Rietveld refinement, including determining a good initial structure for the optimization, is not without issues.
+It is certainly not scalable to the degree required to keep up with advances in throughput and efficiency in other steps of the experimentation pipeline.
+Furthermore, initial guesses for Rietveld refinement are often obtained by matching the observed diffractogram to patterns from a database.
+%But beyond the lack of scalability of this workflow we also wish to highlight the following issue:
+%Providing the analysis software with an initial guess of the structure requires manually or algorithmically
+%consulting a reference library of crystal structures with corresponding simulated or sometimes real diffractograms.
+%From that reference library, a crystal structure must be identified that is likely to be similar
+%to that of the investigated sample which can then serve as the initial guess provided to the analysis software.
+%Since the number of chemically viable crystal structures is infinite, any such library will always be incomplete.
+However, in the case of novel observed structure types, simply matching to a database of known structures will not yield good results.
+Therefore, more complex crystal structure solution approaches need to be applied to find suitable crystal structure guesses.
+%The best that can be done to remedy this is to continually update the library both on the side
+%of the distributor and the user, which is an undesirable overhead.
+%Especially for novel material discovery, it may be impossible to find a sufficiently accurate match in the reference
+%library to apply Rietveld Refinement successfully.
 
-Machine learning is a very promising candidate to eliminate both the need for manual inputs and the need for reference
-libraries in powder x-ray diffraction analysis, resulting in a streamlined and automated analysis process that can
-keep pace with an automated high-throughput experimentation environment. \\
-
-Models can be trained to predict crystal structure properties given only the diffractogram itself, given that
-the model can be trained on dataset consisting of pairs of diffractograms and corresponding crystal structures.
-So far however, due to an absence of sufficiently large labeled datasets with diffractograms from experiment,
-machine learning in this domain has largely relied on simulated diffractograms.
-Note, that reference libraries which are used for Rietveld Refinement by in large use simulated diffractograms
-rather than diffractograms from experiment and hence cannot either provide a dataset with diffractograms coming
-from experiment. \\
+Machine learning has the potential to speed up the manual analysis of powder diffractograms and keep pace with an automated high-throughput experimentation environment.
+%eliminate both the need for manual inputs and the need for reference
+%libraries in powder x-ray diffraction analysis, resulting in a streamlined and automated analysis process that can
+%keep pace with an automated high-throughput experimentation environment.
+Models can be trained to predict crystal structure information directly given a diffractogram, or models can be used to automate the conventional workflow of finding an initial guess and running ML-automated Rietveld refinement.
+%However, the current limiting factor is the availability of sufficiently large labeled datasets.
+So far, however, due to an absence of labeled datasets with experimental diffractograms, machine learning in this domain has largely relied on diffractograms simulated from known structures (such as from the ICSD) or, most recently, from generated synthetic crystals.
 
 Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting
-lattice parameters, spacegroup and crystallite size from simulated diffractograms but the performance drops off immensely when these models are applied to data originating from experiment.
+lattice parameters, spacegroup, and crystallite size from simulated diffractograms.
+However, the performance substantially drops off when these models are applied to data originating from experiments [cite 2-3 papers, incl. ours].
+The patterns observed in experiments suffer from a host of imperfections that need to be modeled accurately in order to be able to transfer models trained on simulated data to experimental patterns.
+This includes noise, background, preferred orientation of the sample, ...
+Most approaches currently model these imperfections as a simple signal based on a heuristic functional form added to the simulated diffractograms.
 
-The patterns observed in experiment suffer from a host of imperfections that may result in a partial loss of information or make extracting the structural information more difficult. \\
-However, there is still reason to believe that the strong performance observed in the domain of purely simulated diffractograms can to a large extent translate to predicting crystal structure properties from diffractograms originating from experiment, given that models are trained on a dataset that accurately reflects diffractograms as they appear in experiment.
+At the moment, the number of openly available datasets for experimental powder data is very limited.
+The main sources for experimental patterns are XXX, YYY, and ZZZ.
+In this publication, we provide a new dataset that collects a broad range of patterns from experiment, exceeding the currently available amount of openly accessible experimental patterns by XX.
+The dataset has been curated by collecting the accumulated powder data from multiple large research groups and institutions with high-throughput XRD facilities.
 
-However due to the scarcity of labeled diffractograms originating from experiment, collecting a labeled dataset
-of real experimental powder x-ray diffraction data of sufficient size to train a model
-that can be useful in practice is next to impossible and is not the aim of this dataset.
-Rather we hope that this dataset can enable machine learning researchers to evaluate how closely their simulations
-represent actual experimental data, identify what features are missing and modify their simulations algorithms accordingly to arrive at a simulation that nearly perfectly mirrors how data appears in reality.\\
+% How to contribute? This aspect deserves a separate paragraph
+We further provide an online platform that can be used to easily submit new pXRD data to the dataset, which will be used to extend the existing dataset in the future (Figure~\ref{fig:overview}).
+TODO: Make this longer?
 
-There are roughly speaking two paths towards improving the simulation: The first is to more accurately model the sample and the processes taking place within it during the simulation of the powder diffraction pattern.
+Collecting a labeled dataset of raw experimental powder X-ray diffraction data of sufficient size to directly train a model that predicts crystal structure information is still out of reach.
+The most advanced ML models currently are trained on approximately x simulated diffractograms, which is x orders of magnitude higher than the initial state of the opXRD database presented here.
+However, our dataset still holds significant value in practice.
+First of all, the dataset can be used to better test and benchmark existing and new automated analysis approaches, which yields a better comparison than when using the very limited currently available datasets, such as the commonly used RRUFF mineral dataset.
+
+Furthermore, we hope that our dataset can enable machine learning researchers to evaluate how closely their simulations
+represent actual experimental data, identify what features are missing, and modify their simulation algorithms accordingly to arrive at a simulation that mirrors how data appears in reality.
+One can bring the simulated patterns closer to experiment by improving the simulation itself by properly modeling more of the observed physical effects such as preferred orientation, peak shapes resulting from different sources, etc.
+Additionally, some of the observed imperfections, such as background and noise, can be modeled using machine learning approaches.
+This is similar to the concept of style transfer\cite{Gatys2016,Ganin2015} found in the domain of machine learning applied to images.
+Here, the objective is to obtain a classification or regression model when only labeled images of one style (simulated data) and unlabeled or partially labeled images of another style (experimental data) are available.
+The broad range of available experimental samples contained in our dataset makes it possible to apply these state-of-the-art ML approaches to the domain of powder XRD analysis.
+We hope that this will yield more advanced automated ML approaches in the future, which can readily be applied in high-throughput experimentation pipelines.
+
+%We hope that this will yield more attempts of ML-guided XRD analysis in the future, where also open challenges such as different angle step sizes, different radiation sources, will be addressed.
+
+%However, there is still reason to believe that the strong performance observed in the domain of purely simulated diffractograms can to a large extent translate to predicting crystal structure properties from diffractograms originating from experiment, given that models are trained on a dataset that accurately reflects diffractograms as they appear in experiment.
+
+%There are roughly speaking two paths towards improving the simulation: The first is to more accurately model the sample and the processes taking place within it during the simulation of the powder diffraction pattern.
 This would for example necessitate modelling absorption of the incoming X-ray radiation and its re-emission in the form of characteristic radiation and modeling the effect of stress and motion within the crystal on the diffractogram.
-\\
-An alternative route to obtaining a more accurate simulation is to try to learn these "imperfections" compared to an idealized simulation using machine learning techniques by extracting them from this dataset.
-\\
-As an example of this could be done we can look at results from other domains, where a transfer of core data between different "styles" of representation was attempted:
-%\cite{Ganin2015} -> Attempted to
+
+%An alternative route to obtaining a more accurate simulation is to try to learn these "imperfections" compared to an idealized simulation using machine learning techniques by extracting them from this dataset.
+
+%As an example of this could be done, we can investigate results from other areas of research, where a transfer of data content between different "styles" of representation was attempted:
+
+%Ganin et al.\cite{Ganin2015} proposed a mechanism for training a classifier which performs well independently of which style the input data is represented in, which only requires labeled data from one style and unlabeled data from other styles.
+This corresponds precisely to the situation at hand, where the number of labeled simulated diffractograms is virtually unlimited and a large amount of unlabeled real diffractograms are available, but labeled real diffractogram are in short suppply.
+This algorithm could be applied to train a "style-invariant" structure properties prediction model using a combination of our dataset and simulated labeled patterns.
+
+%Another approach by Gatys et al.
+\cite{Gatys2016} achieves style transfer of images, by requiring that the content of the output image remain the same, while the style must match that of another image, chosen to represent the target style.
+This procedure could be applied to transfer the style of simulated diffractorams to that of real diffractograms, which could then be used to train the model data that is more faithful to powder diffractograms as they appear in experiment.
 
@@
 
-For a set of given sample properties there is only one corresponding simulated diffractogram, but there may an infinite amount of variations in how the diffractogram of a sample could look like in experiment for example due varying measurement apperatus, noise
-, sample impurities and lattice defects.
-In order for a simulated dataset to closely resemble experiment it is critical that the simulatio also represents these
-statistical variations.
-Determining the nature of these variations is a much a harder task than making individual diffractograms look like they
-originated from experiment and cannot be accomplished by simply training a classifier.
-Nevertheless, we hope that this dataset can at least provide clues of how to take this aspect into account during
-simulation as well.\\
+%For a set of given sample properties there is only one corresponding simulated diffractogram, but there may be an infinite amount of variations in how the diffractogram of a sample could look like in experiment for example due varying measurement apperatus, noise
+%, sample impurities and lattice defects.
+%In order for a simulated dataset to closely resemble experiment it is critical that the simulation also represents these
+%statistical variations.
+%Determining the nature of these variations is a much a harder task than making individual diffractograms look like they
+%originated from experiment and cannot be accomplished by simply training a classifier.
+%Nevertheless, we hope that this dataset can at least provide clues of how to take this aspect into account during
+%simulation as well.
 
-Secondly we would like to also present this dataset as a challenge: The fraction of labeled data can be used to easily numerically evalute the performance of the model on a broad range of data as it appears in experiment, which can be used to gauge how well models would perform in practice.
-\\
-But even the unlabeled portion of the dataset can be used to test the performance of models, if the model attempts to predict the entire crystal structure.
+%But even the unlabeled portion of the dataset can be used to test the performance of models, if the model attempts to predict the entire crystal structure.
 From the predicted crystal structure, the corresponding diffractogram can be simulated, which can then be compared to the real diffractogram that was used as input for the prediction.
 While a nearly matching diffractogram doesn't necessarily mean a correctly predicted structure, a non-matching diffractogram certainly means an incorrectly predicted structure.
-\\
 
-Finally, by inviting the application of models to data as it appears in experiment we would like to bring attention
-to some of the practical difficulties that need to be addressed in applying a model to a wide variety of x-ray diffractograms
-collected from different sources: For example the angle and the step size of the diffractograms will differ between sources, 
-which is something that any model that aims to perform well on this dataset must be equipped to handle.
+%Finally, by inviting the application of models to data as it appears in experiment we would like to bring attention
+%to some of the practical difficulties that need to be addressed in applying a model to a wide variety of X-ray diffractograms
+%collected from different sources: For example the angle and the step size of the diffractograms will differ between sources, 
+%which is something that any model that aims to perform well on this dataset must be equipped to handle.
 

### 2024-05-11 chapters/02_existing_datasets.tex
Added todos in introduction
Commited updates from Henrik/Pascal

--- 2024-04-29 paper/02_existing_datasets.tex
@@
-% Source of png : https://www.sciencedirect.com/topics/biochemistry-genetics-and-molecular-biology/rietveld-refinement
-
 % Sources to mention
@@
 % + 19 k entries with "Experimental raw data digital patterns"; 
-%  Proprietary
+%  Proprietary (% PDF pricing: https://www.icdd.com/assets/files/2023-Purchase-Options.pdf)
 %The PDF itself lists the following data sources:
 
+\paragraph{Existing datasets} TODO: Summarize the existing datasets shortly, including number of entries, etc.
+
 \begin{itemize}
+\item RRUFF
 \item Powder Diffraction File (ICDD)
-\begin{itemize}
-    \item ICDD Powders (00) – International Centre for Diffraction Data Newtown Square, PA, USA
-    \item ICSD (01) – FIZ Karlsruhe – Leibniz Institute for Information Infrastructure (FIZ), Karlsruhe, Germany – Inorganic Crystal Structure Database
-    \item NIST (03) – National Institute of Standards and Technology, Gaithersburg, Maryland, USA
-    \item MPDS (04) – Material Phases Data System, Vitznau, Switzerland – Linus Pauling File (LPF)
-    \item ICDD Single Crystal Data (05) – International Centre for Diffraction Data Newtown Square, PA, USA
-\end{itemize}
+%\begin{itemize}
+%    \item ICDD Powders (00) – International Centre for Diffraction Data Newtown Square, PA, USA
+%    \item ICSD (01) – FIZ Karlsruhe – Leibniz Institute for Information Infrastructure (FIZ), Karlsruhe, Germany – Inorganic Crystal Structure Database
+%    \item NIST (03) – National Institute of Standards and Technology, Gaithersburg, Maryland, USA
+%    \item MPDS (04) – Material Phases Data System, Vitznau, Switzerland – Linus Pauling File (LPF)
+%    \item ICDD Single Crystal Data (05) – International Centre for Diffraction Data Newtown Square, PA, USA
+%    \item TODO: Include how many of those are actually experimental pXRD! If we cannot find this information, we need to ask them directly.
+%\end{itemize}
+\item Scattered experimental datasets on Zenodo (TODO: roughly estimate size)
+\item ...
 \end{itemize}
 
-% ->  COD (https://www.crystallography.net/cod/) -
+% ->  COD (https://www.crystallography.net/cod/)
 % -  513k structures,

### 2024-05-11 chapters/03_SOTA_ML.tex
Commited updates from Henrik/Pascal

\paragraph{ML for pXRD} Several approaches have applied machine learning methods to classification and regression tasks for powder diffractograms.

TODO: We need a short literature review here; needs to also cover papers published after our previous paper!

TODO: briefly discuss here that there are expert models for specific classes of materials which give more information, and more generic models like ours which currently only gives you the space group and not much more; especially focus on how experimental conditions (background, noise, ... is modeled)

TODO: discuss in more detail why domain randomization as we did in the last paper so far failed for background and noise, and thus why we need background and noise statistics from experiment to enhance simulated datasets and transfer them to the experimental data distribution.

### 2024-05-11 chapters/04_our_dataset.tex
Commited updates from Henrik/Pascal

--- 2024-04-28 paper/03_our_dataset.tex
@@
+% Notes:
+% The subsections of this section should be the 
+% individual contributions that we get from labs
+
 % Sharing raw powder x-ray diffraction data: Challenges and benefits https://www.researchgate.net/publication/329256004_Sharing_powder_diffraction_raw_data_Challenges_and_benefits
 
-% The subsections of this section should be the 
-% individual contributions that we get from labs
+\begin{figure*}[!ht]
+    \centering
+    \missingfigure{} 
+    \caption{Statistics, histograms, etc. of our dataset.}
+    \label{fig:statistics}
+\end{figure*}
+
+In collaboration with several other institutions we have collected a dataset of diffractograms stemming from experiment, some of them labeled with corresponding structural information. \\
+Currently, institutions that contributed to our dataset include:
+\begin{itemize}
+    \item Institute of Nanotechnology at Karlsruhe Institute of Technology
+    \item University of Southern California
+    \item Lawrence Berkeley National Laboratory
+\end{itemize}
+
+We are still in the process of growing the dataset and additional contributions are still very welcome.
+To find out more about how to contribute to this dataset, visit our website specially designed to collect this dataset, https://xrd.aimat.science \\

### 2024-05-11 chapters/05_how_to_use.tex
Commited updates from Henrik/Pascal

--- 2024-04-28 paper/03_our_dataset.tex
@@
-% Sharing raw powder x-ray diffraction data: Challenges and benefits https://www.researchgate.net/publication/329256004_Sharing_powder_diffraction_raw_data_Challenges_and_benefits
+TODO: A small tutorial with a few python commands to show how easy it is to use our dataset through the provided API.
 
-% The subsections of this section should be the 
-% individual contributions that we get from labs
+A major advantage of this dataset over other comparable datasets is that it is very easy to handle in python and in PyTorch.
+Our accompanying python library xrdpattern (https://github.com/aimat-lab/xrdpattern) provides a means to read in as much of the dataset as fits in memory simply by using PatternDB.load(data\_dirpath).
+The patterns attribute of this class is a list of the indivdual pattern diffractograms, each of which supports a standardize, plot and to tensor method.
+\\
+The standardize method returns a "standardized" version of the pattern with a fixed angle range, a fixed number of entries and intensities normalized to the $[0,1]$ interval.

### 2024-05-14 chapters/00_abstract.tex
Pushed resolved comments

@@
 Powder X-ray diffraction (pXRD) experiments are a cornerstone for materials structure characterization.
-Despite their widespread application, the analysis of pXRD diffractograms still presents a significant automation challenge and thus a bottleneck in high-throughput experimentation that remains to be solved in order to match increasing throughput of automated materials synthesis in (self-driving) labs.
-In the current landscape of powder X-ray diffraction analysis, machine learning has emerged as a promising research direction to resolve this bottleneck by enabling automated powder diffraction analysis.
+Despite their widespread application, the analysis of pXRD diffractograms still presents a significant challenge to automation and thus a bottleneck in high-throughput experimentation automated materials discovery in self-driving labs.
+Machine learning has emerged as a promising research direction to resolve this bottleneck by enabling automated powder diffraction data analysis.
 A notable difficulty in applying machine learning to this domain is the lack of sufficiently sized experimental datasets, which has relegated machine learning researchers to train primarily on simulated data.

### 2024-05-14 chapters/01_introduction.tex
Pushed resolved comments

@@
 To analyze pXRD data, an initial guess is optimized using Rietveld refinement.
-In this process, an initial crystal structure, including unit cell parameters, provided by the operator of the analysis software, is fitted to the observed diffractogram in an iterative optimization process based on simulated forward prediction of expected pXRD diffractograms given the current crystal structure.
+In this process, an initial crystal structure including unit cell parameters is provided by the operator of the analysis software and fitted to the observed diffractogram in an iterative optimization process based on simulated forward prediction of expected pXRD diffractograms given the current crystal structure.
 %The unit cell parameters of the crystal structure are adjusted to minimize the difference between the simulated diffraction pattern and the pattern observed in the experiment. 
@@
 It is certainly not scalable to the degree required to keep up with advances in throughput and efficiency in other steps of the experimentation pipeline.
-Furthermore, initial guesses for Rietveld refinement are often obtained by matching the observed diffractogram to patterns from a database.
+Automation of the Rietveld refinement requires the automation of finding initial guess structures, which is highly non-trivial, and the automation of the iterative optimization procedure, which can be done using black-box optimization algorithms.
+Initial guesses for Rietveld refinement are often obtained by matching the observed diffractogram to patterns from a database.
 %But beyond the lack of scalability of this workflow we also wish to highlight the following issue:
@@
 The main sources for experimental patterns are XXX, YYY, and ZZZ.
-In this publication, we provide a new dataset that collects a broad range of patterns from experiment, exceeding the currently available amount of openly accessible experimental patterns by XX.
-The dataset has been curated by collecting the accumulated powder data from multiple large research groups and institutions with high-throughput XRD facilities.
+Further commercial datasets such as KKK and LLL exist, but license fees and restrictions in their use to train publishable models or even publish the training data limit their usability.
+Collecting a labeled dataset of raw experimental powder X-ray diffraction data of sufficient size to directly train a model that predicts crystal structure information is likely out of reach.
+The most advanced ML models currently are trained on approximately x simulated diffractograms, which is x orders of magnitude higher than the largest currently curated experimental dataset.
 
-% How to contribute? This aspect deserves a separate paragraph
-We further provide an online platform that can be used to easily submit new pXRD data to the dataset, which will be used to extend the existing dataset in the future (Figure~\ref{fig:overview}).
-TODO: Make this longer?
+However, even small labeled and also unlabeled experimental diffractograms hold significant value in practice.
+First of all, labeled data can be used to better test and benchmark existing and new automated analysis approaches, which yields a better comparison than when using the very limited currently available datasets, such as the commonly used RRUFF mineral dataset.
 
-Collecting a labeled dataset of raw experimental powder X-ray diffraction data of sufficient size to directly train a model that predicts crystal structure information is still out of reach.
-The most advanced ML models currently are trained on approximately x simulated diffractograms, which is x orders of magnitude higher than the initial state of the opXRD database presented here.
-However, our dataset still holds significant value in practice.
-First of all, the dataset can be used to better test and benchmark existing and new automated analysis approaches, which yields a better comparison than when using the very limited currently available datasets, such as the commonly used RRUFF mineral dataset.
-
-Furthermore, we hope that our dataset can enable machine learning researchers to evaluate how closely their simulations
-represent actual experimental data, identify what features are missing, and modify their simulation algorithms accordingly to arrive at a simulation that mirrors how data appears in reality.
+Furthermore, even completely unlabeled data can enable machine learning researchers to evaluate how closely their simulations represent actual experimental data, identify what features are missing, and modify their simulation algorithms accordingly to arrive at a simulation that mirrors how data appears in reality.
 One can bring the simulated patterns closer to experiment by improving the simulation itself by properly modeling more of the observed physical effects such as preferred orientation, peak shapes resulting from different sources, etc.
@@
 Here, the objective is to obtain a classification or regression model when only labeled images of one style (simulated data) and unlabeled or partially labeled images of another style (experimental data) are available.
+
+In this publication, we provide a new open powder X-ray diffraction (opXRD) dataset that collects a broad range of patterns from experiment, exceeding the currently available amount of openly accessible experimental patterns by XX.
+Our opXRD dataset has been curated by collecting the accumulated powder data from multiple large research groups and institutions with high-throughput XRD facilities.
+
+% How to contribute? This aspect deserves a separate paragraph
+The opXRD database is planned as a growing, community-driven initiative.
+The dataset we present here is the first version, but we hope to further increase the dataset size though active engagement with the pXRD community.
+In order to xxx.
+We further provide an online platform that can be used to easily submit new pXRD data to the dataset, which will be used to extend the existing dataset in the future (Figure~\ref{fig:overview}).
+TODO: Make this longer?
+
 The broad range of available experimental samples contained in our dataset makes it possible to apply these state-of-the-art ML approaches to the domain of powder XRD analysis.
@@
 
-% TODOS:
-% - This paper talks about the importance of sharing raw powder x-ray diffraction data. I'd definitely cite it ->  \cite{Aranda2018}
 % - The image diff.png shows the difference between a powder diffractogram calculated from single x-ray diffraction data

### 2024-05-14 chapters/01_introduction.tex
Deleted commented out initial draft material
Deleted unused references
Added Rietveld refinement and high throughput citations

@@
+\begin{figure*}[!ht]
+    \centering
+    \includegraphics{figures/overview.pdf}
+    \caption{Overview of the data collection pipeline.
+Datasets are submitted using an online submission form or our submission helper software.
+After post-processing and data homogenization, we create a Zenodo entry for each user submission and subsequently include the submission in the opXRD database.}
+    \label{fig:overview}
+\end{figure*}
+
 % Autonomous laboratory
@@
 
-\begin{figure*}[!ht]
-    \centering
-    \includegraphics{figures/overview.pdf}
-    \caption{Overview of the data collection pipeline.
-Datasets are submitted using an online submission form or our submission helper software.
-After post-processing and data homogenization, we create a Zenodo entry for each user submission and subsequently include the submission in the opXRD database.}
-    \label{fig:overview}
-\end{figure*}
-
 The advent of high-throughput experiments holds the prospect of significantly accelerating the speed of materials discovery.
 The synthesis and characterization of novel materials is becoming increasingly efficient and automated,
-which increases the throughput of samples in experimentation pipelines.
-
-% Rietveld Refinement (3.9 Profile fitting and rietveld analysis)
-%https://www.sciencedirect.com/topics/biochemistry-genetics-and-molecular-biology/rietveld-refinement
-
-% Rietveld Refinement: Half a century anniversary
-% https://pubs.acs.org/doi/10.1021/acs.cgd.1c00854
+which increases the throughput of samples in experimentation pipelines.\cite{Ozaki2020} \cite{Ludwig2019}
 
@@
 The most important method to determine the crystal structure of a new material is powder X-ray diffraction. 
-%In the case of multi-phase samples, this includes identifying the phases present in the sample and quantifying their fractions.
 To analyze pXRD data, an initial guess is optimized using Rietveld refinement.
-In this process, an initial crystal structure including unit cell parameters is provided by the operator of the analysis software and fitted to the observed diffractogram in an iterative optimization process based on simulated forward prediction of expected pXRD diffractograms given the current crystal structure.
-%The unit cell parameters of the crystal structure are adjusted to minimize the difference between the simulated diffraction pattern and the pattern observed in the experiment. 
+In this process, an initial crystal structure including unit cell parameters is provided by the operator of the analysis software and fitted to the observed diffractogram in an iterative optimization process based on simulated forward prediction of expected pXRD diffractograms given the current crystal structure.\cite{Dinnebier2019} \cite{Cano2021}
 As Rietveld refinement is a local optimization method, the result of the refinement procedure is generally only as good as the initial guess it was provided with.
@@
 Automation of the Rietveld refinement requires the automation of finding initial guess structures, which is highly non-trivial, and the automation of the iterative optimization procedure, which can be done using black-box optimization algorithms.
+\cite{Ozaki2020}
 Initial guesses for Rietveld refinement are often obtained by matching the observed diffractogram to patterns from a database.
@@
 %Since the number of chemically viable crystal structures is infinite, any such library will always be incomplete.
-However, in the case of novel observed structure types, simply matching to a database of known structures will not yield good results.
+However, for novel observed structure types, simply matching to a database of known structures will not yield good results.
 Therefore, more complex crystal structure solution approaches need to be applied to find suitable crystal structure guesses.
@@
 
-%We hope that this will yield more attempts of ML-guided XRD analysis in the future, where also open challenges such as different angle step sizes, different radiation sources, will be addressed.
-
-%However, there is still reason to believe that the strong performance observed in the domain of purely simulated diffractograms can to a large extent translate to predicting crystal structure properties from diffractograms originating from experiment, given that models are trained on a dataset that accurately reflects diffractograms as they appear in experiment.
-
-%There are roughly speaking two paths towards improving the simulation: The first is to more accurately model the sample and the processes taking place within it during the simulation of the powder diffraction pattern.
-This would for example necessitate modelling absorption of the incoming X-ray radiation and its re-emission in the form of characteristic radiation and modeling the effect of stress and motion within the crystal on the diffractogram.
-
-%An alternative route to obtaining a more accurate simulation is to try to learn these "imperfections" compared to an idealized simulation using machine learning techniques by extracting them from this dataset.
-
-%As an example of this could be done, we can investigate results from other areas of research, where a transfer of data content between different "styles" of representation was attempted:
-
-%Ganin et al.\cite{Ganin2015} proposed a mechanism for training a classifier which performs well independently of which style the input data is represented in, which only requires labeled data from one style and unlabeled data from other styles.
-This corresponds precisely to the situation at hand, where the number of labeled simulated diffractograms is virtually unlimited and a large amount of unlabeled real diffractograms are available, but labeled real diffractogram are in short suppply.
-This algorithm could be applied to train a "style-invariant" structure properties prediction model using a combination of our dataset and simulated labeled patterns.
-
-%Another approach by Gatys et al.
-\cite{Gatys2016} achieves style transfer of images, by requiring that the content of the output image remain the same, while the style must match that of another image, chosen to represent the target style.
-This procedure could be applied to transfer the style of simulated diffractorams to that of real diffractograms, which could then be used to train the model data that is more faithful to powder diffractograms as they appear in experiment.
-
-%A first hint of how closely a given simulation reflects experiment could be obtained by
-%training a machine learning model, which attempts to classify between simulated and real diffractograms. \\
-%If the simulation is very accurate, we would expect the classifier to not have few to no features which it can use to distinguish between simulated and real diffractograms and hence perform poorly.
-\\
-%Conversely, if the classifier performs well and is not difficult to train, that would be an indicator that the simulation doesn't reflect experiment sufficiently well. \\
-
-%For a set of given sample properties there is only one corresponding simulated diffractogram, but there may be an infinite amount of variations in how the diffractogram of a sample could look like in experiment for example due varying measurement apperatus, noise
-%, sample impurities and lattice defects.
-%In order for a simulated dataset to closely resemble experiment it is critical that the simulation also represents these
-%statistical variations.
-%Determining the nature of these variations is a much a harder task than making individual diffractograms look like they
-%originated from experiment and cannot be accomplished by simply training a classifier.
-%Nevertheless, we hope that this dataset can at least provide clues of how to take this aspect into account during
-%simulation as well.
-
-%But even the unlabeled portion of the dataset can be used to test the performance of models, if the model attempts to predict the entire crystal structure.
-From the predicted crystal structure, the corresponding diffractogram can be simulated, which can then be compared to the real diffractogram that was used as input for the prediction.
-While a nearly matching diffractogram doesn't necessarily mean a correctly predicted structure, a non-matching diffractogram certainly means an incorrectly predicted structure.
-
-%Finally, by inviting the application of models to data as it appears in experiment we would like to bring attention
-%to some of the practical difficulties that need to be addressed in applying a model to a wide variety of X-ray diffractograms
-%collected from different sources: For example the angle and the step size of the diffractograms will differ between sources, 
-%which is something that any model that aims to perform well on this dataset must be equipped to handle.
-
-% - The image diff.png shows the difference between a powder diffractogram calculated from single x-ray diffraction data
-% and a real powder diffractogram of the same material. The difference is enormous. I'd include it to highlight
-% that simulating powder diffractino patterns from single crystal data only goes so far
-%https://www.sciencedirect.com/topics/biochemistry-genetics-and-molecular-biology/rietveld-refinement
-
 %-----------------------------------------------------------%

### 2024-05-14 chapters/01_introduction.tex
Added machine learning for powder x-ray diffracitno references

@@
 Machine learning has the potential to speed up the manual analysis of powder diffractograms and keep pace with an automated high-throughput experimentation environment.
-%eliminate both the need for manual inputs and the need for reference
-%libraries in powder x-ray diffraction analysis, resulting in a streamlined and automated analysis process that can
-%keep pace with an automated high-throughput experimentation environment.
+\cite{Agrawal2019} \cite{Surdu2023}
 Models can be trained to predict crystal structure information directly given a diffractogram, or models can be used to automate the conventional workflow of finding an initial guess and running ML-automated Rietveld refinement.

### 2024-05-14 chapters/01_introduction.tex
Added references for high-throughput experimentation

@@
 
-% Autonomous laboratory
-% https://www.nature.com/articles/s41586-023-06734-w
-
-% Accelerating material discovery
-% https://www.nature.com/articles/s41524-022-00765-z
-
-%High-throughput experiments facilitate materials innovation: A review
-%https://link.springer.com/article/10.1007/s11431-018-9369-9#:~:text=High%2Dthroughput%20materials%20experiments%20including,at%20a%20fraction%20of%20cost.
-
-The advent of high-throughput experiments holds the prospect of significantly accelerating the speed of materials discovery.
+% High throughput experimentation
+The advent of high-throughput experiments holds the prospect of significantly accelerating the speed of materials discovery. \cite{Liu2019}
 The synthesis and characterization of novel materials is becoming increasingly efficient and automated,
-which increases the throughput of samples in experimentation pipelines.\cite{Ozaki2020} \cite{Ludwig2019}
+increasing the throughput of samples in experimentation pipelines. \cite{MacLeod2019} \cite{Ludwig2019} \cite{Ozaki2020}
 
@@
 Models can be trained to predict crystal structure information directly given a diffractogram, or models can be used to automate the conventional workflow of finding an initial guess and running ML-automated Rietveld refinement.
+\cite{Surdu2023}
 %However, the current limiting factor is the availability of sufficiently large labeled datasets.
@@
 Our opXRD dataset has been curated by collecting the accumulated powder data from multiple large research groups and institutions with high-throughput XRD facilities.
+It contains pXRD patterns from single and multiphase materials from a wide variety of materials classes, including XXX, YYY, and ZZZ.
+Details about the experiments of contributing research groups are discussed in this paper, along with a detailed description of how to contribute further data and how to use the data.
 
@@
 The opXRD database is planned as a growing, community-driven initiative.
-The dataset we present here is the first version, but we hope to further increase the dataset size though active engagement with the pXRD community.
-In order to xxx.
+The dataset we present here is the first version, but we hope to further increase the dataset size through active engagement with the pXRD community.
+Our primary object.
 We further provide an online platform that can be used to easily submit new pXRD data to the dataset, which will be used to extend the existing dataset in the future (Figure~\ref{fig:overview}).

### 2024-05-14 chapters/01_introduction.tex
Included machine learning on Powder Diffraction references

@@
 The synthesis and characterization of novel materials is becoming increasingly efficient and automated,
-increasing the throughput of samples in experimentation pipelines. \cite{MacLeod2019} \cite{Ludwig2019} \cite{Ozaki2020}
+increasing the throughput of samples in experimentation pipelines. \cite{MacLeod2019, Ludwig2019, Ozaki2020}
 
@@
 To analyze pXRD data, an initial guess is optimized using Rietveld refinement.
-In this process, an initial crystal structure including unit cell parameters is provided by the operator of the analysis software and fitted to the observed diffractogram in an iterative optimization process based on simulated forward prediction of expected pXRD diffractograms given the current crystal structure.\cite{Dinnebier2019} \cite{Cano2021}
+In this process, an initial crystal structure including unit cell parameters is provided by the operator of the analysis software and fitted to the observed diffractogram in an iterative optimization process based on simulated forward prediction of expected pXRD diffractograms given the current crystal structure.\cite{Dinnebier2019, Cano2021}
 As Rietveld refinement is a local optimization method, the result of the refinement procedure is generally only as good as the initial guess it was provided with.
@@
 Initial guesses for Rietveld refinement are often obtained by matching the observed diffractogram to patterns from a database.
-%But beyond the lack of scalability of this workflow we also wish to highlight the following issue:
-%Providing the analysis software with an initial guess of the structure requires manually or algorithmically
-%consulting a reference library of crystal structures with corresponding simulated or sometimes real diffractograms.
-%From that reference library, a crystal structure must be identified that is likely to be similar
-%to that of the investigated sample which can then serve as the initial guess provided to the analysis software.
-%Since the number of chemically viable crystal structures is infinite, any such library will always be incomplete.
 However, for novel observed structure types, simply matching to a database of known structures will not yield good results.
 Therefore, more complex crystal structure solution approaches need to be applied to find suitable crystal structure guesses.
-%The best that can be done to remedy this is to continually update the library both on the side
-%of the distributor and the user, which is an undesirable overhead.
-%Especially for novel material discovery, it may be impossible to find a sufficiently accurate match in the reference
-%library to apply Rietveld Refinement successfully.
 
-Machine learning has the potential to speed up the manual analysis of powder diffractograms and keep pace with an automated high-throughput experimentation environment.
-\cite{Agrawal2019} \cite{Surdu2023}
+Machine learning has the potential to speed up the manual analysis of powder diffractograms and keep pace with an automated high-throughput experimentation environment. \cite{Agrawal2019, Surdu2023}
 Models can be trained to predict crystal structure information directly given a diffractogram, or models can be used to automate the conventional workflow of finding an initial guess and running ML-automated Rietveld refinement.
-\cite{Surdu2023}
+\cite{Surdu2023}    \cite{Feng2019}
 %However, the current limiting factor is the availability of sufficiently large labeled datasets.
@@
 Collecting a labeled dataset of raw experimental powder X-ray diffraction data of sufficient size to directly train a model that predicts crystal structure information is likely out of reach.
-The most advanced ML models currently are trained on approximately x simulated diffractograms, which is x orders of magnitude higher than the largest currently curated experimental dataset.
+The most advanced ML models currently are trained on approximately $10^7 - 10^8$ simulated diffractograms, which is x orders of magnitude higher than the largest currently curated experimental dataset and x orders of magnitude larger than the opXRD dataset we present here.
 
@@
 Additionally, some of the observed imperfections, such as background and noise, can be modeled using machine learning approaches.
-This is similar to the concept of style transfer\cite{Gatys2016,Ganin2015} found in the domain of machine learning applied to images.
+This is similar to the concept of style transfer\cite{Gatys2016, Ganin2015} found in the domain of machine learning applied to images.
 Here, the objective is to obtain a classification or regression model when only labeled images of one style (simulated data) and unlabeled or partially labeled images of another style (experimental data) are available.
@@
 The dataset we present here is the first version, but we hope to further increase the dataset size through active engagement with the pXRD community.
-Our primary object.
-We further provide an online platform that can be used to easily submit new pXRD data to the dataset, which will be used to extend the existing dataset in the future (Figure~\ref{fig:overview}).
-TODO: Make this longer?
+Our primary objective is to minimize the effort and thus the barrier to contributing experimental data to the opXRD database.
+Thus, we developed a program that helps to find and share data from e.g.
+pXRD lab computers.
+In detail, users can select their most common pXRD file types, the program lists all files of that type, and users can select or deselect certain folders or files for sharing.
+Selected contributions will be uploaded to opXRD, processed to a common file format, and - if wanted - published on Zenodo on behalf of the contributors, before becoming part of the opXRD database.
+If labels are available, they can be shared with opXRD as well.
+Further details can be found on the opXRD website (Figure~\ref{fig:overview}).
 
-The broad range of available experimental samples contained in our dataset makes it possible to apply these state-of-the-art ML approaches to the domain of powder XRD analysis.
+The broad range of available experimental samples contained in the opXRD v1.0 database makes it possible to apply state-of-the-art ML approaches to the domain of powder XRD analysis.
 We hope that this will yield more advanced automated ML approaches in the future, which can readily be applied in high-throughput experimentation pipelines.

### 2024-05-14 chapters/03_SOTA_ML.tex
Included machine learning on Powder Diffraction references

@@
 is modeled)
+Phase Labeling and Lattice Refinement: https://arxiv.org/abs/2308.07897
+
+TODO: briefly discuss papers about multiphase classification and remixing of diffraction patterns
+Physically-informed Graph-based DRNet (PG-DRNet): https://ieeexplore.ieee.org/abstract/document/10459774/
+https://www.nature.com/articles/s42256-021-00384-1
+non-negative matrix factorization: https://ojs.aaai.org/aimagazine/index.php/aimagazine/article/view/2785
 

### 2024-05-14 chapters/01_introduction.tex
Added reference on sparse experimental data

@@
 
+% Manual vs (ML) automated Refinement
 Employing manually operated Rietveld refinement, including determining a good initial structure for the optimization, is not without issues.
@@
 \cite{Surdu2023}    \cite{Feng2019}
-%However, the current limiting factor is the availability of sufficiently large labeled datasets.
 So far, however, due to an absence of labeled datasets with experimental diffractograms, machine learning in this domain has largely relied on diffractograms simulated from known structures (such as from the ICSD) or, most recently, from generated synthetic crystals.
+\cite{Wang2020, Lee2023}
 
+% Predicting structure information with machine leraning
 Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting

### 2024-05-14 chapters/03_SOTA_ML.tex
Added reference on sparse experimental data

--- 2024-05-11 chapters/03_SOTA_ML.tex
@@
 is modeled)
+Phase Labeling and Lattice Refinement: https://arxiv.org/abs/2308.07897
+crystal-structure phase mapping: https://www.nature.com/articles/s42256-021-00384-1
+
+TODO: briefly discuss papers about multiphase classification and remixing of diffraction patterns
+Non-negative Matrix Factorization 2018: https://link.springer.com/chapter/10.1007/978-3-319-93031-2_4
+non-negative matrix factorization 2018: https://ojs.aaai.org/aimagazine/index.php/aimagazine/article/view/2785
+Physically-informed Graph-based DRNet (PG-DRNet) 2023: https://ieeexplore.ieee.org/abstract/document/10459774/
 

### 2024-05-14 chapters/01_introduction.tex
Included previous work on crystal structure information prediction

@@
 \cite{Surdu2023}    \cite{Feng2019}
-So far, however, due to an absence of labeled datasets with experimental diffractograms, machine learning in this domain has largely relied on diffractograms simulated from known structures (such as from the ICSD) or, most recently, from generated synthetic crystals.
-\cite{Wang2020, Lee2023}
+So far, however, due to an absence of labeled datasets with experimental diffractograms\cite{Wang2020} , machine learning in this domain has largely relied on diffractograms simulated from known structures\cite{Park2017} \cite{Lee2023} or, most recently, from generated synthetic crystals.
+\cite{Schopmans2023}
 

### 2024-05-14 chapters/01_introduction.tex
Removed unused references on rietveld refimenet
Added ICSD, PDF references

@@
 \cite{Ozaki2020}
-Initial guesses for Rietveld refinement are often obtained by matching the observed diffractogram to patterns from a database.
+Initial guesses for Rietveld refinement are often obtained by matching the observed diffractogram to patterns from a database. \cite{Belsky2002, GatesRector2019}
 However, for novel observed structure types, simply matching to a database of known structures will not yield good results.
@@
 Models can be trained to predict crystal structure information directly given a diffractogram, or models can be used to automate the conventional workflow of finding an initial guess and running ML-automated Rietveld refinement.
-\cite{Surdu2023}    \cite{Feng2019}
+\cite{Surdu2023,Feng2019}
 So far, however, due to an absence of labeled datasets with experimental diffractograms\cite{Wang2020} , machine learning in this domain has largely relied on diffractograms simulated from known structures\cite{Park2017} \cite{Lee2023} or, most recently, from generated synthetic crystals.
@@
 
-% Predicting structure information with machine leraning
+% Predicting structure information with machine learning
 Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting
@@
 We hope that this will yield more advanced automated ML approaches in the future, which can readily be applied in high-throughput experimentation pipelines.
-
-%-----------------------------------------------------------%
-% Sources on rietveld refinement
-
-% Computationally intensive -> Probably outdated. If it used to take minutes in 2010
-% it probably only takes seconds by now.
-% https://www.rigaku.com/de/node/707
-
-% A simple solution to the Rietveld refinement recipe problem
-% -> Rietveld refinement requires "expert input"
-% https://www.ncbi.nlm.nih.gov/pmc/articles/PMC10840306/
-
-% Comments -> Why having to do rietveld refinement sucks
-% FullProf tutorial on fitting XRD peaks -> Tedious, requires lots of manual input
-% https://pranabdas.github.io/research/fullprof/
-
-% Topas tutorial -> 44 page tutorial, many boring and potentially frustrating steps, lots of points offer potential for task failure
-% https://topas.webspace.durham.ac.uk/topas_user_menu/
-
-%  Methods and Tutorials – Powder Diffraction -> Overview of rietveld refinement data
-%https://neutrons.ornl.gov/sites/default/files/Methods%20and%20Tutorials%20Powder%20Diffraction.pdf
-
-% Rietveld Refinement from Powder Diffraction Data -> 50 pg tutorial that no one wants to read, overwhelming amout of quantities/figures that needs to be manually looked at in the process
-% https://www.fkf.mpg.de/4112052/cpd26.pdf

### 2024-05-14 chapters/01_introduction.tex
Added machine learning for spacegroup prediction papers copied from Henrik paper

@@
 \cite{Surdu2023,Feng2019}
-So far, however, due to an absence of labeled datasets with experimental diffractograms\cite{Wang2020} , machine learning in this domain has largely relied on diffractograms simulated from known structures\cite{Park2017} \cite{Lee2023} or, most recently, from generated synthetic crystals.
+So far, however, due to an absence of labeled datasets with experimental diffractograms\cite{Wang2020} , machine learning in this domain has largely relied on diffractograms simulated from known structures\cite{Park2017, Lee2023} or, most recently, from generated synthetic crystals.
 \cite{Schopmans2023}
@@
 Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting
-lattice parameters, spacegroup, and crystallite size from simulated diffractograms.
+lattice parameters, spacegroup \cite{Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021}, and crystallite size from simulated diffractograms.
 However, the performance substantially drops off when these models are applied to data originating from experiments [cite 2-3 papers, incl. ours].
@@
 
-Furthermore, even completely unlabeled data can enable machine learning researchers to evaluate how closely their simulations represent actual experimental data, identify what features are missing, and modify their simulation algorithms accordingly to arrive at a simulation that mirrors how data appears in reality.
+Furthermore, even completely unlabeled data can enable machine learning researchers to evaluate how closely their simulations represent experimental data, identify what features are missing, and modify their simulation algorithms accordingly to arrive at a simulation that mirrors how data appears in reality.
 One can bring the simulated patterns closer to experiment by improving the simulation itself by properly modeling more of the observed physical effects such as preferred orientation, peak shapes resulting from different sources, etc.

### 2024-05-14 chapters/03_SOTA_ML.tex
Added machine learning for spacegroup prediction papers copied from Henrik paper

@@
 \paragraph{ML for pXRD} Several approaches have applied machine learning methods to classification and regression tasks for powder diffractograms.
+
+Using simulated diffractograms based on structures from the ICSD, Lee et al.
+managed to train a deep convolutional neural network (CNN) that is able to classify occurring phases in diffractograms of a specific compound pool.
+\cite{Lee2020}
+Park et al. reached a test accuracy of roughly 81\% for a CNN, which classifies space groups of simulated single-phase diffractograms. \cite{Park2017}
+A regression analysis on lattice parameters within a broader framework encompassing all material classes was conducted by Chitturi et al.\cite{chitturi_automated_2021} They developed a distinct CNN for each crystal system, utilizing a merged dataset from both the ICSD and the Cambridge Structural Database, and managed to achieve a mean absolute percentage error of about \SI{10}{\percent} for the lattice lengths, although they encountered difficulties in accurately predicting angles.
 

### 2024-05-14 main.tex
Commited changes by Henrik
Added todos in introduction
Commited updates from Henrik/Pascal
Pushed resolved comments
Added machine learning for spacegroup prediction papers copied from Henrik paper

--- 2024-04-09 paper/_main.tex
@@
 \usepackage[utf8]{inputenc}
-\usepackage[super,sort\&compress,comma]{natbib}
-\renewcommand{\supercite}[1]{\cite{#1}}
+\usepackage[super,sort&compress,comma]{natbib}
+%\renewcommand{\supercite}[1]{\cite{#1}}
 
@@
 
-%% Useful packages
+% Useful packages
 \usepackage{amsmath}
@@
 \interfootnotelinepenalty=10000
-\title{Towards a free high quality powder X-ray diffraction dataset}
+% Changed idea title idea from orignal title of
+% "Towards a free high quality powder x-ray diffractogram database"
+% Name is supposed to be in analogy with Crystallography Open database
+% which is the largest freely available database of crystal structures
+\title{opXRD: Open Experimental Powder X-ray Diffraction Database}
 
@@
 \author[1,2]{Henrik Schopmans}
-\author[1,2]{Pasc1al Friederich}
-\author[2]{Ben Breitung}
+\author[1,2]{Jona Östreicher}
 \author[2]{Simon Schweidler}
-\author[3]{Andrea Hodge}
 \author[3]{Adie Alwen}
-\author[4,5]{Carolin Sutter-Fella}
 \author[4]{Mriganka Singh}
 \author[4]{Tim Kodalle}
+\author[2]{Ben Breitung}
 \author[4]{Maged Abdelsamie}
+\author[4,5]{Carolin Sutter-Fella}
+\author[3]{Andrea Hodge}
+\author[1,2]{Pascal Friederich}
 
@@
 \begin{abstract}
-\input{00_abstract}
+\input{chapters/00_abstract}
 \end{abstract}
 
-\section{Motivation}\label{sec:motivation}
-\input{01_motivation}
+\newpage
+\section{Introduction}\label{sec:Introduction}
+\input{chapters/01_introduction}
 
-\section{Existing datasets}\label{sec:existing_datasets}
-\input{02_existing_datasets}
+\input{chapters/02_existing_datasets}
 
-\section{Our dataset}\label{sec:our_dataset}
-\input{03_our_dataset}
+\input{chapters/03_SOTA_ML}
 
-\section{Contributions}\label{sec:contributions}
-\input{04_contributions}
+\section{opXRD database}\label{sec:our_dataset}
+\input{chapters/04_our_dataset}
+
+\section{How to contribute and how to use}\label{sec:how_to_use}
+\input{chapters/05_how_to_use}
+
+\subsection*{Data availability}
+Data is available on xxx.
+
+\subsection*{Conflicts of interest}
+There are no conflicts of interest to declare.
+
+\subsection*{Acknowledgements}
+We acknowledge financial support by the German Research Foundation (DFG) through the Research Training Group 2450 “Tailored Scale-Bridging Approaches to Computational Nanoscience”.
+We acknowledge support by the Federal Ministry of Education and Research (BMBF) under Grant No.
+01DM21001B (German-Canadian Materials Acceleration Center).
+We acknowledge financial support from the Helmholtz Foundation Model Initiative within Project "SOL-AI".
 
@@
 
-%\listofchanges
-%\clearpage
-
-%\nocite{*}
-%\TheBibliography
-%\printbibliography[heading=bibintoc]
-
-\bibliography{bib}
-\bibliographystyle{rsc} %the RSC's .bst file
+\bibliography{references}
+\bibliographystyle{rsc}
 

### 2024-05-14 chapters/01_introduction.tex
Added structure information prediction references

@@
 Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting
-lattice parameters, spacegroup \cite{Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021}, and crystallite size from simulated diffractograms.
+lattice parameters\cite{Dong2021, Chitturi2021, Habershon2004}, spacegroup \cite{Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021}, and crystallite size \cite{Dong2021, Chakraborty2021}from simulated diffractograms.
 However, the performance substantially drops off when these models are applied to data originating from experiments [cite 2-3 papers, incl. ours].

### 2024-05-14 chapters/03_SOTA_ML.tex
Added structure information prediction references

--- 2024-05-14 chapters/03_SOTA_ML.tex
@@
 \paragraph{ML for pXRD} Several approaches have applied machine learning methods to classification and regression tasks for powder diffractograms.
+
+Using simulated diffractograms based on structures from the ICSD, Lee et al.
+managed to train a deep convolutional neural network (CNN) that is able to classify occurring phases in diffractograms of a specific compound pool.
+\cite{Lee2020}
+Park et al. reached a test accuracy of roughly 81\% for a CNN, which classifies space groups of simulated single-phase diffractograms. \cite{Park2017}
+A regression analysis on lattice parameters within a broader framework encompassing all material classes was conducted by Chitturi et al.\cite{Chitturi2021} They developed a distinct CNN for each crystal system, utilizing a merged dataset from both the ICSD and the Cambridge Structural Database, and managed to achieve a mean absolute percentage error of about \SI{10}{\percent} for the lattice lengths, although they encountered difficulties in accurately predicting angles.
 

### 2024-06-09 chapters/01_introduction.tex
Commited first paragraph in "Our dataset" section from external author

--- 2024-05-14 chapters/01_introduction.tex
@@
 As Rietveld refinement is a local optimization method, the result of the refinement procedure is generally only as good as the initial guess it was provided with.
-Therefore, it strongly relies on the experience of the operator to provide an accurate initial guess of the crystal structure.
+%Therefore, it strongly relies on the experience of the operator to provide an accurate initial guess of the crystal structure.
 
@@
 \cite{Ozaki2020}
-Initial guesses for Rietveld refinement are often obtained by matching the observed diffractogram to patterns from a database. \cite{Belsky2002, GatesRector2019}
-However, for novel observed structure types, simply matching to a database of known structures will not yield good results.
+Initial guesses for Rietveld refinement are often obtained by matching the observed diffractogram to patterns from a database.
+Commonly used databases include the "Powder Diffraction File"\cite{GatesRector2019}, the "Inorganic Crystal Structure Database" \cite{Belsky2002} and the "Crystallography Open Database" \cite{Grazulis2009}.
+An initial guess obtained from such a database is by no means guaranteed to lead to an accurate structure solution, even if it is iteratively revised, either manually or using blackbox optimization.\cite{Ozaki2020}
+Especially for novel observed structure types, simply matching to a database of known structures will not yield good results.
 Therefore, more complex crystal structure solution approaches need to be applied to find suitable crystal structure guesses.
-
 Machine learning has the potential to speed up the manual analysis of powder diffractograms and keep pace with an automated high-throughput experimentation environment. \cite{Agrawal2019, Surdu2023}
-Models can be trained to predict crystal structure information directly given a diffractogram, or models can be used to automate the conventional workflow of finding an initial guess and running ML-automated Rietveld refinement.
-\cite{Surdu2023,Feng2019}
-So far, however, due to an absence of labeled datasets with experimental diffractograms\cite{Wang2020} , machine learning in this domain has largely relied on diffractograms simulated from known structures\cite{Park2017, Lee2023} or, most recently, from generated synthetic crystals.
+Models can be trained to predict crystal structure information directly given a diffractogram, or models can be used to automate the conventional workflow of finding an initial guess \cite{Surdu2023} and running ML-automated Rietveld refinement.
+\cite{Feng2019}
+So far, however, due to an absence of labeled datasets with experimental diffractograms\cite{Wang2020}, machine learning in this domain has largely relied on diffractograms simulated from known structures\cite{Park2017, Lee2023} or, most recently, from generated synthetic crystals.
 \cite{Schopmans2023}
@@
 Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting
-lattice parameters, spacegroup \cite{Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021}, and crystallite size from simulated diffractograms.
+lattice parameters\cite{Dong2021, Chitturi2021, Habershon2004}, spacegroup \cite{Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021}, and crystallite size \cite{Dong2021, Chakraborty2021} from simulated diffractograms.
 However, the performance substantially drops off when these models are applied to data originating from experiments [cite 2-3 papers, incl. ours].
@@
 First of all, labeled data can be used to better test and benchmark existing and new automated analysis approaches, which yields a better comparison than when using the very limited currently available datasets, such as the commonly used RRUFF mineral dataset.
+\cite{Armbruster2015}
 

### 2024-06-09 chapters/03_SOTA_ML.tex
Commited first paragraph in "Our dataset" section from external author

@@
 managed to train a deep convolutional neural network (CNN) that is able to classify occurring phases in diffractograms of a specific compound pool.
-\cite{Lee2020}
+\cite{Lee2020} Building upon their work they also developed models based on fully convolutional neural networks and transformer encoders, that predict the crystal system and the spacegroup and other structural properties, such as the band gap \cite{Lee2022}.
+With their best model for the crystal system prediction on ICSD structures, they achieved a test accuracy of \SI{92.2}{\percent}.
 Park et al. reached a test accuracy of roughly 81\% for a CNN, which classifies space groups of simulated single-phase diffractograms. \cite{Park2017}

### 2024-06-09 chapters/04_our_dataset.tex
Pushed resolved comments
Deleted commented out initial draft material
Included machine learning on Powder Diffraction references
Commited first paragraph in "Our dataset" section from external author

@@
-% Notes:
-% The subsections of this section should be the 
+% TODOS:
+% - This paper talks about the importance of sharing raw powder X-ray diffraction data. I'd definitely cite it ->  \cite{Aranda2018}
+
+%- The subsections of this section should be the 
 % individual contributions that we get from labs
 
-% Sharing raw powder x-ray diffraction data: Challenges and benefits https://www.researchgate.net/publication/329256004_Sharing_powder_diffraction_raw_data_Challenges_and_benefits
+%------------------------------------------$
 
@@
 
+%% Paragraph by FX Coudert and Arthur Hardiagon
+%% !! work in progress !!
+
+We extracted experimental pXRD data from the Crystallography Open Database (COD).\cite{Grazulis2009, Vaitkus2023}  The COD is, to our knowledge, the largest open-access collection of experimental crystal structures of organic, inorganic, metal-organic compounds and minerals, containing more than 500,000 entries.\footnote{Available online at \url{https://www.crystallography.net/cod/}} The data in the COD is placed in the public domain and licensed under the CC0 License.
+Of the entire COD database, 5432 structures (1\% of the database) contained at least one tag from the {CIF\_POW} dictionary, i.e., relating to powder diffraction studies --- which is expected, since most crystal structures are resolved from single crystal diffraction.
+From these 5432 files, most contained only metadata related to the powder diffraction experiment, but did not include the raw data of the pattern itself.
+We could extract raw, experimental pXRD patterns from 1063 files in total, after curation of a small number of files with clearly invalid data.
+
+The pXRD data from the COD database are of high quality, with a median resolution of $\Delta(2\theta) = 0.013$\textdegree and an average number of 9190 points measured per pattern.
+
+% todo : short analysis of the data
+
+%%
+
 We are still in the process of growing the dataset and additional contributions are still very welcome.
-To find out more about how to contribute to this dataset, visit our website specially designed to collect this dataset, https://xrd.aimat.science \\
+To find out more about how to contribute to this dataset, visit our website specially designed to collect this dataset, https://xrd.aimat.science .

### 2024-06-20 main.tex
Commited first paragraph in "Our dataset" section from external author
Commited external author contribution

@@
 \affil[1]{Institute of Theoretical Informatics, Karlsruhe Institute of Technology,\protect \\
-Engler-Bunte-Ring 8, 76131 Karlsruhe, Germany}
-\affil[2]{Institute of Nanotechnology, Karlsruhe Institute of Technology,\protect \\ Hermann-von-Helmholtz-Platz 1, 76344 Eggenstein-Leopoldshafen, Germany}
+Kaiserstr. 12, 76131 Karlsruhe, Germany}
+\affil[2]{Institute of Nanotechnology, Karlsruhe Institute of Technology,\protect \\ Kaiserstr. 12, 76131 Karlsruhe, Germany}
 \affil[3]{Department of Aerospace and Mechanical Engineering, University of Southern California, \protect \\3650 McClintock Ave, Los Angeles, CA 90089, USA}
@@
 
+\section{Existing datasets}
 \input{chapters/02_existing_datasets}
 
+\section{Machine learning for pXRD}
 \input{chapters/03_SOTA_ML}

### 2024-07-05 chapters/04_our_dataset.tex
Commited external author contribution
Test commit
Commited INT contribution paragraph

@@
 %% Paragraph by FX Coudert and Arthur Hardiagon
-%% !! work in progress !!
+%% Chimie Paris Tech, PSL University, CNRS, Institut de Recherche de Chimie Paris, 75005 Paris, France
 
@@
 The pXRD data from the COD database are of high quality, with a median resolution of $\Delta(2\theta) = 0.013$\textdegree and an average number of 9190 points measured per pattern.
-
-% todo : short analysis of the data
+They span a wide chemical space, including organic, inorganic and hybrid structures, including 75 different elements of the periodic table.
 
@@
 
+%% Institute of Theoretical Informatics, Karlsruhe Institute of Technology,
+%Kaiserstr. 12, 76131 Karlsruhe, Germany
+
+The X-ray diffraction data were collected on a Bruker D8 Advance (Cu Kα1 radiation, λ = 1.54056 Å).
+The samples were initially recorded for various research projects over the last ten years and were measured with different step sizes, times per step, and over different angle ranges, but all using Cu Kα1 radiation.
+The samples mostly contained transition metal oxides, sulfides, and fluorides.
+A major part of the research focused on high-entropy materials, which involved incorporating many different elements into single-phase structures, leading to peak shifts or phase separations.
+Most of those multicomponent complex materials appeared in various structures, including rock-salt, spinel, fluorite, perovskite, and delafossite, to name a few.The samples were almost always prepared in powder form; therefore, powder XRD was performed on samples with adjusted height.
+To improve statistics, the samples were rotated during the entire measurement.
+Some air-sensitive samples were measured using a transparent polymer dome for protection.
+This dome led to increased background noise over the first 20° and slightly decreased pattern resolution.
+The samples were prepared using various synthesis techniques, mostly solid-state or wet chemical syntheses, to obtain the desired structures.
+Consequently, particle size and crystallinity varied significantly.
+The sample set also includes samples that were not successfully measured or where phases could not be identified.
+
 We are still in the process of growing the dataset and additional contributions are still very welcome.

### 2024-07-05 main.tex
Corrected sutter-fella affiliation

@@
 \author[4]{Maged Abdelsamie}
-\author[4,5]{Carolin Sutter-Fella}
+\author[4]{Carolin Sutter-Fella}
 \author[3]{Andrea Hodge}
@@
 \affil[2]{Institute of Nanotechnology, Karlsruhe Institute of Technology,\protect \\ Kaiserstr. 12, 76131 Karlsruhe, Germany}
-\affil[3]{Department of Aerospace and Mechanical Engineering, University of Southern California, \protect \\3650 McClintock Ave, Los Angeles, CA 90089, USA}
-\affil[4]{Department of Chemical Engineering and Materials Science, University of Southern California,\protect \\ 925 Bloom Walk, HED 216, Los Angeles, CA 90089, USA}
-\affil[5]{Lawrence Berkeley National Laboratory, Molecular Foundry Division,\protect \\ 1 Cyclotron Rd., Berkeley, 94720 CA, USA}
+\affil[3]{Department of Chemical Engineering and Materials Science, University of Southern California,\protect \\ 925 Bloom Walk, HED 216, Los Angeles, CA 90089, USA}
+\affil[4]{Lawrence Berkeley National Laboratory, Molecular Foundry Division,\protect \\ 1 Cyclotron Rd., Berkeley, 94720 CA, USA}
 

### 2024-11-04 chapters/01_introduction.tex
Updated author + affiliations list

@@
 It is certainly not scalable to the degree required to keep up with advances in throughput and efficiency in other steps of the experimentation pipeline.
-Automation of the Rietveld refinement requires the automation of finding initial guess structures, which is highly non-trivial, and the automation of the iterative optimization procedure, which can be done using black-box optimization algorithms.
-\cite{Ozaki2020}
+Automation of the Rietveld refinement requires the automation of finding initial guess structures, which is highly non-trivial, and the automation of the iterative optimization procedure, which can be done using black-box optimization algorithms \cite{Ozaki2020}.
 Initial guesses for Rietveld refinement are often obtained by matching the observed diffractogram to patterns from a database.
-Commonly used databases include the "Powder Diffraction File"\cite{GatesRector2019}, the "Inorganic Crystal Structure Database" \cite{Belsky2002} and the "Crystallography Open Database" \cite{Grazulis2009}.
+Commonly used databases include the Powder Diffraction File\cite{GatesRector2019}, the Inorganic Crystal Structure Database \cite{Belsky2002}, and the Crystallography Open Database \cite{Grazulis2009}.
 An initial guess obtained from such a database is by no means guaranteed to lead to an accurate structure solution, even if it is iteratively revised, either manually or using blackbox optimization.\cite{Ozaki2020}
@@
 Machine learning has the potential to speed up the manual analysis of powder diffractograms and keep pace with an automated high-throughput experimentation environment. \cite{Agrawal2019, Surdu2023}
-Models can be trained to predict crystal structure information directly given a diffractogram, or models can be used to automate the conventional workflow of finding an initial guess \cite{Surdu2023} and running ML-automated Rietveld refinement.
-\cite{Feng2019}
-So far, however, due to an absence of labeled datasets with experimental diffractograms\cite{Wang2020}, machine learning in this domain has largely relied on diffractograms simulated from known structures\cite{Park2017, Lee2023} or, most recently, from generated synthetic crystals.
-\cite{Schopmans2023}
+Models can be trained to predict crystal structure information directly given a diffractogram, or models can be used to automate the conventional workflow of finding an initial guess \cite{Surdu2023} and running ML-automated Rietveld refinement \cite{Feng2019}.
+So far, however, due to an absence of labeled datasets with experimental diffractograms\cite{Wang2020}, machine learning in this domain has largely relied on diffractograms simulated from known structures\cite{Park2017, Lee2023} or, most recently, from generated synthetic crystals \cite{Schopmans2023}
 
 % Predicting structure information with machine learning
-Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting
-lattice parameters\cite{Dong2021, Chitturi2021, Habershon2004}, spacegroup \cite{Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021}, and crystallite size \cite{Dong2021, Chakraborty2021} from simulated diffractograms.
-However, the performance substantially drops off when these models are applied to data originating from experiments [cite 2-3 papers, incl. ours].
+Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting lattice parameters\cite{Dong2021, Chitturi2021, Habershon2004}, spacegroup \cite{Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021}, and crystallite size \cite{Dong2021, Chakraborty2021} from simulated diffractograms.
+However, the performance substantially drops off when these models are applied to data originating from experiments \cite{Schopmans2023}[cite 2-3 papers, incl. ours].
 The patterns observed in experiments suffer from a host of imperfections that need to be modeled accurately in order to be able to transfer models trained on simulated data to experimental patterns.
@@
 Collecting a labeled dataset of raw experimental powder X-ray diffraction data of sufficient size to directly train a model that predicts crystal structure information is likely out of reach.
-The most advanced ML models currently are trained on approximately $10^7 - 10^8$ simulated diffractograms, which is x orders of magnitude higher than the largest currently curated experimental dataset and x orders of magnitude larger than the opXRD dataset we present here.
+The most advanced ML models currently are trained on approximately $10^7 - 10^8$ simulated diffractograms, which is - to the best of our knowledge - three orders of magnitude larger than the largest currently curated experimental dataset and even x orders of magnitude larger than the number of unlabeled diffractograms in the opXRD dataset we present here.
 

### 2024-11-04 chapters/04_our_dataset.tex
Updated author + affiliations list

@@
 
-The pXRD data from the COD database are of high quality, with a median resolution of $\Delta(2\theta) = 0.013$\textdegree and an average number of 9190 points measured per pattern.
+The pXRD data from the COD database are of high quality, with a median resolution of $\Delta(2\theta) = 0.013${\textdegree} and an average number of 9190 points measured per pattern.
 They span a wide chemical space, including organic, inorganic and hybrid structures, including 75 different elements of the periodic table.

### 2024-11-04 main.tex
Updated author + affiliations list

@@
 \author[4]{Maged Abdelsamie}
+\author[4]{Gregoire Heymans}
+\author[4]{Hanlin Hu}
+\author[5]{Alexander Wieczorek}
+\author[5]{Siarhei Zhuk}
+\author[5]{Sebastiann Siol}
+\author[6]{Moritz Wolf}
+\author[7]{Arthur Hardiagon}
+\author[8]{Bin Cao}
+\author[8]{Tong-yi Zhang}
 \author[4]{Carolin Sutter-Fella}
@@
 
-\affil[1]{Institute of Theoretical Informatics, Karlsruhe Institute of Technology,\protect \\
-Kaiserstr. 12, 76131 Karlsruhe, Germany}
-\affil[2]{Institute of Nanotechnology, Karlsruhe Institute of Technology,\protect \\ Kaiserstr. 12, 76131 Karlsruhe, Germany}
-\affil[3]{Department of Chemical Engineering and Materials Science, University of Southern California,\protect \\ 925 Bloom Walk, HED 216, Los Angeles, CA 90089, USA}
-\affil[4]{Lawrence Berkeley National Laboratory, Molecular Foundry Division,\protect \\ 1 Cyclotron Rd., Berkeley, 94720 CA, USA}
+\affil[1]{Institute of Theoretical Informatics, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany}
+\affil[2]{Institute of Nanotechnology, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany}
+\affil[3]{Department of Chemical Engineering and Materials Science, University of Southern California, Los Angeles CA 90089, USA}
+\affil[4]{Lawrence Berkeley National Laboratory, Molecular Foundry Division, Berkeley 94720 CA, USA}
+\affil[5]{Empa-Swiss Federal Laboratories
+for Materials Science and Technology, CH-8600 Dübendorf,
+Switzerland}
+\affil[6]{Engler-Bunte-Institut \& Institute of Catalysis
+Research and Technology, Karlsruhe Institute of Technology, Karlsruhe, Germany}
+\affil[7]{Chimie ParisTech, PSL University, CNRS, Institut de Recherche de Chimie Paris, 75005 Paris, France}
+\affil[8]{Guangzhou Municipal Key Laboratory of Materials Informatics,  Hong Kong University of
+Science and Technology (Guangzhou), Guangzhou 511400, Guangdong, China}
 
@@
 
-\section{Existing datasets}
+\*section{Existing datasets}
 \input{chapters/02_existing_datasets}
 
-\section{Machine learning for pXRD}
+\*section{Machine learning for pXRD}
 \input{chapters/03_SOTA_ML}

### 2024-11-05 chapters/01_introduction.tex
Added paragraphs in introduction

@@
-\begin{figure*}[!ht]
+\begin{figure*}[!htb]
     \centering
@@
 % High throughput experimentation
-The advent of high-throughput experiments holds the prospect of significantly accelerating the speed of materials discovery. \cite{Liu2019}
-The synthesis and characterization of novel materials is becoming increasingly efficient and automated,
-increasing the throughput of samples in experimentation pipelines. \cite{MacLeod2019, Ludwig2019, Ozaki2020}
+The advent of high-throughput experiments holds the prospect of significantly accelerating the speed of materials discovery.
+\cite{Liu2019} The synthesis and characterization of novel materials are becoming increasingly efficient and automated, increasing the throughput of samples in experimentation pipelines.
+\cite{MacLeod2019, Ludwig2019, Ozaki2020} \\
 
@@
 After fabricating a new material, a number of analysis techniques can be used to characterize the sample.
-The most important method to determine the crystal structure of a new material is powder X-ray diffraction. 
-To analyze pXRD data, an initial guess is optimized using Rietveld refinement.
-In this process, an initial crystal structure including unit cell parameters is provided by the operator of the analysis software and fitted to the observed diffractogram in an iterative optimization process based on simulated forward prediction of expected pXRD diffractograms given the current crystal structure.\cite{Dinnebier2019, Cano2021}
-As Rietveld refinement is a local optimization method, the result of the refinement procedure is generally only as good as the initial guess it was provided with.
-%Therefore, it strongly relies on the experience of the operator to provide an accurate initial guess of the crystal structure.
+One method that can be used to determine the crystal structure of a new material is powder X-ray diffraction.
+To determine crystal structure from pXRD data, an initial guess is optimized using Rietveld refinement.
+In this process, an initial crystal structure including unit cell parameters is provided by the operator of the analysis software and fitted to the observed diffractogram in an iterative optimization process based on simulated forward prediction of expected pXRD diffractograms given the current crystal structure.
+\cite{Dinnebier2019, Cano2021} As Rietveld refinement is a local optimization method, the result of the refinement procedure is only as good as the initial guess it was provided with.
+\\
 
@@
 Commonly used databases include the Powder Diffraction File\cite{GatesRector2019}, the Inorganic Crystal Structure Database \cite{Belsky2002}, and the Crystallography Open Database \cite{Grazulis2009}.
-An initial guess obtained from such a database is by no means guaranteed to lead to an accurate structure solution, even if it is iteratively revised, either manually or using blackbox optimization.\cite{Ozaki2020}
-Especially for novel observed structure types, simply matching to a database of known structures will not yield good results.
+An initial guess obtained from such a database is by no means guaranteed to lead to an accurate structure solution, even if it is iteratively revised, either manually or using blackbox optimization.\cite{Ozaki2020} Especially for novel observed structure types, simply matching to a database of known structures will not yield good results.
 Therefore, more complex crystal structure solution approaches need to be applied to find suitable crystal structure guesses.
+\\
+
 Machine learning has the potential to speed up the manual analysis of powder diffractograms and keep pace with an automated high-throughput experimentation environment. \cite{Agrawal2019, Surdu2023}
 Models can be trained to predict crystal structure information directly given a diffractogram, or models can be used to automate the conventional workflow of finding an initial guess \cite{Surdu2023} and running ML-automated Rietveld refinement \cite{Feng2019}.
-So far, however, due to an absence of labeled datasets with experimental diffractograms\cite{Wang2020}, machine learning in this domain has largely relied on diffractograms simulated from known structures\cite{Park2017, Lee2023} or, most recently, from generated synthetic crystals \cite{Schopmans2023}
+So far, however, due to an absence of labeled datasets with experimental diffractograms\cite{Wang2020}, machine learning in this domain has largely relied on diffractograms simulated from known structures\cite{Park2017, Lee2023} or, most recently, from generated synthetic crystals.
+\cite{Schopmans2023} \\
 
@@
 The most advanced ML models currently are trained on approximately $10^7 - 10^8$ simulated diffractograms, which is - to the best of our knowledge - three orders of magnitude larger than the largest currently curated experimental dataset and even x orders of magnitude larger than the number of unlabeled diffractograms in the opXRD dataset we present here.
+\\
 
@@
 First of all, labeled data can be used to better test and benchmark existing and new automated analysis approaches, which yields a better comparison than when using the very limited currently available datasets, such as the commonly used RRUFF mineral dataset.
-\cite{Armbruster2015}
+\cite{Armbruster2015} \\
 
@@
 Here, the objective is to obtain a classification or regression model when only labeled images of one style (simulated data) and unlabeled or partially labeled images of another style (experimental data) are available.
+\\
 
@@
 Details about the experiments of contributing research groups are discussed in this paper, along with a detailed description of how to contribute further data and how to use the data.
+\\
 

### 2024-11-05 chapters/02_existing_datasets.tex
Commited external author contribution
Added listing of all found Crystallography databases
Test commit
Commited INT contribution paragraph
Added paragraphs in introduction

@@
 % Crystal structures with corresponding simulated diffractograms;
-% ->  ICSD (https://icsd.products.fiz-karlsruhe.de/) 
+% ->  ICSD 
 % - 230k structures;
@@
 
-% -> PDF (https://www.icdd.com/) -> 
+% -> PDF  -> 
 %- 1200k "material data sets" (structures?) 
@@
 
-\paragraph{Existing datasets} TODO: Summarize the existing datasets shortly, including number of entries, etc.
+% ->  COD (https://www.crystallography.net/cod/)
+% -  513k structures,
+% - Fully open
+
+% Datset listing (Done on: 20.06.24)
+% I would list here *all* crystallography databases but evaluate/determine that they lack or don't provide convenient access to raw powder diffraction data
+
+% Some other Lists of Crystallographic databases
+% - IUCR: "Crystallographyic databases and related resources" (https://www.iucr.org/resources/data/databases)
+% - Bruno et al. (2017): "Crystallography and Databases" (https://datascience.codata.org/articles/10.5334/dsj-2017-038)
+% - University of Virginia "Crystallographic data" (https://guides.lib.virginia.edu/c.php?g=514782&p=3568808)
+
+%->1 PDF5+ (https://www.icdd.com/,https://www.icdd.com/pdf-5/)
+% - Data sources of the PDF
+% (1.1): ICDD Powders
+% (1.2): FIZ ICSD (https://icsd.products.fiz-karlsruhe.de/)
+% (1.3): NIST ICSD https://icsd.nist.gov/
+% (1.4): Linus Pauling File (https://paulingfile.com/index.php?p=about%20us)
+% (1.5): ICDD Single Crystal data
+%-> 2 COD (https://www.crystallography.net/cod/)
+%-> 3 RRUFF (American Mineralogist Crystal Structure Database)
+%-> 4 CCDC (https://www.ccdc.cam.ac.uk/)
+%-> 5 The Material Project (https://legacy.materialsproject.org/)
+%-> 6 Crystal Lattice Structures (https://www.atomic-scale-physics.de/lattice/)
+%-> 7 Crystallographic and Crystallochemical Database for Minerals and their Structural Analogues (https://database.iem.ac.ru/mincryst/)
+%-> 8 Mineralogy Database (https://webmineral.com/)
+%-> 9 IUCr Raw data letters (https://iucrdata.iucr.org/)
+%-> 10 Bilbao Crystallographic server  (https://www.cryst.ehu.es/)
+%-> 11 PowBase (http://www.cristal.org/powbase/index.html)
+%-> 12 Athena (https://athena.unige.ch/athena/mineral/mineral.html)
+%-> 13 Protein database (https://www.rcsb.org/)
+%-> 14 SDPD (Structure Determination from Powder Diffraction - Database)
+%-> 15 Crystal Met (https://cds.dl.ac.uk/cds/datasets/crys/mdf/llmdf.html)
+%-> 16 NAKB (https://www.nakb.org/)
+%-> 17 Pearson's Crystal data (https://www.crystalimpact.com/pcd/)
+%-> 18 Biological macromolecule Crystallization Database (http://bmcd.ibbr.umd.edu/)
+
+%TODO: Also include plans for NOMAD and highlight their efforts
+
+%TODO: Summarize the existing datasets shortly, including number of entries, etc.
+
+\paragraph{Existing powder diffraction datasets} To contextualize opXRD within the current environment of experimental powder diffraction data, the following list provides an overview of the largest crystal structure databases that include powder diffraction data.
 
 \begin{itemize}
+\item \textbf{Powder Diffraction File:} \footnote{More information on the Powder Diffraction File can be found under \url{https://www.icdd.com/pdf-5/}.} The Powder Diffraction File (PDF), published and maintained by the International Center for diffraction data, is currently the largest collection of powder diffraction data available to researchers.
+Though its latest release contains over a million entries, almost all are simulated patterns.
+Only 19000 entries represent powder diffraction data originating from experiments, and of those, only about 11000 come with accompanying atomic coordinates.
+It is widely used as a reference database in Rietveld refinement to identify crystal structures from which the refinement process can begin based on their powder diffraction patterns.
+As of November 2024, the latest release of the PDF, the PDF5+, is available to researchers through a purchase starting at a price point of \$6,265.00.
+
+\item \textbf{Crystallography Open Database:}\footnote{More information on the crystallography open database can be found under \url{https://www.crystallography.net/cod/}.} The Crystallography Open Database (COD) is an open-access collection of crystal structures.
+It provides over 300.000 crystal structures in the form of .cif files.
+Of these files, 1063 contained the experimental powder diffraction data that was used to determine the underlying crystal structures of the investigated samples.
+
+\item Scattered experimental datasets on Zenodo (TODO: roughly estimate size)
 \item RRUFF
-\item Powder Diffraction File (ICDD)
-%\begin{itemize}
-%    \item ICDD Powders (00) – International Centre for Diffraction Data Newtown Square, PA, USA
-%    \item ICSD (01) – FIZ Karlsruhe – Leibniz Institute for Information Infrastructure (FIZ), Karlsruhe, Germany – Inorganic Crystal Structure Database
-%    \item NIST (03) – National Institute of Standards and Technology, Gaithersburg, Maryland, USA
-%    \item MPDS (04) – Material Phases Data System, Vitznau, Switzerland – Linus Pauling File (LPF)
-%    \item ICDD Single Crystal Data (05) – International Centre for Diffraction Data Newtown Square, PA, USA
-%    \item TODO: Include how many of those are actually experimental pXRD! If we cannot find this information, we need to ask them directly.
-%\end{itemize}
-\item Scattered experimental datasets on Zenodo (TODO: roughly estimate size)
 \item ...
@@
 
-% ->  COD (https://www.crystallography.net/cod/)
-% -  513k structures,
-% - Fully open
+Some major crystal structure databases, such as the ICSD, do not appear in the above list, as they do not include powder diffraction data.
+This is because most structure solutions are achieved through single-crystal diffraction rather than powder diffraction.

### 2024-11-05 main.tex
Added paragraphs in introduction

@@
 \affil[8]{Guangzhou Municipal Key Laboratory of Materials Informatics,  Hong Kong University of
-Science and Technology (Guangzhou), Guangzhou 511400, Guangdong, China}
+Science and Technology (Guangzhou), Guangzhou 511400 Guangdong, China}
 

### 2024-11-21 chapters/00_abstract.tex
Added subsections in dataset paragraphs

@@
 A notable difficulty in applying machine learning to this domain is the lack of sufficiently sized experimental datasets, which has relegated machine learning researchers to train primarily on simulated data.
-Since simulations largely fail to accurately reflect the experiment, the performance of models trained on only simulated data often lacks transferability to experimental data and thus fails to provide value in practice.
+Since simulations largely fail to accurately reflect the experiment, the performance of models trained on only simulated data often lacks transferability to experimental data and thus fails to provide value in practice.2
 With the Open Experimental Powder X-Ray Diffraction Database (opXRD), we aim to remedy this by providing an openly available and easily accessible dataset of partially labeled experimental powder diffractograms, providing machine learning researchers with a large quantity of real experimental diffractograms collected from a broad range of samples.

### 2024-11-21 chapters/01_introduction.tex
Added subsections in dataset paragraphs

@@
 % Manual vs (ML) automated Refinement
-Employing manually operated Rietveld refinement, including determining a good initial structure for the optimization, is not without issues.
-It is certainly not scalable to the degree required to keep up with advances in throughput and efficiency in other steps of the experimentation pipeline.
-Automation of the Rietveld refinement requires the automation of finding initial guess structures, which is highly non-trivial, and the automation of the iterative optimization procedure, which can be done using black-box optimization algorithms \cite{Ozaki2020}.
-Initial guesses for Rietveld refinement are often obtained by matching the observed diffractogram to patterns from a database.
-Commonly used databases include the Powder Diffraction File\cite{GatesRector2019}, the Inorganic Crystal Structure Database \cite{Belsky2002}, and the Crystallography Open Database \cite{Grazulis2009}.
+Manually performing Rietveld refinement is time-consuming and often requires expert knowledge.
+It is not scalable to the degree required to keep up with advances in throughput and efficiency in other steps of the experimentation pipeline.
+In a Rietveld refinement process, the operator starts by setting parameters that characterize the background and peak shape of the powder diffraction pattern and determining a crystal structure from which the optimization process can start.
+The latter is usually done using search-match software, which identifies crystal structures with similar powder diffraction patterns from a database of crystal structures with accompanying powder diffraction patterns.
+However, attempting to refine all crystal structure parameters at once, starting from this initial structure, is known to lead to unphysical results.
+What makes the refinement process time-consuming and difficult is finding the correct order in which to refine the crystal structure parameters.
+\cite{Ozaki2020}  Additionally, if the fit develops into an unphysical set of parameters in this iterative determination of parameters, the operator must intervene to backtrack to the last step.
+\\
+
+Commonly used databases for obtaining a match include the Powder Diffraction File\cite{GatesRector2019}, the Inorganic Crystal Structure Database \cite{Belsky2002}, and the Crystallography Open Database \cite{Grazulis2009}.
 An initial guess obtained from such a database is by no means guaranteed to lead to an accurate structure solution, even if it is iteratively revised, either manually or using blackbox optimization.\cite{Ozaki2020} Especially for novel observed structure types, simply matching to a database of known structures will not yield good results.
@@
 % Predicting structure information with machine learning
-Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting lattice parameters\cite{Dong2021, Chitturi2021, Habershon2004}, spacegroup \cite{Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021}, and crystallite size \cite{Dong2021, Chakraborty2021} from simulated diffractograms.
-However, the performance substantially drops off when these models are applied to data originating from experiments \cite{Schopmans2023}[cite 2-3 papers, incl. ours].
+Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting lattice parameters\cite{Dong2021, Chitturi2021, Habershon2004, zhang2024crystallographic}, spacegroup \cite{Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021,zhang2024crystallographic}, and crystallite size \cite{Dong2021, Chakraborty2021} from simulated diffractograms.
+However, the performance substantially drops off when these models are applied to data originating from experiments \cite{Schopmans2023,zhang2024crystallographic}[cite 2-3 papers, incl. ours].
 The patterns observed in experiments suffer from a host of imperfections that need to be modeled accurately in order to be able to transfer models trained on simulated data to experimental patterns.

### 2024-11-21 chapters/02_existing_datasets.tex
Added subsections in dataset paragraphs

@@
 
-\item \textbf{Crystallography Open Database:}\footnote{More information on the crystallography open database can be found under \url{https://www.crystallography.net/cod/}.} The Crystallography Open Database (COD) is an open-access collection of crystal structures.
-It provides over 300.000 crystal structures in the form of .cif files.
+\item \textbf{Crystallography Open Database:}\footnote{More information on the crystallography open database can be found under \url{https://www.crystallography.net/cod/}.} The Crystallography Open Database (COD)\cite{Graulis2009cod} is an open-access collection of crystal structures.
+It provides over 500.000 crystal structures in the form of .cif files.
 Of these files, 1063 contained the experimental powder diffraction data that was used to determine the underlying crystal structures of the investigated samples.
@@
 \item Scattered experimental datasets on Zenodo (TODO: roughly estimate size)
-\item RRUFF
+\item \textbf{RRUFF}: The RRUFF Mineral Database \cite{Armbruster2015} provides detailed information on minerals, including their chemical compositions, crystallography, and spectroscopic data.
+Managed by the University of Arizona, it was created to serve as a public repository for mineral identification and research.
+It contains \num{908} distinct structures with experimental x-ray diffraction patterns.
 \item ...

### 2024-11-21 chapters/03_SOTA_ML.tex
Added subsections in dataset paragraphs

@@
 A regression analysis on lattice parameters within a broader framework encompassing all material classes was conducted by Chitturi et al.\cite{Chitturi2021} They developed a distinct CNN for each crystal system, utilizing a merged dataset from both the ICSD and the Cambridge Structural Database, and managed to achieve a mean absolute percentage error of about \SI{10}{\percent} for the lattice lengths, although they encountered difficulties in accurately predicting angles.
+Zhang and Cao et al.\cite{zhang2024crystallographic} introduce a convolutional self-attention neural network trained on simulated patterns to classify crystal types.
+Their model was tested on 23,073 unary, binary, and ternary inorganic crystal structures sourced from the Crystallography Open Database.
+The study observed a noticeable performance drop when the pretrained model was applied to real experimental patterns, as opposed to simulated data.
+However, their recent work \cite{cao2024simxrd} proposes using convolutional peak descriptors that consider the detector's geometry, which significantly reduces the performance gap in their benchmark tests.
+The option of training purely on experimental diffractograms unfortunately is unfeasible because of the limited availability of labeled experimental diffractograms, but Salgado et al.
+\cite{Salgado2023} showed, that adding a fraction of experimental patterns to a simulated training dataset improves the performance on unseen experimental patterns.
+They used \SI{50}{\percent} of the experimental patterns contained in the RRUFF database and added those to their large simulated training set.
+Then they tested their model's performance on the other half of the RRUFF database and achieved a performance increase in the 230-way spacegroup classification accuracy of \num{11} percentage points compared to the same model only trained on simulated patterns.
 
@@
 TODO: briefly discuss papers about multiphase classification and remixing of diffraction patterns
-Non-negative Matrix Factorization 2018: https://link.springer.com/chapter/10.1007/978-3-319-93031-2_4
-non-negative matrix factorization 2018: https://ojs.aaai.org/aimagazine/index.php/aimagazine/article/view/2785
-Physically-informed Graph-based DRNet (PG-DRNet) 2023: https://ieeexplore.ieee.org/abstract/document/10459774/
+Non-negative Matrix Factorization 2018: \url{https://link.springer.com/chapter/10.1007/978-3-319-93031-2_4}
+non-negative matrix factorization 2018: \url{https://ojs.aaai.org/aimagazine/index.php/aimagazine/article/view/2785}
+Physically-informed Graph-based DRNet (PG-DRNet) 2023: \url{https://ieeexplore.ieee.org/abstract/document/10459774/}
 
 TODO: discuss in more detail why domain randomization as we did in the last paper so far failed for background and noise, and thus why we need background and noise statistics from experiment to enhance simulated datasets and transfer them to the experimental data distribution.
+
+The ultimate goal of this ML technology is to make PXRD data identification practical for experimental use, whether for identifying space group symmetry or phase composition.
+However, the current volume of experimental PXRD patterns is insufficient to effectively train ML models, highlighting an urgent need for a comprehensive experimental PXRD database.
+To address this, two key approaches are essential.
+First, developing more sophisticated simulation methods to better approximate experimental patterns\cite{cao2024simxrd}.
+Second, creating an experimental database that enables transfer learning to bridge the gap between simulated and real-world data.
+The development of opXRD is particularly significant, as it will provide a comprehensive experimental benchmark for the community, allowing fair comparison of baseline models and accurate evaluation of their applicability in real experimental situations.

### 2024-11-21 chapters/04_our_dataset.tex
Added subsections in dataset paragraphs

@@
 
-In collaboration with several other institutions we have collected a dataset of diffractograms stemming from experiment, some of them labeled with corresponding structural information. \\
-Currently, institutions that contributed to our dataset include:
+In collaboration with several other institutions we have collected a dataset of diffractograms stemming from experiment, some of them labeled with corresponding structural information.
+
 \begin{itemize}
@@
     \item Lawrence Berkeley National Laboratory
+    \item Hong Kong University of Science and Technology (Guangzhou)
+    \item Swiss Federal Laboratories for Materials Science and Technology
 \end{itemize}
@@
 %% Paragraph by FX Coudert and Arthur Hardiagon
-%% Chimie Paris Tech, PSL University, CNRS, Institut de Recherche de Chimie Paris, 75005 Paris, France
+\subsection*{Chimie Paris Tech, PSL University}
 
-We extracted experimental pXRD data from the Crystallography Open Database (COD).\cite{Grazulis2009, Vaitkus2023}  The COD is, to our knowledge, the largest open-access collection of experimental crystal structures of organic, inorganic, metal-organic compounds and minerals, containing more than 500,000 entries.\footnote{Available online at \url{https://www.crystallography.net/cod/}} The data in the COD is placed in the public domain and licensed under the CC0 License.
-Of the entire COD database, 5432 structures (1\% of the database) contained at least one tag from the {CIF\_POW} dictionary, i.e., relating to powder diffraction studies --- which is expected, since most crystal structures are resolved from single crystal diffraction.
-From these 5432 files, most contained only metadata related to the powder diffraction experiment, but did not include the raw data of the pattern itself.
+We extracted experimental pXRD data from the Crystallography Open Database (COD).\cite{Grazulis2009, Vaitkus2023} The COD is, to our knowledge, the largest open-access collection of experimental crystal structures of organic, inorganic, and metal-organic compounds and minerals, containing more than 500,000 entries.
+\footnote{Available online at \url{https://www.crystallography.net/cod/}.} The data in the COD is placed in the public domain and licensed under the CC0 License.
+Of the entire COD database, 5432 structures 1\% of the database) contained at least one tag from the {CIF\_POW} dictionary, i.e., relating to powder diffraction studies --- which is expected, since most crystal structures are resolved from single crystal diffraction.
+Of these 5432 files, most contained only metadata related to the powder diffraction experiment, but did not include the raw data of the pattern itself.
 We could extract raw, experimental pXRD patterns from 1063 files in total, after curation of a small number of files with clearly invalid data.
+\\
 
 The pXRD data from the COD database are of high quality, with a median resolution of $\Delta(2\theta) = 0.013${\textdegree} and an average number of 9190 points measured per pattern.
-They span a wide chemical space, including organic, inorganic and hybrid structures, including 75 different elements of the periodic table.
+They span a wide chemical space, including organic, inorganic, and hybrid structures, including 75 different elements of the periodic table.
 
-%%
+%% Paragraph by Ben Breitung
+\subsection*{Institute of Nanotechnology, KIT}
 
-%% Institute of Theoretical Informatics, Karlsruhe Institute of Technology,
-%Kaiserstr. 12, 76131 Karlsruhe, Germany
-
-The X-ray diffraction data were collected on a Bruker D8 Advance (Cu Kα1 radiation, λ = 1.54056 Å).
-The samples were initially recorded for various research projects over the last ten years and were measured with different step sizes, times per step, and over different angle ranges, but all using Cu Kα1 radiation.
+The X-ray diffraction data were collected on a Bruker D8 Advance (Cu K$\alpha$1 radiation, $\lambda$ = 1.54056 Å).
+The samples were initially recorded for various research projects over the last ten years and were measured with different step sizes, times per step, and over different angle ranges, but all using Cu K$\alpha$1 radiation.
 The samples mostly contained transition metal oxides, sulfides, and fluorides.
 A major part of the research focused on high-entropy materials, which involved incorporating many different elements into single-phase structures, leading to peak shifts or phase separations.
-Most of those multicomponent complex materials appeared in various structures, including rock-salt, spinel, fluorite, perovskite, and delafossite, to name a few.The samples were almost always prepared in powder form; therefore, powder XRD was performed on samples with adjusted height.
+Most of those multi-component complex materials appeared in various structures, including rock-salt, spinel, fluorite, perovskite, and delafossite, to name a few.The samples were almost always prepared in powder form; therefore, powder XRD was performed on samples with adjusted height.
 To improve statistics, the samples were rotated during the entire measurement.
@@
 
+% Paragraph by Tim Kodalle
+\subsection*{Lawrence Berkeley National Lab}
+
+Experimental XRD data were collected at beamline 12.3.2 of the Advanced Light Source, the synchrotron at Lawrence Berkeley National Laboratory, using a photon energy of 10 keV (corresponding to a wavelength of $\lambda$ = 1.23984193 Å), selected using a Si(111) monochromator.
+Measurements were taken in grazing incidence geometry, i.e.
+using a beam incidence angle of 1$^{\circ}$.
+Two-dimensional diffraction images were recorded using a Dectris Pilatus 1M area detector at an angle between 34$^{\circ}$ and 36$^{\circ}$ with a sample-to-detector distance of roughly 190 mm.
+The two-dimensional data were calibrated using an Al$_{2}$O$_{3}$ calibration standard and integrated along the azimuthal angle.
+Data collection was performed in situ during thin-film deposition using a custom-made spin-coating and annealing stage.
+One dataset was collected spin-coating triple cation metal-halide perovskite precursor solutions with the composition Cs$_{0.05}$(MA$_{0.23}$FA$_{0.77}$)Pb$_{1.1}$(I$_{0.77}$Br$_{0.23}$)$_{3}$ (MA = Methylammonium, FA = Formamidinium) onto various substrates, including glass (amorphous), GaAs wafers (single crystalline) as well as stacks of glass/ITO, GaAs/Mo/Cu(In,Ga)Se$_{2}$/CdS/ZnO, and glass/Mo/Cu(In,Ga)Se$_{2}$/CdS/ZnO (glass/CIGS).
+Some of the substrates were additionally covered with a self-assembling monolayer of MeO-2PACz.
+The GaAs substrates were prepared by Dr.
+Jiro Nishinaga from the National Institute of Advanced Industrial Science and Technology (AIST) in Japan and the glass/CIGS substrates by Dr.
+Christian Kaufmann and his team at Helmholtz-Zentrum Berlin (HZB) in Germany.
+\\
+
+A second dataset was collected from spin-coating metal-halide perovskite precursor solutions with varying compositions of \ce{MAPb(I_{1-x}Br_x)3} (where MA = Methylammonium and x = 0, 0.33, 0.5, 0.67, 1) spin-coated onto glass substrates.
+The substrates were preheated to different temperatures (30°C, 50°C, 70°C, and 90°C), and the spin-coating process was performed at a constant temperature on the preheated substrates.
+For both datasets, diffraction data were continuously measured during spin-coating, chemical induction of crystallization, and annealing of the samples (at 100$^{\circ}$C and 110$^{\circ}$C respectively) with a frequency of about 0.56 1/s and  0.54 1/s.
+Each in situ measurement consisted of about 500 to 1000 individual diffractograms.
+Depending on the substrate, each series of diffractograms shows an evolution from substrate only to a combination of polycrystalline perovskite, PbI$_{2}$ and substrate via several intermediate phases.
+
+% Paragraph by Bin Cao
+\subsection*{Guangzhou Municipal Key Laboratory of Materials Informatics, HKUST}
+
+In the past two years, we established a small-scale experimental PXRD database called the X-Ray Phase Identification Public Experimental Dataset (XRed,  \href{https://github.com/WPEM/XRED}{https://github.com/WPEM/XRED}).
+The data in XRed is generated in our lab using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers under copper targets.
+Hundreds of PXRD patterns have been refined, labeled with CIF files, and organized by elemental systems.
+The dataset includes original experimental files across single-phase to five-phase mixtures.
+This project is progressing toward integration with the comprehensive opXRD database to establish a large-scale, long-term experimental resource.
+
+% Paragraph by Moritz Wolf and Tony-yi 
+\subsection*{Institute of Catalysis Research and Technology, KIT}
+
+X-ray diffraction (XRD) was conducted with an X’Pert Pro MPD (Panalytical) in Bragg-Brentano geometry using a copper X-ray source ($\lambda_{K_{\alpha}} = 1.5406 Å$).
+The patterns were acquired from 5-80° $2\theta$ with a step size of 0.016711° or 0.033420° and a total acquisition time of 40 to 120 min.
+A variety of samples were analyzed including commercial catalysts, bulk reference materials, porous metal oxide particles, and nanoparticles.
+The latter were synthesized via the surfactant-free benzyl alcohol route (10.1039/C9DT01634A, 10.1016/j.matchemphys.2018.04.021).
+The cobalt oxide (CoO or Co3O4) and cerium oxide (CeO2) nanoparticles were in the size range of 4-16 nm according to the Scherrer equation.
+A series of porous Al2O3 materials, which were prepared by calcination of boehmite (AlOOH) at various temperatures, represents crystalline samples with limited long-range structure and various contributions of Al2O3 polymorphs.
+
+% Paragraph by Alexander Wieczorek and Sebastiann Siol
+\subsection*{Laboratory for Surface Science and Coating Technologies, Empa}
+
+XRD data was measured using a Bruker D8 Discover equipped with a Cu~K$\alpha$ radiation source ($\lambda$~=~1.54056~{\AA} for Cu~K$\alpha$\textsubscript{1} and $\lambda$~=~1.54439~{\AA} for Cu~K$\alpha$\textsubscript{2}) in a Bragg-Brentano geometry.
+For the reported data sets the instrument was equipped with a Goebel mirror effectively removing the Cu~K$\beta$ radiation.
+The data set originates from the combinatorial exploration of the Zn–V–N compositional space, as well as data gathered from multiple research activities on more established metal halide perovskite semiconductors.
+All data was collected from thin films deposited on borosilicate glass.
+The Zn–V–N films showed some preferential out of plane orientation, while for the perovskites the preferential orientation was minimal, resulting in the presence of all reflections.
+Each combinatorial Zn–V–N library was synthesized using radio-frequency co-sputtering of Zn and V in a mixed Ar and N\textsubscript{2} plasma.
+An orthogonal deposition temperature and composition gradient was created, resulting in a deposition temperature of 220~{\textdegree}C for samples 1~–~9 and 114~{\textdegree}C for samples 37~–~45.
+The composition for each sample was determined using X-Ray fluorescence (XRF) spectroscopy which was further calibrated through Rutherford backscattering spectroscopy (RBS) based on selected samples.
+The newly identified and isolated semiconductor Zn\textsubscript{2}VN\textsubscript{3} was identified to exhibit a cation-disordered wurtzite structure as verified by additional GI-XRD and SAED measurements.\cite{Zhuk2021}
+Tin halide perovskites were deposited using single-step spin-coating as reported elsewhere.\cite{Wieczorek2023} Methylammonium lead iodide libraries with varying degrees of residual PbI\textsubscript{2} were deposited using a two-step procedure involving both thermal evaporation of PbI\textsubscript{2} and subsequent spin-coating of a methylammonium solution.
+The relative phase fractions were quantified using supplementary azimuthal angle scans coupled with structural factors and geometrical factors as reported elsewhere.\cite{Wieczorek2024} Fully inorganic lead perovskite libraries were prepared using thermal co-evaporation of lead and cesium halide salts.
+All metal halide perovskite libraries were measured within a custom-made X-Ray transparent inert-gas dome, resulting in the presence of minor additional features within the $\theta$~=~19~–~31{\textdegree} range.
+For all combinatorial libraries where any phases are specified, the complete set of phases are reported in the meta data.
+
 We are still in the process of growing the dataset and additional contributions are still very welcome.
-To find out more about how to contribute to this dataset, visit our website specially designed to collect this dataset, https://xrd.aimat.science .
+To find out more about how to contribute to this dataset, visit our website specially designed to collect this dataset, \url{https://xrd.aimat.science} .

### 2024-11-21 chapters/05_how_to_use.tex
Added subsections in dataset paragraphs

@@
 A major advantage of this dataset over other comparable datasets is that it is very easy to handle in python and in PyTorch.
-Our accompanying python library xrdpattern (https://github.com/aimat-lab/xrdpattern) provides a means to read in as much of the dataset as fits in memory simply by using PatternDB.load(data\_dirpath).
+Our accompanying python library xrdpattern (\url{https://github.com/aimat-lab/xrdpattern}) provides a means to read the dataset simply by using PatternDB.load(data\_dirpath).
 The patterns attribute of this class is a list of the indivdual pattern diffractograms, each of which supports a standardize, plot and to tensor method.
@@
 The standardize method returns a "standardized" version of the pattern with a fixed angle range, a fixed number of entries and intensities normalized to the $[0,1]$ interval.
+
+% caobin
+We have launched Xqueryer (http://xqueryer.caobin.asia/), a free online platform for phase identification.
+Users can upload experimental PXRD data to identify and match structures from the Materials Project database.
+Uploaded data is stored and integrated into our opXRD database.

### 2024-11-21 main.tex
Added subsections in dataset paragraphs

@@
 \author[4]{Mriganka Singh}
-\author[4]{Tim Kodalle}
+\author[4,5]{Tim Kodalle}
 \author[2]{Ben Breitung}
-\author[4]{Maged Abdelsamie}
-\author[4]{Gregoire Heymans}
-\author[4]{Hanlin Hu}
-\author[5]{Alexander Wieczorek}
-\author[5]{Siarhei Zhuk}
-\author[5]{Sebastiann Siol}
-\author[6]{Moritz Wolf}
-\author[7]{Arthur Hardiagon}
-\author[8]{Bin Cao}
-\author[8]{Tong-yi Zhang}
-\author[4]{Carolin Sutter-Fella}
-\author[3]{Andrea Hodge}
+\author[11, 12]{Maged Abdelsamie}
+\author[6]{Gregoire Heymans}
+\author[7]{Alexander Wieczorek}
+\author[7]{Siarhei Zhuk}
+\author[7]{Sebastian Siol}
+\author[8]{Moritz Wolf}
+\author[9]{François-Xavier Coudert}
+\author[9]{Arthur Hardiagon}
+\author[10]{Bin Cao}
+\author[13]{Ruth Schwaiger}
+\author[10]{Tong-yi Zhang}
+\author[4]{Carolin M. Sutter-Fella}
+\author[3]{Andrea M. Hodge}
 \author[1,2]{Pascal Friederich}
@@
 \affil[4]{Lawrence Berkeley National Laboratory, Molecular Foundry Division, Berkeley 94720 CA, USA}
-\affil[5]{Empa-Swiss Federal Laboratories
-for Materials Science and Technology, CH-8600 Dübendorf,
+\affil[5]{Lawrence Berkeley National Laboratory, Advanced Light Source, Berkeley 94720 CA, USA}
+\affil[6]{Lawrence Berkeley National Laboratory, Chemical Sciences Division, Berkeley 94720 CA, USA}
+\affil[7]{Empa~–~Swiss Federal Laboratories
+for Materials Science and Technology, 8600 Dübendorf,
 Switzerland}
-\affil[6]{Engler-Bunte-Institut \& Institute of Catalysis
+\affil[8]{Engler-Bunte-Institut \& Institute of Catalysis
 Research and Technology, Karlsruhe Institute of Technology, Karlsruhe, Germany}
-\affil[7]{Chimie ParisTech, PSL University, CNRS, Institut de Recherche de Chimie Paris, 75005 Paris, France}
-\affil[8]{Guangzhou Municipal Key Laboratory of Materials Informatics,  Hong Kong University of
-Science and Technology (Guangzhou), Guangzhou 511400 Guangdong, China}
-
+\affil[9]{Chimie ParisTech, PSL University, CNRS, Institut de Recherche de Chimie Paris, 75005 Paris, France}
+\affil[10]{Guangzhou Municipal Key Laboratory of Materials Informatics,  Hong Kong University of
+Science and Technology (Guangzhou), Guangzhou 511400, China}
+\affil[11]{Material Science and Engineering Department, King Fahd University of Petroleum and Minerals (KFUPM), Dhahran 31261, Saudi Arabia}
+\affil[12]{Interdisciplinary Research Center for Intelligent Manufacturing and Robotics, KFUPM, Dhahran 31261, Saudi Arabia}
+\affil[13]{Institute of Energy Materials and Devices, Forschungszentrum Juelich GmbH, 52425 Juelich, Germany}
 \begin{document}
@@
 We acknowledge financial support from the Helmholtz Foundation Model Initiative within Project "SOL-AI".
+Part of this work was funded under the France 2030 framework by Agence Nationale de la Recherche (project ANR-22-PEXD-0009 of PEPR DIADEM).
+Work at the Molecular Foundry was supported by the Office of Science, Office of Basic Energy Sciences, of the U.S.
+Department of Energy under Contract No.
+DE-AC02-05CH11231.
+Work at the Advanced Light Source (ALS) was done at beamline 12.3.2.
+The ALS is a DOE Office of Science User Facility under contract no.
+DE-AC02-05CH11231.
+The development of the online phase identification platform is supported by the Guangzhou-HKUST(GZ) Joint Funding Program (No.
+2023A03J0003).
+Work by the USC group was supported by the National Science Foundation (NSF) grant numbers DMR-2227178 and OISE-2106597.
+
+%TODO: Include acknowledgements to the following researchers from the Wolf group
+%The author list is fine, unfortunately I did not get real support from my team to provide a bigger consistent dataset.
+Please acknowledge the experimental contributions of Angelina Barthelmeß, Elisabeth Herzinger, Henning Hinrichs.
 

### 2024-11-21 chapters/04_our_dataset.tex
Changed wording on PSL university text

@@
 \footnote{Available online at \url{https://www.crystallography.net/cod/}.} The data in the COD is placed in the public domain and licensed under the CC0 License.
-Of the entire COD database, 5432 structures 1\% of the database) contained at least one tag from the {CIF\_POW} dictionary, i.e., relating to powder diffraction studies --- which is expected, since most crystal structures are resolved from single crystal diffraction.
+Of the entire COD database 5432 structures contained at least one tag from the {CIF\_POW} dictionary, i.e., a tag relating to powder diffraction studies.
+These 5432 structures only account for 1\% of the total COD database, but this is to be expected since most crystal structures are resolved from single-crystal diffraction.
 Of these 5432 files, most contained only metadata related to the powder diffraction experiment, but did not include the raw data of the pattern itself.

### 2024-11-21 chapters/00_abstract.tex
Proofreading of dataset paragraphs + Respond to comments in abstract

--- 2024-05-14 chapters/00_abstract.tex
@@
 Powder X-ray diffraction (pXRD) experiments are a cornerstone for materials structure characterization.
-Despite their widespread application, the analysis of pXRD diffractograms still presents a significant challenge to automation and thus a bottleneck in high-throughput experimentation automated materials discovery in self-driving labs.
+Despite their widespread application, the analysis of pXRD diffractograms still presents a significant challenge to automation and thus a bottleneck in high-throughput experimentation of automated materials discovery in self-driving labs.
 Machine learning has emerged as a promising research direction to resolve this bottleneck by enabling automated powder diffraction data analysis.
 A notable difficulty in applying machine learning to this domain is the lack of sufficiently sized experimental datasets, which has relegated machine learning researchers to train primarily on simulated data.
-Since simulations largely fail to accurately reflect the experiment, the performance of models trained on only simulated data often lacks transferability to experimental data and thus fails to provide value in practice.
-With the Open Experimental Powder X-Ray Diffraction Database (opXRD), we aim to remedy this by providing an openly available and easily accessible dataset of partially labeled experimental powder diffractograms, providing machine learning researchers with a large quantity of real experimental diffractograms collected from a broad range of samples.
-We provide almost barrier-free software tools that allow experimental researchers to find their data and make it accessible - in virtually any widely used format.
+Since simulations largely fail to accurately reflect the experiment, the performance of models trained on only simulated data often lacks transferability to experimental data and thus currently falls short of providing value in practice.
+With the Open Experimental Powder X-Ray Diffraction Database (opXRD), we aim to remedy this by providing an openly available and easily accessible dataset of both labeled and unlabeled experimental powder diffractograms, providing machine learning researchers with a large quantity of real experimental diffractograms collected from a broad range of samples.
+The labeled data can be used as a benchmark to evaluate the performance of models on real experimental data and unlabeled data can help improve the performance of models on experimental data through transfer learning methods.
+%We provide almost barrier-free software tools that allow experimental researchers to find their data and make it accessible - in virtually any widely used format.
 We collected XXX labeled and YYY unlabeled diffractograms from a wide spectrum of materials classes, which establishes the first version of the opXRD database.

### 2024-11-21 chapters/04_our_dataset.tex
Proofreading of dataset paragraphs + Respond to comments in abstract

@@
 
-In collaboration with several other institutions we have collected a dataset of diffractograms stemming from experiment, some of them labeled with corresponding structural information.
-
-\begin{itemize}
-    \item Institute of Nanotechnology at Karlsruhe Institute of Technology
-    \item University of Southern California
-    \item Lawrence Berkeley National Laboratory
-    \item Hong Kong University of Science and Technology (Guangzhou)
-    \item Swiss Federal Laboratories for Materials Science and Technology
-\end{itemize}
+In collaboration with several other institutions we have collected a dataset of diffractograms stemming from experiments, some of them labeled with corresponding structural information.
+The following texts detail the contribution of each research group.
+Most data was collected with copper radiation sources which exhibit a $K_{\alpha1}$ wavelength of $\lambda$~=~1.54056~{\AA} and a $K_{\alpha2}$ wavelength of $\lambda$~=~1.54439~{\AA}.
 
 %% Paragraph by FX Coudert and Arthur Hardiagon
-\subsection*{Chimie Paris Tech, PSL University}
+\subsubsection*{Chimie Paris Tech, PSL University}
 
@@
 Of these 5432 files, most contained only metadata related to the powder diffraction experiment, but did not include the raw data of the pattern itself.
-We could extract raw, experimental pXRD patterns from 1063 files in total, after curation of a small number of files with clearly invalid data.
+We could extract raw experimental pXRD patterns from 1063 files in total, after curation of a small number of files with clearly invalid data.
 \\
@@
 %% Paragraph by Ben Breitung
-\subsection*{Institute of Nanotechnology, KIT}
+\subsubsection*{Institute of Nanotechnology, KIT}
 
-The X-ray diffraction data were collected on a Bruker D8 Advance (Cu K$\alpha$1 radiation, $\lambda$ = 1.54056 Å).
-The samples were initially recorded for various research projects over the last ten years and were measured with different step sizes, times per step, and over different angle ranges, but all using Cu K$\alpha$1 radiation.
+A major part of the research focused on high-entropy materials, which involved incorporating many different elements into single-phase structures, leading to peak shifts or phase separations.
+Most of those multi-component complex materials appeared in various structures, including rock-salt, spinel, fluorite, perovskite, and delafossite.
+The samples were almost always prepared in powder form; therefore, powder XRD was performed on samples with adjusted height.
+The samples were prepared using various synthesis techniques, mostly solid-state or wet chemical syntheses, to obtain the desired structures.
+Consequently, particle size and crystallinity varied significantly.
+The sample set also includes samples that were not successfully measured or where phases could not be identified.
+\\
+
+The X-ray diffraction data were collected on a Bruker D8 Advance using a Cu radiation source  The samples were initially recorded for various research projects over the last ten years and were measured with different step sizes, times per step, and over different angle ranges, but all using Cu K$\alpha$1 radiation.
 The samples mostly contained transition metal oxides, sulfides, and fluorides.
-A major part of the research focused on high-entropy materials, which involved incorporating many different elements into single-phase structures, leading to peak shifts or phase separations.
-Most of those multi-component complex materials appeared in various structures, including rock-salt, spinel, fluorite, perovskite, and delafossite, to name a few.The samples were almost always prepared in powder form; therefore, powder XRD was performed on samples with adjusted height.
 To improve statistics, the samples were rotated during the entire measurement.
@@
 This dome led to increased background noise over the first 20° and slightly decreased pattern resolution.
-The samples were prepared using various synthesis techniques, mostly solid-state or wet chemical syntheses, to obtain the desired structures.
-Consequently, particle size and crystallinity varied significantly.
-The sample set also includes samples that were not successfully measured or where phases could not be identified.
+\\
+
+% Paragraph by Moritz Wolf and Tony-yi 
+\subsubsection*{Institute of Catalysis Research and Technology, KIT}
+
+A variety of samples were analyzed including commercial catalysts, bulk reference materials, porous metal oxide particles, and nanoparticles.
+The latter were synthesized via the surfactant-free benzyl alcohol route.
+\cite{Wolf2019, Wolf2018} The cobalt oxide (CoO or Co3O4) and cerium oxide (CeO2) nanoparticles were in the size range of 4-16 nm according to the Scherrer equation.
+A series of porous Al2O3 materials, which were prepared by calcination of boehmite (AlOOH) at various temperatures, represents crystalline samples with limited long-range structure and various contributions of Al2O3 polymorphs.\\
+
+X-ray diffraction (XRD) was conducted with an X’Pert Pro MPD (Panalytical) in Bragg-Brentano geometry using a Cu X-ray source.
+The patterns were acquired from 5-80° $2\theta$ with a step size of 0.016711° or 0.033420° and a total acquisition time of 40 to 120 min.
+\\
 
 % Paragraph by Tim Kodalle
-\subsection*{Lawrence Berkeley National Lab}
+\subsubsection*{Lawrence Berkeley National Lab}
 
-Experimental XRD data were collected at beamline 12.3.2 of the Advanced Light Source, the synchrotron at Lawrence Berkeley National Laboratory, using a photon energy of 10 keV (corresponding to a wavelength of $\lambda$ = 1.23984193 Å), selected using a Si(111) monochromator.
-Measurements were taken in grazing incidence geometry, i.e.
-using a beam incidence angle of 1$^{\circ}$.
-Two-dimensional diffraction images were recorded using a Dectris Pilatus 1M area detector at an angle between 34$^{\circ}$ and 36$^{\circ}$ with a sample-to-detector distance of roughly 190 mm.
-The two-dimensional data were calibrated using an Al$_{2}$O$_{3}$ calibration standard and integrated along the azimuthal angle.
 Data collection was performed in situ during thin-film deposition using a custom-made spin-coating and annealing stage.
-One dataset was collected spin-coating triple cation metal-halide perovskite precursor solutions with the composition Cs$_{0.05}$(MA$_{0.23}$FA$_{0.77}$)Pb$_{1.1}$(I$_{0.77}$Br$_{0.23}$)$_{3}$ (MA = Methylammonium, FA = Formamidinium) onto various substrates, including glass (amorphous), GaAs wafers (single crystalline) as well as stacks of glass/ITO, GaAs/Mo/Cu(In,Ga)Se$_{2}$/CdS/ZnO, and glass/Mo/Cu(In,Ga)Se$_{2}$/CdS/ZnO (glass/CIGS).
+One dataset was collected from spin-coating triple cation metal-halide perovskite precursor solutions with the composition Cs$_{0.05}$(MA$_{0.23}$FA$_{0.77}$)Pb$_{1.1}$(I$_{0.77}$Br$_{0.23}$)$_{3}$ (MA = Methylammonium, FA = Formamidinium) onto various substrates, including glass (amorphous), GaAs wafers (single crystalline) as well as stacks of glass/ITO, GaAs/Mo/Cu(In,Ga)Se$_{2}$/CdS/ZnO, and glass/Mo/Cu(In,Ga)Se$_{2}$/CdS/ZnO (glass/CIGS).
 Some of the substrates were additionally covered with a self-assembling monolayer of MeO-2PACz.
@@
 
+Experimental XRD data were collected at beamline 12.3.2 of the Advanced Light Source, the synchrotron at Lawrence Berkeley National Laboratory.
+The data were collected using a photon energy of 10 keV ($\lambda$ = 1.23984 \AA), selected using a Si(111) monochromator.
+Measurements were taken in grazing incidence geometry, i.e.
+using a beam incidence angle of 1$^{\circ}$.
+Two-dimensional diffraction images were recorded using a Dectris Pilatus 1M area detector at an angle between 34$^{\circ}$ and 36$^{\circ}$ with a sample-to-detector distance of roughly 190 mm.
+The two-dimensional data were calibrated using an Al$_{2}$O$_{3}$ calibration standard and integrated along the azimuthal angle.
+\\
+
 % Paragraph by Bin Cao
-\subsection*{Guangzhou Municipal Key Laboratory of Materials Informatics, HKUST}
+\subsubsection*{Guangzhou Municipal Key Laboratory of Materials Informatics, HKUST}
 
-In the past two years, we established a small-scale experimental PXRD database called the X-Ray Phase Identification Public Experimental Dataset (XRed,  \href{https://github.com/WPEM/XRED}{https://github.com/WPEM/XRED}).
-The data in XRed is generated in our lab using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers under copper targets.
+In the past two years, we established a small-scale experimental PXRD database called the X-Ray Phase Identification Public Experimental Dataset.
+The data in XRed is generated in our lab using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers under Cu targets.
 Hundreds of PXRD patterns have been refined, labeled with CIF files, and organized by elemental systems.
@@
 
-% Paragraph by Moritz Wolf and Tony-yi 
-\subsection*{Institute of Catalysis Research and Technology, KIT}
-
-X-ray diffraction (XRD) was conducted with an X’Pert Pro MPD (Panalytical) in Bragg-Brentano geometry using a copper X-ray source ($\lambda_{K_{\alpha}} = 1.5406 Å$).
-The patterns were acquired from 5-80° $2\theta$ with a step size of 0.016711° or 0.033420° and a total acquisition time of 40 to 120 min.
-A variety of samples were analyzed including commercial catalysts, bulk reference materials, porous metal oxide particles, and nanoparticles.
-The latter were synthesized via the surfactant-free benzyl alcohol route (10.1039/C9DT01634A, 10.1016/j.matchemphys.2018.04.021).
-The cobalt oxide (CoO or Co3O4) and cerium oxide (CeO2) nanoparticles were in the size range of 4-16 nm according to the Scherrer equation.
-A series of porous Al2O3 materials, which were prepared by calcination of boehmite (AlOOH) at various temperatures, represents crystalline samples with limited long-range structure and various contributions of Al2O3 polymorphs.
+In the past two years, we established a small-scale experimental pXRD database called the X-Ray Phase Identification Public Experimental Dataset (\href{https://github.com/WPEM/XRED}{XRed}).
+The data in XRed is generated in our lab using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers using copper X-ray sources.
+Hundreds of pXRD patterns have been refined, labeled with CIF files, and organized by elemental systems.
+The dataset includes original experimental files across single-phase to five-phase mixtures.
+This project is progressing toward integration with the comprehensive opXRD database to establish a large-scale, long-term experimental resource.
 
 % Paragraph by Alexander Wieczorek and Sebastiann Siol
-\subsection*{Laboratory for Surface Science and Coating Technologies, Empa}
+\subsubsection*{Laboratory for Surface Science and Coating Technologies, Empa}
 
-XRD data was measured using a Bruker D8 Discover equipped with a Cu~K$\alpha$ radiation source ($\lambda$~=~1.54056~{\AA} for Cu~K$\alpha$\textsubscript{1} and $\lambda$~=~1.54439~{\AA} for Cu~K$\alpha$\textsubscript{2}) in a Bragg-Brentano geometry.
+Each combinatorial Zn–V–N library was synthesized using radio-frequency co-sputtering of Zn and V in a mixed Ar and N\textsubscript{2} plasma.
+An orthogonal deposition temperature and composition gradient was created, resulting in a deposition temperature of 220~{\textdegree}C for samples 1~–~9 and 114~{\textdegree}C for samples 37~–~45.
+The composition for each sample was determined using X-ray fluorescence (XRF) spectroscopy which was further calibrated through Rutherford backscattering spectroscopy (RBS) based on selected samples.
+The newly identified and isolated semiconductor Zn\textsubscript{2}VN\textsubscript{3} was identified to exhibit a cation-disordered wurtzite structure as verified by additional GI-XRD and SAED measurements.
+\cite{Zhuk2021} \\
+Tin halide perovskites were deposited using single-step spin-coating as reported elsewhere.\cite{Wieczorek2023} Methylammonium lead iodide libraries with varying degrees of residual PbI\textsubscript{2} were deposited using a two-step procedure involving both thermal evaporation of PbI\textsubscript{2} and subsequent spin-coating of a methylammonium solution.
+The relative phase fractions were quantified using supplementary azimuthal angle scans coupled with structural factors and geometrical factors as reported elsewhere.\cite{Wieczorek2024} Fully inorganic lead perovskite libraries were prepared using thermal co-evaporation of lead and cesium halide salts.
+All metal halide perovskite libraries were measured within a custom-made X-Ray transparent inert-gas dome, resulting in the presence of minor additional features within the $\theta$~=~19~–~31{\textdegree} range.
+For all combinatorial libraries where any phases are specified, the complete set of phases is reported in the metadata.
+\\
+
+XRD data was measured using a Bruker D8 Discover equipped with a Cu radiation source in a Bragg-Brentano geometry.
 For the reported data sets the instrument was equipped with a Goebel mirror effectively removing the Cu~K$\beta$ radiation.
@@
 All data was collected from thin films deposited on borosilicate glass.
-The Zn–V–N films showed some preferential out of plane orientation, while for the perovskites the preferential orientation was minimal, resulting in the presence of all reflections.
-Each combinatorial Zn–V–N library was synthesized using radio-frequency co-sputtering of Zn and V in a mixed Ar and N\textsubscript{2} plasma.
-An orthogonal deposition temperature and composition gradient was created, resulting in a deposition temperature of 220~{\textdegree}C for samples 1~–~9 and 114~{\textdegree}C for samples 37~–~45.
-The composition for each sample was determined using X-Ray fluorescence (XRF) spectroscopy which was further calibrated through Rutherford backscattering spectroscopy (RBS) based on selected samples.
-The newly identified and isolated semiconductor Zn\textsubscript{2}VN\textsubscript{3} was identified to exhibit a cation-disordered wurtzite structure as verified by additional GI-XRD and SAED measurements.\cite{Zhuk2021}
-Tin halide perovskites were deposited using single-step spin-coating as reported elsewhere.\cite{Wieczorek2023} Methylammonium lead iodide libraries with varying degrees of residual PbI\textsubscript{2} were deposited using a two-step procedure involving both thermal evaporation of PbI\textsubscript{2} and subsequent spin-coating of a methylammonium solution.
-The relative phase fractions were quantified using supplementary azimuthal angle scans coupled with structural factors and geometrical factors as reported elsewhere.\cite{Wieczorek2024} Fully inorganic lead perovskite libraries were prepared using thermal co-evaporation of lead and cesium halide salts.
-All metal halide perovskite libraries were measured within a custom-made X-Ray transparent inert-gas dome, resulting in the presence of minor additional features within the $\theta$~=~19~–~31{\textdegree} range.
-For all combinatorial libraries where any phases are specified, the complete set of phases are reported in the meta data.
+The Zn–V–N films showed some preferential out-of-plane orientation, while for the perovskites the preferential orientation was minimal, resulting in the presence of all reflections.
+
+%---------------------------------
 

### 2024-11-22 chapters/00_abstract.tex
Added paragraph by Hodge group

@@
 A notable difficulty in applying machine learning to this domain is the lack of sufficiently sized experimental datasets, which has relegated machine learning researchers to train primarily on simulated data.
-Since simulations largely fail to accurately reflect the experiment, the performance of models trained on only simulated data often lacks transferability to experimental data and thus currently falls short of providing value in practice.
-With the Open Experimental Powder X-Ray Diffraction Database (opXRD), we aim to remedy this by providing an openly available and easily accessible dataset of both labeled and unlabeled experimental powder diffractograms, providing machine learning researchers with a large quantity of real experimental diffractograms collected from a broad range of samples.
+However, models trained on simulated pXRD patterns showed limited generalization ability when applied to experimental patterns, particularly with low-quality experimental pattern that has high noise levels and elevated background.
+%Therefore, developing an experimental database offers at least two key advantages: (1) providing a comprehensive benchmark dataset to clarify model performance in practical applications, and (2) advancing phase identification tasks by supplying target-domain data within a transfer learning framework.
+With the Open Experimental Powder X-Ray Diffraction Database (opXRD), we aim to provide an openly available and easily accessible dataset of both labeled and unlabeled experimental powder diffractograms, providing machine learning researchers with a large quantity of real experimental diffractograms collected from a broad range of samples.
 The labeled data can be used as a benchmark to evaluate the performance of models on real experimental data and unlabeled data can help improve the performance of models on experimental data through transfer learning methods.

### 2024-11-22 chapters/01_introduction.tex
Added paragraph by Hodge group

@@
-\begin{figure*}[!htb]
-    \centering
-    \includegraphics{figures/overview.pdf}
-    \caption{Overview of the data collection pipeline.
-Datasets are submitted using an online submission form or our submission helper software.
-After post-processing and data homogenization, we create a Zenodo entry for each user submission and subsequently include the submission in the opXRD database.}
-    \label{fig:overview}
-\end{figure*}
-
 % High throughput experimentation
@@
 After fabricating a new material, a number of analysis techniques can be used to characterize the sample.
-One method that can be used to determine the crystal structure of a new material is powder X-ray diffraction.
+One method that can be used for phase identification, phase quantification, grain size characterization, and partially to determine the crystal structure of a new material is powder X-ray diffraction.
 To determine crystal structure from pXRD data, an initial guess is optimized using Rietveld refinement.
@@
 Collecting a labeled dataset of raw experimental powder X-ray diffraction data of sufficient size to directly train a model that predicts crystal structure information is likely out of reach.
-The most advanced ML models currently are trained on approximately $10^7 - 10^8$ simulated diffractograms, which is - to the best of our knowledge - three orders of magnitude larger than the largest currently curated experimental dataset and even x orders of magnitude larger than the number of unlabeled diffractograms in the opXRD dataset we present here.
+The most advanced ML models currently are trained on approximately $10^7 - 10^8$ simulated diffractograms, which is - to the best of our knowledge - three orders of magnitude larger than the largest currently curated experimental dataset (PDF-5+ with approximately 20,000 experimental patterns) and even x orders of magnitude larger than the number of unlabeled diffractograms in the opXRD dataset we present here.
 \\
@@
 One can bring the simulated patterns closer to experiment by improving the simulation itself by properly modeling more of the observed physical effects such as preferred orientation, peak shapes resulting from different sources, etc.
-Additionally, some of the observed imperfections, such as background and noise, can be modeled using machine learning approaches.
+Additionally, some of the observed imperfections, such as background and noise, can be modeled using machine-learning approaches.
 This is similar to the concept of style transfer\cite{Gatys2016, Ganin2015} found in the domain of machine learning applied to images.
@@
 
-In this publication, we provide a new open powder X-ray diffraction (opXRD) dataset that collects a broad range of patterns from experiment, exceeding the currently available amount of openly accessible experimental patterns by XX.
+In this publication, we provide a new open powder X-ray diffraction (opXRD) dataset that collects a broad range of patterns from experiments, exceeding the currently available amount of openly accessible experimental patterns by XX.
 Our opXRD dataset has been curated by collecting the accumulated powder data from multiple large research groups and institutions with high-throughput XRD facilities.
@@
 \\
+
+\begin{figure*}[!htb]
+    \centering
+    \includegraphics{figures/overview.pdf}
+    \caption{Overview of the data collection pipeline.
+Datasets are submitted using an online submission form or our submission helper software.
+After post-processing and data homogenization, we create a Zenodo entry for each user submission and subsequently include the submission in the opXRD database.}
+    \label{fig:overview}
+\end{figure*}
 
@@
 We hope that this will yield more advanced automated ML approaches in the future, which can readily be applied in high-throughput experimentation pipelines.
+
+% TODOS: These papers talk about the importance of sharing raw diffraction data/ raw powder diffraction data them
+%  1 | \cite{Aranda2018}
+% 2 | Raw diffraction data and reproducibility

### 2024-11-22 chapters/02_existing_datasets.tex
Added paragraph by Hodge group

@@
-% Sources to mention
-% Zenodo -> A good amount of raw powder data there but spread into smaller datasets; Inhomogenous, not specially prepared to be worked with by others, therefore difficult to work with especially at scale
-
-% Crystal structures with corresponding simulated diffractograms;
-% ->  ICSD 
-% - 230k structures;
-% - Proprietary
-
-% -> PDF  -> 
-%- 1200k "material data sets" (structures?) 
-% + 19 k entries with "Experimental raw data digital patterns"; 
-%  Proprietary (% PDF pricing: https://www.icdd.com/assets/files/2023-Purchase-Options.pdf)
-%The PDF itself lists the following data sources:
-
-% ->  COD (https://www.crystallography.net/cod/)
-% -  513k structures,
-% - Fully open
-
-% Datset listing (Done on: 20.06.24)
-% I would list here *all* crystallography databases but evaluate/determine that they lack or don't provide convenient access to raw powder diffraction data
-
+% Datset listing (List written on 20.06.24)
 % Some other Lists of Crystallographic databases
-% - IUCR: "Crystallographyic databases and related resources" (https://www.iucr.org/resources/data/databases)
+% - IUCR: "Crystallographic databases and related resources" (https://www.iucr.org/resources/data/databases)
 % - Bruno et al. (2017): "Crystallography and Databases" (https://datascience.codata.org/articles/10.5334/dsj-2017-038)
@@
 
-%->1 PDF5+ (https://www.icdd.com/,https://www.icdd.com/pdf-5/)
-% - Data sources of the PDF
+% [X] 1 PDF5+ (https://www.icdd.com/,https://www.icdd.com/pdf-5/) -> 19000 structures with pXRD patterns | I checked with their support
     % (1.1): ICDD Powders
@@
     % (1.5): ICDD Single Crystal data
-%-> 2 COD (https://www.crystallography.net/cod/)
-%-> 3 RRUFF (American Mineralogist Crystal Structure Database)
-%-> 4 CCDC (https://www.ccdc.cam.ac.uk/)
-%-> 5 The Material Project (https://legacy.materialsproject.org/)
-%-> 6 Crystal Lattice Structures (https://www.atomic-scale-physics.de/lattice/)
-%-> 7 Crystallographic and Crystallochemical Database for Minerals and their Structural Analogues (https://database.iem.ac.ru/mincryst/)
-%-> 8 Mineralogy Database (https://webmineral.com/)
-%-> 9 IUCr Raw data letters (https://iucrdata.iucr.org/)
-%-> 10 Bilbao Crystallographic server  (https://www.cryst.ehu.es/)
-%-> 11 PowBase (http://www.cristal.org/powbase/index.html)
-%-> 12 Athena (https://athena.unige.ch/athena/mineral/mineral.html)
-%-> 13 Protein database (https://www.rcsb.org/)
-%-> 14 SDPD (Structure Determination from Powder Diffraction - Database)
-%-> 15 Crystal Met (https://cds.dl.ac.uk/cds/datasets/crys/mdf/llmdf.html)
-%-> 16 NAKB (https://www.nakb.org/)
-%-> 17 Pearson's Crystal data (https://www.crystalimpact.com/pcd/)
-%-> 18 Biological macromolecule Crystallization Database (http://bmcd.ibbr.umd.edu/)
-
-%TODO: Also include plans for NOMAD and highlight their efforts
+% [x]: 2 COD (https://www.crystallography.net/cod/) -> ~ 1000 structures with real pXRD patterns | These patterns were extracted by PSL university
+% [x]: 3 RRUFF (American Mineralogist Crystal Structure Database) -> ~ 1000 structures with real pXRD patterns | We downloaded the entire dataset and discarded single crystal data
+% [x]: 4 CCDC (https://www.ccdc.cam.ac.uk/) -> No exp pXRD | There is only mention of simulated powder diffraction data. Of 20 sampled cifs none contained powder diffraction data
+% [x]: 5 The Material Project (https://legacy.materialsproject.org/) -> No exp pXRD | The only mention of available X-ray diffraction data is of simulated data
+% [x]: 6 Crystal Lattice Structures (https://www.atomic-scale-physics.de/lattice/) -> No exp pXRD | Their entries provide no X-ray diffraction data, and also most links appear broken
+% [x]: 7 Crystallographic and Crystallochemical Database for Minerals and their Structural Analogues (https://database.iem.ac.ru/mincryst/) -> No exp pXRD | The given powder diffraction data is simulated (see: Powder X-ray Diffraction Standards (CPDS) @ https://database.iem.ac.ru/mincryst/descript.htm)
+% [x]: 8 Mineralogy Database (https://webmineral.com/) -> No exp pXRD | I checked with the author of the site
+% [x]: 9 IUCr Raw data letters (https://iucrdata.iucr.org/) -> No exp pXRD | There are only two raw data letters, both of them single crystal data
+% [x]: 10 Bilbao Crystallographic server  (https://www.cryst.ehu.eus/bincstrdb/search/) No exp pXRD | Of 20 sampled CIFs out of 256 entries none contained pXRD data
+% [x]: 11 PowBase (http://www.cristal.org/powbase/index.html) -> 169 real pXRD patterns some of which have at least partial labels; The described submission process and the lack of crystal structures indicate that the patterns are not simulated
+% [x]: 12 Athena (https://athena.unige.ch/athena/mineral/mineral.html) -> No exp pXRD | No mention of diffraction data anywhere
+% [x]: 13 Protein database (https://www.rcsb.org/) / NAKB (https://www.nakb.org/) / BMCD ((http://bmcd.ibbr.umd.edu/)  -> Little to no exp pXRD | Of the 230k structures only 21 showed up when searching experimental method = Powder diffraction
+% [x]: 14 Crystal Met (https://cds.dl.ac.uk/cds/datasets/crys/mdf/llmdf.html) -> No exp pXRD | The only mention of diffraction data is simulated data
+% [x]: 15 Pearson's Crystal data (https://www.crystalimpact.com/pcd/) -> 21700 experimental powder diffraction patterns (As per https://www.crystalimpact.com/pcd/) | The Pearson's crystal data evolved from the Linus Pauling File which is a data source of the PDF, so there may be significant overlap in these patterns
+% [x]: 16 Zenodo powder diffraction data (https://zenodo.org/search?q=powder%20diffraction&l=list&p=1&s=10&sort=bestmatch) -> Some pXRD data but hard to quantify, probably <= 4000 patterns most of which are unlabeled | 1600 datasets matching "powder diffraction".
+From probing among the first 100 matches I estimate only about 1 in 10 contain powder diffraction patterns.
+The probed datasets that did contain powder diffraction data only contained <= 30 patterns and only one of them came with corresponding crystal structures
 
@@
 
-\paragraph{Existing powder diffraction datasets} To contextualize opXRD within the current environment of experimental powder diffraction data, the following list provides an overview of the largest crystal structure databases that include powder diffraction data.
+\paragraph{Existing experimental powder diffraction datasets} To contextualize opXRD within the current environment of experimental powder diffraction data, the following list provides an overview of the largest crystal structure databases that include powder diffraction data.
 
@@
 \item Scattered experimental datasets on Zenodo (TODO: roughly estimate size)
+
 \item \textbf{RRUFF}: The RRUFF Mineral Database \cite{Armbruster2015} provides detailed information on minerals, including their chemical compositions, crystallography, and spectroscopic data.
@@
 It contains \num{908} distinct structures with experimental x-ray diffraction patterns.
-\item ...
+
 \end{itemize}

### 2024-11-22 chapters/03_SOTA_ML.tex
Added paragraph by Hodge group

@@
+%TODO:
+% - Papers to include: https://pubs.acs.org/doi/10.1021/jacs.4c10244
+% - Sagado et al.: Automated classification of big X-ray diffraction data using deep learning models
+
 \paragraph{ML for pXRD} Several approaches have applied machine learning methods to classification and regression tasks for powder diffractograms.
 
-Using simulated diffractograms based on structures from the ICSD, Lee et al.
-managed to train a deep convolutional neural network (CNN) that is able to classify occurring phases in diffractograms of a specific compound pool.
-\cite{Lee2020} Building upon their work they also developed models based on fully convolutional neural networks and transformer encoders, that predict the crystal system and the spacegroup and other structural properties, such as the band gap \cite{Lee2022}.
+Using simulated diffractograms based on structures from the ICSD, Lee {\it et al.} trained a deep convolutional neural network (CNN) that is able to classify occurring phases in diffractograms of a specific compound pool \cite{Lee2020}.
+They furthermore developed models based on fully convolutional neural networks and transformer encoders that predict the crystal system and the spacegroup and other structural properties, such as the band gap \cite{Lee2022}.
 With their best model for the crystal system prediction on ICSD structures, they achieved a test accuracy of \SI{92.2}{\percent}.
-Park et al. reached a test accuracy of roughly 81\% for a CNN, which classifies space groups of simulated single-phase diffractograms. \cite{Park2017}
+Park {\it et al.} reached a test accuracy of roughly 81\% for a CNN, which classifies space groups of simulated single-phase diffractograms \cite{Park2017}.
 A regression analysis on lattice parameters within a broader framework encompassing all material classes was conducted by Chitturi et al.\cite{Chitturi2021} They developed a distinct CNN for each crystal system, utilizing a merged dataset from both the ICSD and the Cambridge Structural Database, and managed to achieve a mean absolute percentage error of about \SI{10}{\percent} for the lattice lengths, although they encountered difficulties in accurately predicting angles.
-Zhang and Cao et al.\cite{zhang2024crystallographic} introduce a convolutional self-attention neural network trained on simulated patterns to classify crystal types.
-Their model was tested on 23,073 unary, binary, and ternary inorganic crystal structures sourced from the Crystallography Open Database.
-The study observed a noticeable performance drop when the pretrained model was applied to real experimental patterns, as opposed to simulated data.
+Zhang {\it et al.} introduce a convolutional self-attention neural network trained on simulated patterns to classify crystal types \cite{zhang2024crystallographic}.
+Their model was tested on 23,073 unary, binary, and ternary inorganic crystal structures sourced from the COD.
+The study observed a noticeable performance drop when the pre-trained model was applied to real experimental patterns, as opposed to simulated data.
 However, their recent work \cite{cao2024simxrd} proposes using convolutional peak descriptors that consider the detector's geometry, which significantly reduces the performance gap in their benchmark tests.
-The option of training purely on experimental diffractograms unfortunately is unfeasible because of the limited availability of labeled experimental diffractograms, but Salgado et al.
-\cite{Salgado2023} showed, that adding a fraction of experimental patterns to a simulated training dataset improves the performance on unseen experimental patterns.
+The option of training purely on experimental diffractograms unfortunately is unfeasible because of the limited availability of labeled experimental diffractograms, but Salgado {\it et al.} \cite{Salgado2023} showed, that adding a fraction of experimental patterns to a simulated training dataset improves the performance on unseen experimental patterns.
 They used \SI{50}{\percent} of the experimental patterns contained in the RRUFF database and added those to their large simulated training set.
@@
 
-The ultimate goal of this ML technology is to make PXRD data identification practical for experimental use, whether for identifying space group symmetry or phase composition.
+The ultimate goal of ML efforts is to make PXRD data identification practical for experimental use, whether for identifying space group symmetry or phase composition.
 However, the current volume of experimental PXRD patterns is insufficient to effectively train ML models, highlighting an urgent need for a comprehensive experimental PXRD database.

### 2024-11-22 chapters/04_our_dataset.tex
Added paragraph by Hodge group

@@
-% TODOS:
-% - This paper talks about the importance of sharing raw powder X-ray diffraction data. I'd definitely cite it ->  \cite{Aranda2018}
-
-%- The subsections of this section should be the 
-% individual contributions that we get from labs
-
-%------------------------------------------$
-
 \begin{figure*}[!ht]
@@
 
-% Paragraph by Moritz Wolf and Tony-yi 
+% Paragraph by Andrea Hodge
+\subsubsection*{Department of Chemical Engineering and Materials Science, USC}
+
+Combinatorial CuNi and CuAl samples were synthesized at the University of Southern California (USC) via magnetron co-sputtering from Cu (99.999\%) and Ni (99.995\%) or Al (99.999\%) targets (Plasmaterials) onto stationary 10 cm Si (100) substrates yielding 169 unique 5 x 5mm squares per wafer.
+These alloys were selected to study compositional effects on stacking fault and growth twin boundary formation in sputtered materials \cite{2024AcMat.27019839A,alwen2024combinatorial}.
+XRD spectra were collected using a Malvern Panalytical Empyrean X-ray diffractometer at the Forschungszentrum Julich (FZJ) Institute for Energy and Climate Research, Structure and Function of Materials (IEK-2), which enabled automated XRD characterization due to its programmable XY stage and laser sensor for Z height correction.
+XRD analysis was used to link changes in defect formation in the sputtered films with varied phase and crystal structures.
+Each composition was characterized using CuK X-rays (1.5406Å wavelength) that were collimated to probe an area of ~4 x 4 mm using a step size of 0.026° and scan duration of 0.3 seconds per step over a 2 range from 30° - 110°.
+Measurement parameters were chosen to resolve at least six or more points above the full-width half maximum of each XRD peak.
+
+% Paragraph by Moritz Wolf 
 \subsubsection*{Institute of Catalysis Research and Technology, KIT}
@@
 A variety of samples were analyzed including commercial catalysts, bulk reference materials, porous metal oxide particles, and nanoparticles.
-The latter were synthesized via the surfactant-free benzyl alcohol route.
-\cite{Wolf2019, Wolf2018} The cobalt oxide (CoO or Co3O4) and cerium oxide (CeO2) nanoparticles were in the size range of 4-16 nm according to the Scherrer equation.
+The latter were synthesized via the surfactant-free benzyl alcohol route \cite{Wolf2019, Wolf2018}.
+The cobalt oxide (CoO or Co3O4) and cerium oxide (CeO2) nanoparticles were in the size range of 4-16 nm according to the Scherrer equation.
 A series of porous Al2O3 materials, which were prepared by calcination of boehmite (AlOOH) at various temperatures, represents crystalline samples with limited long-range structure and various contributions of Al2O3 polymorphs.\\
@@
 
-% Paragraph by Bin Cao
+% Paragraph by Bin Cao and Tong-yi Zhang
 \subsubsection*{Guangzhou Municipal Key Laboratory of Materials Informatics, HKUST}
@@
 
-In the past two years, we established a small-scale experimental pXRD database called the X-Ray Phase Identification Public Experimental Dataset (\href{https://github.com/WPEM/XRED}{XRed}).
-The data in XRed is generated in our lab using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers using copper X-ray sources.
+In the past two years, we established a small-scale experimental pXRD database called the X-Ray Phase Identification Public Experimental Dataset (XRed).
+\footnote{XRed can be found on GitHub at \url{https://github.com/WPEM/XRED}.} The data in XRed is generated in our lab using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers using copper X-ray sources.
 Hundreds of pXRD patterns have been refined, labeled with CIF files, and organized by elemental systems.
@@
 The Zn–V–N films showed some preferential out-of-plane orientation, while for the perovskites the preferential orientation was minimal, resulting in the presence of all reflections.
+\\
 
@@
 
-We are still in the process of growing the dataset and additional contributions are still very welcome.
-To find out more about how to contribute to this dataset, visit our website specially designed to collect this dataset, \url{https://xrd.aimat.science} .
+We will continue growing the opXRD dataset, thus additional contributors are very welcome to contact us and submit data through the website.
+To find out more about how to contribute to this dataset, visit our website \footnote{Datasets can be contributed to opXRD under \url{https://xrd.aimat.science}.} specially designed to collect this dataset.

### 2024-11-22 main.tex
Added paragraph by Hodge group

@@
 \author[4,5]{Tim Kodalle}
+\author[6]{Hanlin Hu}
+\author[7]{Gregoire Heymans}
+\author[8]{Alexander Wieczorek}
+\author[10]{Arthur Hardiagon}
+\author[11]{Bin Cao}
+\author[12, 13]{Maged Abdelsamie}
+\author[8]{Siarhei Zhuk}
 \author[2]{Ben Breitung}
-\author[11, 12]{Maged Abdelsamie}
-\author[6]{Gregoire Heymans}
-\author[7]{Alexander Wieczorek}
-\author[7]{Siarhei Zhuk}
-\author[7]{Sebastian Siol}
-\author[8]{Moritz Wolf}
-\author[9]{François-Xavier Coudert}
-\author[9]{Arthur Hardiagon}
-\author[10]{Bin Cao}
-\author[13]{Ruth Schwaiger}
-\author[10]{Tong-yi Zhang}
+\author[8]{Sebastian Siol}
+\author[9]{Moritz Wolf}
+\author[10]{François-Xavier Coudert}
+\author[14]{Ruth Schwaiger}
+\author[11]{Tong-yi Zhang}
 \author[4]{Carolin M. Sutter-Fella}
@@
 \affil[5]{Lawrence Berkeley National Laboratory, Advanced Light Source, Berkeley 94720 CA, USA}
-\affil[6]{Lawrence Berkeley National Laboratory, Chemical Sciences Division, Berkeley 94720 CA, USA}
-\affil[7]{Empa~–~Swiss Federal Laboratories
+\affil[6]{Hoffmann Institute of Advanced Materials, Shenzhen Polytechnic, Shenzhen 518055, China}
+\affil[7]{Lawrence Berkeley National Laboratory, Chemical Sciences Division, Berkeley 94720 CA, USA}
+\affil[8]{Empa~–~Swiss Federal Laboratories
 for Materials Science and Technology, 8600 Dübendorf,
 Switzerland}
-\affil[8]{Engler-Bunte-Institut \& Institute of Catalysis
+\affil[9]{Engler-Bunte-Institut \& Institute of Catalysis
 Research and Technology, Karlsruhe Institute of Technology, Karlsruhe, Germany}
-\affil[9]{Chimie ParisTech, PSL University, CNRS, Institut de Recherche de Chimie Paris, 75005 Paris, France}
-\affil[10]{Guangzhou Municipal Key Laboratory of Materials Informatics,  Hong Kong University of
+\affil[10]{Chimie ParisTech, PSL University, CNRS, Institut de Recherche de Chimie Paris, 75005 Paris, France}
+\affil[11]{Guangzhou Municipal Key Laboratory of Materials Informatics,  Hong Kong University of
 Science and Technology (Guangzhou), Guangzhou 511400, China}
-\affil[11]{Material Science and Engineering Department, King Fahd University of Petroleum and Minerals (KFUPM), Dhahran 31261, Saudi Arabia}
-\affil[12]{Interdisciplinary Research Center for Intelligent Manufacturing and Robotics, KFUPM, Dhahran 31261, Saudi Arabia}
-\affil[13]{Institute of Energy Materials and Devices, Forschungszentrum Juelich GmbH, 52425 Juelich, Germany}
+\affil[12]{Material Science and Engineering Department, King Fahd University of Petroleum and Minerals (KFUPM), Dhahran 31261, Saudi Arabia}
+\affil[13]{Interdisciplinary Research Center for Intelligent Manufacturing and Robotics, KFUPM, Dhahran 31261, Saudi Arabia}
+\affil[14]{Institute of Energy Materials and Devices, Forschungszentrum Juelich GmbH, 52425 Juelich, Germany}
 \begin{document}
@@
 
-\section{How to contribute and how to use}\label{sec:how_to_use}
+\section{Summary and Outlook}\label{sec:how_to_use}
 \input{chapters/05_how_to_use}

### 2024-11-22 chapters/02_existing_datasets.tex
Overhauled PDF entry in existing datasets

@@
 % Datset listing (List written on 20.06.24)
-% Some other Lists of Crystallographic databases
+% Lists of Crystallographic databases; Our list contains all of the databases given in the lists linked below
 % - IUCR: "Crystallographic databases and related resources" (https://www.iucr.org/resources/data/databases)
@@
 \begin{itemize}
-\item \textbf{Powder Diffraction File:} \footnote{More information on the Powder Diffraction File can be found under \url{https://www.icdd.com/pdf-5/}.} The Powder Diffraction File (PDF), published and maintained by the International Center for diffraction data, is currently the largest collection of powder diffraction data available to researchers.
-Though its latest release contains over a million entries, almost all are simulated patterns.
-Only 19000 entries represent powder diffraction data originating from experiments, and of those, only about 11000 come with accompanying atomic coordinates.
-It is widely used as a reference database in Rietveld refinement to identify crystal structures from which the refinement process can begin based on their powder diffraction patterns.
+\item \textbf{Powder Diffraction File:} \footnote{More information on the Powder Diffraction File can be found under \url{https://www.icdd.com/pdf-5/}.} The Powder Diffraction File (PDF), published and maintained by the International Center for Diffraction Data (ICDD), is a large collection of materials with accompanying powder diffraction data.
+Its latest release as of November 2024, the PDF5+, contains over a million materials with accompanying powder diffraction data.
+However, most of these powder diffraction patterns are simulated.
+After inquiring with the ICDD in April 2024 we were told that only 19000 of the powder diffraction patterns in the PDF5+ stemmed from experiments.
+\footnote{We reached out to the contact address given on their website, info@icdd.com.} Still, this makes the PDF5+, to the best of our knowledge, the largest collection of experimental powder diffraction data available to researchers.
 As of November 2024, the latest release of the PDF, the PDF5+, is available to researchers through a purchase starting at a price point of \$6,265.00.
@@
 Of these files, 1063 contained the experimental powder diffraction data that was used to determine the underlying crystal structures of the investigated samples.
-
-\item Scattered experimental datasets on Zenodo (TODO: roughly estimate size)
 

### 2024-11-22 chapters/02_existing_datasets.tex
Added square to citations to differentiate from footnotes

@@
 
-\paragraph{Existing experimental powder diffraction datasets} To contextualize opXRD within the current environment of experimental powder diffraction data, the following list provides an overview of the largest crystal structure databases that include powder diffraction data.
+\pagebreak
 
-\begin{itemize}
-\item \textbf{Powder Diffraction File:} \footnote{More information on the Powder Diffraction File can be found under \url{https://www.icdd.com/pdf-5/}.} The Powder Diffraction File (PDF), published and maintained by the International Center for Diffraction Data (ICDD), is a large collection of materials with accompanying powder diffraction data.
+\subsubsection*{Existing experimental powder diffraction datasets}
+To contextualize opXRD within the current environment of experimental powder diffraction data, the following list provides an overview of the largest crystal structure databases that include powder diffraction data.
+\\
+
+\textbf{Powder Diffraction File:}\footnote{More information on the Powder Diffraction File can be found under \url{https://www.icdd.com/pdf-5/}.} The Powder Diffraction File (PDF), published and maintained by the International Center for Diffraction Data (ICDD), is a large collection of materials with accompanying powder diffraction data.
 Its latest release as of November 2024, the PDF5+, contains over a million materials with accompanying powder diffraction data.
@@
 \footnote{We reached out to the contact address given on their website, info@icdd.com.} Still, this makes the PDF5+, to the best of our knowledge, the largest collection of experimental powder diffraction data available to researchers.
-As of November 2024, the latest release of the PDF, the PDF5+, is available to researchers through a purchase starting at a price point of \$6,265.00.
+As of November 2024, the PDF5+ is available to researchers through a purchase starting at a price point of \$6,265.00.\\
 
-\item \textbf{Crystallography Open Database:}\footnote{More information on the crystallography open database can be found under \url{https://www.crystallography.net/cod/}.} The Crystallography Open Database (COD)\cite{Graulis2009cod} is an open-access collection of crystal structures.
+\textbf{Crystallography Open Database:}\footnote{More information on the crystallography open database can be found under \url{https://www.crystallography.net/cod/}.} The Crystallography Open Database (COD)\cite{Graulis2009cod} is an open-access collection of crystal structures.
 It provides over 500.000 crystal structures in the form of .cif files.
 Of these files, 1063 contained the experimental powder diffraction data that was used to determine the underlying crystal structures of the investigated samples.
+\\
 
-\item \textbf{RRUFF}: The RRUFF Mineral Database \cite{Armbruster2015} provides detailed information on minerals, including their chemical compositions, crystallography, and spectroscopic data.
+\textbf{RRUFF}: The RRUFF Mineral Database \cite{Armbruster2015} provides detailed information on minerals, including their chemical compositions, crystallography, and spectroscopic data.
 Managed by the University of Arizona, it was created to serve as a public repository for mineral identification and research.
 It contains \num{908} distinct structures with experimental x-ray diffraction patterns.
+\\
 
-\end{itemize}
+\textbf{PowBase}:
+
+\textbf{Pearson's Crystal data}:
 

### 2024-11-22 chapters/02_existing_datasets.tex
Added RRUFF paragraph

@@
 After inquiring with the ICDD in April 2024 we were told that only 19000 of the powder diffraction patterns in the PDF5+ stemmed from experiments.
+Of these 19000 entries, 10954 contain information about the atomic coordinates of the underlying structures.
 \footnote{We reached out to the contact address given on their website, info@icdd.com.} Still, this makes the PDF5+, to the best of our knowledge, the largest collection of experimental powder diffraction data available to researchers.
@@
 Of these files, 1063 contained the experimental powder diffraction data that was used to determine the underlying crystal structures of the investigated samples.
+Hence the experimental powder diffraction data contained in the COD is fully labeled.
+The .cif files include both lattice parameters and the contents of the unit cell of the underlying structure.
 \\
@@
 Managed by the University of Arizona, it was created to serve as a public repository for mineral identification and research.
-It contains \num{908} distinct structures with experimental x-ray diffraction patterns.
-\\
+It contains \num{1290} powder diffraction patterns stemming from experiments each labeled with the lattice parameters and composition of the underlying structures.
 

### 2024-11-22 chapters/02_existing_datasets.tex
Added powbase paragraph

@@
 
-\subsubsection*{Existing experimental powder diffraction datasets}
+\subsubsection*{Existing experimental powder diffraction databases}
 To contextualize opXRD within the current environment of experimental powder diffraction data, the following list provides an overview of the largest crystal structure databases that include powder diffraction data.
@@
 
-\textbf{RRUFF}: The RRUFF Mineral Database \cite{Armbruster2015} provides detailed information on minerals, including their chemical compositions, crystallography, and spectroscopic data.
+\textbf{RRUFF}:\footnote{More information on RRUFF can be found under \url{https://rruff.info/about/about_general.php}.} The RRUFF Mineral Database \cite{Armbruster2015} provides detailed information on minerals, including their chemical compositions, crystallography, and spectroscopic data.
 Managed by the University of Arizona, it was created to serve as a public repository for mineral identification and research.
 It contains \num{1290} powder diffraction patterns stemming from experiments each labeled with the lattice parameters and composition of the underlying structures.
+\\
 
-\textbf{PowBase}:
+\textbf{PowBase}:\footnote{More information on PowBase can be found under \url{http://www.cristal.org/powbase/index.html}.} PowBase is a database of 169 mostly unlabeled experimental powder diffraction patterns collected and maintained by crystallography researcher Armel Le Bail starting in 1999.
+As of November 2024, all 169 patterns are still freely available for download.
+\\
 

### 2024-11-22 chapters/02_existing_datasets.tex
Added year of first publication for databases

@@
 
-\textbf{Powder Diffraction File:}\footnote{More information on the Powder Diffraction File can be found under \url{https://www.icdd.com/pdf-5/}.} The Powder Diffraction File (PDF), published and maintained by the International Center for Diffraction Data (ICDD), is a large collection of materials with accompanying powder diffraction data.
-Its latest release as of November 2024, the PDF5+, contains over a million materials with accompanying powder diffraction data.
+\textbf{Powder Diffraction File:}\footnote{More information on the Powder Diffraction File can be found under \url{https://www.icdd.com/pdf-5/}.} The Powder Diffraction File (PDF), published and maintained by the International Center for Diffraction Data (ICDD), is a large collection of materials with accompanying powder diffraction data first published in 1941.
+\cite{GatesRector2019} Its latest release as of November 2024, the PDF5+, contains over a million materials with accompanying powder diffraction data.
 However, most of these powder diffraction patterns are simulated.
@@
 
-\textbf{Crystallography Open Database:}\footnote{More information on the crystallography open database can be found under \url{https://www.crystallography.net/cod/}.} The Crystallography Open Database (COD)\cite{Graulis2009cod} is an open-access collection of crystal structures.
-It provides over 500.000 crystal structures in the form of .cif files.
-Of these files, 1063 contained the experimental powder diffraction data that was used to determine the underlying crystal structures of the investigated samples.
+\textbf{Crystallography Open Database:}\footnote{More information on the crystallography open database can be found under \url{https://www.crystallography.net/cod/}.} The Crystallography Open Database (COD) is an open-access collection of crystal structures founded in 2003.
+\cite{Graulis2009cod} It currently provides over 500.000 crystal structures in the form of .cif files.
+Of these files, 1063 contain the experimental powder diffraction data that was used to determine the underlying crystal structures of the investigated samples.
 Hence the experimental powder diffraction data contained in the COD is fully labeled.
@@
 
-\textbf{RRUFF}:\footnote{More information on RRUFF can be found under \url{https://rruff.info/about/about_general.php}.} The RRUFF Mineral Database \cite{Armbruster2015} provides detailed information on minerals, including their chemical compositions, crystallography, and spectroscopic data.
+\textbf{RRUFF}:\footnote{More information on RRUFF can be found under \url{https://rruff.info/about/about_general.php}.} The RRUFF Mineral Database, first published in 2006, provides detailed information on minerals, including their chemical compositions, crystallography, and spectroscopic data.
 Managed by the University of Arizona, it was created to serve as a public repository for mineral identification and research.
@@
 
-\textbf{PowBase}:\footnote{More information on PowBase can be found under \url{http://www.cristal.org/powbase/index.html}.} PowBase is a database of 169 mostly unlabeled experimental powder diffraction patterns collected and maintained by crystallography researcher Armel Le Bail starting in 1999.
+\textbf{PowBase}:\footnote{More information on PowBase can be found under \url{http://www.cristal.org/powbase/index.html}. PowBase was an initiative suggested in the Structure Determination by
+Powder Diffractometry (SDPD) mailing list which was hosted on the same site.
+The COD was another community initiative that grew out of this mailing list.} PowBase is a database of 169 mostly unlabeled experimental powder diffraction patterns collected and maintained by crystallography researcher Armel Le Bail starting in 1999.
 As of November 2024, all 169 patterns are still freely available for download.

### 2024-11-22 submission_helper/app.py
Added showing DOI as clickable links in bibliography when available

@@ -1,123 +1,0 @@
-import os
-import threading
-from datetime import datetime
-from typing import Optional
-from kivy.uix.boxlayout import BoxLayout
-from kivy.app import App
-from typing import List
-
-from kivy.clock import Clock
-
-from submission_helper.filesystem import zip_file_list, produce_csv_file
-from submission_helper.layouts import SelectionLayout, FinishLayout, InputDialog
-
-# -------------------------------------------
-
-class DataCollectApp(App):
-    def __init__(self, override_input_folder : Optional[str]  = None,
-                       override_target_folder : Optional[str] = None):
-        self.icon = None    
-        super(DataCollectApp, self).__init__()
-
-        # Properties
-        self.title : str = 'AiMAT XRD data collect'
-        self.input_folder_override : Optional[str]  = override_input_folder
-        self.targt_folder_override : Optional[str] = override_target_folder
-
-        # PathCheckbox
-        self.selection_layout : Optional[SelectionLayout] = None
-        self.finish_layout : Optional[FinishLayout] = None
-
-    def on_start(self):
-        input_dialog = InputDialog(callback=self.selection_layout.calculate_content)
-        self.selection_layout.register_content_done_callback(input_dialog.dismiss)
-
-        if self.input_folder_override is None:
-            input_dialog.open()
-        else:
-            self.selection_layout.calculate_content(path=self.input_folder_override)
-
-
-    def build(self):
-        self.selection_layout = SelectionLayout()
-        self.finish_layout = FinishLayout(callback=self.produce_dataset_files, exit_funct=self.stop)
-
-        root_layout = BoxLayout(orientation='vertical')
-        root_layout.add_widget(self.selection_layout)
-        root_layout.add_widget(self.finish_layout)
-
-        return root_layout
-
-    def _stop(self, *largs):
-        self.root_window.close()
-        super()._stop()
-
-    # -------------------------------------------
-    # callbacks
-
-    def produce_dataset_files(self, *args):
-        _ = args
-        print('Triggered callback of ok button')
-        checked_paths = self.get_checked_filepaths()
-        target_folder_path = self.get_targetfolder_path()
-
-        if target_folder_path is None:
-            print(f'No target folder provided. Check setup')
-            return
-
-        print("Checked Paths:", checked_paths)
-        produce_files_thread = threading.Thread(target=self.produce_files, args=(target_folder_path,checked_paths,))
-        produce_files_thread.start()
-
-
-
[…]

### 2024-11-22 submission_helper/configs/__init__.py

@@ -1,3 +1,0 @@
-from .conf import set_configs, get_line_height
-from .conf import get_scaled_down_app_height, get_scaled_down_app_width
-from .conf import get_true_height, get_true_width

### 2024-11-22 submission_helper/configs/conf.py

@@ -1,81 +1,0 @@
-import sys
-import ctypes
-from kivy.config import Config
-from screeninfo import get_monitors
-import logging
-
-logging.basicConfig(level=logging.INFO)
-# -------------------------------------------
-
-
-def get_scaling_factor():
-    ui_scale = 1
-    if sys.platform == 'win32':
-        try:
-            ui_scale = ctypes.windll.shcore.GetScaleFactorForDevice(0) / 100
-        except:
-            pass
-
-
-    return ui_scale
-
-def set_configs():
-    Config.set('graphics', 'fullscreen', 0)
-    # Config.set('graphics', 'resizable', '1')
-
-    app_width, app_height = get_scaled_down_app_width(), get_scaled_down_app_height()
-
-    print(f'[INFO]: Unscaled width, height: {app_width},{app_height}')
-
-    Config.set('graphics', 'width', str(app_width))
-    Config.set('graphics', 'height', str(app_height))
-
-
-def get_scaled_down_app_width() -> int:
-    width = int(get_true_width() / get_scaling_factor())
-    return width
-
-
-def get_scaled_down_app_height() -> int:
-    height = int(get_true_height() / get_scaling_factor())
-    return height
-
-
-def get_true_width() -> int:
-    primary_monitor = get_primary_monitor()
-    relative_width = 0.75
-
-    app_width = int(primary_monitor.width * relative_width)
-    return app_width
-
-
-def get_true_height() -> int:
-    primary_monitor = get_primary_monitor()
-    relative_height = 0.8
-
-    app_height = int(primary_monitor.height * relative_height)
-    return app_height
-
-
-def get_line_height():
-    std_width = 1080
-    std_line_height = 40
-    line_height =int(std_line_height * get_primary_monitor_height()/std_width)
-    print(line_height)
-
-    return line_height
-
-
-def get_primary_monitor():
-    monitors = get_monitors()
-    for monitor in monitors:
-        if monitor.is_primary:
-            return monitor
-    return None
-
-
-def get_primary_monitor_width():
-    return get_primary_monitor().width
-
-def get_primary_monitor_height():
-    return get_primary_monitor().height

### 2024-11-22 submission_helper/elements/__init__.py

@@ -1,2 +1,0 @@
-from .node_element import NodeElement
-from .types import BlackLabel, ThickVerticalSlider

### 2024-11-22 submission_helper/elements/loadable_element.py

@@ -1,73 +1,0 @@
-from __future__ import annotations
-
-from submission_helper.elements.node_widget import LabeledCheckBox
-from abc import abstractmethod
-from intervals import Interval
-
-from typing import Optional
-from typing import List
-
-# -------------------------------------------
-
-class LoadableElem:
-
-    def __init__(self, height : int):
-        super().__init__()
-
-        self.labeled_checkbox : Optional[LabeledCheckBox] = None
-
-        self.child_container = None
-        self.root_container = None
-        self.height = height
-        self.child_nodes : List[LoadableElem] = []
-        self.parent : Optional[LoadableElem] = None
-
-    # -------------------------------------------
-    # callbacks
-
-    @abstractmethod
-    def unload(self):
-        pass
-
-
-    @abstractmethod
-    def load(self):
-        pass
-
-    # -------------------------------------------
-    # get
-
-    @staticmethod
-    def get_nodes_in_interval(node : LoadableElem, interval : Interval) -> list:
-        start_y = interval.lower
-        end_y = interval.upper
-
-        selected_nodes = []
-
-        def dfs(current_node):
-            ypos = current_node.get_ypos()
-            if start_y <= ypos <= end_y:
-                selected_nodes.append(current_node)
-
-            if ypos <= end_y:
-                for child in current_node.get_gui_child_nodes():
-                    dfs(child)
-
-        dfs(node)
-        return selected_nodes
-
-
-    def get_ypos(self) -> int:
-        if self.parent is None:
-            return 0
-
-        parent_gui_children = self.parent.get_gui_child_nodes()
-        child_index = parent_gui_children.index(self)
-        previous_children = parent_gui_children[:child_index]
-
-        ypos = self.parent.get_ypos() + sum([child.root_container.height for child in previous_children]) + self.parent.height
-        return ypos
-
-    def get_gui_child_nodes(self):
-        return [child for child in self.child_nodes if child.root_container]
-

### 2024-11-22 submission_helper/elements/node_element.py

@@ -1,163 +1,0 @@
-from __future__ import annotations
-import copy
-import os.path
-from PIL import Image
-from typing import Optional, List
-from kivy.uix.boxlayout import  BoxLayout
-
-from submission_helper.elements.types import Placeholder
-from submission_helper.elements.node_widget import LabeledCheckBox, NodeWidget
-from submission_helper.elements.loadable_element import LoadableElem
-from submission_helper.resources import get_foldericon_path, get_fileicon_path
-from submission_helper.filesystem import FsNode
-from submission_helper.configs import get_line_height
-# -------------------------------------------
-
-class NodeElement(FsNode, LoadableElem):
-    dim = get_line_height()
-    file_img = Image.open(fp=get_fileicon_path()).resize((dim, dim))
-    folder_img = Image.open(fp=get_foldericon_path()).resize((dim, dim))
-
-    @classmethod
-    def make_child(cls, name : str, parent : NodeElement) -> NodeElement:
-        path = os.path.join(parent.path, name)
-        return cls(path=path,height=parent.height,scroll_view=parent.scroll_view)
-
-    def __init__(self,path : str, height : int, scroll_view):
-        FsNode.__init__(self,path=path)
-        LoadableElem.__init__(self, height=height)
-
-        self.height : int = height
-        self.labeled_checkbox : Optional[LabeledCheckBox] = None
-        self.scroll_view = scroll_view
-        self.is_loaded = False
-        self.is_initialized = False
-
-        self.node_widget = None
-        self.child_nodes : List[NodeElement] = []
-        self.parent : Optional[NodeElement] = None
-        self.place_holder = None
-
-    def initialize_gui(self, gui_parent, level : int):
-        self.root_container = BoxLayout(orientation='vertical', size_hint=(1, None))
-        self.root_container.bind(minimum_height=self.root_container.setter('height'))
-
-        self.labeled_checkbox = self._make_label_checkbox(level=level)
-        self.place_holder = self.get_placeholder()
-        self.root_container.add_widget(self.place_holder)
-
-        if not self.get_is_file():
-            self.child_container = self._get_child_container()
-            self.root_container.add_widget(self.child_container)
-
-        gui_parent.add_widget(self.root_container)
-        self.is_initialized = True
-
-
-    def add_child(self, name : str):
-        child = self.make_child(name=name, parent=self)
-        child.parent = self
-        self.child_nodes += [child]
-        return child
-
[…]

### 2024-11-22 submission_helper/elements/node_widget.py

@@ -1,129 +1,0 @@
-from __future__ import annotations
-
-from kivy.core.window import Window
-from kivy.uix.boxlayout import BoxLayout
-from kivy.uix.checkbox import CheckBox
-from kivy.uix.image import Image
-from kivy.uix.relativelayout import RelativeLayout
-from kivy.uix.togglebutton import ToggleButton
-from kivy.uix.widget import Widget
-from kivy.graphics import Rectangle
-from kivy.core.image import Image as CoreImage
-from kivy.properties import BooleanProperty, ObjectProperty
-
-from submission_helper.configs import get_line_height
-from submission_helper.elements.types import BlackLabel
-
-from submission_helper.resources import get_unchecked_box_path, get_checked_box_path, get_fileicon_path, \
-    get_foldericon_path, get_collapsed_icon_path, get_expanded_icon_path, get_kivy_image, get_empty_path
-
-# -------------------------------------------
-
-class NodeWidget(BoxLayout):
-    def __init__(self, callback : callable, height : int, labeled_checkbox : LabeledCheckBox, is_file : bool):
-        super().__init__(orientation='horizontal', size_hint=(1, None), height=height)
-
-        if not is_file:
-            icon_toggle_button = IconToggleButton(size_hint=(None, None),
-                                          size=(height,height))
-
-            icon_toggle_button.btn.bind(state=callback)
-            self.add_widget(icon_toggle_button)
-        else:
-            place_holder = Widget(size_hint=(None,None),size=(height,height))
-            self.add_widget(place_holder)
-
-        self.add_widget(labeled_checkbox)
-
-
-class LabeledCheckBox(BoxLayout):
-    indent = get_line_height()*1.5
-
-    def __init__(self, height : int, text :str, check_callback, is_file : bool, indent=0, **kwargs):
-        super().__init__(**kwargs)
-        self.size_hint_y = None
-        self.height = height
-
-        self.check_box = ImageCheckBox(size_hint=(None, None), size=(self.height, self.height))
-        self.check_box.bind(active=check_callback)
-
-        icon_path = get_fileicon_path() if is_file else get_foldericon_path()
-        self.icon = Image(source=icon_path, size_hint=(None, None), size=(self.height, self.height))
-        self.add_widget(self.icon)
-
-
-        self.label = BlackLabel(text=text,
-                                font_size=Window.width * 0.018,
-                                size_hint_x=None,
-                                size_hint_y=1,
-                                halign='left',
[…]

### 2024-11-22 submission_helper/elements/types.py

@@ -1,111 +1,0 @@
-from __future__ import annotations
-
-from kivy.clock import Clock
-from kivy.graphics import Color, Line
-from kivy.uix.boxlayout import BoxLayout
-from kivy.uix.label import Label
-from kivy.uix.slider import Slider
-from kivy.uix.textinput import TextInput
-from kivy.core.window import Window
-
-from kivy.uix.widget import Widget
-from submission_helper.configs import get_scaled_down_app_width
-from submission_helper.resources import get_logo_path, get_kivy_image
-from typing import List
-
-class FocusTextInput(TextInput):
-    def keyboard_on_key_down(self, window, keycode, text, modifiers):
-        if keycode[1] == 'tab':  # Check if the key is Tab
-            if 'shift' in modifiers:  # Check if Shift is also pressed
-                # Focus the previous widget
-                if self.focus_previous:
-                    self.focus_previous.focus = True
-            else:
-                # Focus the next widget
-                if self.focus_next:
-                    self.focus_next.focus = True
-            return True  # Indicate that the key was handled
-        return super(FocusTextInput, self).keyboard_on_key_down(window, keycode, text, modifiers)
-
-
-class AutoSizeLabel(Label):
-    def __init__(self, text: str = '', *args, **kwargs):
-        _ = args
-        super().__init__(**kwargs,text=text)
-        # self.set_label_width()
-
-        if not 'size_hint' in kwargs and not 'size' in kwargs:
-            Clock.schedule_once(lambda dt: self.set_label_size())
-
-
-    def set_label_size(self):
-        self.width = self.texture_size[0]
-        self.height = self.texture_size[1]
-        self.text_size = (self.width, self.height)
-
-
-class BlackLabel(AutoSizeLabel):
-    def __init__(self, text: str = '', *args, **kwargs):
-        _ = args
-        kwargs.setdefault('color',(0,0,0,1))
-        super().__init__(**kwargs, text=text)
-
-
-class ThickVerticalSlider(Slider):
-    def __init__(self, **kwargs):
-        super(ThickVerticalSlider, self).__init__(**kwargs)
-        self.bind(pos=self._update_canvas, size=self._update_canvas)
-
-    def _update_canvas(self, *args):
-        _ = args
-        self.canvas.before.clear()
-        with self.canvas.before:
-            Color(0.5, 0.5, 0.5, 1)  # Grey color for the track
-            line_width = 2  # Adjust width for thickness
[…]

### 2024-11-22 submission_helper/filesystem/__init__.py

@@ -1,3 +1,0 @@
-from .fsnode import FsNode
-from .out_files import zip_file_list, produce_csv_file
-from .lib import get_initial_path

### 2024-11-22 submission_helper/filesystem/fsnode.py

@@ -1,125 +1,0 @@
-from __future__ import annotations
-
-import threading
-import copy
-import os
-from abc import abstractmethod
-from typing import Optional
-import time
-
-from submission_helper.filesystem.lib import make_fsys_dict
-from kivy.clock import Clock
-
-# -------------------------------------------
-
-class FsNode:
-    xrd_formats : list = ['raw', 'dif', 'gdf', 'dat', 'ras', 'cpi', 'txt', 'plv', 'xrdml']
-
-    @classmethod
-    def set_default_formats(cls, format_text : str):
-        entries = format_text.split(',')
-        formats = [entry.strip('.').strip() for entry in entries if '.' in entry]
-        cls.xrd_formats = formats
-
-    def __init__(self, path : str):
-        self.path : str = path
-        self.name : str = os.path.basename(path)
-
-        self.potential_xrd_children : list = []
-        self.potential_xrd_des : list = []
-        self.xrd_node_des : list =[]
-
-        self.fsys_dict : Optional[dict] = None
-        self.cached_is_xrd_relevant : Optional[bool] = None
-
-        self.time_filesystem_searching = 0
-        self.time_relevancy_sorting = 0
-        self.callback = None
-
-
-    def prepare_fsys(self, on_done : callable):
-        self.callback = on_done
-        threading.Thread(target=self.init_fsys).start()
-
-    def init_fsys(self):
-        fsys_start = time.time()
-        self.initialize_potential_des()
-        self.time_filesystem_searching += time.time()-fsys_start
-
-        sorting_start = time.time()
-        self.find_xrd_relevant_nodes()
-        self.time_relevancy_sorting += time.time()-sorting_start
-
-        if self.callback:
-            Clock.schedule_once(callback=self.callback)
-
-    def initialize_potential_des(self):
-        for name in self.get_all_potential_sub():
-            child = self.add_child(name=name)
-            child.fsys_dict = self.fsys_dict[name]
-            self.potential_xrd_children += [child]
-
-        self.potential_xrd_des = copy.copy(self.potential_xrd_children)
-        for child in self.potential_xrd_children:
-            child.init_fsys()
-            self.potential_xrd_des += child.potential_xrd_des
-
-
-    def find_xrd_relevant_nodes(self):
-        for fsys_des in self.potential_xrd_des:
[…]

### 2024-11-22 submission_helper/filesystem/lib.py

@@ -1,29 +1,0 @@
-import platform, sys, os
-
-def get_initial_path():
-    if platform.system() == 'Windows':
-        initial_path = os.path.splitdrive(sys.executable)[0] + '\\'
-    else:
-        initial_path = '/'
-
-    return initial_path
-
-
-
-
-def make_fsys_dict(root_dir : str) -> dict:
-    file_structure = {}
-
-    for root, dirs, files in os.walk(root_dir, followlinks=True):
-        relative_path = root.replace(root_dir, '').lstrip(os.sep)
-        current_level = file_structure
-
-        parts = relative_path.split(os.sep)
-        for part in parts:
-            if part:
-                current_level = current_level.setdefault(part, {})
-
-        for file in files:
-            current_level[file] = None
-
-    return file_structure

### 2024-11-22 submission_helper/filesystem/out_files.py

@@ -1,64 +1,0 @@
-from submission_helper.resources import get_template_csv
-
-import os
-import zipfile, csv
-from typing import List
-
-
-# -------------------------------------------
-
-def zip_file_list(path_list: List[str], zipfile_path : str, root_path : str):
-    zip_path_dict = get_rel_path_dict(abs_path_list=path_list, root_path=root_path)
-
-    if os.path.isfile(zipfile_path):
-        print(f'Target path {zipfile_path} already exists. Aborting ...')
-        return
-
-    with zipfile.ZipFile(zipfile_path, 'a') as zipf:
-        for file_path in path_list:
-            zip_path = zip_path_dict[file_path]
-            zipf.write(file_path, zip_path)
-
-    print(f"ZIP archive created at {zipfile_path}")
-
-
-
-def produce_csv_file(abs_path_list : List[str], target_path : str, root_path : str):
-    rel_path_dict = get_rel_path_dict(abs_path_list=abs_path_list, root_path=root_path)
-    rel_path_list = list(rel_path_dict.values())
-    template_path = get_template_csv()
-
-    if os.path.isfile(target_path):
-        print(f'Target CSV path {target_path} already exists. Aborting ...')
-        return
-
-    with open(template_path, 'r') as file:
-        reader = csv.reader(file)
-        data = list(reader)
-
-    for i, string in enumerate(rel_path_list, start=2):
-        if i < len(data):
-            data[i][0] = string
-        else:
-            data.append([string])
-
-    with open(target_path, 'w', newline='') as file:
-        writer = csv.writer(file)
-        writer.writerows(data)
-
-    print(f"CSV file created at {target_path}")
-
-
-def get_rel_path_dict(abs_path_list: List[str], root_path : str) -> dict:
-    rel_path_dict = {}
-
-    for path in abs_path_list:
-        try:
-            rel_path = os.path.relpath(path, root_path)
-            rel_path_dict[path] = rel_path
-
-        except Exception as e:
-            print(f"Error processing path '{path}': {e}")
-
-    return rel_path_dict
-

### 2024-11-22 submission_helper/layouts/__init__.py

@@ -1,3 +1,0 @@
-from .finish import FinishLayout
-from .selection import SelectionLayout
-from .dialogue import InputDialog

### 2024-11-22 submission_helper/layouts/dialogue.py

@@ -1,177 +1,0 @@
-from __future__ import annotations
-
-import os
-
-from kivy.uix.widget import Widget
-from kivy.clock import Clock
-from kivy.core.window import Window
-from kivy.uix.boxlayout import BoxLayout
-from kivy.uix.button import Button
-from kivy.uix.label import Label
-from kivy.uix.popup import Popup
-
-from submission_helper.configs import get_true_width, get_true_height
-from submission_helper.elements.types import FocusTextInput
-from submission_helper.filesystem import FsNode, get_initial_path
-from submission_helper.resources.resource_manager import get_blended_logo_path, get_kivy_image
-# -------------------------------------------
-
-class InputDialog(Popup):
-    font_size = get_true_width() * 0.0145
-
-    def __init__(self, callback: callable, **kwargs):
-        super(InputDialog, self).__init__(**kwargs)
-        self.size_hint = (0.65, 0.825)
-        self.title = 'XRD data collector'
-        self.title_align = 'center'
-        self.auto_dismiss = False
-        self.callback = callback
-        self.background_color = (100 / 255, 1, 1, 1)  # (R, G, B, A)
-
-        h_space = 12 * get_true_width() / 1600.
-        v_space = 12 * get_true_height() / 900.
-        self.content = BoxLayout(orientation='vertical',
-                                 padding=[h_space, v_space, h_space, v_space],
-                                 spacing=v_space)
-
-        logo_image = get_kivy_image(width=Window.width*0.4, imgPath=get_blended_logo_path(), size_hint =(1,0.25))
-        first_hint = self.make_hint()
-        self.format_input =self.make_format_input()
-        second_hint = self.make_second_hint()
-
-        self.notice = self.make_notice()
-        self.input_folder_hint = self.make_input_folder_hint()
-        self.path_input = self.make_path_input()
-        confirm_button = self.make_confirm_button()
-
-        for widget in [logo_image,first_hint,self.format_input,second_hint,self.notice,self.input_folder_hint,self.path_input,confirm_button]:
-            self.content.add_widget(widget=widget)
-        self.set_focus_chain()
-
-
-    def set_focus_chain(self):
-        self.format_input.focus_next = self.path_input
-        self.format_input.focus_previous = self.path_input
-
-        self.path_input.focus_next = self.format_input
-        self.path_input.focus_previous = self.format_input
-
-
-    # -------------------------------------------
-    # Callbacks
-
-    @staticmethod
[…]

### 2024-11-22 submission_helper/layouts/finish.py

@@ -1,79 +1,0 @@
-from __future__ import annotations
-
-import os
-from kivy.uix.button import Button
-from kivy.uix.boxlayout import BoxLayout
-from kivy.core.window import Window
-from kivy.uix.textinput import TextInput
-from kivy.uix.widget import Widget
-
-from submission_helper.elements.types import BlackLabel
-from kivy.uix.popup import Popup
-from kivy.uix.label import Label
-from submission_helper.configs import get_true_height, get_true_width
-# -------------------------------------------
-
-class FinishLayout(BoxLayout):
-    def __init__(self, callback: callable, exit_funct : callable, **kwargs):
-
-        l = 15 * get_true_height() / 900.
-        super(FinishLayout, self).__init__(orientation='vertical', size_hint=(1, 0.085),padding=(0, l, 0, l),**kwargs)
-
-        s = 10 * get_true_width() / 1600.
-        upper_finish = BoxLayout(orientation='horizontal', size_hint=(1, 0.3),spacing=s)
-        note = BlackLabel(text='Target folder:', size_hint=(0.2, 1), font_size=Window.width * 0.02, bold=True)
-        self.target_path_input = TextInput(text=f'{os.getcwd()}',
-                                           size_hint=(0.7, 1),
-                                           font_size=get_true_width() * 0.02,
-                                           multiline=False)
-
-        buffer = BlackLabel(text='', size_hint=(0.075, 1))
-        self.ok_button = self.get_ok_button()
-        self.ok_button.bind(on_press=callback)
-
-        upper_finish.add_widget(note)
-        upper_finish.add_widget(self.target_path_input)
-        upper_finish.add_widget(self.ok_button)
-        upper_finish.add_widget(buffer)
-
-        self.exit_funct = exit_funct
-        self.default_font_size = get_true_width() * 0.018
-        self.feedback_label = self.get_feedback_label(font_size=self.default_font_size)
-        self.feedback_popup = self.get_feedback_popup()
-
-        self.add_widget(upper_finish)
-
-
-    def show(self):
-        self.feedback_popup.open()
-
-    def get_feedback_popup(self):
-        container = BoxLayout(orientation='vertical')
-        container.add_widget(self.feedback_label)
-
-        button_layout = BoxLayout(size_hint_y=0.2)
-        dismiss_button = Button(text='Dismiss', on_press=lambda instance: self.feedback_popup.dismiss())
-        exit_button = Button(text='Exit', on_press=self.exit_funct)
-
[…]

### 2024-11-22 submission_helper/layouts/selection.py

@@ -1,157 +1,0 @@
-from __future__ import annotations
-
-from typing import Optional
-from kivy.uix.label import Label
-from kivy.uix.scrollview import ScrollView
-
-from kivy.uix.boxlayout import BoxLayout
-from kivy.clock import Clock
-
-
-from submission_helper.filesystem.fsnode import FsNode
-from submission_helper.elements.node_element import NodeElement
-from submission_helper.elements.types import ThickVerticalSlider, HeaderWidget, Placeholder
-
-from submission_helper.configs import get_line_height, get_true_height
-import intervals as I
-
-# -------------------------------------------
-
-class SelectionLayout(BoxLayout):
-    def __init__(self, **kwargs):
-        super(SelectionLayout, self).__init__(orientation='horizontal', size_hint=(1, 0.9), **kwargs)
-
-        # attributes
-        self.root_checkbox : Optional[NodeElement] = None
-        self.last_load_range = I.closed(float('-inf'),float('inf'))
-        self.dismiss_popup = None
-
-        # GUI elements
-        self.slider = ThickVerticalSlider(orientation='vertical', min=0, max=1, value=1, size_hint=(0.1, 1))
-        self.slider.bind(value=self.do_scroll)
-
-        self.scroll_view = self.get_scroll_view()
-        self.scroll_view.bind(scroll_y=self.do_scroll)
-
-        self.header_layout = self.get_header_widget(num_elements=0)
-        self.checkboxes_layout = self.get_checkboxes_layout(file_count_label=self.header_layout,
-                                                            scroll_view=self.scroll_view)
-
-        self.add_widget(self.checkboxes_layout)
-        self.add_widget(self.slider)
-
-
-    def calculate_content(self, path : str):
-        self.root_checkbox: NodeElement = NodeElement(path=path, height=get_line_height(), scroll_view=self.scroll_view)
-        self.root_checkbox.prepare_fsys(on_done=self.set_content)
-
-
-    def set_content(self, *args, **kwargs):
-        _, __ = args, kwargs
-        print(f'Time taken to initialize filesystem: {self.root_checkbox.time_filesystem_searching}')
-        print(f'Time taken to search relevancy: {self.root_checkbox.time_relevancy_sorting}')
-
-        self.header_layout.update_number_of_elements(num_elements=len(self.root_checkbox.get_xrd_file_des()))
-        self.scroll_view.add_widget(widget=self.get_scroll_view_content(root_checkbox=self.root_checkbox))
-
-        update_rate = 0.2
[…]

### 2024-11-22 submission_helper/package.sh

@@ -1,28 +1,0 @@
-#!/bin/bash
-#using: pyhon3.X -m venv legacyVenv
-
-#setup before executing package.sh:
-
-#sudo apt install python 3.X python3.X-dev python3.X-venv
-#python3.X -m venv legacyVenv
-#source legacyVenv/bin/activate
-#sudo apt install binutils
-#pip install -r requirements.txt
-
-set -e
-
-# Define the root directory where the script and build should be located
-ROOT_DIR="$HOME/xrd_data_collect"
-
-if [ ! -d "$ROOT_DIR" ]; then
-    mkdir -p "$ROOT_DIR"
-fi
-
-
-DIST_DIR="$ROOT_DIR/dist"
-BUILD_DIR="$ROOT_DIR/build"
-script_dir="$(cd "$(dirname "$0")" && pwd)"
-
-pyinstaller "prod_run.spec" --distpath "$DIST_DIR" --workpath "$BUILD_DIR"
-echo "-> Built executable under $ROOT_DIR/dist"
-echo "done"

### 2024-11-22 submission_helper/resources/__init__.py

@@ -1,1 +1,0 @@
-from .resource_manager import *

### 2024-11-22 submission_helper/resources/resource_manager.py

@@ -1,76 +1,0 @@
-import os
-import sys
-
-from kivy.core.image import Image as CoreImage
-from kivy.uix.image import Image as kiImage
-from io import BytesIO
-from PIL import Image
-
-
-if getattr(sys, 'frozen', False):
-    # The app is bundled with PyInstaller
-    base_path = sys._MEIPASS  # PyInstaller's base directory
-else:
-    # The app is run as a script
-    base_path = os.path.dirname(os.path.abspath(__file__))
-
-images_folder_path = os.path.join(base_path, 'images')
-doc_folder_path = os.path.join(base_path, 'documents')
-
-# -------------------------------------------
-# images
-
-def get_blended_logo_path() -> str:
-    return os.path.join(images_folder_path, 'blended_logo.png')
-
-def get_logo_path() -> str:
-    return os.path.join(images_folder_path, 'logo.png')
-
-def get_foldericon_path() -> str:
-    return os.path.join(images_folder_path, 'folder.png')
-
-def get_fileicon_path() -> str:
-    return os.path.join(images_folder_path, 'file.png')
-
-def get_collapsed_icon_path() -> str:
-    return os.path.join(images_folder_path, 'collapsed.png')
-
-def get_expanded_icon_path() -> str:
-    return os.path.join(images_folder_path, 'expanded.png')
-
-def get_loading_icon_path() -> str:
-    return os.path.join(images_folder_path, 'loading.png')
-
-def get_checked_box_path() -> str:
-    return os.path.join(images_folder_path, 'checked_box.png')
-
-def get_unchecked_box_path() -> str:
-    return os.path.join(images_folder_path, 'unchecked_box.png')
-
-# -------------------------------------------
-# documents
-
-def get_template_csv() -> str:
-    return os.path.join(doc_folder_path, 'template.csv')
-
-def get_empty_path() -> str:
-    return os.path.join(images_folder_path, 'empty.png')
-
-# -------------------------------------------
-# resizing with PIL instead of kivy
-
-def pil_to_kivy_image(pil_image, **kwargs):
-    data = BytesIO()
-    pil_image.save(data, format='png')
-    data.seek(0)
-    core_image = CoreImage(BytesIO(data.read()), ext='png')
-    kivy_image = kiImage(**kwargs)
-    kivy_image.texture = core_image.texture
-    return kivy_image
-
-def get_kivy_image(width, imgPath,**kwargs):
-    pil_image = Image.open(imgPath)
[…]

### 2024-11-22 submission_helper/run/dev_run.py

@@ -1,29 +1,0 @@
-from submission_helper.configs import set_configs
-set_configs()
-
-from kivy.core.window import Window
-Window.clearcolor = (1, 1, 1, 1)
-
-from kivy.modules import inspector
-from submission_helper.app import DataCollectApp
-# -------------------------------------------
-
-
-def main():
-    win_test_dir = 'C:\\Users\\daniel\\OneDrive\\Desktop\\test_folder'
-    win_target_test = 'C:\\Users\\daniel\\OneDrive\\Desktop'
-
-    linux_input_test = '/home/daniel/OneDrive'
-    linux_target_test = f'/home/daniel/aimat/tool_target_folder'
-
-
-    _, __ = win_test_dir, win_target_test
-    _, __ = linux_input_test, linux_target_test
-
-    app = DataCollectApp(override_input_folder=None,
-                         override_target_folder=linux_target_test)
-    inspector.create_inspector(Window, app)
-
-    app.run()
-
-main()

### 2024-11-22 submission_helper/run/prod_run.py

@@ -1,16 +1,0 @@
-from submission_helper.configs import set_configs
-set_configs()
-
-from kivy.core.window import Window
-Window.clearcolor = (1, 1, 1, 1)
-from submission_helper.app import DataCollectApp
-# -------------------------------------------
-
-
-def main():
-    app = DataCollectApp(override_input_folder=None,
-                         override_target_folder=None)
-
-    app.run()
-
-main()

### 2024-11-22 chapters/02_existing_datasets.tex
Rearranged paragraphs to fit KIT contributions next to each other; Added reference to database history from IUCr
Added paragraph on Pearson's crystal data

@@
 \footnote{We reached out to the contact address given on their website, info@icdd.com.} Still, this makes the PDF5+, to the best of our knowledge, the largest collection of experimental powder diffraction data available to researchers.
-As of November 2024, the PDF5+ is available to researchers through a purchase starting at a price point of \$6,265.00.\\
+As of November 2024, the PDF5+ is available to researchers through a purchase of a one-year license starting at a price point of \$6,265.00.\\
+
+\textbf{Pearson's Crystal data}:\footnote{More information on Pearson's crystal data can be found on the ASM International website under \url{https://www.asminternational.org/materials-resources/online-databases/pearsons-crystal-data-crystal-structure-database-for-inorganic-compounds/?srsltid=AfmBOorjpwJSVNsaWtCO4kmgu-Ox2ncF3tkzwCXRE2GTemYIxsSLp8xr}.} Pearson's crystal data is a commercial database of 395000 crystal structures including 21700 experimental powder diffraction patterns maintained by ASM International.
+Pearson's crystal data is based on the crystallographic data of another database, the Linus Pauling file.
+\cite{kaduk2007} Since the Linus Pauling file is also a database source for the PDF, there is likely a significant overlap between the experimental powder diffraction data found in Pearson's crystal data and that found in the PDF.
+As of November 2024, Pearson's crystal data is available to researchers through a purchase of a one-year license starting at a price point of 2200 \euro.\\
 
@@
 
-\textbf{Pearson's Crystal data}:
-
 Some major crystal structure databases, such as the ICSD, do not appear in the above list, as they do not include powder diffraction data.

### 2024-11-25 chapters/02_existing_datasets.tex
Changed plaintext SI units/degrees into latex symbols

@@
 % Datset listing (List written on 20.06.24)
-% Lists of Crystallographic databases; Our list contains all of the databases given in the lists linked below
+% Lists of Crystallographic databases; The commented list contains all of the databases given in the lists linked below
 % - IUCR: "Crystallographic databases and related resources" (https://www.iucr.org/resources/data/databases)
@@
 The probed datasets that did contain powder diffraction data only contained <= 30 patterns and only one of them came with corresponding crystal structures
-
-%TODO: Summarize the existing datasets shortly, including number of entries, etc.
 
@@
 
+There is also much publicly available powder diffraction data uploaded to datasets on Zenodo.
+However, this data is split into disparate entries that typically only contain the work of a single research project.
+Additionally, extracting powder diffraction data at scale is hindered by the fact that the data is often given in plain text files in non-standardized formats, which are difficult to parse programmatically.
 Some major crystal structure databases, such as the ICSD, do not appear in the above list, as they do not include powder diffraction data.

### 2024-11-25 chapters/04_our_dataset.tex
Overhauled PDF entry in existing datasets
Rearranged paragraphs to fit KIT contributions next to each other; Added reference to database history from IUCr
Changed plaintext SI units/degrees into latex symbols

@@
 The following texts detail the contribution of each research group.
-Most data was collected with copper radiation sources which exhibit a $K_{\alpha1}$ wavelength of $\lambda$~=~1.54056~{\AA} and a $K_{\alpha2}$ wavelength of $\lambda$~=~1.54439~{\AA}.
+Most data was collected with copper sources which have a  $K_{\alpha1}$ wavelength of $\lambda=1.54056\text{\AA}$ and a $K_{\alpha2}$ wavelength of $\lambda=1.54439\text{\AA}$.
 
@@
 
-The pXRD data from the COD database are of high quality, with a median resolution of $\Delta(2\theta) = 0.013${\textdegree} and an average number of 9190 points measured per pattern.
+The pXRD data from the COD database are of high quality, with a median resolution of $\Delta(2\theta) = 0.013\degree$ and an average number of 9190 points measured per pattern.
 They span a wide chemical space, including organic, inorganic, and hybrid structures, including 75 different elements of the periodic table.
@@
 Some air-sensitive samples were measured using a transparent polymer dome for protection.
-This dome led to increased background noise over the first 20° and slightly decreased pattern resolution.
+This dome led to increased background noise over the first $20\degree$ and slightly decreased pattern resolution.
 \\
-
-% Paragraph by Andrea Hodge
-\subsubsection*{Department of Chemical Engineering and Materials Science, USC}
-
-Combinatorial CuNi and CuAl samples were synthesized at the University of Southern California (USC) via magnetron co-sputtering from Cu (99.999\%) and Ni (99.995\%) or Al (99.999\%) targets (Plasmaterials) onto stationary 10 cm Si (100) substrates yielding 169 unique 5 x 5mm squares per wafer.
-These alloys were selected to study compositional effects on stacking fault and growth twin boundary formation in sputtered materials \cite{2024AcMat.27019839A,alwen2024combinatorial}.
-XRD spectra were collected using a Malvern Panalytical Empyrean X-ray diffractometer at the Forschungszentrum Julich (FZJ) Institute for Energy and Climate Research, Structure and Function of Materials (IEK-2), which enabled automated XRD characterization due to its programmable XY stage and laser sensor for Z height correction.
-XRD analysis was used to link changes in defect formation in the sputtered films with varied phase and crystal structures.
-Each composition was characterized using CuK X-rays (1.5406Å wavelength) that were collimated to probe an area of ~4 x 4 mm using a step size of 0.026° and scan duration of 0.3 seconds per step over a 2 range from 30° - 110°.
-Measurement parameters were chosen to resolve at least six or more points above the full-width half maximum of each XRD peak.
 
@@
 The latter were synthesized via the surfactant-free benzyl alcohol route \cite{Wolf2019, Wolf2018}.
-The cobalt oxide (CoO or Co3O4) and cerium oxide (CeO2) nanoparticles were in the size range of 4-16 nm according to the Scherrer equation.
+The cobalt oxide (CoO or Co3O4) and cerium oxide (CeO2) nanoparticles were in the size range of $4-16 \ \si{nm}$ according to the Scherrer equation.
 A series of porous Al2O3 materials, which were prepared by calcination of boehmite (AlOOH) at various temperatures, represents crystalline samples with limited long-range structure and various contributions of Al2O3 polymorphs.\\
@@
 X-ray diffraction (XRD) was conducted with an X’Pert Pro MPD (Panalytical) in Bragg-Brentano geometry using a Cu X-ray source.
-The patterns were acquired from 5-80° $2\theta$ with a step size of 0.016711° or 0.033420° and a total acquisition time of 40 to 120 min.
+The patterns were acquired in the $2\theta$ range of $5-80\degree$ with a step size of $0.016711\degree$ or $0.033420\degree$ and a total acquisition time of 40 to 120 min.
 \\
+
+% Paragraph by Adie Alwen and Andrea Hodge
+\subsubsection*{Department of Chemical Engineering and Materials Science, USC}
+
+Combinatorial CuNi and CuAl samples were synthesized at the University of Southern California (USC) via magnetron co-sputtering from Cu (99.999\%) and Ni (99.995\%) or Al (99.999\%) targets (Plasmaterials) onto stationary $10 \si{cm}$ Si (100) substrates yielding 169 unique $5 \times 5 \ \si{mm}$ squares per wafer.
+These alloys were selected to study compositional effects on stacking fault and growth twin boundary formation in sputtered materials \cite{2024AcMat.27019839A,alwen2024combinatorial}.
+XRD spectra were collected using a Malvern Panalytical Empyrean X-ray diffractometer at the Forschungszentrum Julich (FZJ) Institute for Energy and Climate Research, Structure and Function of Materials (IEK-2), which enabled automated XRD characterization due to its programmable XY stage and laser sensor for Z height correction.
+XRD analysis was used to link changes in defect formation in the sputtered films with varied phases and crystal structures.
+Each composition was characterized using CuK X-rays that were collimated to probe an area of $4 \times 4 \ \si{mm}$ using a step size of $0.026\degree$ and scan duration of $0.3$ seconds per step over a 2 range from $30\degree - 110\degree$.
+Measurement parameters were chosen to resolve at least six or more points above the full-width half maximum of each XRD peak.
 
@@
 A second dataset was collected from spin-coating metal-halide perovskite precursor solutions with varying compositions of \ce{MAPb(I_{1-x}Br_x)3} (where MA = Methylammonium and x = 0, 0.33, 0.5, 0.67, 1) spin-coated onto glass substrates.
-The substrates were preheated to different temperatures (30°C, 50°C, 70°C, and 90°C), and the spin-coating process was performed at a constant temperature on the preheated substrates.
-For both datasets, diffraction data were continuously measured during spin-coating, chemical induction of crystallization, and annealing of the samples (at 100$^{\circ}$C and 110$^{\circ}$C respectively) with a frequency of about 0.56 1/s and 0.54 1/s.
+The substrates were preheated to different temperatures ($30\degree $C, $50\degree $C, $70\degree $C, and $90\degree $C), and the spin-coating process was performed at a constant temperature on the preheated substrates.
+For both datasets, diffraction data were continuously measured during spin-coating, chemical induction of crystallization, and annealing of the samples (at $100\degree$C and $110\degree$C respectively) with a frequency of about $0.56
+ \ 1/\si{s}$ and $0.54 \ 1/\si{s}$.
 Each in situ measurement consisted of about 500 to 1000 individual diffractograms.
@@
 Experimental XRD data were collected at beamline 12.3.2 of the Advanced Light Source, the synchrotron at Lawrence Berkeley National Laboratory.
-The data were collected using a photon energy of 10 keV ($\lambda$ = 1.23984 \AA), selected using a Si(111) monochromator.
+The data were collected using a photon energy of 10 keV ($\lambda = 1.23984 \text{\AA}$), selected using a Si(111) monochromator.
 Measurements were taken in grazing incidence geometry, i.e.
-using a beam incidence angle of 1$^{\circ}$.
-Two-dimensional diffraction images were recorded using a Dectris Pilatus 1M area detector at an angle between 34$^{\circ}$ and 36$^{\circ}$ with a sample-to-detector distance of roughly 190 mm.
+using a beam incidence angle of $1\degree$ Two-dimensional diffraction images were recorded using a Dectris Pilatus 1M area detector at an angle between $34\degree$ and $36\degree$ with a sample-to-detector distance of roughly $190 \ \si{mm}$.
 The two-dimensional data were calibrated using an Al$_{2}$O$_{3}$ calibration standard and integrated along the azimuthal angle.
@@
 Each combinatorial Zn–V–N library was synthesized using radio-frequency co-sputtering of Zn and V in a mixed Ar and N\textsubscript{2} plasma.
-An orthogonal deposition temperature and composition gradient was created, resulting in a deposition temperature of 220~{\textdegree}C for samples 1~–~9 and 114~{\textdegree}C for samples 37~–~45.
+An orthogonal deposition temperature and composition gradient was created, resulting in a deposition temperature of $220\degree$C for samples 1~–~9 and $114\degree$C for samples 37~–~45.
 The composition for each sample was determined using X-ray fluorescence (XRF) spectroscopy which was further calibrated through Rutherford backscattering spectroscopy (RBS) based on selected samples.
@@
 The relative phase fractions were quantified using supplementary azimuthal angle scans coupled with structural factors and geometrical factors as reported elsewhere.\cite{Wieczorek2024} Fully inorganic lead perovskite libraries were prepared using thermal co-evaporation of lead and cesium halide salts.
-All metal halide perovskite libraries were measured within a custom-made X-Ray transparent inert-gas dome, resulting in the presence of minor additional features within the $\theta$~=~19~–~31{\textdegree} range.
+All metal halide perovskite libraries were measured within a custom-made X-Ray transparent inert-gas dome, resulting in the presence of minor additional features within the $\theta=19 \text{~–~}31\degree$ range.
 For all combinatorial libraries where any phases are specified, the complete set of phases is reported in the metadata.

### 2024-11-25 chapters/01_introduction.tex
Moved citations before punctuation

@@
 % High throughput experimentation
-The advent of high-throughput experiments holds the prospect of significantly accelerating the speed of materials discovery.
-\cite{Liu2019} The synthesis and characterization of novel materials are becoming increasingly efficient and automated, increasing the throughput of samples in experimentation pipelines.
-\cite{MacLeod2019, Ludwig2019, Ozaki2020} \\
+The advent of high-throughput experiments holds the prospect of significantly accelerating the speed of materials discovery\cite{Liu2019}.
+The synthesis and characterization of novel materials are becoming increasingly efficient and automated, increasing the throughput of samples in experimentation pipelines\cite{MacLeod2019, Ludwig2019, Ozaki2020}.
+\\
 
@@
 To determine crystal structure from pXRD data, an initial guess is optimized using Rietveld refinement.
-In this process, an initial crystal structure including unit cell parameters is provided by the operator of the analysis software and fitted to the observed diffractogram in an iterative optimization process based on simulated forward prediction of expected pXRD diffractograms given the current crystal structure.
-\cite{Dinnebier2019, Cano2021} As Rietveld refinement is a local optimization method, the result of the refinement procedure is only as good as the initial guess it was provided with.
+In this process, an initial crystal structure including unit cell parameters is provided by the operator of the analysis software and fitted to the observed diffractogram in an iterative optimization process based on simulated forward prediction of expected pXRD diffractograms given the current crystal structure\cite{Dinnebier2019, Cano2021}.
+As Rietveld refinement is a local optimization method, the result of the refinement procedure is only as good as the initial guess it was provided with.
 \\
@@
 However, attempting to refine all crystal structure parameters at once, starting from this initial structure, is known to lead to unphysical results.
-What makes the refinement process time-consuming and difficult is finding the correct order in which to refine the crystal structure parameters.
-\cite{Ozaki2020}  Additionally, if the fit develops into an unphysical set of parameters in this iterative determination of parameters, the operator must intervene to backtrack to the last step.
+What makes the refinement process time-consuming and difficult is finding the correct order in which to refine the crystal structure parameters\cite{Ozaki2020}.
+Additionally, if the fit develops into an unphysical set of parameters in this iterative determination of parameters, the operator must intervene to backtrack to the last step.
 \\
@@
 Commonly used databases for obtaining a match include the Powder Diffraction File\cite{GatesRector2019}, the Inorganic Crystal Structure Database \cite{Belsky2002}, and the Crystallography Open Database \cite{Grazulis2009}.
-An initial guess obtained from such a database is by no means guaranteed to lead to an accurate structure solution, even if it is iteratively revised, either manually or using blackbox optimization.\cite{Ozaki2020} Especially for novel observed structure types, simply matching to a database of known structures will not yield good results.
+An initial guess obtained from such a database is by no means guaranteed to lead to an accurate structure solution, even if it is iteratively revised, either manually or using blackbox optimization\cite{Ozaki2020}.
+Especially for novel observed structure types, simply matching to a database of known structures will not yield good results.
 Therefore, more complex crystal structure solution approaches need to be applied to find suitable crystal structure guesses.
@@
 
-Machine learning has the potential to speed up the manual analysis of powder diffractograms and keep pace with an automated high-throughput experimentation environment. \cite{Agrawal2019, Surdu2023}
+Machine learning has the potential to speed up the manual analysis of powder diffractograms and keep pace with an automated high-throughput experimentation environment\cite{Agrawal2019, Surdu2023}.
 Models can be trained to predict crystal structure information directly given a diffractogram, or models can be used to automate the conventional workflow of finding an initial guess \cite{Surdu2023} and running ML-automated Rietveld refinement \cite{Feng2019}.
-So far, however, due to an absence of labeled datasets with experimental diffractograms\cite{Wang2020}, machine learning in this domain has largely relied on diffractograms simulated from known structures\cite{Park2017, Lee2023} or, most recently, from generated synthetic crystals.
-\cite{Schopmans2023} \\
+So far, however, due to an absence of labeled datasets with experimental diffractograms\cite{Wang2020}, machine learning in this domain has largely relied on diffractograms simulated from known structures\cite{Park2017, Lee2023} or, most recently, from generated synthetic crystals\cite{Schopmans2023}.
+\\
 
@@
 However, even small labeled and also unlabeled experimental diffractograms hold significant value in practice.
-First of all, labeled data can be used to better test and benchmark existing and new automated analysis approaches, which yields a better comparison than when using the very limited currently available datasets, such as the commonly used RRUFF mineral dataset.
-\cite{Armbruster2015} \\
+First of all, labeled data can be used to better test and benchmark existing and new automated analysis approaches, which yields a better comparison than when using the very limited currently available datasets, such as the commonly used RRUFF mineral dataset\cite{Armbruster2015}.
+\\
 

### 2024-11-25 chapters/02_existing_datasets.tex
Moved citations before punctuation

@@
 
-\textbf{Powder Diffraction File:}\footnote{More information on the Powder Diffraction File can be found under \url{https://www.icdd.com/pdf-5/}.} The Powder Diffraction File (PDF), published and maintained by the International Center for Diffraction Data (ICDD), is a large collection of materials with accompanying powder diffraction data first published in 1941.
-\cite{GatesRector2019} Its latest release as of November 2024, the PDF5+, contains over a million materials with accompanying powder diffraction data.
+\textbf{Powder Diffraction File:}\footnote{More information on the Powder Diffraction File can be found under \url{https://www.icdd.com/pdf-5/}.} The Powder Diffraction File (PDF), published and maintained by the International Center for Diffraction Data (ICDD), is a large collection of materials with accompanying powder diffraction data first published in 1941\cite{GatesRector2019}.
+Its latest release as of November 2024, the PDF5+, contains over a million materials with accompanying powder diffraction data.
 However, most of these powder diffraction patterns are simulated.
@@
 \textbf{Pearson's Crystal data}:\footnote{More information on Pearson's crystal data can be found on the ASM International website under \url{https://www.asminternational.org/materials-resources/online-databases/pearsons-crystal-data-crystal-structure-database-for-inorganic-compounds/?srsltid=AfmBOorjpwJSVNsaWtCO4kmgu-Ox2ncF3tkzwCXRE2GTemYIxsSLp8xr}.} Pearson's crystal data is a commercial database of 395000 crystal structures including 21700 experimental powder diffraction patterns maintained by ASM International.
-Pearson's crystal data is based on the crystallographic data of another database, the Linus Pauling file.
-\cite{kaduk2007} Since the Linus Pauling file is also a database source for the PDF, there is likely a significant overlap between the experimental powder diffraction data found in Pearson's crystal data and that found in the PDF.
+Pearson's crystal data is based on the crystallographic data of another database, the Linus Pauling file\cite{kaduk2007}.
+Since the Linus Pauling file is also a database source for the PDF, there is likely a significant overlap between the experimental powder diffraction data found in Pearson's crystal data and that found in the PDF.
 As of November 2024, Pearson's crystal data is available to researchers through a purchase of a one-year license starting at a price point of 2200 \euro.\\
 
-\textbf{Crystallography Open Database:}\footnote{More information on the crystallography open database can be found under \url{https://www.crystallography.net/cod/}.} The Crystallography Open Database (COD) is an open-access collection of crystal structures founded in 2003.
-\cite{Graulis2009cod} It currently provides over 500.000 crystal structures in the form of .cif files.
+\textbf{Crystallography Open Database:}\footnote{More information on the crystallography open database can be found under \url{https://www.crystallography.net/cod/}.} The Crystallography Open Database (COD) is an open-access collection of crystal structures founded in 2003\cite{Graulis2009cod}.
+It currently provides over 500.000 crystal structures in the form of .cif files.
 Of these files, 1063 contain the experimental powder diffraction data that was used to determine the underlying crystal structures of the investigated samples.

### 2024-11-25 chapters/03_SOTA_ML.tex
Rearranged paragraphs to fit KIT contributions next to each other; Added reference to database history from IUCr
Moved citations before punctuation

@@
-%TODO:
-% - Papers to include: https://pubs.acs.org/doi/10.1021/jacs.4c10244
+%TODO: Papers to include:
+% -  https://pubs.acs.org/doi/10.1021/jacs.4c10244
 % - Sagado et al.: Automated classification of big X-ray diffraction data using deep learning models
@@
 Park {\it et al.} reached a test accuracy of roughly 81\% for a CNN, which classifies space groups of simulated single-phase diffractograms \cite{Park2017}.
-A regression analysis on lattice parameters within a broader framework encompassing all material classes was conducted by Chitturi et al.\cite{Chitturi2021} They developed a distinct CNN for each crystal system, utilizing a merged dataset from both the ICSD and the Cambridge Structural Database, and managed to achieve a mean absolute percentage error of about \SI{10}{\percent} for the lattice lengths, although they encountered difficulties in accurately predicting angles.
+A regression analysis on lattice parameters within a broader framework encompassing all material classes was conducted by Chitturi et al\cite{Chitturi2021}.
+They developed a distinct CNN for each crystal system, utilizing a merged dataset from both the ICSD and the Cambridge Structural Database, and managed to achieve a mean absolute percentage error of about \SI{10}{\percent} for the lattice lengths, although they encountered difficulties in accurately predicting angles.
 Zhang {\it et al.} introduce a convolutional self-attention neural network trained on simulated patterns to classify crystal types \cite{zhang2024crystallographic}.

### 2024-11-25 chapters/04_our_dataset.tex
Moved citations before punctuation

@@
 
-We extracted experimental pXRD data from the Crystallography Open Database (COD).\cite{Grazulis2009, Vaitkus2023} The COD is, to our knowledge, the largest open-access collection of experimental crystal structures of organic, inorganic, and metal-organic compounds and minerals, containing more than 500,000 entries.
+We extracted experimental pXRD data from the Crystallography Open Database (COD)\cite{Grazulis2009, Vaitkus2023}.
+The COD is, to our knowledge, the largest open-access collection of experimental crystal structures of organic, inorganic, and metal-organic compounds and minerals, containing more than 500,000 entries.
 \footnote{Available online at \url{https://www.crystallography.net/cod/}.} The data in the COD is placed in the public domain and licensed under the CC0 License.
@@
 The composition for each sample was determined using X-ray fluorescence (XRF) spectroscopy which was further calibrated through Rutherford backscattering spectroscopy (RBS) based on selected samples.
-The newly identified and isolated semiconductor Zn\textsubscript{2}VN\textsubscript{3} was identified to exhibit a cation-disordered wurtzite structure as verified by additional GI-XRD and SAED measurements.
-\cite{Zhuk2021} \\
-Tin halide perovskites were deposited using single-step spin-coating as reported elsewhere.\cite{Wieczorek2023} Methylammonium lead iodide libraries with varying degrees of residual PbI\textsubscript{2} were deposited using a two-step procedure involving both thermal evaporation of PbI\textsubscript{2} and subsequent spin-coating of a methylammonium solution.
-The relative phase fractions were quantified using supplementary azimuthal angle scans coupled with structural factors and geometrical factors as reported elsewhere.\cite{Wieczorek2024} Fully inorganic lead perovskite libraries were prepared using thermal co-evaporation of lead and cesium halide salts.
-All metal halide perovskite libraries were measured within a custom-made X-Ray transparent inert-gas dome, resulting in the presence of minor additional features within the $\theta=19 \text{~–~}31\degree$ range.
+The newly identified and isolated semiconductor Zn\textsubscript{2}VN\textsubscript{3} was identified to exhibit a cation-disordered wurtzite structure as verified by additional GI-XRD and SAED measurements\cite{Zhuk2021}.
+\\
+Tin halide perovskites were deposited using single-step spin-coating as reported elsewhere\cite{Wieczorek2023}.
+Methylammonium lead iodide libraries with varying degrees of residual PbI\textsubscript{2} were deposited using a two-step procedure involving both thermal evaporation of PbI\textsubscript{2} and subsequent spin-coating of a methylammonium solution.
+The relative phase fractions were quantified using supplementary azimuthal angle scans coupled with structural factors and geometrical factors as reported elsewhere\cite{Wieczorek2024}.
+Fully inorganic lead perovskite libraries were prepared using thermal co-evaporation of lead and cesium halide salts.
+All metal halide perovskite libraries were measured within a custom-made X-Ray transparent inert-gas dome, resulting in the presence of minor additional features within the $\theta=19 \text{–}31\degree$ range.
 For all combinatorial libraries where any phases are specified, the complete set of phases is reported in the metadata.

### 2024-11-25 chapters/02_existing_datasets.tex
Delete commented list on crystal structure databases i.e. experimental powder X-ray diffractino candidates
Warpped chemical formulas in \ce{} environemnt

@@
-% Datset listing (List written on 20.06.24)
-% Lists of Crystallographic databases; The commented list contains all of the databases given in the lists linked below
-% - IUCR: "Crystallographic databases and related resources" (https://www.iucr.org/resources/data/databases)
-% - Bruno et al. (2017): "Crystallography and Databases" (https://datascience.codata.org/articles/10.5334/dsj-2017-038)
-% - University of Virginia "Crystallographic data" (https://guides.lib.virginia.edu/c.php?g=514782&p=3568808)
-
-% [X] 1 PDF5+ (https://www.icdd.com/,https://www.icdd.com/pdf-5/) -> 19000 structures with pXRD patterns | I checked with their support
-    % (1.1): ICDD Powders
-    % (1.2): FIZ ICSD (https://icsd.products.fiz-karlsruhe.de/)
-    % (1.3): NIST ICSD https://icsd.nist.gov/
-    % (1.4): Linus Pauling File (https://paulingfile.com/index.php?p=about%20us)
-    % (1.5): ICDD Single Crystal data
-% [x]: 2 COD (https://www.crystallography.net/cod/) -> ~ 1000 structures with real pXRD patterns | These patterns were extracted by PSL university
-% [x]: 3 RRUFF (American Mineralogist Crystal Structure Database) -> ~ 1000 structures with real pXRD patterns | We downloaded the entire dataset and discarded single crystal data
-% [x]: 4 CCDC (https://www.ccdc.cam.ac.uk/) -> No exp pXRD | There is only mention of simulated powder diffraction data. Of 20 sampled cifs none contained powder diffraction data
-% [x]: 5 The Material Project (https://legacy.materialsproject.org/) -> No exp pXRD | The only mention of available X-ray diffraction data is of simulated data
-% [x]: 6 Crystal Lattice Structures (https://www.atomic-scale-physics.de/lattice/) -> No exp pXRD | Their entries provide no X-ray diffraction data, and also most links appear broken
-% [x]: 7 Crystallographic and Crystallochemical Database for Minerals and their Structural Analogues (https://database.iem.ac.ru/mincryst/) -> No exp pXRD | The given powder diffraction data is simulated (see: Powder X-ray Diffraction Standards (CPDS) @ https://database.iem.ac.ru/mincryst/descript.htm)
-% [x]: 8 Mineralogy Database (https://webmineral.com/) -> No exp pXRD | I checked with the author of the site
-% [x]: 9 IUCr Raw data letters (https://iucrdata.iucr.org/) -> No exp pXRD | There are only two raw data letters, both of them single crystal data
-% [x]: 10 Bilbao Crystallographic server  (https://www.cryst.ehu.eus/bincstrdb/search/) No exp pXRD | Of 20 sampled CIFs out of 256 entries none contained pXRD data
-% [x]: 11 PowBase (http://www.cristal.org/powbase/index.html) -> 169 real pXRD patterns some of which have at least partial labels; The described submission process and the lack of crystal structures indicate that the patterns are not simulated
-% [x]: 12 Athena (https://athena.unige.ch/athena/mineral/mineral.html) -> No exp pXRD | No mention of diffraction data anywhere
-% [x]: 13 Protein database (https://www.rcsb.org/) / NAKB (https://www.nakb.org/) / BMCD ((http://bmcd.ibbr.umd.edu/)  -> Little to no exp pXRD | Of the 230k structures only 21 showed up when searching experimental method = Powder diffraction
-% [x]: 14 Crystal Met (https://cds.dl.ac.uk/cds/datasets/crys/mdf/llmdf.html) -> No exp pXRD | The only mention of diffraction data is simulated data
-% [x]: 15 Pearson's Crystal data (https://www.crystalimpact.com/pcd/) -> 21700 experimental powder diffraction patterns (As per https://www.crystalimpact.com/pcd/) | The Pearson's crystal data evolved from the Linus Pauling File which is a data source of the PDF, so there may be significant overlap in these patterns
-% [x]: 16 Zenodo powder diffraction data (https://zenodo.org/search?q=powder%20diffraction&l=list&p=1&s=10&sort=bestmatch) -> Some pXRD data but hard to quantify, probably <= 4000 patterns most of which are unlabeled | 1600 datasets matching "powder diffraction".
-From probing among the first 100 matches I estimate only about 1 in 10 contain powder diffraction patterns.
-The probed datasets that did contain powder diffraction data only contained <= 30 patterns and only one of them came with corresponding crystal structures
-
 \pagebreak
@@
 \subsubsection*{Existing experimental powder diffraction databases}
-To contextualize opXRD within the current environment of experimental powder diffraction data, the following list provides an overview of the largest crystal structure databases that include powder diffraction data.
+To contextualize opXRD within the current environment of experimental powder diffraction data, the following list provides an overview of the largest crystal structure databases that offer access to experimental powder diffraction data.
 \\

### 2024-11-25 chapters/03_SOTA_ML.tex
Warpped chemical formulas in \ce{} environemnt

@@
 %TODO: Papers to include:
-% -  https://pubs.acs.org/doi/10.1021/jacs.4c10244
-% - Sagado et al.: Automated classification of big X-ray diffraction data using deep learning models
+% -  Crystal Structure Determination from Powder Diffraction Patterns with Generative Machine Learning: (https://pubs.acs.org/doi/10.1021/jacs.4c10244)
 
@@
 Park {\it et al.} reached a test accuracy of roughly 81\% for a CNN, which classifies space groups of simulated single-phase diffractograms \cite{Park2017}.
+\\
+
 A regression analysis on lattice parameters within a broader framework encompassing all material classes was conducted by Chitturi et al\cite{Chitturi2021}.
@@
 The study observed a noticeable performance drop when the pre-trained model was applied to real experimental patterns, as opposed to simulated data.
-However, their recent work \cite{cao2024simxrd} proposes using convolutional peak descriptors that consider the detector's geometry, which significantly reduces the performance gap in their benchmark tests.
+However, their recent work \cite{cao2024simxrd} proposes using convolutional peak descriptors that consider the detector's geometry, which significantly reduces the performance gap in their benchmark tests.\\
+
 The option of training purely on experimental diffractograms unfortunately is unfeasible because of the limited availability of labeled experimental diffractograms, but Salgado {\it et al.} \cite{Salgado2023} showed, that adding a fraction of experimental patterns to a simulated training dataset improves the performance on unseen experimental patterns.
@@
 Then they tested their model's performance on the other half of the RRUFF database and achieved a performance increase in the 230-way spacegroup classification accuracy of \num{11} percentage points compared to the same model only trained on simulated patterns.
+\\
 

### 2024-11-25 chapters/04_our_dataset.tex
Warpped chemical formulas in \ce{} environemnt

@@
 The following texts detail the contribution of each research group.
-Most data was collected with copper sources which have a  $K_{\alpha1}$ wavelength of $\lambda=1.54056\text{\AA}$ and a $K_{\alpha2}$ wavelength of $\lambda=1.54439\text{\AA}$.
+Most data was collected using \ce{Cu} radiaton sources which have a  $K_{\alpha1}$ wavelength of $\lambda=1.54056\text{\AA}$ and a $K_{\alpha2}$ wavelength of $\lambda=1.54439\text{\AA}$.
 
@@
 
-The X-ray diffraction data were collected on a Bruker D8 Advance using a Cu radiation source  The samples were initially recorded for various research projects over the last ten years and were measured with different step sizes, times per step, and over different angle ranges, but all using Cu K$\alpha$1 radiation.
+The X-ray diffraction data were collected on a Bruker D8 Advance using a Cu radiation source  The samples were initially recorded for various research projects over the last ten years and were measured with different step sizes, times per step, and over different angle ranges, but all using \ce{Cu} $K_\alpha$ radiation.
 The samples mostly contained transition metal oxides, sulfides, and fluorides.
@@
 The latter were synthesized via the surfactant-free benzyl alcohol route \cite{Wolf2019, Wolf2018}.
-The cobalt oxide (CoO or Co3O4) and cerium oxide (CeO2) nanoparticles were in the size range of $4-16 \ \si{nm}$ according to the Scherrer equation.
-A series of porous Al2O3 materials, which were prepared by calcination of boehmite (AlOOH) at various temperatures, represents crystalline samples with limited long-range structure and various contributions of Al2O3 polymorphs.\\
+The cobalt oxide (\ce{CoO} or \ce{Co3O4}) and cerium oxide (\ce{CeO2}) nanoparticles were in the size range of $4-16 \ \si{nm}$ according to the Scherrer equation.
+A series of porous \ce{Al2O3} materials, which were prepared by calcination of boehmite (\ce{AlOOH}) at various temperatures, represents crystalline samples with limited long-range structure and various contributions of \ce{Al2O3} polymorphs.\\
 
-X-ray diffraction (XRD) was conducted with an X’Pert Pro MPD (Panalytical) in Bragg-Brentano geometry using a Cu X-ray source.
+X-ray diffraction (XRD) was conducted with an X’Pert Pro MPD (Panalytical) in Bragg-Brentano geometry using a \ce{Cu} X-ray source.
 The patterns were acquired in the $2\theta$ range of $5-80\degree$ with a step size of $0.016711\degree$ or $0.033420\degree$ and a total acquisition time of 40 to 120 min.
@@
 
-Combinatorial CuNi and CuAl samples were synthesized at the University of Southern California (USC) via magnetron co-sputtering from Cu (99.999\%) and Ni (99.995\%) or Al (99.999\%) targets (Plasmaterials) onto stationary $10 \si{cm}$ Si (100) substrates yielding 169 unique $5 \times 5 \ \si{mm}$ squares per wafer.
+Combinatorial \ce{CuNi} and \ce{CuAl} samples were synthesized at the University of Southern California (USC) via magnetron co-sputtering from \ce{Cu} (99.999\%) and \ce{Ni} (99.995\%) or \ce{Al} (99.999\%) targets (Plasmaterials) onto stationary $10 \si{cm}$ \ce{Si} (100) substrates yielding 169 unique $5 \times 5 \ \si{mm}$ squares per wafer.
 These alloys were selected to study compositional effects on stacking fault and growth twin boundary formation in sputtered materials \cite{2024AcMat.27019839A,alwen2024combinatorial}.
@@
 XRD analysis was used to link changes in defect formation in the sputtered films with varied phases and crystal structures.
-Each composition was characterized using CuK X-rays that were collimated to probe an area of $4 \times 4 \ \si{mm}$ using a step size of $0.026\degree$ and scan duration of $0.3$ seconds per step over a 2 range from $30\degree - 110\degree$.
+Each composition was characterized using \ce{Cu} $K_\alpha$ X-rays that were collimated to probe an area of $4 \times 4 \ \si{mm}$ using a step size of $0.026\degree$ and scan duration of $0.3$ seconds per step over a 2 range from $30\degree - 110\degree$.
 Measurement parameters were chosen to resolve at least six or more points above the full-width half maximum of each XRD peak.
@@
 Data collection was performed in situ during thin-film deposition using a custom-made spin-coating and annealing stage.
-One dataset was collected from spin-coating triple cation metal-halide perovskite precursor solutions with the composition Cs$_{0.05}$(MA$_{0.23}$FA$_{0.77}$)Pb$_{1.1}$(I$_{0.77}$Br$_{0.23}$)$_{3}$ (MA = Methylammonium, FA = Formamidinium) onto various substrates, including glass (amorphous), GaAs wafers (single crystalline) as well as stacks of glass/ITO, GaAs/Mo/Cu(In,Ga)Se$_{2}$/CdS/ZnO, and glass/Mo/Cu(In,Ga)Se$_{2}$/CdS/ZnO (glass/CIGS).
+One dataset was collected from spin-coating triple cation metal-halide perovskite precursor solutions with the composition
+\ce{Cs}$_{0.05}$(MA$_{0.23}$FA$_{0.77}$)\ce{Pb}$_{1.1}$(\ce{I}$_{0.77}$\ce{Br}$_{0.23}$)$_{3}$ (MA = Methylammonium, FA = Formamidinium) onto various substrates, including glass (amorphous), $\ce{GaAs}$ wafers (single crystalline) as well as stacks of glass/ITO, $\ce{GaAs}$/$\ce{Mo}$/$\ce{Cu}$($\ce{In}$,$\ce{Ga}$)$\ce{Se2}$/$\ce{CdS}$/$\ce{ZnO}$, and glass/$\ce{Mo}$/$\ce{Cu}$($\ce{In}$,$\ce{Ga}$)$\ce{Se2}$/$\ce{CdS}$/$\ce{ZnO}$ (glass/CIGS).
 Some of the substrates were additionally covered with a self-assembling monolayer of MeO-2PACz.
-The GaAs substrates were prepared by Dr.
+The $\ce{GaAs}$ substrates were prepared by Dr.
 Jiro Nishinaga from the National Institute of Advanced Industrial Science and Technology (AIST) in Japan and the glass/CIGS substrates by Dr.
@@
 Each in situ measurement consisted of about 500 to 1000 individual diffractograms.
-Depending on the substrate, each series of diffractograms shows an evolution from substrate only to a combination of polycrystalline perovskite, PbI$_{2}$ and substrate via several intermediate phases.
+Depending on the substrate, each series of diffractograms shows an evolution from substrate only to a combination of polycrystalline perovskite, $\ce{PbI2}$ and substrate via several intermediate phases.
+\\
 
 Experimental XRD data were collected at beamline 12.3.2 of the Advanced Light Source, the synchrotron at Lawrence Berkeley National Laboratory.
-The data were collected using a photon energy of 10 keV ($\lambda = 1.23984 \text{\AA}$), selected using a Si(111) monochromator.
+The data were collected using a photon energy of 10 $\si{keV}$ ($\lambda = 1.23984 \text{\AA}$), selected using a Si(111) monochromator.
 Measurements were taken in grazing incidence geometry, i.e.
@@
 In the past two years, we established a small-scale experimental pXRD database called the X-Ray Phase Identification Public Experimental Dataset (XRed).
-\footnote{XRed can be found on GitHub at \url{https://github.com/WPEM/XRED}.} The data in XRed is generated in our lab using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers using copper X-ray sources.
+\footnote{XRed can be found on GitHub at \url{https://github.com/WPEM/XRED}.} The data in XRed is generated in our lab using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers using \ce{Cu} X-ray sources.
 Hundreds of pXRD patterns have been refined, labeled with CIF files, and organized by elemental systems.
@@
 
-Each combinatorial Zn–V–N library was synthesized using radio-frequency co-sputtering of Zn and V in a mixed Ar and N\textsubscript{2} plasma.
+Each combinatorial Zn–V–N library was synthesized using radio-frequency co-sputtering of Zn and V in a mixed Ar and \ce{N2} plasma.
 An orthogonal deposition temperature and composition gradient was created, resulting in a deposition temperature of $220\degree$C for samples 1~–~9 and $114\degree$C for samples 37~–~45.
 The composition for each sample was determined using X-ray fluorescence (XRF) spectroscopy which was further calibrated through Rutherford backscattering spectroscopy (RBS) based on selected samples.
-The newly identified and isolated semiconductor Zn\textsubscript{2}VN\textsubscript{3} was identified to exhibit a cation-disordered wurtzite structure as verified by additional GI-XRD and SAED measurements\cite{Zhuk2021}.
+The newly identified and isolated semiconductor \ce{Zn2VN3} was identified to exhibit a cation-disordered wurtzite structure as verified by additional GI-XRD and SAED measurements\cite{Zhuk2021}.
 \\
 Tin halide perovskites were deposited using single-step spin-coating as reported elsewhere\cite{Wieczorek2023}.
-Methylammonium lead iodide libraries with varying degrees of residual PbI\textsubscript{2} were deposited using a two-step procedure involving both thermal evaporation of PbI\textsubscript{2} and subsequent spin-coating of a methylammonium solution.
+Methylammonium lead iodide libraries with varying degrees of residual \ce{PbI2} were deposited using a two-step procedure involving both thermal evaporation of \ce{PbI2} and subsequent spin-coating of a methylammonium solution.
 The relative phase fractions were quantified using supplementary azimuthal angle scans coupled with structural factors and geometrical factors as reported elsewhere\cite{Wieczorek2024}.

### 2024-11-26 chapters/04_our_dataset.tex
Merged changes by Ben Breitung

@@
 Most of those multi-component complex materials appeared in various structures, including rock-salt, spinel, fluorite, perovskite, and delafossite.
-The samples were almost always prepared in powder form; therefore, powder XRD was performed on samples with adjusted height.
+The samples were prepared either in powder or in bulk form; therefore, powder XRD was performed on samples with adjusted height.
 The samples were prepared using various synthesis techniques, mostly solid-state or wet chemical syntheses, to obtain the desired structures.
@@
 
-The X-ray diffraction data were collected on a Bruker D8 Advance using a Cu radiation source  The samples were initially recorded for various research projects over the last ten years and were measured with different step sizes, times per step, and over different angle ranges, but all using \ce{Cu} $K_\alpha$ radiation.
+The X-ray diffraction data were collected on a Bruker D8 Advance using a Cu radiation source or a STOE Stadi P diffractometer equipped with a Ga-jet X-ray source.
+The samples were initially recorded for various research projects over the last ten years and were measured with different step sizes, times per step, and over different angle ranges, but all using \ce{Cu} $K_\alpha$ or \ce{Ga} $K_\beta$ radiation.
 The samples mostly contained transition metal oxides, sulfides, and fluorides.

### 2024-11-26 chapters/02_existing_datasets.tex
Added links to other crystal structure resources.

@@
 Additionally, extracting powder diffraction data at scale is hindered by the fact that the data is often given in plain text files in non-standardized formats, which are difficult to parse programmatically.
-Some major crystal structure databases, such as the ICSD, do not appear in the above list, as they do not include powder diffraction data.
-This is because most structure solutions are achieved through single-crystal diffraction rather than powder diffraction.
+\\
+
+Aside from the databases mentioned above, we have also investigated several other crystal structure resources in search of experimental powder diffraction data.
+Crystal structure resources that were investigated but not found to contain any appreciable amount of publicly available experimental powder diffraction data include the Inorganic Crystal Structure Database,\footnote{\url{https://icsd.products.fiz-karlsruhe.de/}}   Cambridge Structural Database,\footnote{\url{https://www.ccdc.cam.ac.uk/structures/}}  the Materials Project database,\footnote{\url{https://next-gen.materialsproject.org/}}  the Crystallographic and Crystallochemical Database\footnote{\url{https://database.iem.ac.ru/mincryst/index.php}},  the Bilbao Incommensurate Crystal Structure Database,\footnote{\url{https://www.cryst.ehu.eus/bincstrdb/search/}}  the Mineralogy Database,\footnote{\url{https://webmineral.com/}} IUCr Raw data letters,\footnote{\url{https://iucrdata.iucr.org/x/index.html}} and U.S.
+Naval Research Laboratory Crystal Lattice-Structures,\footnote{\url{https://www.atomic-scale-physics.de/lattice/}} the Athena Mineral database \footnote{\url{https://athena.unige.ch/athena/mineral/mineral.html}} and the Protein data bank \footnote{\url{https://www.rcsb.org/}}.
+This is to be expected as most structure solutions are achieved through single-crystal diffraction rather than powder diffraction.\\

### 2024-11-26 chapters/02_existing_datasets.tex
Added more information can be found to footnotes

--- 2024-11-25 chapters/02_existing_datasets.tex
@@
 Additionally, extracting powder diffraction data at scale is hindered by the fact that the data is often given in plain text files in non-standardized formats, which are difficult to parse programmatically.
-Some major crystal structure databases, such as the ICSD, do not appear in the above list, as they do not include powder diffraction data.
-This is because most structure solutions are achieved through single-crystal diffraction rather than powder diffraction.
+\\
+
+Aside from the databases mentioned above, we have also investigated several other crystal structure resources in search of experimental powder diffraction data.
+Crystal structure resources that were investigated but not found to contain any appreciable amount of publicly available experimental powder diffraction data include the Inorganic Crystal Structure Database,\footnote{More information on the ICSD can be found under \url{https://icsd.products.fiz-karlsruhe.de/}.}   Cambridge Structural Database,\footnote{More information on the CCDC can be found under \url{https://www.ccdc.cam.ac.uk/structures/}.}  the Materials Project database,\footnote{More information on the Materials Project can be found under\url{https://next-gen.materialsproject.org/}.}  the Crystallographic and Crystallochemical Database\footnote{More information on the Crystallographic and Crystallochemical Database can be found under \url{https://database.iem.ac.ru/mincryst/index.php}.},  the Bilbao Incommensurate Crystal Structure Database,\footnote{More information on the Incommensurate Crystal Structure Database can be found under \url{https://www.cryst.ehu.eus/bincstrdb/search/}.}  the Mineralogy Database,\footnote{More information on the Mineralogy database can be found under \url{https://webmineral.com/}.} IUCr Raw data letters,\footnote{More information on the IUCr Raw Data letters can be found under \url{https://iucrdata.iucr.org/x/index.html}.} and U.S.
+Naval Research Laboratory Crystal Lattice-Structures,\footnote{More information on the Crystal Lattice-Structures database can be found under \url{https://www.atomic-scale-physics.de/lattice/}.} the Athena Mineral database \footnote{More information on the Athena Mineral database can be found under \url{https://athena.unige.ch/athena/mineral/mineral.html}.} and the Protein data bank.
+\footnote{More information on the Protein data bank can be found under \url{https://www.rcsb.org/}.} This is to be expected as most structure solutions are achieved through single-crystal diffraction rather than powder diffraction.\\

### 2024-11-26 chapters/02_existing_datasets.tex
Fixed decimal seperator in existing datasets

@@
 To contextualize opXRD within the current environment of experimental powder diffraction data, the following list provides an overview of the largest crystal structure databases that offer access to experimental powder diffraction data.
-\\
+For an overview of these databases refer to Tab.$\eqref{tab:exp_databases}$ below.\\
+
+\begin{table}[!htb]
+\centering
+\caption{Overview of Experimental Powder Diffraction Databases}
+\label{tab:exp_databases}
+\begin{tabular}{@{}lcc@{}}
+\toprule
+\textbf{Name}                          & \textbf{No. of Patterns} & \textbf{Year Established} \\
+\midrule
+Powder Diffraction File (PDF)          & 19,000                  & 1941                      \\
+Pearson's Crystal Data                 & 21,700                  & [Year Established]        \\
+Crystallography Open Database (COD)    & 1063                    & 2003                      \\
+RRUFF                                  & 1290                    & 2006                      \\
+PowBase                                & 169                     & 1999                      \\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 However, most of these powder diffraction patterns are simulated.
-After inquiring with the ICDD in April 2024 we were told that only 19000 of the powder diffraction patterns in the PDF5+ stemmed from experiments.
-Of these 19000 entries, 10954 contain information about the atomic coordinates of the underlying structures.
+After inquiring with the ICDD in April 2024 we were told that only 19,000 of the powder diffraction patterns in the PDF5+ stemmed from experiments.
+Of these 19,000 entries, 10,954 contain information about the atomic coordinates of the underlying structures.
 \footnote{We reached out to the contact address given on their website, info@icdd.com.} Still, this makes the PDF5+, to the best of our knowledge, the largest collection of experimental powder diffraction data available to researchers.
-As of November 2024, the PDF5+ is available to researchers through a purchase of a one-year license starting at a price point of \$6,265.00.\\
+As of November 2024, the PDF5+ is available to researchers through a purchase of a one-year license starting at a price point of \$6265.00.\\
 
-\textbf{Pearson's Crystal data}:\footnote{More information on Pearson's crystal data can be found on the ASM International website under \url{https://www.asminternational.org/materials-resources/online-databases/pearsons-crystal-data-crystal-structure-database-for-inorganic-compounds/?srsltid=AfmBOorjpwJSVNsaWtCO4kmgu-Ox2ncF3tkzwCXRE2GTemYIxsSLp8xr}.} Pearson's crystal data is a commercial database of 395000 crystal structures including 21700 experimental powder diffraction patterns maintained by ASM International.
+\textbf{Pearson's Crystal data}:\footnote{More information on Pearson's crystal data can be found on the ASM International website under \url{https://www.asminternational.org/materials-resources/online-databases/pearsons-crystal-data-crystal-structure-database-for-inorganic-compounds/?srsltid=AfmBOorjpwJSVNsaWtCO4kmgu-Ox2ncF3tkzwCXRE2GTemYIxsSLp8xr}.} Pearson's crystal data is a commercial database of 395,000 crystal structures including 21,700 experimental powder diffraction patterns maintained by ASM International.
 Pearson's crystal data is based on the crystallographic data of another database, the Linus Pauling file\cite{kaduk2007}.
@@
 \textbf{Crystallography Open Database:}\footnote{More information on the crystallography open database can be found under \url{https://www.crystallography.net/cod/}.} The Crystallography Open Database (COD) is an open-access collection of crystal structures founded in 2003\cite{Graulis2009cod}.
-It currently provides over 500.000 crystal structures in the form of .cif files.
+It currently provides over 500,000 crystal structures in the form of .cif files.
 Of these files, 1063 contain the experimental powder diffraction data that was used to determine the underlying crystal structures of the investigated samples.

### 2024-11-26 chapters/02_existing_datasets.tex
Added year established of pearsons crystal data

@@
 Powder Diffraction File (PDF)          & 19,000                  & 1941                      \\
-Pearson's Crystal Data                 & 21,700                  & [Year Established]        \\
+Pearson's Crystal Data                 & 21,700                  & 2007        \\
 Crystallography Open Database (COD)    & 1063                    & 2003                      \\
@@
 \textbf{Powder Diffraction File:}\footnote{More information on the Powder Diffraction File can be found under \url{https://www.icdd.com/pdf-5/}.} The Powder Diffraction File (PDF), published and maintained by the International Center for Diffraction Data (ICDD), is a large collection of materials with accompanying powder diffraction data first published in 1941\cite{GatesRector2019}.
-Its latest release as of November 2024, the PDF5+, contains over a million materials with accompanying powder diffraction data.
-However, most of these powder diffraction patterns are simulated.
-After inquiring with the ICDD in April 2024 we were told that only 19,000 of the powder diffraction patterns in the PDF5+ stemmed from experiments.
-Of these 19,000 entries, 10,954 contain information about the atomic coordinates of the underlying structures.
-\footnote{We reached out to the contact address given on their website, info@icdd.com.} Still, this makes the PDF5+, to the best of our knowledge, the largest collection of experimental powder diffraction data available to researchers.
+According to the ICDD the PDF5+, the latest release of the PDF as of November 2024, contains over a million materials with accompanying powder diffraction data.
+However, most of these powder diffraction patterns are simulated.\footnote{This information is available under \url{https://www.icdd.com/pdf-5/}.} After inquiring with the ICDD in April 2024 we were told that only 19,000 of the powder diffraction patterns in the PDF5+ stemmed from experiments.
+Of these 19,000 entries, 10,954 contain information about the atomic coordinates of the underlying structures.\footnote{We reached out to the contact address given on their website, info@icdd.com.} Still, this makes the PDF5+, to the best of our knowledge, the largest collection of experimental powder diffraction data available to researchers.
 As of November 2024, the PDF5+ is available to researchers through a purchase of a one-year license starting at a price point of \$6265.00.\\
 
-\textbf{Pearson's Crystal data}:\footnote{More information on Pearson's crystal data can be found on the ASM International website under \url{https://www.asminternational.org/materials-resources/online-databases/pearsons-crystal-data-crystal-structure-database-for-inorganic-compounds/?srsltid=AfmBOorjpwJSVNsaWtCO4kmgu-Ox2ncF3tkzwCXRE2GTemYIxsSLp8xr}.} Pearson's crystal data is a commercial database of 395,000 crystal structures including 21,700 experimental powder diffraction patterns maintained by ASM International.
-Pearson's crystal data is based on the crystallographic data of another database, the Linus Pauling file\cite{kaduk2007}.
+\textbf{Pearson's Crystal data}:\footnote{More information on Pearson's crystal data can be found on the ASM International website under \url{https://www.asminternational.org/materials-resources/online-databases/pearsons-crystal-data-crystal-structure-database-for-inorganic-compounds/?srsltid=AfmBOorjpwJSVNsaWtCO4kmgu-Ox2ncF3tkzwCXRE2GTemYIxsSLp8xr}.} Pearson's crystal data is a commercial database of 395,000 crystal structures including 21,700 experimental powder diffraction patterns maintained by ASM International first published in 2003.\footnote{According to CrystalImpact, who acts as distributor for Pearson's crystal data.
+For more information refer to the CrystalImpact website under \url{https://web.archive.org/web/20241112132017/https://www.crystalimpact.com/news/Default.htm}.
+} Pearson's crystal data is based on the data of another crystal structure database, the Linus Pauling file \cite{kaduk2007}.
 Since the Linus Pauling file is also a database source for the PDF, there is likely a significant overlap between the experimental powder diffraction data found in Pearson's crystal data and that found in the PDF.
@@
 
-There is also much publicly available powder diffraction data uploaded to datasets on Zenodo.
+There is also publicly available powder diffraction data uploaded to datasets on Zenodo\footnote{More information on Zenodo can be found under \url{https://zenodo.org/}.}.
 However, this data is split into disparate entries that typically only contain the work of a single research project.

### 2024-11-26 chapters/02_existing_datasets.tex
Commented out statement that pearson and PDF have huge overlap

@@
 For more information refer to the CrystalImpact website under \url{https://web.archive.org/web/20241112132017/https://www.crystalimpact.com/news/Default.htm}.
-} Pearson's crystal data is based on the data of another crystal structure database, the Linus Pauling file \cite{kaduk2007}.
-Since the Linus Pauling file is also a database source for the PDF, there is likely a significant overlap between the experimental powder diffraction data found in Pearson's crystal data and that found in the PDF.
-As of November 2024, Pearson's crystal data is available to researchers through a purchase of a one-year license starting at a price point of 2200 \euro.\\
+}As of November 2024, Pearson's crystal data is available to researchers through a purchase of a one-year license starting at a price point of 2200 \euro.\\
+
+% TODO: I need to check that by asking
+%Pearson's crystal data is provided by the Pauling File project which also provides the Materials Phases Data System database (MPDS). 
+%Since the MPDS is listed as a database source for the PDF5+ \cite{kaduk2007}, there is likely a significant overlap between the experimental powder diffraction data found in Pearson's crystal data and that found in the PDF.
+ %\tablefootnote{The PDF and Pearson's Crystal data both derive some of their data from the Linus Pauling File, which has a similar number of experimental powder diffraction patterns as both the PDF and Pearson's Crystal Data.
+Hence there is likely significant overlap in the patterns available in the PDF and in Pearson's crystal data.}
 

### 2024-11-27 chapters/02_existing_datasets.tex
Removed unneccessary footnotes in existing datasets; Converted remaining footnotes to web citations

@@
 
-\textbf{Powder Diffraction File:}\footnote{More information on the Powder Diffraction File can be found under \url{https://www.icdd.com/pdf-5/}.} The Powder Diffraction File (PDF), published and maintained by the International Center for Diffraction Data (ICDD), is a large collection of materials with accompanying powder diffraction data first published in 1941\cite{GatesRector2019}.
+\textbf{Powder Diffraction File:} \cite{PDFWeb} The Powder Diffraction File (PDF), published and maintained by the International Center for Diffraction Data (ICDD), is a large collection of materials with accompanying powder diffraction data first published in 1941\cite{GatesRector2019}.
 According to the ICDD the PDF5+, the latest release of the PDF as of November 2024, contains over a million materials with accompanying powder diffraction data.
-However, most of these powder diffraction patterns are simulated.\footnote{This information is available under \url{https://www.icdd.com/pdf-5/}.} After inquiring with the ICDD in April 2024 we were told that only 19,000 of the powder diffraction patterns in the PDF5+ stemmed from experiments.
+However, most of these powder diffraction patterns are simulated.
+After inquiring with the ICDD in April 2024 we were told that only 19,000 of the powder diffraction patterns in the PDF5+ stemmed from experiments.
 Of these 19,000 entries, 10,954 contain information about the atomic coordinates of the underlying structures.\footnote{We reached out to the contact address given on their website, info@icdd.com.} Still, this makes the PDF5+, to the best of our knowledge, the largest collection of experimental powder diffraction data available to researchers.
@@
 
-\textbf{Pearson's Crystal data}:\footnote{More information on Pearson's crystal data can be found on the ASM International website under \url{https://www.asminternational.org/materials-resources/online-databases/pearsons-crystal-data-crystal-structure-database-for-inorganic-compounds/?srsltid=AfmBOorjpwJSVNsaWtCO4kmgu-Ox2ncF3tkzwCXRE2GTemYIxsSLp8xr}.} Pearson's crystal data is a commercial database of 395,000 crystal structures including 21,700 experimental powder diffraction patterns maintained by ASM International first published in 2003.\footnote{According to CrystalImpact, who acts as distributor for Pearson's crystal data.
-For more information refer to the CrystalImpact website under \url{https://web.archive.org/web/20241112132017/https://www.crystalimpact.com/news/Default.htm}.
-}As of November 2024, Pearson's crystal data is available to researchers through a purchase of a one-year license starting at a price point of 2200 \euro.\\
+\textbf{Pearson's Crystal data}:\cite{PearsonWeb} Pearson's crystal data is a commercial database of 395,000 crystal structures including 21,700 experimental powder diffraction patterns maintained by ASM International first published in 2007\cite{PaulingWeb}.
+As of November 2024, Pearson's crystal data is available to researchers through a purchase of a one-year license starting at a price point of 2200 \euro.\\
 
@@
 
-\textbf{Crystallography Open Database:}\footnote{More information on the crystallography open database can be found under \url{https://www.crystallography.net/cod/}.} The Crystallography Open Database (COD) is an open-access collection of crystal structures founded in 2003\cite{Graulis2009cod}.
+\textbf{Crystallography Open Database:} \cite{CODWeb} The Crystallography Open Database (COD) is an open-access collection of crystal structures founded in 2003\cite{Graulis2009cod}.
 It currently provides over 500,000 crystal structures in the form of .cif files.
@@
 
-\textbf{RRUFF}:\footnote{More information on RRUFF can be found under \url{https://rruff.info/about/about_general.php}.} The RRUFF Mineral Database, first published in 2006, provides detailed information on minerals, including their chemical compositions, crystallography, and spectroscopic data.
-Managed by the University of Arizona, it was created to serve as a public repository for mineral identification and research.
+\textbf{RRUFF}: \cite{RRUFFWeb} The RRUFF Mineral Database, first published in 2006, provides detailed information on minerals, including their chemical compositions, crystallography, and spectroscopic data.
+\cite{lafuente2015} Managed by the University of Arizona, it was created to serve as a public repository for mineral identification and research.
 It contains \num{1290} powder diffraction patterns stemming from experiments each labeled with the lattice parameters and composition of the underlying structures.
@@
 
-\textbf{PowBase}:\footnote{More information on PowBase can be found under \url{http://www.cristal.org/powbase/index.html}. PowBase was an initiative suggested in the Structure Determination by
-Powder Diffractometry (SDPD) mailing list which was hosted on the same site.
+\textbf{PowBase}:\footnote{More information on PowBase can be found under \url{http://www.cristal.org/powbase/index.html}.
+PowBase was an initiative suggested in the Structure Determination by Powder Diffractometry (SDPD) mailing list which was hosted on the same site.
 The COD was another community initiative that grew out of this mailing list.} PowBase is a database of 169 mostly unlabeled experimental powder diffraction patterns collected and maintained by crystallography researcher Armel Le Bail starting in 1999.
@@
 
-There is also publicly available powder diffraction data uploaded to datasets on Zenodo\footnote{More information on Zenodo can be found under \url{https://zenodo.org/}.}.
+There is also publicly available powder diffraction data uploaded to datasets on Zenodo.
 However, this data is split into disparate entries that typically only contain the work of a single research project.
@@
 Aside from the databases mentioned above, we have also investigated several other crystal structure resources in search of experimental powder diffraction data.
-Crystal structure resources that were investigated but not found to contain any appreciable amount of publicly available experimental powder diffraction data include the Inorganic Crystal Structure Database,\footnote{More information on the ICSD can be found under \url{https://icsd.products.fiz-karlsruhe.de/}.}   Cambridge Structural Database,\footnote{More information on the CCDC can be found under \url{https://www.ccdc.cam.ac.uk/structures/}.}  the Materials Project database,\footnote{More information on the Materials Project can be found under\url{https://next-gen.materialsproject.org/}.}  the Crystallographic and Crystallochemical Database\footnote{More information on the Crystallographic and Crystallochemical Database can be found under \url{https://database.iem.ac.ru/mincryst/index.php}.},  the Bilbao Incommensurate Crystal Structure Database,\footnote{More information on the Incommensurate Crystal Structure Database can be found under \url{https://www.cryst.ehu.eus/bincstrdb/search/}.}  the Mineralogy Database,\footnote{More information on the Mineralogy database can be found under \url{https://webmineral.com/}.} IUCr Raw data letters,\footnote{More information on the IUCr Raw Data letters can be found under \url{https://iucrdata.iucr.org/x/index.html}.} and U.S.
-Naval Research Laboratory Crystal Lattice-Structures,\footnote{More information on the Crystal Lattice-Structures database can be found under \url{https://www.atomic-scale-physics.de/lattice/}.} the Athena Mineral database \footnote{More information on the Athena Mineral database can be found under \url{https://athena.unige.ch/athena/mineral/mineral.html}.} and the Protein data bank.
-\footnote{More information on the Protein data bank can be found under \url{https://www.rcsb.org/}.} This is to be expected as most structure solutions are achieved through single-crystal diffraction rather than powder diffraction.\\
+Crystal structure resources that were investigated but not found to contain any appreciable amount of publicly available experimental powder diffraction data include the Inorganic Crystal Structure Database,   Cambridge Structural Database,  the Materials Project database,  the Crystallographic and Crystallochemical Database, the Bilbao Incommensurate Crystal Structure Database,  the Mineralogy Database, IUCr Raw data letters, the U.S.
+Naval Research Laboratory Crystal Lattice-Structures, the Athena Mineral database and the Protein data bank.
+This is to be expected as most structure solutions are achieved through single-crystal diffraction rather than powder diffraction.\\

### 2024-11-28 chapters/02_existing_datasets.tex
Described overlap in experimental pXRD data in PDF5+ and Pearson's Crystal data

@@
 \midrule
-Powder Diffraction File (PDF)          & 19,000                  & 1941                      \\
+Powder Diffraction File (PDF)  \tablefootnote{The PDF lists the Materials Phases Data System (MPDS) as a database source.
+The MPDS and Pearson's crystal, both maintained by the Pauling File project, contain largely identical data.
+Hence it is likely that there is significant overlap in the experimental patterns available in the PDF and Pearson's crystal data.}         & 20,800                  & 1941                      \\
 Pearson's Crystal Data                 & 21,700                  & 2007        \\
@@
 However, most of these powder diffraction patterns are simulated.
-After inquiring with the ICDD in April 2024 we were told that only 19,000 of the powder diffraction patterns in the PDF5+ stemmed from experiments.
-Of these 19,000 entries, 10,954 contain information about the atomic coordinates of the underlying structures.\footnote{We reached out to the contact address given on their website, info@icdd.com.} Still, this makes the PDF5+, to the best of our knowledge, the largest collection of experimental powder diffraction data available to researchers.
+After inquiring with the ICDD in April 2024 we were told that only 20,800 of the powder diffraction patterns in the PDF5+ stemmed from experiments.
+Of these 20,800 entries, 10,954 contain information about the atomic coordinates of the underlying structures.\footnote{We reached out to the contact address given on their website, info@icdd.com.} Still, this makes the PDF5+, to the best of our knowledge, the largest collection of experimental powder diffraction data available to researchers.
 As of November 2024, the PDF5+ is available to researchers through a purchase of a one-year license starting at a price point of \$6265.00.\\
 
-\textbf{Pearson's Crystal data}:\cite{PearsonWeb} Pearson's crystal data is a commercial database of 395,000 crystal structures including 21,700 experimental powder diffraction patterns maintained by ASM International first published in 2007\cite{PaulingWeb}.
+\textbf{Pearson's Crystal data}:\cite{PearsonWeb} Pearson's crystal data is a commercial database published and maintained by the Pauling File project \cite{kaduk2007}.
+The database, first published in 2007, currently contains 395,000 crystal structures and 21,700 corresponding experimental powder diffraction patterns.
+The Pauling File project also distributes the Material phases data system, which contains largely identical data.
+Since the PDF5+ lists the MPDS as a database source, there is likely significant overlap between the experimental patterns found in the PDF5+ and those found in Pearson's crystal data.
 As of November 2024, Pearson's crystal data is available to researchers through a purchase of a one-year license starting at a price point of 2200 \euro.\\
-
-% TODO: I need to check that by asking
-%Pearson's crystal data is provided by the Pauling File project which also provides the Materials Phases Data System database (MPDS). 
-%Since the MPDS is listed as a database source for the PDF5+ \cite{kaduk2007}, there is likely a significant overlap between the experimental powder diffraction data found in Pearson's crystal data and that found in the PDF.
- %\tablefootnote{The PDF and Pearson's Crystal data both derive some of their data from the Linus Pauling File, which has a similar number of experimental powder diffraction patterns as both the PDF and Pearson's Crystal Data.
-Hence there is likely significant overlap in the patterns available in the PDF and in Pearson's crystal data.}
 
@@
 Aside from the databases mentioned above, we have also investigated several other crystal structure resources in search of experimental powder diffraction data.
-Crystal structure resources that were investigated but not found to contain any appreciable amount of publicly available experimental powder diffraction data include the Inorganic Crystal Structure Database,   Cambridge Structural Database,  the Materials Project database,  the Crystallographic and Crystallochemical Database, the Bilbao Incommensurate Crystal Structure Database,  the Mineralogy Database, IUCr Raw data letters, the U.S.
-Naval Research Laboratory Crystal Lattice-Structures, the Athena Mineral database and the Protein data bank.
-This is to be expected as most structure solutions are achieved through single-crystal diffraction rather than powder diffraction.\\
+Crystal structure resources that were investigated but not found to contain any appreciable amount of publicly available experimental powder diffraction data include the Inorganic Crystal Structure Database \cite{ICSDWeb}, the Cambridge Structural Database \cite{CambridgeWeb}, the Materials Project database \cite{MatProjWeb}, the Crystallographic and Crystallochemical Database \cite{CrystallochemicalWeb}, the Bilbao Incommensurate Crystal Structure Database \cite{BilbaoWeb}, the Mineralogy Database \cite{MineralogyWeb}, the IUCr Raw data letters \cite{IUCrWeb}, the U.S.
+Naval Research Laboratory Crystal Lattice-Structures \cite{NRLWeb}, the Athena Mineral database \cite{AthenaWeb} and the Protein data bank\cite{PDBWeb}.
+This is to be expected as most structure solutions are achieved through single-crystal diffraction rather than powder diffraction.
+\\

### 2024-11-28 chapters/04_our_dataset.tex
Described overlap in experimental pXRD data in PDF5+ and Pearson's Crystal data

@@
 % Paragraph by Moritz Wolf 
-\subsubsection*{Institute of Catalysis Research and Technology, KIT}
+\subsubsection*{Institute of Catalys1is Research and Technology, KIT}
 
@@
 Measurements were taken in grazing incidence geometry, i.e.
-using a beam incidence angle of $1\degree$ Two-dimensional diffraction images were recorded using a Dectris Pilatus 1M area detector at an angle between $34\degree$ and $36\degree$ with a sample-to-detector distance of roughly $190 \ \si{mm}$.
+using a beam incidence angle of $1\degree$.
+Two-dimensional diffraction images were recorded using a Dectris Pilatus 1M area detector at an angle between $34\degree$ and $36\degree$ with a sample-to-detector distance of roughly $190 \ \si{mm}$.
 The two-dimensional data were calibrated using an Al$_{2}$O$_{3}$ calibration standard and integrated along the azimuthal angle.
@@
 % Paragraph by Bin Cao and Tong-yi Zhang
-\subsubsection*{Guangzhou Municipal Key Laboratory of Materials Informatics, HKUST}
+\subsubsection*{Guangzhou Municipal Key Laboratory of Materials Informatics, HKUST(GZ)}
 
-In the past two years, we established a small-scale experimental PXRD database called the X-Ray Phase Identification Public Experimental Dataset.
+%In the past two years, we established a small-scale experimental PXRD database called the X-Ray Phase Identification Public Experimental Dataset.
 The data in XRed is generated in our lab using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers under Cu targets.
@@
 
-In the past two years, we established a small-scale experimental pXRD database called the X-Ray Phase Identification Public Experimental Dataset (XRed).
-\footnote{XRed can be found on GitHub at \url{https://github.com/WPEM/XRED}.} The data in XRed is generated in our lab using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers using \ce{Cu} X-ray sources.
-Hundreds of pXRD patterns have been refined, labeled with CIF files, and organized by elemental systems.
-The dataset includes original experimental files across single-phase to five-phase mixtures.
-This project is progressing toward integration with the comprehensive opXRD database to establish a large-scale, long-term experimental resource.
+In the past two years, we have established a small-scale experimental powder X-ray diffraction database called the X-Ray Phase Identification Public Experimental Dataset (XRed).
+\footnote{XRed can be accessed on GitHub at \url{https://github.com/WPEM/XRED}.} The primary objective of this database is to support the development of intelligent phase identification technology, serving as a foundation for data collection for future large-scale machine learning calculations.
+
+The data in XRed were generated in our laboratory using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers, all utilizing \ce{Cu} X-ray sources.
+Hundreds of pXRD patterns have been obtained through XRD refinement, enabling the determination of detailed lattice constants, atomic sizes, and other structural parameters.
+These patterns are labeled with CIF files that document the refined structures.
+The dataset is organized by elemental systems and includes original experimental files ranging from single-phase to five-phase mixtures, as well as mixtures studied for different tasks.
+This project is progressing toward integration with the comprehensive opXRD database to create a large-scale, long-term experimental resource.
 

### 2024-11-28 chapters/05_how_to_use.tex
Added paragraph by Hodge group
Described overlap in experimental pXRD data in PDF5+ and Pearson's Crystal data

@@
+\subsubsection*{How to contribute and how to use}
+
 TODO: A small tutorial with a few python commands to show how easy it is to use our dataset through the provided API.
@@
 
-% caobin
-We have launched Xqueryer (http://xqueryer.caobin.asia/), a free online platform for phase identification.
-Users can upload experimental PXRD data to identify and match structures from the Materials Project database.
-Uploaded data is stored and integrated into our opXRD database.
+% 
+We have launched Xqueryer (\url{http://xqueryer.caobin.asia/}), a cutting-edge, free online platform for intelligent structure identification.
+Users can upload experimental pXRD data to identify crystal structures and extract all relevant information available in the Materials Project database.
+All usage complies with Xqueryer's terms of service.
+Uploaded data are securely stored and integrated into the opXRD database.

### 2024-11-28 chapters/01_introduction.tex
Removed unneccessary footnotes in existing datasets; Converted remaining footnotes to web citations
Described overlap in experimental pXRD data in PDF5+ and Pearson's Crystal data
Described and hinted at imperfections that need to be simulated

@@
 However, the performance substantially drops off when these models are applied to data originating from experiments \cite{Schopmans2023,zhang2024crystallographic}[cite 2-3 papers, incl. ours].
-The patterns observed in experiments suffer from a host of imperfections that need to be modeled accurately in order to be able to transfer models trained on simulated data to experimental patterns.
-This includes noise, background, preferred orientation of the sample, ...
+The patterns observed in experiments exhibit imperfections that do not appear when simulating the X-ray diffraction pattern of a powder sample under ideal conditions.
+The physical effects underlying these imperfections need to be modeled accurately in order to be able to transfer models trained on simulated data to experimental patterns.
 Most approaches currently model these imperfections as a simple signal based on a heuristic functional form added to the simulated diffractograms.
+
+%This includes preferred grain orientation, instrument noise, crystallographic defects, internal stress, background originating from various sources, preferred orientation of the sample, ...
 
@@
 Furthermore, even completely unlabeled data can enable machine learning researchers to evaluate how closely their simulations represent experimental data, identify what features are missing, and modify their simulation algorithms accordingly to arrive at a simulation that mirrors how data appears in reality.
-One can bring the simulated patterns closer to experiment by improving the simulation itself by properly modeling more of the observed physical effects such as preferred orientation, peak shapes resulting from different sources, etc.
+In this way, one can bring the simulated patterns closer to those observed in experiments by more accurately modeling the scattering process.
+Improving the simulation of data involves accounting for more of the physical effects that influence the scattering process such as preferred grain orientation, variations in grain size, crystal defects, the impact of temperature on the scattering process, internal stress, and X-ray-induced fluorescence.
+\cite{cao2024simxrd} Additionally, it should be considered that varying experimental setups can produce distinct powder diffraction patterns on the same sample.
+Features that may vary between experimental setups include the shape of diffraction peaks and the employed X-ray source.
+Some experimental setups may also exhibit imperfections such as displacement of the sample from its intended position leading to the recorded scattering angles being slightly erroneous or background events originating from instrument noise or  X-rays scattering off equipment onto the screen.\\
+
+% Polarizaiton and detector geometry
+
 Additionally, some of the observed imperfections, such as background and noise, can be modeled using machine-learning approaches.
@@
     \centering
-    \includegraphics{figures/overview.pdf}
+    \includegraphics[width=\linewidth]{figures/overview2.png}
     \caption{Overview of the data collection pipeline.

### 2024-11-28 chapters/01_introduction.tex
Added polarizationn and detector geometry imperfections

@@
 
-%This includes preferred grain orientation, instrument noise, crystallographic defects, internal stress, background originating from various sources, preferred orientation of the sample, ...
-
 At the moment, the number of openly available datasets for experimental powder data is very limited.
@@
 \cite{cao2024simxrd} Additionally, it should be considered that varying experimental setups can produce distinct powder diffraction patterns on the same sample.
-Features that may vary between experimental setups include the shape of diffraction peaks and the employed X-ray source.
+Features that may vary between experimental setups and influence the recorded pattern include the shape of diffraction peaks, the wavelength and polarization of the employed X-ray source, and the detector geometry.
 Some experimental setups may also exhibit imperfections such as displacement of the sample from its intended position leading to the recorded scattering angles being slightly erroneous or background events originating from instrument noise or  X-rays scattering off equipment onto the screen.\\
-
-% Polarizaiton and detector geometry
 

### 2024-11-28 chapters/01_introduction.tex
Added definition of transfer learning/style transfer

@@
 In this way, one can bring the simulated patterns closer to those observed in experiments by more accurately modeling the scattering process.
-Improving the simulation of data involves accounting for more of the physical effects that influence the scattering process such as preferred grain orientation, variations in grain size, crystal defects, the impact of temperature on the scattering process, internal stress, and X-ray-induced fluorescence.
-\cite{cao2024simxrd} Additionally, it should be considered that varying experimental setups can produce distinct powder diffraction patterns on the same sample.
-Features that may vary between experimental setups and influence the recorded pattern include the shape of diffraction peaks, the wavelength and polarization of the employed X-ray source, and the detector geometry.
-Some experimental setups may also exhibit imperfections such as displacement of the sample from its intended position leading to the recorded scattering angles being slightly erroneous or background events originating from instrument noise or  X-rays scattering off equipment onto the screen.\\
+Improving the simulation of data involves accounting for more of the physical effects that influence the scattering process such as preferred grain orientation, variations in grain size, crystal defects, the impact of temperature on the scattering process, internal stress, and X-ray-induced fluorescence\cite{cao2024simxrd, Waseda2011, Pecharsky2023}.
+Additionally, it should be considered that varying experimental setups can produce distinct powder diffraction patterns on the same sample.
+Features that may vary between experimental setups and influence the recorded pattern include the shape of diffraction peaks, the wavelength and polarization of the employed X-ray source, and the detector geometry
+\cite{Waseda2011, Pecharsky2023}.
+Some experimental setups may also exhibit imperfections such as displacement of the sample from its intended position leading to the recorded scattering angles being slightly erroneous \cite{hulbert2023} or background events originating from instrument noise or X-rays scattering off equipment onto the screen.\\
 
-Additionally, some of the observed imperfections, such as background and noise, can be modeled using machine-learning approaches.
-This is similar to the concept of style transfer\cite{Gatys2016, Ganin2015} found in the domain of machine learning applied to images.
-Here, the objective is to obtain a classification or regression model when only labeled images of one style (simulated data) and unlabeled or partially labeled images of another style (experimental data) are available.
-\\
+Additionally, some of the observed imperfections, such as background and noise, can be accounted for using transfer learning approaches.
+In transfer learning the objective is to transfer the capabilities of a model learned on a source domain in which labeled data is abundant to a target domain in which labeled data is sparse\cite{Zhuang2021}.
+In this context, the source domain is simulated powder diffraction patterns and the target domain is experimental powder diffraction patterns.
+Many approaches to transfer learning have been proposed, particularly in the domain of image classification \cite{Gatys2016, Ganin2015}.
+These existing techniques could potentially be adapted to facilitate transfer learning in the context of pXRD patterns.
 
@@
 The broad range of available experimental samples contained in the opXRD v1.0 database makes it possible to apply state-of-the-art ML approaches to the domain of powder XRD analysis.
+Sharing raw powder diffraction data is in the interest of
 We hope that this will yield more advanced automated ML approaches in the future, which can readily be applied in high-throughput experimentation pipelines.

### 2024-11-28 chapters/04_our_dataset.tex
Added definition of transfer learning/style transfer

@@
 We will continue growing the opXRD dataset, thus additional contributors are very welcome to contact us and submit data through the website.
-To find out more about how to contribute to this dataset, visit our website \footnote{Datasets can be contributed to opXRD under \url{https://xrd.aimat.science}.} specially designed to collect this dataset.
+To find out more about how to contribute to this dataset, visit our website\footnote{Datasets can be contributed to opXRD under \url{https://xrd.aimat.science}.} specially designed to collect this dataset.

### 2024-11-28 chapters/01_introduction.tex
Revised paragraph describing projected impact of opXRD database

@@
 The broad range of available experimental samples contained in the opXRD v1.0 database makes it possible to apply state-of-the-art ML approaches to the domain of powder XRD analysis.
-Sharing raw powder diffraction data is in the interest of
-We hope that this will yield more advanced automated ML approaches in the future, which can readily be applied in high-throughput experimentation pipelines.
+We hope that this can drive ML research in this field towards more advanced automated analysis workflows that can accelerate material science research through ready application in high-throughput experimentation pipelines.
 

### 2024-11-28 chapters/01_introduction.tex
Added paragraph arguing for publication via opXRD in interest of open science

@@
 Further details can be found on the opXRD website (Figure~\ref{fig:overview}).
+\\
+
+As argued by Aranda and Kroon-Batenburg \textit{et al}.\cite{Aranda2018, Kroon-Batenburg2024}, sharing raw powder diffraction data is not only in the interest of furthering machine learning research but is also in line with open science principles.
+It furthers the ability of other researchers to reproduce published work and in turn, adds to the credibility of the publisher of the data.
+Considering our streamlined submission and post-processing framework, a publication of powder diffraction data through the opXRD database is a simple way of realizing these principles.
+Compared to uploading the data to Zenodo it also has the added benefit that the data is contributed towards a large, homogenous dataset with a standardized interface.
+This makes the data more easily accessible to other researchers and provides more value to researchers seeking great quantities of data compared to an isolated submission on Zenodo.
+\\
 
@@
 We hope that this can drive ML research in this field towards more advanced automated analysis workflows that can accelerate material science research through ready application in high-throughput experimentation pipelines.
-
-% TODOS: These papers talk about the importance of sharing raw diffraction data/ raw powder diffraction data them
-%  1 | \cite{Aranda2018}
-% 2 | Raw diffraction data and reproducibility

### 2024-11-28 chapters/04_our_dataset.tex
Moved notice about further contributions to summary and outlook

@@
 %% Paragraph by FX Coudert and Arthur Hardiagon
-\subsubsection*{Chimie Paris Tech, PSL University}
+\subsubsection*{Chimie Paris Tech, PSL}
 
@@
 The COD is, to our knowledge, the largest open-access collection of experimental crystal structures of organic, inorganic, and metal-organic compounds and minerals, containing more than 500,000 entries.
-\footnote{Available online at \url{https://www.crystallography.net/cod/}.} The data in the COD is placed in the public domain and licensed under the CC0 License.
+The data in the COD is placed in the public domain and licensed under the CC0 License.
 Of the entire COD database 5432 structures contained at least one tag from the {CIF\_POW} dictionary, i.e., a tag relating to powder diffraction studies.
@@
 % Paragraph by Moritz Wolf 
-\subsubsection*{Institute of Catalys1is Research and Technology, KIT}
+\subsubsection*{Institute of Catalysis Research and Technology, KIT}
 
@@
 These alloys were selected to study compositional effects on stacking fault and growth twin boundary formation in sputtered materials \cite{2024AcMat.27019839A,alwen2024combinatorial}.
+\\
+
 XRD spectra were collected using a Malvern Panalytical Empyrean X-ray diffractometer at the Forschungszentrum Julich (FZJ) Institute for Energy and Climate Research, Structure and Function of Materials (IEK-2), which enabled automated XRD characterization due to its programmable XY stage and laser sensor for Z height correction.
@@
 % Paragraph by Tim Kodalle
-\subsubsection*{Lawrence Berkeley National Lab}
+\subsubsection*{Molecular Foundry Division \& Advanced Light Source \& Chemical Sciences Division, LBNL}
 
@@
 Christian Kaufmann and his team at Helmholtz-Zentrum Berlin (HZB) in Germany.
-\\
-
 A second dataset was collected from spin-coating metal-halide perovskite precursor solutions with varying compositions of \ce{MAPb(I_{1-x}Br_x)3} (where MA = Methylammonium and x = 0, 0.33, 0.5, 0.67, 1) spin-coated onto glass substrates.
@@
 \footnote{XRed can be accessed on GitHub at \url{https://github.com/WPEM/XRED}.} The primary objective of this database is to support the development of intelligent phase identification technology, serving as a foundation for data collection for future large-scale machine learning calculations.
+\\
 
@@
 \\
-
-%---------------------------------
-
-We will continue growing the opXRD dataset, thus additional contributors are very welcome to contact us and submit data through the website.
-To find out more about how to contribute to this dataset, visit our website\footnote{Datasets can be contributed to opXRD under \url{https://xrd.aimat.science}.} specially designed to collect this dataset.

### 2024-11-28 chapters/05_how_to_use.tex
Moved notice about further contributions to summary and outlook

@@
 \subsubsection*{How to contribute and how to use}
+
+We will continue growing the opXRD dataset, thus additional contributors are very welcome to contact us and submit data through the website.
+To find out more about how to contribute to this dataset, visit our website\footnote{Datasets can be contributed to opXRD under \url{https://xrd.aimat.science}.} specially designed to collect this dataset.
 

### 2024-12-01 chapters/00_abstract.tex
Added dataset overview tables

@@
 Powder X-ray diffraction (pXRD) experiments are a cornerstone for materials structure characterization.
-Despite their widespread application, the analysis of pXRD diffractograms still presents a significant challenge to automation and thus a bottleneck in high-throughput experimentation of automated materials discovery in self-driving labs.
+Despite their widespread application, the analysis of pXRD diffractograms still presents a significant challenge to automation and thus a bottleneck in high-throughput materials discovery in self-driving labs.
 Machine learning has emerged as a promising research direction to resolve this bottleneck by enabling automated powder diffraction data analysis.
-A notable difficulty in applying machine learning to this domain is the lack of sufficiently sized experimental datasets, which has relegated machine learning researchers to train primarily on simulated data.
-However, models trained on simulated pXRD patterns showed limited generalization ability when applied to experimental patterns, particularly with low-quality experimental pattern that has high noise levels and elevated background.
+A notable difficulty in applying machine learning to this domain is the lack of sufficiently sized experimental datasets, which has constrained machine learning researchers to train primarily on simulated data.
+However, models trained on simulated pXRD patterns showed limited generalization ability when applied to experimental patterns, particularly for low-quality experimental patterns with high noise levels and elevated background.
 %Therefore, developing an experimental database offers at least two key advantages: (1) providing a comprehensive benchmark dataset to clarify model performance in practical applications, and (2) advancing phase identification tasks by supplying target-domain data within a transfer learning framework.
-With the Open Experimental Powder X-Ray Diffraction Database (opXRD), we aim to provide an openly available and easily accessible dataset of both labeled and unlabeled experimental powder diffractograms, providing machine learning researchers with a large quantity of real experimental diffractograms collected from a broad range of samples.
-The labeled data can be used as a benchmark to evaluate the performance of models on real experimental data and unlabeled data can help improve the performance of models on experimental data through transfer learning methods.
+With the Open Experimental Powder X-Ray Diffraction Database (opXRD), we aim to provide an openly available and easily accessible dataset of both labeled and unlabeled experimental powder diffractograms, providing machine learning researchers with a large quantity of real experimental diffractograms collected from a broad range of materials systems.
+The labeled data can be used as a benchmark to evaluate the performance of models on real experimental data and the unlabeled data can help improve the performance of models on experimental data, e.g.
+through transfer learning methods.
 %We provide almost barrier-free software tools that allow experimental researchers to find their data and make it accessible - in virtually any widely used format.

### 2024-12-01 chapters/01_introduction.tex
Added dataset overview tables

@@
 After fabricating a new material, a number of analysis techniques can be used to characterize the sample.
-One method that can be used for phase identification, phase quantification, grain size characterization, and partially to determine the crystal structure of a new material is powder X-ray diffraction.
-To determine crystal structure from pXRD data, an initial guess is optimized using Rietveld refinement.
-In this process, an initial crystal structure including unit cell parameters is provided by the operator of the analysis software and fitted to the observed diffractogram in an iterative optimization process based on simulated forward prediction of expected pXRD diffractograms given the current crystal structure\cite{Dinnebier2019, Cano2021}.
+One method that can be used for phase identification, phase quantification, grain size characterization, and to determine the crystal structure of a new material is powder X-ray diffraction.
+To determine the crystal structure from pXRD data, an initial guess is typically optimized using Rietveld refinement.
+In this process, an initial crystal structure including unit cell parameters is provided by the researcher and subsequently fitted to the observed diffractogram in an iterative optimization process based on simulated forward prediction of pXRD diffractograms given the current crystal structure\cite{Dinnebier2019, Cano2021}.
 As Rietveld refinement is a local optimization method, the result of the refinement procedure is only as good as the initial guess it was provided with.
@@
 It is not scalable to the degree required to keep up with advances in throughput and efficiency in other steps of the experimentation pipeline.
-In a Rietveld refinement process, the operator starts by setting parameters that characterize the background and peak shape of the powder diffraction pattern and determining a crystal structure from which the optimization process can start.
+In a Rietveld refinement process, the operator starts by setting parameters that characterize the background and peak shape of the powder diffraction pattern and then determines a crystal structure from which the optimization process can start.
 The latter is usually done using search-match software, which identifies crystal structures with similar powder diffraction patterns from a database of crystal structures with accompanying powder diffraction patterns.
@@
 What makes the refinement process time-consuming and difficult is finding the correct order in which to refine the crystal structure parameters\cite{Ozaki2020}.
-Additionally, if the fit develops into an unphysical set of parameters in this iterative determination of parameters, the operator must intervene to backtrack to the last step.
+\\ %Additionally, if the fit develops into an unphysical set of parameters in this iterative determination of parameters, the operator must intervene to backtrack to the last step.
 \\
@@
 The physical effects underlying these imperfections need to be modeled accurately in order to be able to transfer models trained on simulated data to experimental patterns.
-Most approaches currently model these imperfections as a simple signal based on a heuristic functional form added to the simulated diffractograms.
+Most approaches currently model these imperfections as a simple signal based on a heuristic functional form added to the simulated diffractograms.\\
 
@@
 Further commercial datasets such as KKK and LLL exist, but license fees and restrictions in their use to train publishable models or even publish the training data limit their usability.
-Collecting a labeled dataset of raw experimental powder X-ray diffraction data of sufficient size to directly train a model that predicts crystal structure information is likely out of reach.
+Collecting a labeled dataset of raw experimental powder X-ray diffraction data of sufficient size to directly train a model that predicts crystal structure information is currently out of reach.
 The most advanced ML models currently are trained on approximately $10^7 - 10^8$ simulated diffractograms, which is - to the best of our knowledge - three orders of magnitude larger than the largest currently curated experimental dataset (PDF-5+ with approximately 20,000 experimental patterns) and even x orders of magnitude larger than the number of unlabeled diffractograms in the opXRD dataset we present here.
@@
 However, even small labeled and also unlabeled experimental diffractograms hold significant value in practice.
-First of all, labeled data can be used to better test and benchmark existing and new automated analysis approaches, which yields a better comparison than when using the very limited currently available datasets, such as the commonly used RRUFF mineral dataset\cite{Armbruster2015}.
+First of all, labeled data can be used to test and benchmark existing and new automated analysis approaches, which yields a better comparison than when using the very limited currently available datasets, such as the commonly used RRUFF mineral dataset\cite{Armbruster2015}.
 \\
@@
 
-Additionally, some of the observed imperfections, such as background and noise, can be accounted for using transfer learning approaches.
+Additionally, some of the observed imperfections can be accounted for using transfer learning approaches.
 In transfer learning the objective is to transfer the capabilities of a model learned on a source domain in which labeled data is abundant to a target domain in which labeled data is sparse\cite{Zhuang2021}.
@@
 Many approaches to transfer learning have been proposed, particularly in the domain of image classification \cite{Gatys2016, Ganin2015}.
-These existing techniques could potentially be adapted to facilitate transfer learning in the context of pXRD patterns.
+These existing techniques can be adapted to facilitate transfer learning in the context of pXRD patterns.\\
 
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/overview2.png}
+    \includegraphics[width=\linewidth]{figures/overview.png}
     \caption{Overview of the data collection pipeline.
-Datasets are submitted using an online submission form or our submission helper software.
-After post-processing and data homogenization, we create a Zenodo entry for each user submission and subsequently include the submission in the opXRD database.}
+Datasets are submitted using an online submission form, optionally with the help of our submission helper software.
+After post-processing and data homogenization, we offer creation of a Zenodo entry for each user submission and subsequently include the submission in the opXRD database.}
     \label{fig:overview}
@@
 % How to contribute? This aspect deserves a separate paragraph
-The opXRD database is planned as a growing, community-driven initiative.
+The opXRD database is intended as a growing, community-driven initiative.
 The dataset we present here is the first version, but we hope to further increase the dataset size through active engagement with the pXRD community.
 Our primary objective is to minimize the effort and thus the barrier to contributing experimental data to the opXRD database.
-Thus, we developed a program that helps to find and share data from e.g.
-pXRD lab computers.
+Thus, we developed a program that helps to find and share data from pXRD lab computers.
 In detail, users can select their most common pXRD file types, the program lists all files of that type, and users can select or deselect certain folders or files for sharing.
@@
 Considering our streamlined submission and post-processing framework, a publication of powder diffraction data through the opXRD database is a simple way of realizing these principles.
-Compared to uploading the data to Zenodo it also has the added benefit that the data is contributed towards a large, homogenous dataset with a standardized interface.
-This makes the data more easily accessible to other researchers and provides more value to researchers seeking great quantities of data compared to an isolated submission on Zenodo.
-\\
+Compared to publishing data individually, it also has the added benefit that the data is contributed towards a large, homogenous dataset with a standardized interface.
+This makes the data more easily accessible to other researchers and provides more value to researchers seeking large quantities of data compared to an isolated submission on Zenodo.
+However, further data annotation with metadata is required to fulfill the FAIR data principles.\\
 
 The broad range of available experimental samples contained in the opXRD v1.0 database makes it possible to apply state-of-the-art ML approaches to the domain of powder XRD analysis.
-We hope that this can drive ML research in this field towards more advanced automated analysis workflows that can accelerate material science research through ready application in high-throughput experimentation pipelines.
+We hope that this can drive ML research in this field towards more advanced automated analysis workflows that can accelerate materials science research through ready application in high-throughput experimentation pipelines.

### 2024-12-01 chapters/03_SOTA_ML.tex
Added dataset overview tables

@@
 
-\paragraph{ML for pXRD} Several approaches have applied machine learning methods to classification and regression tasks for powder diffractograms.
+\paragraph{ML for pXRD}
+Several approaches have applied machine learning methods to classification and regression tasks for powder diffractograms.
 
@@
 With their best model for the crystal system prediction on ICSD structures, they achieved a test accuracy of \SI{92.2}{\percent}.
-Park {\it et al.} reached a test accuracy of roughly 81\% for a CNN, which classifies space groups of simulated single-phase diffractograms \cite{Park2017}.
+Park {\it et al.} reached a test accuracy of roughly \SI{81}{\percent} for a CNN, which classifies space groups of simulated single-phase diffractograms \cite{Park2017}.
 \\
 
-A regression analysis on lattice parameters within a broader framework encompassing all material classes was conducted by Chitturi et al\cite{Chitturi2021}.
+A regression analysis on lattice parameters within a broader framework encompassing all material classes was conducted by Chitturi {\it et al} \cite{Chitturi2021}.
 They developed a distinct CNN for each crystal system, utilizing a merged dataset from both the ICSD and the Cambridge Structural Database, and managed to achieve a mean absolute percentage error of about \SI{10}{\percent} for the lattice lengths, although they encountered difficulties in accurately predicting angles.
@@
 
-TODO: We need a short literature review here; needs to also cover papers published after our previous paper!
+Schuetzke {\it et al.} trained a classifier to classify if a diffractogram stems from an amorphous, single-phase, or multi-phase sample \cite{Schuetzke2024}.
+Due to the lack of experimental pXRDs, they build a pipeline to augment simulated diffractograms by, among other things, slightly varying the underlying crystal lattice.
+
+In our 2023 paper, we came up with the approach to generate crystal structures and their corresponding pXRD patterns on the fly during the training process \cite{Schopmans2023}.
+This approach defeats the issue of
+a limited dataset size, which limits the depth of neural networks that can be trained.
+However, when we applied our space group classification model to experimental patterns from the RRUFF database the accuracy dropped substantially.
+Augmenting our simulated patterns with background, noise, and impurities would help to bring simulated diffractograms closer to experimental ones, making models trained on them more performant on experimental diffractograms.
+For this augmentation, background and noise statistics from a broader experimental pXRD database would be very valuable.
 
@@
 
-TODO: briefly discuss papers about multiphase classification and remixing of diffraction patterns
-Non-negative Matrix Factorization 2018: \url{https://link.springer.com/chapter/10.1007/978-3-319-93031-2_4}
-non-negative matrix factorization 2018: \url{https://ojs.aaai.org/aimagazine/index.php/aimagazine/article/view/2785}
 Physically-informed Graph-based DRNet (PG-DRNet) 2023: \url{https://ieeexplore.ieee.org/abstract/document/10459774/}
-
-TODO: discuss in more detail why domain randomization as we did in the last paper so far failed for background and noise, and thus why we need background and noise statistics from experiment to enhance simulated datasets and transfer them to the experimental data distribution.
 

### 2024-12-01 chapters/04_our_dataset.tex
Added dataset overview tables

@@
+In collaboration with several other institutions we have collected a dataset of diffractograms stemming from experiments, some of them labeled with corresponding structural information.
+The following texts detail the contribution of each research group.
+Most data was collected using \ce{Cu} radiaton sources which have a  $K_{\alpha1}$ wavelength of $\lambda=1.54056\text{\AA}$ and a $K_{\alpha2}$ wavelength of $\lambda=1.54439\text{\AA}$.
+
+\begin{table}[!htb]
+\centering
+\caption{Labeled opXRD database entries}
+\label{tab:labeled}
+\begin{tabular}{@{}ccc@{}} % Adjusted columns
+\toprule
+Institution & Research project & No. patterns\footnote{This number is subject to change} \\
+\midrule
+CNRS        & Extracted from the Crystallography open database & 895 \\
+USC         & Study of CuNi and CuAl alloys & 338 \\
+HKUST(GZ)   & Reference database for phase identification & 21 \\
+EMPA        & Tin halide perovskites and combinatorial Zn-V-N library & 790 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}[!htb]
+\centering
+\caption{Unlabeled opXRD database entries}
+\label{tab:unlabeled}
+\begin{tabular}{@{}ccc@{}} % Adjusted columns
+\toprule
+Institution & Research project & No. patterns\footnote{This number is subject to change} \\
+\midrule
+INT         & Compilation of various research projects & 19780 \\
+IKFT        & Compilation of various research projects & 64 \\
+LBNL        & Time series recording of in-situ measurements & n/a \\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \begin{figure*}[!ht]
@@
 
-In collaboration with several other institutions we have collected a dataset of diffractograms stemming from experiments, some of them labeled with corresponding structural information.
-The following texts detail the contribution of each research group.
-Most data was collected using \ce{Cu} radiaton sources which have a  $K_{\alpha1}$ wavelength of $\lambda=1.54056\text{\AA}$ and a $K_{\alpha2}$ wavelength of $\lambda=1.54439\text{\AA}$.
-
 %% Paragraph by FX Coudert and Arthur Hardiagon
-\subsubsection*{Chimie Paris Tech, PSL}
+\subsubsection*{Chimie ParisTech, PSL University}
 

### 2024-12-01 chapters/05_how_to_use.tex
Added dataset overview tables

--- 2024-11-21 chapters/05_how_to_use.tex
@@
-TODO: A small tutorial with a few python commands to show how easy it is to use our dataset through the provided API.
+Next to the availability of the opXRD dataset on Zenodo, we further provide an easy-to-use Python library (\url{https://github.com/aimat-lab/xrdpattern}) to download and interface the dataset.
+The dataset can be accessed using the simple command 
+\pyth{PatternDB.load(data_dirpath)}. Our tool further includes options for standardization, plotting, and the conversion to \emph{PyTorch} tensors.
+We provide the following Jupyter notebook that showcases the functionality of the library in more detail: <Link to github Jupyter notebook on Colab>
 
-A major advantage of this dataset over other comparable datasets is that it is very easy to handle in python and in PyTorch.
-Our accompanying python library xrdpattern (\url{https://github.com/aimat-lab/xrdpattern}) provides a means to read the dataset simply by using PatternDB.load(data\_dirpath).
-The patterns attribute of this class is a list of the indivdual pattern diffractograms, each of which supports a standardize, plot and to tensor method.
-\\
+%The patterns attribute of this class is a list of the indivdual pattern diffractograms, each of which supports a standardize, plot and to tensor method.
 The standardize method returns a "standardized" version of the pattern with a fixed angle range, a fixed number of entries and intensities normalized to the $[0,1]$ interval.
-
-% caobin
-We have launched Xqueryer (http://xqueryer.caobin.asia/), a free online platform for phase identification.
-Users can upload experimental PXRD data to identify and match structures from the Materials Project database.
-Uploaded data is stored and integrated into our opXRD database.

### 2024-12-01 chapters/06_summary_and_outlook.tex
Added dataset overview tables

With the opXRD dataset, a curation of XXX unlabeled and YYY labeled experimental powder XRD patterns from a wide range of different materials systems, we provide the largest currently available source of experimental XRD patterns. With this, we address the need for experimental data that arises when developing algorithms and analysis tools for XRD data, both based on machine learning and classical approaches. The data can be used for the actual method development and for testing. Our dataset is a valuable and so far missing resource to drive further developments in the automated analysis of XRD data.

\paragraph{How to contribute}

Rather than a finished project, the opXRD dataset is an ongoing effort to collect experimental powder XRD data. We invite everyone in possession of experimental powder XRD data to submit it to our dataset, to further improve the dataset and aid further developments in this field. Our submission page (\url{https://xrd.aimat.science/}) and submission helper software will be kept available to collect more data. We will keep updating and maintaining the dataset with incoming new submissions.

%We have launched Xqueryer (\url{http://xqueryer.caobin.asia/}), a cutting-edge, free online platform for intelligent structure identification. Users can upload experimental pXRD data to identify crystal structures and extract all relevant information available in the Materials Project database. All usage complies with Xqueryer's terms of service. Uploaded data are securely stored and integrated into the opXRD database.

### 2024-12-03 chapters/01_introduction.tex
Updated introductory paragraphs on rietveld refinement

@@
 One method that can be used for phase identification, phase quantification, grain size characterization, and to determine the crystal structure of a new material is powder X-ray diffraction.
-To determine the crystal structure from pXRD data, an initial guess is typically optimized using Rietveld refinement.
-In this process, an initial crystal structure including unit cell parameters is provided by the researcher and subsequently fitted to the observed diffractogram in an iterative optimization process based on simulated forward prediction of pXRD diffractograms given the current crystal structure\cite{Dinnebier2019, Cano2021}.
-As Rietveld refinement is a local optimization method, the result of the refinement procedure is only as good as the initial guess it was provided with.
+Typically crystal structures are determined from pXRD data through Rietveld refinement.
+In Rietveld refinement, an initial crystal structure model is fitted to the observed diffractogram by iteratively updating the structural model.
+Each update of the structural model seeks to minimize the difference between the observed diffractogram and the diffractogram simulated from the current structural model \cite{Dinnebier2019, Cano2021}.
+As Rietveld refinement is a local optimization method, the result of the refinement procedure is generally only as good as the initial structural model the process started from.
 \\
 
-% Manual vs (ML) automated Refinement
+% Manual rietveld refinement and its problems
 Manually performing Rietveld refinement is time-consuming and often requires expert knowledge.
 It is not scalable to the degree required to keep up with advances in throughput and efficiency in other steps of the experimentation pipeline.
-In a Rietveld refinement process, the operator starts by setting parameters that characterize the background and peak shape of the powder diffraction pattern and then determines a crystal structure from which the optimization process can start.
-The latter is usually done using search-match software, which identifies crystal structures with similar powder diffraction patterns from a database of crystal structures with accompanying powder diffraction patterns.
-However, attempting to refine all crystal structure parameters at once, starting from this initial structure, is known to lead to unphysical results.
-What makes the refinement process time-consuming and difficult is finding the correct order in which to refine the crystal structure parameters\cite{Ozaki2020}.
-\\ %Additionally, if the fit develops into an unphysical set of parameters in this iterative determination of parameters, the operator must intervene to backtrack to the last step.
+In a Rietveld refinement process, the operator starts by setting parameters that characterize the background and peak shape of the powder diffraction pattern and determines a crystal structure model from which the optimization process can start.
+This model is usually obtained using search-match software, which identifies crystal structures with similar powder diffraction patterns from a database of crystal structures with accompanying powder diffraction patterns.
+Databases commonly used for this purpose include the Powder Diffraction File\cite{GatesRector2019}, the Inorganic Crystal Structure Database \cite{Belsky2002}, and the Crystallography Open Database \cite{Grazulis2009}.
+However, an initial structural model obtained from such a database via search-match software is not guaranteed to lead to an accurate structure solution through Rietveld refinement.
+Especially for novel structures, obtaining an initial structural model through matching to a database of known structures can be problematic.
+\\ %In such cases, more complex crystal structure solution approaches need to be applied to find suitable crystal structure guesses.
 \\
 
-Commonly used databases for obtaining a match include the Powder Diffraction File\cite{GatesRector2019}, the Inorganic Crystal Structure Database \cite{Belsky2002}, and the Crystallography Open Database \cite{Grazulis2009}.
-An initial guess obtained from such a database is by no means guaranteed to lead to an accurate structure solution, even if it is iteratively revised, either manually or using blackbox optimization\cite{Ozaki2020}.
-Especially for novel observed structure types, simply matching to a database of known structures will not yield good results.
-Therefore, more complex crystal structure solution approaches need to be applied to find suitable crystal structure guesses.
-\\
+Additionally, attempting to refine all crystal structure parameters at once, starting from this initial structure, is known to lead to unphysical results\cite{Ozaki2020}.
+Hence parameters are refined iteratively, with each iteration only refining a limited set of parameters.
+If, during the iterative determination of parameters, the structural model becomes unphysical the operator must intervene to backtrack to a previous successful step or start over entirely.
+Finding the correct order in which to refine structure parameters and finding the correct values for fixed parameters both present problems that add to the difficulty of the refinement process.
+
+%What makes the refinement process time-consuming and difficult is finding the correct order in which to refine the crystal structure parameters\cite{Ozaki2020}. \\ % \\
 
@@
 The physical effects underlying these imperfections need to be modeled accurately in order to be able to transfer models trained on simulated data to experimental patterns.
-Most approaches currently model these imperfections as a simple signal based on a heuristic functional form added to the simulated diffractograms.\\
+Most approaches currently model these imperfections as a simple signal based on a heuristic functional form added to the simulated diffractograms.
+Experimental data can provide clues on how to model these imperfections in a more realistic fashion.
+\\
 
@@
 Collecting a labeled dataset of raw experimental powder X-ray diffraction data of sufficient size to directly train a model that predicts crystal structure information is currently out of reach.
-The most advanced ML models currently are trained on approximately $10^7 - 10^8$ simulated diffractograms, which is - to the best of our knowledge - three orders of magnitude larger than the largest currently curated experimental dataset (PDF-5+ with approximately 20,000 experimental patterns) and even x orders of magnitude larger than the number of unlabeled diffractograms in the opXRD dataset we present here.
+The most advanced ML models currently are trained on approximately $10^7 - 10^8$ simulated diffractograms,
+%For example 4M sim patterns\cite{cao2024simxrd}
+which is - to the best of our knowledge - three orders of magnitude larger than the largest currently curated experimental dataset (PDF-5+ with approximately 20,000 experimental patterns) and even x orders of magnitude larger than the number of unlabeled diffractograms in the opXRD dataset we present here.
 \\
@@
 It contains pXRD patterns from single and multiphase materials from a wide variety of materials classes, including XXX, YYY, and ZZZ.
-Details about the experiments of contributing research groups are discussed in this paper, along with a detailed description of how to contribute further data and how to use the data.
+Details about the experiments of contributing research groups are discussed in section $\eqref{sec:our_dataset}$, along with a detailed description of how to contribute further data and how to use the data.
 \\
@@
 Datasets are submitted using an online submission form, optionally with the help of our submission helper software.
-After post-processing and data homogenization, we offer creation of a Zenodo entry for each user submission and subsequently include the submission in the opXRD database.}
+After post-processing and data homogenization, we offer the creation of a Zenodo entry for each user submission and subsequently include the submission in the opXRD database.}
     \label{fig:overview}
@@
 If labels are available, they can be shared with opXRD as well.
-Further details can be found on the opXRD website (Figure~\ref{fig:overview}).
+Further details can be found on the opXRD website\footnote{\url{https://xrd.aimat.science/}}.
 \\

### 2024-12-03 chapters/02_existing_datasets.tex
Added dataset overview tables
Updated introductory paragraphs on rietveld refinement

@@
-\pagebreak
-
 \subsubsection*{Existing experimental powder diffraction databases}
@@
 After inquiring with the ICDD in April 2024 we were told that only 20,800 of the powder diffraction patterns in the PDF5+ stemmed from experiments.
-Of these 20,800 entries, 10,954 contain information about the atomic coordinates of the underlying structures.\footnote{We reached out to the contact address given on their website, info@icdd.com.} Still, this makes the PDF5+, to the best of our knowledge, the largest collection of experimental powder diffraction data available to researchers.
-As of November 2024, the PDF5+ is available to researchers through a purchase of a one-year license starting at a price point of \$6265.00.\\
+Of these 20,800 entries, 10,954 contain information about the atomic coordinates of the underlying structures.%\footnote{We reached out to the contact address given on their website, info@icdd.com.}
+Still, this makes the PDF5+, to the best of our knowledge, the largest collection of experimental powder diffraction data available to researchers.
+As of November 2024, the PDF5+ is available to researchers through a purchase of a one-year license starting at a price point of \$6265.00.
+However, the ICDD does not allow researchers to train machine learning models on PDF5+ data, regardless of whether the resulting models are published.
+\\
 
@@
 \textbf{Crystallography Open Database:} \cite{CODWeb} The Crystallography Open Database (COD) is an open-access collection of crystal structures founded in 2003\cite{Graulis2009cod}.
-It currently provides over 500,000 crystal structures in the form of .cif files.
+It currently provides over 500,000 crystal structures.
 Of these files, 1063 contain the experimental powder diffraction data that was used to determine the underlying crystal structures of the investigated samples.
-Hence the experimental powder diffraction data contained in the COD is fully labeled.
-The .cif files include both lattice parameters and the contents of the unit cell of the underlying structure.
+Hence, the experimental powder diffraction data contained in the COD is labeled with the full crystal structure information.
 \\
@@
 \cite{lafuente2015} Managed by the University of Arizona, it was created to serve as a public repository for mineral identification and research.
-It contains \num{1290} powder diffraction patterns stemming from experiments each labeled with the lattice parameters and composition of the underlying structures.
+It contains \num{1290} powder diffraction patterns stemming from experiments each labeled with the lattice parameters and composition of the underlying structure.
 \\
 
-\textbf{PowBase}:\footnote{More information on PowBase can be found under \url{http://www.cristal.org/powbase/index.html}.
+\textbf{PowBase}:\footnote{More information on PowBase can be found at \url{http://www.cristal.org/powbase/index.html}.
 PowBase was an initiative suggested in the Structure Determination by Powder Diffractometry (SDPD) mailing list which was hosted on the same site.
@@
 Additionally, extracting powder diffraction data at scale is hindered by the fact that the data is often given in plain text files in non-standardized formats, which are difficult to parse programmatically.
-\\
+We are currently planning a systematic large-scale extraction of powder diffraction data from databases like Zenodo with the help of a large language model.
+We will include this data in a future release of our dataset.\\
 

### 2024-12-03 chapters/03_SOTA_ML.tex
Updated introductory paragraphs on rietveld refinement

@@
-%TODO: Papers to include:
-% -  Crystal Structure Determination from Powder Diffraction Patterns with Generative Machine Learning: (https://pubs.acs.org/doi/10.1021/jacs.4c10244)
-
 \paragraph{ML for pXRD}
-Several approaches have applied machine learning methods to classification and regression tasks for powder diffractograms.
+To showcase the need for datasets such as the one presented in this publication, we now shortly discuss some recent
+approaches that apply machine learning methods to classification and regression tasks for powder diffractograms. \\
 
 Using simulated diffractograms based on structures from the ICSD, Lee {\it et al.} trained a deep convolutional neural network (CNN) that is able to classify occurring phases in diffractograms of a specific compound pool \cite{Lee2020}.
-They furthermore developed models based on fully convolutional neural networks and transformer encoders that predict the crystal system and the spacegroup and other structural properties, such as the band gap \cite{Lee2022}.
+They furthermore developed models based on fully convolutional neural networks and transformer encoders that predict the crystal system, the spacegroup, and other structural properties, such as the band gap \cite{Lee2022}.
 With their best model for the crystal system prediction on ICSD structures, they achieved a test accuracy of \SI{92.2}{\percent}.
@@
 
-The option of training purely on experimental diffractograms unfortunately is unfeasible because of the limited availability of labeled experimental diffractograms, but Salgado {\it et al.} \cite{Salgado2023} showed, that adding a fraction of experimental patterns to a simulated training dataset improves the performance on unseen experimental patterns.
+The option of training purely on experimental diffractograms unfortunately is unfeasible because of the limited availability of labeled experimental diffractograms, but Salgado {\it et al.} \cite{Salgado2023} showed that adding a fraction of experimental patterns to a simulated training dataset improves the performance on unseen experimental patterns.
 They used \SI{50}{\percent} of the experimental patterns contained in the RRUFF database and added those to their large simulated training set.
@@
 Schuetzke {\it et al.} trained a classifier to classify if a diffractogram stems from an amorphous, single-phase, or multi-phase sample \cite{Schuetzke2024}.
-Due to the lack of experimental pXRDs, they build a pipeline to augment simulated diffractograms by, among other things, slightly varying the underlying crystal lattice.
+Due to the lack of experimental pXRDs, they built a pipeline to augment simulated diffractograms of a reference structure by, among other things, slightly varying the underlying crystal lattice.
+For spinel structures they reported and accuracy of \SI{100}{\percent} but they also proofed that their approach can be transferred to other datasets.\\
 
-In our 2023 paper, we came up with the approach to generate crystal structures and their corresponding pXRD patterns on the fly during the training process \cite{Schopmans2023}.
+In our 2023 paper, we developed an approach to generate synthetic crystal structures and their corresponding pXRD patterns on the fly during the training process \cite{Schopmans2023}.
 This approach defeats the issue of
 a limited dataset size, which limits the depth of neural networks that can be trained.
-However, when we applied our space group classification model to experimental patterns from the RRUFF database the accuracy dropped substantially.
-Augmenting our simulated patterns with background, noise, and impurities would help to bring simulated diffractograms closer to experimental ones, making models trained on them more performant on experimental diffractograms.
-For this augmentation, background and noise statistics from a broader experimental pXRD database would be very valuable.
+However, the accuracy dropped substantially when we applied our space group classification model to experimental patterns from the RRUFF database.
+Augmenting our simulated patterns with background, noise, and impurities helps to bring simulated diffractograms closer to experimental ones, making models trained on them more performant on experimental diffractograms.
+However, this augmentation process could be improved by incorporating background and noise statistics from a broader experimental pXRD database.\\
 
-TODO: briefly discuss here that there are expert models for specific classes of materials which give more information, and more generic models like ours which currently only gives you the space group and not much more; especially focus on how experimental conditions (background, noise, ...
+%TODO: briefly discuss here that there are expert models for specific classes of materials which give more information, and more generic models like ours which currently only gives you the space group and not much more; especially focus on how experimental conditions (background, noise, ...
 is modeled)
-Phase Labeling and Lattice Refinement: https://arxiv.org/abs/2308.07897
-crystal-structure phase mapping: https://www.nature.com/articles/s42256-021-00384-1
+%Phase Labeling and Lattice Refinement: https://arxiv.org/abs/2308.07897
+%crystal-structure phase mapping: https://www.nature.com/articles/s42256-021-00384-1
 
-Physically-informed Graph-based DRNet (PG-DRNet) 2023: \url{https://ieeexplore.ieee.org/abstract/document/10459774/}
+%Physically-informed Graph-based DRNet (PG-DRNet) 2023: \url{https://ieeexplore.ieee.org/abstract/document/10459774/}
 
-The ultimate goal of ML efforts is to make PXRD data identification practical for experimental use, whether for identifying space group symmetry or phase composition.
+% -  Crystal Structure Determination from Powder Diffraction Patterns with Generative Machine Learning: (https://pubs.acs.org/doi/10.1021/jacs.4c10244)
+
+The ultimate goal of ML efforts is to make PXRD data identification practical for experimental use, automating structure property prediction.
 However, the current volume of experimental PXRD patterns is insufficient to effectively train ML models, highlighting an urgent need for a comprehensive experimental PXRD database.
 To address this, two key approaches are essential.
-First, developing more sophisticated simulation methods to better approximate experimental patterns\cite{cao2024simxrd}.
+First, developing more sophisticated simulation methods to better approximate experimental patterns\cite{cao2024simxrd} by using statistics from experimental diffractograms.
 Second, creating an experimental database that enables transfer learning to bridge the gap between simulated and real-world data.
-The development of opXRD is particularly significant, as it will provide a comprehensive experimental benchmark for the community, allowing fair comparison of baseline models and accurate evaluation of their applicability in real experimental situations.
+For both of these steps, the development of opXRD is particularly significant, as it will provide a comprehensive experimental benchmark for the community, allowing fair comparison of baseline models and accurate evaluation of their applicability in real experimental situations.

### 2024-12-03 chapters/04_our_dataset.tex
Updated introductory paragraphs on rietveld refinement

@@
 In collaboration with several other institutions we have collected a dataset of diffractograms stemming from experiments, some of them labeled with corresponding structural information.
-The following texts detail the contribution of each research group.
+The following paragraphs detail the contribution of each research group.
 Most data was collected using \ce{Cu} radiaton sources which have a  $K_{\alpha1}$ wavelength of $\lambda=1.54056\text{\AA}$ and a $K_{\alpha2}$ wavelength of $\lambda=1.54439\text{\AA}$.
@@
 
+%TODO: A paragraph on how data is correlated
+
 \begin{figure*}[!ht]
@@
 The pXRD data from the COD database are of high quality, with a median resolution of $\Delta(2\theta) = 0.013\degree$ and an average number of 9190 points measured per pattern.
-They span a wide chemical space, including organic, inorganic, and hybrid structures, including 75 different elements of the periodic table.
+They span a wide chemical space, including organic, inorganic, and hybrid structures, and 75 different elements of the periodic table.
 

### 2024-12-03 chapters/01_introduction.tex
Described python + zenodo data availability

@@
 Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting lattice parameters\cite{Dong2021, Chitturi2021, Habershon2004, zhang2024crystallographic}, spacegroup \cite{Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021,zhang2024crystallographic}, and crystallite size \cite{Dong2021, Chakraborty2021} from simulated diffractograms.
-However, the performance substantially drops off when these models are applied to data originating from experiments \cite{Schopmans2023,zhang2024crystallographic}[cite 2-3 papers, incl. ours].
+However, the performance substantially drops off when these models are applied to data originating from experiments \cite{Schopmans2023,zhang2024crystallographic, Wang2020}
 The patterns observed in experiments exhibit imperfections that do not appear when simulating the X-ray diffraction pattern of a powder sample under ideal conditions.
@@
 The most advanced ML models currently are trained on approximately $10^7 - 10^8$ simulated diffractograms,
-%For example 4M sim patterns\cite{cao2024simxrd}
+%For example 4M sim patterns\cite{cao2024simxrd} = 4 Mio. patterns
 which is - to the best of our knowledge - three orders of magnitude larger than the largest currently curated experimental dataset (PDF-5+ with approximately 20,000 experimental patterns) and even x orders of magnitude larger than the number of unlabeled diffractograms in the opXRD dataset we present here.
@@
 \cite{Waseda2011, Pecharsky2023}.
-Some experimental setups may also exhibit imperfections such as displacement of the sample from its intended position leading to the recorded scattering angles being slightly erroneous \cite{hulbert2023} or background events originating from instrument noise or X-rays scattering off equipment onto the screen.\\
+Some experimental setups may also exhibit imperfections such as displacement of the sample from its intended position leading to the recorded scattering angles being slightly off \cite{hulbert2023} or background events such as instrument noise or X-rays scattering off equipment onto the screen.\\
 
@@
 Our opXRD dataset has been curated by collecting the accumulated powder data from multiple large research groups and institutions with high-throughput XRD facilities.
-It contains pXRD patterns from single and multiphase materials from a wide variety of materials classes, including XXX, YYY, and ZZZ.
+It contains pXRD patterns from single and multiphase materials from a wide variety of material classes, including
+high-entropy materials, perovskites, and commercial catalysts.
 Details about the experiments of contributing research groups are discussed in section $\eqref{sec:our_dataset}$, along with a detailed description of how to contribute further data and how to use the data.

### 2024-12-03 chapters/02_SOTA_ML.tex
Described python + zenodo data availability

--- 2024-12-03 chapters/03_SOTA_ML.tex
@@
-\paragraph{ML for pXRD}
+%\paragraph{ML for pXRD}
 To showcase the need for datasets such as the one presented in this publication, we now shortly discuss some recent
@@
 Due to the lack of experimental pXRDs, they built a pipeline to augment simulated diffractograms of a reference structure by, among other things, slightly varying the underlying crystal lattice.
-For spinel structures they reported and accuracy of \SI{100}{\percent} but they also proofed that their approach can be transferred to other datasets.\\
+For spinel structures, they reported an accuracy of \SI{100}{\percent} but they also proved that their approach can be transferred to other datasets.\\
 
@@
 
-%TODO: briefly discuss here that there are expert models for specific classes of materials which give more information, and more generic models like ours which currently only gives you the space group and not much more; especially focus on how experimental conditions (background, noise, ...
-is modeled)
-%Phase Labeling and Lattice Refinement: https://arxiv.org/abs/2308.07897
-%crystal-structure phase mapping: https://www.nature.com/articles/s42256-021-00384-1
-
-%Physically-informed Graph-based DRNet (PG-DRNet) 2023: \url{https://ieeexplore.ieee.org/abstract/document/10459774/}
-
-% -  Crystal Structure Determination from Powder Diffraction Patterns with Generative Machine Learning: (https://pubs.acs.org/doi/10.1021/jacs.4c10244)
+It becomes apparent, that the more general the task is, the more challenging the transfer to experimental data is.
+For example, the space group classification task is very general, for all material systems.
+Therefore, transferring it to the application on experimental diffraction patterns is difficult.
+\cite{Schopmans2023, Lee2022, Park2017} On the other hand, there are some successful approaches that also work well on experimental data, but those are mostly methods that do phase determination in a limited compound space, making the task less complex.
+\cite{Schuetzke2024, Lee2020}
 

### 2024-12-03 chapters/03_existing_datasets.tex
Described python + zenodo data availability

--- 2024-12-03 chapters/02_existing_datasets.tex
@@
 As of November 2024, the PDF5+ is available to researchers through a purchase of a one-year license starting at a price point of \$6265.00.
-However, the ICDD does not allow researchers to train machine learning models on PDF5+ data, regardless of whether the resulting models are published.
-\\
+However, the ICDD does not allow researchers to train machine learning models on PDF5+ data, regardless of whether the resulting models are published.\footnote{This information can be found in the product license agreement for the PDF5+ at \url{https://www.icdd.com/licensing-process/\#1528471154226-933e5cc6-8da7}.} \\
 
@@
 Since the PDF5+ lists the MPDS as a database source, there is likely significant overlap between the experimental patterns found in the PDF5+ and those found in Pearson's crystal data.
-As of November 2024, Pearson's crystal data is available to researchers through a purchase of a one-year license starting at a price point of 2200 \euro.\\
+As of November 2024, Pearson's crystal data is available to researchers through a purchase of a one-year license starting at a price point of \qty{2200}{\myeuro}.\\
 
@@
 \cite{lafuente2015} Managed by the University of Arizona, it was created to serve as a public repository for mineral identification and research.
-It contains \num{1290} powder diffraction patterns stemming from experiments each labeled with the lattice parameters and composition of the underlying structure.
+It contains \num{1290} powder diffraction patterns stemming from experiments each labeled with the lattice parameters and composition of the underlying structures.
 \\

### 2024-12-03 chapters/05_how_to_use.tex
Described python + zenodo data availability

@@
-Next to the availability of the opXRD dataset on Zenodo, we further provide an easy-to-use Python library (\url{https://github.com/aimat-lab/xrdpattern}) to download and interface the dataset.
-The dataset can be accessed using the simple command 
-\pyth{PatternDB.load(data_dirpath)}. Our tool further includes options for standardization, plotting, and the conversion to \emph{PyTorch} tensors.
+The opXRD database is hosted on Zenodo\footnote{The dataset can be found at \url{https://zenodo.org/records/14254271}.} and can be downloaded by any user without any barriers or restrictions.
+The database comes in two files, \"opxrd.zip\" and \"opxrd\_in\_situ.zip\".
+The latter contains the in-situ data which contains highly correlated patterns recorded through time series measurements.
+Patterns belonging to time series measurements have filenames that indicate the measurement series they belong to as well as their order in that series.
+\\
+
+Next to the availability of the opXRD dataset on Zenodo, we also provide a Python library \"opxrd\" to easily download and interface with the dataset.
+The instructions for how to install this library can be found in the repository associated with the library.\footnote{The repository to this library can is located at \url{https://github.com/aimat-lab/opxrd}.} The opxrd library allows the dataset to be accessed using one simple command \pyth{OpXRD.load(root_dirpath)}.
+If the database is locally available under \pyth{root_dirpath} this command loads the library from this location.
+If the database is not available locally at this location, the database is automatically downloaded from Zenodo to \pyth{root_dirpath}.
+\\
+
+Our tool further includes options for standardization, plotting, and the conversion to \emph{PyTorch} tensors.
 We provide the following Jupyter notebook that showcases the functionality of the library in more detail: <Link to github Jupyter notebook on Colab>

### 2024-12-04 chapters/01_introduction.tex
Merged Xqueyer paragraph into contribution entry

@@
 If, during the iterative determination of parameters, the structural model becomes unphysical the operator must intervene to backtrack to a previous successful step or start over entirely.
-Finding the correct order in which to refine structure parameters and finding the correct values for fixed parameters both present problems that add to the difficulty of the refinement process.
-
-%What makes the refinement process time-consuming and difficult is finding the correct order in which to refine the crystal structure parameters\cite{Ozaki2020}. \\ % \\
+Finding the correct order in which to refine structure parameters and finding the correct values for fixed parameters both present problems that add to the difficulty of the refinement process.\\
 
@@
 Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting lattice parameters\cite{Dong2021, Chitturi2021, Habershon2004, zhang2024crystallographic}, spacegroup \cite{Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021,zhang2024crystallographic}, and crystallite size \cite{Dong2021, Chakraborty2021} from simulated diffractograms.
-However, the performance substantially drops off when these models are applied to data originating from experiments \cite{Schopmans2023,zhang2024crystallographic, Wang2020}
-The patterns observed in experiments exhibit imperfections that do not appear when simulating the X-ray diffraction pattern of a powder sample under ideal conditions.
+However, the performance substantially drops off when these models are applied to data originating from experiments \cite{Schopmans2023,zhang2024crystallographic, Wang2020, Vecsei2018} The patterns observed in experiments exhibit imperfections that do not appear when simulating the X-ray diffraction pattern of a powder sample under ideal conditions.
 The physical effects underlying these imperfections need to be modeled accurately in order to be able to transfer models trained on simulated data to experimental patterns.
@@
 Collecting a labeled dataset of raw experimental powder X-ray diffraction data of sufficient size to directly train a model that predicts crystal structure information is currently out of reach.
-The most advanced ML models currently are trained on approximately $10^7 - 10^8$ simulated diffractograms,
-%For example 4M sim patterns\cite{cao2024simxrd} = 4 Mio. patterns
-which is - to the best of our knowledge - three orders of magnitude larger than the largest currently curated experimental dataset (PDF-5+ with approximately 20,000 experimental patterns) and even x orders of magnitude larger than the number of unlabeled diffractograms in the opXRD dataset we present here.
+The most advanced ML models currently are trained on approximately $10^5 - 10^6$ simulated diffractograms.
+\cite{Salgado2023, Schopmans2023}
+This is - to the best of our knowledge - two orders of magnitude larger than the largest currently curated experimental dataset (PDF-5+ with approximately 20,000 experimental patterns) and even x orders of magnitude larger than the number of unlabeled diffractograms in the opXRD dataset we present here.
 \\

### 2024-12-04 chapters/02_SOTA_ML.tex
Merged Xqueyer paragraph into contribution entry

@@
 Therefore, transferring it to the application on experimental diffraction patterns is difficult.
-\cite{Schopmans2023, Lee2022, Park2017} On the other hand, there are some successful approaches that also work well on experimental data, but those are mostly methods that do phase determination in a limited compound space, making the task less complex.
+\cite{Schopmans2023, Lee2022, Vecsei2018} On the other hand, there are some successful approaches that also work well on experimental data, but those are mostly methods that do phase determination in a limited compound space, making the task less complex.
 \cite{Schuetzke2024, Lee2020}
 
-The ultimate goal of ML efforts is to make PXRD data identification practical for experimental use, automating structure property prediction.
-However, the current volume of experimental PXRD patterns is insufficient to effectively train ML models, highlighting an urgent need for a comprehensive experimental PXRD database.
+The ultimate goal of ML efforts is to make pXRD data identification practical for experimental use, automating structure property prediction.
+However, the current volume of experimental pXRD patterns is insufficient to effectively train ML models, highlighting an urgent need for a comprehensive experimental pXRD database.
 To address this, two key approaches are essential.

### 2024-12-04 chapters/04_our_dataset.tex
Described python + zenodo data availability
Merged Xqueyer paragraph into contribution entry

@@
 
-We extracted experimental pXRD data from the Crystallography Open Database (COD)\cite{Grazulis2009, Vaitkus2023}.
+Experimental pXRD data was extracted from the Crystallography Open Database (COD)\cite{Grazulis2009, Vaitkus2023}.
+
+%We extracted experimental pXRD data from the Crystallography Open Database .
+
 The COD is, to our knowledge, the largest open-access collection of experimental crystal structures of organic, inorganic, and metal-organic compounds and minerals, containing more than 500,000 entries.
@@
 
-In the past two years, we have established a small-scale experimental powder X-ray diffraction database called the X-Ray Phase Identification Public Experimental Dataset (XRed).
-\footnote{XRed can be accessed on GitHub at \url{https://github.com/WPEM/XRED}.} The primary objective of this database is to support the development of intelligent phase identification technology, serving as a foundation for data collection for future large-scale machine learning calculations.
+A small-scale experimental powder X-ray database called the X-Ray Phase Identification Public Experimental Dataset (XRed) was established over the past two years.\footnote{XRed can be accessed on GitHub at \url{https://github.com/WPEM/XRED}.} The primary objective of this database is to support the development of intelligent phase identification technology, serving as a foundation for data collection for future large-scale machine learning calculations.
 \\
@@
 The dataset is organized by elemental systems and includes original experimental files ranging from single-phase to five-phase mixtures, as well as mixtures studied for different tasks.
-This project is progressing toward integration with the comprehensive opXRD database to create a large-scale, long-term experimental resource.
+\\
+
+Additionally, Xqueryer,\footnote{Xqueryer can be accessed at \url{http://xqueryer.caobin.asia/}.} a cutting-edge free online platform for intelligent structure identification has been launched.
+Users can upload experimental pXRD data to identify crystal structures and extract all relevant information available in the Materials Project database.
+Uploaded data are securely stored and integrated into the comprehensive opXRD database to progress toward a large-scale, long-term experimental resource.
+All usage complies with Xqueryer's terms of service.
 

### 2024-12-04 chapters/05_how_to_use.tex
Merged Xqueyer paragraph into contribution entry

@@
 The opXRD database is hosted on Zenodo\footnote{The dataset can be found at \url{https://zenodo.org/records/14254271}.} and can be downloaded by any user without any barriers or restrictions.
-The database comes in two files, \"opxrd.zip\" and \"opxrd\_in\_situ.zip\".
+The database comes in two files, ``opxrd.zip'' and ``opxrd\_in\_situ.zip''.
 The latter contains the in-situ data which contains highly correlated patterns recorded through time series measurements.
@@
 Next to the availability of the opXRD dataset on Zenodo, we also provide a Python library \"opxrd\" to easily download and interface with the dataset.
-The instructions for how to install this library can be found in the repository associated with the library.\footnote{The repository to this library can is located at \url{https://github.com/aimat-lab/opxrd}.} The opxrd library allows the dataset to be accessed using one simple command \pyth{OpXRD.load(root_dirpath)}.
+The instructions for how to install this library can be found in the repository associated with the library.\footnote{The repository to this library can is located at \url{https://github.com/aimat-lab/opxrd}.} The opxrd library allows the dataset to be accessed through one simple command: \pyth{OpXRD.load(root_dirpath)}.
 If the database is locally available under \pyth{root_dirpath} this command loads the library from this location.
-If the database is not available locally at this location, the database is automatically downloaded from Zenodo to \pyth{root_dirpath}.
+If the database is not available locally at this location, the database is automatically downloaded to \pyth{root_dirpath}.
 \\

### 2024-12-04 chapters/03_existing_datasets.tex
Commited table version with footnote directly underneath
Added labeled column in dataset overview

@@
 \toprule
-\textbf{Name}                          & \textbf{No. of Patterns} & \textbf{Year Established} \\
+\textbf{Name}                          & \textbf{No. of Patterns} & \textbf{Year established} \\
 \midrule
-Powder Diffraction File (PDF)  \tablefootnote{The PDF lists the Materials Phases Data System (MPDS) as a database source.
-The MPDS and Pearson's crystal, both maintained by the Pauling File project, contain largely identical data.
-Hence it is likely that there is significant overlap in the experimental patterns available in the PDF and Pearson's crystal data.}         & 20,800                  & 1941                      \\
-Pearson's Crystal Data                 & 21,700                  & 2007        \\
+Linus Pauling file                 & 21,700                  & 2002        \\
+Powder Diffraction File (PDF)  \tablefootnote{The PDF lists the Material Platform for Data Science (MPDS) as a database source.
+Since the MPDS is hosted by the Pauling File project, there is likely significant overlap in the experimental patterns available in the PDF and the Linus Pauling File.}         & 20,800                  & 1941                      \\
+RRUFF                                  & 1290                    & 2006                      \\
 Crystallography Open Database (COD)    & 1063                    & 2003                      \\
-RRUFF                                  & 1290                    & 2006                      \\
 PowBase                                & 169                     & 1999                      \\
@@
 
+\textbf{Linus Pauling File}:\cite{PaulingWeb} The Linus Pauling File is a largely commercial crystal structure database published and maintained by the Pauling File project \cite{villars2018}.
+It is currently distributed as Pearson Crystal data \cite{PearsonWeb} and the Materials Platform for Data Science (MPDS)\cite{MPDSWeb}.
+The database, first published in 2002, currently contains more than 534,000 crystal structures\cite{MPDSWeb} and 21,700 corresponding experimental powder diffraction patterns\cite{PearsonWeb}.
+This makes the Pauling file, to the best of our knowledge, the largest collection of experimental powder diffraction data available to researchers.
+As of November 2024, Pearson's crystal data is available to researchers through a purchase of a one-year license starting at a price point of \qty{2200}{\myeuro}.
+The MPDS is partially open, with the open MPDS data accessible through a web interface\cite{MPDSWeb}.
+Access to the full MPDS database can be purchased through a one-year license starting at \qty{2000}.
+\\
+
 \textbf{Powder Diffraction File:} \cite{PDFWeb} The Powder Diffraction File (PDF), published and maintained by the International Center for Diffraction Data (ICDD), is a large collection of materials with accompanying powder diffraction data first published in 1941\cite{GatesRector2019}.
@@
 After inquiring with the ICDD in April 2024 we were told that only 20,800 of the powder diffraction patterns in the PDF5+ stemmed from experiments.
-Of these 20,800 entries, 10,954 contain information about the atomic coordinates of the underlying structures.%\footnote{We reached out to the contact address given on their website, info@icdd.com.}
-Still, this makes the PDF5+, to the best of our knowledge, the largest collection of experimental powder diffraction data available to researchers.
+Of these 20,800 entries, 10,954 contain information about the atomic coordinates of the underlying structures.
+Since the PDF5+ lists the MPDS as a database source, there is likely a significant overlap in the experimental patterns found in the PDF5+ and those found in the Pauling file.
 As of November 2024, the PDF5+ is available to researchers through a purchase of a one-year license starting at a price point of \$6265.00.
@@
 
-\textbf{Pearson's Crystal data}:\cite{PearsonWeb} Pearson's crystal data is a commercial database published and maintained by the Pauling File project \cite{kaduk2007}.
-The database, first published in 2007, currently contains 395,000 crystal structures and 21,700 corresponding experimental powder diffraction patterns.
-The Pauling File project also distributes the Material phases data system, which contains largely identical data.
-Since the PDF5+ lists the MPDS as a database source, there is likely significant overlap between the experimental patterns found in the PDF5+ and those found in Pearson's crystal data.
-As of November 2024, Pearson's crystal data is available to researchers through a purchase of a one-year license starting at a price point of \qty{2200}{\myeuro}.\\
+\textbf{RRUFF}: \cite{RRUFFWeb} The RRUFF Mineral Database, first published in 2006, provides detailed information on minerals, including their chemical compositions, crystallography, and spectroscopic data.
+\cite{lafuente2015} Managed by the University of Arizona, it was created to serve as a public repository for mineral identification and research.
+It contains \num{1290} powder diffraction patterns stemming from experiments each labeled with the lattice parameters and composition of the underlying structures.
+\\
 
@@
 Hence, the experimental powder diffraction data contained in the COD is labeled with the full crystal structure information.
-\\
-
-\textbf{RRUFF}: \cite{RRUFFWeb} The RRUFF Mineral Database, first published in 2006, provides detailed information on minerals, including their chemical compositions, crystallography, and spectroscopic data.
-\cite{lafuente2015} Managed by the University of Arizona, it was created to serve as a public repository for mineral identification and research.
-It contains \num{1290} powder diffraction patterns stemming from experiments each labeled with the lattice parameters and composition of the underlying structures.
 \\

### 2024-12-04 chapters/04_our_dataset.tex
Added labeled column in dataset overview

@@
 \centering
-\caption{Labeled opXRD database entries}
-\label{tab:labeled}
-\begin{tabular}{@{}ccc@{}} % Adjusted columns
+\caption{Merged opXRD database entries}
+\label{tab:merged}
+\begin{tabular}{@{}cccc@{}}
 \toprule
-Institution & Research project & No. patterns\footnote{This number is subject to change} \\
+Institution & Research Project & No. Patterns & Available labels \\ 
 \midrule
-CNRS        & Extracted from the Crystallography open database & 895 \\
-USC         & Study of CuNi and CuAl alloys & 338 \\
-HKUST(GZ)   & Reference database for phase identification & 21 \\
-EMPA        & Tin halide perovskites and combinatorial Zn-V-N library & 790 \\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\caption{Unlabeled opXRD database entries}
-\label{tab:unlabeled}
-\begin{tabular}{@{}ccc@{}} % Adjusted columns
-\toprule
-Institution & Research project & No. patterns\footnote{This number is subject to change} \\
-\midrule
-INT         & Compilation of various research projects & 19780 \\
-IKFT        & Compilation of various research projects & 64 \\
-LBNL        & Time series recording of in-situ measurements & n/a \\
+CNRS        & Extracted from the COD         & 895          & Spg, Lattice, Unit cell, Comp   \\ 
+HKUST(GZ)   & Phase identification reference database              & 21           & Spg, Lattice, Unit cell, Comp   \\ 
+USC         & Study of CuNi and CuAl alloys                            & 338          & Spg, Lattice, Comp   \\ 
+EMPA        & Tin halide perovskites, Zn-V-N library  & 790          & Spg, Comp   \\ 
+INT         & Compilation of various projects                 & 19780        & None \\ 
+IKFT        & Commercial catalysts, bulk materials                 & 64           & None \\ 
+LBNL        & Time series recordings of Perovskites            & n/a          & None \\ 
 \bottomrule
@@
 Experimental pXRD data was extracted from the Crystallography Open Database (COD)\cite{Grazulis2009, Vaitkus2023}.
-
-%We extracted experimental pXRD data from the Crystallography Open Database .
-
 The COD is, to our knowledge, the largest open-access collection of experimental crystal structures of organic, inorganic, and metal-organic compounds and minerals, containing more than 500,000 entries.
@@
 They span a wide chemical space, including organic, inorganic, and hybrid structures, and 75 different elements of the periodic table.
+
+% Paragraph by Bin Cao and Tong-yi Zhang
+\subsubsection*{Guangzhou Municipal Key Laboratory of Materials Informatics, HKUST(GZ)}
+
+%In the past two years, we established a small-scale experimental PXRD database called the X-Ray Phase Identification Public Experimental Dataset.
+The data in XRed is generated in our lab using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers under Cu targets.
+Hundreds of PXRD patterns have been refined, labeled with CIF files, and organized by elemental systems.
+The dataset includes original experimental files across single-phase to five-phase mixtures.
+This project is progressing toward integration with the comprehensive opXRD database to establish a large-scale, long-term experimental resource.
+
+A small-scale experimental powder X-ray database called the X-Ray Phase Identification Public Experimental Dataset (XRed) was established over the past two years.\footnote{XRed can be accessed on GitHub at \url{https://github.com/WPEM/XRED}.} The primary objective of this database is to support the development of intelligent phase identification technology, serving as a foundation for data collection for future large-scale machine learning calculations.
+\\
+
+The data in XRed were generated in our laboratory using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers, all utilizing \ce{Cu} X-ray sources.
+Hundreds of pXRD patterns have been obtained through XRD refinement, enabling the determination of detailed lattice constants, atomic sizes, and other structural parameters.
+These patterns are labeled with CIF files that document the refined structures.
+The dataset is organized by elemental systems and includes original experimental files ranging from single-phase to five-phase mixtures, as well as mixtures studied for different tasks.
+\\
+
+Additionally, Xqueryer,\footnote{Xqueryer can be accessed at \url{http://xqueryer.caobin.asia/}.} a cutting-edge free online platform for intelligent structure identification has been launched.
+Users can upload experimental pXRD data to identify crystal structures and extract all relevant information available in the Materials Project database.
+Uploaded data are securely stored and integrated into the comprehensive opXRD database to progress toward a large-scale, long-term experimental resource.
+All usage complies with Xqueryer's terms of service.
+
+% Paragraph by Adie Alwen and Andrea Hodge
+\subsubsection*{Department of Chemical Engineering and Materials Science, USC}
+
+Combinatorial \ce{CuNi} and \ce{CuAl} samples were synthesized at the University of Southern California (USC) via magnetron co-sputtering from \ce{Cu} (99.999\%) and \ce{Ni} (99.995\%) or \ce{Al} (99.999\%) targets (Plasmaterials) onto stationary $10 \si{cm}$ \ce{Si} (100) substrates yielding 169 unique $5 \times 5 \ \si{mm}$ squares per wafer.
+These alloys were selected to study compositional effects on stacking fault and growth twin boundary formation in sputtered materials \cite{2024AcMat.27019839A,alwen2024combinatorial}.
+\\
+
+XRD spectra were collected using a Malvern Panalytical Empyrean X-ray diffractometer at the Forschungszentrum Julich (FZJ) Institute for Energy and Climate Research, Structure and Function of Materials (IEK-2), which enabled automated XRD characterization due to its programmable XY stage and laser sensor for Z height correction.
+XRD analysis was used to link changes in defect formation in the sputtered films with varied phases and crystal structures.
+Each composition was characterized using \ce{Cu} $K_\alpha$ X-rays that were collimated to probe an area of $4 \times 4 \ \si{mm}$ using a step size of $0.026\degree$ and scan duration of $0.3$ seconds per step over a 2 range from $30\degree - 110\degree$.
+Measurement parameters were chosen to resolve at least six or more points above the full-width half maximum of each XRD peak.
+
+% Paragraph by Alexander Wieczorek and Sebastiann Siol
+\subsubsection*{Laboratory for Surface Science and Coating Technologies, Empa}
+
+Each combinatorial Zn–V–N library was synthesized using radio-frequency co-sputtering of Zn and V in a mixed Ar and \ce{N2} plasma.
+An orthogonal deposition temperature and composition gradient was created, resulting in a deposition temperature of $220\degree$C for samples 1~–~9 and $114\degree$C for samples 37~–~45.
+The composition for each sample was determined using X-ray fluorescence (XRF) spectroscopy which was further calibrated through Rutherford backscattering spectroscopy (RBS) based on selected samples.
+The newly identified and isolated semiconductor \ce{Zn2VN3} was identified to exhibit a cation-disordered wurtzite structure as verified by additional GI-XRD and SAED measurements\cite{Zhuk2021}.
+\\
+Tin halide perovskites were deposited using single-step spin-coating as reported elsewhere\cite{Wieczorek2023}.
+Methylammonium lead iodide libraries with varying degrees of residual \ce{PbI2} were deposited using a two-step procedure involving both thermal evaporation of \ce{PbI2} and subsequent spin-coating of a methylammonium solution.
+The relative phase fractions were quantified using supplementary azimuthal angle scans coupled with structural factors and geometrical factors as reported elsewhere\cite{Wieczorek2024}.
+Fully inorganic lead perovskite libraries were prepared using thermal co-evaporation of lead and cesium halide salts.
+All metal halide perovskite libraries were measured within a custom-made X-Ray transparent inert-gas dome, resulting in the presence of minor additional features within the $\theta=19 \text{–}31\degree$ range.
+For all combinatorial libraries where any phases are specified, the complete set of phases is reported in the metadata.
+\\
+
+XRD data was measured using a Bruker D8 Discover equipped with a Cu radiation source in a Bragg-Brentano geometry.
+For the reported data sets the instrument was equipped with a Goebel mirror effectively removing the Cu~K$\beta$ radiation.
+The data set originates from the combinatorial exploration of the Zn–V–N compositional space, as well as data gathered from multiple research activities on more established metal halide perovskite semiconductors.
+All data was collected from thin films deposited on borosilicate glass.
+The Zn–V–N films showed some preferential out-of-plane orientation, while for the perovskites the preferential orientation was minimal, resulting in the presence of all reflections.
+\\
 
@@
 
-% Paragraph by Adie Alwen and Andrea Hodge
-\subsubsection*{Department of Chemical Engineering and Materials Science, USC}
-
-Combinatorial \ce{CuNi} and \ce{CuAl} samples were synthesized at the University of Southern California (USC) via magnetron co-sputtering from \ce{Cu} (99.999\%) and \ce{Ni} (99.995\%) or \ce{Al} (99.999\%) targets (Plasmaterials) onto stationary $10 \si{cm}$ \ce{Si} (100) substrates yielding 169 unique $5 \times 5 \ \si{mm}$ squares per wafer.
-These alloys were selected to study compositional effects on stacking fault and growth twin boundary formation in sputtered materials \cite{2024AcMat.27019839A,alwen2024combinatorial}.
-\\
-
-XRD spectra were collected using a Malvern Panalytical Empyrean X-ray diffractometer at the Forschungszentrum Julich (FZJ) Institute for Energy and Climate Research, Structure and Function of Materials (IEK-2), which enabled automated XRD characterization due to its programmable XY stage and laser sensor for Z height correction.
-XRD analysis was used to link changes in defect formation in the sputtered films with varied phases and crystal structures.
-Each composition was characterized using \ce{Cu} $K_\alpha$ X-rays that were collimated to probe an area of $4 \times 4 \ \si{mm}$ using a step size of $0.026\degree$ and scan duration of $0.3$ seconds per step over a 2 range from $30\degree - 110\degree$.
-Measurement parameters were chosen to resolve at least six or more points above the full-width half maximum of each XRD peak.
-
 % Paragraph by Tim Kodalle
@@
 \\
-
-% Paragraph by Bin Cao and Tong-yi Zhang
-\subsubsection*{Guangzhou Municipal Key Laboratory of Materials Informatics, HKUST(GZ)}
-
-%In the past two years, we established a small-scale experimental PXRD database called the X-Ray Phase Identification Public Experimental Dataset.
-The data in XRed is generated in our lab using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers under Cu targets.
-Hundreds of PXRD patterns have been refined, labeled with CIF files, and organized by elemental systems.
-The dataset includes original experimental files across single-phase to five-phase mixtures.
-This project is progressing toward integration with the comprehensive opXRD database to establish a large-scale, long-term experimental resource.
-
-A small-scale experimental powder X-ray database called the X-Ray Phase Identification Public Experimental Dataset (XRed) was established over the past two years.\footnote{XRed can be accessed on GitHub at \url{https://github.com/WPEM/XRED}.} The primary objective of this database is to support the development of intelligent phase identification technology, serving as a foundation for data collection for future large-scale machine learning calculations.
-\\
-
-The data in XRed were generated in our laboratory using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers, all utilizing \ce{Cu} X-ray sources.
-Hundreds of pXRD patterns have been obtained through XRD refinement, enabling the determination of detailed lattice constants, atomic sizes, and other structural parameters.
-These patterns are labeled with CIF files that document the refined structures.
-The dataset is organized by elemental systems and includes original experimental files ranging from single-phase to five-phase mixtures, as well as mixtures studied for different tasks.
-\\
-
-Additionally, Xqueryer,\footnote{Xqueryer can be accessed at \url{http://xqueryer.caobin.asia/}.} a cutting-edge free online platform for intelligent structure identification has been launched.
-Users can upload experimental pXRD data to identify crystal structures and extract all relevant information available in the Materials Project database.
-Uploaded data are securely stored and integrated into the comprehensive opXRD database to progress toward a large-scale, long-term experimental resource.
-All usage complies with Xqueryer's terms of service.
-
-% Paragraph by Alexander Wieczorek and Sebastiann Siol
-\subsubsection*{Laboratory for Surface Science and Coating Technologies, Empa}
-
-Each combinatorial Zn–V–N library was synthesized using radio-frequency co-sputtering of Zn and V in a mixed Ar and \ce{N2} plasma.
-An orthogonal deposition temperature and composition gradient was created, resulting in a deposition temperature of $220\degree$C for samples 1~–~9 and $114\degree$C for samples 37~–~45.
-The composition for each sample was determined using X-ray fluorescence (XRF) spectroscopy which was further calibrated through Rutherford backscattering spectroscopy (RBS) based on selected samples.
-The newly identified and isolated semiconductor \ce{Zn2VN3} was identified to exhibit a cation-disordered wurtzite structure as verified by additional GI-XRD and SAED measurements\cite{Zhuk2021}.
-\\
-Tin halide perovskites were deposited using single-step spin-coating as reported elsewhere\cite{Wieczorek2023}.
-Methylammonium lead iodide libraries with varying degrees of residual \ce{PbI2} were deposited using a two-step procedure involving both thermal evaporation of \ce{PbI2} and subsequent spin-coating of a methylammonium solution.
-The relative phase fractions were quantified using supplementary azimuthal angle scans coupled with structural factors and geometrical factors as reported elsewhere\cite{Wieczorek2024}.
-Fully inorganic lead perovskite libraries were prepared using thermal co-evaporation of lead and cesium halide salts.
-All metal halide perovskite libraries were measured within a custom-made X-Ray transparent inert-gas dome, resulting in the presence of minor additional features within the $\theta=19 \text{–}31\degree$ range.
-For all combinatorial libraries where any phases are specified, the complete set of phases is reported in the metadata.
-\\
-
-XRD data was measured using a Bruker D8 Discover equipped with a Cu radiation source in a Bragg-Brentano geometry.
-For the reported data sets the instrument was equipped with a Goebel mirror effectively removing the Cu~K$\beta$ radiation.
-The data set originates from the combinatorial exploration of the Zn–V–N compositional space, as well as data gathered from multiple research activities on more established metal halide perovskite semiconductors.
-All data was collected from thin films deposited on borosilicate glass.
-The Zn–V–N films showed some preferential out-of-plane orientation, while for the perovskites the preferential orientation was minimal, resulting in the presence of all reflections.
-\\

### 2024-12-04 chapters/04_our_dataset.tex
Added labeled columns per feature in database overview

@@
 \centering
-\caption{Merged opXRD database entries}
+\caption{opXRD database entries}
 \label{tab:merged}
-\begin{tabular}{@{}cccc@{}}
+\begin{tabular}{@{}clccccl@{}}
 \toprule
-Institution & Research Project & No. Patterns & Available labels \\ 
+Institution & No. Patterns & Spg & Lattice & Unit Cell & Comp. & Research Project \\ 
 \midrule
-CNRS        & Extracted from the COD         & 895          & Spg, Lattice, Unit cell, Comp   \\ 
-HKUST(GZ)   & Phase identification reference database              & 21           & Spg, Lattice, Unit cell, Comp   \\ 
-USC         & Study of CuNi and CuAl alloys                            & 338          & Spg, Lattice, Comp   \\ 
-EMPA        & Tin halide perovskites, Zn-V-N library  & 790          & Spg, Comp   \\ 
-INT         & Compilation of various projects                 & 19780        & None \\ 
-IKFT        & Commercial catalysts, bulk materials                 & 64           & None \\ 
-LBNL        & Time series recordings of Perovskites            & n/a          & None \\ 
+CNRS        & \qquad 895          & $\times$   & $\times$   & $\times$   & $\times$   & Extracted from the COD                 \\ 
+HKUST(GZ)   & \qquad 21           & $\times$   & $\times$   & $\times$   & $\times$   & Phase identification reference database \\ 
+USC         & \qquad 338          & $\times$   & $\times$   &            & $\times$   & Study of CuNi and CuAl alloys          \\ 
+EMPA        & \qquad 790          & $\times$   &            &            & $\times$   & Tin halide perovskites, Zn-V-N library \\ 
+INT         & \qquad 19780        &            &            &            &            & Compilation of various projects        \\ 
+IKFT        & \qquad 64           &            &            &            &            & Commercial catalysts, bulk materials   \\ 
+LBNL        & \qquad n/a          &            &            &            &            & Time series recordings of Perovskites  \\ 
 \bottomrule

### 2024-12-04 chapters/04_our_dataset.tex
Added overview table with checkmarks idea

@@
 \label{tab:merged}
-\begin{tabular}{@{}clccccl@{}}
+\begin{tabular}{@{}clcccccl@{}}
 \toprule
-Institution & No. Patterns & Spg & Lattice & Unit Cell & Comp. & Research Project \\ 
+Institution & No. Patterns & Comp. & Spg & Lattice & Unit Cell & Research Project \\ 
 \midrule
-CNRS        & \qquad 895          & $\times$   & $\times$   & $\times$   & $\times$   & Extracted from the COD                 \\ 
-HKUST(GZ)   & \qquad 21           & $\times$   & $\times$   & $\times$   & $\times$   & Phase identification reference database \\ 
-USC         & \qquad 338          & $\times$   & $\times$   &            & $\times$   & Study of CuNi and CuAl alloys          \\ 
-EMPA        & \qquad 790          & $\times$   &            &            & $\times$   & Tin halide perovskites, Zn-V-N library \\ 
-INT         & \qquad 19780        &            &            &            &            & Compilation of various projects        \\ 
-IKFT        & \qquad 64           &            &            &            &            & Commercial catalysts, bulk materials   \\ 
-LBNL        & \qquad n/a          &            &            &            &            & Time series recordings of Perovskites  \\ 
+CNRS        & \qquad 895          & \cmark   & \cmark   & \cmark   & \cmark   & Extracted from the COD                 \\ 
+HKUST(GZ)   & \qquad 21           & \cmark   & \cmark   & \cmark   & \cmark   & Phase identification reference database \\ 
+USC         & \qquad 338          & \cmark   & \cmark   & \cmark   & \xmark   & Study of CuNi and CuAl alloys          \\ 
+EMPA        & \qquad 790          & \cmark   & \cmark   & \xmark   & \xmark   & Tin halide perovskites, Zn-V-N library \\ 
+INT         & \qquad 19780        & \xmark   & \xmark   & \xmark   & \xmark   & Compilation of various projects        \\ 
+IKFT        & \qquad 64           & \xmark   & \xmark   & \xmark   & \xmark   & Commercial catalysts, bulk materials   \\ 
+LBNL        & \qquad n/a          & \xmark   & \xmark   & \xmark   & \xmark   & Time series recordings of Perovskites  \\ 
 \bottomrule

### 2024-12-05 chapters/03_existing_datasets.tex
Added histograms, PCA plots + paragraph on notebook; Added labeled column in existing datasets;

@@
 \label{tab:exp_databases}
-\begin{tabular}{@{}lcc@{}}
+\scalebox{0.85}{
+\begin{tabular}{@{}cccccccc@{}}
 \toprule
-\textbf{Name}                          & \textbf{No. of Patterns} & \textbf{Year established} \\
+\textbf{Name} & \textbf{No. of Patterns} & \textbf{O.A.} & \textbf{Comp.} & \textbf{Spg} & \textbf{Lattice} & \textbf{Unit Cell} & \textbf{Year est.} \\
 \midrule
-Linus Pauling file                 & 21,700                  & 2002        \\
-Powder Diffraction File (PDF)  \tablefootnote{The PDF lists the Material Platform for Data Science (MPDS) as a database source.
-Since the MPDS is hosted by the Pauling File project, there is likely significant overlap in the experimental patterns available in the PDF and the Linus Pauling File.}         & 20,800                  & 1941                      \\
-RRUFF                                  & 1290                    & 2006                      \\
-Crystallography Open Database (COD)    & 1063                    & 2003                      \\
-PowBase                                & 169                     & 1999                      \\
+Linus Pauling file                     & 21,700                    & \xmark          & \cmark        & \cmark        & \cmark        & \cmark          & 2002            \\
+Powder Diffraction File \tablefootnote{The PDF lists the Material Platform for Data Science (MPDS) as a database source.
+Since the MPDS is hosted by the Pauling File project, there is likely significant overlap in the experimental patterns available in the PDF and the Linus Pauling File.} & 20,800 & \xmark & \cmark & \cmark & \cmark & \cmark & 1941 \\
+RRUFF                                  & 1290                      & \cmark          & \cmark        & \cmark        & \cmark        & \xmark          & 2006            \\
+Crystallography Open Database          & 1063                      & \cmark          & \cmark        & \cmark        & \cmark        & \cmark          & 2003            \\
+PowBase                                & 169                       & \cmark          & \xmark        & \xmark        & \xmark        & \xmark          & 1999            \\
 \bottomrule
-\end{tabular}
+\end{tabular}}
 \end{table}

### 2024-12-05 chapters/04_our_dataset.tex
Added histograms, PCA plots + paragraph on notebook; Added labeled column in existing datasets;

@@
 \label{tab:merged}
+\scalebox{0.92}{
 \begin{tabular}{@{}clcccccl@{}}
@@
 \midrule
-CNRS        & \qquad 895          & \cmark   & \cmark   & \cmark   & \cmark   & Extracted from the COD                 \\ 
+CNRS        & \qquad 1052          & \cmark   & \cmark   & \cmark   & \cmark   & Extracted from the COD                 \\ 
 HKUST(GZ)   & \qquad 21           & \cmark   & \cmark   & \cmark   & \cmark   & Phase identification reference database \\ 
@@
 EMPA        & \qquad 790          & \cmark   & \cmark   & \xmark   & \xmark   & Tin halide perovskites, Zn-V-N library \\ 
-INT         & \qquad 19780        & \xmark   & \xmark   & \xmark   & \xmark   & Compilation of various projects        \\ 
+INT         & \qquad 19,780        & \xmark   & \xmark   & \xmark   & \xmark   & Compilation of various projects        \\ 
 IKFT        & \qquad 64           & \xmark   & \xmark   & \xmark   & \xmark   & Commercial catalysts, bulk materials   \\ 
@@
 \bottomrule
-\end{tabular}
+\end{tabular}}
 \end{table}
+
+%TODO:
+% The three datasets shown here are (1): USC CuNi, (2): USC CuAl, (3): HKUST
+% They certainly form clusters, but we have to think about how to interpret these results. It might still represent background/angle range more than anything else more interesting like material class.
+% Also if we add a larger dataset like the INT data the variance is so huge that all these three more narrow datasets effectively collapse onto a tiny region (see below)
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=0.5\linewidth]{figures/pca_scatter_smol.png}
+    \caption{Two component PCA scatter plot showing (1): USC CuNi, (2): USC CuAl, (3): HKUST data2}
+    \label{fig:enter-label}
+\end{figure}
+
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=0.5\linewidth]{figures/pca_scatter.png}
+    \caption{Two component PCA scatter plot including all opXRD data}
+    \label{fig:enter-label}
+\end{figure}
 
@@
 
-\begin{figure*}[!ht]
+\begin{figure}[!htb]
     \centering
-    \missingfigure{} 
-    \caption{Statistics, histograms, etc. of our dataset.}
-    \label{fig:statistics}
-\end{figure*}
+    \includegraphics[width=\linewidth]{figures/quantities_hist.png}
+    \caption{Histograms detailing spacegroups of labeled data and number of entries, starting angle value and final angle value for all data}
+    \label{fig:enter-label}
+\end{figure}
 
+\pagebreak
 %% Paragraph by FX Coudert and Arthur Hardiagon
@@
 Of these 5432 files, most contained only metadata related to the powder diffraction experiment, but did not include the raw data of the pattern itself.
-We could extract raw experimental pXRD patterns from 1063 files in total, after curation of a small number of files with clearly invalid data.
+We could extract raw experimental pXRD patterns from 1052 files in total, after curation of a small number of files with clearly invalid data.
 \\

### 2024-12-05 chapters/05_how_to_use.tex
Added histograms, PCA plots + paragraph on notebook; Added labeled column in existing datasets;

@@
-The opXRD database is hosted on Zenodo\footnote{The dataset can be found at \url{https://zenodo.org/records/14254271}.} and can be downloaded by any user without any barriers or restrictions.
+The opXRD database is hosted on Zenodo\footnote{The dataset can be found at \url{https://zenodo.org/records/14278656}.} and can be downloaded by any user without any barriers or restrictions.
 The database comes in two files, ``opxrd.zip'' and ``opxrd\_in\_situ.zip''.
@@
 
-Our tool further includes options for standardization, plotting, and the conversion to \emph{PyTorch} tensors.
-We provide the following Jupyter notebook that showcases the functionality of the library in more detail: <Link to github Jupyter notebook on Colab>
-
-%The patterns attribute of this class is a list of the indivdual pattern diffractograms, each of which supports a standardize, plot and to tensor method.
-The standardize method returns a "standardized" version of the pattern with a fixed angle range, a fixed number of entries and intensities normalized to the $[0,1]$ interval.
+Our library further includes options for standardization, plotting, and the conversion to \emph{PyTorch} tensors.
+We provide a Jupyter Notebook\footnote{This notebook can be found here: \url{https://colab.research.google.com/github/aimat-lab/opXRD/blob/main/opxrd/usage.ipynb}.} that showcases these functionalities in more detail.
+This notebook also illustrates how to interface with the OpXRD database through Python.

### 2024-12-06 chapters/01_introduction.tex
Added labeled column in dataset overview
Eliminated redundancy in author listing

@@
 Especially for novel structures, obtaining an initial structural model through matching to a database of known structures can be problematic.
-\\ %In such cases, more complex crystal structure solution approaches need to be applied to find suitable crystal structure guesses.
-\\
+In such cases, more complex crystal structure solution approaches need to be applied to find suitable crystal structure guesses \cite{structure_solution_review}.\\
 
@@
 The broad range of available experimental samples contained in the opXRD v1.0 database makes it possible to apply state-of-the-art ML approaches to the domain of powder XRD analysis.
-We hope that this can drive ML research in this field towards more advanced automated analysis workflows that can accelerate materials science research through ready application in high-throughput experimentation pipelines.
+We hope that this can drive ML research in this field towards more advanced automated analysis workflows that can accelerate materials science research through ready application in high-throughput experimentation pipelines.\\

### 2024-12-06 chapters/02_SOTA_ML.tex
Eliminated redundancy in author listing

@@
-%\paragraph{ML for pXRD}
+\paragraph{ML for pXRD}
+
 To showcase the need for datasets such as the one presented in this publication, we now shortly discuss some recent
@@
 The study observed a noticeable performance drop when the pre-trained model was applied to real experimental patterns, as opposed to simulated data.
-However, their recent work \cite{cao2024simxrd} proposes using convolutional peak descriptors that consider the detector's geometry, which significantly reduces the performance gap in their benchmark tests.\\
+However, their recent work \cite{cao2024simxrd} proposes using convolutional peak descriptors that consider the detector's geometry, which reduces the performance gap in their benchmark tests.\\
 
@@
 Augmenting our simulated patterns with background, noise, and impurities helps to bring simulated diffractograms closer to experimental ones, making models trained on them more performant on experimental diffractograms.
-However, this augmentation process could be improved by incorporating background and noise statistics from a broader experimental pXRD database.\\
+However, this augmentation process could be improved by incorporating background and noise statistics from a broader experimental pXRD database, such as the one presented in this publication.\\
 
-It becomes apparent, that the more general the task is, the more challenging the transfer to experimental data is.
-For example, the space group classification task is very general, for all material systems.
+It becomes apparent that the more general the task is, the more challenging the transfer to experimental data.
+For example, the space group classification task across all material systems is very general.
 Therefore, transferring it to the application on experimental diffraction patterns is difficult.
 \cite{Schopmans2023, Lee2022, Vecsei2018} On the other hand, there are some successful approaches that also work well on experimental data, but those are mostly methods that do phase determination in a limited compound space, making the task less complex.
-\cite{Schuetzke2024, Lee2020}
+\cite{Schuetzke2024, Lee2020}\\
 
-The ultimate goal of ML efforts is to make pXRD data identification practical for experimental use, automating structure property prediction.
+The ultimate goal of ML efforts is to make pXRD data identification practical for experimental use, automating structure prediction.
 However, the current volume of experimental pXRD patterns is insufficient to effectively train ML models, highlighting an urgent need for a comprehensive experimental pXRD database.
-To address this, two key approaches are essential.
+Two key approaches are essential to address this.
 First, developing more sophisticated simulation methods to better approximate experimental patterns\cite{cao2024simxrd} by using statistics from experimental diffractograms.

### 2024-12-06 chapters/03_existing_datasets.tex
Eliminated redundancy in author listing

@@
 To contextualize opXRD within the current environment of experimental powder diffraction data, the following list provides an overview of the largest crystal structure databases that offer access to experimental powder diffraction data.
-For an overview of these databases refer to Tab.$\eqref{tab:exp_databases}$ below.\\
+For an overview of these databases refer to Tab.$\eqref{tab:exp_databases}$ below.
 
@@
 The MPDS is partially open, with the open MPDS data accessible through a web interface\cite{MPDSWeb}.
-Access to the full MPDS database can be purchased through a one-year license starting at \qty{2000}.
+Access to the full MPDS database can be purchased through a one-year license starting at \qty{2000}{\myeuro}.
 \\

### 2024-12-06 chapters/04_our_dataset.tex
Eliminated redundancy in author listing

@@
 Most data was collected using \ce{Cu} radiaton sources which have a  $K_{\alpha1}$ wavelength of $\lambda=1.54056\text{\AA}$ and a $K_{\alpha2}$ wavelength of $\lambda=1.54439\text{\AA}$.
+
+\newcommand{\partialexclam}[1]{\textcolor{orange}{\ding{52}(#1\%)}}
 
@@
 \scalebox{0.92}{
-\begin{tabular}{@{}clcccccl@{}}
+\begin{tabular}{@{}clccclll@{}}
 \toprule
-Institution & No. Patterns & Comp. & Spg & Lattice & Unit Cell & Research Project \\ 
+Institution & No. Patterns & Comp. & Spg & Lattice & Basis & Research Project \\ 
 \midrule
-CNRS        & \qquad 1052          & \cmark   & \cmark   & \cmark   & \cmark   & Extracted from the COD                 \\ 
+CNRS        & \qquad 1052          & \cmark   & \cmark   & \cmark   &  \partialexclam{79}   & Extracted from the COD                 \\ 
 HKUST(GZ)   & \qquad 21           & \cmark   & \cmark   & \cmark   & \cmark   & Phase identification reference database \\ 

### 2024-12-06 main.tex
Added square to citations to differentiate from footnotes
Added showing DOI as clickable links in bibliography when available
Added paragraph on Pearson's crystal data
Changed plaintext SI units/degrees into latex symbols
Warpped chemical formulas in \ce{} environemnt
Commented out statement that pearson and PDF have huge overlap
Described overlap in experimental pXRD data in PDF5+ and Pearson's Crystal data
Added dataset overview tables
Updated introductory paragraphs on rietveld refinement
Described python + zenodo data availability
Commited table version with footnote directly underneath
Ad

--- 2024-11-21 main.tex
@@
 \usepackage[utf8]{inputenc}
-\usepackage[super,sort&compress,comma]{natbib}
+\usepackage[super,sort&compress,comma, square]{natbib}
 %\renewcommand{\supercite}[1]{\cite{#1}}
@@
 %% Language and font encodings
+\usepackage{tikz}
+\usepackage{tablefootnote}
+\usepackage{gensymb}
 \usepackage[english]{babel}
@@
 \usepackage{csquotes}
-
+\usepackage{doi}
 %% Sets page size and margins
@@
 % Useful packages
+\usepackage[version=4]{mhchem}
 \usepackage{amsmath}
@@
 \usepackage{multirow}
+\usepackage{nicematrix,enumitem,booktabs}
 \usepackage{nomencl}
@@
 \usepackage[version=4]{mhchem}
-\usepackage{siunitx}
+\usepackage{eurosym}
+\usepackage[locale = DE]{siunitx}
 \usepackage{bm}
 \usepackage[normalem]{ulem} %%temporary, for organization purposes
+\usepackage{pythonhighlight}
+\usepackage{graphicx}
+\usepackage{subcaption}
+
+\usepackage{pifont}% http://ctan.org/pkg/pifont
+\newcommand{\cmark}{\textcolor{green!80!black}{\ding{52}}}
+\newcommand{\xmark}{\textcolor{red}{\ding{53}}}
 
@@
 \newcommand{\red}[1]{\textcolor{red}{#1}}
+\DeclareSIUnit{\myeuro}{\text{\euro}}
 \DeclareSIUnit\angstrom{\text {Å}}
@@
 \interfootnotelinepenalty=10000
-% Changed idea title idea from orignal title of
-% "Towards a free high quality powder x-ray diffractogram database"
-% Name is supposed to be in analogy with Crystallography Open database
-% which is the largest freely available database of crystal structures
 \title{opXRD: Open Experimental Powder X-ray Diffraction Database}
 
-\author[1,2]{Daniel Hollarek}
-\author[1,2]{Henrik Schopmans}
-\author[1,2]{Jona Östreicher}
-\author[2]{Simon Schweidler}
-\author[3]{Adie Alwen}
-\author[4]{Mriganka Singh}
-\author[4,5]{Tim Kodalle}
-\author[2]{Ben Breitung}
-\author[11, 12]{Maged Abdelsamie}
-\author[6]{Gregoire Heymans}
-\author[7]{Alexander Wieczorek}
-\author[7]{Siarhei Zhuk}
-\author[7]{Sebastian Siol}
-\author[8]{Moritz Wolf}
-\author[9]{François-Xavier Coudert}
-\author[9]{Arthur Hardiagon}
-\author[10]{Bin Cao}
-\author[13]{Ruth Schwaiger}
-\author[10]{Tong-yi Zhang}
-\author[4]{Carolin M. Sutter-Fella}
-\author[3]{Andrea M. Hodge}
-\author[1,2]{Pascal Friederich}
+\newcommand{\ITI}{1}
+\newcommand{\INT}{2}
+\newcommand{\HKUSTGZ}{3}
+\newcommand{\USC}{4}
+\newcommand{\LBLMF}{5}
+\newcommand{\LBLALS}{6}
+\newcommand{\Hoffmann}{7}
+\newcommand{\LBLCS}{8}
+\newcommand{\EMPA}{9}
+\newcommand{\KITICRT}{10}
+\newcommand{\CPT}{11}
+\newcommand{\KFUPMMSE}{12}
+\newcommand{\KFUPMIRCIMR}{13}
+\newcommand{\FZJIEMD}{14}
 
-% I made the affiliations smaller because there's a lot of them
+\author[\ITI,\INT]{Daniel Hollarek}
+\author[\ITI,\INT]{Henrik Schopmans}
+\author[\INT]{Jona Östreicher}
+\author[\HKUSTGZ]{Bin Cao}
+\author[\INT]{Simon Schweidler}
+\author[\USC]{Adie Alwen}
+\author[\LBLMF]{Mriganka Singh}
+\author[\LBLMF,\LBLALS]{Tim Kodalle}
+\author[\Hoffmann]{Hanlin Hu}
+\author[\LBLCS]{Gregoire Heymans}
+\author[\EMPA]{Alexander Wieczorek}
+\author[\EMPA]{Siarhei Zhuk}
+\author[\CPT]{Arthur Hardiagon}
+\author[\KFUPMMSE,\KFUPMIRCIMR]{Maged Abdelsamie}
+\author[\HKUSTGZ]{Tong-yi Zhang}
+\author[\USC]{Andrea M. Hodge}
+\author[\FZJIEMD]{Ruth Schwaiger}
+\author[\INT]{Ben Breitung}
+\author[\EMPA]{Sebastian Siol}
+\author[\KITICRT]{Moritz Wolf}
+\author[\CPT]{François-Xavier Coudert}
+\author[\LBLMF]{Carolin M. Sutter-Fella}
+\author[\ITI,\INT]{Pascal Friederich}
+
 \renewcommand\Affilfont{\small}
 
-\affil[1]{Institute of Theoretical Informatics, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany}
-\affil[2]{Institute of Nanotechnology, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany}
-\affil[3]{Department of Chemical Engineering and Materials Science, University of Southern California, Los Angeles CA 90089, USA}
-\affil[4]{Lawrence Berkeley National Laboratory, Molecular Foundry Division, Berkeley 94720 CA, USA}
-\affil[5]{Lawrence Berkeley National Laboratory, Advanced Light Source, Berkeley 94720 CA, USA}
-\affil[6]{Lawrence Berkeley National Laboratory, Chemical Sciences Division, Berkeley 94720 CA, USA}
-\affil[7]{Empa~–~Swiss Federal Laboratories
-for Materials Science and Technology, 8600 Dübendorf,
-Switzerland}
-\affil[8]{Engler-Bunte-Institut \& Institute of Catalysis
-Research and Technology, Karlsruhe Institute of Technology, Karlsruhe, Germany}
-\affil[9]{Chimie ParisTech, PSL University, CNRS, Institut de Recherche de Chimie Paris, 75005 Paris, France}
-\affil[10]{Guangzhou Municipal Key Laboratory of Materials Informatics,  Hong Kong University of
-Science and Technology (Guangzhou), Guangzhou 511400, China}
-\affil[11]{Material Science and Engineering Department, King Fahd University of Petroleum and Minerals (KFUPM), Dhahran 31261, Saudi Arabia}
-\affil[12]{Interdisciplinary Research Center for Intelligent Manufacturing and Robotics, KFUPM, Dhahran 31261, Saudi Arabia}
-\affil[13]{Institute of Energy Materials and Devices, Forschungszentrum Juelich GmbH, 52425 Juelich, Germany}
+\affil[\ITI]{Institute of Theoretical Informatics, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany}
+\affil[\INT]{Institute of Nanotechnology, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany}
+\affil[\USC]{Department of Chemical Engineering and Materials Science, University of Southern California, Los Angeles CA 90089, USA}
+\affil[\LBLMF]{Lawrence Berkeley National Laboratory, Molecular Foundry Division, Berkeley 94720 CA, USA}
+\affil[\LBLALS]{Lawrence Berkeley National Laboratory, Advanced Light Source, Berkeley 94720 CA, USA}
+\affil[\Hoffmann]{Hoffmann Institute of Advanced Materials, Shenzhen Polytechnic, Shenzhen 518055, China}
+\affil[\LBLCS]{Lawrence Berkeley National Laboratory, Chemical Sciences Division, Berkeley 94720 CA, USA}
+\affil[\EMPA]{Empa–Swiss Federal Laboratories for Materials Science and Technology, 8600 Dübendorf, Switzerland}
+\affil[\KITICRT]{Engler-Bunte-Institut \& Institute of Catalysis Research and Technology, Karlsruhe Institute of Technology, Karlsruhe, Germany}
+\affil[\CPT]{Chimie ParisTech, PSL University, CNRS, Institut de Recherche de Chimie Paris, 75005 Paris, France}
+\affil[\HKUSTGZ]{Guangzhou Municipal Key Laboratory of Materials Informatics, Hong Kong University of Science and Technology (Guangzhou), Guangzhou 511400, China}
+\affil[\KFUPMMSE]{Material Science and Engineering Department, King Fahd University of Petroleum and Minerals (KFUPM), Dhahran 31261, Saudi Arabia}
+\affil[\KFUPMIRCIMR]{Interdisciplinary Research Center for Intelligent Manufacturing and Robotics, KFUPM, Dhahran 31261, Saudi Arabia}
+\affil[\FZJIEMD]{Institute of Energy Materials and Devices, Forschungszentrum Juelich GmbH, 52425 Juelich, Germany}
+
 \begin{document}
@@
 \input{chapters/01_introduction}
+\input{chapters/02_SOTA_ML}
 
-\*section{Existing datasets}
-\input{chapters/02_existing_datasets}
+\section{Existing datasets}
+\input{chapters/03_existing_datasets}
 
-\*section{Machine learning for pXRD}
-\input{chapters/03_SOTA_ML}
-
+\pagebreak
 \section{opXRD database}\label{sec:our_dataset}
@@
 
-\section{How to contribute and how to use}\label{sec:how_to_use}
+\section{Usage}\label{sec:how_to_use}
 \input{chapters/05_how_to_use}
+
+\section{Summary and Outlook}\label{sec:summary_and_outlook}
+\input{chapters/06_summary_and_outlook}
 
@@
 
+\bibliographystyle{bibstyle/bibstyle.bst}
 \bibliography{references}
-\bibliographystyle{rsc}
 

### 2024-12-08 chapters/04_our_dataset.tex
Added percentages to database overview

--- 2024-12-05 chapters/04_our_dataset.tex
@@
 Most data was collected using \ce{Cu} radiaton sources which have a  $K_{\alpha1}$ wavelength of $\lambda=1.54056\text{\AA}$ and a $K_{\alpha2}$ wavelength of $\lambda=1.54439\text{\AA}$.
+
+\newcommand{\partialcheck}[1]{\textcolor{orange}{\ding{52}(#1\%)}}
 
@@
 \label{tab:merged}
-\scalebox{0.92}{
-\begin{tabular}{@{}clcccccl@{}}
+\scalebox{0.90}{
+\begin{tabular}{@{}lllllll@{}}
 \toprule
-Institution & No. Patterns & Comp. & Spg & Lattice & Unit Cell & Research Project \\ 
+Institution & No. Patterns & Comp. & Spg & Lattice & Atom coords. & Research Project \\ 
 \midrule
-CNRS        & \qquad 1052          & \cmark   & \cmark   & \cmark   & \cmark   & Extracted from the COD                 \\ 
+CNRS        & \qquad 1052          & \cmark   & \partialcheck{85}   & \cmark   &  \partialcheck{85}   & Extracted from the COD \\ 
 HKUST(GZ)   & \qquad 21           & \cmark   & \cmark   & \cmark   & \cmark   & Phase identification reference database \\ 
-USC         & \qquad 338          & \cmark   & \cmark   & \cmark   & \xmark   & Study of CuNi and CuAl alloys          \\ 
-EMPA        & \qquad 790          & \cmark   & \cmark   & \xmark   & \xmark   & Tin halide perovskites, Zn-V-N library \\ 
+USC         & \qquad 338          & \cmark   & \cmark   & \partialcheck{90}   & \xmark   & Study of CuNi and CuAl alloys          \\ 
+EMPA        & \qquad 790          & \cmark   & \partialcheck{63}   & \xmark   & \xmark   & Tin halide perovskites, Zn-V-N library \\ 
 INT         & \qquad 19,780        & \xmark   & \xmark   & \xmark   & \xmark   & Compilation of various projects        \\ 
@@
 \\
+
+% Paragraph by Gregoire Heymans
+
+\subsubsection*{\textit{In situ} XRD (RT $\le$ T$_{annealing}$ $\le$ 850$\degree$C)}
+
+The phase evolution of the Mn-Sb-O system varying with annealing temperature was measured using the \textit{in situ} XRD (Rigaku-SmartLab3kW XRD, Figure \ref{fig:stability} (a)).
+This tool operates with SmartLab Studio II software, which is able to measure the X-ray diffraction during the annealing process.
+This enables to show directly all the phase transitions when annealing in various atmospheres such as O$_2$, Ar and NH$_3$.
+This device can theoretically operate up to 1000$\degree$C in a 20\% O$_2$ atmosphere.
+For safety reasons, phase transitions are analyzed with the \textit{in situ} XRD tool up to 850$\degree$C in this work.
+
+Most of the \textit{in situ} experiments were performed under an air-like 20\% O$_2$ and 80\% argon (Ar) environment is chosen (Ar flow: 50 sccm, O$_2$ flow: 10 sccm).
+When a 100\% Ar environment is fixed, an Ar flow of 60 sccm is input.
+
+Concerning the geometry, the Bragg-Brentano (BB) mode is preferred because it is more adapted in the analysis of scarce phases (i.e., MnSb$_2$O$_6$ rutile).
+The step is 0.01$\degree$ and the scanning rate is 10$\degree$/min.
+The step gives information about the resolution and the scanning rate (or speed duration time) results in the quality of the signal-to-noise ratio.
+
+Figure \ref{fig:temp} shows the temperature profile during the \textit{in situ} XRD measurements.
+The temperatures used to annalyze the crystal structure of the Mn-Sb-O system were chosen depending on the number of phase transitions appearing for a certain temperature range.
+As explained in details in chapter \ref{res}, few changes in the crystal structure appear between RT and 300$\degree$C and phase transitions appeared from 300$\degree$C until 850$\degree$C.
+No phase transition appeared when cooling down.
+Therefore, the crystal structure was measured every 100$\degree$C between RT and 300$\degree$C; every 50$\degree$C between 300$\degree$C and 850$\degree$C; and every 200$\degree$C when cooling down.
+The heating and cooling rates are fixed for all the experiments at 50$\degree$C/min and the holding time is fixed at 2 min.
+
+% end of Gregoire Heymans

### 2024-12-08 chapters/03_existing_datasets.tex
Added paragraphs on data correlation and filtering mechanisms

@@
 RRUFF                                  & 1290                      & \cmark          & \cmark        & \cmark        & \cmark        & \xmark          & 2006            \\
-Crystallography Open Database          & 1063                      & \cmark          & \cmark        & \cmark        & \cmark        & \cmark          & 2003            \\
+Crystallography Open Database          & 1052                      & \cmark          & \cmark        & \cmark        & \cmark        & \cmark          & 2003            \\
 PowBase                                & 169                       & \cmark          & \xmark        & \xmark        & \xmark        & \xmark          & 1999            \\
@@
 It currently provides over 500,000 crystal structures.
-Of these files, 1063 contain the experimental powder diffraction data that was used to determine the underlying crystal structures of the investigated samples.
-Hence, the experimental powder diffraction data contained in the COD is labeled with the full crystal structure information.
+Of these files, 1052 contain the experimental powder diffraction data that was used to determine the underlying crystal structures of the investigated samples.
+Hence, the experimental powder diffraction data contained in the COD is mostly labeled with the full crystal structure information.
 \\

### 2024-12-08 chapters/04_our_dataset.tex
Added paragraphs on data correlation and filtering mechanisms

@@
 Most data was collected using \ce{Cu} radiaton sources which have a  $K_{\alpha1}$ wavelength of $\lambda=1.54056\text{\AA}$ and a $K_{\alpha2}$ wavelength of $\lambda=1.54439\text{\AA}$.
+We filtered the submitted datasets to exclude patterns with invalid features such as patterns with only one unique recorded angle, patterns in which all intensities were zero, or patterns that contained less than 50 recorded angles.
+\\
 
-\newcommand{\partialcheck}[1]{\textcolor{orange}{\ding{52}(#1\%)}}
+The below table Tab.$\eqref{tab:merged}$ provides an overview of the opXRD database.
+The columns to the right of No.
+patterns indicate the availability of properties of the underlying samples.
+The column titled ``Comp.'' indicates the availability of the chemical composition, and the column titled ``Spg'' indicates the availability of the spacegroup.
+The ``Lattice'' and ``Atom coords.'' columns indicate the availability of lattice parameters and atomic coordinates respectively.\\
 
@@
 \centering
-\caption{opXRD database entries}
+\caption{The opXRD database}
 \label{tab:merged}
-\scalebox{0.90}{
+\scalebox{0.9}{
+\begin{adjustbox}{center}
 \begin{tabular}{@{}lllllll@{}}
 \toprule
-Institution & No. Patterns & Comp. & Spg & Lattice & Atom coords. & Research Project \\ 
+Institution & No. patterns & Comp. & Spg. & Lattice & Atom coords. & Research Project \\ 
 \midrule
-CNRS        & \qquad 1052          & \cmark   & \partialcheck{85}   & \cmark   &  \partialcheck{85}   & Extracted from the COD \\ 
-HKUST(GZ)   & \qquad 21           & \cmark   & \cmark   & \cmark   & \cmark   & Phase identification reference database \\ 
-USC         & \qquad 338          & \cmark   & \cmark   & \partialcheck{90}   & \xmark   & Study of CuNi and CuAl alloys          \\ 
-EMPA        & \qquad 790          & \cmark   & \partialcheck{63}   & \xmark   & \xmark   & Tin halide perovskites, Zn-V-N library \\ 
-INT         & \qquad 19,780        & \xmark   & \xmark   & \xmark   & \xmark   & Compilation of various projects        \\ 
-IKFT        & \qquad 64           & \xmark   & \xmark   & \xmark   & \xmark   & Commercial catalysts, bulk materials   \\ 
-LBNL        & \qquad n/a          & \xmark   & \xmark   & \xmark   & \xmark   & Time series recordings of Perovskites  \\ 
+CNRS        &  1052          & \cmark   & \partialcheck{85}   & \cmark   &  \partialcheck{85}   & Extracted from the COD \\ 
+HKUST(GZ)   &  21           & \cmark   & \cmark   & \cmark   & \cmark   & Phase identification reference database \\ 
+USC         &  338          & \cmark   & \cmark   & \partialcheck{90}   & \xmark   & Study of CuNi and CuAl alloys          \\ 
+EMPA        &  790          & \cmark   & \partialcheck{63}   & \xmark   & \xmark   & Tin halide perovskites, Zn-V-N library \\ 
+INT         &  19,780        & \xmark   & \xmark   & \xmark   & \xmark   & Compilation of various projects        \\ 
+IKFT        &  64           & \xmark   & \xmark   & \xmark   & \xmark   & Commercial catalysts, bulk materials   \\ 
+LBNL        &  n/a          & \xmark   & \xmark   & \xmark   & \xmark   & Time series recordings of Perovskites  \\ 
 \bottomrule
-\end{tabular}}
+\end{tabular}
+\end{adjustbox}}
 \end{table}
+
+%TODO: A paragraph on how data is correlated
+The degree of variation between the patterns is different for each contribution.
+For example, the CRNS and the INT contributions each are collections that encompass many research projects over a large period of time and hence exhibit a high degree of variability between individual patterns.
+In contrast, the contributions by the USC and the LBNL contain many very similar patterns.
+The patterns in the USC datasets are similar because the underlying samples are all variations of CuNi and CuAl alloys collected on the same machine.
+The LBNL patterns are similar because they stem from in-situ recordings where several hundred or several thousand patterns were collected over time per sample while they were undergoing physical processes.
+The following paragraphs describe each contribution individually.
+\\
 
@@
 \end{figure}
-
-%TODO: A paragraph on how data is correlated
 

### 2024-12-08 chapters/03_existing_datasets.tex
Added partial percentages in existing dataset table

@@
-\subsubsection*{Existing experimental powder diffraction databases}
 To contextualize opXRD within the current environment of experimental powder diffraction data, the following list provides an overview of the largest crystal structure databases that offer access to experimental powder diffraction data.
 For an overview of these databases refer to Tab.$\eqref{tab:exp_databases}$ below.
+In this table, the column ``O.A.'' indicates whether or not the database is open access.
+The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'' respectively.
 
@@
 \centering
-\caption{Overview of Experimental Powder Diffraction Databases}
+\caption{Overview of experimental powder diffraction databases}
 \label{tab:exp_databases}
 \scalebox{0.85}{
-\begin{tabular}{@{}cccccccc@{}}
+\begin{adjustbox}{center}
+\begin{tabular}{@{}llllllll@{}}
 \toprule
-\textbf{Name} & \textbf{No. of Patterns} & \textbf{O.A.} & \textbf{Comp.} & \textbf{Spg} & \textbf{Lattice} & \textbf{Unit Cell} & \textbf{Year est.} \\
+\textbf{Name} & \textbf{No. patterns} & \textbf{O.A.} & \textbf{Comp.} & \textbf{Spg.} & \textbf{Lattice} & \textbf{Atom coords.} & \textbf{Year est.} \\
 \midrule
@@
 Powder Diffraction File \tablefootnote{The PDF lists the Material Platform for Data Science (MPDS) as a database source.
-Since the MPDS is hosted by the Pauling File project, there is likely significant overlap in the experimental patterns available in the PDF and the Linus Pauling File.} & 20,800 & \xmark & \cmark & \cmark & \cmark & \cmark & 1941 \\
+Since the MPDS is hosted by the Pauling File project, there is likely significant overlap in the experimental patterns available in the PDF and the Linus Pauling File.} & 20,800 & \xmark & \cmark & \cmark & \cmark & \partialcheck{52} & 1941 \\
 RRUFF                                  & 1290                      & \cmark          & \cmark        & \cmark        & \cmark        & \xmark          & 2006            \\
-Crystallography Open Database          & 1052                      & \cmark          & \cmark        & \cmark        & \cmark        & \cmark          & 2003            \\
+Crystallography Open Database          & 1052                      & \cmark          & \cmark        & \cmark        & \partialcheck{85}        & \partialcheck{85}          & 2003            \\
 PowBase                                & 169                       & \cmark          & \xmark        & \xmark        & \xmark        & \xmark          & 1999            \\
 \bottomrule
-\end{tabular}}
+\end{tabular}
+\end{adjustbox}}
 \end{table}

### 2024-12-08 chapters/04_our_dataset.tex
Added partial percentages in existing dataset table

@@
 The below table Tab.$\eqref{tab:merged}$ provides an overview of the opXRD database.
-The columns to the right of No.
-patterns indicate the availability of properties of the underlying samples.
-The column titled ``Comp.'' indicates the availability of the chemical composition, and the column titled ``Spg'' indicates the availability of the spacegroup.
-The ``Lattice'' and ``Atom coords.'' columns indicate the availability of lattice parameters and atomic coordinates respectively.\\
+The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'' respectively.\ \
 
@@
 \toprule
-Institution & No. patterns & Comp. & Spg. & Lattice & Atom coords. & Research Project \\ 
+\textbf{Institution} & \textbf{No. patterns} & \textbf{Comp.} & \textbf{Spg.} & \textbf{Lattice} & \textbf{Atom coords.} & \textbf{Research Project} \\
 \midrule
@@
 In contrast, the contributions by the USC and the LBNL contain many very similar patterns.
-The patterns in the USC datasets are similar because the underlying samples are all variations of CuNi and CuAl alloys collected on the same machine.
-The LBNL patterns are similar because they stem from in-situ recordings where several hundred or several thousand patterns were collected over time per sample while they were undergoing physical processes.
+The patterns in the USC datasets are similar because the underlying samples are all variations of \ce{CuNi} and \ce{CuAl} alloys collected on the same machine.
+The patterns submitted by the LBNL are similar because they stem from in-situ recordings where several hundred or several thousand patterns were collected over time per sample while they were undergoing physical processes.
 The following paragraphs describe each contribution individually.

### 2024-12-09 chapters/03_existing_datasets.tex
Added data availability section

@@
 Crystallography Open Database          & 1052                      & \cmark          & \cmark        & \cmark        & \partialcheck{85}        & \partialcheck{85}          & 2003            \\
-PowBase                                & 169                       & \cmark          & \xmark        & \xmark        & \xmark        & \xmark          & 1999            \\
+PowBase                                & 169                       & \cmark          & \cmark        & \xmark        & \xmark        & \xmark          & 1999            \\
 \bottomrule
@@
 It currently provides over 500,000 crystal structures.
-Of these files, 1052 contain the experimental powder diffraction data that was used to determine the underlying crystal structures of the investigated samples.
+Of these files, 1052 contains the experimental powder diffraction data that was used to determine the underlying crystal structures of the investigated samples.
 Hence, the experimental powder diffraction data contained in the COD is mostly labeled with the full crystal structure information.

### 2024-12-09 chapters/04_our_dataset.tex
Added data availability section

@@
 Most data was collected using \ce{Cu} radiaton sources which have a  $K_{\alpha1}$ wavelength of $\lambda=1.54056\text{\AA}$ and a $K_{\alpha2}$ wavelength of $\lambda=1.54439\text{\AA}$.
-We filtered the submitted datasets to exclude patterns with invalid features such as patterns with only one unique recorded angle, patterns in which all intensities were zero, or patterns that contained less than 50 recorded angles.
+We filtered the submitted datasets to exclude patterns with invalid features such as the pattern containing only one unique recorded angle, the pattern containing less than 50 recorded angles total, or all intensities being zero.
 \\
@@
 The below table Tab.$\eqref{tab:merged}$ provides an overview of the opXRD database.
-The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'' respectively.\ \
+The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'' respectively.
 
@@
 \label{tab:merged}
-\scalebox{0.9}{
+\scalebox{0.81}{
 \begin{adjustbox}{center}
@@
 The COD is, to our knowledge, the largest open-access collection of experimental crystal structures of organic, inorganic, and metal-organic compounds and minerals, containing more than 500,000 entries.
-The data in the COD is placed in the public domain and licensed under the CC0 License.
+The data in the COD are placed in the public domain and licensed under the CC0 License.
 Of the entire COD database 5432 structures contained at least one tag from the {CIF\_POW} dictionary, i.e., a tag relating to powder diffraction studies.
@@
 
-Additionally, Xqueryer,\footnote{Xqueryer can be accessed at \url{http://xqueryer.caobin.asia/}.} a cutting-edge free online platform for intelligent structure identification has been launched.
-Users can upload experimental pXRD data to identify crystal structures and extract all relevant information available in the Materials Project database.
+Additionally, Xqueryer,\footnote{Xqueryer can be accessed at \url{https://xqueryer.caobin.asia/}.} a cutting-edge free online platform for intelligent structure identification has been launched.
+Users can upload experimental pXRD data to identify crystal structures and extract all relevant information available in the Materials Project database\cite{jain2013commentary}.
 Uploaded data are securely stored and integrated into the comprehensive opXRD database to progress toward a large-scale, long-term experimental resource.

### 2024-12-09 chapters/05_how_to_use.tex
Added data availability section

@@
 
-Next to the availability of the opXRD dataset on Zenodo, we also provide a Python library \"opxrd\" to easily download and interface with the dataset.
+Next to the availability of the opXRD dataset on Zenodo, we also provide a Python library ``opxrd'' to easily download and interface with the dataset.
 The instructions for how to install this library can be found in the repository associated with the library.\footnote{The repository to this library can is located at \url{https://github.com/aimat-lab/opxrd}.} The opxrd library allows the dataset to be accessed through one simple command: \pyth{OpXRD.load(root_dirpath)}.

### 2024-12-09 main.tex
Added paragraphs on data correlation and filtering mechanisms
Added data availability section

@@
 \usepackage{subcaption}
+\usepackage{adjustbox}
 
@@
 \newcommand{\xmark}{\textcolor{red}{\ding{53}}}
+\newcommand{\partialcheck}[1]{\textcolor{orange}{\ding{52}(#1\%)}}
 
@@
 \author[\KFUPMMSE,\KFUPMIRCIMR]{Maged Abdelsamie}
+\author[\HKUSTGZ]{Longhan Zhang}
 \author[\HKUSTGZ]{Tong-yi Zhang}
@@
 \affil[\INT]{Institute of Nanotechnology, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany}
+\affil[\HKUSTGZ]{Guangzhou Municipal Key Laboratory of Materials Informatics, Hong Kong University of Science and Technology (Guangzhou), Guangzhou 511400, China}
 \affil[\USC]{Department of Chemical Engineering and Materials Science, University of Southern California, Los Angeles CA 90089, USA}
@@
 \affil[\CPT]{Chimie ParisTech, PSL University, CNRS, Institut de Recherche de Chimie Paris, 75005 Paris, France}
-\affil[\HKUSTGZ]{Guangzhou Municipal Key Laboratory of Materials Informatics, Hong Kong University of Science and Technology (Guangzhou), Guangzhou 511400, China}
+
 \affil[\KFUPMMSE]{Material Science and Engineering Department, King Fahd University of Petroleum and Minerals (KFUPM), Dhahran 31261, Saudi Arabia}
@@
 \subsection*{Data availability}
-Data is available on xxx.
+The opXRD database is available on Zenodo at \url{https://zenodo.org/records/14254270}.
+It can be downloaded by any user without any barriers or restrictions.
+For further details, please refer to section $\eqref{sec:how_to_use}$.
 

### 2024-12-10 chapters/03_existing_datasets.tex
Updated histogram

@@
-To contextualize opXRD within the current environment of experimental powder diffraction data, the following list provides an overview of the largest crystal structure databases that offer access to experimental powder diffraction data.
+To contextualize opXRD within the current environment of experimental powder diffraction data, the below list provides an overview of the largest crystal structure databases that offer access to experimental powder diffraction data.
 For an overview of these databases refer to Tab.$\eqref{tab:exp_databases}$ below.
@@
 \label{tab:exp_databases}
-\scalebox{0.85}{
+\scalebox{0.84}{
 \begin{adjustbox}{center}
@@
 This makes the Pauling file, to the best of our knowledge, the largest collection of experimental powder diffraction data available to researchers.
-As of November 2024, Pearson's crystal data is available to researchers through a purchase of a one-year license starting at a price point of \qty{2200}{\myeuro}.
+As of November 2024, Pearson's crystal data is available to researchers through a purchase of a one-year license starting at a price point of \qty{2200}{\myeuro}\cite{PearsonBuy}.
 The MPDS is partially open, with the open MPDS data accessible through a web interface\cite{MPDSWeb}.
-Access to the full MPDS database can be purchased through a one-year license starting at \qty{2000}{\myeuro}.
+API access to the full MPDS can be purchased through a one-year license starting at \qty{9500}{\myeuro}\cite{MPDSBuy}.
+When inquired about the availability of experimental powder diffraction data in the MPDS API, the Pauling File project stated that this data is not currently included.
+It could however be offered in the future at the request of customers.
 \\
@@
 It contains \num{1290} powder diffraction patterns stemming from experiments each labeled with the lattice parameters and composition of the underlying structures.
+The RRUFF data is openly accessible on the official website \cite{RRUFFWeb}.
 \\
@@
 Hence, the experimental powder diffraction data contained in the COD is mostly labeled with the full crystal structure information.
+The data is openly accessible in the form of .cif files on the official COD website\cite{CODWeb}.
 \\

### 2024-12-10 chapters/04_our_dataset.tex
Updated histogram

@@
 Most data was collected using \ce{Cu} radiaton sources which have a  $K_{\alpha1}$ wavelength of $\lambda=1.54056\text{\AA}$ and a $K_{\alpha2}$ wavelength of $\lambda=1.54439\text{\AA}$.
-We filtered the submitted datasets to exclude patterns with invalid features such as the pattern containing only one unique recorded angle, the pattern containing less than 50 recorded angles total, or all intensities being zero.
+We filtered the submitted datasets to exclude patterns with invalid features such as the pattern containing only one unique recorded angle or negative angles, the pattern containing less than 50 recorded angles total, or all intensities being zero.
 \\
@@
 The patterns submitted by the LBNL are similar because they stem from in-situ recordings where several hundred or several thousand patterns were collected over time per sample while they were undergoing physical processes.
-The following paragraphs describe each contribution individually.
-\\
+
+%The below histograms $\eqref{fig:histograms}$ show the distributions of spacegroups,
+
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=\linewidth]{figures/quantities_hist.png}
+    \caption{Histograms detailing spacegroups of labeled data and number of entries, starting angle value and final angle value for all data}
+    \label{fig:histograms}
+\end{figure}
 
@@
 
-\begin{figure}[!htb]
-    \centering
-    \includegraphics[width=\linewidth]{figures/quantities_hist.png}
-    \caption{Histograms detailing spacegroups of labeled data and number of entries, starting angle value and final angle value for all data}
-    \label{fig:enter-label}
-\end{figure}
+The following paragraphs describe each contribution individually. \\
 
@@
 
-\subsubsection*{\textit{In situ} XRD (RT $\le$ T$_{annealing}$ $\le$ 850$\degree$C)}
+In a second study, the phase evolution of the Mn-Sb-O system varying with annealing temperature was measured using the \textit{in situ}  Rigaku-SmartLab3kW diffractometer.
+This tool operates with SmartLab Studio II software, which can measure the X-ray diffraction during the annealing process.
+This enables directly showing all the phase transitions when annealing in various atmospheres such as \ce{O2}, \ce{Ar}, and \ce{NH3}.
+Phase transitions are analyzed with the \textit{in situ} XRD tool up to 850$\degree$C in this work.
+Most of the \textit{in situ} experiments were performed under an air-like 20\% \ce{O2} and 80\% argon (Ar) environment is chosen (Ar flow: 50 \si{sccm}, O$_2$ flow: 10 \si{sccm}).
+When a 100\% Ar environment is fixed, an Ar flow of 60 \si{sccm} is input.
+The Bragg-Brentano (BB) mode is preferred in terms of geometry because it is more adapted in the analysis of scarce phases such as  \ce{MnSb2O6} rutile.
+The angular step is 0.01$\degree$ and the scanning rate is 10$\degree$/min.
+The step gives information about the resolution and the scanning rate (or speed duration time) results in the quality of the signal-to-noise ratio.
+\\
 
-The phase evolution of the Mn-Sb-O system varying with annealing temperature was measured using the \textit{in situ} XRD (Rigaku-SmartLab3kW XRD, Figure \ref{fig:stability} (a)).
-This tool operates with SmartLab Studio II software, which is able to measure the X-ray diffraction during the annealing process.
-This enables to show directly all the phase transitions when annealing in various atmospheres such as O$_2$, Ar and NH$_3$.
-This device can theoretically operate up to 1000$\degree$C in a 20\% O$_2$ atmosphere.
-For safety reasons, phase transitions are analyzed with the \textit{in situ} XRD tool up to 850$\degree$C in this work.
-
-Most of the \textit{in situ} experiments were performed under an air-like 20\% O$_2$ and 80\% argon (Ar) environment is chosen (Ar flow: 50 sccm, O$_2$ flow: 10 sccm).
-When a 100\% Ar environment is fixed, an Ar flow of 60 sccm is input.
-
-Concerning the geometry, the Bragg-Brentano (BB) mode is preferred because it is more adapted in the analysis of scarce phases (i.e., MnSb$_2$O$_6$ rutile).
-The step is 0.01$\degree$ and the scanning rate is 10$\degree$/min.
-The step gives information about the resolution and the scanning rate (or speed duration time) results in the quality of the signal-to-noise ratio.
-
-Figure \ref{fig:temp} shows the temperature profile during the \textit{in situ} XRD measurements.
-The temperatures used to annalyze the crystal structure of the Mn-Sb-O system were chosen depending on the number of phase transitions appearing for a certain temperature range.
-As explained in details in chapter \ref{res}, few changes in the crystal structure appear between RT and 300$\degree$C and phase transitions appeared from 300$\degree$C until 850$\degree$C.
+The temperatures used to analyze the crystal structure of the Mn-Sb-O system were chosen depending on the number of phase transitions appearing for a certain temperature range.
+Few changes in the crystal structure appear between RT and 300$\degree$C and phase transitions appeared from 300$\degree$C until 850$\degree$C.
 No phase transition appeared when cooling down.

### 2024-12-13 chapters/01_introduction.tex
Added data availability section
Adjusted table footnote to wiley guidelines

@@
 
+% VISUAL ABSTRACT FIGURE
+\begin{figure}[!b]
+    \centering
+    \includegraphics[width=1.0\linewidth]{figures/figure1_v5.pdf}
+    \caption{Overview}
+    \label{fig:overview}
+\end{figure}
+
 Additionally, attempting to refine all crystal structure parameters at once, starting from this initial structure, is known to lead to unphysical results\cite{Ozaki2020}.
@@
 Finding the correct order in which to refine structure parameters and finding the correct values for fixed parameters both present problems that add to the difficulty of the refinement process.\\
+
+%\begin{figure}[!htb]
+%    \centering
+%    \includegraphics[width=0.6\linewidth]{figures/rietveld_flowchart.png}
+%    \caption{Enter Caption}
+%    \label{fig:enter-label}
+%\end{figure}
 
@@
 \cite{Salgado2023, Schopmans2023}
-This is - to the best of our knowledge - two orders of magnitude larger than the largest currently curated experimental dataset (PDF-5+ with approximately 20,000 experimental patterns) and even x orders of magnitude larger than the number of unlabeled diffractograms in the opXRD dataset we present here.
+This is---to the best of our knowledge---two orders of magnitude larger than the largest currently curated experimental dataset (PDF-5+ with approximately 20,000 experimental patterns) and even x orders of magnitude larger than the number of unlabeled diffractograms in the opXRD dataset we present here.
 \\
@@
 high-entropy materials, perovskites, and commercial catalysts.
-Details about the experiments of contributing research groups are discussed in section $\eqref{sec:our_dataset}$, along with a detailed description of how to contribute further data and how to use the data.
+Details about the experiments of contributing research groups are discussed in Section~$\eqref{sec:our_dataset}$, along with a detailed description of how to contribute further data and how to use the data.
 \\
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/overview.png}
+    \includegraphics[width=\linewidth]{figures/overview.pdf}
     \caption{Overview of the data collection pipeline.
@@
 In detail, users can select their most common pXRD file types, the program lists all files of that type, and users can select or deselect certain folders or files for sharing.
-Selected contributions will be uploaded to opXRD, processed to a common file format, and - if wanted - published on Zenodo on behalf of the contributors, before becoming part of the opXRD database.
+Selected contributions will be uploaded to opXRD, processed to a common file format, and---if wanted---published on Zenodo on behalf of the contributors, before becoming part of the opXRD database.
 If labels are available, they can be shared with opXRD as well.
-Further details can be found on the opXRD website\footnote{\url{https://xrd.aimat.science/}}.
-\\
+Further details can be found on the opXRD website.\footnote{The opXRD website is \url{https://xrd.aimat.science/}.} \\
 

### 2024-12-13 chapters/03_existing_datasets.tex
Adjusted table footnote to wiley guidelines

@@
 To contextualize opXRD within the current environment of experimental powder diffraction data, the below list provides an overview of the largest crystal structure databases that offer access to experimental powder diffraction data.
-For an overview of these databases refer to Tab.$\eqref{tab:exp_databases}$ below.
+For an overview of these databases refer to Table$\eqref{tab:exp_databases}$ below.
 In this table, the column ``O.A.'' indicates whether or not the database is open access.
@@
 \begin{adjustbox}{center}
-\begin{tabular}{@{}llllllll@{}}
+\begin{NiceTabular}{@{}llllllll@{}}
 \toprule
@@
 Linus Pauling file                     & 21,700                    & \xmark          & \cmark        & \cmark        & \cmark        & \cmark          & 2002            \\
-Powder Diffraction File \tablefootnote{The PDF lists the Material Platform for Data Science (MPDS) as a database source.
+Powder Diffraction File \tabularnote[a)]{The PDF lists the Material Platform for Data Science (MPDS) as a database source.
 Since the MPDS is hosted by the Pauling File project, there is likely significant overlap in the experimental patterns available in the PDF and the Linus Pauling File.} & 20,800 & \xmark & \cmark & \cmark & \cmark & \partialcheck{52} & 1941 \\
@@
 \bottomrule
-\end{tabular}
+\end{NiceTabular}
 \end{adjustbox}}

### 2024-12-13 chapters/04_our_dataset.tex
Adjusted table footnote to wiley guidelines

@@
-In collaboration with several other institutions we have collected a dataset of diffractograms stemming from experiments, some of them labeled with corresponding structural information.
+33In collaboration with several other institutions we have collected a dataset of diffractograms stemming from experiments, some of them labeled with corresponding structural information.
 The following paragraphs detail the contribution of each research group.
@@
 
-The below table Tab.$\eqref{tab:merged}$ provides an overview of the opXRD database.
+The below table Table$\eqref{tab:merged}$ provides an overview of the opXRD database.
 The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'' respectively.
@@
 The patterns submitted by the LBNL are similar because they stem from in-situ recordings where several hundred or several thousand patterns were collected over time per sample while they were undergoing physical processes.
-
-%The below histograms $\eqref{fig:histograms}$ show the distributions of spacegroups,
-
 \begin{figure}[!htb]
     \centering
-    \includegraphics[width=\linewidth]{figures/quantities_hist.png}
-    \caption{Histograms detailing spacegroups of labeled data and number of entries, starting angle value and final angle value for all data}
+    \includegraphics[width=0.95\linewidth]{figures/quantities_hist.png}
+    \caption{Histograms detailing spacegroups of labeled data and number of entries, starting angle value, and final angle value for all data}
     \label{fig:histograms}
@@
 
-%TODO:
-% The three datasets shown here are (1): USC CuNi, (2): USC CuAl, (3): HKUST
-% They certainly form clusters, but we have to think about how to interpret these results. It might still represent background/angle range more than anything else more interesting like material class.
-% Also if we add a larger dataset like the INT data the variance is so huge that all these three more narrow datasets effectively collapse onto a tiny region (see below)
-\begin{figure}[!htb]
-    \centering
-    \includegraphics[width=0.5\linewidth]{figures/pca_scatter_smol.png}
-    \caption{Two component PCA scatter plot showing (1): USC CuNi, (2): USC CuAl, (3): HKUST data2}
-    \label{fig:enter-label}
-\end{figure}
-
-\begin{figure}[!htb]
-    \centering
-    \includegraphics[width=0.5\linewidth]{figures/pca_scatter.png}
-    \caption{Two component PCA scatter plot including all opXRD data}
-    \label{fig:enter-label}
-\end{figure}
-
-The following paragraphs describe each contribution individually. \\
-
-\pagebreak
 %% Paragraph by FX Coudert and Arthur Hardiagon
@@
 \subsubsection*{Guangzhou Municipal Key Laboratory of Materials Informatics, HKUST(GZ)}
-
-%In the past two years, we established a small-scale experimental PXRD database called the X-Ray Phase Identification Public Experimental Dataset.
-The data in XRed is generated in our lab using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers under Cu targets.
-Hundreds of PXRD patterns have been refined, labeled with CIF files, and organized by elemental systems.
-The dataset includes original experimental files across single-phase to five-phase mixtures.
-This project is progressing toward integration with the comprehensive opXRD database to establish a large-scale, long-term experimental resource.
 
@@
 Combinatorial \ce{CuNi} and \ce{CuAl} samples were synthesized at the University of Southern California (USC) via magnetron co-sputtering from \ce{Cu} (99.999\%) and \ce{Ni} (99.995\%) or \ce{Al} (99.999\%) targets (Plasmaterials) onto stationary $10 \si{cm}$ \ce{Si} (100) substrates yielding 169 unique $5 \times 5 \ \si{mm}$ squares per wafer.
-These alloys were selected to study compositional effects on stacking fault and growth twin boundary formation in sputtered materials \cite{2024AcMat.27019839A,alwen2024combinatorial}.
+These alloys were selected to study compositional effects on stacking fault and growth twin boundary formation in sputtered materials \cite{alwen2024_1,alwen2024_2}.
 \\
@@
 The cobalt oxide (\ce{CoO} or \ce{Co3O4}) and cerium oxide (\ce{CeO2}) nanoparticles were in the size range of $4-16 \ \si{nm}$ according to the Scherrer equation.
-A series of porous \ce{Al2O3} materials, which were prepared by calcination of boehmite (\ce{AlOOH}) at various temperatures, represents crystalline samples with limited long-range structure and various contributions of \ce{Al2O3} polymorphs.\\
+A series of porous \ce{Al2O3} materials, which were prepared by calcination of boehmite (\ce{AlOOH}) at various temperatures, represents crystalline samples with limited long-range structure and various contributions of \ce{Al2O3} polymorphs.
+\\
 
@@
 The patterns were acquired in the $2\theta$ range of $5-80\degree$ with a step size of $0.016711\degree$ or $0.033420\degree$ and a total acquisition time of 40 to 120 min.
+This study has been carried out with the support of Angelina Barthelmeß, Elisabeth Herzinger, and Henning Hinrichs.
 \\
@@
 
-Data collection was performed in situ during thin-film deposition using a custom-made spin-coating and annealing stage.
-One dataset was collected from spin-coating triple cation metal-halide perovskite precursor solutions with the composition
-\ce{Cs}$_{0.05}$(MA$_{0.23}$FA$_{0.77}$)\ce{Pb}$_{1.1}$(\ce{I}$_{0.77}$\ce{Br}$_{0.23}$)$_{3}$ (MA = Methylammonium, FA = Formamidinium) onto various substrates, including glass (amorphous), $\ce{GaAs}$ wafers (single crystalline) as well as stacks of glass/ITO, $\ce{GaAs}$/$\ce{Mo}$/$\ce{Cu}$($\ce{In}$,$\ce{Ga}$)$\ce{Se2}$/$\ce{CdS}$/$\ce{ZnO}$, and glass/$\ce{Mo}$/$\ce{Cu}$($\ce{In}$,$\ce{Ga}$)$\ce{Se2}$/$\ce{CdS}$/$\ce{ZnO}$ (glass/CIGS).
+Data was collected from two separate studies.
+A first study collected data from spin-coating metal-halide perovskite precursor solutions.
+For this study, two datasets were collected.
+The first dataset was collected from spin-coating triple-cation metal-halide perovskite precursor solutions with the composition \ce{Cs}$_{0.05}$(MA$_{0.23}$FA$_{0.77}$)\ce{Pb}$_{1.1}$(\ce{I}$_{0.77}$\ce{Br}$_{0.23}$)$_{3}$ onto various substrates.
+Here, MA stands for Methylammonium and FA stands for Formamidinium.
+The substrates onto which these solutions were coated include glass, which is amorphous, and $\ce{GaAs}$ wafers, which are single crystalline.
+Other substrates were stacks of glass/indium tin oxide, stacks of \ce{GaAs}/CIGS, and stacks of glass/CIGS.
+Here, CIGS stands for a stack of \ce{Mo}, \ce{Cu}(\ce{In}, \ce{Ga})\ce{Se2},\ce{Cds} and \ce{ZnO}.
 Some of the substrates were additionally covered with a self-assembling monolayer of MeO-2PACz.
@@
 Christian Kaufmann and his team at Helmholtz-Zentrum Berlin (HZB) in Germany.
-A second dataset was collected from spin-coating metal-halide perovskite precursor solutions with varying compositions of \ce{MAPb(I_{1-x}Br_x)3} (where MA = Methylammonium and x = 0, 0.33, 0.5, 0.67, 1) spin-coated onto glass substrates.
-The substrates were preheated to different temperatures ($30\degree $C, $50\degree $C, $70\degree $C, and $90\degree $C), and the spin-coating process was performed at a constant temperature on the preheated substrates.
-For both datasets, diffraction data were continuously measured during spin-coating, chemical induction of crystallization, and annealing of the samples (at $100\degree$C and $110\degree$C respectively) with a frequency of about $0.56
+Data collection was performed in situ during thin-film deposition using a custom-made spin-coating and annealing stage.
+\\
+
+A second dataset was collected from spin-coating metal-halide perovskite precursor solutions with varying compositions of \ce{MAPb(I_{1-x}Br_x)3} spin-coated onto glass substrates.
+Here, MA = Methylammonium and x = 0, 0.33, 0.5, 0.67, 1.
+The substrates were preheated to different temperatures including $30\degree $C, $50\degree $C, $70\degree $C, and $90\degree $C, and the spin-coating process was performed at a constant temperature on the preheated substrates.
+For both datasets, diffraction data were continuously measured during spin-coating, chemical induction of crystallization, and annealing of the samples, at $100\degree$C and $110\degree$C respectively.
+The diffraction data was recorded with a frequency of about $0.56
  \ 1/\si{s}$ and $0.54 \ 1/\si{s}$.
@@
 
-Experimental XRD data were collected at beamline 12.3.2 of the Advanced Light Source, the synchrotron at Lawrence Berkeley National Laboratory.
+For this study, experimental XRD data were collected at beamline 12.3.2 of the Advanced Light Source, the synchrotron at Lawrence Berkeley National Laboratory.
 The data were collected using a photon energy of 10 $\si{keV}$ ($\lambda = 1.23984 \text{\AA}$), selected using a Si(111) monochromator.
@@
 
-In a second study, the phase evolution of the Mn-Sb-O system varying with annealing temperature was measured using the \textit{in situ}  Rigaku-SmartLab3kW diffractometer.
+In a second study, the phase evolution of the Mn-Sb-O system varying with annealing temperature was measured.
+The temperatures used to analyze the crystal structure of the Mn-Sb-O system were chosen depending on the number of phase transitions appearing for a certain temperature range.
+Few changes in the crystal structure appear between room temperature and 300$\degree$C and phase transitions appeared from 300$\degree$C until 850$\degree$C.
+No phase transition appeared when cooling down.
+Therefore, the crystal structure was measured every 100$\degree$C between room temperature and 300$\degree$C; every 50$\degree$C between 300$\degree$C and 850$\degree$C; and every 200$\degree$C when cooling down.
+The heating and cooling rates are fixed for all the experiments at 50$\degree$C/min and the holding time is fixed at 2 min.
+\\
+
+The data was collected using the \textit{in situ}  Rigaku-SmartLab3kW diffractometer.
 This tool operates with SmartLab Studio II software, which can measure the X-ray diffraction during the annealing process.
@@
 Phase transitions are analyzed with the \textit{in situ} XRD tool up to 850$\degree$C in this work.
-Most of the \textit{in situ} experiments were performed under an air-like 20\% \ce{O2} and 80\% argon (Ar) environment is chosen (Ar flow: 50 \si{sccm}, O$_2$ flow: 10 \si{sccm}).
+Most of the \textit{in situ} experiments were performed under an air-like 20\% \ce{O2} and 80\% \ce{Ar} environment is chosen (\ce{Ar} flow: 50 \si{sccm}, \ce{O2} flow: 10 \si{sccm}).
 When a 100\% Ar environment is fixed, an Ar flow of 60 \si{sccm} is input.
 The Bragg-Brentano (BB) mode is preferred in terms of geometry because it is more adapted in the analysis of scarce phases such as  \ce{MnSb2O6} rutile.
-The angular step is 0.01$\degree$ and the scanning rate is 10$\degree$/min.
-The step gives information about the resolution and the scanning rate (or speed duration time) results in the quality of the signal-to-noise ratio.
-\\
-
-The temperatures used to analyze the crystal structure of the Mn-Sb-O system were chosen depending on the number of phase transitions appearing for a certain temperature range.
-Few changes in the crystal structure appear between RT and 300$\degree$C and phase transitions appeared from 300$\degree$C until 850$\degree$C.
-No phase transition appeared when cooling down.
-Therefore, the crystal structure was measured every 100$\degree$C between RT and 300$\degree$C; every 50$\degree$C between 300$\degree$C and 850$\degree$C; and every 200$\degree$C when cooling down.
-The heating and cooling rates are fixed for all the experiments at 50$\degree$C/min and the holding time is fixed at 2 min.
+The angular step used in the recording was 0.01$\degree$ and the scanning rate was 10$\degree$/min.
 

### 2024-12-13 main.tex
Updated histogram
Adjusted table footnote to wiley guidelines

@@
 \author[\INT]{Jona Östreicher}
+\author[\ITI,\INT]{Jonas Teufel}
 \author[\HKUSTGZ]{Bin Cao}
@@
 \affil[\CPT]{Chimie ParisTech, PSL University, CNRS, Institut de Recherche de Chimie Paris, 75005 Paris, France}
-
 \affil[\KFUPMMSE]{Material Science and Engineering Department, King Fahd University of Petroleum and Minerals (KFUPM), Dhahran 31261, Saudi Arabia}
@@
 
+\subsection*{Supporting Information}
+Supporting Information is available from the Wiley Online Library or the author.
+
 \subsection*{Data availability}
 The opXRD database is available on Zenodo at \url{https://zenodo.org/records/14254270}.
-It can be downloaded by any user without any barriers or restrictions.
-For further details, please refer to section $\eqref{sec:how_to_use}$.
+It is published under the Creative Commons Attribution 4.0 International license.\footnote{Detailed terms of this license can be found at \url{https://creativecommons.org/licenses/by/4.0/deed.en}.} It can be downloaded by any user without any barriers or restrictions.
+For further details, please refer to Section~$\eqref{sec:how_to_use}$.
 
@@
 
-%TODO: Include acknowledgements to the following researchers from the Wolf group
-%The author list is fine, unfortunately I did not get real support from my team to provide a bigger consistent dataset.
-Please acknowledge the experimental contributions of Angelina Barthelmeß, Elisabeth Herzinger, Henning Hinrichs.
+%\subsection*{Table of contents}
+%\tableofcontents
+
+%A short text and graphic should be provided for the Table of Contents (ToC).
+The ToC text should describe the main results in 50 to 60 words.
+It should be written for a general audience and be written in the third person.
+
+%The ToC figure should convey the main message of the article.
+It does not have to be a figure from the article; it can be a combination of figures or a new, original figure composed to represent the topic.
+The author must be the copyright holder for this figure and any images used to create it.
+The size of the image should be either 55 mm × 50 mm (w × h) or 110 mm × 20 mm (w × h).
 

### 2024-12-13 chapters/00_abstract.tex
Adjusted table footnote to wiley guidelines
Recommended cuts to abstract

@@
 However, models trained on simulated pXRD patterns showed limited generalization ability when applied to experimental patterns, particularly for low-quality experimental patterns with high noise levels and elevated background.
-%Therefore, developing an experimental database offers at least two key advantages: (1) providing a comprehensive benchmark dataset to clarify model performance in practical applications, and (2) advancing phase identification tasks by supplying target-domain data within a transfer learning framework.
-With the Open Experimental Powder X-Ray Diffraction Database (opXRD), we aim to provide an openly available and easily accessible dataset of both labeled and unlabeled experimental powder diffractograms, providing machine learning researchers with a large quantity of real experimental diffractograms collected from a broad range of materials systems.
-The labeled data can be used as a benchmark to evaluate the performance of models on real experimental data and the unlabeled data can help improve the performance of models on experimental data, e.g.
+With the Open Experimental Powder X-Ray Diffraction Database (opXRD), we provide an openly available and easily accessible dataset of both labeled and unlabeled experimental powder diffractograms.
+%, providing machine learning researchers with a large quantity of real experimental diffractograms collected from a broad range of materials systems
+Labeled opXRD data can be used to evaluate the performance of models on real experimental data and unlabeled opXRD data can help improve the performance of models on experimental data, e.g.
 through transfer learning methods.
-%We provide almost barrier-free software tools that allow experimental researchers to find their data and make it accessible - in virtually any widely used format.
 We collected XXX labeled and YYY unlabeled diffractograms from a wide spectrum of materials classes, which establishes the first version of the opXRD database.
-We hope that this ongoing effort can guide machine learning research toward fully automated analysis of pXRD data and thus enable future self-driving materials labs.
+%We hope that this ongoing effort can guide machine learning research toward fully automated analysis of pXRD data and thus enable future self-driving materials labs.

### 2024-12-13 chapters/04_our_dataset.tex
Recommended cuts to abstract

@@
-33In collaboration with several other institutions we have collected a dataset of diffractograms stemming from experiments, some of them labeled with corresponding structural information.
+In collaboration with several other institutions we have collected a dataset of diffractograms stemming from experiments, some of them labeled with corresponding structural information.
 The following paragraphs detail the contribution of each research group.
@@
 
-The below table Table$\eqref{tab:merged}$ provides an overview of the opXRD database.
+The below Table$\eqref{tab:merged}$ provides an overview of the opXRD database.
 The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'' respectively.

### 2024-12-19 chapters/01_introduction.tex
Updated research projects in overview table

@@
     \includegraphics[width=1.0\linewidth]{figures/figure1_v5.pdf}
-    \caption{Overview}
+    \caption{Experimental powder x-ray diffraction (pXRD) patterns from several contributors are collected in the \textit{opXRD} database.
+The proposed open-access database of experimental data aims to support each step in the pXRD-related machine learning workflow by informing better physics simulations, supplying model training data, and providing a foundation for realistic performance evaluations.}
     \label{fig:overview}
@@
 
-However, even small labeled and also unlabeled experimental diffractograms hold significant value in practice.
+However, even small labeled and also unlabeled datasets of experimental diffractograms hold significant value in practice.
 First of all, labeled data can be used to test and benchmark existing and new automated analysis approaches, which yields a better comparison than when using the very limited currently available datasets, such as the commonly used RRUFF mineral dataset\cite{Armbruster2015}.

### 2024-12-19 chapters/03_existing_datasets.tex
Updated research projects in overview table

@@
 To contextualize opXRD within the current environment of experimental powder diffraction data, the below list provides an overview of the largest crystal structure databases that offer access to experimental powder diffraction data.
-For an overview of these databases refer to Table$\eqref{tab:exp_databases}$ below.
-In this table, the column ``O.A.'' indicates whether or not the database is open access.
-The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'' respectively.
+For an overview of these databases refer to Table~$\eqref{tab:exp_databases}$ below.
 
@@
 \centering
-\caption{Overview of experimental powder diffraction databases}
+\caption{\footnotesize Overview of experimental powder diffraction databases: The column ``O.A.'' indicates whether or not the database is open access.
+The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'' respectively.}
 \label{tab:exp_databases}

### 2024-12-19 chapters/04_our_dataset.tex
Updated research projects in overview table

@@
 We filtered the submitted datasets to exclude patterns with invalid features such as the pattern containing only one unique recorded angle or negative angles, the pattern containing less than 50 recorded angles total, or all intensities being zero.
-\\
-
 The below Table$\eqref{tab:merged}$ provides an overview of the opXRD database.
-The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'' respectively.
 
@@
 \centering
-\caption{The opXRD database}
+\caption{\footnotesize The opXRD database: The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'' respectively.}
 \label{tab:merged}
-\scalebox{0.81}{
+\scalebox{0.79}{
 \begin{adjustbox}{center}
@@
 \midrule
-CNRS        &  1052          & \cmark   & \partialcheck{85}   & \cmark   &  \partialcheck{85}   & Extracted from the COD \\ 
-HKUST(GZ)   &  21           & \cmark   & \cmark   & \cmark   & \cmark   & Phase identification reference database \\ 
+CNRS        &  1052          & \cmark   & \partialcheck{85}   & \cmark   &  \partialcheck{85}   & Diffraction data extracted from COD \\ 
+HKUST(GZ)   &  21+499          & \cmark   & \cmark   & \cmark   & \cmark   & Phase/structure identification dataset  \\ 
 USC         &  338          & \cmark   & \cmark   & \partialcheck{90}   & \xmark   & Study of CuNi and CuAl alloys          \\ 
@@
 IKFT        &  64           & \xmark   & \xmark   & \xmark   & \xmark   & Commercial catalysts, bulk materials   \\ 
-LBNL        &  n/a          & \xmark   & \xmark   & \xmark   & \xmark   & Time series recordings of Perovskites  \\ 
+LBNL        &  n/a          & \xmark   & \xmark   & \xmark   & \xmark   & Perovskites precursors, Mn-Sb-O system  \\ 
 \bottomrule
@@
 A small-scale experimental powder X-ray database called the X-Ray Phase Identification Public Experimental Dataset (XRed) was established over the past two years.\footnote{XRed can be accessed on GitHub at \url{https://github.com/WPEM/XRED}.} The primary objective of this database is to support the development of intelligent phase identification technology, serving as a foundation for data collection for future large-scale machine learning calculations.
-\\
-
-The data in XRed were generated in our laboratory using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers, all utilizing \ce{Cu} X-ray sources.
+The data in XRed were generated using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers, all utilizing \ce{Cu} X-ray sources.
 Hundreds of pXRD patterns have been obtained through XRD refinement, enabling the determination of detailed lattice constants, atomic sizes, and other structural parameters.
@@
 \\
+
+Another experimental database integrated with opXRD is the "collected database", which comprises powder XRD data paired with corresponding crystal structures.
+The data were sourced from publicly available open-access publications and our collaborating institutions, with full authorization for research use.
+This database surpasses Xred in size, containing 499 entries.
+However, these entries have not undergone individual verification via pattern refinement.
+In most cases, the XRD patterns are labeled based on powder diffraction file cards rather than detailed pattern refinements.\\
 

### 2024-12-19 chapters/05_how_to_use.tex
Updated research projects in overview table

@@
 The opXRD database is hosted on Zenodo\footnote{The dataset can be found at \url{https://zenodo.org/records/14278656}.} and can be downloaded by any user without any barriers or restrictions.
-The database comes in two files, ``opxrd.zip'' and ``opxrd\_in\_situ.zip''.
+The database comes in two zip archives, ``opxrd.zip'' and ``opxrd\_in\_situ.zip''.
 The latter contains the in-situ data which contains highly correlated patterns recorded through time series measurements.
-Patterns belonging to time series measurements have filenames that indicate the measurement series they belong to as well as their order in that series.
-\\
+Within the .zip archives patterns are saved as .json files grouped in folders indicating the contributing institution.
+Each .json file contains a pattern recorded from an X-ray diffraction experiment.
+If available, the composition and structure of the investigated sample and experiment conditions are also included in this file.
+Patterns belonging to time series measurements are labeled with filenames that indicate the measurement series they belong to as well as their order in that series.
 

### 2024-12-19 main.tex
Recommended cuts to abstract
Updated research projects in overview table

@@
 \newcommand{\partialcheck}[1]{\textcolor{orange}{\ding{52}(#1\%)}}
+\captionsetup[table]{
+  font=footnotesize
+}
 
@@
 \newcommand{\red}[1]{\textcolor{red}{#1}}
+\renewcommand\Authfont{\small\raggedright\itshape}  
+\renewcommand\Affilfont{\normalfont\small}
+
 \DeclareSIUnit{\myeuro}{\text{\euro}}
@@
 \author[\ITI,\INT]{Henrik Schopmans}
-\author[\INT]{Jona Östreicher}
+\author[\ITI,\INT]{Jona Östreicher}
 \author[\ITI,\INT]{Jonas Teufel}
@@
 \author[\LBLMF]{Carolin M. Sutter-Fella}
-\author[\ITI,\INT]{Pascal Friederich}
+\author[\ITI,\INT]{Pascal Friederich$^\ast$}
 
-\renewcommand\Affilfont{\small}
-
-\affil[\ITI]{Institute of Theoretical Informatics, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany}
+\affil[\ITI]{Institute of Theoretical Informatics, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany.  E-mail: pascal.friederich@kit.edu}
 \affil[\INT]{Institute of Nanotechnology, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany}
@@
 \maketitle
+\textbf{Keywords}: powder x-ray diffraction, machine learning, open-access database, crystal structure determination, phase identification
+
 \begin{abstract}

### 2024-12-22 chapters/04_our_dataset.tex
Removed mention of second HKUST dataset which is unpublishable

--- 2024-12-13 chapters/04_our_dataset.tex
@@
 \midrule
-CNRS        &  1052          & \cmark   & \partialcheck{85}   & \cmark   &  \partialcheck{85}   & Extracted from the COD \\ 
-HKUST(GZ)   &  21           & \cmark   & \cmark   & \cmark   & \cmark   & Phase identification reference database \\ 
+CNRS        &  1052          & \cmark   & \partialcheck{85}   & \cmark   &  \partialcheck{85}   & Diffraction data extracted from the COD \\ 
+HKUST(GZ)   &  21          & \cmark   & \cmark   & \cmark   & \cmark   & Phase/structure identification dataset  \\ 
 USC         &  338          & \cmark   & \cmark   & \partialcheck{90}   & \xmark   & Study of CuNi and CuAl alloys          \\ 
-EMPA        &  790          & \cmark   & \partialcheck{63}   & \xmark   & \xmark   & Tin halide perovskites, Zn-V-N library \\ 
+EMPA        &  790          & \cmark   & \partialcheck{63}   & \xmark   & \xmark   & Tin halide perovskites, Zn-V-N libraries \\ 
 INT         &  19,780        & \xmark   & \xmark   & \xmark   & \xmark   & Compilation of various projects        \\ 
 IKFT        &  64           & \xmark   & \xmark   & \xmark   & \xmark   & Commercial catalysts, bulk materials   \\ 
-LBNL        &  n/a          & \xmark   & \xmark   & \xmark   & \xmark   & Time series recordings of Perovskites  \\ 
+LBNL        &  n/a          & \xmark   & \xmark   & \xmark   & \xmark   & Perovskites precursors, Mn-Sb-O system  \\ 
 \bottomrule
@@
 A small-scale experimental powder X-ray database called the X-Ray Phase Identification Public Experimental Dataset (XRed) was established over the past two years.\footnote{XRed can be accessed on GitHub at \url{https://github.com/WPEM/XRED}.} The primary objective of this database is to support the development of intelligent phase identification technology, serving as a foundation for data collection for future large-scale machine learning calculations.
-\\
-
-The data in XRed were generated in our laboratory using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers, all utilizing \ce{Cu} X-ray sources.
+The XRed dataset primarily focuses on metal and metal-oxide particles, with data generated using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers, all utilizing \ce{Cu} X-ray sources.
 Hundreds of pXRD patterns have been obtained through XRD refinement, enabling the determination of detailed lattice constants, atomic sizes, and other structural parameters.
@@
 
-Additionally, Xqueryer,\footnote{Xqueryer can be accessed at \url{https://xqueryer.caobin.asia/}.} a cutting-edge free online platform for intelligent structure identification has been launched.
+Xqueryer,\footnote{Xqueryer can be accessed at \url{https://xqueryer.caobin.asia/}.} a recently launched cutting-edge free online platform for intelligent structure identification, will also serve as a source of data for the opXRD database.
 Users can upload experimental pXRD data to identify crystal structures and extract all relevant information available in the Materials Project database\cite{jain2013commentary}.
@@
 
-Each combinatorial Zn–V–N library was synthesized using radio-frequency co-sputtering of Zn and V in a mixed Ar and \ce{N2} plasma.
+Combinatorial Zn–V–N libraries were synthesized using radio-frequency co-sputtering of Zn and V in a mixed Ar and \ce{N2} plasma.
 An orthogonal deposition temperature and composition gradient was created, resulting in a deposition temperature of $220\degree$C for samples 1~–~9 and $114\degree$C for samples 37~–~45.
@@
 
+X-ray diffraction data was collected from a wide range of research projects conducted at the Institute of Nanotechnology over the past years.
 A major part of the research focused on high-entropy materials, which involved incorporating many different elements into single-phase structures, leading to peak shifts or phase separations.

### 2024-12-22 chapters/00_abstract.tex
Deleted older figure versions

@@
 A notable difficulty in applying machine learning to this domain is the lack of sufficiently sized experimental datasets, which has constrained machine learning researchers to train primarily on simulated data.
-However, models trained on simulated pXRD patterns showed limited generalization ability when applied to experimental patterns, particularly for low-quality experimental patterns with high noise levels and elevated background.
+However, models trained on simulated pXRD patterns showed limited generalization ability when applied to experimental patterns, particularly for low-quality experimental patterns with high noise levels and elevated backgrounds.
 With the Open Experimental Powder X-Ray Diffraction Database (opXRD), we provide an openly available and easily accessible dataset of both labeled and unlabeled experimental powder diffractograms.

### 2024-12-22 chapters/04_our_dataset.tex
Deleted older figure versions

@@
 CNRS        &  1052          & \cmark   & \partialcheck{85}   & \cmark   &  \partialcheck{85}   & Diffraction data extracted from the COD \\ 
-HKUST(GZ)   &  21          & \cmark   & \cmark   & \cmark   & \cmark   & Phase/structure identification dataset  \\ 
+HKUST(GZ)   &  21+499          & \cmark   & \cmark   & \cmark   & \cmark   & Phase/structure identification dataset  \\ 
 USC         &  338          & \cmark   & \cmark   & \partialcheck{90}   & \xmark   & Study of CuNi and CuAl alloys          \\ 
@@
     \centering
-    \includegraphics[width=0.95\linewidth]{figures/quantities_hist.png}
-    \caption{Histograms detailing spacegroups of labeled data and number of entries, starting angle value, and final angle value for all data}
+    \includegraphics[width=0.85\linewidth]{figures/hist.png}
+    \caption{Histograms detailing the distribution of properties in the opXRD dataset: a) distribution of spacegroups present in labeled data; b) distribution of angular resolution in all data; c) distribution of smallest and largest recorded $2\theta$ values for all data.}
     \label{fig:histograms}
@@
 \\
+
+Another experimental dataset integrated with opXRD is a collection of powder diffraction data sourced from open-access publications and collaborating institutions.
+All collaborating institutions provided this data with full authorization for research purposes.
+This dataset provides broader coverage of chemical elements and includes ionic crystals, atomic crystals, and metallic crystals.
+This dataset surpasses Xred in size, containing 499 entries.
+However, contrary to Xred these data do not come paired with corresponding crystal structures.
 
@@
 Data was collected from two separate studies.
-A first study collected data from spin-coating metal-halide perovskite precursor solutions.
+A first study focused on metal-halide perovskite precursor solutions.
 For this study, two datasets were collected.

### 2024-12-22 main.tex
Deleted older figure versions

@@
 
-\affil[\ITI]{Institute of Theoretical Informatics, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany.  E-mail: pascal.friederich@kit.edu}
-\affil[\INT]{Institute of Nanotechnology, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany}
-\affil[\HKUSTGZ]{Guangzhou Municipal Key Laboratory of Materials Informatics, Hong Kong University of Science and Technology (Guangzhou), Guangzhou 511400, China}
-\affil[\USC]{Department of Chemical Engineering and Materials Science, University of Southern California, Los Angeles CA 90089, USA}
-\affil[\LBLMF]{Lawrence Berkeley National Laboratory, Molecular Foundry Division, Berkeley 94720 CA, USA}
-\affil[\LBLALS]{Lawrence Berkeley National Laboratory, Advanced Light Source, Berkeley 94720 CA, USA}
+\affil[\ITI]{Institute of Theoretical Informatics, Karlsruhe Institute of Technology (KIT), 76131 Karlsruhe, Germany.  E-mail: pascal.friederich@kit.edu}
+\affil[\INT]{Institute of Nanotechnology, Karlsruhe Institute of Technology (KIT), 76131 Karlsruhe, Germany}
+\affil[\HKUSTGZ]{Guangzhou Municipal Key Laboratory of Materials Informatics, Hong Kong University of Science and Technology (Guangzhou) (HKUST), Guangzhou 511400, China}
+\affil[\USC]{Department of Chemical Engineering and Materials Science, University of Southern California (USC), Los Angeles CA 90089, USA}
+\affil[\LBLMF]{Molecular Foundry Division, Lawrence Berkeley National Laboratory (LBNL), Berkeley 94720 CA, USA}
+\affil[\LBLALS]{Advanced Light Source, Lawrence Berkeley National Laboratory,  Berkeley 94720 CA, USA}
 \affil[\Hoffmann]{Hoffmann Institute of Advanced Materials, Shenzhen Polytechnic, Shenzhen 518055, China}
-\affil[\LBLCS]{Lawrence Berkeley National Laboratory, Chemical Sciences Division, Berkeley 94720 CA, USA}
-\affil[\EMPA]{Empa–Swiss Federal Laboratories for Materials Science and Technology, 8600 Dübendorf, Switzerland}
-\affil[\KITICRT]{Engler-Bunte-Institut \& Institute of Catalysis Research and Technology, Karlsruhe Institute of Technology, Karlsruhe, Germany}
-\affil[\CPT]{Chimie ParisTech, PSL University, CNRS, Institut de Recherche de Chimie Paris, 75005 Paris, France}
+\affil[\LBLCS]{Lawrence Berkeley National Laboratory (LBNL), Chemical Sciences Division, Berkeley 94720 CA, USA}
+\affil[\EMPA]{Empa–Swiss Federal Laboratories for Materials Science and Technology (EMPA), 8600 Dübendorf, Switzerland}
+\affil[\KITICRT]{Engler-Bunte-Institut \& Institute of Catalysis Research and Technology, Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany}
+\affil[\CPT]{Chimie ParisTech, Paris Sciences et Lettres University University (PSL), CNRS, Institut de Recherche de Chimie Paris, 75005 Paris, France}
 \affil[\KFUPMMSE]{Material Science and Engineering Department, King Fahd University of Petroleum and Minerals (KFUPM), Dhahran 31261, Saudi Arabia}
-\affil[\KFUPMIRCIMR]{Interdisciplinary Research Center for Intelligent Manufacturing and Robotics, KFUPM, Dhahran 31261, Saudi Arabia}
+\affil[\KFUPMIRCIMR]{Interdisciplinary Research Center for Intelligent Manufacturing and Robotics, King Fahd University of Petroleum and Minerals (KFUPM), Dhahran 31261, Saudi Arabia}
 \affil[\FZJIEMD]{Institute of Energy Materials and Devices, Forschungszentrum Juelich GmbH, 52425 Juelich, Germany}

### 2024-12-22 chapters/04_our_dataset.tex
Added alphabetic labeling rule in Usage section

@@
     \centering
-    \includegraphics[width=0.85\linewidth]{figures/hist.png}
+    \includegraphics[width=0.8\linewidth]{figures/hist.png}
     \caption{Histograms detailing the distribution of properties in the opXRD dataset: a) distribution of spacegroups present in labeled data; b) distribution of angular resolution in all data; c) distribution of smallest and largest recorded $2\theta$ values for all data.}
@@
 \end{figure}
+
+\pagebreak
 
@@
 
-Data was collected from two separate studies.
-A first study focused on metal-halide perovskite precursor solutions.
-For this study, two datasets were collected.
+In total three different datasets were collected.
+The first and second datasets focused on metal-halide perovskite precursor solutions.
 The first dataset was collected from spin-coating triple-cation metal-halide perovskite precursor solutions with the composition \ce{Cs}$_{0.05}$(MA$_{0.23}$FA$_{0.77}$)\ce{Pb}$_{1.1}$(\ce{I}$_{0.77}$\ce{Br}$_{0.23}$)$_{3}$ onto various substrates.
@@
 
-For this study, experimental XRD data were collected at beamline 12.3.2 of the Advanced Light Source, the synchrotron at Lawrence Berkeley National Laboratory.
+For these two datasets, experimental XRD data were collected at beamline 12.3.2 of the Advanced Light Source, the synchrotron at Lawrence Berkeley National Laboratory.
 The data were collected using a photon energy of 10 $\si{keV}$ ($\lambda = 1.23984 \text{\AA}$), selected using a Si(111) monochromator.
@@
 
-In a second study, the phase evolution of the Mn-Sb-O system varying with annealing temperature was measured.
+A third dataset was collected by observing the phase evolution of an Mn-Sb-O system with varying annealing temperatures.
 The temperatures used to analyze the crystal structure of the Mn-Sb-O system were chosen depending on the number of phase transitions appearing for a certain temperature range.

### 2024-12-22 main.tex
Added alphabetic labeling rule in Usage section

@@
 
+\pagebreak
+
 \subsection*{Supporting Information}
-Supporting Information is available from the Wiley Online Library or the author.
+%Supporting Information is available from the Wiley Online Library or the author.
+
+The below figure $\eqref{fig:combined}$ displays a randomly selected sample of the X-ray diffraction patterns found in each of the research projects contributed to the opXRD database.
+
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=\linewidth]{figures/combined.png}
+    \caption{50 randomly chosen X-ray diffraction patterns from each contributed dataset.
+The figure shows data from the following datasets: a) EMPA, b) LBNL-A, c) LBNL-B, d) LBNL-C, e) USC, f) INT, g) HKUST-A, h) HKUST-B, i) CNRS, j) IKFT.}
+    \label{fig:combined}
+\end{figure}
+
+\pagebreak
 

### 2024-12-22 chapters/04_our_dataset.tex
Added discussion of histogram

@@
 We filtered the submitted datasets to exclude patterns with invalid features such as the pattern containing only one unique recorded angle or negative angles, the pattern containing less than 50 recorded angles total, or all intensities being zero.
-The below Table$\eqref{tab:merged}$ provides an overview of the opXRD database.
+The below Table~$\eqref{tab:merged}$ provides an overview of the opXRD database.
 
@@
 The patterns submitted by the LBNL are similar because they stem from in-situ recordings where several hundred or several thousand patterns were collected over time per sample while they were undergoing physical processes.
+
+\pagebreak
+
+Figure~$\eqref{fig:histograms}$ below provides an overview of the distributions of pattern and structure properties in the opXRD dataset.
+Nearly all patterns have an angular resolution smaller than $\Delta(2\theta) = 0.1 ^\circ$.
+Here the angular resolution is defined as the range of recorded angles divided by the number of recorded intensity values along that range.
+For most patterns, the lowest recorded angle is smaller than $30 ^\circ$ and the highest recorded angle is smaller than $120 ^\circ$.
+
 \begin{figure}[!htb]
@@
     \includegraphics[width=0.8\linewidth]{figures/hist.png}
-    \caption{Histograms detailing the distribution of properties in the opXRD dataset: a) distribution of spacegroups present in labeled data; b) distribution of angular resolution in all data; c) distribution of smallest and largest recorded $2\theta$ values for all data.}
+    \caption{Histograms detailing the distribution of pattern and structure properties in the opXRD database: a) distribution of spacegroups present in labeled data; b) distribution of angular resolution in all data; c) distribution of smallest and largest recorded $2\theta$ values for all data.}
     \label{fig:histograms}

### 2024-12-22 chapters/04_our_dataset.tex
Centered checkmarks in overview table

@@
 \label{tab:merged}
-\scalebox{0.79}{
+\scalebox{0.75}{
 \begin{adjustbox}{center}
 \begin{tabular}{@{}lllllll@{}}
+%\begin{tabular}{@{}llccccl@{}}
 \toprule
@@
 CNRS        &  1052          & \cmark   & \partialcheck{85}   & \cmark   &  \partialcheck{85}   & Diffraction data extracted from the COD \\ 
-HKUST(GZ)   &  21+499          & \cmark   & \cmark   & \cmark   & \cmark   & Phase/structure identification dataset  \\ 
-USC         &  338          & \cmark   & \cmark   & \partialcheck{90}   & \xmark   & Study of CuNi and CuAl alloys          \\ 
-EMPA        &  790          & \cmark   & \partialcheck{63}   & \xmark   & \xmark   & Tin halide perovskites, Zn-V-N libraries \\ 
-INT         &  19,780        & \xmark   & \xmark   & \xmark   & \xmark   & Compilation of various projects        \\ 
-IKFT        &  64           & \xmark   & \xmark   & \xmark   & \xmark   & Commercial catalysts, bulk materials   \\ 
-LBNL        &  n/a          & \xmark   & \xmark   & \xmark   & \xmark   & Perovskites precursors, Mn-Sb-O system  \\ 
+HKUST(GZ)   &  21+499          & \cmark   & \cmark   & \cmark   & \cmark[\quad]   & Phase/structure identification dataset  \\ 
+USC         &  338          & \cmark   & \cmark   & \partialcheck{90}   & \xmark[\quad]   & Study of CuNi and CuAl alloys          \\ 
+EMPA        &  790          & \cmark   & \partialcheck{63}   & \xmark   & \xmark[\quad]   & Tin halide perovskites, Zn-V-N libraries \\ 
+INT         &  19,780        & \xmark   & \xmark   & \xmark   & \xmark[\quad]   & Compilation of various projects        \\ 
+IKFT        &  64           & \xmark   & \xmark   & \xmark   & \xmark[\quad]   & Commercial catalysts, bulk materials   \\ 
+LBNL        &  n/a          & \xmark   & \xmark   & \xmark   & \xmark[\quad]   & Perovskites precursors, Mn-Sb-O system  \\ 
 \bottomrule

### 2024-12-22 chapters/04_our_dataset.tex
Updated checkmark centering in overview tables

@@
 \toprule
-\textbf{Institution} & \textbf{No. patterns} & \textbf{Comp.} & \textbf{Spg.} & \textbf{Lattice} & \textbf{Atom coords.} & \textbf{Research Project} \\
+\textbf{Institution} & \textbf{No. patterns} & \ \textbf{Comp.} & \ \ \textbf{Spg.} & \textbf{Lattice} & \textbf{Atom coords.} & \textbf{Research Project} \\
 \midrule
-CNRS        &  1052          & \cmark   & \partialcheck{85}   & \cmark   &  \partialcheck{85}   & Diffraction data extracted from the COD \\ 
-HKUST(GZ)   &  21+499          & \cmark   & \cmark   & \cmark   & \cmark[\quad]   & Phase/structure identification dataset  \\ 
-USC         &  338          & \cmark   & \cmark   & \partialcheck{90}   & \xmark[\quad]   & Study of CuNi and CuAl alloys          \\ 
-EMPA        &  790          & \cmark   & \partialcheck{63}   & \xmark   & \xmark[\quad]   & Tin halide perovskites, Zn-V-N libraries \\ 
-INT         &  19,780        & \xmark   & \xmark   & \xmark   & \xmark[\quad]   & Compilation of various projects        \\ 
-IKFT        &  64           & \xmark   & \xmark   & \xmark   & \xmark[\quad]   & Commercial catalysts, bulk materials   \\ 
-LBNL        &  n/a          & \xmark   & \xmark   & \xmark   & \xmark[\quad]   & Perovskites precursors, Mn-Sb-O system  \\ 
+CNRS        &  1052          & \ \cmark   & \partialcheck{85}   & \cmark   &  \ \ \quad \partialcheck{85}   & Diffraction data extracted from the COD \\ 
+HKUST(GZ)   &  21+499          & \ \cmark   & \cmark   & \cmark   & \ \ \quad \cmark   & Phase/structure identification dataset  \\ 
+USC         &  338          & \ \cmark   & \cmark   & \partialcheck{90}   & \ \ \quad \xmark   & Study of CuNi and CuAl alloys          \\ 
+EMPA        &  790          &\ \cmark   & \partialcheck{63}   & \xmark   & \ \ \quad \xmark   & Tin halide perovskites, Zn-V-N libraries \\ 
+INT         &  19,780        &\ \xmark   & \xmark   & \xmark   & \ \ \quad \xmark   & Compilation of various projects        \\ 
+IKFT        &  64           &\ \xmark   & \xmark   & \xmark   & \ \ \quad \xmark   & Commercial catalysts, bulk materials   \\ 
+LBNL        &  n/a          &\ \xmark   & \xmark   & \xmark   & \ \ \quad \xmark   & Perovskites precursors, Mn-Sb-O system  \\ 
 \bottomrule

### 2024-12-23 chapters/04_our_dataset.tex
Added general introduction to PCA

@@
 
+The below figure~$\eqref{fig:components}$ illustrates how varied the data of each contribution is using principle component analysis (PCA).
+PCA is a dimensional reduction technique that can be applied to datasets $X \subset \mathbb{R}^n$ to reduce the number of components needed to describe points $\p \in X$ up to some tolerance in lost accuracy.
+The cumulative explained variance ratio is a measure indicating how much of the variance in the dataset $X$ can be explained using a given number of components.
+For example, if PCA on a dataset $X \subset \mathbb{R}^10$ yields a cumulative variance ratio of 99\%, then each point in the dataset can be represented using only 5 degrees of freedom up to a minor loss in accuracy.
+
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=\linewidth]{figures/component_fractions.png}
+    \caption{Explained variance ratio over the fraction of the maximum number of components for each dataset contributed to opXRD.
+Each dataset is named after the contributing institutions.
+Datasets contributed by the same institution are enumerated alphabetically in the order in which they are described below.
+}
+    \label{fig:components}
+\end{figure}
+
 %TODO: A paragraph on how data is correlated
@@
 The patterns submitted by the LBNL are similar because they stem from in-situ recordings where several hundred or several thousand patterns were collected over time per sample while they were undergoing physical processes.
+\\
 
@@
 
-\pagebreak
+The following texts describe the datasets contributed by each collaborating institution. Each text includes a description of the investigated materials and how X-ray diffraction data was collected.
 
@@
 However, contrary to Xred these data do not come paired with corresponding crystal structures.
+\\
 

### 2024-12-23 chapters/04_our_dataset.tex
Added description of PCA analysis and figure

--- 2024-12-22 chapters/04_our_dataset.tex
@@
 
+The variance of the data was analyzed using a dimensional reduction technique called principal component analysis (PCA).
+PCA can be applied to datasets $X \subset \mathbb{R}^N$ to reduce the number of components needed to describe points $p \in X$ up to some tolerance in lost accuracy.
+In the context of PCA, the cumulative explained variance ratio is a measure of how much of the variance in the dataset $X$ can be explained using a given number of components.
+For example, if performing PCA on a dataset $X \subset \mathbb{R}^{10}$ yields a cumulative variance ratio of 99\%, then each point in the dataset can be represented using only 5 degrees of freedom up to a minor loss in accuracy.
+For a rigorous definition of PCA and the explained variance ratio, we refer to the literature \cite{Jolliffe2016}.
+\\
+
+Here, PCA was performed on datasets of X-ray diffraction patterns.
+These datasets $X$ are subsets of $\mathbb{R}^{N}$ with $N=N_{\text{values}} =512$ since each pattern $p \in X$ was standardized to have 512 intensity values spread out evenly from $0\degree$ to $180\degree$ using zero padding and interpolation with cubic splines.
+Hence the maximal components that could be needed to describe a dataset of diffraction data in this context is $N_{\text{values}}=512$.
+However, the maximal number of components is even lower for datasets that contain less than 512 patterns.
+In this case, the maximal number of components is equal to the number of patterns in the dataset since each pattern can add at most one degree of freedom to the dataset $X \in \mathbb{R}^N$.
+Hence the maximum number of components $N_{\text{max}}$ of a pattern dataset $X$ is given as follows:
+\begin{align}
+    N_{\text{max}} = \min(N_{\text{values}}, N_{\text{patterns}}).
+\end{align}
+Here $N_{\text{values}} = N =512$ is the number of recorded intensity values per pattern and $N_{\text{patterns}}$ is the number of patterns in the dataset $X$.
+
+\pagebreak
+
+The below figure~$\eqref{fig:components}$ shows the cumulative explained variance ratio over the fraction of maximal No.
+components $N_{\text{max}}$ as defined above.
+In this figure, a faster convergence of the cumulative variance ratio towards one indicates that the patterns in this data are relatively similar.
+
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=\linewidth]{figures/component_fractions.png}
+    \caption{Explained variance ratio over the fraction of the maximum number of components for each dataset contributed to opXRD.
+Each dataset is named after the contributing institutions.
+Datasets contributed by the same institution are enumerated alphabetically in the order in which they are described below.
+}
+    \label{fig:components}
+\end{figure}
+
 %TODO: A paragraph on how data is correlated

### 2024-12-23 chapters/04_our_dataset.tex
Updated figurre/texxt formatting in our_dataset

@@
 Here $N_{\text{values}} = N =512$ is the number of recorded intensity values per pattern and $N_{\text{patterns}}$ is the number of patterns in the dataset $X$.
+Figure~$\eqref{fig:components}$ below shows the cumulative explained variance ratio over the fraction of maximal No.
+components $N_{\text{max}}$ as defined above.
+In this figure, a faster convergence of the cumulative variance ratio towards one indicates that the patterns in this data are relatively similar.
+\\
 
+The degree of variation between the patterns is different for each contribution.
+For example, the CRNS and the HKUST contributions each are collections that encompass many research projects over a large period of time and hence exhibit a high degree of variability between individual patterns.
+In contrast, the contributions by the USC and the LBNL contain many very similar patterns.
+The patterns in the USC datasets are similar because the underlying samples are all variations of \ce{CuNi} and \ce{CuAl} alloys collected on the same machine.
+The patterns submitted by the LBNL are similar because they stem from in-situ recordings where several hundred or several thousand patterns were collected over time per sample while they were undergoing physical processes.
 \pagebreak
 
-The below figure~$\eqref{fig:components}$ shows the cumulative explained variance ratio over the fraction of maximal No.
-components $N_{\text{max}}$ as defined above.
-In this figure, a faster convergence of the cumulative variance ratio towards one indicates that the patterns in this data are relatively similar.
+These effects can be observed in figure~$\eqref{fig:components}$: The cumulative explained variance ratios of the datasets contributed by the USC and LBNL converge towards one significantly faster than the datasets contributed by the CNRS and HKUST.
+\\
 
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/component_fractions.png}
+    \includegraphics[width=0.85\linewidth]{figures/component_fractions.png}
     \caption{Explained variance ratio over the fraction of the maximum number of components for each dataset contributed to opXRD.
 Each dataset is named after the contributing institutions.
-Datasets contributed by the same institution are enumerated alphabetically in the order in which they are described below.
-}
+Datasets contributed by the same institution are labeled alphabetically in the order in which they are described in the texts towards the end of this section.}
     \label{fig:components}
 \end{figure}
-
-%TODO: A paragraph on how data is correlated
-The degree of variation between the patterns is different for each contribution.
-For example, the CRNS and the INT contributions each are collections that encompass many research projects over a large period of time and hence exhibit a high degree of variability between individual patterns.
-In contrast, the contributions by the USC and the LBNL contain many very similar patterns.
-The patterns in the USC datasets are similar because the underlying samples are all variations of \ce{CuNi} and \ce{CuAl} alloys collected on the same machine.
-The patterns submitted by the LBNL are similar because they stem from in-situ recordings where several hundred or several thousand patterns were collected over time per sample while they were undergoing physical processes.
-\\
-
-\pagebreak
 

### 2025-03-01 chapters/00_abstract.tex
Commited Feburary changes

@@
 With the Open Experimental Powder X-Ray Diffraction Database (opXRD), we provide an openly available and easily accessible dataset of both labeled and unlabeled experimental powder diffractograms.
-%, providing machine learning researchers with a large quantity of real experimental diffractograms collected from a broad range of materials systems
 Labeled opXRD data can be used to evaluate the performance of models on real experimental data and unlabeled opXRD data can help improve the performance of models on experimental data, e.g.
 through transfer learning methods.
-We collected XXX labeled and YYY unlabeled diffractograms from a wide spectrum of materials classes, which establishes the first version of the opXRD database.
-%We hope that this ongoing effort can guide machine learning research toward fully automated analysis of pXRD data and thus enable future self-driving materials labs.
+We collected \numpatterns diffractograms - 2179 of them labeled - from a wide spectrum of materials classes, establishing the first version of the opXRD database.
+We hope that this ongoing effort can guide machine learning research toward fully automated analysis of pXRD data and thus enable future self-driving materials labs.

### 2025-03-01 chapters/01_introduction.tex
Deleted older figure versions
Commited Feburary changes

@@
     \centering
-    \includegraphics[width=1.0\linewidth]{figures/figure1_v5.pdf}
-    \caption{Experimental powder x-ray diffraction (pXRD) patterns from several contributors are collected in the \textit{opXRD} database.
+    \includegraphics[width=1.0\linewidth]{figures/pipeline.pdf}
+    \caption{Experimental powder X-ray diffraction (pXRD) patterns from several contributors are collected in the \textit{opXRD} database.
 The proposed open-access database of experimental data aims to support each step in the pXRD-related machine learning workflow by informing better physics simulations, supplying model training data, and providing a foundation for realistic performance evaluations.}
@@
 
-%\begin{figure}[!htb]
-%    \centering
-%    \includegraphics[width=0.6\linewidth]{figures/rietveld_flowchart.png}
-%    \caption{Enter Caption}
-%    \label{fig:enter-label}
-%\end{figure}
-
 Machine learning has the potential to speed up the manual analysis of powder diffractograms and keep pace with an automated high-throughput experimentation environment\cite{Agrawal2019, Surdu2023}.
@@
 % Predicting structure information with machine learning
-Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting lattice parameters\cite{Dong2021, Chitturi2021, Habershon2004, zhang2024crystallographic}, spacegroup \cite{Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021,zhang2024crystallographic}, and crystallite size \cite{Dong2021, Chakraborty2021} from simulated diffractograms.
-However, the performance substantially drops off when these models are applied to data originating from experiments \cite{Schopmans2023,zhang2024crystallographic, Wang2020, Vecsei2018} The patterns observed in experiments exhibit imperfections that do not appear when simulating the X-ray diffraction pattern of a powder sample under ideal conditions.
+Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting lattice parameters\cite{Dong2021, Chitturi2021, Habershon2004, zhang2024crystallographic}, spacegroup \cite{cao2024simxrd,Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021,zhang2024crystallographic}, and crystallite size \cite{Dong2021, Chakraborty2021} from simulated diffractograms.
+However, the performance substantially drops off when these models are applied to data originating from experiments \cite{cao2024simxrd,Schopmans2023,zhang2024crystallographic, Wang2020, Vecsei2018} The patterns observed in experiments exhibit imperfections that do not appear when simulating the X-ray diffraction pattern of a powder sample under ideal conditions.
 The physical effects underlying these imperfections need to be modeled accurately in order to be able to transfer models trained on simulated data to experimental patterns.
@@
 At the moment, the number of openly available datasets for experimental powder data is very limited.
-The main sources for experimental patterns are XXX, YYY, and ZZZ.
-Further commercial datasets such as KKK and LLL exist, but license fees and restrictions in their use to train publishable models or even publish the training data limit their usability.
+The main source for experimental patterns is the RRUFF database\cite{lafuente2015}.
+Further commercial datasets such as the PDF5+\cite{GatesRector2019} and the Linus Pauling File\cite{villars2018} exist, but license fees and restrictions in their use to train publishable models or even publish the training data limit their usability.
 Collecting a labeled dataset of raw experimental powder X-ray diffraction data of sufficient size to directly train a model that predicts crystal structure information is currently out of reach.
 The most advanced ML models currently are trained on approximately $10^5 - 10^6$ simulated diffractograms.
-\cite{Salgado2023, Schopmans2023}
-This is---to the best of our knowledge---two orders of magnitude larger than the largest currently curated experimental dataset (PDF-5+ with approximately 20,000 experimental patterns) and even x orders of magnitude larger than the number of unlabeled diffractograms in the opXRD dataset we present here.
-\\
+\cite{Salgado2023, Schopmans2023} This is, to the best of our knowledge, two orders of magnitude larger than the largest currently curated experimental dataset, the PDF-5+ with approximately $2\cdot 10^4$ experimental patterns.
+It is even one order of magnitude larger than the approximately $10^5$ unlabeled diffractograms in the initial version of the opXRD dataset we present here.\\
 
@@
 Features that may vary between experimental setups and influence the recorded pattern include the shape of diffraction peaks, the wavelength and polarization of the employed X-ray source, and the detector geometry
-\cite{Waseda2011, Pecharsky2023}.
-Some experimental setups may also exhibit imperfections such as displacement of the sample from its intended position leading to the recorded scattering angles being slightly off \cite{hulbert2023} or background events such as instrument noise or X-rays scattering off equipment onto the screen.\\
+\cite{cao2024simxrd, Waseda2011, Pecharsky2023}.
+Some experimental setups may also exhibit imperfections such as displacement of the sample from its intended position leading to the recorded scattering angles being slightly off \cite{cao2024simxrd,hulbert2023} or background events such as instrument noise or X-rays scattering off equipment onto the screen.\\
 
@@
 
-In this publication, we provide a new open powder X-ray diffraction (opXRD) dataset that collects a broad range of patterns from experiments, exceeding the currently available amount of openly accessible experimental patterns by XX.
-Our opXRD dataset has been curated by collecting the accumulated powder data from multiple large research groups and institutions with high-throughput XRD facilities.
-It contains pXRD patterns from single and multiphase materials from a wide variety of material classes, including
-high-entropy materials, perovskites, and commercial catalysts.
-Details about the experiments of contributing research groups are discussed in Section~$\eqref{sec:our_dataset}$, along with a detailed description of how to contribute further data and how to use the data.
-\\
+In this publication, we introduce an open powder X-ray diffraction (opXRD) database that collects a broad range of patterns from experiments.
+With a total of $92,552$ patterns, the opXRD database exceeds the size of the previously largest database of openly accessible experimental powder diffraction data by two orders of magnitude.
+To the best of our knowledge, up to now the largest database of this type was the RRUFF database, containing 1,290 experimental powder diffraction patterns.
+Our opXRD dataset has been curated by collecting the accumulated data from multiple research groups and institutions with high-throughput XRD facilities.
+It contains pXRD patterns from single and multiphase materials from a wide variety of material classes, including high-entropy materials, perovskites, and commercial catalysts.
+Details about the experiments of contributing research groups are discussed in Section~$\eqref{sec:our_dataset}$.
+A detailed description of how to use the data is given in Section~$\eqref{sec:how_to_use}$ and Section~$\eqref{sec:summary_and_outlook}$ describes how further data can be contributed.
 
@@
 If labels are available, they can be shared with opXRD as well.
-Further details can be found on the opXRD website.\footnote{The opXRD website is \url{https://xrd.aimat.science/}.} \\
+Further details can be found on the opXRD website (\url{https://xrd.aimat.science/}).
+\\
 
@@
 This makes the data more easily accessible to other researchers and provides more value to researchers seeking large quantities of data compared to an isolated submission on Zenodo.
-However, further data annotation with metadata is required to fulfill the FAIR data principles.\\
+However, further data annotation with metadata is required to fully fulfill the FAIR data principles.\\
 

### 2025-03-01 chapters/02_SOTA_ML.tex
Added partial percentages in existing dataset table
Commited Feburary changes

@@
-\paragraph{ML for pXRD}
-
 To showcase the need for datasets such as the one presented in this publication, we now shortly discuss some recent
@@
 
-The option of training purely on experimental diffractograms unfortunately is unfeasible because of the limited availability of labeled experimental diffractograms, but Salgado {\it et al.} \cite{Salgado2023} showed that adding a fraction of experimental patterns to a simulated training dataset improves the performance on unseen experimental patterns.
+The option of training purely on experimental diffractograms unfortunately is infeasible because of the limited availability of labeled experimental diffractograms, but Salgado {\it et al.} \cite{Salgado2023} showed that adding a fraction of experimental patterns to a simulated training dataset improves the performance on unseen experimental patterns.
 They used \SI{50}{\percent} of the experimental patterns contained in the RRUFF database and added those to their large simulated training set.
@@
 
-It becomes apparent that the more general the task is, the more challenging the transfer to experimental data.
+It becomes apparent that the more general the task is, the more challenging the transfer to experimental data becomes.
 For example, the space group classification task across all material systems is very general.

### 2025-03-01 chapters/03_existing_datasets.tex
Updated checkmark centering in overview tables
Commited Feburary changes

@@
-To contextualize opXRD within the current environment of experimental powder diffraction data, the below list provides an overview of the largest crystal structure databases that offer access to experimental powder diffraction data.
+To contextualize opXRD within the current environment of experimental powder diffraction data, the list below provides an overview of the largest crystal structure databases that offer access to experimental powder diffraction data.
 For an overview of these databases refer to Table~$\eqref{tab:exp_databases}$ below.
@@
 \caption{\footnotesize Overview of experimental powder diffraction databases: The column ``O.A.'' indicates whether or not the database is open access.
-The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'' respectively.}
+The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'', respectively.}
 \label{tab:exp_databases}
-\scalebox{0.84}{
+\scalebox{0.82}{
 \begin{adjustbox}{center}
@@
 Since the PDF5+ lists the MPDS as a database source, there is likely a significant overlap in the experimental patterns found in the PDF5+ and those found in the Pauling file.
-As of November 2024, the PDF5+ is available to researchers through a purchase of a one-year license starting at a price point of \$6265.00.
-However, the ICDD does not allow researchers to train machine learning models on PDF5+ data, regardless of whether the resulting models are published.\footnote{This information can be found in the product license agreement for the PDF5+ at \url{https://www.icdd.com/licensing-process/\#1528471154226-933e5cc6-8da7}.} \\
+As of November 2024, the PDF5+ is available to researchers through a purchase of a one-year license starting at a price point of \$6265.
+However, the ICDD does not allow researchers to train machine learning models on PDF5+ data, regardless of whether the resulting models are published (see product license agreement for the PDF5+ at \url{https://www.icdd.com/licensing-process/\#1528471154226-933e5cc6-8da7}).\\
 
@@
 
-\textbf{PowBase}:\footnote{More information on PowBase can be found at \url{http://www.cristal.org/powbase/index.html}.
-PowBase was an initiative suggested in the Structure Determination by Powder Diffractometry (SDPD) mailing list which was hosted on the same site.
-The COD was another community initiative that grew out of this mailing list.} PowBase is a database of 169 mostly unlabeled experimental powder diffraction patterns collected and maintained by crystallography researcher Armel Le Bail starting in 1999.
+\textbf{PowBase} (\url{http://www.cristal.org/powbase/index.html}): PowBase was an initiative suggested in the Structure Determination by Powder Diffractometry (SDPD) mailing list which was hosted on the same site.
+The COD was another community initiative that grew out of this mailing list.
+PowBase is a database of 169 mostly unlabeled experimental powder diffraction patterns collected and maintained by crystallography researcher Armel Le Bail starting in 1999.
 As of November 2024, all 169 patterns are still freely available for download.
@@
 We are currently planning a systematic large-scale extraction of powder diffraction data from databases like Zenodo with the help of a large language model.
-We will include this data in a future release of our dataset.\\
+This data will be included in a future release of our database.
+\\
 

### 2025-03-01 chapters/04_our_dataset.tex
Commited Feburary changes

@@
-In collaboration with several other institutions we have collected a dataset of diffractograms stemming from experiments, some of them labeled with corresponding structural information.
-The following paragraphs detail the contribution of each research group.
-Most data was collected using \ce{Cu} radiaton sources which have a  $K_{\alpha1}$ wavelength of $\lambda=1.54056\text{\AA}$ and a $K_{\alpha2}$ wavelength of $\lambda=1.54439\text{\AA}$.
-We filtered the submitted datasets to exclude patterns with invalid features such as the pattern containing only one unique recorded angle or negative angles, the pattern containing less than 50 recorded angles total, or all intensities being zero.
-The below Table~$\eqref{tab:merged}$ provides an overview of the opXRD database.
+In collaboration with several other research institutions, we have collected a database of $\numpatterns$ experimental patterns of which 2179 are at least partially labeled with structural information of the underlying material.
+The following research institutions contributed data to the opXRD database: Paris Sciences et Lettres University University (PSL), Hong Kong University of Science and Technology (Guangzhou) (HKUST), University of Southern California (USC), Lawrence Berkeley National Laboratory (LBNL), Empa–Swiss Federal Laboratories for Materials Science and Technology (EMPA) and the Karlsruhe Institute of Technology (KIT).
+\newline
+
+We filtered the submitted datasets to exclude patterns with invalid features such only one unique recorded angle, negative angles, less than 50 recorded angles total, or all intensities being zero.
+Table~$\eqref{tab:merged}$ provides an overview of the opXRD database.
 
@@
 \centering
-\caption{\footnotesize The opXRD database: The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'' respectively.}
+\caption{\footnotesize Overview of the contributions to the opXRD database: The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'' respectively.}
 \label{tab:merged}
@@
 CNRS        &  1052          & \ \cmark   & \partialcheck{85}   & \cmark   &  \ \ \quad \partialcheck{85}   & Diffraction data extracted from the COD \\ 
-HKUST(GZ)   &  21+499          & \ \cmark   & \cmark   & \cmark   & \ \ \quad \cmark   & Phase/structure identification dataset  \\ 
+HKUST(GZ)   &  520          & \ \cmark   & \cmark   & \cmark   & \ \ \quad \cmark   & Phase identification dataset  \\ 
 USC         &  338          & \ \cmark   & \cmark   & \partialcheck{90}   & \ \ \quad \xmark   & Study of CuNi and CuAl alloys          \\ 
-EMPA        &  790          &\ \cmark   & \partialcheck{63}   & \xmark   & \ \ \quad \xmark   & Tin halide perovskites, Zn-V-N libraries \\ 
-INT         &  19,780        &\ \xmark   & \xmark   & \xmark   & \ \ \quad \xmark   & Compilation of various projects        \\ 
-IKFT        &  64           &\ \xmark   & \xmark   & \xmark   & \ \ \quad \xmark   & Commercial catalysts, bulk materials   \\ 
-LBNL        &  n/a          &\ \xmark   & \xmark   & \xmark   & \ \ \quad \xmark   & Perovskites precursors, Mn-Sb-O system  \\ 
+EMPA        &  770          &\ \cmark   & \partialcheck{63}   & \xmark   & \ \ \quad \xmark   & Metal halide perovskites, Zn-V-N libraries \\ 
+INT         &  19,796        &\ \xmark   & \xmark   & \xmark   & \ \ \quad \xmark   & Compilation of various projects        \\ 
+IKFT        &  64           &\ \xmark   & \xmark   & \xmark   & \ \ \quad \xmark   & Commercial catalysts, metals, metal oxides   \\ 
+LBNL        &  70,012          &\ \xmark   & \xmark   & \xmark   & \ \ \quad \xmark   & Perovskites precursors, Mn-Sb-O system  \\ 
 \bottomrule
@@
 
-The variance of the data was analyzed using a dimensional reduction technique called principal component analysis (PCA).
+The variance of the data was analyzed using a principal component analysis (PCA).
 PCA can be applied to datasets $X \subset \mathbb{R}^N$ to reduce the number of components needed to describe points $p \in X$ up to some tolerance in lost accuracy.
 In the context of PCA, the cumulative explained variance ratio is a measure of how much of the variance in the dataset $X$ can be explained using a given number of components.
-For example, if performing PCA on a dataset $X \subset \mathbb{R}^{10}$ yields a cumulative variance ratio of 99\%, then each point in the dataset can be represented using only 5 degrees of freedom up to a minor loss in accuracy.
 For a rigorous definition of PCA and the explained variance ratio, we refer to the literature \cite{Jolliffe2016}.
-\\
+\newline
 
 Here, PCA was performed on datasets of X-ray diffraction patterns.
-These datasets $X$ are subsets of $\mathbb{R}^{N}$ with $N=N_{\text{values}} =512$ since each pattern $p \in X$ was standardized to have 512 intensity values spread out evenly from $0\degree$ to $180\degree$ using zero padding and interpolation with cubic splines.
-Hence the maximal components that could be needed to describe a dataset of diffraction data in this context is $N_{\text{values}}=512$.
+These datasets $X$ are subsets of $\mathbb{R}^{N}$ with $N=512$ since each pattern $p \in X$ was standardized to have 512 intensity values spread out evenly from $0\degree$ to $180\degree$ using zero padding and interpolation with cubic splines.
+Hence the maximal components that could be needed to describe a dataset of diffraction data in this context is $N=512$.
 However, the maximal number of components is even lower for datasets that contain less than 512 patterns.
@@
     N_{\text{max}} = \min(N_{\text{values}}, N_{\text{patterns}}).
+      \label{eq:nmax}
 \end{align}
@@
 For example, the CRNS and the HKUST contributions each are collections that encompass many research projects over a large period of time and hence exhibit a high degree of variability between individual patterns.
-In contrast, the contributions by the USC and the LBNL contain many very similar patterns.
+In contrast, the contributions by USC and LBNL contain many very similar patterns.
 The patterns in the USC datasets are similar because the underlying samples are all variations of \ce{CuNi} and \ce{CuAl} alloys collected on the same machine.
-The patterns submitted by the LBNL are similar because they stem from in-situ recordings where several hundred or several thousand patterns were collected over time per sample while they were undergoing physical processes.
+The patterns submitted by LBNL are similar because they stem from in-situ recordings where several hundred or several thousand patterns were collected over time per sample while they were undergoing physical conversion processes.
 \pagebreak
 
-These effects can be observed in figure~$\eqref{fig:components}$: The cumulative explained variance ratios of the datasets contributed by the USC and LBNL converge towards one significantly faster than the datasets contributed by the CNRS and HKUST.
+These effects can be observed in Figure~$\eqref{fig:components}$: The cumulative explained variance ratios of the datasets contributed by USC and LBNL converge towards one significantly faster than the datasets contributed by CNRS and HKUST.
 \\
@@
     \includegraphics[width=0.85\linewidth]{figures/component_fractions.png}
-    \caption{Explained variance ratio over the fraction of the maximum number of components for each dataset contributed to opXRD.
+    \caption{Explained variance ratio over the fraction of the maximum number of components for each dataset contributed to the opXRD database.
+Here the maximal No.
+components refers to $N_{\text{max}}$ as defined in equation $\eqref{eq:nmax}$.
 Each dataset is named after the contributing institutions.
@@
 
-Figure~$\eqref{fig:histograms}$ below provides an overview of the distributions of pattern and structure properties in the opXRD dataset.
+Figure~$\eqref{fig:histograms}$ provides an overview of the distributions of pattern and structure properties in the opXRD dataset.
 Nearly all patterns have an angular resolution smaller than $\Delta(2\theta) = 0.1 ^\circ$.
@@
 For most patterns, the lowest recorded angle is smaller than $30 ^\circ$ and the highest recorded angle is smaller than $120 ^\circ$.
+The start-to-end angle distribution reveals that all diffractograms start in a narrow window between $0\degree$ and approximately $50\degree$, while they end between $50\degree$ and $150\degree$, with the majority of patterns going from $0\degree$ to approximately $70\degree$.
+This means that, unlike in most ML approaches using synthetic data over the full angle range with fixed resolution, the opXRD dataset has a strongly varying angle range and resolution, which requires additional pre-processing methods such as padding and interpolation, or more flexible ML models beyond standard CNNs.
 
@@
 
-The following texts describe the datasets contributed by each collaborating institution. Each text includes a description of the investigated materials and how X-ray diffraction data was collected.
+In the following, we will describe the datasets contributed by each of the collaborating research groups and institutions.
+Each paragraph includes a description of the investigated materials and how X-ray diffraction data was collected.
+Most data was collected using \ce{Cu} radiaton sources which has a $K_{\alpha1}$ wavelength of $\lambda=1.54056\text{\AA}$ and a $K_{\alpha2}$ wavelength of $\lambda=1.54439\text{\AA}$.
 
@@
 
-A small-scale experimental powder X-ray database called the X-Ray Phase Identification Public Experimental Dataset (XRed) was established over the past two years.\footnote{XRed can be accessed on GitHub at \url{https://github.com/WPEM/XRED}.} The primary objective of this database is to support the development of intelligent phase identification technology, serving as a foundation for data collection for future large-scale machine learning calculations.
-The XRed dataset primarily focuses on metal and metal-oxide particles, with data generated using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers, all utilizing \ce{Cu} X-ray sources.
-Hundreds of pXRD patterns have been obtained through XRD refinement, enabling the determination of detailed lattice constants, atomic sizes, and other structural parameters.
-These patterns are labeled with CIF files that document the refined structures.
-The dataset is organized by elemental systems and includes original experimental files ranging from single-phase to five-phase mixtures, as well as mixtures studied for different tasks.
+A small-scale experimental powder X-ray database, the X-Ray Phase Identification Public Experimental Dataset (XRed), has been developed over the past two years (\url{https://github.com/WPEM/XRED}).
+The primary goal of XRed is to support the advancement of intelligent phase identification technology by providing a foundation for data collection in future large-scale machine learning applications.
+XRed primarily focuses on metal and metal-oxide particles, with data collected using diffractometers such as the Empyrean 3.0, Aeris, and Bruker D8 Advance, all employing \ce{Cu} X-ray sources.
+The dataset contains hundreds of pXRD patterns, obtained through XRD refinement, which enable the determination of lattice constants, atomic sizes, and other structural parameters.
+Each pattern is labeled with a corresponding CIF that documents the refined structure.
+Data are categorized by elemental systems and include original experimental files, spanning single-phase to five-phase mixtures, as well as mixtures designed for various research tasks.
 \\
 
-Another experimental dataset integrated with opXRD is a collection of powder diffraction data sourced from open-access publications and collaborating institutions.
-All collaborating institutions provided this data with full authorization for research purposes.
-This dataset provides broader coverage of chemical elements and includes ionic crystals, atomic crystals, and metallic crystals.
-This dataset surpasses Xred in size, containing 499 entries.
-However, contrary to Xred these data do not come paired with corresponding crystal structures.
+In addition to XRed, the opXRD database integrates an experimental dataset composed of powder diffraction data sourced from open-access publications and collaborating institutions.
+These institutions have provided the data with full authorization for research purposes.
+Compared to XRed, this dataset offers broader chemical element coverage, encompassing ionic, atomic, and metallic crystals.
+It is also larger, containing 499 entries.
+However, unlike XRed, these data entries are not accompanied by CIF files.
 \\
 
-Xqueryer,\footnote{Xqueryer can be accessed at \url{https://xqueryer.caobin.asia/}.} a recently launched cutting-edge free online platform for intelligent structure identification, will also serve as a source of data for the opXRD database.
-Users can upload experimental pXRD data to identify crystal structures and extract all relevant information available in the Materials Project database\cite{jain2013commentary}.
-Uploaded data are securely stored and integrated into the comprehensive opXRD database to progress toward a large-scale, long-term experimental resource.
-All usage complies with Xqueryer's terms of service.
+Another data source for opXRD is Xqueryer (\url{https://xqueryer.caobin.asia/}), a recently launched free online platform for intelligent structure identification.
+Users can upload experimental pXRD data to identify crystal structures and extract relevant information from the Materials Project database\cite{jain2013commentary}.
+Uploaded data are securely stored and integrated into the opXRD database, contributing to its development as a large-scale, long-term experimental resource.
+All data usage adheres to Xqueryer's terms of service.
 
@@
 
-X-ray diffraction data was collected from a wide range of research projects conducted at the Institute of Nanotechnology over the past years.
+X-ray diffraction data was collected from a wide range of research projects conducted at the Institute of Nanotechnology over the past 10 years.
 A major part of the research focused on high-entropy materials, which involved incorporating many different elements into single-phase structures, leading to peak shifts or phase separations.
@@
 
-In total three different datasets were collected.
-The first and second datasets focused on metal-halide perovskite precursor solutions.
-The first dataset was collected from spin-coating triple-cation metal-halide perovskite precursor solutions with the composition \ce{Cs}$_{0.05}$(MA$_{0.23}$FA$_{0.77}$)\ce{Pb}$_{1.1}$(\ce{I}$_{0.77}$\ce{Br}$_{0.23}$)$_{3}$ onto various substrates.
+In total four different datasets were collected. The first and second datasets focused on metal-halide perovskite thin-films, (...). \\
+
+The first dataset (LBNL-A) was collected from spin-coating and annealing triple-cation metal-halide perovskite precursor solutions with the composition \ce{Cs}$_{0.05}$(MA$_{0.23}$FA$_{0.77}$)\ce{Pb}$_{1.1}$(\ce{I}$_{0.77}$\ce{Br}$_{0.23}$)$_{3}$ onto various substrates.
 Here, MA stands for Methylammonium and FA stands for Formamidinium.
@@
 Christian Kaufmann and his team at Helmholtz-Zentrum Berlin (HZB) in Germany.
-Data collection was performed in situ during thin-film deposition using a custom-made spin-coating and annealing stage.
-\\
+Data collection was performed in situ during thin-film deposition using a custom-made spin-coating and annealing stage \cite{song2019}.
 
-A second dataset was collected from spin-coating metal-halide perovskite precursor solutions with varying compositions of \ce{MAPb(I_{1-x}Br_x)3} spin-coated onto glass substrates.
+A second dataset (LBNL-B) was collected from spin-coating metal-halide perovskite precursor solutions with varying compositions of \ce{MAPb(I_{1-x}Br_x)3} spin-coated onto glass substrates.
 Here, MA = Methylammonium and x = 0, 0.33, 0.5, 0.67, 1.
@@
 
-A third dataset was collected by observing the phase evolution of an Mn-Sb-O system with varying annealing temperatures.
+A third dataset (LBNL-C) was collected by observing the phase evolution of an Mn-Sb-O system with varying annealing temperatures.
 The temperatures used to analyze the crystal structure of the Mn-Sb-O system were chosen depending on the number of phase transitions appearing for a certain temperature range.
@@
 Therefore, the crystal structure was measured every 100$\degree$C between room temperature and 300$\degree$C; every 50$\degree$C between 300$\degree$C and 850$\degree$C; and every 200$\degree$C when cooling down.
-The heating and cooling rates are fixed for all the experiments at 50$\degree$C/min and the holding time is fixed at 2 min.
+The heating and cooling rates were fixed for all the experiments at 50$\degree$C/min and the holding time was fixed to 2 min.
 \\
 
-The data was collected using the \textit{in situ}  Rigaku-SmartLab3kW diffractometer.
+This data was collected using the \textit{in situ}  Rigaku-SmartLab3kW diffractometer.
 This tool operates with SmartLab Studio II software, which can measure the X-ray diffraction during the annealing process.
@@
 % end of Gregoire Heymans
+
+The fourth and final dataset (LBNL-D) was collected from (...).
+To prepare these compounds thiol-functionalized UiO-66-type Zr-based MOFs were added as additives to the PbI2 precursor solution followed by a two-step synthesis method to form FAMAPb(I,Br)3 films.
+XRD data was collected (...).

### 2025-03-01 chapters/05_how_to_use.tex
Added alphabetic labeling rule in Usage section
Commited Feburary changes

--- 2024-12-09 chapters/05_how_to_use.tex
@@
-The opXRD database is hosted on Zenodo\footnote{The dataset can be found at \url{https://zenodo.org/records/14278656}.} and can be downloaded by any user without any barriers or restrictions.
-The database comes in two files, ``opxrd.zip'' and ``opxrd\_in\_situ.zip''.
-The latter contains the in-situ data which contains highly correlated patterns recorded through time series measurements.
-Patterns belonging to time series measurements have filenames that indicate the measurement series they belong to as well as their order in that series.
+The opXRD database is hosted on Zenodo (\url{https://zenodo.org/records/14278656}) and can be downloaded by any user without any barriers or restrictions.
 \\
@@
 Next to the availability of the opXRD dataset on Zenodo, we also provide a Python library ``opxrd'' to easily download and interface with the dataset.
-The instructions for how to install this library can be found in the repository associated with the library.\footnote{The repository to this library can is located at \url{https://github.com/aimat-lab/opxrd}.} The opxrd library allows the dataset to be accessed through one simple command: \pyth{OpXRD.load(root_dirpath)}.
-If the database is locally available under \pyth{root_dirpath} this command loads the library from this location.
-If the database is not available locally at this location, the database is automatically downloaded to \pyth{root_dirpath}.
+The instructions for how to install this library can be found in the repository associated with the library.
+The repository to this library is located at \url{https://github.com/aimat-lab/opxrd}.
 \\
 
-Our library further includes options for standardization, plotting, and the conversion to \emph{PyTorch} tensors.
-We provide a Jupyter Notebook\footnote{This notebook can be found here: \url{https://colab.research.google.com/github/aimat-lab/opXRD/blob/main/opxrd/usage.ipynb}.} that showcases these functionalities in more detail.
-This notebook also illustrates how to interface with the OpXRD database through Python.
+The opxrd library includes options for data-loading, for standardization, plotting, and the conversion to \emph{PyTorch} tensors.
+We provide a Jupyter Notebook %\footnote{This notebook can be found here: \url{https://colab.research.google.com/github/aimat-lab/opXRD/blob/main/opxrd/usage.ipynb}.}
+that showcases these functionalities in more detail. This notebook also illustrates how to interface with the opXRD database through Python.

### 2025-03-01 chapters/06_summary_and_outlook.tex
Updated introductory paragraphs on rietveld refinement
Merged Xqueyer paragraph into contribution entry
Commited table version with footnote directly underneath
Added histograms, PCA plots + paragraph on notebook; Added labeled column in existing datasets;
Updated research projects in overview table
Commited Feburary changes

@@
-With the opXRD dataset, a curation of XXX unlabeled and YYY labeled experimental powder XRD patterns from a wide range of different materials systems, we provide the largest currently available source of experimental XRD patterns.
+With the opXRD dataset, a curation of $\numpatterns$ unlabeled and 2179 at least partially labeled experimental powder XRD patterns from a wide range of different materials systems, we provide the largest currently available source of experimental XRD patterns.
 With this, we address the need for experimental data that arises when developing algorithms and analysis tools for XRD data, both based on machine learning and classical approaches.
@@
 
-\paragraph{How to contribute}
-
+% How to contribute
 Rather than a finished project, the opXRD dataset is an ongoing effort to collect experimental powder XRD data.
-We invite everyone in possession of experimental powder XRD data to submit it to our dataset, to further improve the dataset and aid further developments in this field.
+We invite everyone who is working in the area of experimental powder XRD to submit it to the dataset, in order to further improve the utility of the dataset and thus aid further developments in this field.
 Our submission page (\url{https://xrd.aimat.science/}) and submission helper software will be kept available to collect more data.
-We will keep updating and maintaining the dataset with incoming new submissions.
-
-%We have launched Xqueryer (\url{http://xqueryer.caobin.asia/}), a cutting-edge, free online platform for intelligent structure identification.
-Users can upload experimental pXRD data to identify crystal structures and extract all relevant information available in the Materials Project database.
-All usage complies with Xqueryer's terms of service.
-Uploaded data are securely stored and integrated into the opXRD database.
+We will keep updating and maintaining the dataset with new incoming submissions.

### 2025-03-01 main.tex
Centered checkmarks in overview table
Commited Feburary changes

@@
 \usepackage{eurosym}
-\usepackage[locale = DE]{siunitx}
+\usepackage[locale = US]{siunitx}
 \usepackage{bm}
@@
 \usepackage{pifont}% http://ctan.org/pkg/pifont
-\newcommand{\cmark}{\textcolor{green!80!black}{\ding{52}}}
-\newcommand{\xmark}{\textcolor{red}{\ding{53}}}
-\newcommand{\partialcheck}[1]{\textcolor{orange}{\ding{52}(#1\%)}}
-\captionsetup[table]{
-  font=footnotesize
-}
+\newcommand{\cmark}[1][]{\textcolor{green!80!black}{#1\quad\ding{52}}}
+\newcommand{\xmark}[1][]{\textcolor{red}{#1\quad\ding{53}}}
+\newcommand{\partialcheck}[1]{\textcolor{orange}{\quad\ding{52}(#1\%)}}
+\captionsetup[table]{font=footnotesize}
 
@@
 \renewcommand\Affilfont{\normalfont\small}
+\newcommand{\numpatterns}{92,552 }
 
@@
 \author[\KFUPMMSE,\KFUPMIRCIMR]{Maged Abdelsamie}
-\author[\HKUSTGZ]{Longhan Zhang}
+%\author[\HKUSTGZ]{Longhan Zhang}
 \author[\HKUSTGZ]{Tong-yi Zhang}
@@
 \affil[\INT]{Institute of Nanotechnology, Karlsruhe Institute of Technology (KIT), 76131 Karlsruhe, Germany}
-\affil[\HKUSTGZ]{Guangzhou Municipal Key Laboratory of Materials Informatics, Hong Kong University of Science and Technology (Guangzhou) (HKUST), Guangzhou 511400, China}
+\affil[\HKUSTGZ]{Guangzhou Municipal Key Laboratory of Materials Informatics, Advanced Materials Thrust, Hong Kong University of Science and Technology (Guangzhou) (HKUST), Guangzhou 511400, China}
 \affil[\USC]{Department of Chemical Engineering and Materials Science, University of Southern California (USC), Los Angeles CA 90089, USA}
@@
 
-\pagebreak
-
 \subsection*{Supporting Information}
-%Supporting Information is available from the Wiley Online Library or the author.
-
-The below figure $\eqref{fig:combined}$ displays a randomly selected sample of the X-ray diffraction patterns found in each of the research projects contributed to the opXRD database.
-
-\begin{figure}[!htb]
-    \centering
-    \includegraphics[width=\linewidth]{figures/combined.png}
-    \caption{50 randomly chosen X-ray diffraction patterns from each contributed dataset.
-The figure shows data from the following datasets: a) EMPA, b) LBNL-A, c) LBNL-B, d) LBNL-C, e) USC, f) INT, g) HKUST-A, h) HKUST-B, i) CNRS, j) IKFT.}
-    \label{fig:combined}
-\end{figure}
-
-\pagebreak
+Supporting Information is available from the Wiley Online Library or the author.
 
@@
 The opXRD database is available on Zenodo at \url{https://zenodo.org/records/14254270}.
-It is published under the Creative Commons Attribution 4.0 International license.\footnote{Detailed terms of this license can be found at \url{https://creativecommons.org/licenses/by/4.0/deed.en}.} It can be downloaded by any user without any barriers or restrictions.
-For further details, please refer to Section~$\eqref{sec:how_to_use}$.
+It is published under the Creative Commons Attribution 4.0 International license.%\footnote{Detailed terms of this license can be found at \url{https://creativecommons.org/licenses/by/4.0/deed.en}.}
+It can be downloaded by any user without any barriers or restrictions. For further details, please refer to Section~$\eqref{sec:how_to_use}$.
 
@@
 
+\setcounter{section}{0}
+\renewcommand{\thesection}{S\arabic{section}}
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{S\arabic{figure}}
+\setcounter{table}{0}
+\renewcommand{\thetable}{S\arabic{table}}
+
+\section*{Supporting Information}
+%Supporting Information is available from the Wiley Online Library or the author.
+
+\section{Description of opXRD files on Zenodo}
+The database comes in two zip archives, ``opxrd.zip'' and ``opxrd\_in\_situ.zip''.
+The latter contains the in-situ data with highly correlated patterns recorded through time series measurements.
+Within the .zip archives patterns are saved as .json files grouped in folders indicating the contributing institution.
+If an institution contributed data from several projects, the contributed data is further divided into folders indicating the research project.
+These research project folders are labeled alphabetically in the order they are introduced in section $\eqref{sec:our_dataset}$.
+Each .json file contains a pattern recorded from an X-ray diffraction experiment.
+If available, the composition and structure of the investigated sample and experiment conditions are also included in this file.
+Patterns belonging to time series measurements are labeled with filenames that indicate the measurement series they belong to as well as their order in that series.
+
+\section{opXRD Python library usage}
+
+The opXRD Python library allows the dataset to be accessed through one simple command: \pyth{OpXRD.load(root_dirpath)}.
+If the database is locally available under \pyth{root_dirpath} this command loads the library from this location.
+If the database is not available locally at this location, the database is automatically downloaded to \pyth{root_dirpath}.
+
+\section{Fourier analysis}
+The Figure below $\eqref{fig:combined}$ shows 50 randomly selected samples of the X-ray diffraction patterns found in each of the research projects contributed to the opXRD database.
+
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=\linewidth]{figures/combined.png}
+    \caption{50 randomly chosen X-ray diffraction patterns from each contributed dataset.
+The figure shows data from the following datasets: a) EMPA, b) LBNL-A, c) LBNL-B, d) LBNL-C, e) USC, f) INT, g) HKUST-A, h) HKUST-B, i) CNRS, j) IKFT.}
+    \label{fig:combined}
+\end{figure}
+
+\pagebreak
+
 \end{document}

### 2025-03-04 chapters/00_abstract.tex
Took on board co-author feedback/inputs

@@
 Powder X-ray diffraction (pXRD) experiments are a cornerstone for materials structure characterization.
-Despite their widespread application, the analysis of pXRD diffractograms still presents a significant challenge to automation and thus a bottleneck in high-throughput materials discovery in self-driving labs.
-Machine learning has emerged as a promising research direction to resolve this bottleneck by enabling automated powder diffraction data analysis.
-A notable difficulty in applying machine learning to this domain is the lack of sufficiently sized experimental datasets, which has constrained machine learning researchers to train primarily on simulated data.
-However, models trained on simulated pXRD patterns showed limited generalization ability when applied to experimental patterns, particularly for low-quality experimental patterns with high noise levels and elevated backgrounds.
-With the Open Experimental Powder X-Ray Diffraction Database (opXRD), we provide an openly available and easily accessible dataset of both labeled and unlabeled experimental powder diffractograms.
-Labeled opXRD data can be used to evaluate the performance of models on real experimental data and unlabeled opXRD data can help improve the performance of models on experimental data, e.g.
+Despite their widespread application, analyzing pXRD diffractograms still presents a significant challenge to automation and a bottleneck in high-throughput discovery in self-driving labs.
+Machine learning promises to resolve this bottleneck by enabling automated powder diffraction analysis.
+A notable difficulty in applying machine learning to this domain is the lack of sufficiently sized experimental datasets, which has constrained researchers to train primarily on simulated data.
+However, models trained on simulated pXRD patterns showed limited generalization to experimental patterns, particularly for low-quality experimental patterns with high noise levels and elevated backgrounds.
+With the Open Experimental Powder X-Ray Diffraction Database (opXRD), we provide an openly available and easily accessible dataset of labeled and unlabeled experimental powder diffractograms.
+Labeled opXRD data can be used to evaluate the performance of models on experimental data and unlabeled opXRD data can help improve the performance of models on experimental data, e.g.
 through transfer learning methods.
-We collected \numpatterns diffractograms - 2179 of them labeled - from a wide spectrum of materials classes, establishing the first version of the opXRD database.
-We hope that this ongoing effort can guide machine learning research toward fully automated analysis of pXRD data and thus enable future self-driving materials labs.
+We collected \numpatterns diffractograms - 2179 of them labeled - from a wide spectrum of materials classes.
+We hope this ongoing effort can guide machine learning research toward fully automated analysis of pXRD data and thus enable future self-driving materials labs.

### 2025-03-04 chapters/01_introduction.tex
Took on board co-author feedback/inputs

@@
 After fabricating a new material, a number of analysis techniques can be used to characterize the sample.
-One method that can be used for phase identification, phase quantification, grain size characterization, and to determine the crystal structure of a new material is powder X-ray diffraction.
-Typically crystal structures are determined from pXRD data through Rietveld refinement.
+One method that can be used for phase identification, phase quantification, grain size characterization, and to determine the crystal structure of a new material is powder X-ray diffraction (pXRD).
+When using pXRD measurements, crystal structures are typically determined through Rietveld refinement.
 In Rietveld refinement, an initial crystal structure model is fitted to the observed diffractogram by iteratively updating the structural model.
@@
 Additionally, it should be considered that varying experimental setups can produce distinct powder diffraction patterns on the same sample.
-Features that may vary between experimental setups and influence the recorded pattern include the shape of diffraction peaks, the wavelength and polarization of the employed X-ray source, and the detector geometry
+Features that may vary between experimental setups and influence the recorded pattern include the shape of diffraction peaks, the wavelength and polarization of the employed X-ray source as well as potentially non-monochromatic radiation, and the detector geometry
 \cite{cao2024simxrd, Waseda2011, Pecharsky2023}.
@@
 It contains pXRD patterns from single and multiphase materials from a wide variety of material classes, including high-entropy materials, perovskites, and commercial catalysts.
-Details about the experiments of contributing research groups are discussed in Section~$\eqref{sec:our_dataset}$.
-A detailed description of how to use the data is given in Section~$\eqref{sec:how_to_use}$ and Section~$\eqref{sec:summary_and_outlook}$ describes how further data can be contributed.
+Some of the XRD data was collected on thin-films rather than on true powder samples, which may influence the quality of the data in regards to full structure resolution.
+Additionally some of the data was collected in grazing-angle geometry rather than in the usual Bragg-Brentano geometry employed in powder diffraction.
+Details about the experiments of contributing research groups, including the presence of thin-films samples and the usage of atypical diffraction geometries are discussed in Section~$\eqref{sec:our_dataset}$.
+A detailed description of how to use the data is given in Section~$\eqref{sec:how_to_use}$, and Section~$\eqref{sec:summary_and_outlook}$ describes how further data can be contributed.
 

### 2025-03-04 chapters/02_SOTA_ML.tex
Took on board co-author feedback/inputs

@@
 
-Using simulated diffractograms based on structures from the ICSD, Lee {\it et al.} trained a deep convolutional neural network (CNN) that is able to classify occurring phases in diffractograms of a specific compound pool \cite{Lee2020}.
-They furthermore developed models based on fully convolutional neural networks and transformer encoders that predict the crystal system, the spacegroup, and other structural properties, such as the band gap \cite{Lee2022}.
+In 2020, Lee {\it et al.} trained a deep convolutional neural network (CNN) using simulated diffractograms based on structures from the ICSD, which is able to classify occurring phases in diffractograms of a specific compound pool \cite{Lee2020}.
+In 2022, they furthermore developed models based on fully convolutional neural networks and transformer encoders that predict the crystal system, the spacegroup, and other structural properties, such as the band gap \cite{Lee2022}.
 With their best model for the crystal system prediction on ICSD structures, they achieved a test accuracy of \SI{92.2}{\percent}.
-Park {\it et al.} reached a test accuracy of roughly \SI{81}{\percent} for a CNN, which classifies space groups of simulated single-phase diffractograms \cite{Park2017}.
+In 2017, Park {\it et al.} reached a test accuracy of roughly \SI{81}{\percent} for a CNN, which classifies space groups of simulated single-phase diffractograms \cite{Park2017}.
 \\
 
-A regression analysis on lattice parameters within a broader framework encompassing all material classes was conducted by Chitturi {\it et al} \cite{Chitturi2021}.
+A regression analysis on lattice parameters within a broader framework encompassing all material classes was conducted by Chitturi {\it et al} \cite{Chitturi2021} in 2021.
 They developed a distinct CNN for each crystal system, utilizing a merged dataset from both the ICSD and the Cambridge Structural Database, and managed to achieve a mean absolute percentage error of about \SI{10}{\percent} for the lattice lengths, although they encountered difficulties in accurately predicting angles.
-Zhang {\it et al.} introduce a convolutional self-attention neural network trained on simulated patterns to classify crystal types \cite{zhang2024crystallographic}.
+In 2024, Zhang {\it et al.} introduced a convolutional self-attention neural network trained on simulated patterns to classify crystal types \cite{zhang2024crystallographic}.
 Their model was tested on 23,073 unary, binary, and ternary inorganic crystal structures sourced from the COD.
-The study observed a noticeable performance drop when the pre-trained model was applied to real experimental patterns, as opposed to simulated data.
+The study observed a noticeable performance drop when the pre-trained model was applied to real experimental patterns as opposed to simulated data.
 However, their recent work \cite{cao2024simxrd} proposes using convolutional peak descriptors that consider the detector's geometry, which reduces the performance gap in their benchmark tests.\\
 
-The option of training purely on experimental diffractograms unfortunately is infeasible because of the limited availability of labeled experimental diffractograms, but Salgado {\it et al.} \cite{Salgado2023} showed that adding a fraction of experimental patterns to a simulated training dataset improves the performance on unseen experimental patterns.
+The option of training purely on experimental diffractograms is infeasible for neural networks because of the limited availability of labeled experimental diffractograms.
+However, in 2023, Salgado {\it et al.} \cite{Salgado2023} showed that adding a fraction of experimental patterns to a simulated training dataset improves the performance on unseen experimental patterns.
 They used \SI{50}{\percent} of the experimental patterns contained in the RRUFF database and added those to their large simulated training set.
@@
 
-Schuetzke {\it et al.} trained a classifier to classify if a diffractogram stems from an amorphous, single-phase, or multi-phase sample \cite{Schuetzke2024}.
+In 2024, Schuetzke {\it et al.} trained a classifier to classify if a diffractogram stems from an amorphous, single-phase, or multi-phase sample \cite{Schuetzke2024}.
 Due to the lack of experimental pXRDs, they built a pipeline to augment simulated diffractograms of a reference structure by, among other things, slightly varying the underlying crystal lattice.
@@
 
-In our 2023 paper, we developed an approach to generate synthetic crystal structures and their corresponding pXRD patterns on the fly during the training process \cite{Schopmans2023}.
-This approach defeats the issue of
-a limited dataset size, which limits the depth of neural networks that can be trained.
+In 2023, Schopmans {\it et al.} presented an approach to generate synthetic crystal structures and their corresponding pXRD patterns on the fly during the training process \cite{Schopmans2023}.
+This approach defeats the issue of a limited dataset size, which limits the depth of neural networks that can be trained.
 However, the accuracy dropped substantially when we applied our space group classification model to experimental patterns from the RRUFF database.

### 2025-03-04 chapters/04_our_dataset.tex
Took on board co-author feedback/inputs

@@
 In collaboration with several other research institutions, we have collected a database of $\numpatterns$ experimental patterns of which 2179 are at least partially labeled with structural information of the underlying material.
-The following research institutions contributed data to the opXRD database: Paris Sciences et Lettres University University (PSL), Hong Kong University of Science and Technology (Guangzhou) (HKUST), University of Southern California (USC), Lawrence Berkeley National Laboratory (LBNL), Empa–Swiss Federal Laboratories for Materials Science and Technology (EMPA) and the Karlsruhe Institute of Technology (KIT).
+The following research institutions contributed data to the opXRD database: French National Centre for Scientific Research (CNRS), Hong Kong University of Science and Technology (Guangzhou) (HKUST), University of Southern California (USC), Lawrence Berkeley National Laboratory (LBNL), Empa–Swiss Federal Laboratories for Materials Science and Technology (EMPA) and the Karlsruhe Institute of Technology (KIT).
 \newline
 
-We filtered the submitted datasets to exclude patterns with invalid features such only one unique recorded angle, negative angles, less than 50 recorded angles total, or all intensities being zero.
+We filtered the submitted datasets to exclude patterns with invalid features such as only one unique recorded angle, negative angles, less than 50 recorded angles total, or all intensities being zero.
 Table~$\eqref{tab:merged}$ provides an overview of the opXRD database.
@@
 Each paragraph includes a description of the investigated materials and how X-ray diffraction data was collected.
+If applicable, the presence of thin-film samples or atypical diffraction geometrices is indicated.
 Most data was collected using \ce{Cu} radiaton sources which has a $K_{\alpha1}$ wavelength of $\lambda=1.54056\text{\AA}$ and a $K_{\alpha2}$ wavelength of $\lambda=1.54439\text{\AA}$.
@@
 %% Paragraph by FX Coudert and Arthur Hardiagon
-\subsubsection*{Chimie ParisTech, PSL University}
+\subsubsection*{Institut de Recherche de Chimie Paris, CNRS}
 
@@
 
-A small-scale experimental powder X-ray database, the X-Ray Phase Identification Public Experimental Dataset (XRed), has been developed over the past two years (\url{https://github.com/WPEM/XRED}).
+Two datasets were contributed to the opXRD database.
+The first dataset (HKUST-A) is a small-scale experimental powder X-ray database developed over the past two years, called the X-Ray Phase Identification Public Experimental Dataset (XRed) (\url{https://github.com/WPEM/XRED}).
 The primary goal of XRed is to support the advancement of intelligent phase identification technology by providing a foundation for data collection in future large-scale machine learning applications.
@@
 
-In addition to XRed, the opXRD database integrates an experimental dataset composed of powder diffraction data sourced from open-access publications and collaborating institutions.
+In addition to XRed, the opXRD database integrates an experimental dataset composed of powder diffraction data sourced from open-access publications and collaborating institutions (HKUST-B).
 These institutions have provided the data with full authorization for research purposes.
@@
 
-Another data source for opXRD is Xqueryer (\url{https://xqueryer.caobin.asia/}), a recently launched free online platform for intelligent structure identification.
+%Another data source for opXRD is Xqueryer (\url{https://xqueryer.caobin.asia/}), a recently launched free online platform for intelligent structure identification.
 Users can upload experimental pXRD data to identify crystal structures and extract relevant information from the Materials Project database\cite{jain2013commentary}.
@@
 
-In total four different datasets were collected. The first and second datasets focused on metal-halide perovskite thin-films, (...). \\
-
+In total four different datasets were collected.
 The first dataset (LBNL-A) was collected from spin-coating and annealing triple-cation metal-halide perovskite precursor solutions with the composition \ce{Cs}$_{0.05}$(MA$_{0.23}$FA$_{0.77}$)\ce{Pb}$_{1.1}$(\ce{I}$_{0.77}$\ce{Br}$_{0.23}$)$_{3}$ onto various substrates.
@@
 
-% end of Gregoire Heymans
-
-The fourth and final dataset (LBNL-D) was collected from (...).
-To prepare these compounds thiol-functionalized UiO-66-type Zr-based MOFs were added as additives to the PbI2 precursor solution followed by a two-step synthesis method to form FAMAPb(I,Br)3 films.
-XRD data was collected (...).
+% Paragraph by Mriganka Singh
+A fourth dataset (LBNL-D) was collected from a two-step spin-coating process using metal-organic frameworks (MOFs) in perovskite precursor solutions, deposited onto glass substrates.
+In the first step, a nanoscale thiol-functionalized UiO-66-type Zr-based MOF (\ce{UiO}-66-\ce{(SH)2)} was added to the \ce{PbI2} precursor.
+This was followed by the deposition of an organic mixture solution containing \ce{FAI}, \ce{MACl}, and \ce{MABr} in the second step.
+The incorporation of MOFs aids in suppressing perovskite vacancy defects, thereby enhancing device stability and efficiency.
+To further investigate the influence of \ce{UiO}-66-\ce{(SH)2)} on perovskite thin-film formation during the annealing process, a time-resolved GIWAXS experiment was conducted.
+The measurements were performed using a setup similar to that of LBNL-A and B.

### 2025-03-04 main.tex
Took on board co-author feedback/inputs

@@
 
-\affil[\ITI]{Institute of Theoretical Informatics, Karlsruhe Institute of Technology (KIT), 76131 Karlsruhe, Germany.  E-mail: pascal.friederich@kit.edu}
-\affil[\INT]{Institute of Nanotechnology, Karlsruhe Institute of Technology (KIT), 76131 Karlsruhe, Germany}
-\affil[\HKUSTGZ]{Guangzhou Municipal Key Laboratory of Materials Informatics, Advanced Materials Thrust, Hong Kong University of Science and Technology (Guangzhou) (HKUST), Guangzhou 511400, China}
-\affil[\USC]{Department of Chemical Engineering and Materials Science, University of Southern California (USC), Los Angeles CA 90089, USA}
-\affil[\LBLMF]{Molecular Foundry Division, Lawrence Berkeley National Laboratory (LBNL), Berkeley 94720 CA, USA}
-\affil[\LBLALS]{Advanced Light Source, Lawrence Berkeley National Laboratory,  Berkeley 94720 CA, USA}
-\affil[\Hoffmann]{Hoffmann Institute of Advanced Materials, Shenzhen Polytechnic, Shenzhen 518055, China}
-\affil[\LBLCS]{Lawrence Berkeley National Laboratory (LBNL), Chemical Sciences Division, Berkeley 94720 CA, USA}
-\affil[\EMPA]{Empa–Swiss Federal Laboratories for Materials Science and Technology (EMPA), 8600 Dübendorf, Switzerland}
-\affil[\KITICRT]{Engler-Bunte-Institut \& Institute of Catalysis Research and Technology, Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany}
-\affil[\CPT]{Chimie ParisTech, Paris Sciences et Lettres University University (PSL), CNRS, Institut de Recherche de Chimie Paris, 75005 Paris, France}
-\affil[\KFUPMMSE]{Material Science and Engineering Department, King Fahd University of Petroleum and Minerals (KFUPM), Dhahran 31261, Saudi Arabia}
-\affil[\KFUPMIRCIMR]{Interdisciplinary Research Center for Intelligent Manufacturing and Robotics, King Fahd University of Petroleum and Minerals (KFUPM), Dhahran 31261, Saudi Arabia}
-\affil[\FZJIEMD]{Institute of Energy Materials and Devices, Forschungszentrum Juelich GmbH, 52425 Juelich, Germany}
+\affil[\ITI]{Institute of Theoretical Informatics, Karlsruhe Institute of Technology (KIT), 76131 Karlsruhe, Germany.
+E-mail: pascal.friederich@kit.edu} % Informatik-Hauptgebäude, Am Fasanengarten 5, 76131 Karlsruhe
+\affil[\INT]{Institute of Nanotechnology, Karlsruhe Institute of Technology (KIT), 76131 Karlsruhe, Germany} % Hermann-von-Helmholtz-Platz 1
+\affil[\HKUSTGZ]{Guangzhou Municipal Key Laboratory of Materials Informatics, Advanced Materials Thrust, Hong Kong University of Science and Technology (Guangzhou) (HKUST), Guangzhou 511400, China} %1 Duxue Road,
+\affil[\USC]{Department of Chemical Engineering and Materials Science, University of Southern California (USC), Los Angeles CA 90089, USA} % 925 Bloom Walk
+\affil[\LBLMF]{Molecular Foundry Division, Lawrence Berkeley National Laboratory (LBNL), Berkeley 94720 CA, USA}  %67 Cyclotron Rd, Berkeley, CA 94720, United States
+\affil[\LBLALS]{Advanced Light Source, Lawrence Berkeley National Laboratory,  Berkeley 94720 CA, USA}  % 6 Cyclotron Rd, Berkeley, CA 94720, United States
+\affil[\Hoffmann]{Hoffmann Institute of Advanced Materials, Shenzhen Polytechnic, Shenzhen 518055, China} % ??
+\affil[\LBLCS]{Lawrence Berkeley National Laboratory (LBNL), Chemical Sciences Division, Berkeley 94720 CA, USA}  % 1 Cyclotron Rd, Berkeley, CA 94720, United States
+\affil[\EMPA]{Empa–Swiss Federal Laboratories for Materials Science and Technology (EMPA), 8600 Dübendorf, Switzerland}  % Ueberlandstrasse 129, 8600 Dübendorf, Switzerland
+\affil[\KITICRT]{Engler-Bunte-Institut \& Institute of Catalysis Research and Technology, Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany}  %Hermann-von-Helmholtz-Platz 1
+\affil[\CPT]{Chimie ParisTech, PSL University, CNRS, Institut de Recherche de Chimie Paris, 75005 Paris, France} % 11 Rue Pierre et Marie Curie, 75005 Paris, France
+\affil[\KFUPMMSE]{Material Science and Engineering Department, King Fahd University of Petroleum and Minerals (KFUPM), Dhahran 31261, Saudi Arabia} % 846R+8C2, Academic Belt Rd, KFUPM, Dhahran 34463, Saudi Arabia
+\affil[\KFUPMIRCIMR]{Interdisciplinary Research Center for Intelligent Manufacturing and Robotics, King Fahd University of Petroleum and Minerals (KFUPM), Dhahran 31261, Saudi Arabia} %Academic Belt Road, Dhahran 31261, Saudi Arabia
+\affil[\FZJIEMD]{Institute of Energy Materials and Devices, Forschungszentrum Juelich GmbH, 52425 Juelich, Germany} %Wilhelm-Johnen-Straße, 52428 Jülich
 
@@
 Work by the USC group was supported by the National Science Foundation (NSF) grant numbers DMR-2227178 and OISE-2106597.
+Funding by the Helmholtz Research Program “Materials and Technologies for the Energy Transition (MTET), Topic 3: Chemical Energy Carriers" is kindly acknowledged.
 
@@
 
-\section{Description of opXRD files on Zenodo}
+\section*{Description of opXRD files on Zenodo}
 The database comes in two zip archives, ``opxrd.zip'' and ``opxrd\_in\_situ.zip''.
@@
 
-\section{opXRD Python library usage}
+\section*{opXRD Python library usage}
 
@@
 
-\section{Fourier analysis}
+\section*{Combined pattern plots}
 The Figure below $\eqref{fig:combined}$ shows 50 randomly selected samples of the X-ray diffraction patterns found in each of the research projects contributed to the opXRD database.

### 2025-03-06 chapters/01_introduction.tex
Restructured paragraph discussing significance of data

@@
 
-However, even small labeled and also unlabeled datasets of experimental diffractograms hold significant value in practice.
-First of all, labeled data can be used to test and benchmark existing and new automated analysis approaches, which yields a better comparison than when using the very limited currently available datasets, such as the commonly used RRUFF mineral dataset\cite{Armbruster2015}.
-\\
+Both labeled and unlabeled datasets of experimental diffractograms hold significant value for machine learning-based pXRD analysis.
+Labeled opXRD data can be used to test and benchmark existing and new automated analysis approaches, which yields a better comparison than when using the very limited currently available datasets, such as the commonly used RRUFF mineral dataset\cite{Armbruster2015}.
+Even completely unlabeled data can enable machine learning researchers to evaluate how closely their simulations represent experimental data, identify what features are missing, and modify their simulation algorithms accordingly to arrive at a simulation that mirrors how data appears in reality.
+In this way, one can bring the simulated patterns closer to those observed in experiments by more accurately modeling the scattering process.
 
-Furthermore, even completely unlabeled data can enable machine learning researchers to evaluate how closely their simulations represent experimental data, identify what features are missing, and modify their simulation algorithms accordingly to arrive at a simulation that mirrors how data appears in reality.
-In this way, one can bring the simulated patterns closer to those observed in experiments by more accurately modeling the scattering process.
+\begin{figure*}[!htb]
+    \centering
+    \includegraphics[width=\linewidth]{figures/overview.pdf}
+    \caption{Overview of the data collection pipeline.
+Datasets are submitted using an online submission form, optionally with the help of our submission helper software.
+After post-processing and data homogenization, we offer the creation of a Zenodo entry for each user submission and subsequently include the submission in the opXRD database.}
+    \label{fig:overview}
+\end{figure*}
+
 Improving the simulation of data involves accounting for more of the physical effects that influence the scattering process such as preferred grain orientation, variations in grain size, crystal defects, the impact of temperature on the scattering process, internal stress, and X-ray-induced fluorescence\cite{cao2024simxrd, Waseda2011, Pecharsky2023}.
 Additionally, it should be considered that varying experimental setups can produce distinct powder diffraction patterns on the same sample.
-Features that may vary between experimental setups and influence the recorded pattern include the shape of diffraction peaks, the wavelength and polarization of the employed X-ray source as well as potentially non-monochromatic radiation, and the detector geometry
+Features that may vary between experimental setups and influence the recorded pattern include the shape of diffraction peaks, the wavelength, and polarization of the employed X-ray source as well as potentially non-monochromatic radiation, and the detector geometry
 \cite{cao2024simxrd, Waseda2011, Pecharsky2023}.
@@
 Many approaches to transfer learning have been proposed, particularly in the domain of image classification \cite{Gatys2016, Ganin2015}.
-These existing techniques can be adapted to facilitate transfer learning in the context of pXRD patterns.\\
+These existing techniques can be adapted to facilitate transfer learning in the context of pXRD patterns.
+Seddiki et al.
+have successfully applied transfer learning in the domain of mass spectrometry to boost the accuracy of mass spectrum classification models\cite{Seddiki2020}.
+Since both mass spectrometry data and pXRD data are one-dimensional, this work demonstrates the merit of transfer learning in a setting similar to pXRD.
+\\
 
@@
 A detailed description of how to use the data is given in Section~$\eqref{sec:how_to_use}$, and Section~$\eqref{sec:summary_and_outlook}$ describes how further data can be contributed.
-
-\begin{figure*}[!htb]
-    \centering
-    \includegraphics[width=\linewidth]{figures/overview.pdf}
-    \caption{Overview of the data collection pipeline.
-Datasets are submitted using an online submission form, optionally with the help of our submission helper software.
-After post-processing and data homogenization, we offer the creation of a Zenodo entry for each user submission and subsequently include the submission in the opXRD database.}
-    \label{fig:overview}
-\end{figure*}
 

### 2025-03-06 chapters/02_SOTA_ML.tex
Restructured paragraph discussing significance of data

@@
 
-The option of training purely on experimental diffractograms is infeasible for neural networks because of the limited availability of labeled experimental diffractograms.
+Neural networks trained purely on experimental diffractograms can perform well when the range of samples is narrow and the data is collected only on a single machine \cite{Lee2023, hattrick-simpers2021}.
+However, in a more general setting with a wide range of investigated samples and employed diffractometers training neural networks purely on experimental diffractograms becomes infeasible.
+This is because of the limited availability of labeled experimental diffractograms relative to the scope of the task.
 However, in 2023, Salgado {\it et al.} \cite{Salgado2023} showed that adding a fraction of experimental patterns to a simulated training dataset improves the performance on unseen experimental patterns.

### 2025-03-06 chapters/03_existing_datasets.tex
Restructured paragraph discussing significance of data

@@
 \centering
-\caption{\footnotesize Overview of experimental powder diffraction databases: The column ``O.A.'' indicates whether or not the database is open access.
+\caption{\footnotesize Overview of experimental powder diffraction databases: The column ``O.A.'' indicates whether or not the database is open-access.
 The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'', respectively.}
@@
 
+While not strictly speaking a powder diffraction database, the High Throughput Experimental Materials Database (HTEM) by the National Renewable Energy Laboratory (NREL) is also worth mentioning\cite{zakutayev2018}.
+As of the time of writing, the HTEM database contains 65,779 (https://htem.nrel.gov/stats) thin-film samples with corresponding X-ray diffraction data.
+Each database entry includes the elemental composition of the underlying sample but does not provide any information on its structure.
+HTEM data is open-access and can be downloaded through an API provided by NREL.
+\\
+
 Aside from the databases mentioned above, we have also investigated several other crystal structure resources in search of experimental powder diffraction data.

### 2025-03-06 main.tex
Restructured paragraph discussing significance of data

@@
 \author[\HKUSTGZ]{Bin Cao}
+\author[\USC]{Adie Alwen}
 \author[\INT]{Simon Schweidler}
-\author[\USC]{Adie Alwen}
 \author[\LBLMF]{Mriganka Singh}
@@
 \author[\LBLCS]{Gregoire Heymans}
+\author[\KFUPMMSE,\KFUPMIRCIMR]{Maged Abdelsamie}
 \author[\EMPA]{Alexander Wieczorek}
@@
 \author[\CPT]{Arthur Hardiagon}
-\author[\KFUPMMSE,\KFUPMIRCIMR]{Maged Abdelsamie}
-%\author[\HKUSTGZ]{Longhan Zhang}
+\author[\FZJIEMD]{Ruth Schwaiger}
+\author[\CPT]{François-Xavier Coudert}
+\author[\KITICRT]{Moritz Wolf}
+\author[\EMPA]{Sebastian Siol}
+\author[\LBLMF]{Carolin M. Sutter-Fella}
+\author[\INT]{Ben Breitung}
+\author[\USC]{Andrea M. Hodge}
 \author[\HKUSTGZ]{Tong-yi Zhang}
-\author[\USC]{Andrea M. Hodge}
-\author[\FZJIEMD]{Ruth Schwaiger}
-\author[\INT]{Ben Breitung}
-\author[\EMPA]{Sebastian Siol}
-\author[\KITICRT]{Moritz Wolf}
-\author[\CPT]{François-Xavier Coudert}
-\author[\LBLMF]{Carolin M. Sutter-Fella}
 \author[\ITI,\INT]{Pascal Friederich$^\ast$}
@@
 \affil[\INT]{Institute of Nanotechnology, Karlsruhe Institute of Technology (KIT), 76131 Karlsruhe, Germany} % Hermann-von-Helmholtz-Platz 1
-\affil[\HKUSTGZ]{Guangzhou Municipal Key Laboratory of Materials Informatics, Advanced Materials Thrust, Hong Kong University of Science and Technology (Guangzhou) (HKUST), Guangzhou 511400, China} %1 Duxue Road,
+\affil[\HKUSTGZ]{Guangzhou Municipal Key Laboratory of Materials Informatics, Advanced Materials Thrust, Hong Kong University of Science and Technology (Guangzhou) (HKUST), Guangzhou 511400, China} % 1 Du Xue Road,
 \affil[\USC]{Department of Chemical Engineering and Materials Science, University of Southern California (USC), Los Angeles CA 90089, USA} % 925 Bloom Walk
@@
 \affil[\LBLALS]{Advanced Light Source, Lawrence Berkeley National Laboratory,  Berkeley 94720 CA, USA}  % 6 Cyclotron Rd, Berkeley, CA 94720, United States
-\affil[\Hoffmann]{Hoffmann Institute of Advanced Materials, Shenzhen Polytechnic, Shenzhen 518055, China} % ??
+\affil[\Hoffmann]{Hoffmann Institute of Advanced Materials, Shenzhen Polytechnic, Shenzhen 518055, China} % Liuxian Street NO.7098, Shenzhen 518055, China
 \affil[\LBLCS]{Lawrence Berkeley National Laboratory (LBNL), Chemical Sciences Division, Berkeley 94720 CA, USA}  % 1 Cyclotron Rd, Berkeley, CA 94720, United States
@@
 \maketitle
-\textbf{Keywords}: powder x-ray diffraction, machine learning, open-access database, crystal structure determination, phase identification
+\textbf{Keywords}: powder x-ray diffraction, machine learning, open-access data, data sharing, crystal structure determination, phase identification
 
@@
 
-\pagebreak
+%\pagebreak
 \section{opXRD database}\label{sec:our_dataset}
@@
 Funding by the Helmholtz Research Program “Materials and Technologies for the Energy Transition (MTET), Topic 3: Chemical Energy Carriers" is kindly acknowledged.
+Work by the Empa group was supported by the Strategic Focus Area–Advanced Manufacturing (SFA–AM) through the project Advancing manufacturability of hybrid organic–inorganic semiconductors for large area optoelectronics (AMYS) as well as the Empa internal research call 2020.
 

### 2025-03-06 chapters/01_introduction.tex
Cut Problems Rietveld refinement section in half

@@
 It is not scalable to the degree required to keep up with advances in throughput and efficiency in other steps of the experimentation pipeline.
-In a Rietveld refinement process, the operator starts by setting parameters that characterize the background and peak shape of the powder diffraction pattern and determines a crystal structure model from which the optimization process can start.
-This model is usually obtained using search-match software, which identifies crystal structures with similar powder diffraction patterns from a database of crystal structures with accompanying powder diffraction patterns.
-Databases commonly used for this purpose include the Powder Diffraction File\cite{GatesRector2019}, the Inorganic Crystal Structure Database \cite{Belsky2002}, and the Crystallography Open Database \cite{Grazulis2009}.
-However, an initial structural model obtained from such a database via search-match software is not guaranteed to lead to an accurate structure solution through Rietveld refinement.
+The refinement process requires the operator to determine an initial structural model from which the refinement can start and initial values for parameters that characterize the background.
+%In a Rietveld refinement process, the operator starts by setting parameters that characterize the background and peak shape of the powder diffraction pattern and determines an initial crystal structure model from which the optimization process can start.
+The structural model is usually obtained using search-match software, which identifies crystal structures with similar powder diffraction patterns from a database of crystal structures with accompanying powder diffraction patterns.
+%Databases commonly used for this purpose include the Powder Diffraction File\cite{GatesRector2019}, the Inorganic Crystal Structure Database \cite{Belsky2002}, and the Crystallography Open Database \cite{Grazulis2009}.
+However, an initial structural model obtained from such a database is not guaranteed to lead to an accurate structure solution through Rietveld refinement, especially not for novel structures. 
+%In such cases, more complex crystal structure solution approaches need to be applied to find suitable crystal structure guesses \cite{structure_solution_review}.\\%.
 Especially for novel structures, obtaining an initial structural model through matching to a database of known structures can be problematic.
-In such cases, more complex crystal structure solution approaches need to be applied to find suitable crystal structure guesses \cite{structure_solution_review}.\\
-
-% VISUAL ABSTRACT FIGURE
-\begin{figure}[!b]
-    \centering
-    \includegraphics[width=1.0\linewidth]{figures/pipeline.pdf}
-    \caption{Experimental powder X-ray diffraction (pXRD) patterns from several contributors are collected in the \textit{opXRD} database.
-The proposed open-access database of experimental data aims to support each step in the pXRD-related machine learning workflow by informing better physics simulations, supplying model training data, and providing a foundation for realistic performance evaluations.}
-    \label{fig:overview}
-\end{figure}
-
-Additionally, attempting to refine all crystal structure parameters at once, starting from this initial structure, is known to lead to unphysical results\cite{Ozaki2020}.
-Hence parameters are refined iteratively, with each iteration only refining a limited set of parameters.
-If, during the iterative determination of parameters, the structural model becomes unphysical the operator must intervene to backtrack to a previous successful step or start over entirely.
+Additionally, attempting to refine all crystal structure parameters at once is known to lead to unphysical results\cite{Ozaki2020}.
+Hence parameters are refined iteratively, with each iteration only refining a limited set of parameters.%If during the iterative determination of parameters, the structural model becomes unphysical the operator must intervene to backtrack to a previous successful step or start over entirely.
 Finding the correct order in which to refine structure parameters and finding the correct values for fixed parameters both present problems that add to the difficulty of the refinement process.\\
@@
 Even completely unlabeled data can enable machine learning researchers to evaluate how closely their simulations represent experimental data, identify what features are missing, and modify their simulation algorithms accordingly to arrive at a simulation that mirrors how data appears in reality.
-In this way, one can bring the simulated patterns closer to those observed in experiments by more accurately modeling the scattering process.
+Unlabeled data could also find applications in transfer learning approaches to transfer model capabilities from the domain of simulated diffractograms to the domain of experimental diffractograms.
+Below Fig.~$\eqref{fig:overview}$ provides an overview of the uses of the opXRD database in the context of machine learning.
+\\
 
-\begin{figure*}[!htb]
+% VISUAL ABSTRACT FIGURE
+\begin{figure}[!htb]
     \centering
-    \includegraphics[width=\linewidth]{figures/overview.pdf}
-    \caption{Overview of the data collection pipeline.
-Datasets are submitted using an online submission form, optionally with the help of our submission helper software.
-After post-processing and data homogenization, we offer the creation of a Zenodo entry for each user submission and subsequently include the submission in the opXRD database.}
+    \includegraphics[width=1.0\linewidth]{figures/pipeline.pdf}
+    \caption{Experimental powder X-ray diffraction (pXRD) patterns from several contributors are collected in the \textit{opXRD} database.
+The proposed open-access database of experimental data aims to support each step in the pXRD-related machine learning workflow by informing better physics simulations, supplying model training data, and providing a foundation for realistic performance evaluations.}
     \label{fig:overview}
-\end{figure*}
+\end{figure}
 
@@
 
+\begin{figure*}[!htb]
+    \centering
+    \includegraphics[width=\linewidth]{figures/overview.pdf}
+    \caption{Overview of the data collection pipeline.
+Datasets are submitted using an online submission form, optionally with the help of our submission helper software.
+After post-processing and data homogenization, we offer the creation of a Zenodo entry for each user submission and subsequently include the submission in the opXRD database.}
+    \label{fig:overview}
+\end{figure*}
+
 As argued by Aranda and Kroon-Batenburg \textit{et al}.\cite{Aranda2018, Kroon-Batenburg2024}, sharing raw powder diffraction data is not only in the interest of furthering machine learning research but is also in line with open science principles.

### 2025-03-06 chapters/01_introduction.tex
Added test data comparison

@@
 It is not scalable to the degree required to keep up with advances in throughput and efficiency in other steps of the experimentation pipeline.
-The refinement process requires the operator to determine an initial structural model from which the refinement can start and initial values for parameters that characterize the background.
-%In a Rietveld refinement process, the operator starts by setting parameters that characterize the background and peak shape of the powder diffraction pattern and determines an initial crystal structure model from which the optimization process can start.
+The refinement process requires the operator to determine an initial structural model from which the refinement can start and as well as initial values for parameters that characterize the background \cite{mccusker1999}.
 The structural model is usually obtained using search-match software, which identifies crystal structures with similar powder diffraction patterns from a database of crystal structures with accompanying powder diffraction patterns.
-%Databases commonly used for this purpose include the Powder Diffraction File\cite{GatesRector2019}, the Inorganic Crystal Structure Database \cite{Belsky2002}, and the Crystallography Open Database \cite{Grazulis2009}.
 However, an initial structural model obtained from such a database is not guaranteed to lead to an accurate structure solution through Rietveld refinement, especially not for novel structures.
-%In such cases, more complex crystal structure solution approaches need to be applied to find suitable crystal structure guesses \cite{structure_solution_review}.\\%.
-Especially for novel structures, obtaining an initial structural model through matching to a database of known structures can be problematic.
 Additionally, attempting to refine all crystal structure parameters at once is known to lead to unphysical results\cite{Ozaki2020}.
-Hence parameters are refined iteratively, with each iteration only refining a limited set of parameters.%If during the iterative determination of parameters, the structural model becomes unphysical the operator must intervene to backtrack to a previous successful step or start over entirely.
-Finding the correct order in which to refine structure parameters and finding the correct values for fixed parameters both present problems that add to the difficulty of the refinement process.\\
+Hence parameters are refined iteratively, with each iteration only refining a limited set of parameters.
+Finding the correct order in which to refine structure parameters and finding the correct values for initial background parameters both present problems that add to the difficulty of the refinement process.\\
 
+% Simulated pattern machine learning successes + Sim/Exp performance gap
 Machine learning has the potential to speed up the manual analysis of powder diffractograms and keep pace with an automated high-throughput experimentation environment\cite{Agrawal2019, Surdu2023}.
-Models can be trained to predict crystal structure information directly given a diffractogram, or models can be used to automate the conventional workflow of finding an initial guess \cite{Surdu2023} and running ML-automated Rietveld refinement \cite{Feng2019}.
-So far, however, due to an absence of labeled datasets with experimental diffractograms\cite{Wang2020}, machine learning in this domain has largely relied on diffractograms simulated from known structures\cite{Park2017, Lee2023} or, most recently, from generated synthetic crystals\cite{Schopmans2023}.
+Models can be trained to predict crystal structure information directly given a diffractogram, or models can be used to automate the conventional refinement workflow.
+In the latter case a model would first predict an initial crystal structure \cite{Surdu2023} which is then refined by a second model trained to perform the refinement process \cite{Feng2019}.
+So far due to an absence of labeled datasets with experimental diffractograms\cite{Wang2020}, machine learning in this domain has largely relied on diffractograms simulated from known structures\cite{Park2017, Lee2023} or, most recently, from generated synthetic crystals\cite{Schopmans2023}.
+Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting lattice parameters\cite{Dong2021, Chitturi2021, Habershon2004, zhang2024crystallographic}, spacegroup \cite{cao2024simxrd, Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021,zhang2024crystallographic}, and crystallite size \cite{Dong2021, Chakraborty2021} from simulated diffractograms.
+However, the performance substantially drops off when these models are applied to data originating from experiments \cite{cao2024simxrd, Schopmans2023,zhang2024crystallographic, Wang2020, Vecsei2018}.
+This discrepancy in performance arises due to imperfections in experimental data which are not present in diffraction patterns modeled under ideal conditions.
+This is discussed in more detail below.
 \\
 
-% Predicting structure information with machine learning
-Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting lattice parameters\cite{Dong2021, Chitturi2021, Habershon2004, zhang2024crystallographic}, spacegroup \cite{cao2024simxrd,Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021,zhang2024crystallographic}, and crystallite size \cite{Dong2021, Chakraborty2021} from simulated diffractograms.
-However, the performance substantially drops off when these models are applied to data originating from experiments \cite{cao2024simxrd,Schopmans2023,zhang2024crystallographic, Wang2020, Vecsei2018} The patterns observed in experiments exhibit imperfections that do not appear when simulating the X-ray diffraction pattern of a powder sample under ideal conditions.
-The physical effects underlying these imperfections need to be modeled accurately in order to be able to transfer models trained on simulated data to experimental patterns.
-Most approaches currently model these imperfections as a simple signal based on a heuristic functional form added to the simulated diffractograms.
-Experimental data can provide clues on how to model these imperfections in a more realistic fashion.
+Both labeled and unlabeled datasets of experimental powder diffractograms hold significant value for machine learning-based pXRD analysis, particularly with regard to bridging the performance gap between simulated and experimental domains.
+Labeled experimental data can be used to test and benchmark existing and new automated analysis approaches.
+This enables researchers to gauge how well a given model would perform under real-world conditions if integrated into an automated experimentation pipeline.
+Unlabeled experimental data enables machine learning researchers to evaluate how closely their simulations represent experimental data and modify their simulation algorithms accordingly.
+Unlabeled data could also find applications in transfer learning approaches to transfer model capabilities from the domain of simulated diffractograms to the domain of experimental diffractograms.
+While some experimental powder databases exist, their utility is limited by the fact that they are either small or not openly accessible.
 \\
 
-At the moment, the number of openly available datasets for experimental powder data is very limited.
-The main source for experimental patterns is the RRUFF database\cite{lafuente2015}.
-Further commercial datasets such as the PDF5+\cite{GatesRector2019} and the Linus Pauling File\cite{villars2018} exist, but license fees and restrictions in their use to train publishable models or even publish the training data limit their usability.
-Collecting a labeled dataset of raw experimental powder X-ray diffraction data of sufficient size to directly train a model that predicts crystal structure information is currently out of reach.
-The most advanced ML models currently are trained on approximately $10^5 - 10^6$ simulated diffractograms.
-\cite{Salgado2023, Schopmans2023} This is, to the best of our knowledge, two orders of magnitude larger than the largest currently curated experimental dataset, the PDF-5+ with approximately $2\cdot 10^4$ experimental patterns.
-It is even one order of magnitude larger than the approximately $10^5$ unlabeled diffractograms in the initial version of the opXRD dataset we present here.\\
-
-Both labeled and unlabeled datasets of experimental diffractograms hold significant value for machine learning-based pXRD analysis.
-Labeled opXRD data can be used to test and benchmark existing and new automated analysis approaches, which yields a better comparison than when using the very limited currently available datasets, such as the commonly used RRUFF mineral dataset\cite{Armbruster2015}.
-Even completely unlabeled data can enable machine learning researchers to evaluate how closely their simulations represent experimental data, identify what features are missing, and modify their simulation algorithms accordingly to arrive at a simulation that mirrors how data appears in reality.
-Unlabeled data could also find applications in transfer learning approaches to transfer model capabilities from the domain of simulated diffractograms to the domain of experimental diffractograms.
-Below Fig.~$\eqref{fig:overview}$ provides an overview of the uses of the opXRD database in the context of machine learning.
+In this work, we introduce an open powder X-ray diffraction (opXRD) database featuring a broad range of patterns collected from experiments.
+With a total of $92,552$ patterns collected from 6 contributing institutions, the opXRD database exceeds the size of the previously largest database of openly accessible experimental powder diffraction data by two orders of magnitude.
+To the best of our knowledge, the largest database of this type is the RRUFF database, containing 1290 experimental powder diffraction patterns \cite{lafuente2015}.
+Further commercial datasets such as the PDF5+\cite{GatesRector2019} and the Linus Pauling File\cite{villars2018} exist, but license fees and restrictions in their use to train publishable models or publish the training data limit their usability.
+Fig.~$\eqref{fig:overview}$ provides an overview of machine learning workflows enabled and supported by the opXRD database.
 \\
@@
 
-Improving the simulation of data involves accounting for more of the physical effects that influence the scattering process such as preferred grain orientation, variations in grain size, crystal defects, the impact of temperature on the scattering process, internal stress, and X-ray-induced fluorescence\cite{cao2024simxrd, Waseda2011, Pecharsky2023}.
+Of the \numpatterns patterns in the opXRD database \numlabeled pattern are at least partially labeled with structural data and more than 900 patterns possess full structural labels including atomic coordinates.
+This constitutes an experimental pXRD test dataset that is larger in size, richer in labels and broader in represented experimental diffraction setups than the RRUFF database which only provides lattice parameters as labels\cite{Armbruster2015}.
+However, since the majority of the opXRD database is unlabeled we also want to further discuss the uses of unlabeled data with regards to improving pattern simulations and transfer learning.
+\\
+
+These imperfections include preferred crystallations orientation, variations in grain size, crystal defects, the impact of temperature on the scattering process, internal stress, the non-monochromacity of the X-ray source and X-ray-induced fluorescence\cite{cao2024simxrd, Waseda2011, Pecharsky2023}.
 Additionally, it should be considered that varying experimental setups can produce distinct powder diffraction patterns on the same sample.
-Features that may vary between experimental setups and influence the recorded pattern include the shape of diffraction peaks, the wavelength, and polarization of the employed X-ray source as well as potentially non-monochromatic radiation, and the detector geometry
-\cite{cao2024simxrd, Waseda2011, Pecharsky2023}.
-Some experimental setups may also exhibit imperfections such as displacement of the sample from its intended position leading to the recorded scattering angles being slightly off \cite{cao2024simxrd,hulbert2023} or background events such as instrument noise or X-rays scattering off equipment onto the screen.\\
+Features that may vary between experimental setups include the shape of diffraction peaks, the wavelength and polarization of the employed X-ray source, and the detector geometry \cite{cao2024simxrd, Waseda2011, Pecharsky2023}.
+The recorded scattering angles may also be slightly falsified if the sample is displaced from its intended position\cite{cao2024simxrd,hulbert2023}.
 
@@
 
-In this publication, we introduce an open powder X-ray diffraction (opXRD) database that collects a broad range of patterns from experiments.
-With a total of $92,552$ patterns, the opXRD database exceeds the size of the previously largest database of openly accessible experimental powder diffraction data by two orders of magnitude.
-To the best of our knowledge, up to now the largest database of this type was the RRUFF database, containing 1,290 experimental powder diffraction patterns.
 Our opXRD dataset has been curated by collecting the accumulated data from multiple research groups and institutions with high-throughput XRD facilities.
@@
 Some of the XRD data was collected on thin-films rather than on true powder samples, which may influence the quality of the data in regards to full structure resolution.
-Additionally some of the data was collected in grazing-angle geometry rather than in the usual Bragg-Brentano geometry employed in powder diffraction.
+Additionally, some of the data was collected in grazing-angle geometry rather than in the usual Bragg-Brentano geometry employed in powder diffraction.
 Details about the experiments of contributing research groups, including the presence of thin-films samples and the usage of atypical diffraction geometries are discussed in Section~$\eqref{sec:our_dataset}$.
 A detailed description of how to use the data is given in Section~$\eqref{sec:how_to_use}$, and Section~$\eqref{sec:summary_and_outlook}$ describes how further data can be contributed.
+\\
+
+\begin{figure*}[!htb]
+    \centering
+    \includegraphics[width=\linewidth]{figures/overview.pdf}
+    \caption{Overview of the data collection pipeline.
+Datasets are submitted using an online submission form, optionally with the help of our submission helper software.
+After post-processing and data homogenization, we offer the creation of a Zenodo entry for each user submission and subsequently include the submission in the opXRD database.}
+    \label{fig:overview}
+\end{figure*}
 
@@
 
-\begin{figure*}[!htb]
-    \centering
-    \includegraphics[width=\linewidth]{figures/overview.pdf}
-    \caption{Overview of the data collection pipeline.
-Datasets are submitted using an online submission form, optionally with the help of our submission helper software.
-After post-processing and data homogenization, we offer the creation of a Zenodo entry for each user submission and subsequently include the submission in the opXRD database.}
-    \label{fig:overview}
-\end{figure*}
-
 As argued by Aranda and Kroon-Batenburg \textit{et al}.\cite{Aranda2018, Kroon-Batenburg2024}, sharing raw powder diffraction data is not only in the interest of furthering machine learning research but is also in line with open science principles.

### 2025-03-06 chapters/02_SOTA_ML.tex
Added test data comparison

@@
+\subsection*{Review of machine learning-based pXRD analysis}
 To showcase the need for datasets such as the one presented in this publication, we now shortly discuss some recent
@@
 
+At the moment, the number of openly available datasets for experimental powder data is very limited.
+The main source for experimental patterns is the RRUFF database\cite{lafuente2015}.
+Collecting a labeled dataset of raw experimental powder X-ray diffraction data of sufficient size to directly train a model that predicts crystal structure information is currently out of reach.
+The most advanced ML models currently are trained on approximately $10^5 - 10^6$ simulated diffractograms.
+\cite{Salgado2023, Schopmans2023} This is, to the best of our knowledge, two orders of magnitude larger than the largest currently curated experimental dataset, the PDF-5+ with approximately $2\cdot 10^4$ experimental patterns.
+It is even one order of magnitude larger than the approximately $10^5$ unlabeled diffractograms in the initial version of the opXRD dataset we present here.\\
+
 The ultimate goal of ML efforts is to make pXRD data identification practical for experimental use, automating structure prediction.

### 2025-03-06 chapters/03_existing_datasets.tex
Added test data comparison

--- 2025-03-01 chapters/03_existing_datasets.tex
@@
 There is also publicly available powder diffraction data uploaded to datasets on Zenodo.
-However, this data is split into disparate entries that typically only contain the work of a single research project.
-Additionally, extracting powder diffraction data at scale is hindered by the fact that the data is often given in plain text files in non-standardized formats, which are difficult to parse programmatically.
+However, this data is split into disparate entries that typically only contains the work of a single research project.
+Additionally, extracting powder diffraction data at scale is hindered by the fact that the data is often given in plain text files in non-standardized formats, which are difficult to parse automatically.
 We are currently planning a systematic large-scale extraction of powder diffraction data from databases like Zenodo with the help of a large language model.
-This data will be included in a future release of our database.
+This data will be included in a future release of the opXRD database.
 \\
+
+While not strictly speaking a powder diffraction database, the High-Throughput Experimental Materials Database (HTEM) by the National Renewable Energy Laboratory (NREL) is a valuable source of X-ray diffraction data \cite{zakutayev2018}.
+Currently, the HTEM database (https://htem.nrel.gov/stats) contains 65,779 thin-film samples with corresponding X-ray diffraction data.
+Each database entry includes the elemental composition of the underlying sample but does not provide any information on its structure.
+HTEM data is open-access and can be downloaded through an API provided by NREL.\\
 

### 2025-03-06 chapters/01_introduction.tex
Updated state of the art ML

@@
 Unlabeled experimental data enables machine learning researchers to evaluate how closely their simulations represent experimental data and modify their simulation algorithms accordingly.
-Unlabeled data could also find applications in transfer learning approaches to transfer model capabilities from the domain of simulated diffractograms to the domain of experimental diffractograms.
+Unlabeled data can also find applications in transfer learning approaches to transfer model capabilities from the domain of simulated diffractograms to the domain of experimental diffractograms.
 While some experimental powder databases exist, their utility is limited by the fact that they are either small or not openly accessible.
@@
 To the best of our knowledge, the largest database of this type is the RRUFF database, containing 1290 experimental powder diffraction patterns \cite{lafuente2015}.
-Further commercial datasets such as the PDF5+\cite{GatesRector2019} and the Linus Pauling File\cite{villars2018} exist, but license fees and restrictions in their use to train publishable models or publish the training data limit their usability.
-Fig.~$\eqref{fig:overview}$ provides an overview of machine learning workflows enabled and supported by the opXRD database.
+Larger commerical datasets such as the PDF5+\cite{GatesRector2019} and the Linus Pauling File\cite{villars2018} exist, but their utility is limited by fees and restrictive licenses.
+License terms of commercial datasets such as the PDF5+ and the Linus Pauling File prohibit or restrict the publication of models trained on their data.
+In contrast, the opXRD database is both free and imposes no restrictions on how its data is used.
+Fig.~$\eqref{fig:ml_uses}$ provides an overview of machine learning workflows enabled and supported by the opXRD database.
 \\
@@
 The proposed open-access database of experimental data aims to support each step in the pXRD-related machine learning workflow by informing better physics simulations, supplying model training data, and providing a foundation for realistic performance evaluations.}
-    \label{fig:overview}
+    \label{fig:ml_uses}
 \end{figure}
@@
 
-These imperfections include preferred crystallations orientation, variations in grain size, crystal defects, the impact of temperature on the scattering process, internal stress, the non-monochromacity of the X-ray source and X-ray-induced fluorescence\cite{cao2024simxrd, Waseda2011, Pecharsky2023}.
-Additionally, it should be considered that varying experimental setups can produce distinct powder diffraction patterns on the same sample.
+The neglected effects that lead to discrepancies between simulated patterns and patterns stemming from experiments are largely known.
+Unaccounted effects may include preferred crystallite orientation, variations in grain size, crystal defects, the impact of temperature on the scattering process, internal stress, the non-monochromaticity of the X-ray source, and X-ray-induced fluorescence\cite{cao2024simxrd, Waseda2011, Pecharsky2023}.
+Additionally, varying experimental setups produce distinct powder diffraction patterns on the same sample.
 Features that may vary between experimental setups include the shape of diffraction peaks, the wavelength and polarization of the employed X-ray source, and the detector geometry \cite{cao2024simxrd, Waseda2011, Pecharsky2023}.
 The recorded scattering angles may also be slightly falsified if the sample is displaced from its intended position\cite{cao2024simxrd,hulbert2023}.
+As these and more neglected effects are integrated into the simulation process, real powder diffraction data can be used to evaluate how closely simulated data matches up with real data.
+While direct comparisons are only possible on labeled patterns, comparing the strength and prevalence of features between simulated and real data can nevertheless provide clues about the fidelity of the situation.
+Taking on board all neglected effects without making approximations will incur significant computational costs that will lower the size of the generated training data.
+A more efficient approach could be to use real experimental data to identify the effects that have the largest impact in practice and model them heuristically.
+\\
 
-Additionally, some of the observed imperfections can be accounted for using transfer learning approaches.
-In transfer learning the objective is to transfer the capabilities of a model learned on a source domain in which labeled data is abundant to a target domain in which labeled data is sparse\cite{Zhuang2021}.
+The second way in which unlabeled experimental data can serve to bridge the performance gap between simulated and experimental domains is through transfer learning.
+The objective of transfer learning is to transfer the capabilities of a model learned on a source domain in which labeled data is abundant to a target domain in which labeled data is sparse\cite{Zhuang2021}.
 In this context, the source domain is simulated powder diffraction patterns and the target domain is experimental powder diffraction patterns.
@@
 These existing techniques can be adapted to facilitate transfer learning in the context of pXRD patterns.
-Seddiki et al.
-have successfully applied transfer learning in the domain of mass spectrometry to boost the accuracy of mass spectrum classification models\cite{Seddiki2020}.
+Additionally, Seddiki et al.
+have already successfully applied transfer learning in the domain of mass spectrometry to boost the accuracy of mass spectrum classification models\cite{Seddiki2020}.
 Since both mass spectrometry data and pXRD data are one-dimensional, this work demonstrates the merit of transfer learning in a setting similar to pXRD.
@@
 
-Our opXRD dataset has been curated by collecting the accumulated data from multiple research groups and institutions with high-throughput XRD facilities.
-It contains pXRD patterns from single and multiphase materials from a wide variety of material classes, including high-entropy materials, perovskites, and commercial catalysts.
-Some of the XRD data was collected on thin-films rather than on true powder samples, which may influence the quality of the data in regards to full structure resolution.
-Additionally, some of the data was collected in grazing-angle geometry rather than in the usual Bragg-Brentano geometry employed in powder diffraction.
-Details about the experiments of contributing research groups, including the presence of thin-films samples and the usage of atypical diffraction geometries are discussed in Section~$\eqref{sec:our_dataset}$.
-A detailed description of how to use the data is given in Section~$\eqref{sec:how_to_use}$, and Section~$\eqref{sec:summary_and_outlook}$ describes how further data can be contributed.
+% Contributing + community driven
+The database is intended as a growing, community-driven initiative.
+The database we present here is the first version, but we hope to further increase the database size through active engagement with the pXRD community.
+Our primary objective is to minimize the effort and thus the barrier to contributing experimental data to the opXRD database.
+Thus, we developed a program that helps to find and share data from pXRD lab computers.
+Users can select their most common pXRD file types, the program lists all files of that type, and users can select or deselect certain folders or files for sharing.
+Selected contributions will be uploaded to opXRD, processed to a common file format, and---if wanted---published on Zenodo on behalf of the contributors, before becoming part of the opXRD database.
+If labels are available, they can be shared with opXRD as well.
+Further details can be found on the opXRD website (\url{https://xrd.aimat.science/}).
+An overview of this process is given in below Fig.~$\eqref{fig:submission}$.
 \\
@@
 After post-processing and data homogenization, we offer the creation of a Zenodo entry for each user submission and subsequently include the submission in the opXRD database.}
-    \label{fig:overview}
+    \label{fig:submission}
 \end{figure*}
-
-% How to contribute? This aspect deserves a separate paragraph
-The opXRD database is intended as a growing, community-driven initiative.
-The dataset we present here is the first version, but we hope to further increase the dataset size through active engagement with the pXRD community.
-Our primary objective is to minimize the effort and thus the barrier to contributing experimental data to the opXRD database.
-Thus, we developed a program that helps to find and share data from pXRD lab computers.
-In detail, users can select their most common pXRD file types, the program lists all files of that type, and users can select or deselect certain folders or files for sharing.
-Selected contributions will be uploaded to opXRD, processed to a common file format, and---if wanted---published on Zenodo on behalf of the contributors, before becoming part of the opXRD database.
-If labels are available, they can be shared with opXRD as well.
-Further details can be found on the opXRD website (\url{https://xrd.aimat.science/}).
-\\
 
@@
 It furthers the ability of other researchers to reproduce published work and in turn, adds to the credibility of the publisher of the data.
-Considering our streamlined submission and post-processing framework, a publication of powder diffraction data through the opXRD database is a simple way of realizing these principles.
 Compared to publishing data individually, it also has the added benefit that the data is contributed towards a large, homogenous dataset with a standardized interface.
@@
 
+The opXRD database contains pXRD patterns from single and multiphase materials from a wide variety of material classes, including high-entropy materials, perovskites, and commercial catalysts.
+Some of the XRD data was collected on thin-films rather than on true powder samples, which may influence the quality of the data in regards to full structure resolution.
+Additionally, some of the data was collected in grazing-angle geometry rather than in the usual Bragg-Brentano geometry employed in powder diffraction.
 The broad range of available experimental samples contained in the opXRD v1.0 database makes it possible to apply state-of-the-art ML approaches to the domain of powder XRD analysis.
-We hope that this can drive ML research in this field towards more advanced automated analysis workflows that can accelerate materials science research through ready application in high-throughput experimentation pipelines.\\
+We hope that the opXRD database can drive ML research in this field towards more advanced automated analysis workflows that can accelerate materials science research through ready application in high-throughput experimentation pipelines.
+Details of the experiments of research groups contributing to the opXRD database are discussed in Section~$\eqref{sec:our_dataset}$.
+A detailed description of how to acquire and use opXRD data is given in Section~$\eqref{sec:how_to_use}$, and Section~$\eqref{sec:summary_and_outlook}$ describes how further data can be contributed.

### 2025-03-06 chapters/02_SOTA_ML.tex
Updated state of the art ML

--- 2025-03-06 chapters/02_SOTA_ML.tex
@@
-To showcase the need for datasets such as the one presented in this publication, we now shortly discuss some recent
-approaches that apply machine learning methods to classification and regression tasks for powder diffractograms. \\
+\subsection*{Review of machine learning-based pXRD analysis}
+To showcase the need for datasets such as the one presented in this publication, we now discuss some recent approaches that apply machine learning methods to classification and regression tasks for powder diffractograms.
+\\
 
@@
 
-The ultimate goal of ML efforts is to make pXRD data identification practical for experimental use, automating structure prediction.
-However, the current volume of experimental pXRD patterns is insufficient to effectively train ML models, highlighting an urgent need for a comprehensive experimental pXRD database.
-Two key approaches are essential to address this.
+The current volume of experimental pXRD patterns is insufficient to effectively train ML models, highlighting an urgent need for a comprehensive experimental pXRD database.
+The most advanced ML models currently are trained on approximately $10^5 - 10^6$ simulated diffractograms.
+\cite{Salgado2023, Schopmans2023} This is, to the best of our knowledge, two orders of magnitude larger than the largest currently curated experimental dataset, the PDF-5+ with approximately $2\cdot 10^4$ experimental patterns.
+It is even one order of magnitude larger than the approximately $10^5$ unlabeled diffractograms in the initial version of the opXRD dataset we present here.
+\\
+
+To make ML-based pXRD data identification practical for experimental use despite lacking experimental training data two key approaches are essential.
 First, developing more sophisticated simulation methods to better approximate experimental patterns\cite{cao2024simxrd} by using statistics from experimental diffractograms.
@@
 For both of these steps, the development of opXRD is particularly significant, as it will provide a comprehensive experimental benchmark for the community, allowing fair comparison of baseline models and accurate evaluation of their applicability in real experimental situations.
+
+\pagebreak

### 2025-03-07 chapters/02_SOTA_ML.tex
Updated SI formatting

@@
 
-To make ML-based pXRD data identification practical for experimental use despite lacking experimental training data two key approaches are essential.
+To make ML-based pXRD data identification practical for experimental use and automate structure prediction despite lacking experimental training data two key approaches are essential.
 First, developing more sophisticated simulation methods to better approximate experimental patterns\cite{cao2024simxrd} by using statistics from experimental diffractograms.

### 2025-03-07 chapters/03_existing_datasets.tex
Updated state of the art ML
Updated SI formatting

@@
 To contextualize opXRD within the current environment of experimental powder diffraction data, the list below provides an overview of the largest crystal structure databases that offer access to experimental powder diffraction data.
-For an overview of these databases refer to Table~$\eqref{tab:exp_databases}$ below.
+For an overview of these databases refer to Tab.~$\eqref{tab:exp_databases}$ below.
 
@@
 \textbf{Powder Diffraction File:} \cite{PDFWeb} The Powder Diffraction File (PDF), published and maintained by the International Center for Diffraction Data (ICDD), is a large collection of materials with accompanying powder diffraction data first published in 1941\cite{GatesRector2019}.
-According to the ICDD the PDF5+, the latest release of the PDF as of November 2024, contains over a million materials with accompanying powder diffraction data.
-However, most of these powder diffraction patterns are simulated.
-After inquiring with the ICDD in April 2024 we were told that only 20,800 of the powder diffraction patterns in the PDF5+ stemmed from experiments.
-Of these 20,800 entries, 10,954 contain information about the atomic coordinates of the underlying structures.
+According to the ICDD the latest release of the PDF, the PDF5+, contains over a million materials with accompanying powder diffraction data.
+However, since most of these powder diffraction patterns are simulated we inquired the ICDD about the number of experimental diffraction patterns in the PDF5+.
+We were told that 20,800 of the patterns in the PDF5+ stemmed from experiments and that 10,954 of these patterns were accompanied by the atomic coordinates of the underlying structures.
 Since the PDF5+ lists the MPDS as a database source, there is likely a significant overlap in the experimental patterns found in the PDF5+ and those found in the Pauling file.
-As of November 2024, the PDF5+ is available to researchers through a purchase of a one-year license starting at a price point of \$6265.
-However, the ICDD does not allow researchers to train machine learning models on PDF5+ data, regardless of whether the resulting models are published (see product license agreement for the PDF5+ at \url{https://www.icdd.com/licensing-process/\#1528471154226-933e5cc6-8da7}).\\
+Currently, the PDF5+ is available to researchers through a purchase of a one-year license starting at a price point of \$6265.
+However, the ICDD does not allow researchers to train machine learning models on PDF5+ data, regardless of whether the resulting models are published \cite{PDFLicenseWeb}.\\
 
@@
 
-\textbf{PowBase} (\url{http://www.cristal.org/powbase/index.html}): PowBase was an initiative suggested in the Structure Determination by Powder Diffractometry (SDPD) mailing list which was hosted on the same site.
+\textbf{PowBase:} \cite{PowBaseWeb} PowBase was an initiative suggested in the Structure Determination by Powder Diffractometry (SDPD) mailing list which was hosted on the same site.
 The COD was another community initiative that grew out of this mailing list.
@@
 There is also publicly available powder diffraction data uploaded to datasets on Zenodo.
-However, this data is split into disparate entries that typically only contains the work of a single research project.
+However, this data is split into disparate entries that typically only contain the work of a single research project.
 Additionally, extracting powder diffraction data at scale is hindered by the fact that the data is often given in plain text files in non-standardized formats, which are difficult to parse automatically.
@@
 This is to be expected as most structure solutions are achieved through single-crystal diffraction rather than powder diffraction.
-\\

### 2025-03-07 chapters/04_our_dataset.tex
Updated state of the art ML
Updated SI formatting

@@
 The following research institutions contributed data to the opXRD database: French National Centre for Scientific Research (CNRS), Hong Kong University of Science and Technology (Guangzhou) (HKUST), University of Southern California (USC), Lawrence Berkeley National Laboratory (LBNL), Empa–Swiss Federal Laboratories for Materials Science and Technology (EMPA) and the Karlsruhe Institute of Technology (KIT).
-\newline
-
+Tab.~$\eqref{tab:merged}$ provides an overview of the contribution made by each institution opXRD database.
 We filtered the submitted datasets to exclude patterns with invalid features such as only one unique recorded angle, negative angles, less than 50 recorded angles total, or all intensities being zero.
-Table~$\eqref{tab:merged}$ provides an overview of the opXRD database.
+\\
 
@@
 CNRS        &  1052          & \ \cmark   & \partialcheck{85}   & \cmark   &  \ \ \quad \partialcheck{85}   & Diffraction data extracted from the COD \\ 
-HKUST(GZ)   &  520          & \ \cmark   & \cmark   & \cmark   & \ \ \quad \cmark   & Phase identification dataset  \\ 
 USC         &  338          & \ \cmark   & \cmark   & \partialcheck{90}   & \ \ \quad \xmark   & Study of CuNi and CuAl alloys          \\
+HKUST(GZ)   &  520          & \ \partialcheck{4}   & \partialcheck{4}   & \partialcheck{4}   & \ \ \quad \partialcheck{4}   & Phase identification dataset  \\ 
 EMPA        &  770          &\ \cmark   & \partialcheck{63}   & \xmark   & \ \ \quad \xmark   & Metal halide perovskites, Zn-V-N libraries \\ 
@@
 For a rigorous definition of PCA and the explained variance ratio, we refer to the literature \cite{Jolliffe2016}.
-\newline
-
 Here, PCA was performed on datasets of X-ray diffraction patterns.
@@
 Here $N_{\text{values}} = N =512$ is the number of recorded intensity values per pattern and $N_{\text{patterns}}$ is the number of patterns in the dataset $X$.
-Figure~$\eqref{fig:components}$ below shows the cumulative explained variance ratio over the fraction of maximal No.
+Fig~$\eqref{fig:components}$ below shows the cumulative explained variance ratio over the fraction of maximal No.
 components $N_{\text{max}}$ as defined above.
 In this figure, a faster convergence of the cumulative variance ratio towards one indicates that the patterns in this data are relatively similar.
-\\
-
 The degree of variation between the patterns is different for each contribution.
@@
 The patterns submitted by LBNL are similar because they stem from in-situ recordings where several hundred or several thousand patterns were collected over time per sample while they were undergoing physical conversion processes.
-\pagebreak
-
-These effects can be observed in Figure~$\eqref{fig:components}$: The cumulative explained variance ratios of the datasets contributed by USC and LBNL converge towards one significantly faster than the datasets contributed by CNRS and HKUST.
 \\
@@
 
-Figure~$\eqref{fig:histograms}$ provides an overview of the distributions of pattern and structure properties in the opXRD dataset.
+Fig.~$\eqref{fig:histograms}$ provides an overview of the distributions of pattern and structure properties in the opXRD dataset.
 Nearly all patterns have an angular resolution smaller than $\Delta(2\theta) = 0.1 ^\circ$.
@@
 Two datasets were contributed to the opXRD database.
-The first dataset (HKUST-A) is a small-scale experimental powder X-ray database developed over the past two years, called the X-Ray Phase Identification Public Experimental Dataset (XRed) (\url{https://github.com/WPEM/XRED}).
+The first dataset (HKUST-A) is a selected subset of a small-scale experimental powder X-ray database developed over the past two years, called the X-Ray Phase Identification Public Experimental Dataset (XRed) (\url{https://github.com/WPEM/XRED}).
 The primary goal of XRed is to support the advancement of intelligent phase identification technology by providing a foundation for data collection in future large-scale machine learning applications.
 XRed primarily focuses on metal and metal-oxide particles, with data collected using diffractometers such as the Empyrean 3.0, Aeris, and Bruker D8 Advance, all employing \ce{Cu} X-ray sources.
-The dataset contains hundreds of pXRD patterns, obtained through XRD refinement, which enable the determination of lattice constants, atomic sizes, and other structural parameters.
-Each pattern is labeled with a corresponding CIF that documents the refined structure.
+The dataset HKUST-A contains 21 pXRD patterns each labeled with a corresponding CIF file that documents the refined structure.
 Data are categorized by elemental systems and include original experimental files, spanning single-phase to five-phase mixtures, as well as mixtures designed for various research tasks.
@@
 \\
-
-%Another data source for opXRD is Xqueryer (\url{https://xqueryer.caobin.asia/}), a recently launched free online platform for intelligent structure identification.
-Users can upload experimental pXRD data to identify crystal structures and extract relevant information from the Materials Project database\cite{jain2013commentary}.
-Uploaded data are securely stored and integrated into the opXRD database, contributing to its development as a large-scale, long-term experimental resource.
-All data usage adheres to Xqueryer's terms of service.
-
-% Paragraph by Adie Alwen and Andrea Hodge
-\subsubsection*{Department of Chemical Engineering and Materials Science, USC}
-
-Combinatorial \ce{CuNi} and \ce{CuAl} samples were synthesized at the University of Southern California (USC) via magnetron co-sputtering from \ce{Cu} (99.999\%) and \ce{Ni} (99.995\%) or \ce{Al} (99.999\%) targets (Plasmaterials) onto stationary $10 \si{cm}$ \ce{Si} (100) substrates yielding 169 unique $5 \times 5 \ \si{mm}$ squares per wafer.
-These alloys were selected to study compositional effects on stacking fault and growth twin boundary formation in sputtered materials \cite{alwen2024_1,alwen2024_2}.
-\\
-
-XRD spectra were collected using a Malvern Panalytical Empyrean X-ray diffractometer at the Forschungszentrum Julich (FZJ) Institute for Energy and Climate Research, Structure and Function of Materials (IEK-2), which enabled automated XRD characterization due to its programmable XY stage and laser sensor for Z height correction.
-XRD analysis was used to link changes in defect formation in the sputtered films with varied phases and crystal structures.
-Each composition was characterized using \ce{Cu} $K_\alpha$ X-rays that were collimated to probe an area of $4 \times 4 \ \si{mm}$ using a step size of $0.026\degree$ and scan duration of $0.3$ seconds per step over a 2 range from $30\degree - 110\degree$.
-Measurement parameters were chosen to resolve at least six or more points above the full-width half maximum of each XRD peak.
 
@@
 The $\ce{GaAs}$ substrates were prepared by Dr.
-Jiro Nishinaga from the National Institute of Advanced Industrial Science and Technology (AIST) in Japan and the glass/CIGS substrates by Dr.
-Christian Kaufmann and his team at Helmholtz-Zentrum Berlin (HZB) in Germany.
+Jiro Nishinaga from the National Institute of Advanced Industrial Science and Technology (AIST) in Japan \cite{nishinaga2018} and the glass/CIGS substrates by Dr.
+Christian Kaufmann and his team at Helmholtz-Zentrum Berlin (HZB) in Germany \cite{heinemann2017}.
 Data collection was performed in situ during thin-film deposition using a custom-made spin-coating and annealing stage \cite{song2019}.

### 2025-03-07 chapters/05_how_to_use.tex
Updated SI formatting

@@
-The opXRD database is hosted on Zenodo (\url{https://zenodo.org/records/14278656}) and can be downloaded by any user without any barriers or restrictions.
-\\
+The opXRD database is hosted on Zenodo (\url{https://zenodo.org/records/14278656}) and can be downloaded by any user without any barriers or restrictions. \\
 
@@
 The repository to this library is located at \url{https://github.com/aimat-lab/opxrd}.
-\\
-
 The opxrd library includes options for data-loading, for standardization, plotting, and the conversion to \emph{PyTorch} tensors.
-We provide a Jupyter Notebook %\footnote{This notebook can be found here: \url{https://colab.research.google.com/github/aimat-lab/opXRD/blob/main/opxrd/usage.ipynb}.}
+We provide a Jupyter Notebook (\url{https://colab.research.google.com/github/aimat-lab/opXRD/blob/main/opxrd/usage.ipynb})
 that showcases these functionalities in more detail. This notebook also illustrates how to interface with the opXRD database through Python.

### 2025-03-07 main.tex
Added test data comparison
Updated SI formatting

@@
 \newcommand{\numpatterns}{92,552 }
+\newcommand{\numlabeled}{2179 }
 
@@
 The opXRD database is available on Zenodo at \url{https://zenodo.org/records/14254270}.
-It is published under the Creative Commons Attribution 4.0 International license.%\footnote{Detailed terms of this license can be found at \url{https://creativecommons.org/licenses/by/4.0/deed.en}.}
-It can be downloaded by any user without any barriers or restrictions. For further details, please refer to Section~$\eqref{sec:how_to_use}$.
+It is published under the Creative Commons Attribution 4.0 International license.It can be downloaded by any user without any barriers or restrictions.
+For further details, please refer to Section~$\eqref{sec:how_to_use}$.
 
@@
 
-%\subsection*{Table of contents}
-%\tableofcontents
-
-%A short text and graphic should be provided for the Table of Contents (ToC).
-The ToC text should describe the main results in 50 to 60 words.
-It should be written for a general audience and be written in the third person.
-
-%The ToC figure should convey the main message of the article.
-It does not have to be a figure from the article; it can be a combination of figures or a new, original figure composed to represent the topic.
-The author must be the copyright holder for this figure and any images used to create it.
-The size of the image should be either 55 mm × 50 mm (w × h) or 110 mm × 20 mm (w × h).
-
 \printnomenclature
@@
 
+\pagebreak
+
+\renewcommand{\thesubsection}{S\arabic{subsection}}
 \section*{Supporting Information}
-%Supporting Information is available from the Wiley Online Library or the author.
+
+\setcounter{page}{1}
+\renewcommand{\thesubsection}{S\arabic{subsection}}
+
+\subsection{Combined pattern plots}
+The Figure below $\eqref{fig:combined}$ shows 50 randomly selected samples of the X-ray diffraction patterns found in each of the research projects contributed to the opXRD database.
+
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=0.95\linewidth]{figures/combined.png}
+    \caption{50 randomly chosen X-ray diffraction patterns from each contributed dataset.
+The figure shows data from the following datasets: a) EMPA, b) LBNL-A, c) LBNL-B, d) LBNL-C, e) USC, f) INT, g) HKUST-A, h) HKUST-B, i) CNRS, j) IKFT.}
+    \label{fig:combined}
+\end{figure}
+
+\pagebreak
 
@@
 
-\pagebreak
-
 \end{document}

### 2025-03-07 SI.tex
Merged final changes

--- 2025-03-07 main.tex
@@
 \usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
+\usepackage{titling}
 
@@
 \interfootnotelinepenalty=10000
-\title{opXRD: Open Experimental Powder X-ray Diffraction Database}
+\title{Supporting information for opXRD: Open Experimental Powder X-ray Diffraction Database}
 
-\newcommand{\ITI}{1}
-\newcommand{\INT}{2}
-\newcommand{\HKUSTGZ}{3}
-\newcommand{\USC}{4}
-\newcommand{\LBLMF}{5}
-\newcommand{\LBLALS}{6}
-\newcommand{\Hoffmann}{7}
-\newcommand{\LBLCS}{8}
-\newcommand{\EMPA}{9}
-\newcommand{\KITICRT}{10}
-\newcommand{\CPT}{11}
-\newcommand{\KFUPMMSE}{12}
-\newcommand{\KFUPMIRCIMR}{13}
-\newcommand{\FZJIEMD}{14}
-
-\author[\ITI,\INT]{Daniel Hollarek}
-\author[\ITI,\INT]{Henrik Schopmans}
-\author[\ITI,\INT]{Jona Östreicher}
-\author[\ITI,\INT]{Jonas Teufel}
-\author[\HKUSTGZ]{Bin Cao}
-\author[\USC]{Adie Alwen}
-\author[\INT]{Simon Schweidler}
-\author[\LBLMF]{Mriganka Singh}
-\author[\LBLMF,\LBLALS]{Tim Kodalle}
-\author[\Hoffmann]{Hanlin Hu}
-\author[\LBLCS]{Gregoire Heymans}
-\author[\KFUPMMSE,\KFUPMIRCIMR]{Maged Abdelsamie}
-\author[\EMPA]{Alexander Wieczorek}
-\author[\EMPA]{Siarhei Zhuk}
-\author[\CPT]{Arthur Hardiagon}
-\author[\FZJIEMD]{Ruth Schwaiger}
-\author[\CPT]{François-Xavier Coudert}
-\author[\KITICRT]{Moritz Wolf}
-\author[\EMPA]{Sebastian Siol}
-\author[\LBLMF]{Carolin M. Sutter-Fella}
-\author[\INT]{Ben Breitung}
-\author[\USC]{Andrea M. Hodge}
-\author[\HKUSTGZ]{Tong-yi Zhang}
-\author[\ITI,\INT]{Pascal Friederich$^\ast$}
-
-\affil[\ITI]{Institute of Theoretical Informatics, Karlsruhe Institute of Technology (KIT), 76131 Karlsruhe, Germany.
-E-mail: pascal.friederich@kit.edu} % Informatik-Hauptgebäude, Am Fasanengarten 5, 76131 Karlsruhe
-\affil[\INT]{Institute of Nanotechnology, Karlsruhe Institute of Technology (KIT), 76131 Karlsruhe, Germany} % Hermann-von-Helmholtz-Platz 1
-\affil[\HKUSTGZ]{Guangzhou Municipal Key Laboratory of Materials Informatics, Advanced Materials Thrust, Hong Kong University of Science and Technology (Guangzhou) (HKUST), Guangzhou 511400, China} % 1 Du Xue Road,
-\affil[\USC]{Department of Chemical Engineering and Materials Science, University of Southern California (USC), Los Angeles CA 90089, USA} % 925 Bloom Walk
-\affil[\LBLMF]{Molecular Foundry Division, Lawrence Berkeley National Laboratory (LBNL), Berkeley 94720 CA, USA}  %67 Cyclotron Rd, Berkeley, CA 94720, United States
-\affil[\LBLALS]{Advanced Light Source, Lawrence Berkeley National Laboratory,  Berkeley 94720 CA, USA}  % 6 Cyclotron Rd, Berkeley, CA 94720, United States
-\affil[\Hoffmann]{Hoffmann Institute of Advanced Materials, Shenzhen Polytechnic, Shenzhen 518055, China} % Liuxian Street NO.7098, Shenzhen 518055, China
-\affil[\LBLCS]{Lawrence Berkeley National Laboratory (LBNL), Chemical Sciences Division, Berkeley 94720 CA, USA}  % 1 Cyclotron Rd, Berkeley, CA 94720, United States
-\affil[\EMPA]{Empa–Swiss Federal Laboratories for Materials Science and Technology (EMPA), 8600 Dübendorf, Switzerland}  % Ueberlandstrasse 129, 8600 Dübendorf, Switzerland
-\affil[\KITICRT]{Engler-Bunte-Institut \& Institute of Catalysis Research and Technology, Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany}  %Hermann-von-Helmholtz-Platz 1
-\affil[\CPT]{Chimie ParisTech, PSL University, CNRS, Institut de Recherche de Chimie Paris, 75005 Paris, France} % 11 Rue Pierre et Marie Curie, 75005 Paris, France
-\affil[\KFUPMMSE]{Material Science and Engineering Department, King Fahd University of Petroleum and Minerals (KFUPM), Dhahran 31261, Saudi Arabia} % 846R+8C2, Academic Belt Rd, KFUPM, Dhahran 34463, Saudi Arabia
-\affil[\KFUPMIRCIMR]{Interdisciplinary Research Center for Intelligent Manufacturing and Robotics, King Fahd University of Petroleum and Minerals (KFUPM), Dhahran 31261, Saudi Arabia} %Academic Belt Road, Dhahran 31261, Saudi Arabia
-\affil[\FZJIEMD]{Institute of Energy Materials and Devices, Forschungszentrum Juelich GmbH, 52425 Juelich, Germany} %Wilhelm-Johnen-Straße, 52428 Jülich
+\affil[]{}
+\author[ ]{Daniel Hollarek}
+\author[ ]{Henrik Schopmans}
+\author[ ]{Jona Östreicher}
+\author[ ]{Jonas Teufel}
+\author[ ]{Bin Cao}
+\author[ ]{Adie Alwen}
+\author[ ]{Simon Schweidler}
+\author[ ]{Mriganka Singh}
+\author[ ]{Tim Kodalle}
+\author[ ]{Hanlin Hu}
+\author[ ]{Gregoire Heymans}
+\author[ ]{Maged Abdelsamie}
+\author[ ]{Alexander Wieczorek}
+\author[ ]{Siarhei Zhuk}
+\author[ ]{Arthur Hardiagon}
+\author[ ]{Ruth Schwaiger}
+\author[ ]{François-Xavier Coudert}
+\author[ ]{Moritz Wolf}
+\author[ ]{Sebastian Siol}
+\author[ ]{Carolin M. Sutter-Fella}
+\author[ ]{Ben Breitung}
+\author[ ]{Andrea M. Hodge}
+\author[ ]{Tong-yi Zhang}
+\author[ ]{Pascal Friederich}
 
@@
 \maketitle
-\textbf{Keywords}: powder x-ray diffraction, machine learning, open-access data, data sharing, crystal structure determination, phase identification
+\renewcommand{\thesection}{S\arabic{section}}
 
-\begin{abstract}
-\input{chapters/00_abstract}
-\end{abstract}
+\section{Description of opXRD files on Zenodo}
+The database comes in two zip archives, ``opxrd.zip'' and ``opxrd\_in\_situ.zip''.
+The latter contains the in-situ data with highly correlated patterns recorded through time series measurements.
+Within the .zip archives patterns are saved as .json files grouped in folders indicating the contributing institution.
+If an institution contributed data from several projects, the contributed data is further divided into folders indicating the research project.
+These research project folders are labeled alphabetically in the order they are introduced in section $3$.
+Each .json file contains a pattern recorded from an X-ray diffraction experiment.
+If available, the composition and structure of the investigated sample and experiment conditions are also included in this file.
+Patterns belonging to time series measurements are labeled with filenames that indicate the measurement series they belong to as well as their order in that series.
 
-\newpage
-\section{Introduction}\label{sec:Introduction}
-\input{chapters/01_introduction}
-\input{chapters/02_SOTA_ML}
+\section{opXRD Python library usage}
 
-\section{Existing datasets}
-\input{chapters/03_existing_datasets}
-
-%\pagebreak
-\section{opXRD database}\label{sec:our_dataset}
-\input{chapters/04_our_dataset}
-
-\section{Usage}\label{sec:how_to_use}
-\input{chapters/05_how_to_use}
-
-\section{Summary and Outlook}\label{sec:summary_and_outlook}
-\input{chapters/06_summary_and_outlook}
-
-\subsection*{Supporting Information}
-Supporting Information is available from the Wiley Online Library or the author.
-
-\subsection*{Data availability}
-The opXRD database is available on Zenodo at \url{https://zenodo.org/records/14254270}.
-It is published under the Creative Commons Attribution 4.0 International license.It can be downloaded by any user without any barriers or restrictions.
-For further details, please refer to Section~$\eqref{sec:how_to_use}$.
-
-\subsection*{Conflicts of interest}
-There are no conflicts of interest to declare.
-
-\subsection*{Acknowledgements}
-We acknowledge financial support by the German Research Foundation (DFG) through the Research Training Group 2450 “Tailored Scale-Bridging Approaches to Computational Nanoscience”.
-We acknowledge support by the Federal Ministry of Education and Research (BMBF) under Grant No.
-01DM21001B (German-Canadian Materials Acceleration Center).
-We acknowledge financial support from the Helmholtz Foundation Model Initiative within Project "SOL-AI".
-Part of this work was funded under the France 2030 framework by Agence Nationale de la Recherche (project ANR-22-PEXD-0009 of PEPR DIADEM).
-Work at the Molecular Foundry was supported by the Office of Science, Office of Basic Energy Sciences, of the U.S.
-Department of Energy under Contract No.
-DE-AC02-05CH11231.
-Work at the Advanced Light Source (ALS) was done at beamline 12.3.2.
-The ALS is a DOE Office of Science User Facility under contract no.
-DE-AC02-05CH11231.
-The development of the online phase identification platform is supported by the Guangzhou-HKUST(GZ) Joint Funding Program (No.
-2023A03J0003).
-Work by the USC group was supported by the National Science Foundation (NSF) grant numbers DMR-2227178 and OISE-2106597.
-Funding by the Helmholtz Research Program “Materials and Technologies for the Energy Transition (MTET), Topic 3: Chemical Energy Carriers" is kindly acknowledged.
-Work by the Empa group was supported by the Strategic Focus Area–Advanced Manufacturing (SFA–AM) through the project Advancing manufacturability of hybrid organic–inorganic semiconductors for large area optoelectronics (AMYS) as well as the Empa internal research call 2020.
-
-\printnomenclature
-
-\clearpage
-
-\bibliographystyle{bibstyle/bibstyle.bst}
-\bibliography{references}
-
-\setcounter{section}{0}
-\renewcommand{\thesection}{S\arabic{section}}
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{S\arabic{figure}}
-\setcounter{table}{0}
-\renewcommand{\thetable}{S\arabic{table}}
+The opXRD Python library allows the dataset to be accessed through one simple command: \pyth{OpXRD.load(root_dirpath)}.
+If the database is locally available under \pyth{root_dirpath} this command loads the library from this location.
+If the database is not available locally at this location, the database is automatically downloaded to \pyth{root_dirpath}.
 
@@
 
-\renewcommand{\thesubsection}{S\arabic{subsection}}
-\section*{Supporting Information}
-
-\setcounter{page}{1}
-\renewcommand{\thesubsection}{S\arabic{subsection}}
-
-\subsection{Combined pattern plots}
-The Figure below $\eqref{fig:combined}$ shows 50 randomly selected samples of the X-ray diffraction patterns found in each of the research projects contributed to the opXRD database.
+\section{Combined pattern plots}
+Figure~$\eqref{fig:combined}$ belowshows 50 randomly selected samples of the X-ray diffraction patterns found in each of the research projects contributed to the opXRD database.
 
@@
 
-\pagebreak
-
-\section*{Description of opXRD files on Zenodo}
-The database comes in two zip archives, ``opxrd.zip'' and ``opxrd\_in\_situ.zip''.
-The latter contains the in-situ data with highly correlated patterns recorded through time series measurements.
-Within the .zip archives patterns are saved as .json files grouped in folders indicating the contributing institution.
-If an institution contributed data from several projects, the contributed data is further divided into folders indicating the research project.
-These research project folders are labeled alphabetically in the order they are introduced in section $\eqref{sec:our_dataset}$.
-Each .json file contains a pattern recorded from an X-ray diffraction experiment.
-If available, the composition and structure of the investigated sample and experiment conditions are also included in this file.
-Patterns belonging to time series measurements are labeled with filenames that indicate the measurement series they belong to as well as their order in that series.
-
-\section*{opXRD Python library usage}
-
-The opXRD Python library allows the dataset to be accessed through one simple command: \pyth{OpXRD.load(root_dirpath)}.
-If the database is locally available under \pyth{root_dirpath} this command loads the library from this location.
-If the database is not available locally at this location, the database is automatically downloaded to \pyth{root_dirpath}.
-
-\section*{Combined pattern plots}
-The Figure below $\eqref{fig:combined}$ shows 50 randomly selected samples of the X-ray diffraction patterns found in each of the research projects contributed to the opXRD database.
-
-\begin{figure}[!htb]
-    \centering
-    \includegraphics[width=\linewidth]{figures/combined.png}
-    \caption{50 randomly chosen X-ray diffraction patterns from each contributed dataset.
-The figure shows data from the following datasets: a) EMPA, b) LBNL-A, c) LBNL-B, d) LBNL-C, e) USC, f) INT, g) HKUST-A, h) HKUST-B, i) CNRS, j) IKFT.}
-    \label{fig:combined}
-\end{figure}
-
 \end{document}

### 2025-03-07 chapters/01_introduction.tex
Updated SI formatting
Merged final changes

@@
 The synthesis and characterization of novel materials are becoming increasingly efficient and automated, increasing the throughput of samples in experimentation pipelines\cite{MacLeod2019, Ludwig2019, Ozaki2020}.
-\\
 
@@
 As Rietveld refinement is a local optimization method, the result of the refinement procedure is generally only as good as the initial structural model the process started from.
-\\
 
@@
 Hence parameters are refined iteratively, with each iteration only refining a limited set of parameters.
-Finding the correct order in which to refine structure parameters and finding the correct values for initial background parameters both present problems that add to the difficulty of the refinement process.\\
+Finding the correct order in which to refine structure parameters and finding the correct values for initial background parameters both present problems that add to the difficulty of the refinement process.
 
@@
 Machine learning has the potential to speed up the manual analysis of powder diffractograms and keep pace with an automated high-throughput experimentation environment\cite{Agrawal2019, Surdu2023}.
-Models can be trained to predict crystal structure information directly given a diffractogram, or models can be used to automate the conventional refinement workflow.
+Models can be either trained to predict crystal structure information directly given a diffractogram, or they can be used to automate the conventional refinement workflow.
 In the latter case a model would first predict an initial crystal structure \cite{Surdu2023} which is then refined by a second model trained to perform the refinement process \cite{Feng2019}.
 So far due to an absence of labeled datasets with experimental diffractograms\cite{Wang2020}, machine learning in this domain has largely relied on diffractograms simulated from known structures\cite{Park2017, Lee2023} or, most recently, from generated synthetic crystals\cite{Schopmans2023}.
-Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting lattice parameters\cite{Dong2021, Chitturi2021, Habershon2004, zhang2024crystallographic}, spacegroup \cite{cao2024simxrd, Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021,zhang2024crystallographic}, and crystallite size \cite{Dong2021, Chakraborty2021} from simulated diffractograms.
+Models trained on datasets with simulated diffractograms have already shown strong performance in predicting lattice parameters\cite{Dong2021, Chitturi2021, Habershon2004, zhang2024crystallographic}, spacegroup \cite{cao2024simxrd, Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021,zhang2024crystallographic}, and crystallite size \cite{Dong2021, Chakraborty2021} from simulated diffractograms.
 However, the performance substantially drops off when these models are applied to data originating from experiments \cite{cao2024simxrd, Schopmans2023,zhang2024crystallographic, Wang2020, Vecsei2018}.
@@
 This is discussed in more detail below.
-\\
 
@@
 While some experimental powder databases exist, their utility is limited by the fact that they are either small or not openly accessible.
-\\
 
@@
 Larger commerical datasets such as the PDF5+\cite{GatesRector2019} and the Linus Pauling File\cite{villars2018} exist, but their utility is limited by fees and restrictive licenses.
-License terms of commercial datasets such as the PDF5+ and the Linus Pauling File prohibit or restrict the publication of models trained on their data.
+License terms of commercial datasets, such as the PDF5+ and the Linus Pauling File, prohibit or restrict the publication of models trained on their data.
 In contrast, the opXRD database is both free and imposes no restrictions on how its data is used.
 Fig.~$\eqref{fig:ml_uses}$ provides an overview of machine learning workflows enabled and supported by the opXRD database.
-\\
 
@@
 
-Of the \numpatterns patterns in the opXRD database \numlabeled pattern are at least partially labeled with structural data and more than 900 patterns possess full structural labels including atomic coordinates.
-This constitutes an experimental pXRD test dataset that is larger in size, richer in labels and broader in represented experimental diffraction setups than the RRUFF database which only provides lattice parameters as labels\cite{Armbruster2015}.
-However, since the majority of the opXRD database is unlabeled we also want to further discuss the uses of unlabeled data with regards to improving pattern simulations and transfer learning.
-\\
+Of the \numpatterns patterns in the opXRD database \numlabeled patterns come with at least partial structural information of the underlying sample.
+Of these \numlabeled labeled patterns, more than 900 have full structural labels including atomic coordinates.
+This constitutes an experimental pXRD test dataset that is larger in size, richer in labels, and broader in represented experimental setups than the RRUFF database, which only provides lattice parameters as labels\cite{Armbruster2015}.
+However, since the majority of the opXRD database is unlabeled we also want to further discuss the uses of unlabeled data, including its role in improving pattern simulations and its application in transfer learning approaches.
 
@@
 As these and more neglected effects are integrated into the simulation process, real powder diffraction data can be used to evaluate how closely simulated data matches up with real data.
-While direct comparisons are only possible on labeled patterns, comparing the strength and prevalence of features between simulated and real data can nevertheless provide clues about the fidelity of the situation.
-Taking on board all neglected effects without making approximations will incur significant computational costs that will lower the size of the generated training data.
+While direct comparisons are only possible on labeled patterns, comparing the strength and prevalence of features between simulated and real data can nevertheless provide information about the fidelity of the simulation.
+Taking into account all neglected effects without making approximations will incur significant computational costs that will lower the size of the generated training data.
 A more efficient approach could be to use real experimental data to identify the effects that have the largest impact in practice and model them heuristically.
-\\
 
@@
 These existing techniques can be adapted to facilitate transfer learning in the context of pXRD patterns.
-Additionally, Seddiki et al.
+Seddiki \textit{et al}.
 have already successfully applied transfer learning in the domain of mass spectrometry to boost the accuracy of mass spectrum classification models\cite{Seddiki2020}.
 Since both mass spectrometry data and pXRD data are one-dimensional, this work demonstrates the merit of transfer learning in a setting similar to pXRD.
-\\
 
 % Contributing + community driven
-The database is intended as a growing, community-driven initiative.
+The opXRD database is intended as a growing, community-driven initiative.
 The database we present here is the first version, but we hope to further increase the database size through active engagement with the pXRD community.
@@
 Further details can be found on the opXRD website (\url{https://xrd.aimat.science/}).
-An overview of this process is given in below Fig.~$\eqref{fig:submission}$.
-\\
+An overview of this process is given in Fig.~$\eqref{fig:submission}$ below.
 
@@
 It furthers the ability of other researchers to reproduce published work and in turn, adds to the credibility of the publisher of the data.
-Compared to publishing data individually, it also has the added benefit that the data is contributed towards a large, homogenous dataset with a standardized interface.
-This makes the data more easily accessible to other researchers and provides more value to researchers seeking large quantities of data compared to an isolated submission on Zenodo.
-However, further data annotation with metadata is required to fully fulfill the FAIR data principles.\\
+Compared to publishing data individually, publishing data on the opXRD database has the added benefit of contributing to a large, homogenous dataset with a standardized interface.
+This makes the data more easily accessible to other researchers and provides more value to researchers seeking large quantities of data.
+However, further data annotation with metadata is required to fully fulfill the FAIR data principles.
 
@@
 Additionally, some of the data was collected in grazing-angle geometry rather than in the usual Bragg-Brentano geometry employed in powder diffraction.
-The broad range of available experimental samples contained in the opXRD v1.0 database makes it possible to apply state-of-the-art ML approaches to the domain of powder XRD analysis.
+The broad range of available experimental samples contained in the opXRD v1.0 database makes it possible to apply state-of-the-art ML approaches to the domain of pXRD analysis.
 We hope that the opXRD database can drive ML research in this field towards more advanced automated analysis workflows that can accelerate materials science research through ready application in high-throughput experimentation pipelines.

### 2025-03-07 chapters/02_SOTA_ML.tex
Merged final changes

@@
-\subsection*{Review of machine learning-based pXRD analysis}
+\subsubsection*{Review of machine learning-based pXRD analysis}
 To showcase the need for datasets such as the one presented in this publication, we now discuss some recent approaches that apply machine learning methods to classification and regression tasks for powder diffractograms.
-\\
 
@@
 In 2017, Park {\it et al.} reached a test accuracy of roughly \SI{81}{\percent} for a CNN, which classifies space groups of simulated single-phase diffractograms \cite{Park2017}.
-\\
 
@@
 The study observed a noticeable performance drop when the pre-trained model was applied to real experimental patterns as opposed to simulated data.
-However, their recent work \cite{cao2024simxrd} proposes using convolutional peak descriptors that consider the detector's geometry, which reduces the performance gap in their benchmark tests.\\
+However, their recent work \cite{cao2024simxrd} proposes using convolutional peak descriptors that consider the detector's geometry, which reduces the performance gap in their benchmark tests.
 
@@
 Then they tested their model's performance on the other half of the RRUFF database and achieved a performance increase in the 230-way spacegroup classification accuracy of \num{11} percentage points compared to the same model only trained on simulated patterns.
-\\
 
@@
 Due to the lack of experimental pXRDs, they built a pipeline to augment simulated diffractograms of a reference structure by, among other things, slightly varying the underlying crystal lattice.
-For spinel structures, they reported an accuracy of \SI{100}{\percent} but they also proved that their approach can be transferred to other datasets.\\
+For spinel structures, they reported an accuracy of \SI{100}{\percent} but they also proved that their approach can be transferred to other datasets.
 
@@
 Augmenting our simulated patterns with background, noise, and impurities helps to bring simulated diffractograms closer to experimental ones, making models trained on them more performant on experimental diffractograms.
-However, this augmentation process could be improved by incorporating background and noise statistics from a broader experimental pXRD database, such as the one presented in this publication.\\
+However, this augmentation process could be improved by incorporating background and noise statistics from a broader experimental pXRD database, such as the one presented in this publication.
 
@@
 Therefore, transferring it to the application on experimental diffraction patterns is difficult.
-\cite{Schopmans2023, Lee2022, Vecsei2018} On the other hand, there are some successful approaches that also work well on experimental data, but those are mostly methods that do phase determination in a limited compound space, making the task less complex.
-\cite{Schuetzke2024, Lee2020}\\
+\cite{Schopmans2023, Lee2022, Vecsei2018} On the other hand, there are some successful approaches that also work well on experimental data, but those are mostly methods that do phase determination in a limited compound space, making the task less complex \cite{Schuetzke2024, Lee2020}.
 
 The current volume of experimental pXRD patterns is insufficient to effectively train ML models, highlighting an urgent need for a comprehensive experimental pXRD database.
-The most advanced ML models currently are trained on approximately $10^5 - 10^6$ simulated diffractograms.
-\cite{Salgado2023, Schopmans2023} This is, to the best of our knowledge, two orders of magnitude larger than the largest currently curated experimental dataset, the PDF-5+ with approximately $2\cdot 10^4$ experimental patterns.
+The most advanced ML models currently are trained on approximately $10^5 - 10^6$ simulated diffractograms \cite{Salgado2023, Schopmans2023}.
+This is, to the best of our knowledge, two orders of magnitude larger than the largest currently curated experimental dataset, the PDF-5+ with approximately $2\cdot 10^4$ experimental patterns.
 It is even one order of magnitude larger than the approximately $10^5$ unlabeled diffractograms in the initial version of the opXRD dataset we present here.
-\\
 
@@
 For both of these steps, the development of opXRD is particularly significant, as it will provide a comprehensive experimental benchmark for the community, allowing fair comparison of baseline models and accurate evaluation of their applicability in real experimental situations.
-
-\pagebreak

### 2025-03-07 chapters/03_existing_datasets.tex
Merged final changes

@@
 As of November 2024, Pearson's crystal data is available to researchers through a purchase of a one-year license starting at a price point of \qty{2200}{\myeuro}\cite{PearsonBuy}.
-The MPDS is partially open, with the open MPDS data accessible through a web interface\cite{MPDSWeb}.
+The MPDS is partially open, with the open portion of the MPDS data accessible through a web interface\cite{MPDSWeb}.
 API access to the full MPDS can be purchased through a one-year license starting at \qty{9500}{\myeuro}\cite{MPDSBuy}.
-When inquired about the availability of experimental powder diffraction data in the MPDS API, the Pauling File project stated that this data is not currently included.
-It could however be offered in the future at the request of customers.
-\\
+We asked the Pauling File project whether the experimental powder diffraction data is accessible through the MPDS API.
+The Pauling File project responded that this data is not currently provided through the API, but could be offered in the future at the request of customers.
 
@@
 According to the ICDD the latest release of the PDF, the PDF5+, contains over a million materials with accompanying powder diffraction data.
-However, since most of these powder diffraction patterns are simulated we inquired the ICDD about the number of experimental diffraction patterns in the PDF5+.
+However, since most of these powder diffraction patterns are simulated we asked the ICDD about the number of experimental diffraction patterns in the PDF5+.
 We were told that 20,800 of the patterns in the PDF5+ stemmed from experiments and that 10,954 of these patterns were accompanied by the atomic coordinates of the underlying structures.
@@
 Currently, the PDF5+ is available to researchers through a purchase of a one-year license starting at a price point of \$6265.
-However, the ICDD does not allow researchers to train machine learning models on PDF5+ data, regardless of whether the resulting models are published \cite{PDFLicenseWeb}.\\
+However, the ICDD does not allow researchers to train machine learning models on PDF5+ data, regardless of whether the resulting models are published \cite{PDFLicenseWeb}.
 
-\textbf{RRUFF}: \cite{RRUFFWeb} The RRUFF Mineral Database, first published in 2006, provides detailed information on minerals, including their chemical compositions, crystallography, and spectroscopic data.
-\cite{lafuente2015} Managed by the University of Arizona, it was created to serve as a public repository for mineral identification and research.
+\textbf{RRUFF}: \cite{RRUFFWeb} The RRUFF Mineral Database, first published in 2006, provides detailed information on minerals, including their chemical compositions, crystallography, and spectroscopic data \cite{lafuente2015}.
+Managed by the University of Arizona, it was created to serve as a public repository for mineral identification and research.
 It contains \num{1290} powder diffraction patterns stemming from experiments each labeled with the lattice parameters and composition of the underlying structures.
-The RRUFF data is openly accessible on the official website \cite{RRUFFWeb}.
-\\
+The RRUFF data is openly accessible on its official website \cite{RRUFFWeb}.
 
@@
 The data is openly accessible in the form of .cif files on the official COD website\cite{CODWeb}.
-\\
 
-\textbf{PowBase:} \cite{PowBaseWeb} PowBase was an initiative suggested in the Structure Determination by Powder Diffractometry (SDPD) mailing list which was hosted on the same site.
-The COD was another community initiative that grew out of this mailing list.
-PowBase is a database of 169 mostly unlabeled experimental powder diffraction patterns collected and maintained by crystallography researcher Armel Le Bail starting in 1999.
-As of November 2024, all 169 patterns are still freely available for download.
-\\
+\textbf{PowBase:} \cite{PowBaseWeb} PowBase is a database of 169 mostly unlabeled experimental powder diffraction patterns collected and maintained by crystallography researcher Armel Le Bail starting in 1999.
+PowBase is an initiative suggested in the Structure Determination by Powder Diffractometry (SDPD) mailing list which was co-maintained by Le Bail.
+The COD is another community initiative that grew out of this mailing list.
+As of March 2025, all 169 patterns are still freely available for download on the official website \cite{PowBaseWeb}.
 
@@
 This data will be included in a future release of the opXRD database.
-\\
 
 While not strictly speaking a powder diffraction database, the High-Throughput Experimental Materials Database (HTEM) by the National Renewable Energy Laboratory (NREL) is a valuable source of X-ray diffraction data \cite{zakutayev2018}.
-Currently, the HTEM database (https://htem.nrel.gov/stats) contains 65,779 thin-film samples with corresponding X-ray diffraction data.
+Currently, the HTEM database contains 65,779 thin-film samples with corresponding X-ray diffraction data\cite{HTEMWeb}.
 Each database entry includes the elemental composition of the underlying sample but does not provide any information on its structure.
-HTEM data is open-access and can be downloaded through an API provided by NREL.\\
+HTEM data is open-access and can be downloaded through an API provided by NREL.
 
@@
 Naval Research Laboratory Crystal Lattice-Structures \cite{NRLWeb}, the Athena Mineral database \cite{AthenaWeb} and the Protein data bank\cite{PDBWeb}.
-This is to be expected as most structure solutions are achieved through single-crystal diffraction rather than powder diffraction.
+The lack of experimental powder diffraction data in these databases is to be expected as most structure solutions are achieved through single-crystal diffraction.

### 2025-03-07 chapters/04_our_dataset.tex
Merged final changes

@@
 In collaboration with several other research institutions, we have collected a database of $\numpatterns$ experimental patterns of which 2179 are at least partially labeled with structural information of the underlying material.
-The following research institutions contributed data to the opXRD database: French National Centre for Scientific Research (CNRS), Hong Kong University of Science and Technology (Guangzhou) (HKUST), University of Southern California (USC), Lawrence Berkeley National Laboratory (LBNL), Empa–Swiss Federal Laboratories for Materials Science and Technology (EMPA) and the Karlsruhe Institute of Technology (KIT).
-Tab.~$\eqref{tab:merged}$ provides an overview of the contribution made by each institution opXRD database.
+The following research institutions contributed data to the opXRD database: The French National Centre for Scientific Research (CNRS), Hong Kong University of Science and Technology (Guangzhou) (HKUST), University of Southern California (USC), Lawrence Berkeley National Laboratory (LBNL), Empa–Swiss Federal Laboratories for Materials Science and Technology (EMPA) and the Karlsruhe Institute of Technology (KIT).
+Tab.~$\eqref{tab:merged}$ provides an overview of the contributions of each institution.
 We filtered the submitted datasets to exclude patterns with invalid features such as only one unique recorded angle, negative angles, less than 50 recorded angles total, or all intensities being zero.
-\\
 
@@
 \label{tab:merged}
-\scalebox{0.75}{
+\scalebox{0.735}{
 \begin{adjustbox}{center}
@@
 
-The variance of the data was analyzed using a principal component analysis (PCA).
+The variance of the data was analyzed using principal component analysis (PCA).
 PCA can be applied to datasets $X \subset \mathbb{R}^N$ to reduce the number of components needed to describe points $p \in X$ up to some tolerance in lost accuracy.
@@
 \end{align}
-Here $N_{\text{values}} = N =512$ is the number of recorded intensity values per pattern and $N_{\text{patterns}}$ is the number of patterns in the dataset $X$.
+Here $N_{\text{values}} =512$ is the number of recorded intensity values per pattern and $N_{\text{patterns}}$ is the number of patterns in the dataset $X$.
 Fig~$\eqref{fig:components}$ below shows the cumulative explained variance ratio over the fraction of maximal No.
@@
 The degree of variation between the patterns is different for each contribution.
-For example, the CRNS and the HKUST contributions each are collections that encompass many research projects over a large period of time and hence exhibit a high degree of variability between individual patterns.
+For example, the CRNS and the HKUST contributions each are collections that encompass many research projects over a large period of time and thus exhibit a high degree of variability between individual patterns.
 In contrast, the contributions by USC and LBNL contain many very similar patterns.
-The patterns in the USC datasets are similar because the underlying samples are all variations of \ce{CuNi} and \ce{CuAl} alloys collected on the same machine.
+The patterns in the USC datasets are similar because the underlying samples are all variations of \ce{CuNi} and \ce{CuAl} alloys.
 The patterns submitted by LBNL are similar because they stem from in-situ recordings where several hundred or several thousand patterns were collected over time per sample while they were undergoing physical conversion processes.
-\\
 
@@
 components refers to $N_{\text{max}}$ as defined in equation $\eqref{eq:nmax}$.
-Each dataset is named after the contributing institutions.
 Datasets contributed by the same institution are labeled alphabetically in the order in which they are described in the texts towards the end of this section.}
@@
 The start-to-end angle distribution reveals that all diffractograms start in a narrow window between $0\degree$ and approximately $50\degree$, while they end between $50\degree$ and $150\degree$, with the majority of patterns going from $0\degree$ to approximately $70\degree$.
-This means that, unlike in most ML approaches using synthetic data over the full angle range with fixed resolution, the opXRD dataset has a strongly varying angle range and resolution, which requires additional pre-processing methods such as padding and interpolation, or more flexible ML models beyond standard CNNs.
+Unlike most ML approaches using synthetic data over the full angle range with fixed resolution, the opXRD dataset has a strongly varying angle range and resolution.
+Hence, working with this data requires additional pre-processing methods such as padding and interpolation, or more flexible ML models beyond standard CNNs.
 
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{figures/hist.png}
+    \includegraphics[width=0.78\linewidth]{figures/hist.png}
     \caption{Histograms detailing the distribution of pattern and structure properties in the opXRD database: a) distribution of spacegroups present in labeled data; b) distribution of angular resolution in all data; c) distribution of smallest and largest recorded $2\theta$ values for all data.}
@@
 Each paragraph includes a description of the investigated materials and how X-ray diffraction data was collected.
-If applicable, the presence of thin-film samples or atypical diffraction geometrices is indicated.
+If applicable, the presence of thin-film samples or atypical diffraction geometries is indicated.
 Most data was collected using \ce{Cu} radiaton sources which has a $K_{\alpha1}$ wavelength of $\lambda=1.54056\text{\AA}$ and a $K_{\alpha2}$ wavelength of $\lambda=1.54439\text{\AA}$.
@@
 We could extract raw experimental pXRD patterns from 1052 files in total, after curation of a small number of files with clearly invalid data.
-\\
 
@@
 Data are categorized by elemental systems and include original experimental files, spanning single-phase to five-phase mixtures, as well as mixtures designed for various research tasks.
-\\
 
@@
 However, unlike XRed, these data entries are not accompanied by CIF files.
-\\
 
@@
 The newly identified and isolated semiconductor \ce{Zn2VN3} was identified to exhibit a cation-disordered wurtzite structure as verified by additional GI-XRD and SAED measurements\cite{Zhuk2021}.
-\\
+
 Tin halide perovskites were deposited using single-step spin-coating as reported elsewhere\cite{Wieczorek2023}.
@@
 For all combinatorial libraries where any phases are specified, the complete set of phases is reported in the metadata.
-\\
 
@@
 The Zn–V–N films showed some preferential out-of-plane orientation, while for the perovskites the preferential orientation was minimal, resulting in the presence of all reflections.
-\\
 
@@
 The sample set also includes samples that were not successfully measured or where phases could not be identified.
-\\
 
@@
 This dome led to increased background noise over the first $20\degree$ and slightly decreased pattern resolution.
-\\
 
@@
 A series of porous \ce{Al2O3} materials, which were prepared by calcination of boehmite (\ce{AlOOH}) at various temperatures, represents crystalline samples with limited long-range structure and various contributions of \ce{Al2O3} polymorphs.
-\\
 
@@
 This study has been carried out with the support of Angelina Barthelmeß, Elisabeth Herzinger, and Henning Hinrichs.
-\\
 
@@
 Depending on the substrate, each series of diffractograms shows an evolution from substrate only to a combination of polycrystalline perovskite, $\ce{PbI2}$ and substrate via several intermediate phases.
-\\
 
@@
 The two-dimensional data were calibrated using an Al$_{2}$O$_{3}$ calibration standard and integrated along the azimuthal angle.
-\\
 
@@
 The heating and cooling rates were fixed for all the experiments at 50$\degree$C/min and the holding time was fixed to 2 min.
-\\
 

### 2025-03-07 chapters/05_how_to_use.tex
Merged final changes

@@
-The opXRD database is hosted on Zenodo (\url{https://zenodo.org/records/14278656}) and can be downloaded by any user without any barriers or restrictions. \\
+The opXRD database is hosted on Zenodo (\url{https://zenodo.org/records/14254270}) and can be downloaded by any user without any barriers or restrictions.
 
@@
 The repository to this library is located at \url{https://github.com/aimat-lab/opxrd}.
-The opxrd library includes options for data-loading, for standardization, plotting, and the conversion to \emph{PyTorch} tensors.
+The opxrd library includes options for data-loading, standardization, plotting, and the conversion to \emph{PyTorch} tensors.
 We provide a Jupyter Notebook (\url{https://colab.research.google.com/github/aimat-lab/opXRD/blob/main/opxrd/usage.ipynb})

### 2025-03-07 chapters/06_summary_and_outlook.tex
Merged final changes

@@
-With the opXRD dataset, a curation of $\numpatterns$ unlabeled and 2179 at least partially labeled experimental powder XRD patterns from a wide range of different materials systems, we provide the largest currently available source of experimental XRD patterns.
-With this, we address the need for experimental data that arises when developing algorithms and analysis tools for XRD data, both based on machine learning and classical approaches.
+With the opXRD dataset, a curation of $\numpatterns$ unlabeled and 2179 at least partially labeled experimental powder X-ray diffraction patterns from a wide range of different materials systems, we provide the largest currently available source of experimental XRD patterns.
+With this, we address the need for experimental data that arises when developing algorithms and analysis tools for pXRD data, both based on machine learning and classical approaches.
 The data can be used for the actual method development and for testing.

### 2025-03-07 main.tex
Merged final changes

@@
 \usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
+\usepackage{titling}
+\usepackage{parskip}
 
@@
 \newcommand{\LBLCS}{8}
-\newcommand{\EMPA}{9}
-\newcommand{\KITICRT}{10}
+\newcommand{\KFUPMMSE}{9}
+\newcommand{\KFUPMIRCIMR}{10}
 \newcommand{\CPT}{11}
-\newcommand{\KFUPMMSE}{12}
-\newcommand{\KFUPMIRCIMR}{13}
-\newcommand{\FZJIEMD}{14}
+\newcommand{\EMPA}{12}
+\newcommand{\FZJIEMD}{13}
+\newcommand{\KITICRT}{14}
 
@@
 \author[\KFUPMMSE,\KFUPMIRCIMR]{Maged Abdelsamie}
+\author[\CPT]{Arthur Hardiagon}
 \author[\EMPA]{Alexander Wieczorek}
 \author[\EMPA]{Siarhei Zhuk}
-\author[\CPT]{Arthur Hardiagon}
 \author[\FZJIEMD]{Ruth Schwaiger}
+\author[\EMPA]{Sebastian Siol}
 \author[\CPT]{François-Xavier Coudert}
 \author[\KITICRT]{Moritz Wolf}
-\author[\EMPA]{Sebastian Siol}
 \author[\LBLMF]{Carolin M. Sutter-Fella}
@@
 \maketitle
-\textbf{Keywords}: powder x-ray diffraction, machine learning, open-access data, data sharing, crystal structure determination, phase identification
 
@@
 The opXRD database is available on Zenodo at \url{https://zenodo.org/records/14254270}.
-It is published under the Creative Commons Attribution 4.0 International license.It can be downloaded by any user without any barriers or restrictions.
+It is published under the Creative Commons Attribution 4.0 International license.
+It can be downloaded by any user without any barriers or restrictions.
 For further details, please refer to Section~$\eqref{sec:how_to_use}$.
@@
 
-\clearpage
-
 \bibliographystyle{bibstyle/bibstyle.bst}
@@
 
-\pagebreak
-
-\renewcommand{\thesubsection}{S\arabic{subsection}}
-\section*{Supporting Information}
-
-\setcounter{page}{1}
-\renewcommand{\thesubsection}{S\arabic{subsection}}
-
-\subsection{Combined pattern plots}
-The Figure below $\eqref{fig:combined}$ shows 50 randomly selected samples of the X-ray diffraction patterns found in each of the research projects contributed to the opXRD database.
-
-\begin{figure}[!htb]
-    \centering
-    \includegraphics[width=0.95\linewidth]{figures/combined.png}
-    \caption{50 randomly chosen X-ray diffraction patterns from each contributed dataset.
-The figure shows data from the following datasets: a) EMPA, b) LBNL-A, c) LBNL-B, d) LBNL-C, e) USC, f) INT, g) HKUST-A, h) HKUST-B, i) CNRS, j) IKFT.}
-    \label{fig:combined}
-\end{figure}
-
-\pagebreak
-
-\section*{Description of opXRD files on Zenodo}
-The database comes in two zip archives, ``opxrd.zip'' and ``opxrd\_in\_situ.zip''.
-The latter contains the in-situ data with highly correlated patterns recorded through time series measurements.
-Within the .zip archives patterns are saved as .json files grouped in folders indicating the contributing institution.
-If an institution contributed data from several projects, the contributed data is further divided into folders indicating the research project.
-These research project folders are labeled alphabetically in the order they are introduced in section $\eqref{sec:our_dataset}$.
-Each .json file contains a pattern recorded from an X-ray diffraction experiment.
-If available, the composition and structure of the investigated sample and experiment conditions are also included in this file.
-Patterns belonging to time series measurements are labeled with filenames that indicate the measurement series they belong to as well as their order in that series.
-
-\section*{opXRD Python library usage}
-
-The opXRD Python library allows the dataset to be accessed through one simple command: \pyth{OpXRD.load(root_dirpath)}.
-If the database is locally available under \pyth{root_dirpath} this command loads the library from this location.
-If the database is not available locally at this location, the database is automatically downloaded to \pyth{root_dirpath}.
-
-\section*{Combined pattern plots}
-The Figure below $\eqref{fig:combined}$ shows 50 randomly selected samples of the X-ray diffraction patterns found in each of the research projects contributed to the opXRD database.
-
-\begin{figure}[!htb]
-    \centering
-    \includegraphics[width=\linewidth]{figures/combined.png}
-    \caption{50 randomly chosen X-ray diffraction patterns from each contributed dataset.
-The figure shows data from the following datasets: a) EMPA, b) LBNL-A, c) LBNL-B, d) LBNL-C, e) USC, f) INT, g) HKUST-A, h) HKUST-B, i) CNRS, j) IKFT.}
-    \label{fig:combined}
-\end{figure}
-
 \end{document}

### 2025-04-24 SI.tex
Pre-review changes commit

@@
 
-\affil[]{}
 \author[ ]{Daniel Hollarek}
@@
 \author[ ]{Tong-yi Zhang}
-\author[ ]{Pascal Friederich}
+\author[*]{Pascal Friederich}
+
+\affil[*]{Corresponding author: pascal.friederich@kit.edu}
 
@@
 If an institution contributed data from several projects, the contributed data is further divided into folders indicating the research project.
-These research project folders are labeled alphabetically in the order they are introduced in section $3$.
+These research project folders are labeled alphabetically in the order they are introduced in Section~$3$.
 Each .json file contains a pattern recorded from an X-ray diffraction experiment.
 If available, the composition and structure of the investigated sample and experiment conditions are also included in this file.
-Patterns belonging to time series measurements are labeled with filenames that indicate the measurement series they belong to as well as their order in that series.
+Patterns belonging to time series measurements are labeled with filenames that indicate the measurement series they belong to and their order in that series.
 
@@
 \section{Combined pattern plots}
-Figure~$\eqref{fig:combined}$ belowshows 50 randomly selected samples of the X-ray diffraction patterns found in each of the research projects contributed to the opXRD database.
+Figure~$\eqref{fig:combined}$ shows 50 randomly selected samples of the X-ray diffraction patterns found in each of the research projects contributed to the opXRD database.
 

### 2025-04-24 chapters/01_introduction.tex
Pre-review changes commit

@@
 So far due to an absence of labeled datasets with experimental diffractograms\cite{Wang2020}, machine learning in this domain has largely relied on diffractograms simulated from known structures\cite{Park2017, Lee2023} or, most recently, from generated synthetic crystals\cite{Schopmans2023}.
-Models trained on datasets with simulated diffractograms have already shown strong performance in predicting lattice parameters\cite{Dong2021, Chitturi2021, Habershon2004, zhang2024crystallographic}, spacegroup \cite{cao2024simxrd, Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021,zhang2024crystallographic}, and crystallite size \cite{Dong2021, Chakraborty2021} from simulated diffractograms.
+Models trained on datasets with simulated diffractograms have already shown strong performance in predicting phases \cite{Park2017,chenAutomatingCrystalstructurePhase2021, changProbabilisticPhaseLabeling2023}, lattice parameters\cite{Dong2021, Chitturi2021, Habershon2004, zhang2024crystallographic}, spacegroup \cite{cao2024simxrd, Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021,zhang2024crystallographic}, and crystallite size \cite{Dong2021, Chakraborty2021} from simulated diffractograms.
 However, the performance substantially drops off when these models are applied to data originating from experiments \cite{cao2024simxrd, Schopmans2023,zhang2024crystallographic, Wang2020, Vecsei2018}.

### 2025-04-24 chapters/04_our_dataset.tex
Pre-review changes commit

@@
 In contrast, the contributions by USC and LBNL contain many very similar patterns.
-The patterns in the USC datasets are similar because the underlying samples are all variations of \ce{CuNi} and \ce{CuAl} alloys.
+The patterns in the USC dataset are similar because the underlying samples are all variations of \ce{CuNi} and \ce{CuAl} alloys.
 The patterns submitted by LBNL are similar because they stem from in-situ recordings where several hundred or several thousand patterns were collected over time per sample while they were undergoing physical conversion processes.
@@
 \begin{figure}[!htb]
-    \centering
-    \includegraphics[width=0.85\linewidth]{figures/component_fractions.png}
+    \centering\includegraphics[width=0.85\linewidth]{figures/component_fractions.png}
     \caption{Explained variance ratio over the fraction of the maximum number of components for each dataset contributed to the opXRD database.
@@
 
-Fig.~$\eqref{fig:histograms}$ provides an overview of the distributions of pattern and structure properties in the opXRD dataset.
+Fig.~$\eqref{fig:histograms}$ provides an overview of the distributions of pattern and structure properties in the opXRD database.
 Nearly all patterns have an angular resolution smaller than $\Delta(2\theta) = 0.1 ^\circ$.
@@
     \centering
-    \includegraphics[width=0.78\linewidth]{figures/hist.png}
+    \includegraphics[width=0.8\linewidth]{figures/hist.png}
     \caption{Histograms detailing the distribution of pattern and structure properties in the opXRD database: a) distribution of spacegroups present in labeled data; b) distribution of angular resolution in all data; c) distribution of smallest and largest recorded $2\theta$ values for all data.}

### 2025-04-24 chapters/06_summary_and_outlook.tex
Pre-review changes commit

@@
-With the opXRD dataset, a curation of $\numpatterns$ unlabeled and 2179 at least partially labeled experimental powder X-ray diffraction patterns from a wide range of different materials systems, we provide the largest currently available source of experimental XRD patterns.
+With the opXRD database, a curation of $\numpatterns$ unlabeled and 2179 at least partially labeled experimental powder X-ray diffraction patterns from a wide range of different materials systems, we provide the largest currently available source of experimental XRD patterns.
 With this, we address the need for experimental data that arises when developing algorithms and analysis tools for pXRD data, both based on machine learning and classical approaches.
@@
 % How to contribute
-Rather than a finished project, the opXRD dataset is an ongoing effort to collect experimental powder XRD data.
+Rather than a finished project, the opXRD database is an ongoing effort to collect experimental powder XRD data.
 We invite everyone who is working in the area of experimental powder XRD to submit it to the dataset, in order to further improve the utility of the dataset and thus aid further developments in this field.

### 2025-04-24 main.tex
Pre-review changes commit

@@
 \author[\HKUSTGZ]{Tong-yi Zhang}
-\author[\ITI,\INT]{Pascal Friederich$^\ast$}
+\author[\ITI,\INT,*]{Pascal Friederich}
 
@@
 \affil[\LBLCS]{Lawrence Berkeley National Laboratory (LBNL), Chemical Sciences Division, Berkeley 94720 CA, USA}  % 1 Cyclotron Rd, Berkeley, CA 94720, United States
-\affil[\EMPA]{Empa–Swiss Federal Laboratories for Materials Science and Technology (EMPA), 8600 Dübendorf, Switzerland}  % Ueberlandstrasse 129, 8600 Dübendorf, Switzerland
-\affil[\KITICRT]{Engler-Bunte-Institut \& Institute of Catalysis Research and Technology, Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany}  %Hermann-von-Helmholtz-Platz 1
-\affil[\CPT]{Chimie ParisTech, PSL University, CNRS, Institut de Recherche de Chimie Paris, 75005 Paris, France} % 11 Rue Pierre et Marie Curie, 75005 Paris, France
 \affil[\KFUPMMSE]{Material Science and Engineering Department, King Fahd University of Petroleum and Minerals (KFUPM), Dhahran 31261, Saudi Arabia} % 846R+8C2, Academic Belt Rd, KFUPM, Dhahran 34463, Saudi Arabia
 \affil[\KFUPMIRCIMR]{Interdisciplinary Research Center for Intelligent Manufacturing and Robotics, King Fahd University of Petroleum and Minerals (KFUPM), Dhahran 31261, Saudi Arabia} %Academic Belt Road, Dhahran 31261, Saudi Arabia
+\affil[\CPT]{Chimie ParisTech, PSL University, CNRS, Institut de Recherche de Chimie Paris, 75005 Paris, France} % 11 Rue Pierre et Marie Curie, 75005 Paris, France
+\affil[\EMPA]{Empa–Swiss Federal Laboratories for Materials Science and Technology (EMPA), 8600 Dübendorf, Switzerland}  % Ueberlandstrasse 129, 8600 Dübendorf, Switzerland
 \affil[\FZJIEMD]{Institute of Energy Materials and Devices, Forschungszentrum Juelich GmbH, 52425 Juelich, Germany} %Wilhelm-Johnen-Straße, 52428 Jülich
+\affil[\KITICRT]{Engler-Bunte-Institut \& Institute of Catalysis Research and Technology, Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany}  %Hermann-von-Helmholtz-Platz 1
+\affil[*]{Corresponding author: pascal.friederich@kit.edu}
 
 \begin{document}
-
 \maketitle
 
+\textbf{Keywords}: powder x-ray diffraction, machine learning, open-access data, crystal structure determination, phase identification
 \begin{abstract}
@@
 \subsection*{Acknowledgements}
-We acknowledge financial support by the German Research Foundation (DFG) through the Research Training Group 2450 “Tailored Scale-Bridging Approaches to Computational Nanoscience”.
-We acknowledge support by the Federal Ministry of Education and Research (BMBF) under Grant No.
+H.S.
+acknowledges financial support by the German Research Foundation (DFG) through the Research Training Group 2450 “Tailored Scale-Bridging Approaches to Computational Nanoscience”.
+P.F.
+and D.H.
+acknowledge support by the Federal Ministry of Education and Research (BMBF) under Grant No.
 01DM21001B (German-Canadian Materials Acceleration Center).
-We acknowledge financial support from the Helmholtz Foundation Model Initiative within Project "SOL-AI".
+J.Oe.
+and P.F.
+acknowledge financial support from the Helmholtz Foundation Model Initiative within Project "SOL-AI".
 Part of this work was funded under the France 2030 framework by Agence Nationale de la Recherche (project ANR-22-PEXD-0009 of PEPR DIADEM).
@@
 Work by the USC group was supported by the National Science Foundation (NSF) grant numbers DMR-2227178 and OISE-2106597.
-Funding by the Helmholtz Research Program “Materials and Technologies for the Energy Transition (MTET), Topic 3: Chemical Energy Carriers" is kindly acknowledged.
+M.W.
+acknowledges funding by the Helmholtz Research Program “Materials and Technologies for the Energy Transition (MTET), Topic 3: Chemical Energy Carriers".
 Work by the Empa group was supported by the Strategic Focus Area–Advanced Manufacturing (SFA–AM) through the project Advancing manufacturability of hybrid organic–inorganic semiconductors for large area optoelectronics (AMYS) as well as the Empa internal research call 2020.
+We thank BWCloud, funded by the Ministry of Science, Research and Arts Baden-Württemberg, for providing cloud server infrastructure.
+
+\clearpage
+
+\begin{figure}
+\centering
+\textbf{Table of Contents}\\
+\medskip
+  \includegraphics{figures/TOC.png}
+  \medskip
+  \caption*{We introduce the Open Experimental Powder X-ray Diffraction Database (opXRD), the largest openly accessible collection of experimental powder diffractograms, comprising over 92,000 patterns collected across diverse material classes and experimental setups.
+Our ongoing effort aims to guide machine learning research toward fully automated analysis of pXRD data and to enable future self-driving materials labs.}
+\end{figure}
 

### 2025-04-28 SI.tex
Added .json description

@@
 \usepackage{adjustbox}
+\usepackage{parskip}
 
@@
 \DeclareSIUnit\angstrom{\text {Å}}
-\usepackage[final]{changes}
+\usepackage{changes}
 
@@
 \section{Description of opXRD files on Zenodo}
-The database comes in two zip archives, ``opxrd.zip'' and ``opxrd\_in\_situ.zip''.
-The latter contains the in-situ data with highly correlated patterns recorded through time series measurements.
-Within the .zip archives patterns are saved as .json files grouped in folders indicating the contributing institution.
+\replaced{The opXRD database is packaged into a single zip file ``opxrd.zip'' available for download on Zenodo.}{The database comes in two zip archives, ``opxrd.zip'' and ``opxrd\_in\_situ.zip''.
+The latter contains the in-situ data with highly correlated patterns recorded through time series measurements.} Within the .zip archive patterns are saved as .json files grouped in folders indicating the contributing institution.
 If an institution contributed data from several projects, the contributed data is further divided into folders indicating the research project.
 These research project folders are labeled alphabetically in the order they are introduced in Section~$3$.
-Each .json file contains a pattern recorded from an X-ray diffraction experiment.
-If available, the composition and structure of the investigated sample and experiment conditions are also included in this file.
-Patterns belonging to time series measurements are labeled with filenames that indicate the measurement series they belong to and their order in that series.
+\added{Patterns belonging to time series measurements are labeled with filenames that indicate the measurement series they belong to and their order in that series.
+}
 
-\section{opXRD Python library usage}
+Each .json file contains a pattern recorded from an X-ray diffraction experiment\added{, as well as any available structural information of the underlying sample}.
+\added{At its topmost layer the .json file contains 4 mandatory fields: ``two\_theta\_values'', ``intensities'', ``powder\_experiment'' and ``metadata''.
+The first two together define the content of the diffraction pattern.
+The powder experiment field contains all available details about the experiment, including the wavelength of the employed X-ray source and everything known about the sample underlying the experiment.
+Insofar as these properties are available, stored sample properties include the chemical composition, lattice parameters, space group, atomic coordinates, and phase fraction of each phase.
+Finally, the metadata field contains information about the origin of the data including contributing institution, the name of the contributor as well as the original file format and measurement date if available.}
 
-The opXRD Python library allows the dataset to be accessed through one simple command: \pyth{OpXRD.load(root_dirpath)}.
+\added{For further details on the structure of the JSON file, we refer to the custom Python class ``XrdPattern'' in the \textsc{xrdpattern} library.
+Since the .json file serializes the ``XrdPattern'' class, the fields of the .json file are in direct correspondence with the attributes of the python class.}\deleted{If available, the composition and structure of the investigated sample and experiment conditions are also included in this file.
+Patterns belonging to time series measurements are labeled with filenames that indicate the measurement series they belong to and their order in that series.}
+
+\section{opXRD Python library \deleted{usage}}
+
+\added{Next to the dataset on Zenodo we also offer a dedicated Python library interface with the dataset directly through Python.} The opXRD Python library allows the dataset to be accessed through one simple command: \pyth{OpXRD.load(root_dirpath)}.
 If the database is locally available under \pyth{root_dirpath} this command loads the library from this location.
-If the database is not available locally at this location, the database is automatically downloaded to \pyth{root_dirpath}.
+If the database is not available locally at this location, the database is automatically downloaded from Zenodo to \pyth{root_dirpath}.
+\added{The opXRD library also contains the code used to generate the tables and figures of this work: \url{https://github.com/aimat-lab/opXRD/blob/main/opxrd/plots.ipynb}.}
 

### 2025-04-28 chapters/01_introduction.tex
Added .json description

@@
 In this work, we introduce an open powder X-ray diffraction (opXRD) database featuring a broad range of patterns collected from experiments.
-With a total of $92,552$ patterns collected from 6 contributing institutions, the opXRD database exceeds the size of the previously largest database of openly accessible experimental powder diffraction data by two orders of magnitude.
+\added{The objective of this work is to introduce and disseminate a large, open experimental pXRD dataset, paving the way for future studies that will evaluate and benchmark its impact on model performance.} With a total of $92,552$ patterns collected from 6 contributing institutions, the opXRD database exceeds the size of the previously largest database of openly accessible experimental powder diffraction data by two orders of magnitude.
 To the best of our knowledge, the largest database of this type is the RRUFF database, containing 1290 experimental powder diffraction patterns \cite{lafuente2015}.
@@
 
-Of the \numpatterns patterns in the opXRD database \numlabeled patterns come with at least partial structural information of the underlying sample.
+Of the \numpatterns patterns in the opXRD database\added{,} \numlabeled patterns come with at least partial structural information of the underlying sample.
 Of these \numlabeled labeled patterns, more than 900 have full structural labels including atomic coordinates.
-This constitutes an experimental pXRD test dataset that is larger in size, richer in labels, and broader in represented experimental setups than the RRUFF database, which only provides lattice parameters as labels\cite{Armbruster2015}.
-However, since the majority of the opXRD database is unlabeled we also want to further discuss the uses of unlabeled data, including its role in improving pattern simulations and its application in transfer learning approaches.
+\deleted{This constitutes an experimental pXRD test dataset that is larger in size, richer in labels, and broader in represented experimental setups than the RRUFF database, which only provides lattice parameters as labels\cite{Armbruster2015}.}
+\added{While the fraction of samples with structural labels (\SI{2.35}{\percent}) may appear small, this fraction represents the largest openly available collection of experimentally derived, structure-annotated pXRD patterns.
+As a comparison, the RRUFF dataset, often used for benchmarking ML models, contains partial labels for \num{1290} patterns, but no atomic coordinates.
+The pXRD dataset is larger in size, richer in labels, and broader in represented experimental setups compared to previous openly available datasets.
+Given the inherently labor-intensive nature of manual labeling in pXRD analysis, it is impractical to expect a fully labeled dataset at the scale of simulated training datasets (exceeding $10^5$ patterns).
+Therefore, next to the option of benchmarking models and methods on the labeled fraction, opXRD is designed to complement vast simulated datasets.
+This can be achieved through the adjustment of simulation parameters by comparing with the unlabeled fraction of the dataset, and through transfer learning strategies.
+We now want to discuss these two options of utilizing the unlabeled portion of our dataset in more detail.}
+\deleted{However, since the majority of the opXRD database is unlabeled we also want to further discuss the uses of unlabeled data, including its role in improving pattern simulations and its application in transfer learning approaches.}
 

### 2025-04-28 chapters/02_SOTA_ML.tex
Added .json description

@@
 For example, the space group classification task across all material systems is very general.
-Therefore, transferring it to the application on experimental diffraction patterns is difficult.
-\cite{Schopmans2023, Lee2022, Vecsei2018} On the other hand, there are some successful approaches that also work well on experimental data, but those are mostly methods that do phase determination in a limited compound space, making the task less complex \cite{Schuetzke2024, Lee2020}.
+Therefore, transferring it to the application on experimental diffraction patterns is difficult \cite{Schopmans2023, Lee2022, Vecsei2018}.
+On the other hand, there are some successful approaches that also work well on experimental data, but those are mostly methods that do phase determination in a limited compound space, making the task less complex \cite{Schuetzke2024, Lee2020}.
 

### 2025-04-28 chapters/05_how_to_use.tex
Added .json description

@@
 
-Next to the availability of the opXRD dataset on Zenodo, we also provide a Python library ``opxrd'' to easily download and interface with the dataset.
+\added{The opXRD library is designed for easy integration with common machine learning frameworks such as \emph{PyTorch}.
+This makes it an ideal resource for researchers developing and benchmarking sim-to-real transfer strategies in pXRD data analysis.}
+\replaced{We}{Next to the availability of the opXRD dataset on Zenodo, we also} provide a Python library ``opxrd'' to easily download and interface with the dataset.
 The instructions for how to install this library can be found in the repository associated with the library.

### 2025-04-28 chapters/06_summary_and_outlook.tex
Added .json description

@@
 Our dataset is a valuable and so far missing resource to drive further developments in the automated analysis of XRD data.
+\added{Looking forward, opXRD is expected to play an important role in the development of advanced transfer learning approaches that integrate large-scale simulated data with real experimental patterns, ultimately driving the automation and accuracy of pXRD analysis.}
+\added{Future work will also include comprehensive benchmark evaluations to quantify the performance improvements achieved by incorporating opXRD into transfer learning pipelines.}
 

### 2025-04-28 chapters/03_existing_datasets.tex
Improved table formatting

@@
 \caption{\footnotesize Overview of experimental powder diffraction databases: The column ``O.A.'' indicates whether or not the database is open-access.
-The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'', respectively.}
+The availability of the chemical composition, spacegroups, lattice parameters, and \replaced{basis}{atomic coordinates} of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'', respectively.}
 \label{tab:exp_databases}
-\scalebox{0.82}{
+\scalebox{0.85}{
 \begin{adjustbox}{center}
-\begin{NiceTabular}{@{}llllllll@{}}
+\begin{NiceTabular}{@{}llllllll@{}}[columns-width=1.2cm]
 \toprule
-\textbf{Name} & \textbf{No. patterns} & \textbf{O.A.} & \textbf{Comp.} & \textbf{Spg.} & \textbf{Lattice} & \textbf{Atom coords.} & \textbf{Year est.} \\
+\textbf{Name} & \textbf{No. patterns} & \textbf{O.A.} & \textbf{Comp.} & \textbf{Spg.} & \textbf{Lattice} & \textbf{Basis} & \textbf{Year est.} \\
 \midrule
@@
 Powder Diffraction File \tabularnote[a)]{The PDF lists the Material Platform for Data Science (MPDS) as a database source.
-Since the MPDS is hosted by the Pauling File project, there is likely significant overlap in the experimental patterns available in the PDF and the Linus Pauling File.} & 20,800 & \xmark & \cmark & \cmark & \cmark & \partialcheck{52} & 1941 \\
+Since the MPDS is hosted by the Pauling File project, there is likely significant overlap in the experimental patterns available in the PDF and the Linus Pauling File.} & 20,800 & \xmark & \cmark & \cmark & \cmark & \pchk{52} & 1941 \\
 RRUFF                                  & 1290                      & \cmark          & \cmark        & \cmark        & \cmark        & \xmark          & 2006            \\
-Crystallography Open Database          & 1052                      & \cmark          & \cmark        & \cmark        & \partialcheck{85}        & \partialcheck{85}          & 2003            \\
+Crystallography Open Database          & 1052                      & \cmark          & \cmark        & \cmark        & \pchk{85}     & \pchk{85}       & 2003            \\
 PowBase                                & 169                       & \cmark          & \cmark        & \xmark        & \xmark        & \xmark          & 1999            \\

### 2025-04-28 chapters/04_our_dataset.tex
Added .json description
Improved table formatting

@@
 We filtered the submitted datasets to exclude patterns with invalid features such as only one unique recorded angle, negative angles, less than 50 recorded angles total, or all intensities being zero.
+\added{We standardized the associated structural information according to the standards described by Setyawan {\it et al.}\cite{Setyawan2010} using \textsc{pymatgen}.
+In particular, this standardization enforces a single crystal axis convention throughout all opXRD data.
+}
 
@@
 \centering
-\caption{\footnotesize Overview of the contributions to the opXRD database: The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'' respectively.}
+\caption{\footnotesize Overview of the contributions to the opXRD database: The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``\replaced{Basis}{Atom coords}.'' respectively.}
 \label{tab:merged}
-\scalebox{0.735}{
+\scalebox{0.82}{
 \begin{adjustbox}{center}
-\begin{tabular}{@{}lllllll@{}}
+\begin{NiceTabular}{@{}llllllll@{}}[columns-width=1.2cm]
 %\begin{tabular}{@{}llccccl@{}}
 \toprule
-\textbf{Institution} & \textbf{No. patterns} & \ \textbf{Comp.} & \ \ \textbf{Spg.} & \textbf{Lattice} & \textbf{Atom coords.} & \textbf{Research Project} \\
+\textbf{Institution} & \textbf{No. patterns} & \textbf{Comp.} & \textbf{Spg.} & \textbf{Lattice} & \textbf{Basis} & \textbf{Research Project} \\
 \midrule
-CNRS        &  1052          & \ \cmark   & \partialcheck{85}   & \cmark   &  \ \ \quad \partialcheck{85}   & Diffraction data extracted from the COD \\ 
-USC         &  338          & \ \cmark   & \cmark   & \partialcheck{90}   & \ \ \quad \xmark   & Study of CuNi and CuAl alloys          \\
-HKUST(GZ)   &  520          & \ \partialcheck{4}   & \partialcheck{4}   & \partialcheck{4}   & \ \ \quad \partialcheck{4}   & Phase identification dataset  \\ 
-EMPA        &  770          &\ \cmark   & \partialcheck{63}   & \xmark   & \ \ \quad \xmark   & Metal halide perovskites, Zn-V-N libraries \\ 
-INT         &  19,796        &\ \xmark   & \xmark   & \xmark   & \ \ \quad \xmark   & Compilation of various projects        \\ 
-IKFT        &  64           &\ \xmark   & \xmark   & \xmark   & \ \ \quad \xmark   & Commercial catalysts, metals, metal oxides   \\ 
-LBNL        &  70,012          &\ \xmark   & \xmark   & \xmark   & \ \ \quad \xmark   & Perovskites precursors, Mn-Sb-O system  \\ 
+CNRS        &  1052         &  \cmark   & \pchk{85}  & \cmark    &   \pchk{85}   & Diffraction data extracted from the COD \\ 
+USC         &  338          &  \cmark   & \cmark     & \pchk{90} &   \xmark   & Study of CuNi and CuAl alloys          \\
+HKUST(GZ)   &  520          &  \pchk{4} & \pchk{4}   & \pchk{4}  &   \pchk{4}   & Phase identification dataset  \\ 
+EMPA        &  770          & \cmark    & \pchk{63}  & \xmark    &   \xmark   & Metal halide perovskites, Zn-V-N libraries \\ 
+INT         &  19,796       & \xmark    & \xmark     & \xmark    &   \xmark   & Compilation of various projects        \\ 
+IKFT        &  64           & \xmark    & \xmark     & \xmark    &   \xmark   & Commercial catalysts, metals, metal oxides   \\ 
+LBNL        &  70,012       & \xmark    & \xmark     & \xmark    &   \xmark   & Perovskites precursors, Mn-Sb-O system  \\ 
 \bottomrule
-\end{tabular}
+\end{NiceTabular}
 \end{adjustbox}}

### 2025-04-28 chapters/04_our_dataset.tex
Described new xrd/structure figures

@@
 \centering
-\caption{\footnotesize Overview of the contributions to the opXRD database: The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``\replaced{Basis}{Atom coords}.'' respectively.}
+\caption{\footnotesize Overview of the contributions to the opXRD database: The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``\replaced{Basis}{ Atom coords}.'' respectively.}
 \label{tab:merged}
@@
 \toprule
-\textbf{Institution} & \textbf{No. patterns} & \textbf{Comp.} & \textbf{Spg.} & \textbf{Lattice} & \textbf{Basis} & \textbf{Research Project} \\
+\textbf{\ Institution} & \textbf{No. patterns} & \textbf{Comp.} & \textbf{Spg.} & \textbf{Lattice} & \textbf{Basis} & \textbf{Research Project} \\
 \midrule
-CNRS        &  1052         &  \cmark   & \pchk{85}  & \cmark    &   \pchk{85}   & Diffraction data extracted from the COD \\ 
-USC         &  338          &  \cmark   & \cmark     & \pchk{90} &   \xmark   & Study of CuNi and CuAl alloys          \\
-HKUST(GZ)   &  520          &  \pchk{4} & \pchk{4}   & \pchk{4}  &   \pchk{4}   & Phase identification dataset  \\ 
-EMPA        &  770          & \cmark    & \pchk{63}  & \xmark    &   \xmark   & Metal halide perovskites, Zn-V-N libraries \\ 
-INT         &  19,796       & \xmark    & \xmark     & \xmark    &   \xmark   & Compilation of various projects        \\ 
-IKFT        &  64           & \xmark    & \xmark     & \xmark    &   \xmark   & Commercial catalysts, metals, metal oxides   \\ 
-LBNL        &  70,012       & \xmark    & \xmark     & \xmark    &   \xmark   & Perovskites precursors, Mn-Sb-O system  \\ 
+\ CNRS        &  1052         &  \cmark   & \pchk{85}  & \cmark    &   \pchk{85}   & Diffraction data extracted from the COD \\ 
+\ USC         &  338          &  \cmark   & \cmark     & \pchk{90} &   \xmark   & Study of CuNi and CuAl alloys          \\
+\ HKUST(GZ)   &  520          &  \pchk{4} & \pchk{4}   & \pchk{4}  &   \pchk{4}   & Phase identification dataset  \\ 
+\ EMPA        &  770          & \cmark    & \pchk{63}  & \xmark    &   \xmark   & Metal halide perovskites, Zn-V-N libraries \\ 
+\ INT         &  19,796       & \xmark    & \xmark     & \xmark    &   \xmark   & Compilation of various projects        \\ 
+\ IKFT        &  64           & \xmark    & \xmark     & \xmark    &   \xmark   & Commercial catalysts, metals, metal oxides   \\ 
+\ LBNL        &  70,012       & \xmark    & \xmark     & \xmark    &   \xmark   & Perovskites precursors, Mn-Sb-O system  \\ 
+\midrule
+\midrule
+\added{\ $\sum$ Labeled}   & 2,179  & \pchk{79}  & \pchk{66}  & \pchk{63}  & \pchk{42} & Partially and fully labeled opXRD data \\
+\added{\ $\sum$ Unlabeled }& 90,373 & \xmark     & \xmark     & \xmark     & \xmark    & Unlabeled opXRD data\\
 \bottomrule
@@
 components $N_{\text{max}}$ as defined above.
-In this figure, a faster convergence of the cumulative variance ratio towards one indicates that the patterns in this data are relatively similar.
-The degree of variation between the patterns is different for each contribution.
-For example, the CRNS and the HKUST contributions each are collections that encompass many research projects over a large period of time and thus exhibit a high degree of variability between individual patterns.
-In contrast, the contributions by USC and LBNL contain many very similar patterns.
+In this figure, a faster convergence of the cumulative variance ratio towards one indicates that the patterns in this dataset are \replaced{largely}{relatively} similar.
+\deleted{The degree of variation between the patterns is different for each contribution.} \replaced{For example}{In contrast}, the contributions by USC and LBNL contain many very similar patterns.
 The patterns in the USC dataset are similar because the underlying samples are all variations of \ce{CuNi} and \ce{CuAl} alloys.
-The patterns submitted by LBNL are similar because they stem from in-situ recordings where several hundred or several thousand patterns were collected over time per sample while they were undergoing physical conversion processes.
+The patterns submitted by LBNL are similar because they stem from in-situ recordings where several hundred or several thousand patterns were collected over time per sample.
+\deleted{ while they were undergoing physical conversion processes.} \replaced{In contrast}{For example}, the CRNS and the HKUST contributions each are collections that encompass many research projects over a large period of time and thus exhibit a high degree of variability between individual patterns.
 
@@
 
-Fig.~$\eqref{fig:histograms}$ provides an overview of the distributions of pattern and structure properties in the opXRD database.
+Fig.~$\eqref{fig:hist_XRD}$ \replaced{illustrates}{provides an overview of} the distributions of pattern \deleted{and structure} properties in the opXRD database.
 Nearly all patterns have an angular resolution smaller than $\Delta(2\theta) = 0.1 ^\circ$.
@@
 
+\added{Fig.~$\eqref{fig:hist_CRYSTAL}$ illustrates how structural properties are distributed among the structures underlying the labeled subset of the opXRD database.
+Most structures include either \ce{N}, \ce{C}, or \ce{O} atoms, and have unit cells that contain less than 100 atoms and are smaller than $10^3$\AA.
+The most common spacegroups include the orthorhombic $\text{Pnma}$, the monoclinic $\text{P}2_1/c$, and the cubic $\text{Fm}3\text{m}$.}
+
 \begin{figure}[!htb]
     \centering
-    \includegraphics[width=0.8\linewidth]{figures/hist.png}
-    \caption{Histograms detailing the distribution of pattern and structure properties in the opXRD database: a) distribution of spacegroups present in labeled data; b) distribution of angular resolution in all data; c) distribution of smallest and largest recorded $2\theta$ values for all data.}
-    \label{fig:histograms}
+    \includegraphics[width=\linewidth]{figures/hist_XRD.png}
+    \caption{Histograms detailing properties of all diffraction patterns in the opXRD database: a) distribution of angular resolutions; b) distribution of smallest and largest recorded $2\theta$ values.}
+    \label{fig:hist_XRD}
 \end{figure}
+
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=\linewidth]{figures/hist_CRYSTAL.png}
+    \caption{Histograms detailing properties of the structures underlying labeled diffraction patterns in the opXRD database: a) distribution of structures containing specified elements;b) distribution of structures containing spacegroups; c) distribution of No.
+atoms $N_{\text{atom}}$ contained in unit cell; d) distribution of unit cell volume $V_{\text{uc}}$}
+    \label{fig:hist_CRYSTAL}
+\end{figure}
+
+\pagebreak
 

### 2025-04-29 SI.tex
Finalized changes @ 14:41 29.04.25

@@
 \section{Description of opXRD files on Zenodo}
-\replaced{The opXRD database is packaged into a single zip file ``opxrd.zip'' available for download on Zenodo.}{The database comes in two zip archives, ``opxrd.zip'' and ``opxrd\_in\_situ.zip''.
+\replaced{The opXRD database is packaged as a single zip file ``opxrd.zip'' available for download on Zenodo.}{The database comes in two zip archives, ``opxrd.zip'' and ``opxrd\_in\_situ.zip''.
 The latter contains the in-situ data with highly correlated patterns recorded through time series measurements.} Within the .zip archive patterns are saved as .json files grouped in folders indicating the contributing institution.
@@
 Each .json file contains a pattern recorded from an X-ray diffraction experiment\added{, as well as any available structural information of the underlying sample}.
-\added{At its topmost layer the .json file contains 4 mandatory fields: ``two\_theta\_values'', ``intensities'', ``powder\_experiment'' and ``metadata''.
-The first two together define the content of the diffraction pattern.
-The powder experiment field contains all available details about the experiment, including the wavelength of the employed X-ray source and everything known about the sample underlying the experiment.
-Insofar as these properties are available, stored sample properties include the chemical composition, lattice parameters, space group, atomic coordinates, and phase fraction of each phase.
-Finally, the metadata field contains information about the origin of the data including contributing institution, the name of the contributor as well as the original file format and measurement date if available.}
+\added{At its topmost layer, the .json file contains 4 mandatory fields: ``two\_theta\_values'', ``intensities'', ``powder\_experiment'', and ``metadata''.
+The first two define the content of the diffraction pattern.
+The ``powder\_experiment'' field contains all available details about the experiment, including the wavelength of the employed X-ray source and everything known about the sample underlying the experiment.
+Insofar as these properties are available, stored sample properties include the chemical composition, lattice parameters, space group, atomic coordinates, and the phase fraction of each phase of the material.
+Finally, the metadata field contains information about the origin of the data, including the contributing institution, the name of the contributor as well as the original file format and measurement date if available.
+The metadata field will in the future also display the version of the opXRD database in which the pattern was added.}
 
@@
 
-\section{opXRD Python library \deleted{usage}}
+\section{Figure source code\deleted{opXRD Python library usage}}
 
-\added{Next to the dataset on Zenodo we also offer a dedicated Python library interface with the dataset directly through Python.} The opXRD Python library allows the dataset to be accessed through one simple command: \pyth{OpXRD.load(root_dirpath)}.
+\deleted{Next to the dataset on Zenodo we also offer a dedicated Python library interface with the dataset directly through Python.
+The opXRD Python library allows the dataset to be accessed through one simple command: \pyth{OpXRD.load(root_dirpath)}.
 If the database is locally available under \pyth{root_dirpath} this command loads the library from this location.
-If the database is not available locally at this location, the database is automatically downloaded from Zenodo to \pyth{root_dirpath}.
-\added{The opXRD library also contains the code used to generate the tables and figures of this work: \url{https://github.com/aimat-lab/opXRD/blob/main/opxrd/plots.ipynb}.}
+If the database is not available locally at this location, the database is automatically downloaded from Zenodo to \pyth{root_dirpath}.}
+\added{The opXRD library also contains the source code that was used to generate the tables and figures seen in Section$~3$ and the appendix.
+All figures, including the associated code to generate them, are located in the following notebook: \url{https://github.com/aimat-lab/opXRD/blob/9a6050ea8b62c9d67f72a126a33bf18f70cebefb/opxrd/plots.ipynb}.}
 

### 2025-04-29 chapters/01_introduction.tex
Finalized changes @ 14:41 29.04.25

@@
 
-In this work, we introduce an open powder X-ray diffraction (opXRD) database featuring a broad range of patterns collected from experiments.
-\added{The objective of this work is to introduce and disseminate a large, open experimental pXRD dataset, paving the way for future studies that will evaluate and benchmark its impact on model performance.} With a total of $92,552$ patterns collected from 6 contributing institutions, the opXRD database exceeds the size of the previously largest database of openly accessible experimental powder diffraction data by two orders of magnitude.
+In this work, we introduce an open powder X-ray diffraction (opXRD) database featuring a broad range of patterns collected from experiments.\added{The objective of this work is to introduce and disseminate a large, open experimental pXRD dataset, paving the way for future studies that will evaluate and benchmark its impact on model performance.} With a total of $\numpatterns$ patterns collected from 6 contributing institutions, the opXRD database exceeds the size of the previously largest database of openly accessible experimental powder diffraction data by two orders of magnitude.
 To the best of our knowledge, the largest database of this type is the RRUFF database, containing 1290 experimental powder diffraction patterns \cite{lafuente2015}.
@@
 
-Of the \numpatterns patterns in the opXRD database\added{,} \numlabeled patterns come with at least partial structural information of the underlying sample.
-Of these \numlabeled labeled patterns, more than 900 have full structural labels including atomic coordinates.
+Of the $\numpatterns$ patterns in the opXRD database\added{,} \numlabeled patterns come with at least partial structural information of the underlying sample.
+Of these \numlabeled labeled patterns, \numfull have full structural labels including lattice and basis.
 \deleted{This constitutes an experimental pXRD test dataset that is larger in size, richer in labels, and broader in represented experimental setups than the RRUFF database, which only provides lattice parameters as labels\cite{Armbruster2015}.}
 \added{While the fraction of samples with structural labels (\SI{2.35}{\percent}) may appear small, this fraction represents the largest openly available collection of experimentally derived, structure-annotated pXRD patterns.
-As a comparison, the RRUFF dataset, often used for benchmarking ML models, contains partial labels for \num{1290} patterns, but no atomic coordinates.
+As a comparison, the RRUFF database, often used for benchmarking ML models, contains partial labels for \num{1290} patterns, but no atomic coordinates.
 The pXRD dataset is larger in size, richer in labels, and broader in represented experimental setups compared to previous openly available datasets.
-Given the inherently labor-intensive nature of manual labeling in pXRD analysis, it is impractical to expect a fully labeled dataset at the scale of simulated training datasets (exceeding $10^5$ patterns).
-Therefore, next to the option of benchmarking models and methods on the labeled fraction, opXRD is designed to complement vast simulated datasets.
+Given the inherently labor-intensive nature of manual labeling in pXRD analysis, it is impractical to expect a fully labeled dataset at the scale of simulated training datasets, which commonly exceed $10^6$ patterns \cite{Salgado2023, Schopmans2023}.
+Therefore, next to the option of benchmarking models and methods on the labeled subset, opXRD is designed to complement vast simulated datasets.
 This can be achieved through the adjustment of simulation parameters by comparing with the unlabeled fraction of the dataset, and through transfer learning strategies.

### 2025-04-29 chapters/04_our_dataset.tex
Finalized changes @ 14:41 29.04.25

@@
-In collaboration with several other research institutions, we have collected a database of $\numpatterns$ experimental patterns of which 2179 are at least partially labeled with structural information of the underlying material.
+In collaboration with several other research institutions, we have collected a database of $\numpatterns$ experimental \added{powder diffraction} patterns.
+\added{Of these patterns, $\numlabeled$} are at least partially labeled with structural information \added{and $\numfull$ are fully labeled through the lattice and basis of the underlying material}.
 The following research institutions contributed data to the opXRD database: The French National Centre for Scientific Research (CNRS), Hong Kong University of Science and Technology (Guangzhou) (HKUST), University of Southern California (USC), Lawrence Berkeley National Laboratory (LBNL), Empa–Swiss Federal Laboratories for Materials Science and Technology (EMPA) and the Karlsruhe Institute of Technology (KIT).
-Tab.~$\eqref{tab:merged}$ provides an overview of the contributions of each institution.
-We filtered the submitted datasets to exclude patterns with invalid features such as only one unique recorded angle, negative angles, less than 50 recorded angles total, or all intensities being zero.
+\added{We have taken measures to ensure data validation through both manual and automated processes.
+Before parsing the data we established the file formats and data organization of each submission to ensure that the files were compatible with our custom parsing mechanism.
+As part of the automatic parsing process,} we filtered the submitted datasets to exclude patterns with invalid features such as only one unique recorded angle, negative angles, less than 50 recorded angles total, or all intensities being zero.
+\added{After the parsing, we also manually inspected a random selection of patterns from each submitted dataset for any anomalies that would warrant further investigation.}
+
 \added{We standardized the associated structural information according to the standards described by Setyawan {\it et al.}\cite{Setyawan2010} using \textsc{pymatgen}.
 In particular, this standardization enforces a single crystal axis convention throughout all opXRD data.
-}
+} Tab.~$\eqref{tab:merged}$ provides an overview of the contributions of each institution.
 
@@
 \label{tab:merged}
-\scalebox{0.82}{
+\scalebox{0.81}{
 \begin{adjustbox}{center}
 \begin{NiceTabular}{@{}llllllll@{}}[columns-width=1.2cm]
-%\begin{tabular}{@{}llccccl@{}}
 \toprule
@@
 \midrule
-\added{\ $\sum$ Labeled}   & 2,179  & \pchk{79}  & \pchk{66}  & \pchk{63}  & \pchk{42} & Partially and fully labeled opXRD data \\
-\added{\ $\sum$ Unlabeled }& 90,373 & \xmark     & \xmark     & \xmark     & \xmark    & Unlabeled opXRD data\\
+\added{\ $\sum$ Labeled}   & \added{2,179}  & \pchk{79}  & \pchk{66}  & \pchk{63}  & \pchk{42} & \added{Partially and fully labeled opXRD data} \\
+\added{\ $\sum$ Unlabeled }& \added{90,373} & \xmark     & \xmark     & \xmark     & \xmark    & \added{Unlabeled opXRD data}\\
 \bottomrule
@@
     \includegraphics[width=\linewidth]{figures/hist_XRD.png}
-    \caption{Histograms detailing properties of all diffraction patterns in the opXRD database: a) distribution of angular resolutions; b) distribution of smallest and largest recorded $2\theta$ values.}
+    \caption{\added{Histograms detailing properties of all diffraction patterns in the opXRD database: a) distribution of angular resolutions, b) distribution of smallest and largest recorded $2\theta$ values.}}
     \label{fig:hist_XRD}
@@
     \includegraphics[width=\linewidth]{figures/hist_CRYSTAL.png}
-    \caption{Histograms detailing properties of the structures underlying labeled diffraction patterns in the opXRD database: a) distribution of structures containing specified elements;b) distribution of structures containing spacegroups; c) distribution of No.
-atoms $N_{\text{atom}}$ contained in unit cell; d) distribution of unit cell volume $V_{\text{uc}}$}
+    \caption{\added{Histograms detailing properties of the structures underlying labeled diffraction patterns in the opXRD database: a) distribution of structures containing specified elements, b) distribution of structures containing spacegroups, c) distribution of No.
+atoms $N_{\text{atom}}$ contained in unit cell, d) distribution of unit cell volume $V_{\text{uc}}$.}}
     \label{fig:hist_CRYSTAL}

### 2025-04-29 chapters/05_how_to_use.tex
Finalized changes @ 14:41 29.04.25

--- 2025-03-07 chapters/05_how_to_use.tex
@@
 
-Next to the availability of the opXRD dataset on Zenodo, we also provide a Python library ``opxrd'' to easily download and interface with the dataset.
+\replaced{We}{Next to the availability of the opXRD dataset on Zenodo, we} also provide a Python library ``opxrd'' to easily download and interface with the dataset.
+\added{The opXRD library is designed for easy integration with common machine learning frameworks such as \emph{PyTorch}.
+This makes it an ideal resource for researchers developing and benchmarking sim-to-real transfer strategies in pXRD data analysis.}
 The instructions for how to install this library can be found in the repository associated with the library.

### 2025-04-29 chapters/06_summary_and_outlook.tex
Finalized changes @ 14:41 29.04.25

--- 2025-04-24 chapters/06_summary_and_outlook.tex
@@
 Our dataset is a valuable and so far missing resource to drive further developments in the automated analysis of XRD data.
+\added{Looking forward, opXRD is expected to play an important role in the development of advanced transfer learning approaches that integrate large-scale simulated data with real experimental patterns, ultimately driving the automation and accuracy of pXRD analysis.
+Future work will also include comprehensive benchmark evaluations to quantify the performance improvements achieved by incorporating opXRD into transfer learning pipelines.}
 
@@
 Rather than a finished project, the opXRD database is an ongoing effort to collect experimental powder XRD data.
-We invite everyone who is working in the area of experimental powder XRD to submit it to the dataset, in order to further improve the utility of the dataset and thus aid further developments in this field.
-Our submission page (\url{https://xrd.aimat.science/}) and submission helper software will be kept available to collect more data.
+We invite everyone \deleted{who is} working in the area of experimental powder XRD to submit it to the dataset, \deleted{in order} to further improve the utility of the dataset and thus aid further developments in this field.
+Our submission page (\url{https://xrd.aimat.science/}) \deleted{and submission helper software} will \replaced{continue to stay}{ be kept available} \added{available for the submission of data.
+As new submissions come in, newer versions of the opXRD database incorporating these submitted datasets will be released.}
+\added{As the opXRD database grows further, we look forward to expanding this website to become a comprehensive community resource from which the database can be governed.
+Planned resources on this site include a contributor list, version list, and changelogs as well as comprehensive versioning, license, citation, and attribution practice statements.
+}
+
 We will keep updating and maintaining the dataset with new incoming submissions.

### 2025-04-29 chapters/03_existing_datasets.tex
Update

@@
 \caption{\footnotesize Overview of experimental powder diffraction databases: The column ``O.A.'' indicates whether or not the database is open-access.
-The availability of the chemical composition, spacegroups, lattice parameters, and \replaced{basis}{atomic coordinates} of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'', respectively.}
+The availability of the chemical composition, spacegroups, lattice parameters, and \replaced{the full structure}{atomic coordinates} of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Full structure'', respectively.}
 \label{tab:exp_databases}
-\scalebox{0.85}{
+\scalebox{0.82}{
 \begin{adjustbox}{center}
@@
 \toprule
-\textbf{Name} & \textbf{No. patterns} & \textbf{O.A.} & \textbf{Comp.} & \textbf{Spg.} & \textbf{Lattice} & \textbf{Basis} & \textbf{Year est.} \\
+\textbf{Name} & \textbf{No. Patterns} & \textbf{O.A.} & \textbf{Comp.} & \textbf{Spg.} & \textbf{Lattice} & \textbf{Full} & \textbf{Year est.} \\
+& & & & & & \textbf{Structure} & \\
 \midrule

### 2025-04-29 chapters/04_our_dataset.tex
Update

@@
 In collaboration with several other research institutions, we have collected a database of $\numpatterns$ experimental \added{powder diffraction} patterns.
-\added{Of these patterns, $\numlabeled$} are at least partially labeled with structural information \added{and $\numfull$ are fully labeled through the lattice and basis of the underlying material}.
+\added{Of these patterns, $\numlabeled$} are at least partially labeled with structural information \added{and $\numfull$ are labeled with the full crystal structure of the underlying material}.
 The following research institutions contributed data to the opXRD database: The French National Centre for Scientific Research (CNRS), Hong Kong University of Science and Technology (Guangzhou) (HKUST), University of Southern California (USC), Lawrence Berkeley National Laboratory (LBNL), Empa–Swiss Federal Laboratories for Materials Science and Technology (EMPA) and the Karlsruhe Institute of Technology (KIT).
@@
 \centering
-\caption{\footnotesize Overview of the contributions to the opXRD database: The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``\replaced{Basis}{ Atom coords}.'' respectively.}
+\caption{\footnotesize Overview of the contributions to the opXRD database: The availability of the chemical composition, spacegroups, lattice parameters, and \replaced{the full structure}{atomic coordinates} of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``\replaced{Full structure}{ Atom coords}.'' respectively.}
 \label{tab:merged}
@@
 \toprule
-\textbf{\ Institution} & \textbf{No. patterns} & \textbf{Comp.} & \textbf{Spg.} & \textbf{Lattice} & \textbf{Basis} & \textbf{Research Project} \\
+\textbf{\ Institution} & \textbf{No. Patterns} & \textbf{Comp.} & \textbf{Spg.} & \textbf{Lattice} & \textbf{Full} & \textbf{Research Project} \\
+& & & & & \textbf{Structure} & \\
 \midrule

### 2025-04-29 main.tex
Added .json description
Improved table formatting
Finalized changes @ 14:41 29.04.25
Update

@@
 \usepackage{pifont}% http://ctan.org/pkg/pifont
-\newcommand{\cmark}[1][]{\textcolor{green!80!black}{#1\quad\ding{52}}}
-\newcommand{\xmark}[1][]{\textcolor{red}{#1\quad\ding{53}}}
-\newcommand{\partialcheck}[1]{\textcolor{orange}{\quad\ding{52}(#1\%)}}
+\newcommand{\cmark}[1][]{\textcolor{green!80!black}{#1\ding{52}}}
+\newcommand{\xmark}[1][]{\textcolor{red}{#1\ding{53}}}
+\newcommand{\pchk}[1]{\textcolor{orange}{#1\%}}
 \captionsetup[table]{font=footnotesize}
@@
 \newcommand{\numlabeled}{2179 }
+\newcommand{\numfull}{912 }
 
@@
 \subsection*{Data availability}
-The opXRD database is available on Zenodo at \url{https://zenodo.org/records/14254270}.
-It is published under the Creative Commons Attribution 4.0 International license.
+\replaced{The opXRD database is available on Zenodo at \url{https://zenodo.org/records/15298026}.}{The opXRD database is available on Zenodo at \url{https://zenodo.org/records/14254270}.}\added{The database version this manuscript refers to is v11 on Zenodo.} \replaced{opXRD}{It} is published under the Creative Commons Attribution 4.0 International license.
 It can be downloaded by any user without any barriers or restrictions.

### 2025-04-29 chapters/04_our_dataset.tex
Switch to Wiley Latex Template

@@
 \label{tab:merged}
-\scalebox{0.81}{
+\scalebox{0.9}{
 \begin{adjustbox}{center}
@@
 \toprule
-\textbf{\ Institution} & \textbf{No. Patterns} & \textbf{Comp.} & \textbf{Spg.} & \textbf{Lattice} & \textbf{Full} & \textbf{Research Project} \\
-& & & & & \textbf{Structure} & \\
+\textbf{\ Institution} & \textbf{No. Patterns} & \textbf{Comp.} & \textbf{Spg.} & \textbf{Lattice} & \textbf{Full Structure} & \textbf{Research Project} \\
 \midrule
@@
     \includegraphics[width=\linewidth]{figures/hist_CRYSTAL.png}
-    \caption{\added{Histograms detailing properties of the structures underlying labeled diffraction patterns in the opXRD database: a) distribution of structures containing specified elements, b) distribution of structures containing spacegroups, c) distribution of No.
-atoms $N_{\text{atom}}$ contained in unit cell, d) distribution of unit cell volume $V_{\text{uc}}$.}}
+    \caption{\added{Histograms detailing properties of the structures underlying labeled diffraction patterns in the opXRD database: a) distribution of structures containing specified elements, b) distribution of structures containing spacegroups, c) distribution of number of atoms $N_{\text{atom}}$ contained in unit cell, d) distribution of unit cell volume $V_{\text{uc}}$.}}
     \label{fig:hist_CRYSTAL}

### 2025-04-29 main.tex
Switch to Wiley Latex Template

@@
-\documentclass[a4paper]{article}
+\documentclass{WileyMSP-template}
 
@@
 %% Sets page size and margins
-\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
+%\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
 \usepackage{titling}
@@
 \begin{document}
+
+\pagestyle{fancy}
+\rhead{\includegraphics[width=2.5cm]{vch-logo.png}}
+
 \maketitle
 
-\textbf{Keywords}: powder x-ray diffraction, machine learning, open-access data, crystal structure determination, phase identification
+\keywords{powder x-ray diffraction, machine learning, open-access data, crystal structure determination, phase identification}
+
 \begin{abstract}
@@
 
-\bibliographystyle{bibstyle/bibstyle.bst}
+\bibliographystyle{bibstyle/MSP}
 \bibliography{references}
